### 2019-10-04 proposal/paper.tex
started researching approaches

\documentclass[12pt]{scrartcl}

\usepackage[american]{babel}

\usepackage{graphicx}
\graphicspath{{images/}}

\usepackage{paralist}
\usepackage{csquotes}
\usepackage[T1]{fontenc}
\usepackage{lmodern}

\usepackage{geometry}
% \geometry{a4paper,body={5.8in,9in}}
\geometry{a4paper}

\usepackage{amsmath, amsfonts, amssymb}
\usepackage{placeins}
\usepackage{subcaption}

\usepackage{setspace}

\setlength{\parindent}{0pt}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ab hier steht der eigentliche Text:
\begin{document}
\title{Thesis Title}
\subtitle{Master Thesis Proposal \& Work Plan}
% bei Masterarbeit ggf. ändern
\author{Clemens Damke\\ \small{Matriculation Number: 7011488}\\ \small{cdamke@mail.uni-paderborn.de}}
\date{\today}
\maketitle

\section{Motivation}\label{sec:motivation}

\section{Task Objectives}\label{sec:task-objectives}

\section{Task Description}\label{sec:task-description}

\section{Time-Schedule and Subtasks}\label{sec:schedule-and-subtasks}

% % %
\begin{figure}[!ht]
	\centering
	%\includegraphics[width=.9\textwidth]{timeschedule}
	\caption{Sketch of the time schedule for the work on the thesis}\label{fig:time-schedule}
\end{figure}
% % %
\section{Initial Document Structure}\label{sec:doc-structure}
\begin{enumerate}
	\item Introduction
	\item \dots
\end{enumerate}
% % %
\newpage
\bibliography{literature}
\bibliographystyle{myalphadin}

\vspace{6cm}

\begin{center}
	\begin{tabular}{l p{0.1\textwidth} r}
		\cline{1-1} \cline{3-3}
		\begin{minipage}[t]{0.4\textwidth}
			\centering
			\vspace{0cm}Supervisor
		\end{minipage}
		&
		\begin{minipage}[t]{0.2\textwidth}
		\end{minipage}
		&
		\begin{minipage}[t]{0.4\textwidth}
			\centering
			\vspace{0cm}Student
		\end{minipage}
	\end{tabular}
\end{center}

\end{document}

### 2019-10-06 proposal/paper.tex
started working on the proposal

@@
+% chktex-file 46
 \documentclass[12pt]{scrartcl}
 
+\PassOptionsToPackage{utf8}{inputenc}
+\usepackage{inputenc}
 \usepackage[american]{babel}
+\usepackage{csquotes}
+\usepackage{microtype}
 
@@
 \geometry{a4paper}
+\renewcommand{\baselinestretch}{1.1}
+\usepackage{parskip}
+\setlength{\parindent}{0pt}
 
 \usepackage{amsmath, amsfonts, amssymb}
+\usepackage{bm}
 \usepackage{placeins}
@@
 
-\setlength{\parindent}{0pt}
-%
+\usepackage{hyperref}
+
+\usepackage[nameinlink]{cleveref}
+\newcommand{\crefrangeconjunction}{--}
+
+\usepackage[						% use biblatex for bibliography
+	backend=bibtex,					% 	- use biber backend (bibtex replacement) or bibtex
+	style=numeric,					% 	- use alphabetic (or numeric) bib style
+	natbib=true,					% 	- allow natbib commands
+	hyperref=true,					% 	- activate hyperref support
+	backref=true,					% 	- activate backrefs
+	isbn=false,						% 	- don't show isbn tags
+	url=false,						% 	- don't show url tags
+	doi=false,						% 	- don't show doi tags
+	urldate=long,					% 	- display type for dates
+	maxnames=3,%
+	minnames=1,%
+	maxbibnames=5,%
+	minbibnames=3,%
+	maxcitenames=2,%
+	mincitenames=1%
+]{biblatex}
+\bibliography{literature}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ab hier steht der eigentliche Text:
+
+\newcommand{\thesisTitle}{Learning to Aggregate on Structured Data}
+\newcommand{\thesisSubject}{Master Thesis Proposal \& Work Plan}
+\newcommand{\thesisName}{Clemens Damke}
+\newcommand{\thesisMail}{cdamke@mail.uni-paderborn.de}
+\newcommand{\thesisMatNr}{7011488}
+\hypersetup{% setup the hyperref-package options
+    pdftitle={\thesisTitle},    %   - title (PDF meta)
+    pdfsubject={\thesisSubject},%   - subject (PDF meta)
+    pdfauthor={\thesisName},    %   - author (PDF meta)
+    plainpages=false,           %   -
+    colorlinks=false,           %   - colorize links?
+    pdfborder={0 0 0},          %   -
+    breaklinks=true,            %   - allow line break inside links
+    bookmarksnumbered=true,     %
+    bookmarksopen=true          %
+}
+
 \begin{document}
-\title{Thesis Title}
-\subtitle{Master Thesis Proposal \& Work Plan}
-% bei Masterarbeit ggf. ändern
-\author{Clemens Damke\\ \small{Matriculation Number: 7011488}\\ \small{cdamke@mail.uni-paderborn.de}}
+
+\title{\thesisTitle}
+\subtitle{\thesisSubject}
+\author{{\thesisName}\\\small{Matriculation Number: \thesisMatNr}\\\small{\href{mailto:\thesisMail}{\thesisMail}}}
 \date{\today}
@@
 
-\section{Task Objectives}\label{sec:task-objectives}
+Most of the commonly used supervised machine learning techniques assume that instances are represented by $d$-dimensional feature vectors $x \in \mathcal{X} = \mathcal{X}_1 \times \cdots \times \mathcal{X}_d$ for which some target value $y \in \mathcal{Y}$ should be predicted.
+In the regression setting the target domain $\mathcal{Y}$ is continuous, typically $\mathcal{Y} = \mathbb{R}$, whereas $\mathcal{Y}$ is some discrete set of classes in the classification setting.
 
-\section{Task Description}\label{sec:task-description}
+Since not all data is well-suited for a fixed-dimensional vector representation, approaches that directly consider the structure of the input data might be more appropriate in such cases.
+One such case is the class of so-called \textit{``learning to aggregate''} (LTA) problems as described by \citet{Melnikov2016}.
+There the instances are represented by compositions $\bm{c}$ of constituents $c_i \in \bm{c}$, i.e.\@ variable-size multisets.
+The assumption in LTA problems is that for all constituents $c_i$ a local valuation $y_i \in \mathcal{Y}$ is either given or computable.
+The set of those local valuations should be indicative of the overall valuation $y \in \mathcal{Y}$ of the entire composition $\bm{c}$.
+The goal of LTA is to learn a variadic aggregation function $A: \mathcal{Y}^{*} \to \mathcal{Y}$ that estimates such composite valuations, i.e.\@ $\hat{y} = A(y_1, \dots, y_n)$ for a composition with $n$ constituents.
+Additionally the aggregation function $A$ should be associative and commutative to fit with the multiset-structure of compositions.
 
-\section{Time-Schedule and Subtasks}\label{sec:schedule-and-subtasks}
+\section{Related Work}\label{sec:related-work}
 
-% % %
-\begin{figure}[!ht]
-	\centering
-	%\includegraphics[width=.9\textwidth]{timeschedule}
-	\caption{Sketch of the time schedule for the work on the thesis}\label{fig:time-schedule}
-\end{figure}
-% % %
-\section{Initial Document Structure}\label{sec:doc-structure}
+\section{Goals}\label{sec:goals}
+
+\subsection{Required Goals}\label{sec:goals:req}
+
+\subsection{Optional Goals}\label{sec:goals:opt}
+
+\section{Approach}\label{sec:approach}
+
+\section{Preliminary Document Structure}\label{sec:doc-structure}
+
 \begin{enumerate}
@@
 \end{enumerate}
+
+\section{Time-Schedule}\label{sec:schedule}
+
+\begin{figure}[!ht]
+	\centering
+	%\includegraphics[width=.9\textwidth]{timeschedule}
+	\caption{Sketch of the time schedule for the work on the thesis}\label{fig:schedule}
+\end{figure}
+
 % % %
 \newpage
-\bibliography{literature}
-\bibliographystyle{myalphadin}
+{%
+\renewcommand{\bibfont}{\normalfont\small}
+\setlength{\biblabelsep}{5pt}
+\setlength{\bibitemsep}{0.5\baselineskip plus 0.5\baselineskip} % chktex 1
+\setcounter{biburllcpenalty}{9000}
+\setcounter{biburlucpenalty}{9999}
+\printbibliography%
+}
 

### 2019-10-07 proposal/paper.tex
continued with the proposal

@@
 \usepackage{hyperref}
-
 \usepackage[nameinlink]{cleveref}
@@
 \bibliography{literature}
+
+\usepackage[inline]{enumitem}
 
@@
 
-\section{Motivation}\label{sec:motivation}
+\section{Motivation}%
+\label{sec:motivation}
 
@@
 Since not all data is well-suited for a fixed-dimensional vector representation, approaches that directly consider the structure of the input data might be more appropriate in such cases.
-One such case is the class of so-called \textit{``learning to aggregate''} (LTA) problems as described by \citet{Melnikov2016}.
+One such case is the class of so-called \textit{learning to aggregate} (LTA) problems as described by \citet{Melnikov2016}.
 There the instances are represented by compositions $\bm{c}$ of constituents $c_i \in \bm{c}$, i.e.\@ variable-size multisets.
@@
 
-\section{Related Work}\label{sec:related-work}
+Current LTA approaches only work with multiset inputs.
+In practice there is however often some relational structure among the constituents of a composition.
+This effectively turns LTA into a graph classification or regression problem.
+The overall aim of this thesis is to look into the question of how aggregation function learning methods might be generalized to the graph setting.
 
-\section{Goals}\label{sec:goals}
+\section{Related Work}%
+\label{sec:related-work}
 
-\subsection{Required Goals}\label{sec:goals:req}
+This thesis will be based on two currently mostly unrelated fields of research:
+\begin{enumerate*}
+	\item Aggregation function learning
+	\item Graph classification
+\end{enumerate*}.
+A short overview of the current state-of-the-art approaches in both fields will be given now.
 
-\subsection{Optional Goals}\label{sec:goals:opt}
+\subsection{Aggregation Function Learning}%
+\label{sec:related-work:lta}
 
-\section{Approach}\label{sec:approach}
+\subsection{Graph Classification}%
+\label{sec:related-work:gc}
 
-\section{Preliminary Document Structure}\label{sec:doc-structure}
+\section{Goals}%
+\label{sec:goals}
+
+\subsection{Required Goals}%
+\label{sec:goals:req}
+
+\subsection{Optional Goals}%
+\label{sec:goals:opt}
+
+\section{Approach}%
+\label{sec:approach}
+
+\section{Preliminary Document Structure}%
+\label{sec:doc-structure}
 
@@
 
-\section{Time-Schedule}\label{sec:schedule}
+\section{Time-Schedule}%
+\label{sec:schedule}
 

### 2019-10-08 proposal/paper.tex
started describing the lta uninorms approach

@@
 \begin{enumerate*}
-	\item Aggregation function learning
+	\item Learning to Aggregate
 	\item Graph classification
@@
 
-\subsection{Aggregation Function Learning}%
+\subsection{Learning to Aggregate}%
 \label{sec:related-work:lta}
+
+Two main approaches to represent the aggregation function in LTA problems have been explored.
+The first approach uses so-called uninorms~\cite{Melnikov2016} to do so.
+There the basic idea is to express composite valuations as fuzzy logical assignments $y \in [0, 1]$.
+Such a composite assignment $y$ is modeled as the result of a parameterized logical expression of constituent assignments $y_i \in [0, 1]$.
+As the logical expression that thus effectively aggregates the constituents, a uninorm $U_{\lambda}$ is used.
+Depending on the parameter $\lambda$, a uninorm interpolates between t-norms and t-conorms which correspond to logical conjunction and disjunction respectively.
+\begin{align*}
+
+\end{align*}
 

### 2019-10-09 proposal/paper.tex
a bit more on lta

@@
 	maxcitenames=2,%
-	mincitenames=1%
+	mincitenames=1,%
+	sorting=none
 ]{biblatex}
@@
 \usepackage[inline]{enumitem}
+\usepackage{todonotes}
 
@@
 There the instances are represented by compositions $\bm{c}$ of constituents $c_i \in \bm{c}$, i.e.\@ variable-size multisets.
-The assumption in LTA problems is that for all constituents $c_i$ a local valuation $y_i \in \mathcal{Y}$ is either given or computable.
-The set of those local valuations should be indicative of the overall valuation $y \in \mathcal{Y}$ of the entire composition $\bm{c}$.
-The goal of LTA is to learn a variadic aggregation function $A: \mathcal{Y}^{*} \to \mathcal{Y}$ that estimates such composite valuations, i.e.\@ $\hat{y} = A(y_1, \dots, y_n)$ for a composition with $n$ constituents.
+The assumption in LTA problems is that for all constituents $c_i$ a local score $y_i \in \mathcal{Y}$ is either given or computable.
+The set of those local scores should be indicative of the overall score $y \in \mathcal{Y}$ of the entire composition $\bm{c}$.
+The goal of LTA is to learn a variadic aggregation function $A: \mathcal{Y}^{*} \to \mathcal{Y}$ that estimates such composite scores, i.e.\@ $\hat{y} = A(y_1, \dots, y_n)$ for a composition with $n$ constituents.
 Additionally the aggregation function $A$ should be associative and commutative to fit with the multiset-structure of compositions.
@@
 Two main approaches to represent the aggregation function in LTA problems have been explored.
-The first approach uses so-called uninorms~\cite{Melnikov2016} to do so.
-There the basic idea is to express composite valuations as fuzzy logical assignments $y \in [0, 1]$.
+The first approach uses \textit{uninorms}~\cite{Melnikov2016} to do so.
+There the basic idea is to express composite scores as fuzzy truth assignments $y \in [0, 1]$.
 Such a composite assignment $y$ is modeled as the result of a parameterized logical expression of constituent assignments $y_i \in [0, 1]$.
 As the logical expression that thus effectively aggregates the constituents, a uninorm $U_{\lambda}$ is used.
-Depending on the parameter $\lambda$, a uninorm interpolates between t-norms and t-conorms which correspond to logical conjunction and disjunction respectively.
-\begin{align*}
+Depending on the parameter $\lambda$, $U_{\lambda}$ interpolates between t-norms and t-conorms which are continuous generalizations of logical conjunction and disjunction respectively.
+\todo{TODO:\@ Details}
 
-\end{align*}
+Recently \citet{Melnikov2019} have also looked at an alternative class of aggregation function.
+Instead of using fuzzy logic to describe score aggregation, \textit{ordered weighted average} (OWA) operators were used.
+\todo{TODO:\@ Details}
 

### 2019-10-10 proposal/paper.tex
added related work for graph classification

@@
 
-Most of the commonly used supervised machine learning techniques assume that instances are represented by $d$-dimensional feature vectors $x \in \mathcal{X} = \mathcal{X}_1 \times \cdots \times \mathcal{X}_d$ for which some target value $y \in \mathcal{Y}$ should be predicted.
+Most of the commonly used supervised machine learning techniques assume that instances are represented by $d$-dimensional feature vectors $x_i \in \mathcal{X} = \mathcal{X}_1 \times \cdots \times \mathcal{X}_d$ for which some target value $y_i \in \mathcal{Y}$ should be predicted.
 In the regression setting the target domain $\mathcal{Y}$ is continuous, typically $\mathcal{Y} = \mathbb{R}$, whereas $\mathcal{Y}$ is some discrete set of classes in the classification setting.
@@
 One such case is the class of so-called \textit{learning to aggregate} (LTA) problems as described by \citet{Melnikov2016}.
-There the instances are represented by compositions $\bm{c}$ of constituents $c_i \in \bm{c}$, i.e.\@ variable-size multisets.
-The assumption in LTA problems is that for all constituents $c_i$ a local score $y_i \in \mathcal{Y}$ is either given or computable.
-The set of those local scores should be indicative of the overall score $y \in \mathcal{Y}$ of the entire composition $\bm{c}$.
-The goal of LTA is to learn a variadic aggregation function $A: \mathcal{Y}^{*} \to \mathcal{Y}$ that estimates such composite scores, i.e.\@ $\hat{y} = A(y_1, \dots, y_n)$ for a composition with $n$ constituents.
-Additionally the aggregation function $A$ should be associative and commutative to fit with the multiset-structure of compositions.
+There the instances are represented by compositions $\bm{c}_i$ of constituents $c_{i,j} \in \bm{c}_i$, i.e.\@ variable-size multisets with $n_i = |\bm{c}_i|$.
+The assumption in LTA problems is that for all constituents $c_{i,j}$ a local score $y_{i,j} \in \mathcal{Y}$ is either given or computable.
+The set of those local scores should be indicative of the overall score $y_i \in \mathcal{Y}$ of the entire composition $\bm{c}_i$.
+LTA problems typically require two subproblems to be solved:
+\begin{enumerate}[label=\textbf{\arabic*.}]
+	\item \textbf{Aggregation:}
+		A variadic aggregation function $A: \mathcal{Y}^{*} \to \mathcal{Y}$ that estimates composite scores has to be learned, i.e.\@ $y_i \approx \hat{y}_i = A(y_{i,1}, \dots, y_{i,n_i})$.
+		Typically the aggregation function $A$ should be associative and commutative to fit with the multiset-structure of compositions.
+	\item \textbf{Disaggregation:}
+		In case the constituent scores $y_{i,j}$ are not given, they have to be derived from a constituent representation, e.g.\ a vector $v_{i,j} \in \mathcal{V}$.
+		To learn this derivation function $f: \mathcal{V} \to \mathcal{Y}$, only the constituent vectors ${\{v_{i,j}\}}_{j = 1}^{n_i}$ and the composite score $y_i$ is given.
+		Thus the constituent scores $y_{i,j}$ need to be \textit{disaggregated} from $y_i$ in order to learn $f$.
+\end{enumerate}
+Overall LTA can be understood as the joint problem of learning the aggregation function $A$ and the local score derivation function $f$.
 
@@
 In practice there is however often some relational structure among the constituents of a composition.
-This effectively turns LTA into a graph classification or regression problem.
-The overall aim of this thesis is to look into the question of how aggregation function learning methods might be generalized to the graph setting.
+This effectively turns LTA into a graph regression problem.
+The goal of this thesis is to look into the question of how aggregation function learning methods might be generalized to the graph setting.
 
@@
 
-This thesis will be based on two currently mostly unrelated fields of research:
-\begin{enumerate*}
+This thesis will be based on two currently mostly separate fields of research:
+\begin{enumerate*}[label=\textbf{\arabic*.}]
 	\item Learning to Aggregate
@@
 The first approach uses \textit{uninorms}~\cite{Melnikov2016} to do so.
-There the basic idea is to express composite scores as fuzzy truth assignments $y \in [0, 1]$.
-Such a composite assignment $y$ is modeled as the result of a parameterized logical expression of constituent assignments $y_i \in [0, 1]$.
+There the basic idea is to express composite scores as fuzzy truth assignments $y_i \in [0, 1]$.
+Such a composite assignment $y_i$ is modeled as the result of a parameterized logical expression of constituent assignments $y_{i,j} \in [0, 1]$.
 As the logical expression that thus effectively aggregates the constituents, a uninorm $U_{\lambda}$ is used.
@@
 \label{sec:related-work:gc}
+
+As previously mentioned, the addition of relations between constituents turns LTA into a graph regression problem.
+Most of the recent research in the field of learning from graph structured data however focused on the closely related graph classification problem.
+Since many ideas from the classification setting are also applicable in the regression setting, a brief overview of those ideas is given.
+
+At a high level graph classification methods can be taxonomized into two main families:
+\begin{enumerate}[label=\textbf{\arabic*.}]
+	\item \textbf{Vector representation approaches:}
+		One way to tackle the graph classification problem is to map an input graph $G$ to a vectorial representation.
+		This can be done
+		\begin{enumerate*}
+			\item by either handpicking global graph features like vertex/edge count, degree distribution or graph diameter,
+			\item via a graph embedding algorithm like node2vec~\cite{Grover2016}, sub2vec~\cite{Adhikari2018} or graph2vec~\cite{Narayanan2017},
+			\item implicitly by using a graph kernel that computes the structural similarity between graphs, e.g.\ the Weisfeiler-Lehman kernel~\cite{Shervashidze2011} or the multiscale Laplacian graph kernel~\cite{Kondor2016}
+		\end{enumerate*}.
+		Graphs can then be classified via any classification algorithm that works with vectors and/or kernels.
+	\item \textbf{Graph neural networks (GNNs):}
+
+\end{enumerate}
 

### 2019-10-12 proposal/paper.tex
added gnn related work

@@
 		Graphs can then be classified via any classification algorithm that works with vectors and/or kernels.
-	\item \textbf{Graph neural networks (GNNs):}
+	\item \textbf{Graph neural networks:}
+		An alternative approach is to adapt neural networks to graph inputs.
+		The notion of a \textit{graph neural network} (GNN) was first introduced by \citet{Gori2005}.\@
+		There a message-passing architecture is used to iteratively propagate vertex information to neighbors until a fixed point is reached.
+		This process is computationally expensive.
 
+		Recently the class of \textit{convolutional graph neural networks} (ConvGNNs) gained traction.
+		It generalizes the convolution operator that is used by \textit{convolutional neural networks} (CNNs) to graphs.
+		There are two main variants of ConvGNNs:
+		\begin{enumerate}[label=\textbf{\alph*)}] % chktex 9 chktex 10
+			\item \textbf{Spatial variant:}
+				Standard CNNs use a convolution operator that is applied to a grid graph with diagonal edges.
+				Spatial ConvGNNs directly generalize this idea by defining a more flexible convolution that aggregates the local neighborhood of each vertex.
+				Spatial graph convolution aggregates the direct neighborhood of each vertex where each vertex $v_i$ is typically described by a feature vector $x_i \in \mathbb{R}^k$.
+				This process returns a new aggregate feature vector $x'_i$.
+				By stacking multiple convolution layers, features from indirect neighbors become part of the aggregate.
+				Analogous to how CNNs learn kernel matrices, spatial ConvGNNs learn vertex neighborhood feature aggregation functions.
+				This approach shares similarities with the previously mentioned message-passing architecture~\cite{Gori2005} with the major different being that the number of message-passing, i.e.\ convolution, steps is fixed.
+			\item \textbf{Spectral variant:}
+				Motivated by the theoretical foundation provided by spectral graph theory, spectral ConvGNNs learn a filter function $\hat{g}$ that is applied in the Fourier domain of a graph.
+				Based on the convolutional theorem, a convolution operator can not only be expressed in terms of the neighborhood of vertices but also as a filter on the eigenvectors of a graph's Laplacian.
+				Formally this can be expressed as
+				\begin{align*}
+					g *_G \bm{x} = U \hat{g}(\Lambda) U^{\top} \bm{x}
+				\end{align*}
+				where $\bm{x} \in \mathbb{R}^{n \times k}$ is the matrix of all vertex features and $U \Lambda U^{\top}$ is the eigendecomposition of the graph Laplacian $L_G$.
+				In the previously described spatial ConvGNN variant, $g$ is learned directly in form of the feature aggregation function with a neighborhood locality constraint.
+				In the spectral variant however the Fourier-transformed filter $\hat{g}$ is learned.
+				Intuitively this means that vertex features $x_i$ are not aggregated with the features of their their direct neighborhood but with the features of vertices with which they share certain structural similarities.
+				This allows spectral ConvGNNs to incorporate the global graph structure at the cost of having to perform the computationally expensive decomposition of every input graph's Laplacian.
+
+				To tackle the performance impact of the spectral convolution approach, various simplifications of the filter $\hat{g}$ have been proposed.
+				The so-called \textit{graph convolutional network}~(GCN) does this by reducing the expressivity of the filter.
+				Then $\hat{g}$ can be applied to the Laplacian directly via $g *_G \bm{x} = \hat{g}(L_G) \bm{x}$ which saves the decomposition costs.
+				This simplification of $\hat{g}$ implicitly causes its Fourier inverse $g$ to be locality constrained just like in the spatial ConvGNN variant.
+				Therefore the spectral GCN approach also allows for a spatial/message-passing interpretation.
+				Recently various variants of GCN have been proposed that apply additional approximations to improve performance in certain scenarios.
+		\end{enumerate}
+		After the application of either spatial or spectral graph convolutions, each vertex will have an updated aggregate feature vector.
+		To obtain an aggregate score for the entire graph, a graph pooling layer is applied.
+		A primitive pooling approach is to simply merge the vertex features via a fixed aggregation function like $\min$, $\max$ or $+$.
+		Better results can be obtained by learning the pooling aggregation function.
+		Two notable graph pooling approaches are SortPooling and Self-Attention Pooling.
+		SortPooling works by determining a top-k ordering of the vertices and then aggregates the remaining fixed-size vertex features via a standard multilayer perceptron (MLP).
 \end{enumerate}

### 2019-10-14 proposal/paper.tex
added graph regression related work

@@
 	\item Learning to Aggregate
-	\item Graph classification
+	\item Graph classification \& regression
 \end{enumerate*}.
@@
 Most of the recent research in the field of learning from graph structured data however focused on the closely related graph classification problem.
-Since many ideas from the classification setting are also applicable in the regression setting, a brief overview of those ideas is given.
+Since many ideas from the classification setting are also useful in the regression setting, a brief overview of those ideas is given.
 
@@
 				Standard CNNs use a convolution operator that is applied to a grid graph with diagonal edges.
-				Spatial ConvGNNs directly generalize this idea by defining a more flexible convolution that aggregates the local neighborhood of each vertex.
-				Spatial graph convolution aggregates the direct neighborhood of each vertex where each vertex $v_i$ is typically described by a feature vector $x_i \in \mathbb{R}^k$.
+				Spatial ConvGNNs~\cite{Micheli2009} directly generalize this idea by defining a more flexible convolution that aggregates the local neighborhood of each vertex.
+				Spatial graph convolution aggregates the direct neighborhood of each vertex where each vertex $v_i$ is typically described by a feature vector $x_i \in \mathbb{R}^d$.
 				This process returns a new aggregate feature vector $x'_i$.
@@
 			\item \textbf{Spectral variant:}
-				Motivated by the theoretical foundation provided by spectral graph theory, spectral ConvGNNs learn a filter function $\hat{g}$ that is applied in the Fourier domain of a graph.
+				Motivated by the theoretical foundation provided by spectral graph theory~\cite{Shuman2013}, spectral ConvGNNs~\cite{Bruna2013} learn a filter function $\hat{g}$ that is applied in the Fourier domain of a graph.
 				Based on the convolutional theorem, a convolution operator can not only be expressed in terms of the neighborhood of vertices but also as a filter on the eigenvectors of a graph's Laplacian.
@@
 				To tackle the performance impact of the spectral convolution approach, various simplifications of the filter $\hat{g}$ have been proposed.
-				The so-called \textit{graph convolutional network}~(GCN) does this by reducing the expressivity of the filter.
+				The so-called \textit{graph convolutional network}~(GCN)~\cite{Kipf2017} does this by reducing the expressivity of the filter.
 				Then $\hat{g}$ can be applied to the Laplacian directly via $g *_G \bm{x} = \hat{g}(L_G) \bm{x}$ which saves the decomposition costs.
@@
 				Therefore the spectral GCN approach also allows for a spatial/message-passing interpretation.
-				Recently various variants of GCN have been proposed that apply additional approximations to improve performance in certain scenarios.
+				Recently various variants of GCNs~\cite{Hamilton2017}\cite{Chen2018}\cite{Chen2017}\cite{Chiang2019} have been proposed that apply additional approximations and sampling methods to improve the runtime as well as the accuracy in certain scenarios.
 		\end{enumerate}
@@
 		Better results can be obtained by learning the pooling aggregation function.
-		Two notable graph pooling approaches are SortPooling and Self-Attention Pooling.
-		SortPooling works by determining a top-k ordering of the vertices and then aggregates the remaining fixed-size vertex features via a standard multilayer perceptron (MLP).
+		Two notable graph pooling approaches are \textit{SortPooling}~\cite{Zhang2018} and \textit{Self-Attention Pooling}~\cite{Lee2019}.
+		\begin{enumerate}
+			\item SortPooling assumes that the vertex feature vectors are obtained using a stack of spectral convolution layers that learn filters on the random walk Laplacian $L = I - D^{-1} A$.
+				This assumption makes it possible to interpret the outputs of the graph convolution layers as a continuous generalization of \textit{Weisfeiler-Lehman} (WL) colors.
+				WL colors are lexicographically sortable signatures that encode the structural role of vertices; they can be used for efficient graph isomorphism falsification~\cite{Weisfeiler1968}.
+				Using this interpretation of the vertex feature vectors, SortPooling determines a top-$k$ ranking of the vertices' WL colors and then aggregates the thus selected fixed-size vertex feature subset via a standard \textit{multilayer perceptron} (MLP).
+			\item Self-Attention Pooling adds an additional graph convolution layer after an arbitrary ConvGNN.\@
+				The added layer takes the aggregated vertex feature vectors as input and outputs a score for each vertex.
+				This score is used to determine a top-$k$ vertex ranking.
+				Unlike SortPooling the feature vectors of the resulting subset are first added up.
+				The resulting aggregate graph feature vector is then fed into a MLP to obtain a class.
 		\end{enumerate}
+\end{enumerate}
+
+\subsection{Graph Regression}%
+\label{sec:related-work:gr}
+
+An overview of the recent work regarding graph classification was just given.
+The generalization of LTA to the graph domain is however better described as a graph regression problem since the predicted composite scores are continuous.
+As there currently are relatively few graph regression approaches, only few references can be provided here.
+
+Graph regression problems have for the most part only been discussed in the context of specific domain problems.
+\textit{RASAR}~\cite{Luechtefeld2018} is a graph regression method to predict the toxicity of chemicals as an alternative for animal tests.
+A standard logistic regression model is trained with manually chosen features that are given for the training molecule graphs.
+To predict the toxicity of a new chemical, $k$-means is used to average the features of chemicals with similar molecule graphs in the training set.
+The structural similarity between molecule graphs is measured by the Tanimoto coefficient of their chemical fingerprints.
+Those fingerprints encode the presence or absence of certain meaningful chemical substructures.
+The implementation and used dataset is unfortunately not publicly available because it is part of a commercial product\footnote{ToxTrack Inc., \url{https://toxtrack.com/}}.
+
+\textit{ProTox}~\cite{Drwal2014}\cite{Banerjee2018} is another toxicity prediction approach.
+It works similarly to RASAR but leaves out the logistic regression step and instead directly aggregates the toxicity of similar chemicals.
+ProTox uses the freely available \textit{SuperToxic} dataset~\cite{Schmidt2009}.
+While the implementation is again not publicly available, the trained model can be freely queried online\footnote{ProTox-II, \url{http://tox.charite.de/protox_II/}}.
 
@@
 \label{sec:goals}
+
+As mentioned in \cref{sec:motivation}, the overall goal of the thesis is to extend the LTA approach described in \cref{sec:related-work:lta} to the graph setting.
+Since this goal strongly overlaps with the work in graph classification described in \cref{sec:related-work:gc}, both branches of research should be combined.
+The practical applications described in \cref{sec:related-work:gr} could be used for comparison and a potential source of a graph regression test dataset.
+Now follows a concrete list of required and optional goals that should be achieved.
 
@@
 
-\end{document}
+\end{document} % chktex 17

### 2019-10-15 proposal/paper.tex
added required goals

@@
 Depending on the parameter $\lambda$, $U_{\lambda}$ interpolates between t-norms and t-conorms which are continuous generalizations of logical conjunction and disjunction respectively.
-\todo{TODO:\@ Details}
 
@@
 Instead of using fuzzy logic to describe score aggregation, \textit{ordered weighted average} (OWA) operators were used.
-\todo{TODO:\@ Details}
+OWA aggregators work by sorting the input scores and then weighting them based on their sort position, i.e.\ %
+\begin{align*}
+	A_{\lambda}(y_1, \dots, y_n) := \sum_{i = 1}^n \lambda_i y_{\pi(i)},
+\end{align*}
+where $\lambda$ is a weight vector with ${\|\lambda\|}_1 = 1$ and $\pi$ is a sorting permutation of the input scores. % chktex 21
+To deal with varying composite sizes $n$ the weights $\lambda_i$ are interpolated using a \textit{basic unit interval monotone} (BUM) function $q: [0, 1] \to [0, 1]$.
+It is used by normalizing every constituent's position $\pi(i)$ to the unit interval via $\frac{\pi(i)}{n}$.
+The BUM function $q$ is then used to interpolate a weight for any normalized sort position.
+Therefore the learning objective boils down to optimizing the shape of $q$.
+The details of this are left out here.
 
@@
 Since this goal strongly overlaps with the work in graph classification described in \cref{sec:related-work:gc}, both branches of research should be combined.
-The practical applications described in \cref{sec:related-work:gr} could be used for comparison and a potential source of a graph regression test dataset.
+The practical applications described in \cref{sec:related-work:gr} could be used for comparison and as a potential source of a graph regression test dataset.
 Now follows a concrete list of required and optional goals that should be achieved.
@@
 
+\begin{enumerate}[label=\textbf{\arabic*.}]
+	\item \textbf{Aggregation:}
+		Define a class of aggregation functions that takes vertex scores as its input.
+		It should fulfill the role of the uninorms and OWA functions in the existing LTA papers.
+	\item \textbf{Disaggregation:}
+		Describe how vertex features should be used to compute a constituent score that will be combined by the aggregation function.
+	\item \textbf{Learning:}
+		Combine the chosen aggregation and disaggregation methods into a joint LTA model with a suitable loss function.
+		This model should be optimizable reasonably efficiently on modern hardware.
+	\item \textbf{Implementation:}
+		A reference implementation for the designed LTA model should be developed.
+	\item \textbf{Evaluation:}
+		The reference implementation should be used to evaluate the model on at least one test dataset.
+		Due to the scarcity of graph regression datasets, the used test data might be randomly generated.
+		Generated dataset should be obtained using a variety of representative deterministic graph scoring measures to check which kinds of local and global structural features are learnable;
+		some examples of measures that might be part of a graph scoring functions are the count of $k$-cliques in a graph, the graph diameter, the presence of cycles or various properties of a graph's degree distribution.
+\end{enumerate}
+
 \subsection{Optional Goals}%
 \label{sec:goals:opt}
+
+In addition to the previous goals, the following goals should be pursued if possible:
+\begin{itemize}
+	\item
+\end{itemize}
 

### 2019-10-16 proposal/paper.tex
added optional goals

@@
 		Define a class of aggregation functions that takes vertex scores as its input.
-		It should fulfill the role of the uninorms and OWA functions in the existing LTA papers.
+		It should fulfill the role of the uninorms and OWA functions in the existing LTA papers (see \cref{sec:related-work:lta}).
 	\item \textbf{Disaggregation:}
@@
 		Combine the chosen aggregation and disaggregation methods into a joint LTA model with a suitable loss function.
-		This model should be optimizable reasonably efficiently on modern hardware.
 	\item \textbf{Implementation:}
@@
 		some examples of measures that might be part of a graph scoring functions are the count of $k$-cliques in a graph, the graph diameter, the presence of cycles or various properties of a graph's degree distribution.
+		The results should be compared with the results of at least one other approach for learning graph scores, e.g.\ a ConvCNN with a fixed aggregation function instead of a learned one.
 \end{enumerate}
@@
 
-In addition to the previous goals, the following goals should be pursued if possible:
-\begin{itemize}
-	\item
-\end{itemize}
+If possible, the previous goals could be extended.
+For each of the previous five goal categories, optional additional ideas are listed below.
+\begin{enumerate}[label=\textbf{\arabic*.}]
+	\item \textbf{Aggregation:}
+		Multiple classes of aggregation functions could be considered and compared.
+	\item \textbf{Disaggregation:}
+		As describen in \cref{sec:related-work:gc} there are multiple methods to learn vertex scores.
+		Depending on the chosen overall approach, a few of those methods might be suitable to be combined with an aggregation function learner.
+		Those suitable combinations could be evaluated and compared.
+	\item \textbf{Learning:}
+		The joint model of aggregation and disaggregation function learning can most likely be optimized using various algorithms.
+		Those algorithms could be compared.
+		In case their runtime performance is impractical in certain situations, approximations and simplifications of the model could be considered.
+	\item \textbf{Implementation:}
+		The chosen approach should ideally align well with the practical constraints of implementation.
+		It should be suitable for an implementation on modern GPUs and/or distributed cluster environments.
+	\item \textbf{Evaluation:}
+		It would be interesting to not only use generated data but also a real dataset for evaluation (see \cref{sec:related-work:gr}).
+\end{enumerate}
 
@@
 	\item Introduction
-	\item \dots
+	\item
+	\item Evaluation
+	\item Conclusion
+	\item Literature
+	\item Appendix
 \end{enumerate}

### 2019-10-17 proposal/paper.tex
extended related works

@@
 				Therefore the spectral GCN approach also allows for a spatial/message-passing interpretation.
-				Recently various variants of GCNs~\cite{Hamilton2017}\cite{Chen2018}\cite{Chen2017}\cite{Chiang2019} have been proposed that apply additional approximations and sampling methods to improve the runtime as well as the accuracy in certain scenarios.
+				Recently various variants and extensions of GCNs~\cite{Hamilton2017}\cite{Chen2018}\cite{Chen2017}\cite{Chiang2019}\cite{Du2017} have been proposed that apply additional approximations and sampling methods to improve the runtime as well as the accuracy in certain scenarios.
+
+				One notable extension of GCNs are the so-called \textit{adadptive graph convolutional networks} (AGCNs)~\cite{Li2018}.
+				AGCNs do not use a fixed Laplacian $L$ for each input but instead learn a so-called residual Laplacian $L_\mathit{res}$.
+				The residual Laplacian is constructed by learning a vertex similarity kernel $\mathbb{G}_{x_i,x_j}$ which is used to determine the residual edge weights between vertices.
+				By overlaying the learned residual Laplacian $L_\mathit{res}$ on top of the static intrinsic Laplacian $L$, similar nodes with a large spatial distance in $L$ become connected.
+				This can improve the accuracy especially if a locality constrained spectral filter $\hat{g}$ is used.
 		\end{enumerate}
@@
 Graph regression problems have for the most part only been discussed in the context of specific domain problems.
-\textit{RASAR}~\cite{Luechtefeld2018} is a graph regression method to predict the toxicity of chemicals as an alternative for animal tests.
-A standard logistic regression model is trained with manually chosen features that are given for the training molecule graphs.
+One of those domains is the prediction of chemical toxicity, typically the $\text{LD}_{50}$ or $\text{LC}_{50}$ doses which measure the smallest amount of a given chemical that is lethal for at least $50\%$ of tested subjects.
+The following approaches could serve as inspiration, baselines and/or sources of training data:
+\begin{itemize}
+	\item \textit{RASAR}~\cite{Luechtefeld2018} uses a logistic regression model.
+		It is trained with manually chosen features that are given for training molecule graphs.
 		To predict the toxicity of a new chemical, $k$-means is used to average the features of chemicals with similar molecule graphs in the training set.
@@
 		The implementation and used dataset is unfortunately not publicly available because it is part of a commercial product\footnote{ToxTrack Inc., \url{https://toxtrack.com/}}.
-
-\textit{ProTox}~\cite{Drwal2014}\cite{Banerjee2018} is another toxicity prediction approach.
-It works similarly to RASAR but leaves out the logistic regression step and instead directly aggregates the toxicity of similar chemicals.
+	\item \textit{ProTox}~\cite{Drwal2014}\cite{Banerjee2018} is another toxicity prediction approach.
+		It works similarly to RASAR but leaves out the logistic regression step and instead directly aggregates the toxicity of similar chemicals via $k$-means.
 		ProTox uses the freely available \textit{SuperToxic} dataset~\cite{Schmidt2009}.
 		While the implementation is again not publicly available, the trained model can be freely queried online\footnote{ProTox-II, \url{http://tox.charite.de/protox_II/}}.
+	\item The \textit{Toxicity Estimation Software Tools}~(T.E.S.T.)\footnote{\url{https://www.epa.gov/chemical-research/toxicity-estimation-software-tool-test}} is a free collection of models for toxicity prediction.
+		Like RASAR and ProTox it includes similarity-based prediction models but also linear regression and decision tree models that are trained on vector representations of chemicals.
+		The \textit{ChemIDplus}\footnote{\url{https://chem.nlm.nih.gov/chemidplus/}} database was used for training.
+\end{itemize}
+
+Recently first toxicity prediction approaches using ConvGNNs have been published.
+The previously mentioned AGCN~\cite{Li2018} was in fact evaluated on a discrete toxicity classification problem;
+this could serve as a starting point for a continuous graph regression method.
+\citet{Pope2018} describe how functional groups of toxic chemicals can be learned using GCNs;
+there a functional group refers to a relevant subgraph of a molecule that can be interpreted as a toxicity indicator.
 
@@
 	\item \textbf{Evaluation:}
-		It would be interesting to not only use generated data but also a real dataset for evaluation (see \cref{sec:related-work:gr}).
+		Not only generated data but also real datasets should be used for evaluation (see \cref{sec:related-work:gr}).
 \end{enumerate}

### 2019-10-17 proposal/paper.tex
first proposal draft

@@
 \usepackage{todonotes}
+\usepackage{pgfgantt}
+
+% Colors:
+\definecolor{blau}{HTML}{355FB3}
+\definecolor{rot}{HTML}{B33535}
+\definecolor{gruen}{HTML}{3BB335}
+\definecolor{hellblau}{HTML}{8ea7d7}
+\definecolor{hellgrau}{HTML}{cccccc}
 
@@
 The assumption in LTA problems is that for all constituents $c_{i,j}$ a local score $y_{i,j} \in \mathcal{Y}$ is either given or computable.
-The set of those local scores should be indicative of the overall score $y_i \in \mathcal{Y}$ of the entire composition $\bm{c}_i$.
+The set of those local scores should be indicative of the overall score $y_i \in \mathcal{Y}$ of the composition $\bm{c}_i$.
 LTA problems typically require two subproblems to be solved:
@@
 
-Recently \citet{Melnikov2019} have also looked at an alternative class of aggregation function.
+Recently \citet{Melnikov2019} have also looked at an alternative class of aggregation functions.
 Instead of using fuzzy logic to describe score aggregation, \textit{ordered weighted average} (OWA) operators were used.
@@
 To deal with varying composite sizes $n$ the weights $\lambda_i$ are interpolated using a \textit{basic unit interval monotone} (BUM) function $q: [0, 1] \to [0, 1]$.
-It is used by normalizing every constituent's position $\pi(i)$ to the unit interval via $\frac{\pi(i)}{n}$.
-The BUM function $q$ is then used to interpolate a weight for any normalized sort position.
+It takes constituent positions that are normalized to the unit interval, i.e.\ $\frac{\pi(i)}{n}$.
+The BUM function $q$ then is then used to interpolate a weight for any normalized sort position via $\lambda_i = q(\frac{\pi(i)}{n}) - q(\frac{\pi(i - 1)}{n})$.
 Therefore the learning objective boils down to optimizing the shape of $q$.
@@
 As previously mentioned, the addition of relations between constituents turns LTA into a graph regression problem.
-Most of the recent research in the field of learning from graph structured data however focused on the closely related graph classification problem.
-Since many ideas from the classification setting are also useful in the regression setting, a brief overview of those ideas is given.
+Most of the recent research in the field of learning from graph structured data has however focused on the closely related graph classification problem.
+Since many ideas from the classification setting are also useful in the regression setting, a brief overview of those ideas is given first.
 
@@
 			\item \textbf{Spatial variant:}
-				Standard CNNs use a convolution operator that is applied to a grid graph with diagonal edges.
-				Spatial ConvGNNs~\cite{Micheli2009} directly generalize this idea by defining a more flexible convolution that aggregates the local neighborhood of each vertex.
-				Spatial graph convolution aggregates the direct neighborhood of each vertex where each vertex $v_i$ is typically described by a feature vector $x_i \in \mathbb{R}^d$.
+				Standard CNNs use a convolution operator that is typically applied to a grid graph of pixels with added diagonal edges.
+				Spatial ConvGNNs~\cite{Micheli2009} directly generalize this idea by defining a more flexible convolution that works with arbitrary graph structures.
+				Spatial graph convolution aggregates the direct neighborhood of a vertex $v_i$, where $v_i$ is typically described by a feature vector $x_i \in \mathbb{R}^d$.
 				This process returns a new aggregate feature vector $x'_i$.
@@
 				Formally this can be expressed as
-				\begin{align*}
-					g *_G \bm{x} = U \hat{g}(\Lambda) U^{\top} \bm{x}
-				\end{align*}
-				where $\bm{x} \in \mathbb{R}^{n \times k}$ is the matrix of all vertex features and $U \Lambda U^{\top}$ is the eigendecomposition of the graph Laplacian $L_G$.
+				\begin{align}
+					g *_G \bm{x} = U \hat{g}(\Lambda) U^{\top} \bm{x}\label{eq:related-work:conv-filter}
+				\end{align}
+				where $\bm{x} \in \mathbb{R}^{n \times k}$ is the matrix of all vertex features and $U \Lambda U^{\top}$ is the eigendecomposition of the graph Laplacian $L_G$\footnote{
+					To see the connection to the convolutional theorem, note that $\mathcal{F} = U$ and $\mathcal{F}^{-1} = U^{\top}$ in \cref{eq:related-work:conv-filter} with $\mathcal{F}$ denoting the graph Fourier transform.
+				}.
 				In the previously described spatial ConvGNN variant, $g$ is learned directly in form of the feature aggregation function with a neighborhood locality constraint.
@@
 				This simplification of $\hat{g}$ implicitly causes its Fourier inverse $g$ to be locality constrained just like in the spatial ConvGNN variant.
-				Therefore the spectral GCN approach also allows for a spatial/message-passing interpretation.
+				Hence the spectral GCN approach also allows for a spatial/message-passing interpretation.
 				Recently various variants and extensions of GCNs~\cite{Hamilton2017}\cite{Chen2018}\cite{Chen2017}\cite{Chiang2019}\cite{Du2017} have been proposed that apply additional approximations and sampling methods to improve the runtime as well as the accuracy in certain scenarios.
@@
 				AGCNs do not use a fixed Laplacian $L$ for each input but instead learn a so-called residual Laplacian $L_\mathit{res}$.
-				The residual Laplacian is constructed by learning a vertex similarity kernel $\mathbb{G}_{x_i,x_j}$ which is used to determine the residual edge weights between vertices.
-				By overlaying the learned residual Laplacian $L_\mathit{res}$ on top of the static intrinsic Laplacian $L$, similar nodes with a large spatial distance in $L$ become connected.
+				The residual Laplacian is constructed by learning a vertex similarity kernel $\mathbb{G}_{x_i,x_j}$ which is used to determine edge weights between vertices.
+				By overlaying the learned residual Laplacian $L_\mathit{res}$ on top of the given intrinsic Laplacian $L$, similar nodes with a large spatial distance in $L$ become connected.
 				This can improve the accuracy especially if a locality constrained spectral filter $\hat{g}$ is used.
@@
 As mentioned in \cref{sec:motivation}, the overall goal of the thesis is to extend the LTA approach described in \cref{sec:related-work:lta} to the graph setting.
-Since this goal strongly overlaps with the work in graph classification described in \cref{sec:related-work:gc}, both branches of research should be combined.
-The practical applications described in \cref{sec:related-work:gr} could be used for comparison and as a potential source of a graph regression test dataset.
+Since this goal strongly overlaps with the work on graph classification described in \cref{sec:related-work:gc}, both branches of research should be combined.
+The practical applications described in \cref{sec:related-work:gr} could be used for comparison and as a potential source of graph regression training/test datasets.
 Now follows a concrete list of required and optional goals that should be achieved.
@@
 		The reference implementation should be used to evaluate the model on at least one test dataset.
-		Due to the scarcity of graph regression datasets, the used test data might be randomly generated.
-		Generated dataset should be obtained using a variety of representative deterministic graph scoring measures to check which kinds of local and global structural features are learnable;
+		Due to the relative scarcity of different graph regression datasets, the used test data might also be randomly generated.
+		A generated dataset should be obtained using a variety of representative deterministic graph scoring measures to check which kinds of local and global structural features are learnable;
 		some examples of measures that might be part of a graph scoring functions are the count of $k$-cliques in a graph, the graph diameter, the presence of cycles or various properties of a graph's degree distribution.
@@
 
-If possible, the previous goals could be extended.
+If time permits, the previous goals can be extended.
 For each of the previous five goal categories, optional additional ideas are listed below.
@@
 	\item \textbf{Disaggregation:}
-		As describen in \cref{sec:related-work:gc} there are multiple methods to learn vertex scores.
+		As described in \cref{sec:related-work:gc} there are multiple methods to learn vertex scores.
 		Depending on the chosen overall approach, a few of those methods might be suitable to be combined with an aggregation function learner.
@@
 	\item \textbf{Evaluation:}
-		Not only generated data but also real datasets should be used for evaluation (see \cref{sec:related-work:gr}).
+		If possible, real datasets should be used for evaluation (see \cref{sec:related-work:gr}) as an alternative or in addition to generated data.
+		Aspects like runtime performance and scalability could also be evaluated.
 \end{enumerate}
@@
 \label{sec:approach}
+
+As described in \cref{sec:motivation}, to extend LTA to graph inputs the aggregation and disaggregation problems have to be solved.
+Since the various approaches to learn vertex features described in \cref{sec:related-work:gc} can already be interpreted as solutions for the disaggregation problem, only a compatible vertex score aggregation function learning method has to be developed.
+This corresponds to learning a pooling function in ConvGNNs.
+The previously mentioned SortPooling and Self-Attention Pooling methods already do this in the classification setting.
+A method for the regression setting has to be found.
+
+To do so, training and test data is needed in the first step.
+Therefore one or multiple of the chemical toxicity datasets mentioned in \cref{sec:related-work:gr} should be prepared for the regression pipeline.
+This includes the following tasks:
+\begin{enumerate*}
+	\item The toxicity data has to be read from various formats (CSV, JSON or XML from a REST-API, chemical SD files, etc.),
+	\item serialized SMILES strings of molecular structure need to be translated into adjacency matrices\footnote{e.g.\ using RDKit (\url{https://github.com/rdkit/rdkit}), pysmiles (\url{https://github.com/pckroon/pysmiles}) or Open Babel (\url{http://openbabel.org})} and
+	\item the various units of measurement for toxicity need to be normalized.
+\end{enumerate*}
+In case the toxicity data turns out to be unsuitable, an artificial dataset might alternatively have to be generated as described in \cref{sec:goals:req}.
+Whether this is necessary will only become apparent after the first evaluations which could show that the chosen approach is completely ill suited to the toxicity prediction task.
+
+In the next step the development of a vertex aggregation method can begin.
+For the disaggregation problem the open source reference implementation of GCNs\footnote{\url{https://github.com/tkipf/gcn}} or more generally the Graph Nets library\footnote{\url{https://github.com/deepmind/graph_nets}} for TensorFlow might be useful.
+The development should start by building a simple ConvGNN with a fixed aggregation function.
+Then an aggregation function learner can be added.
+Since the existing GNN tools and reference implementations are mostly available in Python, the implementation part of this thesis should probably also use this language.
+
+Lastly the chosen method should be evaluated using the dataset that was chosen in the first step.
+The results on the toxicity dataset should be compared with the results of selected approaches from \cref{sec:related-work:gr}.
+Additionally the learned aggregation functions should be compared with na{\"\i}ve aggregation functions like $\min$, $\max$ or $\text{avg}$ since those aggregation functions are typically used in the pooling layer of current ConvGNN approaches.
 
@@
 	\item Introduction
-	\item
+	\item Previous Work
+		\begin{enumerate}[label*=\arabic*.]
+			\item Learning to Aggregate
+			\item Convolutional Graph Neural Networks
+		\end{enumerate}
+	\item Learning to Aggregate Graphs
+	\begin{enumerate}[label*=\arabic*.]
+		\item A Vertex Aggregation Function Learner
+		\item Compatible Disaggregation Methods
+	\end{enumerate}
 	\item Evaluation
+		\begin{enumerate}[label*=\arabic*.]
+			\item Learning Characteristics of Generated Graphs
+			\item Learning Chemical Toxicity
+			\item Scalability
+		\end{enumerate}
 	\item Conclusion
@@
 	\centering
-	%\includegraphics[width=.9\textwidth]{timeschedule}
+	\begin{ganttchart}[
+		x unit = 0.55cm,
+		y unit title = 0.6cm,
+		y unit chart = 0.7cm,
+		vgrid,
+		canvas/.style = {
+			draw = none
+		},
+		title label font = \sffamily,
+		title height = 1,
+		title/.style = {
+			draw = none,
+			fill = none
+		},
+		group label font = \sffamily\bfseries,
+		group label node/.append style = {
+			left = 0.2cm
+		},
+		group/.append style = {
+			fill = blau,
+			draw = none
+		},
+		group peaks tip position = 0,
+		bar label font = \sffamily\bfseries,
+		bar label node/.append style = {
+			left = 0.2cm
+		},
+		bar/.append style = {
+			fill = blau,
+			draw = none
+		}
+	]{1}{20}
+		\gantttitle{week}{20} \\
+		\gantttitlelist{1,...,20}{1} \\ % ignore triple dots: chktex 11
+
+		\ganttgroup{Implementation}{1}{11} \\
+		\ganttbar[bar label font = \sffamily, bar/.append style = {fill = hellblau}]
+			{Data Preparation}{1}{1} \\
+		\ganttbar[bar label font = \sffamily, bar/.append style = {fill = hellblau}]
+			{Basic ConvGNN}{2}{4} \\
+		\ganttbar[bar label font = \sffamily, bar/.append style = {fill = hellblau}]
+			{Aggregation Learner}{5}{8} \\
+			\ganttbar[bar label font = \sffamily, bar/.append style = {fill = hellgrau}]
+				{Buffer}{9}{9} \\
+		\ganttbar[bar label font = \sffamily, bar/.append style = {fill = hellblau}]
+			{Evaluation}{10}{11} \\
+
+		\ganttgroup{Writing}{2}{19} \\
+		\ganttbar[bar label font = \sffamily, bar/.append style = {fill = hellblau}]
+			{Previous Work}{2}{5} \\
+		\ganttbar[bar label font = \sffamily, bar/.append style = {fill = hellblau}]
+			{L.t.A. Graphs}{6}{10} \\
+		\ganttbar[bar label font = \sffamily, bar/.append style = {fill = hellblau}]
+			{Evaluation}{11}{13} \\
+		\ganttbar[bar label font = \sffamily, bar/.append style = {fill = hellblau}]
+			{Conclusion}{14}{14} \\
+		\ganttbar[bar label font = \sffamily, bar/.append style = {fill = hellgrau}]
+			{Buffer}{15}{17} \\
+		\ganttbar[bar label font = \sffamily, bar/.append style = {fill = hellblau}]
+			{Corrections}{18}{19} \\
+
+		\ganttbar{Presentation}{19}{20}
+	\end{ganttchart}
 	\caption{Sketch of the time schedule for the work on the thesis}\label{fig:schedule}

### 2019-10-20 proposal/paper.tex
minor proposal fixes

@@
 To deal with varying composite sizes $n$ the weights $\lambda_i$ are interpolated using a \textit{basic unit interval monotone} (BUM) function $q: [0, 1] \to [0, 1]$.
-It takes constituent positions that are normalized to the unit interval, i.e.\ $\frac{\pi(i)}{n}$.
-The BUM function $q$ then is then used to interpolate a weight for any normalized sort position via $\lambda_i = q(\frac{\pi(i)}{n}) - q(\frac{\pi(i - 1)}{n})$.
+It takes constituent positions that are normalized to the unit interval, i.e.\ $\frac{i}{n}$.
+The BUM function $q$ then is then used to interpolate a weight for any normalized sort position via $\lambda_i = q(\frac{i}{n}) - q(\frac{i - 1}{n})$.
 Therefore the learning objective boils down to optimizing the shape of $q$.
@@
 		To obtain an aggregate score for the entire graph, a graph pooling layer is applied.
-		A primitive pooling approach is to simply merge the vertex features via a fixed aggregation function like $\min$, $\max$ or $+$.
+		A primitive pooling approach is to simply merge the vertex features via a fixed aggregation function like $\min$, $\max$ or $\text{avg}$.
 		Better results can be obtained by learning the pooling aggregation function.
@@
 		The reference implementation should be used to evaluate the model on at least one test dataset.
-		Due to the relative scarcity of different graph regression datasets, the used test data might also be randomly generated.
+		Due to the relative scarcity of different graph regression datasets, the used dataset might also be synthetically generated.
 		A generated dataset should be obtained using a variety of representative deterministic graph scoring measures to check which kinds of local and global structural features are learnable;
@@
 
-Lastly the chosen method should be evaluated using the dataset that was chosen in the first step.
+Lastly the developed method should be evaluated using the dataset that was chosen in the first step.
 The results on the toxicity dataset should be compared with the results of selected approaches from \cref{sec:related-work:gr}.

### 2019-10-24 proposal/paper.tex
updated related work + minor fixes

@@
 As the logical expression that thus effectively aggregates the constituents, a uninorm $U_{\lambda}$ is used.
-Depending on the parameter $\lambda$, $U_{\lambda}$ interpolates between t-norms and t-conorms which are continuous generalizations of logical conjunction and disjunction respectively.
+Depending on the parameter $\lambda$, $U_{\lambda}$ combines t-norms and t-conorms which are continuous generalizations of logical conjunction and disjunction respectively.
 
@@
 
-\subsection{Graph Classification}%
-\label{sec:related-work:gc}
+\subsection{Graph Classification and Regression}%
+\label{sec:related-work:gcr}
 
@@
 		\end{enumerate}
+
+		The just described ConvGNN approaches generally assume that an input consists of an undirected graph with positive real edge weights.
+		Those approaches are not directly applicable in scenarios with multiple edge types, i.e.\ multi-relational scenarios.
+		One class of spatial multi-relational GNN methods was proposed by \citet{Battaglia2018}.
+		There a shared function $\phi^e$ is used to describe how a vertex feature vector $v_i$ is transformed as it passes through an edge with features $e_{ij} \in \mathbb{R}^p$.
+		In a uni-relational spatial GNN with $e_{ij} \in \mathbb{R}^{1}$, typically $\phi^e(e_{ij}, v_i) = e_k \cdot v_i$ is used.
+		In the multi-relational setting, $\phi^e$ can instead be learned via a standard \textit{multilayer perceptron} (MLP).
+		An alternative spectral multi-relational ConvGNN approach was proposed by \citet{Gong2018}.
+		Their so-called \textit{edge enhanced graph neural network} (EGNN) essentially models each relation as a separate graph with its own Laplacian.
+		A convolutional filter $g$ is shared across all relations and used to convolve the input signal separately via each relation's Laplacian.
+		To share vertex features across relations, the separately convolved signals of all relations are then concatenated.
+		Those concatenated vertex features are used as the input signal of the next layer.
+
+		This concludes the short overview of approaches to learn vertex feature vectors via ConvGNNs.
 		After the application of either spatial or spectral graph convolutions, each vertex will have an updated aggregate feature vector.
@@
 				WL colors are lexicographically sortable signatures that encode the structural role of vertices; they can be used for efficient graph isomorphism falsification~\cite{Weisfeiler1968}.
-				Using this interpretation of the vertex feature vectors, SortPooling determines a top-$k$ ranking of the vertices' WL colors and then aggregates the thus selected fixed-size vertex feature subset via a standard \textit{multilayer perceptron} (MLP).
+				Using this interpretation of the vertex feature vectors, SortPooling determines a top-$k$ ranking of the vertices' WL colors and then aggregates the thus selected fixed-size vertex feature subset via a MLP.\@
 			\item Self-Attention Pooling adds an additional graph convolution layer after an arbitrary ConvGNN.\@
@@
 
-\subsection{Graph Regression}%
-\label{sec:related-work:gr}
+While all of the previously mentioned problems were evaluated on graph classification problems, they can naturally be extended to regression tasks.
+In the case of ConvGNNs for example, the described pooling layers already produce continuous outputs.
 
-An overview of the recent work regarding graph classification was just given.
-The generalization of LTA to the graph domain is however better described as a graph regression problem since the predicted composite scores are continuous.
-As there currently are relatively few graph regression approaches, only few references can be provided here.
+The pooling aggregation functions used in ConvGNNs are however typically not part of the learning objective.
+This shows the fundamental methodological difference between GNNs and LTA.\@
+A GNN uses the relational structure of its input to try to determine meaningful feature representations or scores for each vertex.
+Those vertex scores are then combined using a simple aggregation function.
+LTA on the other hand currently does not use any relational structure and therefore computes vertex/constituent scores independently.
+Those comparatively less meaningful local scores are then combined using a more expressive learned aggregation function.
+Generally speaking current GNN approaches focus on the disaggregation problem while LTA approaches focus on the aggregation problem.
+The described Self-Attention Pooling~\cite{Lee2019} can be seen as an exception to this generalization since it learns an aggregation function in form of the self-attention vertex weights.
+
+\subsection{Molecular Analysis Problems and Datasets}%
+\label{sec:related-work:tox}
 
@@
 One of those domains is the prediction of chemical toxicity, typically the $\text{LD}_{50}$ or $\text{LC}_{50}$ doses which measure the smallest amount of a given chemical that is lethal for at least $50\%$ of tested subjects.
+Another regression task from the domain of molecular analysis is the prediction of compound solubility.
 The following approaches could serve as inspiration, baselines and/or sources of training data:
@@
 		It works similarly to RASAR but leaves out the logistic regression step and instead directly aggregates the toxicity of similar chemicals via $k$-means.
-		ProTox uses the freely available \textit{SuperToxic} dataset~\cite{Schmidt2009}.
+		ProTox uses the freely available \textit{SuperToxic} dataset~\cite{Schmidt2009} which contains 13400 molecules with known $\text{LD}_{50}$ values.
 		While the implementation is again not publicly available, the trained model can be freely queried online\footnote{ProTox-II, \url{http://tox.charite.de/protox_II/}}.
@@
 		Like RASAR and ProTox it includes similarity-based prediction models but also linear regression and decision tree models that are trained on vector representations of chemicals.
-		The \textit{ChemIDplus}\footnote{\url{https://chem.nlm.nih.gov/chemidplus/}} database was used for training.
-\end{itemize}
-
-Recently first toxicity prediction approaches using ConvGNNs have been published.
+		The publicly available \textit{ChemIDplus}\footnote{\url{https://chem.nlm.nih.gov/chemidplus/}} database was used for training.
+		It contains 28023 and 13541 molecules with known oral $\text{LD}_{50}$ values for rats and mice respectively.
+		Additionally about 1000 molecules with known $\text{LC}_{50}$ values are available for both species.
+	\item Recently first toxicity prediction approaches using ConvGNNs have been published.
 		The previously mentioned AGCN~\cite{Li2018} was in fact evaluated on a discrete toxicity classification problem;
@@
 		there a functional group refers to a relevant subgraph of a molecule that can be interpreted as a toxicity indicator.
+	\item The previously described EGNN method~\cite{Gong2018} for multi-relational graphs was evaluated on a compound solubility prediction task.
+		For this the Lipophilicity~\cite{Wu2017} and FreeSolv~\cite{Mobley2014} datasets were used.
+		They contain 4200 and 642 labeled molecular graphs respectively.
+\end{itemize}
 
@@
 As mentioned in \cref{sec:motivation}, the overall goal of the thesis is to extend the LTA approach described in \cref{sec:related-work:lta} to the graph setting.
-Since this goal strongly overlaps with the work on graph classification described in \cref{sec:related-work:gc}, both branches of research should be combined.
-The practical applications described in \cref{sec:related-work:gr} could be used for comparison and as a potential source of graph regression training/test datasets.
+Since this goal strongly overlaps with the work on graph classification described in \cref{sec:related-work:gcr}, both branches of research should be combined.
+The practical applications described in \cref{sec:related-work:tox} could be used for comparison and as a potential source of graph regression training/test datasets.
 Now follows a concrete list of required and optional goals that should be achieved.
@@
 	\item \textbf{Disaggregation:}
-		As described in \cref{sec:related-work:gc} there are multiple methods to learn vertex scores.
+		As described in \cref{sec:related-work:gcr} there are multiple methods to learn vertex scores.
 		Depending on the chosen overall approach, a few of those methods might be suitable to be combined with an aggregation function learner.
@@
 	\item \textbf{Evaluation:}
-		If possible, real datasets should be used for evaluation (see \cref{sec:related-work:gr}) as an alternative or in addition to generated data.
+		If possible, real datasets should be used for evaluation (see \cref{sec:related-work:tox}) as an alternative or in addition to generated data.
 		Aspects like runtime performance and scalability could also be evaluated.
@@
 As described in \cref{sec:motivation}, to extend LTA to graph inputs the aggregation and disaggregation problems have to be solved.
-Since the various approaches to learn vertex features described in \cref{sec:related-work:gc} can already be interpreted as solutions for the disaggregation problem, only a compatible vertex score aggregation function learning method has to be developed.
+Since the various approaches to learn vertex features described in \cref{sec:related-work:gcr} can already be interpreted as solutions for the disaggregation problem, only a compatible vertex score aggregation function learning method has to be developed.
 This corresponds to learning a pooling function in ConvGNNs.
@@
 To do so, training and test data is needed in the first step.
-Therefore one or multiple of the chemical toxicity datasets mentioned in \cref{sec:related-work:gr} should be prepared for the regression pipeline.
+Therefore one or multiple of the chemical toxicity datasets mentioned in \cref{sec:related-work:tox} should be prepared for the regression pipeline.
 This includes the following tasks:
@@
 Lastly the developed method should be evaluated using the dataset that was chosen in the first step.
-The results on the toxicity dataset should be compared with the results of selected approaches from \cref{sec:related-work:gr}.
+The results on the toxicity dataset should be compared with the results of selected approaches from \cref{sec:related-work:tox}.
 Additionally the learned aggregation functions should be compared with na{\"\i}ve aggregation functions like $\min$, $\max$ or $\text{avg}$ since those aggregation functions are typically used in the pooling layer of current ConvGNN approaches.
@@
 
-\vspace{6cm}
+\null\vfill
 

### 2019-10-26 proposal/paper.tex
different topic candidates

@@
 		One class of spatial multi-relational GNN methods was proposed by \citet{Battaglia2018}.
+		They describe a general framework for the construction of GNNs; for simplicity only the class message-passing variants is mentioned here.
 		There a shared function $\phi^e$ is used to describe how a vertex feature vector $v_i$ is transformed as it passes through an edge with features $e_{ij} \in \mathbb{R}^p$.
@@
 	\item The previously described EGNN method~\cite{Gong2018} for multi-relational graphs was evaluated on a compound solubility prediction task.
-		For this the Lipophilicity~\cite{Wu2017} and FreeSolv~\cite{Mobley2014} datasets were used.
+		For this the Lipophil\-icity~\cite{Wu2017} and FreeSolv~\cite{Mobley2014} datasets were used.
 		They contain 4200 and 642 labeled molecular graphs respectively.
@@
 
-\section{Goals}%
-\label{sec:goals}
+\section{Research Question}%
+\label{sec:question}
 
-As mentioned in \cref{sec:motivation}, the overall goal of the thesis is to extend the LTA approach described in \cref{sec:related-work:lta} to the graph setting.
-Since this goal strongly overlaps with the work on graph classification described in \cref{sec:related-work:gcr}, both branches of research should be combined.
-The practical applications described in \cref{sec:related-work:tox} could be used for comparison and as a potential source of graph regression training/test datasets.
-Now follows a concrete list of required and optional goals that should be achieved.
+\todo[inline,color=green]{Since the final research question is not yet fixed, I only sketched out the main candidates.}
 
-\subsection{Required Goals}%
-\label{sec:goals:req}
+\todo[inline]{Candidate 1: Comparison of LTA and GNNs}
+As described, current graph classification and regression approaches mostly use simple fixed aggregation functions like $\min$, $\max$ or $\text{avg}$.
+Self-Attention Pooling~\cite{Lee2019} is a notable exception that learns the weights for a weighted average that is computed over convolved vertex features.
+An open question for research is whether more complex aggregation functions are able to outperform simple aggregation functions in the graph setting.
+\todo[inline,color=green]{Not sure how to approach this yet.}
 
-\begin{enumerate}[label=\textbf{\arabic*.}]
-	\item \textbf{Aggregation:}
-		Define a class of aggregation functions that takes vertex scores as its input.
-		It should fulfill the role of the uninorms and OWA functions in the existing LTA papers (see \cref{sec:related-work:lta}).
-	\item \textbf{Disaggregation:}
-		Describe how vertex features should be used to compute a constituent score that will be combined by the aggregation function.
-	\item \textbf{Learning:}
-		Combine the chosen aggregation and disaggregation methods into a joint LTA model with a suitable loss function.
-	\item \textbf{Implementation:}
-		A reference implementation for the designed LTA model should be developed.
-	\item \textbf{Evaluation:}
-		The reference implementation should be used to evaluate the model on at least one test dataset.
-		Due to the relative scarcity of different graph regression datasets, the used dataset might also be synthetically generated.
-		A generated dataset should be obtained using a variety of representative deterministic graph scoring measures to check which kinds of local and global structural features are learnable;
-		some examples of measures that might be part of a graph scoring functions are the count of $k$-cliques in a graph, the graph diameter, the presence of cycles or various properties of a graph's degree distribution.
-		The results should be compared with the results of at least one other approach for learning graph scores, e.g.\ a ConvCNN with a fixed aggregation function instead of a learned one.
+\todo[inline]{Candidate 2: Structural Restrictions}
+The previously described graph classification and regression methods assume that arbitrary graphs are provided as input.
+The question this thesis aims to answer is whether a restriction of the input graph structure makes it possible to use a more specialized method that outperforms the more general approaches.
+Possible restrictions to look at could for example be vertex degree or tree structure constraints.\todo[inline,color=green]{
+	I suspect that such restrictions will mostly improve upon runtime (either automatically because most GNN methods depend on $\mathcal{O}(\left|E\right|)$ or because expensive matrix multiplications can be replaced).
+	Those runtime savings might then of course be translated into regression performance improvements, e.g.\ maybe by using a deeper model.
+	As we discussed a restriction to tree structures may make hierarchical aggregation functions feasible but as I understand those can in principle be expressed via a message-passing GNN as well~\cite{Battaglia2018}.
+}
+
+\todo[inline]{Candidate 3: Multi-relational ConvGNNs}
+As mentioned in \cref{sec:related-work:gcr}, most ConvGNN variants are not directly applicable to multi-relational inputs.
+Two more flexible approaches were described:
+\begin{enumerate}
+	\item The general graph network framework described by \citet{Battaglia2018} that can be interpreted as a spatial message-passing approach.
+		They share information across relations via an arbitrary learned function $\phi^e(e_{ij}, e_i)$.
+	\item EGNNs~\cite{Gong2018}, a spectral ConvGNN variant.
+		They share information across relations by concatenating vertex features that are separately convolved in each relation's graph.
 \end{enumerate}
-
-\subsection{Optional Goals}%
-\label{sec:goals:opt}
-
-If time permits, the previous goals can be extended.
-For each of the previous five goal categories, optional additional ideas are listed below.
-\begin{enumerate}[label=\textbf{\arabic*.}]
-	\item \textbf{Aggregation:}
-		Multiple classes of aggregation functions could be considered and compared.
-	\item \textbf{Disaggregation:}
-		As described in \cref{sec:related-work:gcr} there are multiple methods to learn vertex scores.
-		Depending on the chosen overall approach, a few of those methods might be suitable to be combined with an aggregation function learner.
-		Those suitable combinations could be evaluated and compared.
-	\item \textbf{Learning:}
-		The joint model of aggregation and disaggregation function learning can most likely be optimized using various algorithms.
-		Those algorithms could be compared.
-		In case their runtime performance is impractical in certain situations, approximations and simplifications of the model could be considered.
-	\item \textbf{Implementation:}
-		The chosen approach should ideally align well with the practical constraints of implementation.
-		It should be suitable for an implementation on modern GPUs and/or distributed cluster environments.
-	\item \textbf{Evaluation:}
-		If possible, real datasets should be used for evaluation (see \cref{sec:related-work:tox}) as an alternative or in addition to generated data.
-		Aspects like runtime performance and scalability could also be evaluated.
-\end{enumerate}
+The first approach does not discuss which specific choice of edge propagation function $\phi^e$ is suitable for a multi-relational regression problem like those described in \cref{sec:related-work:tox}.
+In the second approach the concatenation operator implicitly acts as the edge propagation function.
+As we will see in \cref{sec:approach}, this assumes that every relation shares information with the other relations in the same way.
+Currently there are no approaches that explicitly model the correlation structure between relations, i.e.\ if and how information propagates from one relation to another.
+The research question for this thesis is to find out how a model that incorporates cross-relational correlations might look like and to compare it theoretically and empirically with another multi-relational method like the EGNN.\@
 
@@
 
-As described in \cref{sec:motivation}, to extend LTA to graph inputs the aggregation and disaggregation problems have to be solved.
-Since the various approaches to learn vertex features described in \cref{sec:related-work:gcr} can already be interpreted as solutions for the disaggregation problem, only a compatible vertex score aggregation function learning method has to be developed.
-This corresponds to learning a pooling function in ConvGNNs.
-The previously mentioned SortPooling and Self-Attention Pooling methods already do this in the classification setting.
-A method for the regression setting has to be found.
+\todo[inline,color=green]{As discussed on Monday, I added a rough overview of the formal background for candidate 3.}
+A possible approach to define a multi-relational ConvGNN is to directly extend an existing method, e.g.\ GCNs~\cite{Kipf2017}, from the one-dimensional to the multi-dimensional setting.
+Typically a graph is modeled via a symmetric adjacency matrix $E \in \mathbb{R}^{N \times N}$ with non-negative edge weights $e_{ij} = e_{ji} \in \mathbb{R}$.
+An edge weight $e_{ij}$ can be interpreted as a member of the multiplicative group on $\mathbb{R}$ and encodes a spatial signal strength scaling operation between the vertices $i$ and $j$.
+If we now go from a single relation to $M$ relations, the operator encoded by an edge can be generalized from $e_{ij} \in \mathbb{R}^{1 \times 1}$ to $\bm{e}_{ij} \in \mathbb{R}^{M \times M}$.
+The non-negativity constraint $e_{ij} \geq 0$ on the edge weights then generalizes to a positive semidefiniteness constraint $\bm{e}_{ij} \succeq 0$ on the edge matrices; therefore $\bm{e}_{ij}$ is symmetric.
+$\bm{e}_{ijkl} = \bm{e}_{ijlk} = \bm{e}_{jikl} = \bm{e}_{jilk} \in \mathbb{R}$ describes the connection strength between the signals of the relations $k$ and $l$ at vertex $i$ and the signals of the same relations at vertex $j$.
 
-To do so, training and test data is needed in the first step.
-Therefore one or multiple of the chemical toxicity datasets mentioned in \cref{sec:related-work:tox} should be prepared for the regression pipeline.
-This includes the following tasks:
-\begin{enumerate*}
-	\item The toxicity data has to be read from various formats (CSV, JSON or XML from a REST-API, chemical SD files, etc.),
-	\item serialized SMILES strings of molecular structure need to be translated into adjacency matrices\footnote{e.g.\ using RDKit (\url{https://github.com/rdkit/rdkit}), pysmiles (\url{https://github.com/pckroon/pysmiles}) or Open Babel (\url{http://openbabel.org})} and
-	\item the various units of measurement for toxicity need to be normalized.
-\end{enumerate*}
-In case the toxicity data turns out to be unsuitable, an artificial dataset might alternatively have to be generated as described in \cref{sec:goals:req}.
-Whether this is necessary will only become apparent after the first evaluations which could show that the chosen approach is completely ill suited to the toxicity prediction task.
+Using multi-dimensional edge operators as described above turns $E \in \mathbb{R}^{(N \cdot M) \times (N \cdot M)} \equiv {(\mathbb{R}^{M \times M})}^{N \times N} \equiv \mathbb{R}^{N \times N} \otimes \mathbb{R}^{M \times M}$ into a $(2,2)$-tensor.
+The notion of a one-dimensional graph signal $x \in \mathbb{R}^{N}$ then generalizes to $x \in \mathbb{R}^{N \cdot M} \equiv {(\mathbb{R}^M)}^N \equiv \mathbb{R}^N \otimes \mathbb{R}^M$, a $(2, 0)$-tensor that describes the signal strength of each relation at each vertex.
 
-In the next step the development of a vertex aggregation method can begin.
-For the disaggregation problem the open source reference implementation of GCNs\footnote{\url{https://github.com/tkipf/gcn}} or more generally the Graph Nets library\footnote{\url{https://github.com/deepmind/graph_nets}} for TensorFlow might be useful.
-The development should start by building a simple ConvGNN with a fixed aggregation function.
-Then an aggregation function learner can be added.
-Since the existing GNN tools and reference implementations are mostly available in Python, the implementation part of this thesis should probably also use this language.
+Another way to look at the resulting adjacency structure is as a linear operator $E: \mathbb{R}^{N \cdot M} \to \mathbb{R}^{N \cdot M}$ that acts on the $N$-dimensional module over the division ring $R$ of positive semidefinite real $M \times M$ matrices.\todo{TODO:\@ Formalize the inverse and missing covariant factor.}
 
-Lastly the developed method should be evaluated using the dataset that was chosen in the first step.
-The results on the toxicity dataset should be compared with the results of selected approaches from \cref{sec:related-work:tox}.
-Additionally the learned aggregation functions should be compared with na{\"\i}ve aggregation functions like $\min$, $\max$ or $\text{avg}$ since those aggregation functions are typically used in the pooling layer of current ConvGNN approaches.
+To motivate a ConvGNN on this more abstract definition of $E$ we first define the multi-relational combinatorial Laplacian $\mathcal{L} := D - E$ via $D_{ii} := \sum \bm{e}_{ii}$ and $D_{ij} := \bm{0}$ for $i \neq j$, i.e.\ $D$ is a block matrix with summed $M \times M$ blocks on its diagonal and zeros everywhere else.
+The normalized multi-relational Laplacian can then be defined as $L := I - D^{-\frac{1}{2}} E D^{-\frac{1}{2}}$ using the fact that all $D_{ii}$ are elements of the division ring $R$.
+A spectral filter $\hat{g}$ can therefore be learned for $L$, analogous to GCNs~\cite{Kipf2017}.
+
+The main remaining question now is how the entries of the edge operators $\bm{e}_{ij} \in \mathbb{R}^{M \times M}$ are determined.
+In uni-relational methods the scalar edge weight $e_{ij} \in \mathbb{R}$ is typically assumed to be given for each input.
+In the multi-relational setting we can assume that a scalar connection strength is given for each relation, i.e.\ the self-correlations $\bm{e}_{ijkk} \in \mathbb{R}$ on the diagonal are known.
+The off-diagonal entries of an edge operator however represent the cross-relational correlations and cannot be assumed to be known.
+If we set the off-diagonal entries to zero, i.e.\ $\forall i, j, k, l: k \neq l \rightarrow e_{ijkl} = 0$, all relations would be assumed to be uncorrelated and $M$ spectral filters would be learned independently on the graphs that are described by the relations.\todo{TODO:\@ Add a formal proof?}
+
+Since we cannot generally assume that all relations are uncorrelated, a \textit{relational correlation kernel} $r$ has to be learned together with the spectral filter $\hat{g}$.
+$r$ should describe how strong one relation is correlated to another and is used to compute off-diagonal edge operator entries.
+
+The previously mentioned EGNNs~\cite{Gong2018} can in fact be interpreted using the notion of edge operators.
+They implicitly assume that the rows in all $e_{ij}$ are equal, i.e.\ the off-diagonal entries are set to\todo{TODO:\@ Explain this.}
+\begin{align*}
+	\begin{pmatrix}
+		\bm{e}_{ij11} & & \\
+		& \ddots & \\
+		& & \bm{e}_{ijMM}
+	\end{pmatrix} \rightarrow \begin{pmatrix}
+		\bm{e}_{ij11} & \cdots & \bm{e}_{ijMM} \\
+		\vdots & \vdots & \vdots \\
+		\bm{e}_{ij11} & \cdots & \bm{e}_{ijMM}
+	\end{pmatrix}\text{.}
+\end{align*}
+A more flexible correlation structure can potentially yield an improvement over the current multi-relational approaches.
+
+One problem that arises from the described approach is that the adjacency tensor $E$ is changed after each change of $r$ and therefore $\mathcal{L}$ needs to be renormalized after every training step.
+This can potentially incur a significant performance cost and should be mitigated by employing results from perturbation theory.\todo{TODO:\@ Elaborate.}
 

### 2019-10-28 proposal/paper.tex
first draft of chosen research question & approach

--- 2019-10-24 proposal/paper.tex
@@
 \usepackage{graphicx}
-\graphicspath{{images/}}
+\graphicspath{{gfx/}}
 
@@
 
-\section{Goals}%
-\label{sec:goals}
+\section{Research Questions}%
+\label{sec:question}
 
-As mentioned in \cref{sec:motivation}, the overall goal of the thesis is to extend the LTA approach described in \cref{sec:related-work:lta} to the graph setting.
-Since this goal strongly overlaps with the work on graph classification described in \cref{sec:related-work:gcr}, both branches of research should be combined.
-The practical applications described in \cref{sec:related-work:tox} could be used for comparison and as a potential source of graph regression training/test datasets.
-Now follows a concrete list of required and optional goals that should be achieved.
+When looking only at the input and output of the described LTA and graph methods, they share certain similarities.
+Both take a variable size and permutation invariant set of constituents/vertices as input and compute an aggregate class or score from it.
+There is however a difference between LTA and the described graph methods in the way aggregate scores are computed.
 
-\subsection{Required Goals}%
-\label{sec:goals:req}
+LTA approaches are based on the assumption that a composite score can be modeled as an aggregate of constituent scores where constituent and composite scores are required to be from the same domain, e.g.\ $\mathcal{Y} = [0, 1]$.
+One consequence of this assumption is that an aggregate score is explainable by looking at the constituents and their associated scores.
 
-\begin{enumerate}[label=\textbf{\arabic*.}]
-	\item \textbf{Aggregation:}
-		Define a class of aggregation functions that takes vertex scores as its input.
-		It should fulfill the role of the uninorms and OWA functions in the existing LTA papers (see \cref{sec:related-work:lta}).
-	\item \textbf{Disaggregation:}
-		Describe how vertex features should be used to compute a constituent score that will be combined by the aggregation function.
-	\item \textbf{Learning:}
-		Combine the chosen aggregation and disaggregation methods into a joint LTA model with a suitable loss function.
-	\item \textbf{Implementation:}
-		A reference implementation for the designed LTA model should be developed.
-	\item \textbf{Evaluation:}
-		The reference implementation should be used to evaluate the model on at least one test dataset.
-		Due to the relative scarcity of different graph regression datasets, the used dataset might also be synthetically generated.
-		A generated dataset should be obtained using a variety of representative deterministic graph scoring measures to check which kinds of local and global structural features are learnable;
-		some examples of measures that might be part of a graph scoring functions are the count of $k$-cliques in a graph, the graph diameter, the presence of cycles or various properties of a graph's degree distribution.
-		The results should be compared with the results of at least one other approach for learning graph scores, e.g.\ a ConvCNN with a fixed aggregation function instead of a learned one.
-\end{enumerate}
+A graph classification or regression method on the other hand does not necessarily output explainable classes or scores.
+ConvGNNs for example perform multiple neighborhood feature aggregations around each vertex and then combine the aggregated vertex features into a final output using either a fixed aggregation function (e.g.\ $\min$, $\max$ or $\text{avg}$) or a standard predictor that gets a fixed-size top-$k$ ranking of the vertex features.
+A given aggregate score is generally not explained by the presence or absence of certain constituents.
 
-\subsection{Optional Goals}%
-\label{sec:goals:opt}
+Intuitively the difference between LTA and current graph methods can be formulated as \textit{first evaluate, then aggregate} versus \textit{first aggregate, then evaluate}, see \cref{fig:gnnvslta}.
+This general difference is however not a strict separator of LTA and current graph methods.
+As described by \citet{Pope2018} for example, relevant subgraphs with a high influence on the output of a ConvGNN can be identified and interpreted similarly to LTA constituents.
+The line between both approaches is therefore blurry.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.9\linewidth]{GNNvsLTA.pdf}
+	\caption{Intuition for the difference between ConvGNNs and LTA methods.}\label{fig:gnnvslta}
+\end{figure}
 
-If time permits, the previous goals can be extended.
-For each of the previous five goal categories, optional additional ideas are listed below.
-\begin{enumerate}[label=\textbf{\arabic*.}]
-	\item \textbf{Aggregation:}
-		Multiple classes of aggregation functions could be considered and compared.
-	\item \textbf{Disaggregation:}
-		As described in \cref{sec:related-work:gcr} there are multiple methods to learn vertex scores.
-		Depending on the chosen overall approach, a few of those methods might be suitable to be combined with an aggregation function learner.
-		Those suitable combinations could be evaluated and compared.
-	\item \textbf{Learning:}
-		The joint model of aggregation and disaggregation function learning can most likely be optimized using various algorithms.
-		Those algorithms could be compared.
-		In case their runtime performance is impractical in certain situations, approximations and simplifications of the model could be considered.
-	\item \textbf{Implementation:}
-		The chosen approach should ideally align well with the practical constraints of implementation.
-		It should be suitable for an implementation on modern GPUs and/or distributed cluster environments.
-	\item \textbf{Evaluation:}
-		If possible, real datasets should be used for evaluation (see \cref{sec:related-work:tox}) as an alternative or in addition to generated data.
-		Aspects like runtime performance and scalability could also be evaluated.
+To address this, the thesis should try to answer three central questions:
+\begin{enumerate}
+	\item \textit{What are the essential characteristics of a Learning to Aggregate method?}
+		Those characteristics should formally capture the differences and similarities between LTA and convolutional or kernel based graph methods.
+		Currently there is no comprehensive formulation of the relation between both fields of research;
+		this should be addressed by the first question.
+	\item \textit{How could a Learning to Aggregate method, that has the previously defined characteristics, look like for graph inputs?}
+		Currently LTA has only been applied to sets of constituents and not yet to graph structured data;
+		this should be addressed by the second question.
+	\item \textit{What are the assumptions of graph LTA methods and under which circumstances are they applicable?}
+		This aims to clarify in which kinds of scenarios the use of LTA is justified and expected to improve results compared to more general methods.
 \end{enumerate}
@@
 
-As described in \cref{sec:motivation}, to extend LTA to graph inputs the aggregation and disaggregation problems have to be solved.
-Since the various approaches to learn vertex features described in \cref{sec:related-work:gcr} can already be interpreted as solutions for the disaggregation problem, only a compatible vertex score aggregation function learning method has to be developed.
-This corresponds to learning a pooling function in ConvGNNs.
-The previously mentioned SortPooling and Self-Attention Pooling methods already do this in the classification setting.
-A method for the regression setting has to be found.
+To answer the three research questions, the following approach will be taken.
+First the notion of composites and constituents should be generalized based on the definitions found in the previous works on LTA~\cite{Melnikov2016}\cite{Melnikov2019}, so that they are also applicable to graph inputs.
+Next the notion of an aggregation function has to be defined and distinguished from the more general concept of an evaluation function (compare \cref{fig:gnnvslta}).
 
-To do so, training and test data is needed in the first step.
-Therefore one or multiple of the chemical toxicity datasets mentioned in \cref{sec:related-work:tox} should be prepared for the regression pipeline.
-This includes the following tasks:
-\begin{enumerate*}
-	\item The toxicity data has to be read from various formats (CSV, JSON or XML from a REST-API, chemical SD files, etc.),
-	\item serialized SMILES strings of molecular structure need to be translated into adjacency matrices\footnote{e.g.\ using RDKit (\url{https://github.com/rdkit/rdkit}), pysmiles (\url{https://github.com/pckroon/pysmiles}) or Open Babel (\url{http://openbabel.org})} and
-	\item the various units of measurement for toxicity need to be normalized.
-\end{enumerate*}
-In case the toxicity data turns out to be unsuitable, an artificial dataset might alternatively have to be generated as described in \cref{sec:goals:req}.
-Whether this is necessary will only become apparent after the first evaluations which could show that the chosen approach is completely ill suited to the toxicity prediction task.
+The resulting general definition of LTA methods should then be used for a comparison with a representative selection of existing graph methods (see \cref{sec:related-work:gcr}).
+It should be stated how both fields of research are related and under which circumstances a graph method can be interpreted as an LTA method.
+When formulating a graph method from the LTA perspective, it must be explicitly stated how the constituents and the aggregation function looks like.
+Depending on the graph method a constituent could consist of a single vertex but could also represent a certain subgraph or characteristic graph pattern.
 
-In the next step the development of a vertex aggregation method can begin.
-For the disaggregation problem the open source reference implementation of GCNs\footnote{\url{https://github.com/tkipf/gcn}} or more generally the Graph Nets library\footnote{\url{https://github.com/deepmind/graph_nets}} for TensorFlow might be useful.
-The development should start by building a simple ConvGNN with a fixed aggregation function.
-Then an aggregation function learner can be added.
-Since the existing GNN tools and reference implementations are mostly available in Python, the implementation part of this thesis should probably also use this language.
+The comparison between LTA and existing graph methods should serve as an overview of the state-of-the-art in graph classification and regression from the perspective of LTA.\@
+Based on this comparison, a specific graph LTA method or class of graph LTA methods should be developed.
+Depending on the interpretability of existing graph methods as LTA methods, the developed method will potentially be based on those existing methods.
 
-Lastly the developed method should be evaluated using the dataset that was chosen in the first step.
-The results on the toxicity dataset should be compared with the results of selected approaches from \cref{sec:related-work:tox}.
-Additionally the learned aggregation functions should be compared with na{\"\i}ve aggregation functions like $\min$, $\max$ or $\text{avg}$ since those aggregation functions are typically used in the pooling layer of current ConvGNN approaches.
+Finally the proposed graph LTA method should be evaluated.
+For this synthetically generated data or some of the datasets listed in \cref{sec:related-work:tox} could be used.
+If time permits, a comparison between the developed graph LTA method and other graph methods should be conducted to empirically demonstrate the potential advantages and disadvantages of the LTA assumptions.
 
@@
 		\end{enumerate}
-	\item Learning to Aggregate Graphs
+	\item Learning to Aggregate on Graphs
 	\begin{enumerate}[label*=\arabic*.]
-		\item A Vertex Aggregation Function Learner
-		\item Compatible Disaggregation Methods
+		\item Characteristics of LTA methods
+		\item An LTA perspective on GNNs
+		\item A Graph LTA Method
 	\end{enumerate}
 	\item Evaluation
-		\begin{enumerate}[label*=\arabic*.]
-			\item Learning Characteristics of Generated Graphs
-			\item Learning Chemical Toxicity
-			\item Scalability
-		\end{enumerate}
 	\item Conclusion
@@
 
-		\ganttgroup{Implementation}{1}{11} \\
+		\ganttgroup{Implementation}{2}{12} \\
 		\ganttbar[bar label font = \sffamily, bar/.append style = {fill = hellblau}]
-			{Data Preparation}{1}{1} \\
+			{Data Preparation}{2}{2} \\
 		\ganttbar[bar label font = \sffamily, bar/.append style = {fill = hellblau}]
-			{Basic ConvGNN}{2}{4} \\
+			{Graph LTA}{3}{8} \\
+			\ganttbar[bar label font = \sffamily, bar/.append style = {fill = hellgrau}]
+				{Buffer}{9}{10} \\
 		\ganttbar[bar label font = \sffamily, bar/.append style = {fill = hellblau}]
-			{Aggregation Learner}{5}{8} \\
-			\ganttbar[bar label font = \sffamily, bar/.append style = {fill = hellgrau}]
-				{Buffer}{9}{9} \\
+			{Evaluation}{11}{12} \\
+
+		\ganttgroup{Writing}{1}{19} \\
 		\ganttbar[bar label font = \sffamily, bar/.append style = {fill = hellblau}]
-			{Evaluation}{10}{11} \\
-
-		\ganttgroup{Writing}{2}{19} \\
+			{Previous Work}{1}{4} \\
 		\ganttbar[bar label font = \sffamily, bar/.append style = {fill = hellblau}]
-			{Previous Work}{2}{5} \\
-		\ganttbar[bar label font = \sffamily, bar/.append style = {fill = hellblau}]
-			{L.t.A. Graphs}{6}{10} \\
+			{Graph LTA}{5}{10} \\
 		\ganttbar[bar label font = \sffamily, bar/.append style = {fill = hellblau}]

### 2019-10-29 proposal/paper.tex
minor corrections

@@
 				In the previously described spatial ConvGNN variant, $g$ is learned directly in form of the feature aggregation function with a neighborhood locality constraint.
-				In the spectral variant however the Fourier-transformed filter $\hat{g}$ is learned.
+				In the spectral variant however, the Fourier-transformed filter $\hat{g}$ is learned.
 				Intuitively this means that vertex features $x_i$ are not aggregated with the features of their their direct neighborhood but with the features of vertices with which they share certain structural similarities.
-				This allows spectral ConvGNNs to incorporate the global graph structure at the cost of having to perform the computationally expensive decomposition of every input graph's Laplacian.
+				This allows spectral ConvGNNs to incorporate the global graph structure at the cost of having to perform the computationally expensive eigendecomposition of every input graph's Laplacian.
 
@@
 		One class of spatial multi-relational GNN methods was proposed by \citet{Battaglia2018}.
-		They describe a general framework for the construction of GNNs; for simplicity only the class message-passing variants is mentioned here.
+		They describe a general framework for the construction of GNNs; for simplicity only the class of message-passing variants is mentioned here.
 		There a shared function $\phi^e$ is used to describe how a vertex feature vector $v_i$ is transformed as it passes through an edge with features $e_{ij} \in \mathbb{R}^p$.
@@
 
-Intuitively the difference between LTA and current graph methods can be formulated as \textit{first evaluate, then aggregate} versus \textit{first aggregate, then evaluate}, see \cref{fig:gnnvslta}.
+Intuitively the difference between LTA and current graph methods can be formulated as \textit{first evaluate, then aggregate} versus \textit{first aggregate, then evaluate} (see \cref{fig:gnnvslta}).
 This general difference is however not a strict separator of LTA and current graph methods.
@@
 	\item \textit{What are the essential characteristics of a Learning to Aggregate method?}
-		Those characteristics should formally capture the differences and similarities between LTA and convolutional or kernel based graph methods.
+		Those characteristics should formally capture the differences and similarities between LTA and convolutional or kernel-based graph methods.
 		Currently there is no comprehensive formulation of the relation between both fields of research;
@@
 	\item \textit{How could a Learning to Aggregate method, that has the previously defined characteristics, look like for graph inputs?}
-		Currently LTA has only been applied to sets of constituents and not yet to graph structured data;
+		Currently LTA has only been applied to multisets of constituents and not yet to graph structured data;
 		this should be addressed by the second question.
@@
 The resulting general definition of LTA methods should then be used for a comparison with a representative selection of existing graph methods (see \cref{sec:related-work:gcr}).
-It should be stated how both fields of research are related and under which circumstances a graph method can be interpreted as an LTA method.
-When formulating a graph method from the LTA perspective, it must be explicitly stated how the constituents and the aggregation function looks like.
+It should be stated how both fields of research are related and under which circumstances a graph method can also be interpreted as an LTA method.
+When formulating a graph method from the LTA perspective, it must be explicitly stated how the constituents and the aggregation function look like.
 Depending on the graph method a constituent could consist of a single vertex but could also represent a certain subgraph or characteristic graph pattern.
@@
 Finally the proposed graph LTA method should be evaluated.
-For this synthetically generated data or some of the datasets listed in \cref{sec:related-work:tox} could be used.
-If time permits, a comparison between the developed graph LTA method and other graph methods should be conducted to empirically demonstrate the potential advantages and disadvantages of the LTA assumptions.
+Synthetically generated data or some of the datasets listed in \cref{sec:related-work:tox} could be used for this.
+Alternatively the sentiment analysis problem from \textit{natural language processing} (NLP) could be used by interpreting the dependency parse graph of a text as the input and the text's sentiment as its aggregate score.
+If time permits, a comparison between the developed graph LTA method and other graph methods should be conducted to demonstrate the potential advantages and disadvantages of the LTA assumptions.
 

### 2019-10-29 thesis/content/abstract.tex
added thesis wireframe

% !TEX root = ../main.tex
%
\pdfbookmark[0]{Abstract}{Abstract}
\chapter*{Abstract}%
\label{sec:abstract}
\vspace*{-10mm}

### 2019-10-29 thesis/content/chapter-appendix.tex
added thesis wireframe

% !TEX root = ../main.tex
% chktex-file 46

\chapter{Appendix}%
\label{sec:appendix}

### 2019-10-29 thesis/content/chapter-conclusion.tex
added thesis wireframe

% !TEX root = ../main.tex
% chktex-file 46
\chapter{Conclusion}%
\label{sec:conclusion}

\section{Review}%
\label{sec:conclusion:review}

\section{Future Directions}%
\label{sec:conclusion:todo}

### 2019-10-29 thesis/content/chapter-introduction.tex
added thesis wireframe

% !TEX root = ../main.tex
% chktex-file 46
\chapter{Introduction}%
\label{sec:intro}

\pagenumbering{arabic}			% arabic page numbering
\setcounter{page}{1}			% set page counter

\section{Motivation}%
\label{sec:intro:motivation}

\section{Goals}%
\label{sec:intro:goals}

\section{Structure}%
\label{sec:intro:structure}

\paragraph{\Cref{sec:conclusion}: \nameref{sec:conclusion}}

### 2019-10-29 thesis/content/declaration.tex
added thesis wireframe

% !TEX root = ../main.tex
%
\begin{otherlanguage}{ngerman}

\pdfbookmark[0]{Declaration}{Declaration}
\chapter*{Erklärung zur Masterarbeit}%
\label{sec:declaration}
\thispagestyle{empty}

Ich, \thesisName\ (Matrikel-Nr. \thesisMatNr), versichere, dass ich die Masterarbeit mit dem Thema \textit{\thesisTitle} selbstständig verfasst und keine anderen als die angegebenen Quellen und Hilfsmittel benutzt habe.
Die Stellen der Arbeit, die ich anderen Werken dem Wortlaut oder dem Sinn nach entnommen habe, wurden in jedem Fall unter Angabe der Quellen der Entlehnung kenntlich gemacht.
Das Gleiche gilt auch für Tabellen, Skizzen, Zeichnungen, bildliche Darstellungen usw.
Die Masterarbeit habe ich nicht, auch nicht auszugsweise, für eine andere abgeschlossene Prüfung angefertigt.
Auf §~63~Abs.~5 HZG wird hingewiesen.

\bigskip

\noindent\textit{\thesisUniversityCity, \thesisDate}

\smallskip

\begin{flushright}
	\begin{minipage}{7cm}
		\rule{\textwidth}{1pt}
		\centering\thesisName\
	\end{minipage}
\end{flushright}

\clearpage
\newpage

\end{otherlanguage}

### 2019-10-29 thesis/main.tex
added thesis wireframe

% **************************************************
% Document Class Definition
% **************************************************
\documentclass[%
    paper=A4,               % paper size --> A4 is default in Germany
    twoside=true,           % onesite or twoside printing
    openright,              % doublepage cleaning ends up right side
    parskip=full,           % spacing value / method for paragraphs
    chapterprefix=true,     % prefix for chapter marks
    11pt,                   % font size
    headings=normal,        % size of headings
    bibliography=totoc,     % include bib in toc
    listof=totoc,           % include listof entries in toc
    titlepage=on,           % own page for each title page
    captions=tableabove,    % display table captions above the float env
    draft=false,            % value for draft version
]{scrreprt}%

\input{config}

% **************************************************
% Document CONTENT
% **************************************************
\begin{document}

% --------------------------
% rename document parts
% --------------------------

\newcommand{\fixedinput}{\input}

% --------------------------
% Front matter
% --------------------------
\pagenumbering{roman}			% roman page numbing (invisible for empty page style)
\pagestyle{empty}				% no header or footers
\input{content/titlepages}		% INCLUDE: all titlepages
\cleardoublepage\

\pagestyle{plain}				% display just page numbers
\input{content/abstract}		% INCLUDE: the abstracts (english and german)
\cleardoublepage\
%
\setcounter{tocdepth}{2}		% define depth of toc
\tableofcontents				% display table of contents
\cleardoublepage\
% --------------------------
% Body matter
% --------------------------
\pagestyle{maincontentstyle} 	% fancy header and footer

\input{content/chapter-introduction}
\input{content/chapter-conclusion}

% --------------------------
% Back matter
% --------------------------
\appendix\cleardoublepage\
\input{content/chapter-appendix}
%
{%
\setstretch{1.1}
\renewcommand{\bibfont}{\normalfont\small}
\setlength{\biblabelsep}{0pt}
\setlength{\bibitemsep}{0.5\baselineskip plus 0.5\baselineskip} % chktex 1
\setcounter{biburllcpenalty}{9000}
\setcounter{biburlucpenalty}{9999}
\printbibliography[nottype=online]
\printbibliography[heading=subbibliography,title={Webseiten},type=online]
}
\cleardoublepage\

\listoffigures
\cleardoublepage\

\input{content/declaration}

% **************************************************
% End of Document CONTENT
% **************************************************
\end{document}

### 2019-10-30 proposal/paper.tex
proposal rc1

@@
 where $\lambda$ is a weight vector with ${\|\lambda\|}_1 = 1$ and $\pi$ is a sorting permutation of the input scores. % chktex 21
-To deal with varying composite sizes $n$ the weights $\lambda_i$ are interpolated using a \textit{basic unit interval monotone} (BUM) function $q: [0, 1] \to [0, 1]$.
+To deal with varying composition sizes $n$ the weights $\lambda_i$ are interpolated using a \textit{basic unit interval monotone} (BUM) function $q: [0, 1] \to [0, 1]$.
 It takes constituent positions that are normalized to the unit interval, i.e.\ $\frac{i}{n}$.
@@
 
-When looking only at the input and output of the described LTA and graph methods, they share certain similarities.
+The described LTA and graph methods are related via their input and output types:
 Both take a variable size and permutation invariant set of constituents/vertices as input and compute an aggregate class or score from it.
@@
 To answer the three research questions, the following approach will be taken.
-First the notion of composites and constituents should be generalized based on the definitions found in the previous works on LTA~\cite{Melnikov2016}\cite{Melnikov2019}, so that they are also applicable to graph inputs.
+First the notion of compositions and constituents should be generalized based on the definitions found in the previous works on LTA~\cite{Melnikov2016}\cite{Melnikov2019}, so that they are also applicable to graph inputs.
 Next the notion of an aggregation function has to be defined and distinguished from the more general concept of an evaluation function (compare \cref{fig:gnnvslta}).
@@
 Finally the proposed graph LTA method should be evaluated.
-Synthetically generated data or some of the datasets listed in \cref{sec:related-work:tox} could be used for this.
-Alternatively the sentiment analysis problem from \textit{natural language processing} (NLP) could be used by interpreting the dependency parse graph of a text as the input and the text's sentiment as its aggregate score.
+To do so a domain problem has to be found that matches the LTA assumption that composite scores can be expressed as aggregations of local scores.
+Synthetically generated data or some of the datasets listed in \cref{sec:related-work:tox} could be considered here.
+The sentiment analysis problem from \textit{natural language processing} (NLP) could be an alternative candidate by interpreting the dependency parse graph of a text as the input and the text's sentiment as its aggregate score.
 If time permits, a comparison between the developed graph LTA method and other graph methods should be conducted to demonstrate the potential advantages and disadvantages of the LTA assumptions.

### 2019-11-04 thesis/content/chapter-graph-lta.tex
added thesis structure

% !TEX root = ../main.tex
% chktex-file 46
\chapter{Learning to Aggregate on Graphs}%
\label{sec:ltag}

\section{Formalization of LTA Methods}%
\label{sec:ltag:formal}

\section{An LTA Interpretation of Graph Methods}%
\label{sec:ltag:interpretation}

\section{LTA on Dynamically Decomposed Graphs}%
\label{sec:ltag:dyndecomp}

### 2019-11-04 thesis/content/chapter-related-work.tex
added thesis structure

% !TEX root = ../main.tex
% chktex-file 46
\chapter{Related Work}%
\label{sec:related}

\section{Learning to Aggregate}%
\label{sec:related:lta}

\section{Graph Kernels}%
\label{sec:related:gk}

\section{Graph Neural Networks}%
\label{sec:related:gnn}

\subsection{Spatial GNNs}%
\label{sec:related:gnn:spatial}

\subsection{Spectral GNNs}%
\label{sec:related:gnn:spectral}

### 2019-11-07 thesis/content/chapter-introduction.tex
started with motivation

@@
 
+The field of \ac{ml} on graph-structured data has applications in many domains due to the general expressive power of graphs.
+Three common types of graph~\ac{ml} problems are
+\begin{enumerate}[label=\textbf{\arabic*.}]
+	\item \textbf{Link prediction:}
+		A graph with an incomplete edge set is given and the missing edges have to be predicted.
+		The generation of friendship suggestions in a social network is a typical example for this.
+	\item \textbf{Vertex classification \& regression:}
+		Here a class or a score has to be predicted for each vertex of a graph.
+		In social graphs this corresponds to the prediction of properties of individuals.
+		Another example is the prediction of the amount of traffic at the intersections of a street network.
+	\item \textbf{Graph classification \& regression:}
+		In this final problem type a single global class or continuous value has to be predicted for an input graph.
+		The canonical example for this is the prediction of properties of molecule graphs, e.g.\ the toxicity or solubility of a chemical.
+\end{enumerate}
+In this thesis we will focus on the last problem type, graph classification and regression.
+A \ac{ml} method for this problem has to accept graphs of varying size and should be permutation invariant wrt.\ the vertices.
+Those requirements are not met by most of the commonly used learners that only accept fixed-size feature vectors as their input, e.g.\ \acp{lm}, \acp{svm} or \acp{mlp}.
+
 \section{Goals}%
@@
 
+\paragraph{\Cref{sec:related}: \nameref{sec:related}}
+
+\paragraph{\Cref{sec:ltag}: \nameref{sec:ltag}}
+
+\paragraph{\Cref{sec:eval}: \nameref{sec:eval}}
+
 \paragraph{\Cref{sec:conclusion}: \nameref{sec:conclusion}}

### 2019-11-08 thesis/content/chapter-introduction.tex
added motivation, started goals

@@
 \end{enumerate}
-In this thesis we will focus on the last problem type, graph classification and regression.
+In this thesis we will focus on the last problem type, \ac{gcr}.
 A \ac{ml} method for this problem has to accept graphs of varying size and should be permutation invariant wrt.\ the vertices.
-Those requirements are not met by most of the commonly used learners that only accept fixed-size feature vectors as their input, e.g.\ \acp{lm}, \acp{svm} or \acp{mlp}.
+Those requirements are not met by the commonly used learners that only accept fixed-size feature vectors as their input, e.g.\ \acp{lm}, \acp{svm} or \acp{mlp}.
+
+A \ac{gcr} method has to account for two central aspects of the problem:
+\begin{enumerate*}
+	\item Local structural analysis and
+	\item global aggregation
+\end{enumerate*}.
+The first aspect is about the extraction of relevant features of substructures of the input graph.
+The latter is about the way in which the local features are combined into a final class or regression value.
+The existing \ac{gcr} methods are mostly motivated by local structural graph analysis.
+The aspect of global aggregation on the other hand is less emphasized by those methods.
+
+There is however a separate branch of research that specifically looks at the problem of learning aggregation functions, called \ac{lta}.
+Current \ac{lta} approaches explicitly learn an aggregation functions for sets which can be interpreted as graphs without edges.
+The motivation for this thesis is to generalize \ac{lta} from sets to arbitrary graphs.
+The overall goal is to combine the aggregation learning perspective with existing \ac{gcr} methods.
 
@@
 \label{sec:intro:goals}
+
+To extend \ac{lta} to graphs, three goals have to be achieved:
+\begin{enumerate}[label=\textbf{\arabic*.}]
+	\item \textbf{Formalization of \ac{lta}:}
+		Before \ac{lta} can be extended, its essential characteristics have to be defined.
+		Those characteristics should provide the terminology to formally capture the differences and similarities between \ac{lta} and existing \ac{gcr} methods.
+	\item \textbf{Give an \ac{lta} interpretation of \ac{gcr} methods:}
+		Using the \ac{lta} formalization, representative \ac{gcr} approaches should be stated in terms of an \ac{lta} method.
+		Currently there is no comprehensive formulation of the relation between both fields of research;
+		this is addressed by the the second goal.
+	\item \textbf{Define an \ac{lta} method for graphs:}
+
+\end{enumerate}
 

### 2019-11-09 thesis/content/chapter-introduction.tex
added lta description from proposal as a starting point

@@
 	\item \textbf{Give an \ac{lta} interpretation of \ac{gcr} methods:}
-		Using the \ac{lta} formalization, representative \ac{gcr} approaches should be stated in terms of an \ac{lta} method.
+		Using the \ac{lta} formalization, representative \ac{gcr} approaches should be restated as \ac{lta} instances.
 		Currently there is no comprehensive formulation of the relation between both fields of research;
@@
 	\item \textbf{Define an \ac{lta} method for graphs:}
-
+		Using the \ac{lta} perspective on \ac{gcr}, hidden assumptions of the existing approaches should become clear and in which way they share the assumptions of \ac{lta}.
+		The last goal is to use those insights to formulate an \ac{lta}-\ac{gcr} method that combines ideas from the existing approaches with the \ac{lta} assumptions.
 \end{enumerate}

### 2019-11-09 thesis/content/chapter-related-work.tex
added lta description from proposal as a starting point

@@
 
+Before combining \ac{lta} and \ac{gcr} as described in \cref{sec:intro:goals}, we first give an overview of the state-of-the-art in both fields of research.
+First the existing \ac{lta} methods for unstructured set inputs are described.
+Then we will look at the two main families of \ac{gcr} methods:
+\begin{enumerate*}
+	\item Vector representation methods
+	\item Graph neural networks
+\end{enumerate*}.
+
 \section{Learning to Aggregate}%
@@
 
-\section{Graph Kernels}%
-\label{sec:related:gk}
+The class of \ac{lta} problems was first described by \citet{Melnikov2016}.
+There an input instance is understood as a composition $\bm{c}_i$ of so-called constituents $c_{i, j} \in \bm{c}_i$, i.e.\ as a variable-size multiset with $n_i = |\bm{c}_i|$.
+The assumption in \ac{lta} problems is that for all constituents $c_{i,j}$ a local score $y_{i,j} \in \mathcal{Y}$ is either given or computable.
+The set of those local scores should be indicative of the overall score $y_i \in \mathcal{Y}$ of the composition $\bm{c}_i$.
+\ac{lta} problems typically require two subproblems to be solved:
+\begin{enumerate}[label=\textbf{\arabic*.}]
+	\item \textbf{Aggregation:}
+		A variadic aggregation function $A: \mathcal{Y}^{*} \to \mathcal{Y}$ that estimates composite scores has to be learned, i.e.\@ $y_i \approx \hat{y}_i = A(y_{i,1}, \dots, y_{i,n_i})$.
+		Typically the aggregation function $A$ should be associative and commutative to fit with the multiset-structure of compositions.
+	\item \textbf{Disaggregation:}
+		In case the constituent scores $y_{i,j}$ are not given, they have to be derived from a constituent representation, e.g.\ a vector $x_{i,j} \in \mathcal{X}$.
+		To learn this derivation function $f: \mathcal{X} \to \mathcal{Y}$, only the constituent vectors ${\{x_{i,j}\}}_{j = 1}^{n_i}$ and the composite score $y_i$ is given.
+		Thus the constituent scores $y_{i,j}$ need to be \textit{disaggregated} from $y_i$ in order to learn $f$.
+\end{enumerate}
+Overall \ac{lta} can be understood as the joint problem of learning the aggregation function $A$ and the local score derivation function $f$.
+Two main approaches to represent the aggregation function in \ac{lta} problems have been explored.
 
-\section{Graph Neural Networks}%
-\label{sec:related:gnn}
+\paragraph{Uninorm-Aggregation}
+The first approach uses \textit{uninorms}~\cite{Melnikov2016} to do so.
+There the basic idea is to express composite scores as fuzzy truth assignments $y_i \in [0, 1]$.
+Such a composite assignment $y_i$ is modeled as the result of a parameterized logical expression of constituent assignments $y_{i,j} \in [0, 1]$.
+As the logical expression that thus effectively aggregates the constituents, a uninorm $U_{\lambda}$ is used.
+Depending on the parameter $\lambda$, $U_{\lambda}$ combines t-norms and t-conorms which are continuous generalizations of logical conjunction and disjunction respectively.
 
-\subsection{Spatial GNNs}%
-\label{sec:related:gnn:spatial}
+\paragraph{OWA-Aggregation}
+Recently \citet{Melnikov2019} have also looked at an alternative class of aggregation functions.
+Instead of using fuzzy logic to describe score aggregation, \textit{ordered weighted average} (OWA) operators were used.
+OWA aggregators work by sorting the input scores and then weighting them based on their sort position, i.e.\ %
+\begin{align*}
+	A_{\lambda}(y_1, \dots, y_n) := \sum_{i = 1}^n \lambda_i y_{\pi(i)},
+\end{align*}
+where $\lambda$ is a weight vector with ${\|\lambda\|}_1 = 1$ and $\pi$ is a sorting permutation of the input scores. % chktex 21
+To deal with varying composition sizes $n$ the weights $\lambda_i$ are interpolated using a \textit{basic unit interval monotone} (BUM) function $q: [0, 1] \to [0, 1]$.
+It takes constituent positions that are normalized to the unit interval, i.e.\ $\frac{i}{n}$.
+The BUM function $q$ then is then used to interpolate a weight for any normalized sort position via $\lambda_i = q(\frac{i}{n}) - q(\frac{i - 1}{n})$.
+Therefore the learning objective boils down to optimizing the shape of $q$.
 
-\subsection{Spectral GNNs}%
-\label{sec:related:gnn:spectral}
+\section{Graph Regression and Classification}%
+\label{sec:related:gcr}
+
+\subsection{Graph Embeddings}%
+\label{sec:related:gcr:embed}
+
+\subsection{Graph Kernels}%
+\label{sec:related:gcr:kernel}
+
+\subsection{Graph Neural Networks}%
+\label{sec:related:gcr:nn}
+
+\subsubsection{Spatial GNNs}
+
+\subsubsection{Spectral GNNs}

### 2019-11-11 thesis/content/chapter-related-work.tex
extended lta description

--- 2019-11-04 thesis/content/chapter-related-work.tex
@@
 
-\section{Graph Kernels}%
-\label{sec:related:gk}
+The class of \ac{lta} problems was first described by \citet{Melnikov2016}.
+There an input instance is understood as a composition $\bm{c}$ of so-called constituents $c_i \in \bm{c}$, i.e.\ as a variable-size multiset with $n = |\bm{c}|$.
+The assumption in \ac{lta} problems is that for all constituents $c_i$ a local score $y_i \in \mathcal{Y}$ is either given or computable.
+The set of those local scores should be indicative of the overall score $y \in \mathcal{Y}$ of the composition $\bm{c}$.
+\ac{lta} problems typically require two subproblems to be solved:
+\begin{enumerate}[label=\textbf{\arabic*.}]
+	\item \textbf{Aggregation:}
+		A variadic aggregation function $\mathcal{A}: \mathcal{Y}^{*} \to \mathcal{Y}$ that estimates composite scores has to be learned, i.e.\@ $y_i \approx \hat{y} = \mathcal{A}(y_{1}, \dots, y_{n})$.
+		Typically the aggregation function $\mathcal{A}$ should be associative and commutative to fit with the multiset-structure of compositions.
+	\item \textbf{Disaggregation:}
+		In case the constituent scores $y_i$ are not given, they have to be derived from a constituent representation, e.g.\ a vector $x_i \in \mathcal{X}$.
+		To learn this derivation function $f: \mathcal{X} \to \mathcal{Y}$, only the constituent vectors ${\{x_i\}}_{i = 1}^{n}$ and the composite score $y$ is given.
+		Thus the constituent scores $y_i$ need to be \textit{disaggregated} from $y$ in order to learn $f$.
+\end{enumerate}
+\begin{figure}
+	\centering
+	\includegraphics[width=0.7\linewidth]{gfx/related-work/lta-overview.pdf}
+	\caption{Overview of the structure of LTA for multiset compositions.}\label{fig:related:lta-overview}
+\end{figure}
+Overall \ac{lta} can be understood as the joint problem of learning the aggregation function $\mathcal{A}$ and the local score derivation function $f$.
+Two main approaches to represent the aggregation function in \ac{lta} problems have been explored.
 
-\section{Graph Neural Networks}%
-\label{sec:related:gnn}
+\subsection{Uninorm-Aggregation}%
+\label{sec:related:lta:uninorm}
 
-\subsection{Spatial GNNs}%
-\label{sec:related:gnn:spatial}
+The first approach uses \textit{uninorms}~\cite{Melnikov2016} to do so.
+There the basic idea is to express composite scores as fuzzy truth assignments $y \in [0, 1]$.
+Such a composite assignment $y$ is modeled as the result of a parameterized logical expression of constituent assignments $y_i \in [0, 1]$.
+As the logical expression that thus effectively aggregates the constituents, a uninorm $U_{\lambda}$ is used.
+Depending on the parameter $\lambda \in [0, 1]$, $U_{\lambda}$ combines a t-norm $T$ and a t-conorm $S$ which are continuous generalizations of logical conjunction and disjunction respectively.
+One popular choice of norms are the so-called Łukasiewicz norms:
+\begin{align}
+	\text{t-norm } T(a, b) &:= \max \{ 0, a + b - 1 \}, \quad\text{t-conorm } S(a, b) := \min \{ a + b, 1 \} \nonumber \\
+	\text{uninorm } U_\lambda(a, b) &:= \begin{cases}
+		\lambda T\left(\frac{a}{\lambda}, \frac{b}{\lambda}\right) & \text{if } a, b \in [0, \lambda] \\
+		\lambda + (1 - \lambda) S\left(\frac{a - \lambda}{1 - \lambda}, \frac{b - \lambda}{1 - \lambda}\right) & \text{if } a, b \in [\lambda, 1] \\
+		\lambda \min \{ a, b \} & \text{else}
+	\end{cases}
+\end{align}
+At the extreme points $(0, 0)$, $(0, 1)$, $(1, 0)$ and $(1, 1)$, $T$ and $S$ coincide with the Boolean operators $\land$ and $\lor$;
+the values at all other points are interpolated as shown in \cref{fig:related:logic-norms}.
+The uninorm $U_\lambda$ uses the conjunctive t-norm $T$ for values below the threshold $\lambda$ and the disjunctive t-conorm $S$ for values above the threshold.
+$U_\lambda$ therefore smoothly interpolates between a conjunctive and disjunctive operator with the extreme points $U_1 = T$ and $U_0 = S$.
+\begin{figure}[h]
+	\centering
+	\begin{tikzpicture}
+		\begin{axis}[
+			title={t-norm $T$ ($\land$)},
+			xlabel=$a$,
+			ylabel=$b$,
+			xlabel style={xshift=0.2cm, yshift=0.15cm},
+			ylabel style={xshift=-0.2cm, yshift=0.15cm},
+			tick label style={font=\scriptsize},
+			width=0.33\textwidth,
+			colormap = {bluered}{color(0cm) = (t_blue); color(1cm) = (t_red)}
+		]
+			\addplot3[
+				mesh,
+				samples=12,
+				domain=0:1,
+				domain y=0:1
+			]{max(x + y - 1, 0)};
+		\end{axis}
+	\end{tikzpicture}
+	\begin{tikzpicture}
+		\begin{axis}[
+			title={t-conorm $S$ ($\lor$)},
+			xlabel=$a$,
+			ylabel=$b$,
+			xlabel style={xshift=0.2cm, yshift=0.15cm},
+			ylabel style={xshift=-0.2cm, yshift=0.15cm},
+			tick label style={font=\scriptsize},
+			width=0.33\textwidth,
+			colormap = {bluered}{color(0cm) = (t_blue); color(1cm) = (t_red)}
+		]
+			\addplot3[
+				mesh,
+				samples=12,
+				domain=0:1,
+				domain y=0:1
+			]{min(x + y, 1)};
+		\end{axis}
+	\end{tikzpicture}
+	\begin{tikzpicture}
+		\begin{axis}[
+			title={uninorm $U_{0.5}$ ($\land$/$\lor$)},
+			xlabel=$a$,
+			ylabel=$b$,
+			xlabel style={xshift=0.2cm, yshift=0.15cm},
+			ylabel style={xshift=-0.2cm, yshift=0.15cm},
+			tick label style={font=\scriptsize},
+			width=0.33\textwidth,
+			colormap = {bluered}{color(0cm) = (t_blue); color(1cm) = (t_red)}
+		]
+			\addplot3[
+				mesh,
+				samples=12,
+				domain=0:1,
+				domain y=0:1
+			]{(x <= 0.5 && y <= 0.5) * 0.5 * max(2 * (x + y) - 1, 0) + (x > 0.5 && y > 0.5) * (0.5 + 0.5 * min(2 * (x + y - 1), 1)) + (!(x <= 0.5 && y <= 0.5) && !(x > 0.5 && y > 0.5)) * min(x, y)};
+		\end{axis}
+	\end{tikzpicture}
+	\caption{The Łukasiewicz norms and the corresponding uninorm for $\lambda = 0.5$.}\label{fig:related:logic-norms}
+\end{figure}
 
-\subsection{Spectral GNNs}%
-\label{sec:related:gnn:spectral}
+Since t-norms and t-conorms are commutative and associative they can also be applied to non-empty sets of arbitrary size, i.e.
+$T(\{ y_1, \dots, y_n \}) = T(y_1, T(\{ y_2, \dots, y_n \}))$ with fixpoint $T(\{ y \}) = y$.
+Using this extension, a uninorm $U_\lambda$ can be applied to sets which turns it into a parameterized aggregation function $\mathcal{A}_\lambda: {[0, 1]}^* \to [0, 1]$.
+In this simple model the \ac{lta} aggregation problem boils down to the optimization of $\lambda$.
+The \ac{lta} disaggregation problem was not considered in the uninorm approach~\cite{Melnikov2016}, i.e.\ the constituent scores ${\{ y_i \}}_{c_i \in \bm{c}}$ are assumed to be given.
+
+\subsection{\acs*{owa}-Aggregation}%
+\label{sec:related:lta:owa}
+
+Recently \citet{Melnikov2019} have looked at an alternative class of aggregation functions.
+Instead of using fuzzy logic to describe score aggregation, \ac{owa} operators were used.
+\ac{owa} aggregators work by sorting the input scores and then weighting them based on their sort position, i.e.\ %
+\begin{align*}
+	\mathcal{A}_{\lambda}(y_1, \dots, y_n) := \sum_{i = 1}^n \lambda_i y_{\pi(i)},
+\end{align*}
+where $\lambda \in \mathbb{R}^n$ is a weight vector with ${\|\lambda\|}_1 = 1$ and $\pi$ is a sorting permutation of the input scores with $y_i < y_j \Rightarrow \pi(i) < \pi(j)$. % chktex 21
+Depending on the choice of the vector $\lambda$, the \ac{owa} function $\mathcal{A}_\lambda$ can express common aggregation functions like
+$\min$ (if $\lambda = (1, 0, \dots, 0)$),
+$\max$ (if $\lambda = (0, \dots, 0, 1)$)
+or the arithmetic mean (if $\lambda = \left(\frac{1}{n}, \dots, \frac{1}{n}\right)$).
+
+To deal with varying composition sizes $n$, the weights ${\{ \lambda_i \}}_{i = 1}^{n}$ can however not be statically assigned.
+Instead they are interpolated using a so-called \ac{bum} function $q: [0, 1] \to [0, 1]$.
+It takes constituent positions that are normalized to the unit interval, i.e.\ $\frac{i}{n} \in [0, 1]$.
+The \ac{bum} function $q$ is then used to interpolate a weight for any normalized sort position via $\lambda_i := q\left(\frac{i}{n}\right) - q\left(\frac{i - 1}{n}\right)$.
+Because $q$ is monotone with $q(0) = 0$ and $q(1) = 1$, it always holds that ${\|\lambda\|}_1 = q(1) - q(0) = 1$. % chktex 21
+Using this model, the aggregation problem boils down to optimizing the shape of $q$.
+
+In the described \ac{owa} approach~\citet{Melnikov2019} the \ac{bum} function $q$ is modeled as a piecewise linear spline.
+
+\section{Graph Regression and Classification}%
+\label{sec:related:gcr}
+
+\subsection{Graph Embeddings}%
+\label{sec:related:gcr:embed}
+
+\subsection{Graph Kernels}%
+\label{sec:related:gcr:kernel}
+
+\subsection{Graph Neural Networks}%
+\label{sec:related:gcr:nn}
+
+\subsubsection{Spatial GNNs}
+
+\subsubsection{Spectral GNNs}

### 2019-11-14 thesis/content/chapter-introduction.tex
first lta description completed

@@
 A \ac{ml} method for this problem has to accept graphs of varying size and should be permutation invariant wrt.\ the vertices.
-Those requirements are not met by the commonly used learners that only accept fixed-size feature vectors as their input, e.g.\ \acp{lm}, \acp{svm} or \acp{mlp}.
+Those requirements are not met by the commonly used learners that only accept fixed-size feature vectors as their input, e.g.\ \acp*{lm}, \acp{svm} or \acp{mlp}.
 

### 2019-11-14 thesis/content/chapter-related-work.tex
first lta description completed

@@
 In this simple model the \ac{lta} aggregation problem boils down to the optimization of $\lambda$.
-The \ac{lta} disaggregation problem was not considered in the uninorm approach~\cite{Melnikov2016}, i.e.\ the constituent scores ${\{ y_i \}}_{c_i \in \bm{c}}$ are assumed to be given.
+The \ac{lta} disaggregation problem is solved by jointly optimizing a \ac{lm}, i.e.\ the constituent scores ${\{ y_i \in [0, 1] \}}_{c_i \in \bm{c}}$ are described by $y_i = {\left( 1 + \exp(-\theta^\top x_i) \right)}^{-1}$.
+Overall an \ac{lta} model is therefore described by the uninorm parameter $\lambda$ and the regression coefficients $\theta$.
 
@@
 
-In the described \ac{owa} approach~\citet{Melnikov2019} the \ac{bum} function $q$ is modeled as a piecewise linear spline.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.75\linewidth]{gfx/related-work/bum.pdf}
+	\caption[Illustration of how a \ac{bum} function is described as a linear spline and its relation to the \ac{owa} weights.]{
+		Illustration of how \textcolor{t_blue}{$a$} describes \textcolor{t_blue}{$q$} and its relation to \textcolor{t_red}{$\lambda$} ($n = 4$, $m = 7$).
+	}\label{fig:related:bum}
+\end{figure}
+In the \ac{owa} approach the \ac{bum} function $q$ is modeled as a piecewise linear spline.
+This spline is described by $m+1$ points ${\left\{ \left( \frac{j}{m}, a_j \right) \right\}}_{j = 0}^{m}$, the so-called knots of the spline. % chktex 21
+The curve of $q$ is obtained by linearly interpolating between neighboring knots as shown in \cref{fig:related:bum}.
+If $0 = a_0 \leq a_1 \leq \cdots \leq a_m = 1$, $q$ is a \ac{bum} function.
+The \ac{lta} aggregation problem is therefore solved by optimizing $a \in \mathbb{R}^{m + 1}$ under this constraint.
+The disaggregation problem is tackled by adding the scores $(y_1, \dots, y_M) \in \mathbb{R}^M$ to the learnable parameters of the model where $M$ is assumed to be the finite number of constituents.
+Currently the \ac{owa} approach requires all possible constituents to be part of the training dataset since it does not consider constituent features $x_i$ to predict the scores of unseen constituents.
 

### 2019-11-15 thesis/content/chapter-related-work.tex
started with gcr related work

@@
 First the existing \ac{lta} methods for unstructured set inputs are described.
-Then we will look at the two main families of \ac{gcr} methods:
-\begin{enumerate*}
-	\item Vector representation methods
-	\item Graph neural networks
-\end{enumerate*}.
+Then we will look at the current approaches to tackle the \ac{gcr} problem.
 
@@
 \begin{align}
-	\text{t-norm } T(a, b) &:= \max \{ 0, a + b - 1 \}, \quad\text{t-conorm } S(a, b) := \min \{ a + b, 1 \} \nonumber \\
+	\text{t-norm } T(a, b) &:= \max \{ 0, a + b - 1 \}, \quad\text{t-conorm } S(a, b) := \min \{ a + b, 1 \}, \nonumber \\
 	\text{uninorm } U_\lambda(a, b) &:= \begin{cases}
@@
 \ac{owa} aggregators work by sorting the input scores and then weighting them based on their sort position, i.e.\ %
-\begin{align*}
+\begin{align}
 	\mathcal{A}_{\lambda}(y_1, \dots, y_n) := \sum_{i = 1}^n \lambda_i y_{\pi(i)},
-\end{align*}
+\end{align}
 where $\lambda \in \mathbb{R}^n$ is a weight vector with ${\|\lambda\|}_1 = 1$ and $\pi$ is a sorting permutation of the input scores with $y_i < y_j \Rightarrow \pi(i) < \pi(j)$. % chktex 21
@@
 The \ac{lta} aggregation problem is therefore solved by optimizing $a \in \mathbb{R}^{m + 1}$ under this constraint.
-The disaggregation problem is tackled by adding the scores $(y_1, \dots, y_M) \in \mathbb{R}^M$ to the learnable parameters of the model where $M$ is assumed to be the finite number of constituents.
-Currently the \ac{owa} approach requires all possible constituents to be part of the training dataset since it does not consider constituent features $x_i$ to predict the scores of unseen constituents.
+The disaggregation problem is tackled by adding the scores $y_1, \dots, y_M \in \mathbb{R}$ to the learnable parameters of the model where $M$ is assumed to be the finite number of constituents.
+Currently the \ac{owa} approach requires all possible constituents to be part of the training dataset since it does not consider constituent features $x_i \in \mathcal{X}$ to predict the scores of previously unseen constituents.
 
@@
 
-\subsection{Graph Embeddings}%
+The existing approaches to tackle the \ac{gcr} problem can be categorized into three main families:
+\begin{enumerate*}
+	\item Explicit graph embeddings,
+	\item Graph kernels and
+	\item Graph neural networks
+\end{enumerate*}.
+We will now look at the characteristics of those families and give an overview of specific methods.
+
+\subsection{Explicit Graph Embeddings}%
 \label{sec:related:gcr:embed}
+
+The basic idea of graph embedding approaches is to map a graph $G$ to some finite vector space $\mathbb{R}^d$.
+By embedding a graph any classification or regression algorithm that works with vectors can then be applied.
+There are three main types of graph embedding approaches.
+
+\paragraph{Fingerprint embeddings}
+
+\paragraph{Unsupervised embeddings}
+
+\paragraph{Supervised embeddings}
 

### 2019-11-18 thesis/content/chapter-related-work.tex
extended gcr related work

@@
 Before combining \ac{lta} and \ac{gcr} as described in \cref{sec:intro:goals}, we first give an overview of the state-of-the-art in both fields of research.
-First the existing \ac{lta} methods for unstructured set inputs are described.
-Then we will look at the current approaches to tackle the \ac{gcr} problem.
+This is done in three steps:
+\begin{enumerate}
+	\item We begin with an overview of the existing \ac{lta} methods for set inputs.
+	\item Before the existing \ac{gcr} methods are described, we look at the graph theoretical fundamentals on which those methods are based.
+	\item Using the previously introduced fundamentals, the current approaches to tackle the \ac{gcr} problem are then described.
+\end{enumerate}
 
@@
 
+\section{Graph Theoretical Fundamentals}%
+\label{sec:related:theory}
+
+\subsection{Graph Isomorphism Checks}%
+\label{sec:related:theory:wl}
+
+\subsection{Spectral Graph Theory}%
+\label{sec:related:theory:spectral}
+
 \section{Graph Regression and Classification}%
@@
 
-The basic idea of graph embedding approaches is to map a graph $G$ to some finite vector space $\mathbb{R}^d$.
+The basic idea of graph embedding approaches is to map a graph $G$ to some finite vector space $\mathcal{X} = \mathbb{R}^d$.
 By embedding a graph any classification or regression algorithm that works with vectors can then be applied.
-There are three main types of graph embedding approaches.
+There are three main types of explicit graph embedding approaches.
 
-\paragraph{Fingerprint embeddings}
+\subsubsection{Fingerprint embeddings}
+The first works on graph embeddings were motivated by the study of chemical structures~\cite{Adamson1973}\cite{Willett1986}.
+There a molecule can be interpreted as a labeled graph for which the \ac{gcr} problem corresponds to the prediction of some chemical property, e.g.\ toxicity or solubility.
+Fingerprint embeddings try to find a fixed set of subgraphs $S_1, \dots, S_d$ in an input graph.
+The embedding of the input graph $G$ is a binary vector $x \in {\{0, 1\}}^d$ with $x_i = \mathds{1}[\text{$S_i$ is subgraph of $G$}]$, where $\mathds{1}$ denotes the indicator function.
+This simple approach requires a careful choice of subgraphs by domain experts but can still be competitive with the other more recent approaches we will look at next.
+Fingerprint embeddings are for example used in multiple state-of-the-art toxicity prediction tools like RASAR~\cite{Luechtefeld2018}\cite{ToxTrack}, ProTox~\cite{Drwal2014}\cite{Banerjee2018}\cite{ProTox} or the \ac{toxTest}~\cite{TEST}.
 
-\paragraph{Unsupervised embeddings}
+\subsubsection{Unsupervised embeddings}
+While fingerprint embeddings use a fixed set of subgraphs to characterize a given graph, unsupervised embedding methods determine relevant substructures dynamically.
+One such method is \texttt{graph2vec}~\cite{Narayanan2017} which is based on the \texttt{doc2vec}~\cite{Le2014} document embedding method from \ac{nlp}.
+\texttt{graph2vec} takes a set of graphs $\mathbb{G} = \{ G_1, \dots, G_N \}$ as input and finds an embedding $\varphi: \mathbb{G} \to \mathbb{R}^d$ in which graphs that share isomorphic subgraphs are mapped closer to each other than those that do not.
 
-\paragraph{Supervised embeddings}
-
-\subsection{Graph Kernels}%
+\subsection{Implicit Graph Embeddings via Kernels}%
 \label{sec:related:gcr:kernel}

### 2019-11-20 thesis/content/chapter-related-work.tex
added skip-gram related work

--- 2019-11-15 thesis/content/chapter-related-work.tex
@@
 	\item Explicit graph embeddings,
-	\item Graph kernels and
-	\item Graph neural networks
+	\item graph kernels and
+	\item graph neural networks
 \end{enumerate*}.
@@
 
-The basic idea of graph embedding approaches is to map a graph $G$ to some finite vector space $\mathbb{R}^d$.
-By embedding a graph any classification or regression algorithm that works with vectors can then be applied.
-There are three main types of graph embedding approaches.
+The basic idea of explicit graph embedding approaches is to map a graph $G \in \mathbb{G}$ to some vector in a finite vector space $\mathcal{X} = \mathbb{R}^d$.
+A function $f: \mathbb{G} \to \mathcal{X}$ is called a \textit{graph embedding function}.
+By embedding a graph into $\mathcal{X}$, any classification or regression algorithm that works with vectors can then be applied to solve the \ac{gcr} problem.
 
-\paragraph{Fingerprint embeddings}
+The so-called vertex embedding problem is closely related to the graph embedding problem.
+As the name suggests, a \textit{vertex embedding function} $f_G: V \to \mathcal{X}$ maps all vertices $v \in V$ of a graph $G$ into $\mathcal{X}$.
+The embedding vector $f_G(v)$ ideally encodes relevant information about a vertex and its structural position in $G$.
+It can be used to solve the vertex classification and regression problem via arbitrary \ac{ml} methods for vectors.
+We will now look at three main families of explicit graph and vertex embedding approaches.
 
-\paragraph{Unsupervised embeddings}
+\subsubsection{Fingerprint embeddings}
 
-\paragraph{Supervised embeddings}
+The first works on graph embeddings were motivated by the study of chemical structures~\cite{Adamson1973}\cite{Willett1986}.
+There a molecule can be interpreted as a labeled graph for which the \ac{gcr} problem corresponds to the prediction of some chemical property, e.g.\ toxicity or solubility.
+So-called \textit{fingerprint embeddings} try to match a fixed set of subgraphs $S_1, \dots, S_d$ to the input graph.
+The embedding of a graph $G$ is a binary vector $f(G) = x \in {\{0, 1\}}^d$ with $x_i = \mathds{1}[\text{$S_i$ is subgraph of $G$}]$, where $\mathds{1}$ denotes the indicator function.
+This simple approach usually requires a careful choice of subgraphs but can still be competitive with the other more recent approaches we will look at in the following sections.
+Fingerprint embeddings are for example used in multiple state-of-the-art toxicity prediction tools like RASAR~\cite{Luechtefeld2018}\cite{ToxTrack}, ProTox~\cite{Drwal2014}\cite{Banerjee2018}\cite{ProTox} or the \ac{toxTest}~\cite{TEST}.
+
+\subsubsection{Matrix factorization embeddings}
+
+\subsubsection{Skip-gram inspired embeddings}
+
+Skip-gram embeddings were introduced by \citeauthor{Mikolov2013} as part of the well-known \texttt{word2vec}~\cite{Mikolov2013} word embedding method from \ac{nlp}.
+While a fingerprint embedding explicitly assigns an interpretation to each embedding dimension (i.e.\ to each standard basis vector), a skip-gram embedding only optimizes the distance between embedding vectors based on the similarity of the embedded instances without providing an interpretation of the embedding dimensions.
+
+\paragraph{word2vec}
+Let us first look at the \texttt{word2vec} skip-gram method.
+It gets a sequence of words $(w_0, \dots, w_n)$ as input and outputs embedding vectors $f(w_0), \dots, f(w_n) \in \mathbb{R}^d$.
+To do this the context $C_k(w_i) = \{ w_{i-k}, \dots, w_{i + k} \}$ is computed for all words where $w_i$ is the so-called \textit{context root}.
+The word contexts are then used to optimize the following log-likelihood objective:
+\begin{gather}
+	\max_{f, f_C} \sum_{i = 1}^n \log P(C_k(w_i) | w_i)\ =\ \max_{f, f_C} \sum_{i = 1}^n \smashoperator[r]{\sum_{w_j \in C_k(w_i)}}\, \left[ {f(w_i)}^\top f_C(w_j) - \log Z_{w_i} \right] \label{eq:related:gcr:embed:word2vec}\\
+	\text{with } P(C_k(w_i) | w_i) := \prod_{\mathclap{w_j \in C_k(w_i)}}\, \overbrace{\frac{\exp\left({f(w_i)}^\top f_C(w_j)\right)}{Z_{w_i}}}^{P(w_j | w_i)}
+	\text{ and } Z_{w_i} := \sum_{j = 1}^n \exp\left({f(w_i)}^\top f_C(w_j)\right) \nonumber
+\end{gather}
+\texttt{word2vec} essentially uses an expectation maximization scheme to maximize the probabilities $P(w_j | w_i)$ of observing the context words $w_j \in C_k(w_i)$ of all words $w_i$.
+Those probabilities are described by the overlap of the embeddings $f(w_i)$ of words $w_i$ and the embeddings $f_C(w_j)$ of their context words $w_j$.
+Intuitively this means that words with similar contexts will be mapped close to each other in the embedding space.
+Note that \texttt{word2vec} actually finds two embeddings $f(w)$ and $f_C(w)$ for each word of which only the first is returned.
+The two embeddings represent two different perspectives on words: $f$ describes a word $w_i$ as the root of a context $C_k(w_i)$, $f_C$ on the other hand describes a word $w_j$ as part of a context $C_k(w_i) \ni w_j$.
+
+\paragraph{Vertex embeddings}
+Skip-gram embeddings can be  na{\"\i}vely extended to graphs by realizing that \texttt{word2vec} effectively already is a vertex embedding method for linear graphs in which $C_k(v)$ is simply the $k$-neighborhood of the vertex/word $v$.
+The problem with this na{\"\i}ve extension is that the sizes of $k$-neighborhoods in arbitrary graphs can be much larger and often tend to grow exponentially with $k$.
+To deal with this computational problem the so-called DeepWalk~\cite{Perozzi2014} and \texttt{node2vec}~\cite{Grover2016} methods perform random walks of fixed length to effectively take samples from the neighborhood of vertices.
+Both methods only differ in the transition matrix that is used for the random walk.
+Another difference of DeekWalk and \texttt{node2vec} compared to \texttt{word2vec} is the so-called \textit{feature space symmetry} which states that the context root interpretation ($f$) of a vertex should be symmetric to its context element interpretation ($f_C$), i.e.\ $f = f_C$.
+The combination of random walk context sampling and and the feature space symmetry assumption can be used to compute vertex embeddings even for very large graphs.
+
+\paragraph{Graph embeddings}
+Skip-gram methods can not only be used for vertex embeddings but also to embed entire graphs.
+One way to do this is via the \texttt{graph2vec}~\cite{Narayanan2017} method.
+It is inspired by \texttt{doc2vec}~\cite{Le2014} which in turn is based on \texttt{word2vec}.
+\texttt{graph2vec} gets a set of graphs $\mathbb{G} = \{ G_1, \dots G_N \}$ as input and outputs graph embeddings $f(G_1), \dots, f(G_N)$.
+While in \texttt{word2vec} every word can be a context root as well as a context element, \texttt{graph2vec} uses the graphs $\mathbb{G}$ as context roots.
+The context $C_k(G_i)$ of a graph $G_i = (\mathcal{V}_i, \mathcal{E}_i)$ is defined as
+\begin{equation}
+	C_k(G_i) := \bigcup_{l = 0}^k {\left\{ {\textrm{WL}_l(G_i)}_j \right\}}_{v_j \in \mathcal{V}_i} % chktex 21
+	\text{ with } \textrm{WL}_l(G) := \text{\acs{wl} colors after $l$ iter.\ on $G$.}
+\end{equation}
+Intuitively this context can be understood as the set of isomorphically distinct subgraphs of $G_i$ with diameter $\leq 2k$.
+Since \ac{wl} is used to identify isomorphic subgraphs, \texttt{graph2vec} can only be applied to graphs with discrete vertex labels.
+Using the previous definitions, the context root embedding function has the signature $f: \mathbb{G} \to \mathbb{R}^d$ while the context element embedding function is of type $f_C: \bigcup_{G_i \in \mathbb{G}} C_k(G_i) \to \mathbb{R}^d$.
+To find those embeddings the \texttt{word2vec} objective from \cref{eq:related:gcr:embed:word2vec} is reused.
+Analogous to \texttt{word2vec}, \texttt{graph2vec} therefore embeds graphs that share isomorphic subgraphs close to each other, whereas graphs that do not share substructures tend to be embedded further away from each other.
 

### 2019-11-21 thesis/content/chapter-related-work.tex
leave out matrix factorization for now
probably not relevant for chapter 3

@@
 It can be used to solve the vertex classification and regression problem via arbitrary \ac{ml} methods for vectors.
-We will now look at three main families of explicit graph and vertex embedding approaches.
+We will now look at two main families of explicit graph and vertex embedding approaches.
 
@@
 Fingerprint embeddings are for example used in multiple state-of-the-art toxicity prediction tools like RASAR~\cite{Luechtefeld2018}\cite{ToxTrack}, ProTox~\cite{Drwal2014}\cite{Banerjee2018}\cite{ProTox} or the \ac{toxTest}~\cite{TEST}.
-
-\subsubsection{Matrix factorization embeddings}
 
@@
 Both methods only differ in the transition matrix that is used for the random walk.
-Another difference of DeekWalk and \texttt{node2vec} compared to \texttt{word2vec} is the so-called \textit{feature space symmetry} which states that the context root interpretation ($f$) of a vertex should be symmetric to its context element interpretation ($f_C$), i.e.\ $f = f_C$.
+Another difference of DeepWalk and \texttt{node2vec} compared to \texttt{word2vec} is the so-called \textit{feature space symmetry} which states that the context root interpretation ($f$) of a vertex should be symmetric to its context element interpretation ($f_C$), i.e.\ $f = f_C$.
 The combination of random walk context sampling and and the feature space symmetry assumption can be used to compute vertex embeddings even for very large graphs.
@@
 \begin{equation}
-	C_k(G_i) := \bigcup_{l = 0}^k {\left\{ {\textrm{WL}_l(G_i)}_j \right\}}_{v_j \in \mathcal{V}_i} % chktex 21
+	C_k(G_i) := \bigcup_{\mathclap{v_j \in \mathcal{V}_i}}\, {\left\{ {\textrm{WL}_l(G_i)}_j \right\}}_{l = 0}^k % chktex 21
 	\text{ with } \textrm{WL}_l(G) := \text{\acs{wl} colors after $l$ iter.\ on $G$.}

### 2019-11-22 thesis/content/chapter-related-work.tex
minor fixes

@@
 \end{equation}
-Intuitively this context can be understood as the set of isomorphically distinct subgraphs of $G_i$ with diameter $\leq 2k$.
-Since \ac{wl} is used to identify isomorphic subgraphs, \texttt{graph2vec} can only be applied to graphs with discrete vertex labels.
+Intuitively this context can be understood as the set of \acs{wl}-distinguishable subgraphs of $G_i$ with diameter $\leq 2k$.
+Since \ac{wl} is used to identify distinct subgraphs, \texttt{graph2vec} can only be applied to graphs with discrete vertex labels.
 Using the previous definitions, the context root embedding function has the signature $f: \mathbb{G} \to \mathbb{R}^d$ while the context element embedding function is of type $f_C: \bigcup_{G_i \in \mathbb{G}} C_k(G_i) \to \mathbb{R}^d$.
 To find those embeddings the \texttt{word2vec} objective from \cref{eq:related:gcr:embed:word2vec} is reused.
-Analogous to \texttt{word2vec}, \texttt{graph2vec} therefore embeds graphs that share isomorphic subgraphs close to each other, whereas graphs that do not share substructures tend to be embedded further away from each other.
+Analogous to \texttt{word2vec}, \texttt{graph2vec} therefore embeds graphs that share subgraphs close to each other, whereas graphs that do not share substructures tend to be embedded further away from each other.
 
@@
 \label{sec:related:gcr:kernel}
+
+Instead of explicity mapping a graph or vertices into a vector space, one can also do so implicity by employing the kernel trick.
+There is a large variety of so-called \acp{gk} to do this.
+\acp{gk} can be used in combination with any kernel method to solve the \ac{gcr} problem.
 

### 2019-12-16 approach/paper.tex
started writing a brief approach outline

% chktex-file 46
\documentclass[12pt]{scrartcl}

\PassOptionsToPackage{utf8}{inputenc}
\usepackage{inputenc}
\usepackage[american]{babel}
\usepackage{csquotes}
\usepackage{microtype}

\usepackage{graphicx}
\graphicspath{{gfx/}}

\usepackage{paralist}
\usepackage{csquotes}
\usepackage[T1]{fontenc}
\usepackage{lmodern}

\usepackage{geometry}
% \geometry{a4paper,body={5.8in,9in}}
\geometry{a4paper}
\renewcommand{\baselinestretch}{1.1}
\usepackage{parskip}
\setlength{\parindent}{0pt}

\usepackage{amsmath, amsfonts, amssymb}
\usepackage{bm}
\usepackage{placeins}
\usepackage{subcaption}

\usepackage{setspace}

\usepackage{hyperref}
\usepackage[nameinlink]{cleveref}
\newcommand{\crefrangeconjunction}{--}

\usepackage[						% use biblatex for bibliography
	backend=bibtex,					% 	- use biber backend (bibtex replacement) or bibtex
	style=numeric,					% 	- use alphabetic (or numeric) bib style
	natbib=true,					% 	- allow natbib commands
	hyperref=true,					% 	- activate hyperref support
	backref=true,					% 	- activate backrefs
	isbn=false,						% 	- don't show isbn tags
	url=false,						% 	- don't show url tags
	doi=false,						% 	- don't show doi tags
	urldate=long,					% 	- display type for dates
	maxnames=3,%
	minnames=1,%
	maxbibnames=5,%
	minbibnames=3,%
	maxcitenames=2,%
	mincitenames=1,%
	sorting=none
]{biblatex}
\bibliography{literature}

\usepackage[inline]{enumitem}
\usepackage{todonotes}
\usepackage{pgfgantt}

% Colors:
\definecolor{blau}{HTML}{355FB3}
\definecolor{rot}{HTML}{B33535}
\definecolor{gruen}{HTML}{3BB335}
\definecolor{hellblau}{HTML}{8ea7d7}
\definecolor{hellgrau}{HTML}{cccccc}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\thesisTitle}{Learning to Aggregate on Structured Data}
\newcommand{\thesisSubject}{Approach Outline}
\newcommand{\thesisName}{Clemens Damke}
\newcommand{\thesisMail}{cdamke@mail.uni-paderborn.de}
\newcommand{\thesisMatNr}{7011488}
\hypersetup{% setup the hyperref-package options
    pdftitle={\thesisTitle},    %   - title (PDF meta)
    pdfsubject={\thesisSubject},%   - subject (PDF meta)
    pdfauthor={\thesisName},    %   - author (PDF meta)
    plainpages=false,           %   -
    colorlinks=false,           %   - colorize links?
    pdfborder={0 0 0},          %   -
    breaklinks=true,            %   - allow line break inside links
    bookmarksnumbered=true,     %
    bookmarksopen=true          %
}

\begin{document}

\title{\thesisTitle}
\subtitle{\thesisSubject}
\author{{\thesisName}\\\small{Matriculation Number: \thesisMatNr}\\\small{\href{mailto:\thesisMail}{\thesisMail}}}
\date{\today}
\maketitle

This is a brief outline of the approach chosen to tackle the problem of \textit{Learning to Aggregate}~(LTA) on structured data.
The problem was split into three parts:
\begin{enumerate}[label=\textbf{\arabic*.}]
	\item \textbf{Formalization of LTA:}
		Before LTA can be extended, its essential characteristics have to be defined.
		Those characteristics should provide the terminology to formally capture the differences and similarities between LTA and existing \textit{graph classification and regression}~(GC/GR) methods.
	\item \textbf{Give an LTA interpretation of existing GC/GR methods:}
		Using the LTA formalization, representative GC/GR approaches should be restated as LTA instances.
		Currently there is no comprehensive formulation of the relation between both fields of research;
		this is addressed by the the second goal.
	\item \textbf{Define an LTA method for graphs:}
		Using the LTA perspective on GC/GR, hidden assumptions of the existing approaches should become clear and in which way they share the assumptions of LTA.\@
		The last goal is to use those insights to formulate an LTA-GC/GR method that combines ideas from the existing approaches with the LTA assumptions.
\end{enumerate}

\section{LTA Formalization}%
\label{sec:lta}

First a brief sketch of the proposed general formalization of LTA.\@
It consists of three properties that each LTA approach has to fulfill:
\begin{enumerate}[label=\textbf{\arabic*.}]
	\item \textbf{Decomposition:}
		The input composition first has to be decomposed into its constituents.
		In the existing approaches for multisets this property is trivially fulfilled by considering each set element as a constituent.
		For structured data, i.e.\ graphs, the decomposition problem becomes more interesting.
		While one could consider the set of vertices as a graph's constituents, an approach that makes use of the relational structure is most likely more suitable.

		Formally a solution to the decomposition problem is described by a function $\varphi: \mathcal{G} \to \mathcal{P}(\mathcal{C})$, where $\mathcal{G}$ is the set of all structured inputs and $\mathcal{C}$ is the set of all constituents that may be found in those inputs.
	\item \textbf{Disaggregation:}
		The constituents that are determined by $\varphi$ are scored via a function $f: \mathcal{C} \to \mathcal{Y}$.
		The disaggregation problem is solved by learning this function.
	\item \textbf{Aggregation:}
		Finally the aggregation problem is about finding an aggregation function $\mathcal{A}_{\text{SI}}: \mathcal{Y}^* \to \mathcal{Y}$ that combines the constituent scores.
		This is the definition of a \textit{structurally independent} (SI) aggregation function, i.e.\ it only depends on the local constituent scores and does not use global structural information to potentially weight the scores.
		The existing LTA methods necessarily learn SI-aggregators since their inputs do not contain any structural information.

		For structured inputs, a more general definition of aggregation functions can be considered.
		A \textit{structurally dependent} (SD) aggregator not only gets constituent scores $f(c) \in \mathcal{Y}$ but also structural features $g(c, G) \in \mathcal{Z}$ for all constituents $c \in \varphi(G)$, i.e.\ $\mathcal{A}_{\text{SD}}: {(\mathcal{Y} \times \mathcal{Z})}^* \to \mathcal{Y}$.
		To guarantee that the aggregated score is based on the constituent scores, an SD-aggregator is only allowed to use the structural features to weight and filter the constituent scores:
		\begin{align}
			\mathcal{A}_{\text{SD}}(S)
			= \mathcal{A}_{\text{SI}}(\{ w_i \circ y_i \,|\, (y_i, z_i) \in S \,\land\, w_i = h((y_i, z_i), S) \neq \texttt{nil} \}) \text{.} % chktex 21
		\end{align}
		Here $\circ: W \times \mathcal{Y} \to \mathcal{Y}$ is some score scaling operator on $\mathcal{Y}$, e.g.\ real multiplication if $\mathcal{Y} = \mathbb{R}$.
		Also note that structural constituent filtering is described via $w_i = \texttt{nil}$ which excludes the constituent $c_i$ from the aggregated score.
\end{enumerate}
To summarize, the output $y_i$ of an LTA model for an input graph $G_i$ has to be expressible via
\begin{align}
	y_i = \mathcal{A}_{\text{SD}}(\{ (f(c), g(c, G_i)) \,|\, c \in \varphi(G_i) \}) \text{.}
\end{align}
One advantage of this class of models is the explainability of the outputs.
An aggregated score can always be tracked back to the individual constituents that went into it.

\section{LTA Interpretation of GC/GR Methods}%
\label{sec:lta-gcr}

Next a very brief sketch of existing GC/GR methods and their relation to LTA is provided here.
\begin{enumerate}[label=\textbf{\arabic*.}]
	\item \textbf{Graph Kernels:}
		Graph kernels do not directly allow for an LTA interpretation;
		for example an SVM with a graph kernel does not fulfill the previously described disaggregation and aggregation properties of LTA methods.
		Many graph kernels do however effectively compute graph decompositions and can therefore serve as starting points for graph LTA methods.
		The following graph kernels are interesting from an LTA perspective:
		\begin{itemize}
			\item The \textit{Weisfeiler-Lehman} (WL) subtree kernel~\cite{Shervashidze2011} decomposes a graph into subtrees of depth $k$ with each vertex being the root of exactly one subtree.
				The subtrees correspond to the breadth-first-walk trees with vertex repetition.
				Subtrees are identified via their isomorphism class.
			\item The Direct-Product kernel conceptually decomposes a graph into the set of all random walks with arbitrary finite length.
				The walks are identified by the sequence of vertex labels along each walk.
			\item Other relevant kernels include the WL edge kernel, the WL shortest path kernel, the graphlet kernel and the GraphHopper kernel.
				Those kernels also compute graph decompositions.
				The details of those decompositions are not described in this summary.
		\end{itemize}
	\item \textbf{Graph Neural Networks:}
		Most existing \textit{graph neural network} (GNN) approaches fulfill the three previously described LTA properties.
		All GNN methods that are based vertex-neighborhood aggregation can be understood as continuous variants of the \textit{1-dimensional WL algorithm} (1-WL)~\cite{Weisfeiler1968}.
		They are therefore closely related to the WL subtree kernel and in fact upper bounded by it in terms of their discriminative power.
		They effectively also compute a subtree decomposition.
		The constituent scoring function $f$ is modeled as a \textit{multilayer perceptron} (MLP).\@
		Between each MLP layer a weighted average of the features of all vertices in the constituent subtree is taken.
		The last layer of the MLP has to output a value which can be interpreted as an element of $\mathcal{Y}$.

		Graph pooling layers like SortPooling~\cite{Zhang2018} or the self-attention based SAGPooling~\cite{Lee2019} can then be added to solve the aggregation problem.
		SortPooling can be understood as an SI-aggregator, while SAGPooling is an SD-aggregator with $z_i = g(c, G)$ being self-attention scores for the subtrees.
		In SAGPool $g$ is modeled as a second neighborhood-averaging MLP similar to $f$.

		The details of the relation between LTA and GNNs are not described here.
\end{enumerate}

\section{Proposed Graph LTA method}%
\label{sec:graph-lta}

The proposed graph LTA method is based on two main ideas:
\begin{enumerate}
	\item While most existing GNN methods can be interpreted as LTA approaches, they use a fairly static decomposition approach.
		In this approach each constituent is a tree that includes all vertices with a distance of at most $k$ from some root vertex.
		The vertices that are part of a constituent therefore do not necessarily have any significant relation to each other.
		The decompositions of existing GNN methods therefore do not have any semantic.
		A learnable decomposition strategy could improve upon this problem.
	\item As previously mentioned, the discriminative power of most GNNs is bounded by the 1-WL algorithm.
		The limits of this algorithm are well understood and while 1-WL is able to distinguish most graphs, it cannot detect many graph properties that are considered to be important in various domains of graph analysis, e.g.\ triangle counts in the context social networks or cycle counts in the context of molecular analysis.
		1-WL also fails to distinguish regular graphs with the same degree and vertex count.
		There are higher dimensional variants of the WL algorithm that are able to capture such aspects.
		The 2-dimensional WL algorithm, which aggregates edge neighborhoods instead of vertex neighborhoods, is already powerful enough to recognize triangles, cycles of length $\leq 6$ and most regular graphs.
		A GNN based on a higher dimensional WL algorithm could therefore at least theoretically improve upon the expressive power of most existing state-of-the-art approaches.
\end{enumerate}

% % %
\newpage
{%
\renewcommand{\bibfont}{\normalfont\small}
\setlength{\biblabelsep}{5pt}
\setlength{\bibitemsep}{0.5\baselineskip plus 0.5\baselineskip} % chktex 1
\setcounter{biburllcpenalty}{9000}
\setcounter{biburlucpenalty}{9999}
\printbibliography%
}

\end{document} % chktex 17

### 2019-12-17 approach/paper.tex
completed approach outline

@@
 		Those characteristics should provide the terminology to formally capture the differences and similarities between LTA and existing \textit{graph classification and regression}~(GC/GR) methods.
-	\item \textbf{Give an LTA interpretation of existing GC/GR methods:}
+	\item \textbf{Giving an LTA interpretation of existing GC/GR methods:}
 		Using the LTA formalization, representative GC/GR approaches should be restated as LTA instances.
@@
 		this is addressed by the the second goal.
-	\item \textbf{Define an LTA method for graphs:}
+	\item \textbf{Defining an LTA method for graphs:}
 		Using the LTA perspective on GC/GR, hidden assumptions of the existing approaches should become clear and in which way they share the assumptions of LTA.\@
@@
 		Finally the aggregation problem is about finding an aggregation function $\mathcal{A}_{\text{SI}}: \mathcal{Y}^* \to \mathcal{Y}$ that combines the constituent scores.
-		This is the definition of a \textit{structurally independent} (SI) aggregation function, i.e.\ it only depends on the local constituent scores and does not use global structural information to potentially weight the scores.
+		This is the definition of a \textit{structurally independent}~(SI) aggregation function, i.e.\ it only depends on the local constituent scores and does not use global structural information to potentially weight the scores.
 		The existing LTA methods necessarily learn SI-aggregators since their inputs do not contain any structural information.
@@
 		For structured inputs, a more general definition of aggregation functions can be considered.
-		A \textit{structurally dependent} (SD) aggregator not only gets constituent scores $f(c) \in \mathcal{Y}$ but also structural features $g(c, G) \in \mathcal{Z}$ for all constituents $c \in \varphi(G)$, i.e.\ $\mathcal{A}_{\text{SD}}: {(\mathcal{Y} \times \mathcal{Z})}^* \to \mathcal{Y}$.
-		To guarantee that the aggregated score is based on the constituent scores, an SD-aggregator is only allowed to use the structural features to weight and filter the constituent scores:
+		A \textit{structurally dependent}~(SD) aggregator not only gets constituent scores $f(c) \in \mathcal{Y}$ but also structural features $g(c, G) \in \mathcal{Z}$ for all constituents $c \in \varphi(G)$, i.e.\ $\mathcal{A}_{\text{SD}}: {(\mathcal{Y} \times \mathcal{Z})}^* \to \mathcal{Y}$.
+		To guarantee that the aggregated score is based on some combination of the constituent scores and not arbitrarily derived from the structural features, an SD-aggregator is only allowed to use the structural features to weight and filter the constituent scores:
 		\begin{align}
@@
 
-Next a very brief sketch of existing GC/GR methods and their relation to LTA is provided here.
+Next a very brief sketch of existing GC/GR methods and their relation to LTA is provided.
 \begin{enumerate}[label=\textbf{\arabic*.}]
@@
 		Graph kernels do not directly allow for an LTA interpretation;
-		for example an SVM with a graph kernel does not fulfill the previously described disaggregation and aggregation properties of LTA methods.
-		Many graph kernels do however effectively compute graph decompositions and can therefore serve as starting points for graph LTA methods.
+		an SVM with a graph kernel for example does not fulfill the previously described disaggregation and aggregation properties of LTA methods.
+		Many graph kernels do however effectively compute graph decompositions $\varphi(G)$ and can therefore serve as starting points for graph LTA methods.
 		The following graph kernels are interesting from an LTA perspective:
 		\begin{itemize}
-			\item The \textit{Weisfeiler-Lehman} (WL) subtree kernel~\cite{Shervashidze2011} decomposes a graph into subtrees of depth $k$ with each vertex being the root of exactly one subtree.
+			\item The \textit{Weisfeiler-Lehman}~(WL) subtree kernel~\cite{Shervashidze2011} decomposes a graph into subtrees of depth $k$ with each vertex being the root of exactly one subtree.
 				The subtrees correspond to the breadth-first-walk trees with vertex repetition.
@@
 	\item \textbf{Graph Neural Networks:}
-		Most existing \textit{graph neural network} (GNN) approaches fulfill the three previously described LTA properties.
-		All GNN methods that are based vertex-neighborhood aggregation can be understood as continuous variants of the \textit{1-dimensional WL algorithm} (1-WL)~\cite{Weisfeiler1968}.
-		They are therefore closely related to the WL subtree kernel and in fact upper bounded by it in terms of their discriminative power.
+		Most existing \textit{graph neural network}~(GNN) approaches fulfill the three previously described LTA properties.
+		All GNN methods that are based vertex-neighborhood aggregation can be understood as continuous variants of the \textit{1-dimensional WL algorithm}~(1-WL)~\cite{Weisfeiler1968}.
+		They are therefore closely related to the WL subtree kernel and in fact upper bounded by it in terms of their discriminative power~\cite{Xu2018}\cite{Morris2018}.
 		They effectively also compute a subtree decomposition.
-		The constituent scoring function $f$ is modeled as a \textit{multilayer perceptron} (MLP).\@
+		The constituent scoring function $f$ is modeled as a \textit{multilayer perceptron}~(MLP).\@
 		Between each MLP layer a weighted average of the features of all vertices in the constituent subtree is taken.
@@
 The proposed graph LTA method is based on two main ideas:
-\begin{enumerate}
-	\item While most existing GNN methods can be interpreted as LTA approaches, they use a fairly static decomposition approach.
-		In this approach each constituent is a tree that includes all vertices with a distance of at most $k$ from some root vertex.
-		The vertices that are part of a constituent therefore do not necessarily have any significant relation to each other.
-		The decompositions of existing GNN methods therefore do not have any semantic.
-		A learnable decomposition strategy could improve upon this problem.
-	\item As previously mentioned, the discriminative power of most GNNs is bounded by the 1-WL algorithm.
-		The limits of this algorithm are well understood and while 1-WL is able to distinguish most graphs, it cannot detect many graph properties that are considered to be important in various domains of graph analysis, e.g.\ triangle counts in the context social networks or cycle counts in the context of molecular analysis.
+\begin{enumerate}[label=\textbf{\arabic*.}]
+	\item \textbf{Extending discriminative power:}
+		As previously mentioned, the discriminative power of most GNNs is bounded by the 1-WL algorithm.
+		The limits of this algorithm are well understood and while 1-WL is able to distinguish most graphs, it cannot detect many graph properties that are considered to be important in various practical domains of graph analysis, e.g.\ triangle counts in the context social networks or cycle counts in the context of molecular analysis.
 		1-WL also fails to distinguish regular graphs with the same degree and vertex count.
 		There are higher dimensional variants of the WL algorithm that are able to capture such aspects.
-		The 2-dimensional WL algorithm, which aggregates edge neighborhoods instead of vertex neighborhoods, is already powerful enough to recognize triangles, cycles of length $\leq 6$ and most regular graphs.
+		The \textit{2-dimensional WL algorithm}~(2-WL), which aggregates edge neighborhoods instead of vertex neighborhoods, is already powerful enough to recognize triangles, cycles of length $\leq 6$ and also many regular graphs~\cite{Fuerer2017}.
 		A GNN based on a higher dimensional WL algorithm could therefore at least theoretically improve upon the expressive power of most existing state-of-the-art approaches.
+	\item \textbf{Learned decomposition:}
+		While most existing GNN methods can be interpreted as LTA approaches, they use a fairly static decomposition approach.
+		In the static approach each constituent is a tree that includes all vertices with a distance of at most $k$ from some root vertex.
+		The vertices that are part of a constituent therefore do not necessarily have any significant relation to each other.
+		The decompositions of existing GNN methods thus do not have any semantic attached to them, they are purely structural.
+		A learnable decomposition strategy could improve upon this problem.
+\end{enumerate}
+
+The proposed method incorporates the first idea by defining a graph convolution operator inspired by the 2-WL algorithm.
+Recently \citet{Morris2018} have already introduced a GNN architecture based on $k$-WL which does however use a simplified notion of neighborhood.
+This simplification reduces the discriminative power of the architecture; their 2-GNN for example cannot recognize triangles in graphs.
+The proposed method on the other hand is closer to the original 2-WL algorithm.
+
+While 1-WL-based GNNs propagate vertex features via the adjacency structure of an input graph, a 2-WL-based GNN propagates edge features to neighboring edges.
+The aggregation of edge features performed by one network layer $l$ is described by
+\begin{align}
+	h^{(l)}_{i j} &= \sigma\left(\, W^{(l)} h^{(l - 1)}_{i j} + W^{(l)}_{\Gamma} \sum_{k = 1}^{n} \kappa\left(h^{(l - 1)}_{i k}, h^{(l - 1)}_{k j}\right)\right), \\
+	\text{with } h^{(0)}_{i j} &= A_{i j} \oplus \begin{cases}
+		X_i & \text{if } i = j \\
+		\bm{0} & \text{else}
+	\end{cases}
+\end{align}
+where ${\left\{ h^{(l)} \in \mathbb{R}^{n \times n \times F^{(l)}} \right\}}_{l = 0}^{L}$ are the edge feature tensors after each convolution layer. % chktex 21
+The initial edge features $h^{(0)}$ are constructed by concatenating ($\oplus$) the input edge features/weights $A$ with the input vertex features $X$ along the diagonal, i.e.\ self-loops are used to carry the vertex features.
+
+The 2-WL inspired edge convolution combines each edge $(i, j)$ with the edges in all walks of length $2$ between $i$ and $j$.
+First each walk is aggregated via $\kappa: \mathbb{R}^{F} \times \mathbb{R}^{F} \to \mathbb{R}^{F}$ which has to satisfy $\kappa(a, b) = \bm{0}$ if $a = \bm{0} \lor b = \bm{0}$ to preserve the sparsity of $h^{(l)}$ for performance reasons\footnote{
+	$\kappa$ should not be a linear operator, since this would reduce the discriminative power of the model back to 1-WL.\@
+	Element-wise multiplication would be a simple suitable choice for $\kappa$ but a learned nonlinear combinator can also be used.
+}.
+Then the sum of all aggregated walks is used as a neighborhood aggregate.
+Finally $W^{(l)}, W^{(l)}_{\Gamma} \in \mathbb{R}^{F^{(l)} \times F^{(l - 1)}}$ and some activation function $\sigma$ are used to compute the output edge features of the layer.
+
+The choice of 2-WL as a basis for the just described GNN model was not only motivated by its larger discriminative power but also by the idea of learned graph decompositions.
+If the final layer of the described model produces outputs that are interpretable as elements of $\mathcal{Y}$, the graph is effectively decomposed into up to one constituent per edge\footnote{
+	Edges with a zero feature vector are not considered to be constituents.
+}.
+Every convolution layer increases the radius of the neighborhood that is considered in the final feature aggregate of each edge/constituent.
+By learning a filter on the edges one can selectively remove elements from edge neighborhoods since disconnected parts of a graph cannot become part of the same neighborhood.
+The proposed 2-WL convolution model is well suited for a combination with such an edge filter model since all edges already have feature vectors that can be used as inputs for the filter model.
+The details for this still have to be worked out.
+
+To aggregate the edges/constituents, existing graph pooling approaches like SortPooling or SAGPooling can be used.
+
+\section{Remarks and Current Status}%
+\label{sec:remarks}
+
+I have currently only evaluated the proposed 2-WL inspired GNN architecture on synthetic datasets and the MUTAG dataset\footnote{
+	The implementation does not yet use sparse tensors and therefore cannot handle other real datasets that contain large graphs.
+}.
+On the synthetic datasets I was able to clearly demonstrate that most state-of-the-art GNNs fail to learn on datasets consisting of regular graphs as well as datasets with graph labels derived from triangle counts.
+The proposed architecture on the other hand is able to learn the labeling function on those synthetic datasets.
+On the MUTAG dataset it performed more or less like the standard GCN model proposed by \citet{Kipf2017}.
+However the implementation currently only uses a static arithmetic average as the aggregation function, so there is probably still room for improvement.
+
+The next planned steps are:
+\begin{enumerate}
+	\item Use sparse tensors in the implementation in order to be able to evaluate on larger real datasets.
+	\item Integrate the previously described edge filtering idea and formalize how exactly this affects the learned constituents.
+	\item Swap out the constituent averaging with a learned aggregation/pooling layer.
 \end{enumerate}
@@
 % % %
-\newpage
 {%

### 2020-01-13 implementation/evaluate_datasets.sh
added hyperparameter grid search to evaluation

@@ -1,0 +1,5 @@
+#!/usr/bin/env bash
+
+USER="-u $(id -u):$(id -g)"
+
+docker exec -it $USER $(docker ps -aqf "name=ltag") python3 ./ltag/evaluate_datasets.py

### 2020-01-13 implementation/run.sh

@@ -43,14 +43,16 @@ echo "Using additional ARGS='$ARGS' with variant $VARIANT."
 echo "Type \"rm\" to clean logs."
 
 mkdir -p logs
-mkdir -p data
+mkdir -p evaluations
 mkdir -p libs
+mkdir -p data
 
 docker run --runtime=nvidia --rm --name ltag \
 	-p 8888:8888 \
 	-p 6006:6006 \
 	-v $(pwd)/src:/ltag \
 	-v $(pwd)/logs:/logs \
+	-v $(pwd)/evaluations:/evaluations \
 	-v $(pwd)/libs:/libs \
 	-v $(pwd)/data:/data \
 	-e "JUPYTER_TOKEN=$JUPYTER_TOKEN" \

### 2020-01-13 implementation/run_board.sh

@@ -2,9 +2,7 @@
 
 USER="-u $(id -u):$(id -g)"
 
-if [ "$1" == "root" ]; then
-	USER="-u 0:0"
-fi
+LOGDIR="/${1:-logs}"
 
 echo "Starting board at http://localhost:6006/"
-docker exec -it $USER $(docker ps -aqf "name=ltag") tensorboard --logdir /logs --bind_all
+docker exec -it $USER $(docker ps -aqf "name=ltag") tensorboard --logdir $LOGDIR --bind_all

### 2020-01-13 implementation/src/ltag/chaining/model.py

@@ -48,6 +48,7 @@ def create_model(name, steps, **kwargs):
     return create_model(name, steps + additional_steps, **ext_kwargs)
 
   modelFactory.extend = extend
+  modelFactory.name = name
 
   return modelFactory
 

### 2020-01-13 implementation/src/ltag/datasets/manager.py

@@ -1,9 +1,14 @@
+from __future__ import absolute_import, division, print_function,\
+  unicode_literals
+
 import os
 import json
 import numpy as np
 import pickle
 import funcy as fy
 from sklearn.model_selection import train_test_split, StratifiedKFold
+
+import gc
 
 import ltag.chaining.pipeline as cp
 import ltag.datasets.utils as ds_utils
@@ -351,9 +356,11 @@ class StoredGraphDatasetManager(GraphDatasetManager):
       for ik in range(self.inner_k or 1):
         print(f"Preparing train fold {ok} {ik} of {self.name}...")
         self.get_train_fold(ok, ik, output_type="wl2")
+        gc.collect()
 
       print(f"Preparing test fold {ok} of {self.name}...")
       self.get_test_fold(ok, output_type="wl2")
+      gc.collect()
 
     print("Prepared all batch folds.")
 

### 2020-01-13 implementation/src/ltag/evaluate_datasets.py

@@ -1,0 +1,14 @@
+from __future__ import absolute_import, division, print_function,\
+  unicode_literals
+
+import ltag.evaluation.evaluate as evaluate
+import ltag.evaluation.models as models
+import ltag.evaluation.datasets as ds
+
+def run():
+  evaluate.evaluate(
+    models.AvgWL2GCN_Binary, ds.Proteins())
+
+
+if __name__ == "__main__":
+  run()

### 2020-01-13 implementation/src/ltag/evaluation/datasets.py

@@ -40,8 +40,15 @@ chemical = [
   DD
 ]
 
-# All:
+# Other categories:
 
-all_stored = chemical
+stored = chemical
 
-all = all_stored
+binary = [
+  Mutag,
+  NCI1,
+  Proteins,
+  DD
+]
+
+all = stored

### 2020-01-13 implementation/src/ltag/evaluation/evaluate.py

@@ -1,0 +1,72 @@
+from __future__ import absolute_import, division, print_function,\
+  unicode_literals
+
+import warnings
+from pathlib import Path
+from datetime import datetime
+from tensorflow import keras
+
+# Sparse gradient updates don't work for some reason. Disable the warning:
+warnings.filterwarnings(
+  "ignore",
+  "Converting sparse IndexedSlices to a dense Tensor of unknown shape.*",
+  UserWarning)
+
+eval_dir_base = Path("../evaluations")
+log_dir_base = Path("../logs")
+
+def time_str():
+  return datetime.now().strftime("%Y-%m-%d_%H-%M-%S")
+
+def train(
+  model, train_ds, val_ds=None, label=None,
+  log_dir_base=log_dir_base,
+  epochs=500, patience=50):
+  label = "_" + label if label is not None else ""
+
+  t = time_str()
+  tb = keras.callbacks.TensorBoard(
+    log_dir=log_dir_base / f"{t}{label}/",
+    histogram_freq=1,
+    write_images=False)
+  es = keras.callbacks.EarlyStopping(
+    monitor="loss" if val_ds is None else "val_loss",
+    patience=patience,
+    min_delta=0.0001,
+    restore_best_weights=True)
+
+  return model.fit(
+    train_ds, validation_data=val_ds,
+    epochs=epochs, callbacks=[tb, es])
+
+def evaluate(
+  model_factory, ds_manager,
+  outer_k=None, repeat=10, epochs=500):
+  outer_k = outer_k or ds_manager.outer_k
+
+  ds_type = model_factory.input_type
+  ds_name = ds_manager.name
+  t = time_str()
+  eval_dir = eval_dir_base / f"{t}_{ds_name}_{model_factory.name}"
+  log_dir_base = eval_dir / "logs"
+
+  for k in range(outer_k):
+    test_ds = ds_manager.get_test_fold(k, output_type=ds_type)
+    train_ds, val_ds = ds_manager.get_train_fold(k, output_type=ds_type)
+
+    for hp_i, model_ctr in enumerate(model_factory(train_ds)):
+      hp = model_ctr.hyperparams
+      print("\nSplit", k, "with hyperparams:", hp)
+
+      for i in range(repeat):
+        print("\nIteration", i, "of split", k, "...")
+
+        model = model_ctr()
+        label = f"{k}-{hp_i}-{i}"
+        h = train(
+          model, train_ds, val_ds, label,
[…]

### 2020-01-13 implementation/src/ltag/evaluation/models.py

@@ -1,4 +1,82 @@
 from __future__ import absolute_import, division, print_function,\
   unicode_literals
 
+from tensorflow import keras
+import itertools
+import funcy as fy
+
 import ltag.models as models
+
+def model_factory(mf):
+  def model_factory_instance(model_class):
+    def model_factory_with_class(hyperparam_gen):
+      @fy.wraps(hyperparam_gen)
+      def wrapper(ds, **params):
+        hps = hyperparam_gen(ds, **params)
+
+        def mf_caller(hp):
+          model_ctr = fy.partial(mf, model_class, **hp)
+          model_ctr.hyperparams = hp
+          return model_ctr
+
+        return fy.map(mf_caller, hps)
+
+      wrapper.input_type = model_class.input_type
+      wrapper.name = model_class.name
+
+      return wrapper
+
+    return model_factory_with_class
+
+  return model_factory_instance
+
+@model_factory
+def binary_classifier(model_class, learning_rate=0.001, **params):
+  model = model_class(act="sigmoid", squeeze_output=True, **params)
+
+  opt = keras.optimizers.Adam(learning_rate)
+
+  model.compile(
+    optimizer=opt,
+    loss="binary_crossentropy",
+    metrics=["accuracy"])
+
+  return model
+
+@model_factory
+def logistic_regression(
+  model_class, act="tanh", learning_rate=0.001, **params):
+  model = model_class(act=act, squeeze_output=True, **params)
+
+  opt = keras.optimizers.Adam(learning_rate)
+
+  model.compile(
+    optimizer=opt,
+    loss="mse",
+    metrics=["mae"])
+
+  return model
+
+def cart(*pos_params, **params):
+  if len(pos_params) > 0:
+    return itertools.product(*pos_params)
+
+  return (dict(zip(params, x)) for x in itertools.product(*params.values()))
+
+def wl2_in_dim(ds):
+  return ds.element_spec[0][0].shape[-1]
+
+
+@binary_classifier(models.AvgWL2GCN)
+def AvgWL2GCN_Binary(ds):
+  in_dim = wl2_in_dim(ds)
+
+  hidden = [
+    [b] * l
+    for b, l in cart([8, 32], [1, 3])]
+
+  return cart(
+    layer_dims=[[in_dim, *h, 1] for h in hidden],
+    bias=[True, False],
+    learning_rate=[0.01, 0.001, 0.0001]
+  )

### 2020-01-13 implementation/src/ltag/evaluation/scenarios.py

@@ -1,33 +1,0 @@
-from __future__ import absolute_import, division, print_function,\
-  unicode_literals
-
-from tensorflow import keras
-
-import ltag.models as models
-
-import ltag.datasets.disk.tu.datasets as tu
-import ltag.datasets.synthetic as synthetic
-
-learning_rates = [0.001]
-
-adams = [keras.optimizers.Adam(lr) for lr in learning_rates]
-
-Proteins = {
-  "dataset": {
-    "manager": tu.Proteins,
-    "wl2_neighborhood": 6,
-    "wl2_batch_size": {
-      "fuzzy_batch_edge_count": 10000,
-      "upper_batch_edge_count": 30000,
-      "batch_graph_count": 20
-    }
-  },
-  "model": {
-    "class": models.SortWL2GCN,
-    "loss": "binary_crossentropy",
-    "metrics": ["accuracy"]
-  },
-  "hyperparams": {
-    "optimizer": adams,
-  }
-}

### 2020-01-13 implementation/src/ltag/main.py

@@ -1,102 +1,10 @@
 from __future__ import absolute_import, division, print_function,\
   unicode_literals
 
-import tensorflow as tf
 from tensorflow import keras
 import numpy as np
 from datetime import datetime
 
-import ltag.models as models
-import ltag.datasets.synthetic as synthetic
-import ltag.datasets.disk.tu.datasets as tu
+import ltag.evaluate_datasets as evaluate
 
-def binary_classifier(model_class, learning_rate=0.001, **params):
-  model = model_class(act="sigmoid", squeeze_output=True, **params)
-
-  opt = keras.optimizers.Adam(learning_rate)
-
-  model.compile(
-    optimizer=opt,
-    loss="binary_crossentropy",
-    metrics=["accuracy"])
-
-  return model
-
-def logistic_regression(
-  model_class, act="tanh", learning_rate=0.001, **params):
-  model = model_class(act=act, squeeze_output=True, **params)
-
-  opt = keras.optimizers.Adam(learning_rate)
-
-  model.compile(
-    optimizer=opt,
-    loss="mse",
-    metrics=["mae"])
-
-  return model
-
-
-log_dir = "../logs"
-model_class = models.SortWL2GCN
-
-# ds_raw = synthetic.triangle_dataset()(wl2_neighborhood=3).get_all(
-#   model_class.input_type, shuffle=True)
-ds_manager = tu.Proteins(
-  wl2_neighborhood=6,
-  wl2_batch_size={
-    "fuzzy_batch_edge_count": 10000,
-    "upper_batch_edge_count": 30000,
-    "batch_graph_count": 20
-  })
-
-def model_factory(ds):
-  in_dim = ds.element_spec[0][0].shape[-1]
-
-  return binary_classifier(
-    model_class, layer_dims=[in_dim, 8, 8, 8, 1],
-    bias=True, k_pool=512)
-
-
-model_factory.input_type = model_class.input_type
-
-
-def train(model, train_ds, val_ds=None, label=None):
-  label = "_" + label if label is not None else ""
-
-  t = datetime.now().strftime("%Y-%m-%d_%H-%M-%S")
-  tb = keras.callbacks.TensorBoard(
-    log_dir=f"{log_dir}/{t}{label}/",
-    histogram_freq=1,
-    write_images=False)
-  es = tf.keras.callbacks.EarlyStopping(
-    monitor="loss" if val_ds is None else "val_loss",
-    patience=50,
-    min_delta=0.0001,
[…]

### 2020-01-13 implementation/src/ltag/models.py

@@ -44,12 +44,10 @@ def gnn_model(name, steps, input_type="dense"):
 
   def add_attrs(m):
     m.input_type = input_type
+    m.extend = fy.rcompose(m.extend, add_attrs)
     return m
 
-  add_attrs(m)
-  m.extend = fy.rcompose(m.extend, add_attrs)
-
-  return m
+  return add_attrs(m)
 
 
 # Non-aggregating GNNs:

### 2020-01-13 implementation/src/ltag/prepare_datasets.py

@@ -6,7 +6,7 @@ import multiprocessing as mp
 import ltag.evaluation.datasets as datasets
 
 def prepare_ds(i):
-  manager = datasets.all_stored[i](no_wl2_load=True)
+  manager = datasets.stored[i](no_wl2_load=True)
   print(f"Preparing {manager.name}...")
   manager.prepare_wl2_batches()
 
@@ -15,6 +15,6 @@ if __name__ == "__main__":
   print("Preparing all stored datasets...")
 
   with mp.Pool(mp.cpu_count()) as p:
-    p.map(prepare_ds, range(len(datasets.all_stored)))
+    p.map(prepare_ds, range(len(datasets.stored)))
 
   print("Prepared all stored datasets.")

### 2020-01-14 implementation/evaluate_datasets.sh
implemented evaluation logging and continuation
- the evaluation loop now logs all observed train, validation and test metrics
- it also stores the required metadata to continue an interrupted evaluation

@@ -2,4 +2,7 @@
 
 USER="-u $(id -u):$(id -g)"
 
-docker exec -it $USER $(docker ps -aqf "name=ltag") python3 ./ltag/evaluate_datasets.py
+docker exec -it $USER $(docker ps -aqf "name=ltag") \
+	python3 ./ltag/evaluate_datasets.py \
+	| grep --line-buffered -vE \
+	"BaseCollectiveExecutor::StartAbort|IteratorGetNext|Shape/"

### 2020-01-14 implementation/src/ltag/datasets/manager.py

@@ -12,14 +12,9 @@ import gc
 
 import ltag.chaining.pipeline as cp
 import ltag.datasets.utils as ds_utils
+from ltag.utils import NumpyEncoder
 
 # Implementation adapted from https://github.com/diningphil/gnn-comparison.
-
-class NumpyEncoder(json.JSONEncoder):
-  def default(self, obj):
-    if isinstance(obj, np.ndarray):
-      return obj.tolist()
-    return json.JSONEncoder.default(self, obj)
 
 
 class GraphDatasetManager:
@@ -110,6 +105,10 @@ class GraphDatasetManager:
   @property
   def dim_edge_features(self):
     return self._dim_edge_features
+
+  @classmethod
+  def dim_wl2_features(cls):
+    return 3 + cls._dim_node_features + cls._dim_edge_features
 
   @property
   def max_num_nodes(self):
@@ -203,7 +202,10 @@ class GraphDatasetManager:
       targets = targets[idxs]
 
     batches = ds_utils.to_wl2_ds(
-      graphs, targets, as_list=True,
+      graphs, targets,
+      dim_node_features=self._dim_node_features,
+      dim_edge_features=self._dim_edge_features,
+      as_list=True,
       preencoded=self.wl2_cache,
       neighborhood=self.wl2_neighborhood,
       **self.wl2_batch_size)
@@ -351,6 +353,10 @@ class StoredGraphDatasetManager(GraphDatasetManager):
 
   def prepare_wl2_batches(self):
     assert self.wl2_batch_cache
+
+    print(f"Preparing full dataset batches of {self.name}...")
+    self.get_all(output_type="wl2")
+    gc.collect()
 
     for ok in range(self.outer_k or 1):
       for ik in range(self.inner_k or 1):

### 2020-01-14 implementation/src/ltag/datasets/utils.py

@@ -181,6 +181,8 @@ def wl2_batches_to_dataset(batches, dim_wl2, y_dim):
 @cp.tolerant
 def to_wl2_ds(
   graphs, ys,
+  dim_node_features,
+  dim_edge_features,
   fuzzy_batch_edge_count=100000,
   upper_batch_edge_count=120000,
   batch_graph_count=100,
@@ -191,23 +193,7 @@ def to_wl2_ds(
   il = np.arange(ds_size)
   y_shape = ys.shape
   y_dim = y_shape[1:] if len(y_shape) > 1 else []
-
-  dim_node_features = 0
-  dim_edge_features = 0
-  dim_wl2 = 0
-
-  if ds_size > 0:
-    g = graphs[0]
-
-    if preencoded:
-      dim_node_features = "?"
-      dim_edge_features = "?"
-      dim_wl2 = len(g[0][0])
-    else:
-      dim_wl2 = 3
-      dim_node_features, dim_edge_features = get_graph_feature_dims(g)
-      dim_wl2 += dim_node_features
-      dim_wl2 += dim_edge_features
+  dim_wl2 = 3 + dim_node_features + dim_edge_features
 
   if log:
     print(

### 2020-01-14 implementation/src/ltag/evaluate_datasets.py

@@ -5,9 +5,21 @@ import ltag.evaluation.evaluate as evaluate
 import ltag.evaluation.models as models
 import ltag.evaluation.datasets as ds
 
-def run():
+epochs = 500
+repeat = 5  # fewer repeats for now.
+
+mf = models.AvgWL2GCN_Binary
+dsm = ds.Proteins
+
+def run(**kwargs):
   evaluate.evaluate(
-    models.AvgWL2GCN_Binary, ds.Proteins())
+    mf, dsm(), epochs=epochs, repeat=repeat,
+    **kwargs)
+
+def resume(eval_dir_name, **kwargs):
+  return evaluate.resume_evaluation(
+    mf, dsm(), evaluate.eval_dir_base / eval_dir_name,
+    **kwargs)
 
 
 if __name__ == "__main__":

### 2020-01-14 implementation/src/ltag/evaluation/evaluate.py

@@ -1,10 +1,16 @@
 from __future__ import absolute_import, division, print_function,\
   unicode_literals
 
+import os
+import json
+from timeit import default_timer as timer
 import warnings
 from pathlib import Path
 from datetime import datetime
 from tensorflow import keras
+import funcy as fy
+
+from ltag.utils import NumpyEncoder
 
 # Sparse gradient updates don't work for some reason. Disable the warning:
 warnings.filterwarnings(
@@ -21,7 +27,9 @@ def time_str():
 def train(
   model, train_ds, val_ds=None, label=None,
   log_dir_base=log_dir_base,
-  epochs=500, patience=50):
+  epochs=500, patience=50, restore_best=False,
+  stopping_min_delta=0.0001,
+  verbose=2):
   label = "_" + label if label is not None else ""
 
   t = time_str()
@@ -32,41 +40,187 @@ def train(
   es = keras.callbacks.EarlyStopping(
     monitor="loss" if val_ds is None else "val_loss",
     patience=patience,
-    min_delta=0.0001,
-    restore_best_weights=True)
+    min_delta=stopping_min_delta,
+    restore_best_weights=restore_best)
 
   return model.fit(
     train_ds, validation_data=val_ds,
-    epochs=epochs, callbacks=[tb, es])
+    epochs=epochs, callbacks=[tb, es],
+    verbose=verbose)
+
+def summarize_evaluation(eval_dir):
+  pass
 
 def evaluate(
   model_factory, ds_manager,
-  outer_k=None, repeat=10, epochs=500):
+  outer_k=None, repeat=10, epochs=500,
+  patience=50, stopping_min_delta=0.0001,
+  restore_best=False,
+  eval_dir=None, verbose=2):
   outer_k = outer_k or ds_manager.outer_k
+  inner_k = None
 
+  mf_name = model_factory.name
   ds_type = model_factory.input_type
   ds_name = ds_manager.name
   t = time_str()
-  eval_dir = eval_dir_base / f"{t}_{ds_name}_{model_factory.name}"
+  if eval_dir is None:
+    eval_dir = eval_dir_base / f"{t}_{ds_name}_{mf_name}"
+    if not eval_dir.exists():
+      os.makedirs(eval_dir)
+
+    with open(eval_dir / "config.json", "w") as f:
+      config = {
+        "outer_k": outer_k,
+        "inner_k": inner_k,
+        "repeat": repeat,
+        "epochs": epochs,
+        "patience": patience,
+        "stopping_min_delta": stopping_min_delta,
[…]

### 2020-01-14 implementation/src/ltag/evaluation/models.py

@@ -11,15 +11,9 @@ def model_factory(mf):
   def model_factory_instance(model_class):
     def model_factory_with_class(hyperparam_gen):
       @fy.wraps(hyperparam_gen)
-      def wrapper(ds, **params):
-        hps = hyperparam_gen(ds, **params)
-
-        def mf_caller(hp):
-          model_ctr = fy.partial(mf, model_class, **hp)
-          model_ctr.hyperparams = hp
-          return model_ctr
-
-        return fy.map(mf_caller, hps)
+      def wrapper(ds_manager, **params):
+        hps = list(hyperparam_gen(ds_manager, **params))
+        return fy.partial(mf, model_class), hps
 
       wrapper.input_type = model_class.input_type
       wrapper.name = model_class.name
@@ -63,20 +57,22 @@ def cart(*pos_params, **params):
 
   return (dict(zip(params, x)) for x in itertools.product(*params.values()))
 
-def wl2_in_dim(ds):
-  return ds.element_spec[0][0].shape[-1]
-
 
 @binary_classifier(models.AvgWL2GCN)
-def AvgWL2GCN_Binary(ds):
-  in_dim = wl2_in_dim(ds)
+def AvgWL2GCN_Binary(dsm):
+  in_dim = dsm.dim_wl2_features()
 
   hidden = [
     [b] * l
     for b, l in cart([8, 32], [1, 3])]
 
+  hidden = [
+    [8], [8, 8, 8],
+    [32], [32, 32]
+  ]
+
   return cart(
     layer_dims=[[in_dim, *h, 1] for h in hidden],
-    bias=[True, False],
-    learning_rate=[0.01, 0.001, 0.0001]
+    bias=[True],
+    learning_rate=[0.0001]
   )

### 2020-01-14 implementation/src/ltag/main.py

@@ -5,6 +5,40 @@ from tensorflow import keras
 import numpy as np
 from datetime import datetime
 
-import ltag.evaluate_datasets as evaluate
+import ltag.models as models
+import ltag.datasets.disk.tu.datasets as tu
+import ltag.datasets.synthetic as synthetic
+import ltag.evaluation.datasets as eval_ds
+import ltag.evaluation.evaluate as evaluate
 
-evaluate.run()
+import ltag.evaluate_datasets as eval_main
+
+
+def proteins_experient():
+  model_class = models.AvgWL2GCN
+  dsm_class = eval_ds.Proteins
+  dsm = dsm_class()
+
+  model = model_class(
+    act="sigmoid", squeeze_output=True,
+    layer_dims=[dsm.dim_wl2_features(), 32, 32, 1],
+    bias=True)
+
+  opt = keras.optimizers.Adam(0.001)
+
+  model.compile(
+    optimizer=opt,
+    loss="binary_crossentropy",
+    metrics=["accuracy"])
+
+  ds = dsm.get_all(
+    output_type=model_class.input_type)
+
+  evaluate.train(
+    model, ds, verbose=1,
+    label=f"{dsm.name}_{model.name}")
+
+eval_main.run(verbose=1)
+# eval_main.resume(
+#   "2020-01-14_22-03-00_PROTEINS_full_AvgWL2GCN",
+#   verbose=1)

### 2020-01-14 implementation/src/ltag/prepare_datasets.py

@@ -2,6 +2,7 @@ from __future__ import absolute_import, division, print_function,\
   unicode_literals
 
 import multiprocessing as mp
+import argparse
 
 import ltag.evaluation.datasets as datasets
 
@@ -12,9 +13,17 @@ def prepare_ds(i):
 
 
 if __name__ == "__main__":
-  print("Preparing all stored datasets...")
+  parser = argparse.ArgumentParser(
+    description='Prebatch WL2 encoded datasets.')
+  parser.add_argument(
+    "-p", "--parallel", action="store_true",
+    help="Preprocess datasets in parallel.")
+  args = parser.parse_args()
+  p = mp.cpu_count() if args.parallel else 1
 
-  with mp.Pool(mp.cpu_count()) as p:
+  print(f"Preparing all stored datasets with parallelism {p}...")
+
+  with mp.Pool(p) as p:
     p.map(prepare_ds, range(len(datasets.stored)))
 
   print("Prepared all stored datasets.")

### 2020-01-14 implementation/src/ltag/utils.py

@@ -1,0 +1,17 @@
+from __future__ import absolute_import, division, print_function,\
+  unicode_literals
+
+import json
+import numpy as np
+
+class NumpyEncoder(json.JSONEncoder):
+  def default(self, obj):
+    if isinstance(obj, np.ndarray):
+      return obj.tolist()
+    if (
+      isinstance(obj, np.float32)
+      or isinstance(obj, np.float64)
+      or isinstance(obj, np.int32)
+      or isinstance(obj, np.int64)):
+      return np.asscalar(obj)
+    return json.JSONEncoder.default(self, obj)

### 2020-01-15 implementation/src/ltag/evaluate_datasets.py
added evaluation summarization
+ quick evaluation feature to assess whether the wl2 batches fit in memory when evaluated with the relevant set of model configurations for an evaluation.

@@ -2,6 +2,7 @@ from __future__ import absolute_import, division, print_function,\
   unicode_literals
 
 import ltag.evaluation.evaluate as evaluate
+import ltag.evaluation.summary as summary
 import ltag.evaluation.models as models
 import ltag.evaluation.datasets as ds
 
@@ -9,7 +10,10 @@ epochs = 500
 repeat = 5  # fewer repeats for now.
 
 mf = models.AvgWL2GCN_Binary
-dsm = ds.Proteins
+dsm = ds.NCI1_8
+
+def quick_run(**kwargs):
+  evaluate.quick_evaluate(mf, dsm(), **kwargs)
 
 def run(**kwargs):
   evaluate.evaluate(
@@ -21,6 +25,9 @@ def resume(eval_dir_name, **kwargs):
     mf, dsm(), evaluate.eval_dir_base / eval_dir_name,
     **kwargs)
 
+
+def summarize(eval_dir_name):
+  return summary.summarize_evaluation(evaluate.eval_dir_base / eval_dir_name)
 
 if __name__ == "__main__":
   run()

### 2020-01-15 implementation/src/ltag/evaluation/datasets.py

@@ -7,15 +7,15 @@ import ltag.datasets.disk.tu.datasets as tu
 
 # Chemical:
 
-Mutag = fy.partial(
+Mutag_8 = fy.partial(
   tu.Mutag,
   wl2_neighborhood=8)
 
-NCI1 = fy.partial(
+NCI1_8 = fy.partial(
   tu.NCI1,
   wl2_neighborhood=8)
 
-Proteins = fy.partial(
+Proteins_6 = fy.partial(
   tu.Proteins,
   wl2_neighborhood=6,
   wl2_batch_size={
@@ -24,7 +24,7 @@ Proteins = fy.partial(
     "batch_graph_count": 20
   })
 
-DD = fy.partial(
+DD_2 = fy.partial(
   tu.DD,
   wl2_neighborhood=2,
   wl2_batch_size={
@@ -34,10 +34,10 @@ DD = fy.partial(
   })
 
 chemical = [
-  Mutag,
-  NCI1,
-  Proteins,
-  DD
+  Mutag_8,
+  NCI1_8,
+  Proteins_6,
+  DD_2
 ]
 
 # Other categories:
@@ -45,10 +45,10 @@ chemical = [
 stored = chemical
 
 binary = [
-  Mutag,
-  NCI1,
-  Proteins,
-  DD
+  Mutag_8,
+  NCI1_8,
+  Proteins_6,
+  DD_2
 ]
 
 all = stored

### 2020-01-15 implementation/src/ltag/evaluation/evaluate.py

@@ -11,6 +11,7 @@ from tensorflow import keras
 import funcy as fy
 
 from ltag.utils import NumpyEncoder
+import ltag.evaluation.summary as summary
 
 # Sparse gradient updates don't work for some reason. Disable the warning:
 warnings.filterwarnings(
@@ -48,15 +49,13 @@ def train(
     epochs=epochs, callbacks=[tb, es],
     verbose=verbose)
 
-def summarize_evaluation(eval_dir):
-  pass
-
 def evaluate(
   model_factory, ds_manager,
   outer_k=None, repeat=10, epochs=500,
   patience=50, stopping_min_delta=0.0001,
-  restore_best=False,
+  restore_best=False, label=None,
   eval_dir=None, verbose=2):
+  label = "_" + label if label is not None else ""
   outer_k = outer_k or ds_manager.outer_k
   inner_k = None
 
@@ -65,7 +64,7 @@ def evaluate(
   ds_name = ds_manager.name
   t = time_str()
   if eval_dir is None:
-    eval_dir = eval_dir_base / f"{t}_{ds_name}_{mf_name}"
+    eval_dir = eval_dir_base / f"{t}_{ds_name}_{mf_name}{label}"
     if not eval_dir.exists():
       os.makedirs(eval_dir)
 
@@ -205,7 +204,7 @@ def evaluate(
       config["end_time"] = time_str()
       json.dump(config, f, indent="\t", sort_keys=True, cls=NumpyEncoder)
 
-  summarize_evaluation(eval_dir)
+  summary.summarize_evaluation(eval_dir)
   print(
     time_str(),
     f"- Evaluation of {ds_name} using {mf_name} completed in {dur_eval}s.")
@@ -223,4 +222,10 @@ def resume_evaluation(model_factory, ds_manager, eval_dir, **kwargs):
     stopping_min_delta=config["stopping_min_delta"],
     restore_best=config["restore_best"],
     eval_dir=eval_dir,
+    **kwargs)
+
+def quick_evaluate(model_factory, ds_manager, **kwargs):
+  return evaluate(
+    model_factory, ds_manager,
+    epochs=1, repeat=1, label="quick",
     **kwargs)

### 2020-01-15 implementation/src/ltag/evaluation/summary.py

@@ -1,0 +1,105 @@
+from __future__ import absolute_import, division, print_function,\
+  unicode_literals
+
+import os
+import json
+from collections import defaultdict
+import funcy as fy
+import numpy as np
+
+from ltag.utils import NumpyEncoder
+
+selection_metrics = {
+  "accuracy": "max",
+  "loss": "min",
+  "mse": "min",
+  "val_accuracy": "max",
+  "val_loss": "min",
+  "val_mse": "min"
+}
+
+def dict_map(f, d):
+  return {k: f(v) for k, v in d.items()}
+
+def statistics(vals):
+  return {
+    "mean": np.mean(vals),
+    "std": np.std(vals),
+    "min": np.min(vals),
+    "max": np.max(vals)
+  }
+
+def summarize_evaluation(
+  eval_dir, selection_metric="val_accuracy"):
+  assert eval_dir.exists(), f"No evalutation '{eval_dir}' found."
+  with open(eval_dir / "config.json") as f:
+    config = json.load(f)
+
+  with open(eval_dir / "hyperparams.json") as f:
+    hps = json.load(f)
+
+  results_dir = eval_dir / "results"
+  assert results_dir.exists(), f"No results found for '{eval_dir}'."
+  summary_dir = eval_dir / "summary"
+
+  if not summary_dir.exists():
+    os.makedirs(summary_dir)
+
+  result_files = [
+    (list(fy.map(int, f[:-5].split("-"))), results_dir / f)
+    for f in os.listdir(results_dir)]
+
+  fold_files = fy.group_by(lambda f: f[0][0], result_files)
+  fold_param_files = {
+    fold: fy.group_by(lambda f: f[0][1], files)
+    for fold, files in fold_files.items()}
+
+  best_goal = selection_metrics[selection_metric]
+
+  results = []
+
+  for fold_i, param_files in fold_param_files.items():
+    best_result = None
+
+    for hp_i, files in param_files.items():
+      hp_results = defaultdict(list)
+      selection_vals = []
+      for _, file in files:
+        with open(file, "r") as f:
+          result = json.load(f)
+
+        selection_vals.append(result["train"][selection_metric][-1])
+
+        for metric, val in result["test"].items():
+          hp_results[metric].append(val)
+
+      hp_result = (
+        dict_map(np.mean, hp_results),
+        hp_i, np.mean(selection_vals))
+
+      if (
+        best_result is None
+        or (best_goal == "max" and best_result[2] < hp_result[2])
[…]

### 2020-01-15 implementation/src/ltag/main.py

@@ -38,7 +38,10 @@ def proteins_experient():
     model, ds, verbose=1,
     label=f"{dsm.name}_{model.name}")
 
-eval_main.run(verbose=1)
+
+# eval_main.run(verbose=1)
+# eval_main.quick_run()
 # eval_main.resume(
-#   "2020-01-14_22-03-00_PROTEINS_full_AvgWL2GCN",
+#   "2020-01-15_14-26-15_NCI1_AvgWL2GCN_quick",
 #   verbose=1)
+eval_main.summarize("2020-01-15_14-31-53_NCI1_AvgWL2GCN")

### 2020-02-09 implementation/src/ltag/evaluation/datasets.py
use 5 instead of 6 neighborhood for proteins dataset
to deal with OOM issues

@@ -23,6 +23,15 @@ NCI1_8 = fy.partial(
   tu.NCI1,
   wl2_neighborhood=8)
 
+Proteins_5 = fy.partial(
+  tu.Proteins,
+  wl2_neighborhood=5,
+  wl2_batch_size={
+    "fuzzy_batch_edge_count": 20000,
+    "upper_batch_edge_count": 50000,
+    "batch_graph_count": 20
+  })
+
 Proteins_6 = fy.partial(
   tu.Proteins,
   wl2_neighborhood=6,
@@ -43,7 +52,7 @@ DD_2 = fy.partial(
 
 chemical_binary = [
   "NCI1_8",
-  "Proteins_6",
+  "Proteins_5",
   "DD_2"
 ]
 chemical = chemical_binary
@@ -76,9 +85,9 @@ noisy_triangle_classification_2 = fy.partial(
   evaluation_args={
     "epochs": 3000,
     "hp_args": {
-      "cwl2_local_act": "relu",
       # other widths ignored based on previous experiments:
-      "layer_widths": [32]
+      "cwl2_layer_widths": [32],
+      "cwl2_local_act": "relu"
     }
   })
 

### 2020-02-09 implementation/src/ltag/evaluation/models.py

@@ -66,18 +66,21 @@ def AvgWL2GCN_FC_Binary(dsm):
 
 @binary_classifier(gnn_models.AvgCWL2GCN)
 def AvgCWL2GCN_Binary(
-  dsm, cwl2_local_act="sigmoid", layer_widths=[32, 64]):
+  dsm,
+  cwl2_local_act="sigmoid",
+  cwl2_layer_widths=[32, 64],
+  cwl2_stack_tfs=[None, "keep_input"]):
   in_dim = dsm.dim_wl2_features()
 
   hidden = [
    [b] * l
-   for b, l in cart(layer_widths, [1, 3])]
+   for b, l in cart(cwl2_layer_widths, [1, 3])]
 
   return cart(
     conv_layer_dims=[[in_dim, *h, 1] for h in hidden],
     conv_act=["sigmoid"],
     conv_local_act=[cwl2_local_act],
-    conv_stack_tf=[None, "keep_input"],
+    conv_stack_tf=cwl2_stack_tfs,
     conv_bias=[True],
     learning_rate=[0.01, 0.001, 0.0001],
     squeeze_output=[True]
@@ -85,18 +88,21 @@ def AvgCWL2GCN_Binary(
 
 @binary_classifier(gnn_models.SagCWL2GCN)
 def SagCWL2GCN_Binary(
-  dsm, cwl2_local_act="sigmoid", layer_widths=[32, 64]):
+  dsm,
+  cwl2_local_act="sigmoid",
+  cwl2_layer_widths=[32, 64],
+  cwl2_stack_tfs=[None, "keep_input"]):
   in_dim = dsm.dim_wl2_features()
 
   hidden = [
    [b] * l
-   for b, l in cart(layer_widths, [1, 3])]
+   for b, l in cart(cwl2_layer_widths, [1, 3])]
 
   hps = cart(
     conv_layer_dims=[[in_dim, *h, 1] for h in hidden],
     conv_act=["sigmoid"],
     conv_local_act=[cwl2_local_act],
-    conv_stack_tf=[None, "keep_input"],
+    conv_stack_tf=cwl2_stack_tfs,
     conv_bias=[True],
     learning_rate=[0.01, 0.001, 0.0001],
     squeeze_output=[True]
@@ -114,18 +120,21 @@ def SagCWL2GCN_Binary(
 
 @binary_classifier(gnn_models.SagCWL2GCN)
 def SagCWL2GCN_Binary_quick_max(
[…]

### 2020-02-09 implementation/evaluate_datasets.sh
better eval output filter

@@ -8,4 +8,4 @@ fi
 
 docker exec -it $USER $(docker ps -aqf "name=^$LTAG_CONTAINER_NAME\$") python3 ./ltag/evaluate_datasets.py $@ \
 	| grep --line-buffered -vE \
-	"BaseCollectiveExecutor::StartAbort|IteratorGetNext|Shape/"
+	"BaseCollectiveExecutor::StartAbort|IteratorGetNext|Shape/|Shape_[0-9]+/"

### 2020-02-10 implementation/src/ltag/evaluation/datasets.py
added imdb splits + minor eval config changes

@@ -84,9 +84,11 @@ noisy_triangle_classification_2 = fy.partial(
   },
   evaluation_args={
     "epochs": 3000,
+    "patience": 600,
     "hp_args": {
       # other widths ignored based on previous experiments:
       "cwl2_layer_widths": [32],
+      "cwl2_layer_depths": [3],
       "cwl2_local_act": "relu"
     }
   })

### 2020-02-10 implementation/src/ltag/evaluation/evaluate.py

@@ -116,7 +116,7 @@ def find_eval_dir(model_factory, ds_manager, label=None):
 def evaluate(
   model_factory, ds_manager,
   outer_k=None, repeat=1, winner_repeat=3, epochs=1000,
-  patience=50, stopping_min_delta=0.0001,
+  patience=100, stopping_min_delta=0.0001,
   restore_best=False, hp_args=None, label=None,
   eval_dir=None, verbose=2, dry=False):
   outer_k = outer_k or ds_manager.outer_k

### 2020-02-10 implementation/src/ltag/evaluation/models.py

@@ -69,12 +69,13 @@ def AvgCWL2GCN_Binary(
   dsm,
   cwl2_local_act="sigmoid",
   cwl2_layer_widths=[32, 64],
+  cwl2_layer_depths=[1, 3],
   cwl2_stack_tfs=[None, "keep_input"]):
   in_dim = dsm.dim_wl2_features()
 
   hidden = [
    [b] * l
-   for b, l in cart(cwl2_layer_widths, [1, 3])]
+   for b, l in cart(cwl2_layer_widths, cwl2_layer_depths)]
 
   return cart(
     conv_layer_dims=[[in_dim, *h, 1] for h in hidden],
@@ -91,12 +92,13 @@ def SagCWL2GCN_Binary(
   dsm,
   cwl2_local_act="sigmoid",
   cwl2_layer_widths=[32, 64],
+  cwl2_layer_depths=[1, 3],
   cwl2_stack_tfs=[None, "keep_input"]):
   in_dim = dsm.dim_wl2_features()
 
   hidden = [
    [b] * l
-   for b, l in cart(cwl2_layer_widths, [1, 3])]
+   for b, l in cart(cwl2_layer_widths, cwl2_layer_depths)]
 
   hps = cart(
     conv_layer_dims=[[in_dim, *h, 1] for h in hidden],
@@ -123,12 +125,13 @@ def SagCWL2GCN_Binary_quick_max(
   dsm,
   cwl2_local_act="sigmoid",
   cwl2_layer_widths=[64],
+  cwl2_layer_depths=[3],
   cwl2_stack_tfs=["keep_input"]):
   in_dim = dsm.dim_wl2_features()
 
   hidden = [
    [b] * l
-   for b, l in cart(cwl2_layer_widths, [3])]
+   for b, l in cart(cwl2_layer_widths, cwl2_layer_depths)]
 
   hps = cart(
     conv_layer_dims=[[in_dim, *h, 1] for h in hidden],

### 2020-02-10 thesis/content/chapter-introduction.tex
continued related works chapter

@@
 \end{enumerate}
-In this thesis we will focus on the last problem type, \ac{gcr}.
-A \ac{ml} method for this problem has to accept graphs of varying size and should be permutation invariant wrt.\ the vertices.
+In this thesis we will focus on the last problem type, \acf{gcr}.
+An \ac{ml} method for this problem has to accept graphs of varying size and should be permutation invariant wrt.\ the vertices.
 Those requirements are not met by the commonly used learners that only accept fixed-size feature vectors as their input, e.g.\ \acp*{lm}, \acp{svm} or \acp{mlp}.

### 2020-02-10 thesis/content/chapter-related-work.tex
implemented evaluation logging and continuation
- the evaluation loop now logs all observed train, validation and test metrics
- it also stores the required metadata to continue an interrupted evaluation
continued related works chapter

@@
 \begin{enumerate}
-	\item We begin with an overview of the existing \ac{lta} methods for set inputs.
-	\item Before the existing \ac{gcr} methods are described, we look at the graph theoretical fundamentals on which those methods are based.
-	\item Using the previously introduced fundamentals, the current approaches to tackle the \ac{gcr} problem are then described.
+	\item We begin with an overview of the existing \ac{lta} methods for unstructured inputs.
+	\item Then we look at the domain of structured inputs.
+		To solve the \ac{gcr} problem, relevant graphs characteristics have to be defined in order to determine the similarity and dissimilarity of graphs.
+		We will look at two common approaches for graph characterization: The \acl{wl} algorithm and the notion of graph spectra.
+	\item Using the described graph characterization approaches, a brief overview of current \ac{gcr} methods will then be given.
 \end{enumerate}
@@
 
-\section{Graph Theoretical Fundamentals}%
-\label{sec:related:theory}
+\section{Graph Characterization}%
+\label{sec:related:character}
 
-\subsection{Graph Isomorphism Checks}%
-\label{sec:related:theory:wl}
+To classify or score a graph, it first needs to be characterized by a set of relevant properties.
+Two commonly used graph characterization approaches are the so called \acf{wl} colors and the notion of the graph spectrum.
+Both will now be briefly introduced since they form the foundation of most current \ac{gcr} methods.
+
+\subsection{Weisfeiler-Lehman Graph Colorings}%
+\label{sec:related:character:wl}
+
+The Weisfeiler-Lehman algorithm characterizes a graph $G = (\mathcal{V}, \mathcal{E})$ by assigning discrete labels $c \in \mathcal{C}$, called \textit{colors}, to vertex $k$-tuples $(v_1, \dots, v_k) \in \mathcal{V}^k$, where $k \in \mathbb{N}$ is the freely choosable \textit{\ac{wl}-dimension}.
+A mapping $\chi: \mathcal{V}^k \to \mathcal{C}$ is called a coloring of $G = (\mathcal{V}, \mathcal{E})$.
+
+\paragraph{The 1-dimensional \acs{wl} algorithm}
+For the most simple case of $k = 1$, a color is assigned to each vertex of a graph.
 
 \subsection{Spectral Graph Theory}%
-\label{sec:related:theory:spectral}
+\label{sec:related:character:spectral}
 
-\section{Graph Regression and Classification}%
+\section{Graph Classification and Regression}%
 \label{sec:related:gcr}
@@
 \end{enumerate*}.
-We will now look at the characteristics of those families and give an overview of specific methods.
+We will now look at the characteristics of those families and give a brief overview of specific methods.
 
@@
 
-\subsubsection{Fingerprint embeddings}
+\subsubsection{Fingerprint Embeddings}
 
@@
 
-\subsubsection{Skip-gram inspired embeddings}
+\subsubsection{Skip-gram inspired Embeddings}
 
@@
 
-\paragraph{Vertex embeddings}
+\paragraph{Vertex Embeddings}
 Skip-gram embeddings can be  na{\"\i}vely extended to graphs by realizing that \texttt{word2vec} effectively already is a vertex embedding method for linear graphs in which $C_k(v)$ is simply the $k$-neighborhood of the vertex/word $v$.
@@
 
-\paragraph{Graph embeddings}
+\paragraph{Graph Embeddings}
 Skip-gram methods can not only be used for vertex embeddings but also to embed entire graphs.
@@
 	C_k(G_i) := \bigcup_{\mathclap{v_j \in \mathcal{V}_i}}\, {\left\{ {\textrm{WL}_l(G_i)}_j \right\}}_{l = 0}^k % chktex 21
-	\text{ with } \textrm{WL}_l(G) := \text{\acs{wl} colors after $l$ iter.\ on $G$.}
+	\text{ with } \textrm{WL}_l(G) := \text{$l$ iter.\ \acs{wl}-1 coloring of $G$.}
 \end{equation}
@@
 
-Instead of explicity mapping a graph or vertices into a vector space, one can also do so implicity by employing the kernel trick.
+Instead of explicitly mapping a graph or vertices into a vector space, one can also do so implicitly by employing the kernel trick.
 There is a large variety of so-called \acp{gk} to do this.
-\acp{gk} can be used in combination with any kernel method to solve the \ac{gcr} problem.
+\acp{gk} can be used in combination with any kernel method, typically a \ac{svm}, to solve the \ac{gcr} problem.
+While there is a large variety of different \acp{gk}, we will focus on those that are based on the previously described family of \ac{wl} algorithms.
+
+\paragraph{\ac{wl} subtree kernel}
+
+\paragraph{\ac{wl} shortest path kernel}
+
+\paragraph{Higher dimensional \ac{wl} kernels}
 

### 2020-02-11 implementation/run.sh
added balanced_triangle_classification_dataset
more clearly shows the advantage of WL-2 over WL-1 bounded methods

+ added splits for IMDB-BINARY

@@ -33,7 +33,7 @@ if [ "$2" == "rebuild" ]; then
 fi
 
 if [ "$VARIANT" == "tensorflow" ]; then
-	ARGS="-p 6006:6006 -u $(id -u):$(id -g)"
+	ARGS="-p 6006:6006 -u $(id -u):$(id -g) -e TF_FORCE_GPU_ALLOW_GROWTH=$TF_FORCE_GPU_ALLOW_GROWTH"
 fi
 
 trap 'kill %1; exit 0' SIGINT

### 2020-02-11 implementation/src/ltag/datasets/synthetic/datasets.py

@@ -237,3 +237,26 @@ def noisy_triangle_classification_dataset(seed=1337):
         fy.partial(noisy_triangle_graph, *config), repeat)])
 
     return graphs, np.array(ys)
+
+@synthetic_dataset
+def balanced_triangle_classification_dataset(seed=1337):
+  with local_seed(seed):
+    sizes = range(3, 10)
+    balances = [(1, 1), (1, 3), (3, 1)]
+    dens = [0.25, 0.5]
+    classes = [0, 1]
+
+    max_verts = 33
+    repeat = 3
+    configs = [
+      (s * l, s * r, d, y)
+      for s, (l, r), d, y in cart(sizes, balances, dens, classes)
+      if s * (l + r) <= max_verts]
+
+    graphs, ys = unzip([
+      entry
+      for config in configs
+      for entry in fy.repeatedly(
+        fy.partial(noisy_triangle_graph, *config), repeat)])
+
+    return graphs, np.array(ys)

### 2020-02-11 implementation/src/ltag/evaluate_datasets.py

@@ -78,7 +78,7 @@ if __name__ == "__main__":
     else:
       ms = [
         "WL_st", "WL_sp", "LWL2", "GWL2",
-        "AvgCWL2GCN_Binary"]
+        "AvgCWL2GCN_Binary", "SagCWL2GCN_Binary"]
   else:
     ms = args.model
 

### 2020-02-11 implementation/src/ltag/evaluation/datasets.py

@@ -93,8 +93,26 @@ noisy_triangle_classification_2 = fy.partial(
     }
   })
 
+balanced_triangle_classification_2 = fy.partial(
+  syn.balanced_triangle_classification_dataset(stored=True),
+  wl2_neighborhood=2,
+  wl2_batch_size={
+    "batch_graph_count": 228
+  },
+  evaluation_args={
+    "epochs": 5000,
+    "patience": 1000,
+    "hp_args": {
+      # other widths ignored based on previous experiments:
+      "cwl2_layer_widths": [32],
+      "cwl2_layer_depths": [3],
+      "cwl2_local_act": "relu"
+    }
+  })
+
 synthetic_binary = [
-  "noisy_triangle_classification_2"
+  "noisy_triangle_classification_2",
+  "balanced_triangle_classification_2"
 ]
 synthetic = synthetic_binary
 stored_synthetic = synthetic

### 2020-02-11 implementation/src/ltag/playground.py

@@ -192,10 +192,10 @@ def wl2_power_experiment():
 
 def synthetic_experiment2():
   model_class = gnn_models.SagCWL2GCN
-  dsm = synthetic.noisy_triangle_classification_dataset(stored=True)(
+  dsm = synthetic.balanced_triangle_classification_dataset(stored=True)(
     with_holdout=False,
     wl2_neighborhood=2,
-    wl2_batch_size=dict(batch_graph_count=208))
+    wl2_batch_size=dict(batch_graph_count=228))
 
   if model_class.input_type == "wl2c":
     in_dim = dsm.dim_wl2_features()
@@ -205,19 +205,13 @@ def synthetic_experiment2():
   if in_dim == 0:
     in_dim = 1
 
-  opt = keras.optimizers.Adam(0.0003)
-  # reg = keras.regularizers.l1(0.00001)
+  opt = keras.optimizers.Adam(0.0005)
 
   model = model_class(
     act="sigmoid", local_act="relu",
     squeeze_output=True,
     layer_dims=[in_dim, 32, 32, 32, 1],
-    att_conv_layer_dims=[in_dim, 32, 32, 1],
-    W_regularizer=None,
-    W_prop_regularizer=None,
-    W_back_regularizer=None,
-    b_regularizer=None,
-    b_prop_regularizer=None,
+    att_conv_layer_dims=[in_dim, 32, 32, 32, 1],
     bias=True)
 
   model.compile(
@@ -225,7 +219,7 @@ def synthetic_experiment2():
     loss="binary_crossentropy",
     metrics=["accuracy"])
 
-  i = 9
+  i = 5
   ds = dsm.get_train_fold(
     i, output_type=model_class.input_type)
   ds_test = dsm.get_test_fold(
@@ -233,14 +227,14 @@ def synthetic_experiment2():
 
   evaluate.train(
     model, ds, ds_test, verbose=2,
-    epochs=2000, patience=2000,
+    epochs=5000, patience=2000,
     label=f"{dsm.name}_{model.name}")
 
 
 def kernel_experiment():
-  model_class = kernel_models.LWL2
+  model_class = kernel_models.WL_sp
   model = model_class(C=0.001)
-  dsm = synthetic.noisy_triangle_classification_dataset(stored=True)(
+  dsm = synthetic.balanced_triangle_classification_dataset(stored=True)(
     with_holdout=False)
 
   for i in range(10):
@@ -250,11 +244,11 @@ def kernel_experiment():
[…]

### 2020-02-12 implementation/src/ltag/evaluation/datasets.py
replacing noisy triangle class dataset with the balanced variant
since it shows the advantages of WL2 more clearly

@@ -113,7 +113,6 @@ balanced_triangle_classification_2 = fy.partial(
   })
 
 synthetic_binary = [
-  "noisy_triangle_classification_2",
   "balanced_triangle_classification_2"
 ]
 synthetic = synthetic_binary

### 2020-02-13 implementation/src/ltag/models/gnn.py
minor fixes

@@ -121,7 +121,7 @@ AvgDenseWL2GCN = DenseWL2GCN.extend("AvgDenseWL2GCN", [
 
 AvgWL2GCN = WL2GCN.extend("AvgWL2GCN", [
   cm.with_layer(lwl2.AvgPooling, with_inputs=True)])
-AvgCWL2GCN = CWL2GCN.extend("AvgWCL2GCN", [
+AvgCWL2GCN = CWL2GCN.extend("AvgCWL2GCN", [
   cm.with_layer(lwl2.AvgPooling, with_inputs=True)])
 
 # Max GNNs:

### 2020-02-13 thesis/content/chapter-introduction.tex
started with wl section

--- 2019-11-09 thesis/content/chapter-introduction.tex
@@
 The field of \ac{ml} on graph-structured data has applications in many domains due to the general expressive power of graphs.
-Three common types of graph~\ac{ml} problems are
+The three most common types of graph~\ac{ml} problems are
 \begin{enumerate}[label=\textbf{\arabic*.}]
@@
 		A graph with an incomplete edge set is given and the missing edges have to be predicted.
-		The generation of friendship suggestions in a social network is a typical example for this.
+		The suggestion of potential friends in a social network is a typical example for this.
 	\item \textbf{Vertex classification \& regression:}
 		Here a class or a score has to be predicted for each vertex of a graph.
-		In social graphs this corresponds to the prediction of properties of individuals.
+		In social graphs this corresponds to the prediction of properties of individuals, e.g.\ personal preferences or gender.
 		Another example is the prediction of the amount of traffic at the intersections of a street network.
@@
 \end{enumerate}
-In this thesis we will focus on the last problem type, \ac{gcr}.
-A \ac{ml} method for this problem has to accept graphs of varying size and should be permutation invariant wrt.\ the vertices.
+In this thesis we will focus on the last problem type, \acf{gcr}.
+An \ac{ml} method for this problem has to accept graphs of varying size and should be permutation invariant wrt.\ the order in which graph vertices are provided.
 Those requirements are not met by the commonly used learners that only accept fixed-size feature vectors as their input, e.g.\ \acp{lm}, \acp{svm} or \acp{mlp}.

### 2020-02-13 thesis/content/chapter-related-work.tex
started with wl section

@@
 In this simple model the \ac{lta} aggregation problem boils down to the optimization of $\lambda$.
-The \ac{lta} disaggregation problem is solved by jointly optimizing a \ac{lm}, i.e.\ the constituent scores ${\{ y_i \in [0, 1] \}}_{c_i \in \bm{c}}$ are described by $y_i = {\left( 1 + \exp(-\theta^\top x_i) \right)}^{-1}$.
+The \ac{lta} disaggregation problem is solved by jointly optimizing a \acf{lm}, i.e.\ the constituent scores ${\{ y_i \in [0, 1] \}}_{c_i \in \bm{c}}$ are described by $y_i = {\left( 1 + \exp(-\theta^\top x_i) \right)}^{-1}$.
 Overall an \ac{lta} model is therefore described by the uninorm parameter $\lambda$ and the regression coefficients $\theta$.
@@
 
-\subsection{Weisfeiler-Lehman Graph Colorings}%
+\subsection{\acl{wl} Graph Colorings}%
 \label{sec:related:character:wl}
 
-The Weisfeiler-Lehman algorithm characterizes a graph $G = (\mathcal{V}, \mathcal{E})$ by assigning discrete labels $c \in \mathcal{C}$, called \textit{colors}, to vertex $k$-tuples $(v_1, \dots, v_k) \in \mathcal{V}^k$, where $k \in \mathbb{N}$ is the freely choosable \textit{\ac{wl}-dimension}.
-A mapping $\chi: \mathcal{V}^k \to \mathcal{C}$ is called a coloring of $G = (\mathcal{V}, \mathcal{E})$.
+The \acf{wl} algorithm~\cite{Weisfeiler1968} characterizes a graph $G = (\mathcal{V}, \mathcal{E})$ by assigning discrete labels $c \in \mathcal{C}$, called \textit{colors}, to vertex $k$-tuples $(v_1, \dots, v_k) \in \mathcal{V}^k$, where $k \in \mathbb{N}$ is the freely choosable \textit{\ac{wl}-dimension}.
+A mapping $\chi_{G, k}: \mathcal{V}^k \to \mathcal{C}$ is called a \textit{$k$-coloring} of $G = (\mathcal{V}, \mathcal{E})$.
+We say  $\chi'$ \textit{refines} $\chi$ ($\chi' \succeq \chi$) iff.\ $\forall a, b \in \mathcal{V}^k: \chi(a) \neq \chi(b) \rightarrow \chi'(a) \neq \chi'(b)$, i.e.\ $\chi'$ distinguishes at least those tuples that are distinguished by $\chi$.
 
-\paragraph{The 1-dimensional \acs{wl} algorithm}
-For the most simple case of $k = 1$, a color is assigned to each vertex of a graph.
+The $k$-dimensional \ac{wl} algorithm works by iteratively refining $k$-colorings $\chi_{G, k}^0 \preceq \chi_{G, k}^1 \preceq \dots$ of a given graph $G$ until the convergence criterion $\chi_{G, k}^i \equiv \chi_{G, k}^{i+1}$ is satisfied;
+the criterion denotes that $\chi_{G, k}^i$ and $\chi_{G, k}^{i+1}$ must be identical up to color/label substitutions.
+The final \ac{wl} coloring $\chi_{G, k}$ that is obtained after convergence can be used to quickly check whether two graphs are not isomorphic.
+If for two graphs $G$ and $H$ there exists a color $c \in \mathcal{C}$ s.t.\ $\left|\left\{ a \in \mathcal{V}_G^k\, |\, \chi_{G, k}(a) = c \right\}\right|\neq \left|\left\{ b \in \mathcal{V}_H^k\, |\, \chi_{H, k}(b) = c \right\}\right|$, the graphs are $k$-\acs{wl} \textit{distinguishable} ($G \mathrel{{\not\simeq}_k} H$) and therefore non-isomorphic ($G \not\simeq H$).
+% chktex 21
+The opposite does however not necessarily hold;
+two $k$-\acs{wl} indistinguishable graphs are not always isomorphic, i.e.\ $G \mathrel{{\simeq}_k} H \centernot\implies G \simeq H$.
+To understand which types of graphs are $k$-\acs{wl} distinguishable, we will proceed in three steps:
+\begin{enumerate}
+	\item First the color refinement algorithm for the most simple case of $k = 1$ is described.
+	\item Then we extend the definitions and intuitions from the 1-dimensional \ac{wl} algorithm to its higher-dimensional generalization.
+	\item Lastly we will discuss the discriminative power of the \acs{wl} algorithm and its relation to the \acs{wl}-dimension $k$.
+\end{enumerate}
+
+\subsubsection{The 1-dimensional \acs{wl} algorithm}
+In the 1-dimensional \ac{wl} algorithm, a color is assigned to each vertex of a graph.
+If the vertices $\mathcal{V}_G$ of the input graph $G$ are already labeled via some labeling $l_G: \mathcal{V}_G \to L$, those labels can be used as the initial graph coloring $\chi_{G,1}^0 = l_G$.
+For unlabeled graphs a constant coloring is used instead, i.e.\ $\forall v \in \mathcal{V}_G: \chi_{G,1}^0(v) = \texttt{A}$ for some initial color $\texttt{A} \in \mathcal{C}$.
+In each iteration of the 1-\acs{wl} color refinement algorithm, the following neighborhood aggregation scheme is used to compute a new color for all vertices:
+\begin{align}
+	\chi_{G,1}^{i+1}(v) := h\left(\chi_{G,1}^{i}(v), \{ \chi_{G,1}^{i}(u) | u \in \Gamma_{G,1}(v) \}\right)
+\end{align}
 

### 2020-02-14 thesis/content/chapter-related-work.tex
minor additions to wl section
currently still working on WL illustrations...

@@
 The class of \ac{lta} problems was first described by \citet{Melnikov2016}.
-There an input instance is understood as a composition $\bm{c}$ of so-called constituents $c_i \in \bm{c}$, i.e.\ as a variable-size multiset with $n = |\bm{c}|$.
-The assumption in \ac{lta} problems is that for all constituents $c_i$ a local score $y_i \in \mathcal{Y}$ is either given or computable.
+There an input instance is understood as a composition $\bm{c} = \ldblbrace c_1, \dots, c_n \rdblbrace$ of so-called constituents, i.e.\ as a variable-size multiset (denoted as $\ldblbrace \cdot \rdblbrace$).
+The assumption in \ac{lta} problems is that for all constituents $c_i \in \bm{c}$ a local score $y_i \in \mathcal{Y}$ is either given or computable.
 The set of those local scores should be indicative of the overall score $y \in \mathcal{Y}$ of the composition $\bm{c}$.
@@
 		In case the constituent scores $y_i$ are not given, they have to be derived from a constituent representation, e.g.\ a vector $x_i \in \mathcal{X}$.
-		To learn this derivation function $f: \mathcal{X} \to \mathcal{Y}$, only the constituent vectors ${\{x_i\}}_{i = 1}^{n}$ and the composite score $y$ is given.
+		To learn this derivation function $f: \mathcal{X} \to \mathcal{Y}$, only the constituent vectors ${\ldblbrace x_i \rdblbrace}_{i = 1}^{n}$ and the composite score $y$ is given.
 		Thus the constituent scores $y_i$ need to be \textit{disaggregated} from $y$ in order to learn $f$.
@@
 Since t-norms and t-conorms are commutative and associative they can also be applied to non-empty sets of arbitrary size, i.e.
-$T(\{ y_1, \dots, y_n \}) = T(y_1, T(\{ y_2, \dots, y_n \}))$ with fixpoint $T(\{ y \}) = y$.
+$T(\ldblbrace y_1, \dots, y_n \rdblbrace) = T(y_1, T(\ldblbrace y_2, \dots, y_n \rdblbrace))$ with fixpoint $T(\ldblbrace y \rdblbrace) = y$.
 Using this extension, a uninorm $U_\lambda$ can be applied to sets which turns it into a parameterized aggregation function $\mathcal{A}_\lambda: {[0, 1]}^* \to [0, 1]$.
 In this simple model the \ac{lta} aggregation problem boils down to the optimization of $\lambda$.
-The \ac{lta} disaggregation problem is solved by jointly optimizing a \acf{lm}, i.e.\ the constituent scores ${\{ y_i \in [0, 1] \}}_{c_i \in \bm{c}}$ are described by $y_i = {\left( 1 + \exp(-\theta^\top x_i) \right)}^{-1}$.
+The \ac{lta} disaggregation problem is solved by jointly optimizing a \acf{lm}, i.e.\ the constituent scores ${\ldblbrace y_i \in [0, 1] \rdblbrace}_{c_i \in \bm{c}}$ are described by $y_i = {\left( 1 + \exp(-\theta^\top x_i) \right)}^{-1}$.
 Overall an \ac{lta} model is therefore described by the uninorm parameter $\lambda$ and the regression coefficients $\theta$.
@@
 
-To deal with varying composition sizes $n$, the weights ${\{ \lambda_i \}}_{i = 1}^{n}$ can however not be statically assigned.
+To deal with varying composition sizes $n$, the weights $\lambda_1, \dots, \lambda_n$ can however not be statically assigned.
 Instead they are interpolated using a so-called \ac{bum} function $q: [0, 1] \to [0, 1]$.
@@
 \begin{align}
-	\chi_{G,1}^{i+1}(v) := h\left(\chi_{G,1}^{i}(v), \{ \chi_{G,1}^{i}(u) | u \in \Gamma_{G,1}(v) \}\right)
+	\chi_{G,1}^{i+1}(v) := h\left(\chi_{G,1}^{i}(v), \ldblbrace \chi_{G,1}^{i}(u)\, |\, u \in \Gamma_{G}(v) \rdblbrace\right) \text{,}
 \end{align}
+with $\Gamma_G(v)$ denoting the set of adjecent vertices of $v \in \mathcal{V}_G$ and $h: \mathcal{C}^* \to \mathcal{C}$ denoting a perfect hash function that assigns a unique color to each finite combination of colors.
+In practice $h$ is usually defined lazily by using $\mathcal{C} = \mathbb{N}$ and non-deterministically enumerating color combinations in the order in which they are hashed s.t.\ a new color is introduced every time a previously unseen color combination is queried.
 

### 2020-02-15 implementation/src/ltag/evaluation/evaluate.py
fixed bug that caused eval debug logs to be written to wrong directory

@@ -55,7 +55,7 @@ def train(
 @cp.tolerant
 def evaluation_step(
   model_ctr, train_ds, val_ds, test_ds, k, hp_i, i, hp,
-  res_dir, fold_str, hp_str, verbose,
+  res_dir, fold_str, hp_str, verbose, log_dir_base,
   epochs, patience, stopping_min_delta, restore_best,
   repeat, winner_repeat, pos_hp_i=None):
   if i >= repeat:
@@ -247,7 +247,7 @@ def evaluate(
         for i in range(curr_i_start, repeat):
           completed_evaluation_step |= evaluation_step(
             model_ctr, train_ds, val_ds, test_ds, k, hp_i, i, hp,
-            res_dir, fold_str, hp_str, verbose,
+            res_dir, fold_str, hp_str, verbose, log_dir_base,
             **config)
           pos_file.write_text(f"{k},{hp_i},{i}")
 
@@ -274,7 +274,7 @@ def evaluate(
           for i in range(repeat, winner_repeat):
             completed_evaluation_step = evaluation_step(
               model_ctr, train_ds, val_ds, test_ds, k, best_hp_i, i, best_hp,
-              res_dir, fold_str, hp_str, verbose,
+              res_dir, fold_str, hp_str, verbose, log_dir_base,
               **config)
             pos_file.write_text(f"{k},{hpc},{i}")
           print(

### 2020-02-15 thesis/content/chapter-related-work.tex
continued wl section

@@
 To classify or score a graph, it first needs to be characterized by a set of relevant properties.
-Two commonly used graph characterization approaches are the so called \acf{wl} colors and the notion of the graph spectrum.
+Two commonly used graph characterization approaches are the so called \acl{wl} coloring and the notion of the graph spectrum.
 Both will now be briefly introduced since they form the foundation of most current \ac{gcr} methods.
@@
 
-The $k$-dimensional \ac{wl} algorithm works by iteratively refining $k$-colorings $\chi_{G, k}^0 \preceq \chi_{G, k}^1 \preceq \dots$ of a given graph $G$ until the convergence criterion $\chi_{G, k}^i \equiv \chi_{G, k}^{i+1}$ is satisfied;
-the criterion denotes that $\chi_{G, k}^i$ and $\chi_{G, k}^{i+1}$ must be identical up to color/label substitutions.
-The final \ac{wl} coloring $\chi_{G, k}$ that is obtained after convergence can be used to quickly check whether two graphs are not isomorphic.
-If for two graphs $G$ and $H$ there exists a color $c \in \mathcal{C}$ s.t.\ $\left|\left\{ a \in \mathcal{V}_G^k\, |\, \chi_{G, k}(a) = c \right\}\right|\neq \left|\left\{ b \in \mathcal{V}_H^k\, |\, \chi_{H, k}(b) = c \right\}\right|$, the graphs are $k$-\acs{wl} \textit{distinguishable} ($G \mathrel{{\not\simeq}_k} H$) and therefore non-isomorphic ($G \not\simeq H$).
+The $k$-dimensional \ac{wl} algorithm ($k$-\acs{wl}) works by iteratively refining $k$-colorings $\chi_{G, k}^0 \preceq \chi_{G, k}^1 \preceq \dots$ of a given graph $G$ until the convergence criterion $\chi_{G, k}^i \equiv \chi_{G, k}^{i+1}$ is satisfied;
+here the equivalence operator $\equiv$ denotes that $\chi_{G, k}^i$ and $\chi_{G, k}^{i+1}$ must be identical up to color/label substitutions.
+The final \ac{wl} coloring $\bar{\chi}_{G, k}$ that is obtained after convergence can be used to quickly check whether two graphs are not isomorphic.
+If for two graphs $G$ and $H$ there exists a color $c \in \mathcal{C}$ s.t.\ $\left|\left\{ a \in \mathcal{V}_G^k\, |\, \bar{\chi}_{G, k}(a) = c \right\}\right|\neq \left|\left\{ b \in \mathcal{V}_H^k\, |\, \bar{\chi}_{H, k}(b) = c \right\}\right|$, the graphs are $k$-\acs{wl} \textit{distinguishable} ($G \mathrel{{\not\simeq}_k} H$) and therefore non-isomorphic ($G \not\simeq H$).
 % chktex 21
@@
 two $k$-\acs{wl} indistinguishable graphs are not always isomorphic, i.e.\ $G \mathrel{{\simeq}_k} H \centernot\implies G \simeq H$.
-To understand which types of graphs are $k$-\acs{wl} distinguishable, we will proceed in three steps:
-\begin{enumerate}
-	\item First the color refinement algorithm for the most simple case of $k = 1$ is described.
-	\item Then we extend the definitions and intuitions from the 1-dimensional \ac{wl} algorithm to its higher-dimensional generalization.
-	\item Lastly we will discuss the discriminative power of the \acs{wl} algorithm and its relation to the \acs{wl}-dimension $k$.
-\end{enumerate}
+To understand why this is the case and which types of graphs are $k$-\acs{wl} distinguishable, we will now look at the details of \ac{wl} color refinement step.
+First the color refinement algorithm for the most simple case of $k = 1$ is described.
+Then the definitions and intuitions from the 1-dimensional case are extended to its higher-dimensional generalization.
+Lastly we will discuss the discriminative power of the \acs{wl} algorithm and its relation to the \acs{wl}-dimension $k$.
 
@@
 If the vertices $\mathcal{V}_G$ of the input graph $G$ are already labeled via some labeling $l_G: \mathcal{V}_G \to L$, those labels can be used as the initial graph coloring $\chi_{G,1}^0 = l_G$.
-For unlabeled graphs a constant coloring is used instead, i.e.\ $\forall v \in \mathcal{V}_G: \chi_{G,1}^0(v) = \texttt{A}$ for some initial color $\texttt{A} \in \mathcal{C}$.
+For unlabeled graphs a constant coloring is used instead, e.g.\ $\forall v \in \mathcal{V}_G: \chi_{G,1}^0(v) = \textcolor{t_blue}{\texttt{A}}$ for some initial color $\textcolor{t_blue}{\texttt{A}} \in \mathcal{C}$.
+% chktex 25
 In each iteration of the 1-\acs{wl} color refinement algorithm, the following neighborhood aggregation scheme is used to compute a new color for all vertices:
 \begin{align}
-	\chi_{G,1}^{i+1}(v) := h\left(\chi_{G,1}^{i}(v), \ldblbrace \chi_{G,1}^{i}(u)\, |\, u \in \Gamma_{G}(v) \rdblbrace\right) \text{,}
+	\chi_{G,1}^{i+1}(v) := h\left(\chi_{G,1}^{i}(v), \ldblbrace \chi_{G,1}^{i}(u)\, |\, u \in \Gamma_{G}(v) \rdblbrace\right) \text{,}\label{eq:related:wl1-refine}
 \end{align}
@@
 In practice $h$ is usually defined lazily by using $\mathcal{C} = \mathbb{N}$ and non-deterministically enumerating color combinations in the order in which they are hashed s.t.\ a new color is introduced every time a previously unseen color combination is queried.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.75\linewidth]{gfx/related-work/wl1-refine.pdf}
+	\caption[Example 1-WL color refinement steps.]{
+		Example 1-WL color refinement steps.
+		After two iterations the coloring stabilizes.
+		Each vertex is labeled with its current color and has the aggregated neighbors, as defined in \cref{eq:related:wl1-refine}, written next to it.
+	}\label{fig:related:wl1-refine}
+\end{figure}
+
+\subsubsection{The $k$-dimensional \acs{wl} algorithm}
+As we just saw, the 1-\acs{wl} algorithm iteratively refines colorings of single vertices.
+While the obtained colorings differ for most non-isomorphic graphs $G \not\simeq H$, 1-\acs{wl} does not generally solve the graph isomorphism problem as illustrated in \cref{fig:related:wl1-problem}.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=\linewidth]{gfx/related-work/wl1-problem.pdf}
+	\caption[Two simple non-isomorphic graphs that are indistinguishable via 1-\acs{wl}.]{
+		Two simple non-isomorphic graphs that are indistinguishable via 1-\acs{wl}.
+	}\label{fig:related:wl1-problem}
+\end{figure}
+By extending \ac{wl} to higher dimensions such 1-\acs{wl} indistinguishable cases can however be handled.
+Analogous to the 1-dimensional case, the $k$-dimensional color refinement step is defined by
+\begin{align}
+	\chi_{G,k}^{i+1}(v) := &\ h\left(\chi_{G,k}^{i}(v), \ldblbrace (\chi_{G,k}^{i}(v[u/1]), \dots, \chi_{G,k}^{i}(v[u/k]))\, |\, u \in \mathcal{V}_G \rdblbrace\right)\label{eq:related:wl1-refine}\\
+	\text{with } v = &\ (v_1, \dots, v_k) \in \mathcal{V}^k \quad\text{and}\quad v[u/j] := (v_1, \dots, v_{j-1}, u, v_{j+1}, \dots, v_k)\text{.} \nonumber
+\end{align}
+In 1-\acs{wl} a vertex color is refined by combining the colors of neighboring vertices.
+In $k$-\acs{wl} the color of a vertex $k$-tuple $v$ is refined by combining the colors of its neighborhood which is defined as the set of all $k$-tuples in which at most one vertex differs from $v$.
+
+\subsubsection{Discriminative Power of \acs{wl}}
 

### 2020-02-16 thesis/content/chapter-related-work.tex
continued wl section

@@
 \end{enumerate}
-\begin{figure}
+\begin{figure}[ht]
 	\centering
@@
 $U_\lambda$ therefore smoothly interpolates between a conjunctive and disjunctive operator with the extreme points $U_1 = T$ and $U_0 = S$.
-\begin{figure}[h]
+\begin{figure}[ht]
 	\centering
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
 	\centering
@@
 In practice $h$ is usually defined lazily by using $\mathcal{C} = \mathbb{N}$ and non-deterministically enumerating color combinations in the order in which they are hashed s.t.\ a new color is introduced every time a previously unseen color combination is queried.
-\begin{figure}[h]
+\begin{figure}[ht]
 	\centering
@@
 While the obtained colorings differ for most non-isomorphic graphs $G \not\simeq H$, 1-\acs{wl} does not generally solve the graph isomorphism problem as illustrated in \cref{fig:related:wl1-problem}.
-\begin{figure}[h]
+\begin{figure}[ht]
 	\centering
 	\includegraphics[width=\linewidth]{gfx/related-work/wl1-problem.pdf}
-	\caption[Two simple non-isomorphic graphs that are indistinguishable via 1-\acs{wl}.]{
-		Two simple non-isomorphic graphs that are indistinguishable via 1-\acs{wl}.
+	\caption[Two simple non-isomorphic graphs that are indistinguishable by 1-\acs{wl}.]{
+		Two simple non-isomorphic graphs that are indistinguishable by 1-\acs{wl}.
 	}\label{fig:related:wl1-problem}
@@
 By extending \ac{wl} to higher dimensions such 1-\acs{wl} indistinguishable cases can however be handled.
-Analogous to the 1-dimensional case, the $k$-dimensional color refinement step is defined by
+Analogous to the 1-dimensional definition from \cref{eq:related:wl1-refine}, the $k$-dimensional color refinement step is defined by
 \begin{align}
-	\chi_{G,k}^{i+1}(v) := &\ h\left(\chi_{G,k}^{i}(v), \ldblbrace (\chi_{G,k}^{i}(v[u/1]), \dots, \chi_{G,k}^{i}(v[u/k]))\, |\, u \in \mathcal{V}_G \rdblbrace\right)\label{eq:related:wl1-refine}\\
+	\chi_{G,k}^{i+1}(v) := &\ h\left(\chi_{G,k}^{i}(v), \ldblbrace (\chi_{G,k}^{i}(v[u/1]), \dots, \chi_{G,k}^{i}(v[u/k]))\, |\, u \in \mathcal{V}_G \rdblbrace\right)\label{eq:related:wlk-refine}\\
 	\text{with } v = &\ (v_1, \dots, v_k) \in \mathcal{V}^k \quad\text{and}\quad v[u/j] := (v_1, \dots, v_{j-1}, u, v_{j+1}, \dots, v_k)\text{.} \nonumber
@@
 In 1-\acs{wl} a vertex color is refined by combining the colors of neighboring vertices.
-In $k$-\acs{wl} the color of a vertex $k$-tuple $v$ is refined by combining the colors of its neighborhood which is defined as the set of all $k$-tuples in which at most one vertex differs from $v$.
+In $k$-\acs{wl} the color of a $k$-tuple $v \in \mathcal{V}^k$ is refined by combining the colors of its neighborhood which is defined as the set of all $k$-tuples in which at most one vertex differs from $v$.
+Note that each vertex $k$-tuple has one neighbor for each $u \in \mathcal{V}_G$, each of which is a $k$-tuple of vertex $k$-tuples.
+This more abstract notion of neighborhood is illustrated in \cref{fig:related:wl-neighbors}.
+\begin{figure}[ht]
+	\centering
+	\begin{subfigure}{0.33\textwidth}
+		\centering
+		\includegraphics[width=0.8\linewidth]{gfx/related-work/wl1-neighbors.pdf}
+		\subcaption{1-WL $v_1$ neighbors}\label{fig:related:wl-neighbors:1}
+	\end{subfigure}%
+	\begin{subfigure}{0.33\textwidth}
+		\centering
+		\includegraphics[width=0.8\linewidth]{gfx/related-work/wl2-neighbors.pdf}
+		\subcaption{2-WL $(v_1, v_2)$ neighbors}\label{fig:related:wl-neighbors:2}
+	\end{subfigure}%
+	\begin{subfigure}{0.33\textwidth}
+		\centering
+		\includegraphics[width=0.8\linewidth]{gfx/related-work/wl3-neighbors.pdf}
+		\subcaption{3-WL $(v_1, v_2, v_3)$ neighbors}\label{fig:related:wl-neighbors:3}
+	\end{subfigure}
+	\caption[\ac{wl} neighborhoods for different values of $k$.]{
+		Tuple neighborhoods for different values of $k$.
+		The vertices highlighted in \textcolor{t_blue}{blue} form the root tuple whose neighbors are shown.
+		Each neighbor is highlighted with a different color, except for 3-WL where the \textcolor{t_red}{red}, \textcolor{t_green}{green} and \textcolor{t_darkyellow}{yellow} triples actually form the single neighbor for $u = v_4$ (see~\cref{eq:related:wlk-refine}).
+	}\label{fig:related:wl-neighbors}
+\end{figure}
+For $k = 2$ this means that each potential edge $(v, w) \in \mathcal{V}_G^2$ has all possible paths of length $2$ from $v$ to $w$ as its neighbors (see~\cref{fig:related:wl-neighbors:2}).
+Also note that, even though $k$-\acs{wl} refines $k$-tuple colors, lower-dimensional structures still get their own colors since a tuple does not have to consist of distinct vertices, i.e.\ in $k$-\acs{wl} the color of a single vertex $v \in \mathcal{V}_G$ is described by $\bar{\chi}_{G, k}(s)$ for $s = (v, \dots, v) \in \mathcal{V}_G^k$.
+
+Let us now look at how the tuple colors are initialized.
+For this we use subgraph \textit{isomorphism types} which determine the initial color $\chi_{G,k}^0(v)$ of each $k$-tuple $v$.
+For $k = 1$ the isomorphism type of a vertex $v$ directly corresponds to its label $l_G(v)$.
+For $k > 1$ the isomorphism types more generally correspond to the classes of isomorphic ordered subgraphs of $G$ induced by the $k$-tuples $\mathcal{V}_G^k$.
+Formally this means that
+\begin{align}
+	\chi_{G,k}^0(v_1, \dots, v_k) = \chi_{G,k}^0(w_1, \dots, w_k) \label{eq:related:wlk-init}
+	\iff &\ \forall i: l_G(v_i) = l_G(w_i) \\
+	\land &\ \forall i, j: v_i = v_j \leftrightarrow w_i = w_j \nonumber \\
+	\land &\ \forall i, j: (v_i, v_j) \in \mathcal{E}_G \leftrightarrow (w_i, w_j) \in \mathcal{E}_G \text{.} \nonumber
+\end{align}
+If no explicit vertex labeling $l_G$ is given, a single constant label for all vertices is assumed.
+Also note that there is a fundamental difference in how the adjacency information encoded in $\mathcal{E}_G$ is used in 1-\acs{wl} vs.\ $k$-\acs{wl}:
+In 1-\acs{wl} a vertex coloring by itself cannot encode adjacency which is why this information is explicitly incorporated in each refinement step via $\Gamma_G$ (see~\cref{eq:related:wl1-refine}).
+In $k$-\acs{wl} on the other hand each pair of vertices $(v, u) \in \mathcal{V}_G^2$ appears in at least one $k$-tuple (assuming $k \geq 2$) and therefore has at least one color which can implicitly encode the adjacency information.
+Edges and non-edges are colored differently in the initial coloring as defined in \cref{eq:related:wlk-init};
+thus no explicit adjacency information is needed in the $k$-\acs{wl} color refinement step defined in \cref{eq:related:wlk-refine}.
 
 \subsubsection{Discriminative Power of \acs{wl}}
+Now we will look at the types of graphs that can be distinguished by \acs{wl} in relation to the \acs{wl}-dimension $k$.
+\begin{thm}
+	$G \mathrel{\not\simeq_k} H \implies G \mathrel{\not\simeq_{k+1}} H$, i.e.\ the discriminative power of $k$-\acs{wl} grows with $k$.
+	\begin{hproof}
+		All $k$-tuples can be mapped to $(k+1)$-tuples, e.g.\ via $\varphi(v_1, \dots, v_k) = (v_1, \dots, v_k, v_k)$.
+	\end{hproof}
+\end{thm}
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=\linewidth]{gfx/related-work/wl2-problem-solution.pdf}
+	\caption[Two simple non-isomorphic graphs that are indistinguishable by 1-\acs{wl}.]{
+		Two simple non-isomorphic graphs that are indistinguishable by 1-\acs{wl}.
+	}\label{fig:related:wl2-problem-solution}
+\end{figure}
 

### 2020-02-17 implementation/run.sh
various implementation improvements and fixes

@@ -55,7 +55,7 @@ mkdir -p evaluations
 mkdir -p libs
 mkdir -p data
 
-docker run --runtime=nvidia --rm --name $LTAG_CONTAINER_NAME \
+docker run --gpus all --rm --name $LTAG_CONTAINER_NAME \
 	-p $JUPYTER_PORT:8888 \
 	-v $(pwd)/src:/ltag \
 	-v $(pwd)/logs:/logs \

### 2020-02-17 implementation/src/ltag/evaluate_datasets.py

@@ -33,6 +33,8 @@ def resume(mf, dsm, **kwargs):
 def summarize(mf, dsm):
   if dsm.evaluation_args is not None:
     kwargs = fy.project(dsm.evaluation_args, ["ignore_worst"])
+  else:
+    kwargs = {}
 
   return summary.summarize_evaluation(
     evaluate.find_eval_dir(mf, dsm), **kwargs)
@@ -90,7 +92,7 @@ if __name__ == "__main__":
   msl = len(ms)
 
   print(f"Starting {type}...")
-  print(f"Will evaluate the following {dsl} datasets:")
+  print(f"Will use the following {dsl} datasets:")
   for d in ds:
     print(f"- {d}")
   print(f"Will use the following {msl} models:")

### 2020-02-17 implementation/src/ltag/evaluation/evaluate.py

@@ -101,6 +101,9 @@ def evaluation_step(
       "test_duration": test_dur
     }, f, cls=NumpyEncoder)
 
+  tf.keras.backend.clear_session()
+  gc.collect()
+
   print(
     f"\nTest results in {train_dur}s/{test_dur}s for",
     f"it {rep_str}, fold {fold_str}, hps {hp_str}:",
@@ -281,9 +284,6 @@ def evaluate(
             time_str(),
             f"- Completed additional {add_rep} evals of fold {fold_str}",
             f"and winning hp {hp_str}.")
-
-      tf.keras.backend.clear_session()
-      gc.collect()
   finally:
     t_end_eval = timer()
     dur_eval = t_end_eval - t_start_eval

### 2020-02-17 implementation/src/ltag/evaluation/summary.py

@@ -3,7 +3,7 @@ from __future__ import absolute_import, division, print_function,\
 
 import os
 import json
-from collections import defaultdict
+from collections import defaultdict, Sized
 import funcy as fy
 import numpy as np
 
@@ -27,7 +27,7 @@ def statistics(vals):
     "std": np.std(vals),
     "min": np.min(vals),
     "max": np.max(vals),
-    "count": len(vals)
+    "count": len(vals) if isinstance(vals, Sized) else 1
   }
 
 def summarize_evaluation(

### 2020-02-17 thesis/content/chapter-related-work.tex
continued wl section

@@
 \begin{align}
-	\text{t-norm } T(a, b) &:= \max \{ 0, a + b - 1 \}, \quad\text{t-conorm } S(a, b) := \min \{ a + b, 1 \}, \nonumber \\
-	\text{uninorm } U_\lambda(a, b) &:= \begin{cases}
+	\text{t-norm } T(a, b) &\coloneqq \max \{ 0, a + b - 1 \}, \quad\text{t-conorm } S(a, b) \coloneqq \min \{ a + b, 1 \}, \nonumber \\
+	\text{uninorm } U_\lambda(a, b) &\coloneqq \begin{cases}
 		\lambda T\left(\frac{a}{\lambda}, \frac{b}{\lambda}\right) & \text{if } a, b \in [0, \lambda] \\
@@
 \begin{align}
-	\mathcal{A}_{\lambda}(y_1, \dots, y_n) := \sum_{i = 1}^n \lambda_i y_{\pi(i)},
+	\mathcal{A}_{\lambda}(y_1, \dots, y_n) \coloneqq \sum_{i = 1}^n \lambda_i y_{\pi(i)},
 \end{align}
@@
 It takes constituent positions that are normalized to the unit interval, i.e.\ $\frac{i}{n} \in [0, 1]$.
-The \ac{bum} function $q$ is then used to interpolate a weight for any normalized sort position via $\lambda_i := q\left(\frac{i}{n}\right) - q\left(\frac{i - 1}{n}\right)$.
+The \ac{bum} function $q$ is then used to interpolate a weight for any normalized sort position via $\lambda_i \coloneqq q\left(\frac{i}{n}\right) - q\left(\frac{i - 1}{n}\right)$.
 Because $q$ is monotone with $q(0) = 0$ and $q(1) = 1$, it always holds that ${\|\lambda\|}_1 = q(1) - q(0) = 1$. % chktex 21
@@
 here the equivalence operator $\equiv$ denotes that $\chi_{G, k}^i$ and $\chi_{G, k}^{i+1}$ must be identical up to color/label substitutions.
-The final \ac{wl} coloring $\bar{\chi}_{G, k}$ that is obtained after convergence can be used to quickly check whether two graphs are not isomorphic.
-If for two graphs $G$ and $H$ there exists a color $c \in \mathcal{C}$ s.t.\ $\left|\left\{ a \in \mathcal{V}_G^k\, |\, \bar{\chi}_{G, k}(a) = c \right\}\right|\neq \left|\left\{ b \in \mathcal{V}_H^k\, |\, \bar{\chi}_{H, k}(b) = c \right\}\right|$, the graphs are $k$-\acs{wl} \textit{distinguishable} ($G \mathrel{{\not\simeq}_k} H$) and therefore non-isomorphic ($G \not\simeq H$).
+The final \ac{wl} coloring $\bar{\chi}_{G, k}$ that is obtained after convergence can be used to quickly check whether two graphs $G, H$ are not isomorphic ($G \not\simeq H$).
+\begin{defn}\label{defn:related:wl-distinguishable}
+	Two graphs $G$ and $H$ are are $k$-\acs{wl} \textit{distinguishable} ($G \mathrel{{\not\simeq}_k} H$) iff.\ there exists a color $c \in \mathcal{C}$ s.t.\ $\left|\left\{ a \in \mathcal{V}_G^k\, |\, \bar{\chi}_{G, k}(a) = c \right\}\right|\neq \left|\left\{ b \in \mathcal{V}_H^k\, |\, \bar{\chi}_{H, k}(b) = c \right\}\right|$.
 % chktex 21
+\end{defn}
+It is easy to show that $G \mathrel{{\not\simeq}_k} H \implies G \not\simeq H$.
 The opposite does however not necessarily hold;
@@
 In each iteration of the 1-\acs{wl} color refinement algorithm, the following neighborhood aggregation scheme is used to compute a new color for all vertices:
-\begin{align}
-	\chi_{G,1}^{i+1}(v) := h\left(\chi_{G,1}^{i}(v), \ldblbrace \chi_{G,1}^{i}(u)\, |\, u \in \Gamma_{G}(v) \rdblbrace\right) \text{,}\label{eq:related:wl1-refine}
-\end{align}
+\begin{defn}\label{defn:related:wl1-refine}
+	$\chi_{G,1}^{i+1}(v) \coloneqq h\left(\chi_{G,1}^{i}(v), \ldblbrace \chi_{G,1}^{i}(u)\, |\, u \in \Gamma_{G}(v) \rdblbrace\right)$,
 	with $\Gamma_G(v)$ denoting the set of adjecent vertices of $v \in \mathcal{V}_G$ and $h: \mathcal{C}^* \to \mathcal{C}$ denoting a perfect hash function that assigns a unique color to each finite combination of colors.
 	In practice $h$ is usually defined lazily by using $\mathcal{C} = \mathbb{N}$ and non-deterministically enumerating color combinations in the order in which they are hashed s.t.\ a new color is introduced every time a previously unseen color combination is queried.
+\end{defn}
 \begin{figure}[ht]
@@
 		After two iterations the coloring stabilizes.
-		Each vertex is labeled with its current color and has the aggregated neighbors, as defined in \cref{eq:related:wl1-refine}, written next to it.
+		Each vertex is labeled with its current color and has the aggregated neighbors, as in \cref{defn:related:wl1-refine}, written next to it.
 	}\label{fig:related:wl1-refine}
@@
 By extending \ac{wl} to higher dimensions such 1-\acs{wl} indistinguishable cases can however be handled.
-Analogous to the 1-dimensional definition from \cref{eq:related:wl1-refine}, the $k$-dimensional color refinement step is defined by
-\begin{align}
-	\chi_{G,k}^{i+1}(v) := &\ h\left(\chi_{G,k}^{i}(v), \ldblbrace (\chi_{G,k}^{i}(v[u/1]), \dots, \chi_{G,k}^{i}(v[u/k]))\, |\, u \in \mathcal{V}_G \rdblbrace\right)\label{eq:related:wlk-refine}\\
-	\text{with } v = &\ (v_1, \dots, v_k) \in \mathcal{V}^k \quad\text{and}\quad v[u/j] := (v_1, \dots, v_{j-1}, u, v_{j+1}, \dots, v_k)\text{.} \nonumber
-\end{align}
+Analogous to the 1-dimensional \cref{defn:related:wl1-refine}, the $k$-dimensional color refinement step is defined by
+\begin{defn}\label{defn:related:wlk-refine}
+	$\begin{aligned}[t]
+		\chi_{G,k}^{i+1}(v) \coloneqq &\ h\left(\chi_{G,k}^{i}(v), \ldblbrace (\chi_{G,k}^{i}(v[u/1]), \dots, \chi_{G,k}^{i}(v[u/k]))\, |\, u \in \mathcal{V}_G \rdblbrace\right)\\
+		\text{with } v = &\ (v_1, \dots, v_k) \in \mathcal{V}^k, v[u/j] \coloneqq (v_1, \dots, v_{j-1}, u, v_{j+1}, \dots, v_k) \text{.}
+	\end{aligned}$
+\end{defn}
 In 1-\acs{wl} a vertex color is refined by combining the colors of neighboring vertices.
@@
 		Tuple neighborhoods for different values of $k$.
-		The vertices highlighted in \textcolor{t_blue}{blue} form the root tuple whose neighbors are shown.
-		Each neighbor is highlighted with a different color, except for 3-WL where the \textcolor{t_red}{red}, \textcolor{t_green}{green} and \textcolor{t_darkyellow}{yellow} triples actually form the single neighbor for $u = v_4$ (see~\cref{eq:related:wlk-refine}).
+		The vertices highlighted in \textcolor{t_blue}{blue} form the root tuple $v \in \mathcal{V}_G^k$ whose neighbors are shown;
+		for simplicity neighbors with $u \in v$ are left out (see~\cref{defn:related:wlk-refine}).
+		Each neighbor is highlighted with a different color, except for 3-WL where the \textcolor{t_red}{red}, \textcolor{t_green}{green} and \textcolor{t_darkyellow}{yellow} triples actually form the single neighbor for $u = v_4$.
 	}\label{fig:related:wl-neighbors}
@@
 Also note that there is a fundamental difference in how the adjacency information encoded in $\mathcal{E}_G$ is used in 1-\acs{wl} vs.\ $k$-\acs{wl}:
-In 1-\acs{wl} a vertex coloring by itself cannot encode adjacency which is why this information is explicitly incorporated in each refinement step via $\Gamma_G$ (see~\cref{eq:related:wl1-refine}).
+In 1-\acs{wl} a vertex coloring by itself cannot encode adjacency which is why this information is explicitly incorporated in each refinement step via $\Gamma_G$ (see~\cref{defn:related:wl1-refine}).
 In $k$-\acs{wl} on the other hand each pair of vertices $(v, u) \in \mathcal{V}_G^2$ appears in at least one $k$-tuple (assuming $k \geq 2$) and therefore has at least one color which can implicitly encode the adjacency information.
 Edges and non-edges are colored differently in the initial coloring as defined in \cref{eq:related:wlk-init};
-thus no explicit adjacency information is needed in the $k$-\acs{wl} color refinement step defined in \cref{eq:related:wlk-refine}.
+thus no explicit adjacency information is needed in the $k$-\acs{wl} color refinement step in \cref{defn:related:wlk-refine}.
 
@@
 Now we will look at the types of graphs that can be distinguished by \acs{wl} in relation to the \acs{wl}-dimension $k$.
-\begin{thm}
-	$G \mathrel{\not\simeq_k} H \implies G \mathrel{\not\simeq_{k+1}} H$, i.e.\ the discriminative power of $k$-\acs{wl} grows with $k$.
+\begin{lem}
+	$G \mathrel{\not\simeq_k} H \implies G \mathrel{\not\simeq_{k+1}} H$ (the \acs{wl} power is never lower for higher $k$).
+\end{lem}
 \begin{hproof}
-		All $k$-tuples can be mapped to $(k+1)$-tuples, e.g.\ via $\varphi(v_1, \dots, v_k) = (v_1, \dots, v_k, v_k)$.
+	All $k$-tuples can be mapped to $(k+1)$-tuples via $\varphi(v_1, \dots, v_k) \coloneqq (v_1, \dots,\allowbreak v_k, v_k)$.
+	For each neighbor $(v[u/1], \dots,\allowbreak v[u/k])$ of $v$, there is a corresponding neighbor $(\varphi(v[u/1]), \dots \varphi(v[u/k]), \varphi(v[u/k]))$ of $\varphi(v)$.
+	Using \cref{eq:related:wlk-init} (for $i = 0$) and \cref{defn:related:wlk-refine} (for $i > 0$) it follows that $\forall v, w \in \mathcal{V}_G^k: \chi_{G, k}^i(v) \neq \chi_{G, k}^i(w) \rightarrow \chi_{G, k+1}^i(\varphi(v)) \neq \chi_{G, k+1}^i(\varphi(w))$.
+	The lemma then follows by \cref{defn:related:wl-distinguishable}.
 \end{hproof}
-\end{thm}
 \begin{figure}[ht]
@@
 	\includegraphics[width=\linewidth]{gfx/related-work/wl2-problem-solution.pdf}
-	\caption[Two simple non-isomorphic graphs that are indistinguishable by 1-\acs{wl}.]{
-		Two simple non-isomorphic graphs that are indistinguishable by 1-\acs{wl}.
+	\caption{
+		Two non-isomorphic graphs with $G \mathrel{\simeq_1} H$ and $G \mathrel{\not\simeq_2} H$.
 	}\label{fig:related:wl2-problem-solution}
 \end{figure}
+\Cref{fig:related:wl2-problem-solution} illustrates the higher discriminative power of higher \acs{wl}-dimensions;
+it shows how 2-WL is able to distinguish the two 1-WL indistinguishable graphs from \cref{fig:related:wl1-problem}.
 
@@
 	\max_{f, f_C} \sum_{i = 1}^n \log P(C_k(w_i) | w_i)\ =\ \max_{f, f_C} \sum_{i = 1}^n \smashoperator[r]{\sum_{w_j \in C_k(w_i)}}\, \left[ {f(w_i)}^\top f_C(w_j) - \log Z_{w_i} \right] \label{eq:related:gcr:embed:word2vec}\\
-	\text{with } P(C_k(w_i) | w_i) := \prod_{\mathclap{w_j \in C_k(w_i)}}\, \overbrace{\frac{\exp\left({f(w_i)}^\top f_C(w_j)\right)}{Z_{w_i}}}^{P(w_j | w_i)}
-	\text{ and } Z_{w_i} := \sum_{j = 1}^n \exp\left({f(w_i)}^\top f_C(w_j)\right) \nonumber
+	\text{with } P(C_k(w_i) | w_i) \coloneqq \prod_{\mathclap{w_j \in C_k(w_i)}}\, \overbrace{\frac{\exp\left({f(w_i)}^\top f_C(w_j)\right)}{Z_{w_i}}}^{P(w_j | w_i)}
+	\text{ and } Z_{w_i} \coloneqq \sum_{j = 1}^n \exp\left({f(w_i)}^\top f_C(w_j)\right) \nonumber
 \end{gather}
@@
 \begin{equation}
-	C_k(G_i) := \bigcup_{\mathclap{v_j \in \mathcal{V}_i}}\, {\left\{ {\textrm{WL}_l(G_i)}_j \right\}}_{l = 0}^k % chktex 21
-	\text{ with } \textrm{WL}_l(G) := \text{$l$ iter.\ \acs{wl}-1 coloring of $G$.}
+	C_k(G_i) \coloneqq \bigcup_{\mathclap{v_j \in \mathcal{V}_i}}\, {\left\{ {\textrm{WL}_l(G_i)}_j \right\}}_{l = 0}^k % chktex 21
+	\text{ with } \textrm{WL}_l(G) \coloneqq \text{$l$ iter.\ \acs{wl}-1 coloring of $G$.}
 \end{equation}

### 2020-02-18 thesis/content/chapter-related-work.tex
continued wl section

@@
 
-The \acf{wl} algorithm~\cite{Weisfeiler1968} characterizes a graph $G = (\mathcal{V}, \mathcal{E})$ by assigning discrete labels $c \in \mathcal{C}$, called \textit{colors}, to vertex $k$-tuples $(v_1, \dots, v_k) \in \mathcal{V}^k$, where $k \in \mathbb{N}$ is the freely choosable \textit{\ac{wl}-dimension}.
+The \acf{wl} algorithm~\cite{Weisfeiler1968}\cite{Cai1992} characterizes a graph $G = (\mathcal{V}, \mathcal{E})$ by assigning discrete labels $c \in \mathcal{C}$, called \textit{colors}, to vertex $k$-tuples $(v_1, \dots, v_k) \in \mathcal{V}^k$, where $k \in \mathbb{N}$ is the freely choosable \textit{\ac{wl}-dimension}.
 A mapping $\chi_{G, k}: \mathcal{V}^k \to \mathcal{C}$ is called a \textit{$k$-coloring} of $G = (\mathcal{V}, \mathcal{E})$.
@@
 Now we will look at the types of graphs that can be distinguished by \acs{wl} in relation to the \acs{wl}-dimension $k$.
-\begin{lem}
-	$G \mathrel{\not\simeq_k} H \implies G \mathrel{\not\simeq_{k+1}} H$ (the \acs{wl} power is never lower for higher $k$).
+\begin{lem}\label{lem:related:wl-lower}
+	$G \mathrel{\not\simeq_k} H \implies G \mathrel{\not\simeq_{k+1}} H$ ($(k+1)$-\acs{wl} is at least as powerful as $k$-\acs{wl}).
 \end{lem}
@@
 \end{hproof}
+\begin{fact}[see~\citet{Immerman1990}]\label{fact:related:wl-upper}
+	For all $k$ there are non-isomorphic graphs $G \not\simeq H$ of size $\mathcal{O}(k)$ with $G \mathrel{\simeq_k} H$.
+\end{fact}
 \begin{figure}[ht]
@@
 \end{figure}
-\Cref{fig:related:wl2-problem-solution} illustrates the higher discriminative power of higher \acs{wl}-dimensions;
-it shows how 2-WL is able to distinguish the two 1-WL indistinguishable graphs from \cref{fig:related:wl1-problem}.
+\begin{cor}\label{cor:related:wl-monotonous}
+	The discriminative power of $k$-\acs{wl} grows monotonously with $k$ and never converges, i.e.\ for all $k \in \mathbb{N}$ there is an $l \in \mathbb{N}$ s.t.\ the set of $k$-\ac{wl} distinguishable graphs is a true subset of the $(k+l)$-\ac{wl} distinguishable graphs.
+\end{cor}
+\begin{hproof}
+	Note that $k$-\acs{wl} trivially solves the graph isomorphism problem for all graphs of size $\leq k$ via the initial coloring (see~\cref{eq:related:wlk-init}).
+	Using \cref{lem:related:wl-lower} and \cref{fact:related:wl-upper} the corollary directly follows.
+\end{hproof}
+\Cref{fig:related:wl2-problem-solution} illustrates \cref{cor:related:wl-monotonous} by showing how 2-WL is able to distinguish the two 1-WL indistinguishable graphs from \cref{fig:related:wl1-problem}.
+Since the time complexity of $k$-\acs{wl} grows exponentially with $k$~\cite[corr.~1.9.7]{Immerman1990}, a trade-off between the runtime and the discriminative power has to be made.
 
@@
 \begin{equation}
-	C_k(G_i) \coloneqq \bigcup_{\mathclap{v_j \in \mathcal{V}_i}}\, {\left\{ {\textrm{WL}_l(G_i)}_j \right\}}_{l = 0}^k % chktex 21
-	\text{ with } \textrm{WL}_l(G) \coloneqq \text{$l$ iter.\ \acs{wl}-1 coloring of $G$.}
+	C_k(G_i) \coloneqq \bigcup_{\mathclap{v_j \in \mathcal{V}_i}}\, {\left\{ {\chi_{G_i, 1}^l(v_j) \right\}}_{l = 0}^k % chktex 21
+	\text{ with $k \in \mathbb{N}$ and $\chi_{G_i, 1}^l$ as in \cref{defn:related:wl1-refine}.}
 \end{equation}
-Intuitively this context can be understood as the set of \acs{wl}-distinguishable subgraphs of $G_i$ with diameter $\leq 2k$.
+Intuitively this context can be understood as the set of 1-\acs{wl}-distinguishable subgraphs of $G_i$ with diameter $\leq 2k$.
 Since \ac{wl} is used to identify distinct subgraphs, \texttt{graph2vec} can only be applied to graphs with discrete vertex labels.

### 2020-02-19 thesis/content/chapter-related-work.tex
continued wl section

@@
 
-To classify or score a graph, it first needs to be characterized by a set of relevant properties.
-Two commonly used graph characterization approaches are the so called \acl{wl} coloring and the notion of the graph spectrum.
-Both will now be briefly introduced since they form the foundation of most current \ac{gcr} methods.
+\begin{defn}
+	A graph $G \coloneqq (\mathcal{V}_G, \mathcal{E}_G)$ consists of a set of vertices $v_i \in \mathcal{V}_G$ and a set of edges $e_{ij} = (v_i, v_j) \in \mathcal{E}_G \subseteq \mathcal{V}_G^2$.
+	Optionally discrete vertex labels $l_i \in L_{\mathcal{V}}$ or edge labels $l_{ij} \in L_{\mathcal{E}}$ may be associated with all vertices $v_i \in \mathcal{V}_G$ and edges $e_{ij}\in \mathcal{E}_G$ respectively.
+	Also continuous feature vectors $x_{i} \in \mathcal{X}_{\mathcal{V}}$, $x_{ij} \in \mathcal{X}_{\mathcal{E}}$ may be given.
+	If $\mathcal{X}_{\mathcal{E}} = \mathbb{R}$, $x_{ij}$ can for example be interpreted as an edge weight of $e_{ij}$.
+\end{defn}
+In this thesis only undirected graphs $G$ will be considered, i.e.\ ${e_{ij} \in \mathcal{E}_G} \leftrightarrow\allowbreak {e_{ji} \in \mathcal{E}_G} \allowbreak\land l_{ij} = l_{ji} \allowbreak\land x_{ij} = x_{ji}$.
+To classify or score such a graph, it first needs to be characterized by a set of relevant properties.
+The most strict characterization of a graph is its so-called \textit{isomorphism type}.
+It uniquely identifies a graph but lacks any notion of similarity between non-identical graphs.
+We will begin with a brief definition of this strict isomorphism-based graph characterization.
+Then two less strict characterization approaches are described; the so called \acl{wl} coloring and the notion of graph spectra.
+They are the theoretical foundation of many current \ac{gcr} methods.
+
+\subsection{The Graph Isomorphism Problem}%
+\label{sec:related:character:iso}
 
@@
 
-The \acf{wl} algorithm~\cite{Weisfeiler1968}\cite{Cai1992} characterizes a graph $G = (\mathcal{V}, \mathcal{E})$ by assigning discrete labels $c \in \mathcal{C}$, called \textit{colors}, to vertex $k$-tuples $(v_1, \dots, v_k) \in \mathcal{V}^k$, where $k \in \mathbb{N}$ is the freely choosable \textit{\ac{wl}-dimension}.
-A mapping $\chi_{G, k}: \mathcal{V}^k \to \mathcal{C}$ is called a \textit{$k$-coloring} of $G = (\mathcal{V}, \mathcal{E})$.
-We say  $\chi'$ \textit{refines} $\chi$ ($\chi' \succeq \chi$) iff.\ $\forall a, b \in \mathcal{V}^k: \chi(a) \neq \chi(b) \rightarrow \chi'(a) \neq \chi'(b)$, i.e.\ $\chi'$ distinguishes at least those tuples that are distinguished by $\chi$.
-
-The $k$-dimensional \ac{wl} algorithm ($k$-\acs{wl}) works by iteratively refining $k$-colorings $\chi_{G, k}^0 \preceq \chi_{G, k}^1 \preceq \dots$ of a given graph $G$ until the convergence criterion $\chi_{G, k}^i \equiv \chi_{G, k}^{i+1}$ is satisfied;
-here the equivalence operator $\equiv$ denotes that $\chi_{G, k}^i$ and $\chi_{G, k}^{i+1}$ must be identical up to color/label substitutions.
-The final \ac{wl} coloring $\bar{\chi}_{G, k}$ that is obtained after convergence can be used to quickly check whether two graphs $G, H$ are not isomorphic ($G \not\simeq H$).
+The \acf{wl} algorithm~\cite{Weisfeiler1968}\cite{Cai1992} characterizes a graph $G$ by assigning discrete labels $c \in \mathcal{C}$, called \textit{colors}, to vertex $k$-tuples $(v_1, \dots, v_k) \in \mathcal{V}_G^k$, where $k \in \mathbb{N}$ is the freely choosable \textit{\ac{wl}-dimension}.
+A mapping $\chi_{G, k}: \mathcal{V}_G^k \to \mathcal{C}$ is called a \textit{$k$-coloring} of $G$.
+\begin{defn}
+	A coloring $\chi'$ \textit{refines} $\chi$ ($\chi' \succeq \chi$) iff.\ $\forall a, b \in \mathcal{V}_G^k: \chi(a) \neq \chi(b) \rightarrow \chi'(a) \neq \chi'(b)$, i.e.\ $\chi'$ distinguishes at least those tuples that are distinguished by $\chi$.
+\end{defn}
+\begin{defn}
+	Two colorings $\chi$ and $\chi'$ are \textit{equivalent} ($\chi \equiv \chi'$) iff.\ $\forall a, b \in \mathcal{V}_G^k: {\chi(a) = \chi(b)} \leftrightarrow {\chi'(a) = \chi'(b)}$, i.e.\ $\chi$ is identical to $\chi'$ up to color substitutions.
+\end{defn}
+The $k$-dimensional \ac{wl} algorithm ($k$-\acs{wl}) works by iteratively refining $k$-colorings $\chi_{G, k}^{(0)} \preceq \chi_{G, k}^{(1)} \preceq \dots$ of a given graph $G$ until the convergence criterion $\chi_{G, k}^{(i)} \equiv \chi_{G, k}^{(i+1)}$ is satisfied.
+We denote the final, maximally refined, $k$-\acs{wl} coloring with $\chi^{*}_{G, k}$.
 \begin{defn}\label{defn:related:wl-distinguishable}
-	Two graphs $G$ and $H$ are are $k$-\acs{wl} \textit{distinguishable} ($G \mathrel{{\not\simeq}_k} H$) iff.\ there exists a color $c \in \mathcal{C}$ s.t.\ $\left|\left\{ a \in \mathcal{V}_G^k\, |\, \bar{\chi}_{G, k}(a) = c \right\}\right|\neq \left|\left\{ b \in \mathcal{V}_H^k\, |\, \bar{\chi}_{H, k}(b) = c \right\}\right|$.
+	Two graphs $G$ and $H$ are are $k$-\acs{wl} \textit{distinguishable} ($G \mathrel{{\not\simeq}_k} H$) iff.\ there exists a color $c \in \mathcal{C}$ s.t.\ $\left|\left\{ a \in \mathcal{V}_G^k\, |\, \chi^{*}_{G, k}(a) = c \right\}\right|\neq \left|\left\{ b \in \mathcal{V}_H^k\, |\, \chi^{*}_{H, k}(b) = c \right\}\right|$.
 % chktex 21
@@
 In the 1-dimensional \ac{wl} algorithm, a color is assigned to each vertex of a graph.
-If the vertices $\mathcal{V}_G$ of the input graph $G$ are already labeled via some labeling $l_G: \mathcal{V}_G \to L$, those labels can be used as the initial graph coloring $\chi_{G,1}^0 = l_G$.
-For unlabeled graphs a constant coloring is used instead, e.g.\ $\forall v \in \mathcal{V}_G: \chi_{G,1}^0(v) = \textcolor{t_blue}{\texttt{A}}$ for some initial color $\textcolor{t_blue}{\texttt{A}} \in \mathcal{C}$.
+If the vertices $\mathcal{V}_G$ of the input graph $G$ are already labeled via some labeling $l_G: \mathcal{V}_G \to L$, those labels can be used as the initial graph coloring $\chi_{G,1}^{(0)} = l_G$.
+For unlabeled graphs a constant coloring is used instead, e.g.\ $\forall v \in \mathcal{V}_G: \chi_{G,1}^{(0)}(v) = \textcolor{t_blue}{\texttt{A}}$ for some initial color $\textcolor{t_blue}{\texttt{A}} \in \mathcal{C}$.
 % chktex 25
@@
 \begin{defn}\label{defn:related:wl1-refine}
-	$\chi_{G,1}^{i+1}(v) \coloneqq h\left(\chi_{G,1}^{i}(v), \ldblbrace \chi_{G,1}^{i}(u)\, |\, u \in \Gamma_{G}(v) \rdblbrace\right)$,
+	$\chi_{G,1}^{(i+1)}(v) \coloneqq h\left(\chi_{G,1}^{(i)}(v), \ldblbrace \chi_{G,1}^{(i)}(u)\, |\, u \in \Gamma_{G}(v) \rdblbrace\right)$,
 	with $\Gamma_G(v)$ denoting the set of adjecent vertices of $v \in \mathcal{V}_G$ and $h: \mathcal{C}^* \to \mathcal{C}$ denoting a perfect hash function that assigns a unique color to each finite combination of colors.
@@
 	$\begin{aligned}[t]
-		\chi_{G,k}^{i+1}(v) \coloneqq &\ h\left(\chi_{G,k}^{i}(v), \ldblbrace (\chi_{G,k}^{i}(v[u/1]), \dots, \chi_{G,k}^{i}(v[u/k]))\, |\, u \in \mathcal{V}_G \rdblbrace\right)\\
-		\text{with } v = &\ (v_1, \dots, v_k) \in \mathcal{V}^k, v[u/j] \coloneqq (v_1, \dots, v_{j-1}, u, v_{j+1}, \dots, v_k) \text{.}
+		\chi_{G,k}^{(i+1)}(v) \coloneqq &\ h\left(\chi_{G,k}^{(i)}(v), \ldblbrace (\chi_{G,k}^{(i)}(v[u/1]), \dots, \chi_{G,k}^{(i)}(v[u/k]))\, |\, u \in \mathcal{V}_G \rdblbrace\right)\\
+		\text{with } v = &\ (v_1, \dots, v_k) \in \mathcal{V}_G^k, v[u/j] \coloneqq (v_1, \dots, v_{j-1}, u, v_{j+1}, \dots, v_k) \text{.}
 	\end{aligned}$
@@
 In 1-\acs{wl} a vertex color is refined by combining the colors of neighboring vertices.
-In $k$-\acs{wl} the color of a $k$-tuple $v \in \mathcal{V}^k$ is refined by combining the colors of its neighborhood which is defined as the set of all $k$-tuples in which at most one vertex differs from $v$.
+In $k$-\acs{wl} the color of a $k$-tuple $v \in \mathcal{V}_G^k$ is refined by combining the colors of its neighborhood which is defined as the set of all $k$-tuples in which at most one vertex differs from $v$.
 Note that each vertex $k$-tuple has one neighbor for each $u \in \mathcal{V}_G$, each of which is a $k$-tuple of vertex $k$-tuples.
@@
 For $k = 2$ this means that each potential edge $(v, w) \in \mathcal{V}_G^2$ has all possible paths of length $2$ from $v$ to $w$ as its neighbors (see~\cref{fig:related:wl-neighbors:2}).
-Also note that, even though $k$-\acs{wl} refines $k$-tuple colors, lower-dimensional structures still get their own colors since a tuple does not have to consist of distinct vertices, i.e.\ in $k$-\acs{wl} the color of a single vertex $v \in \mathcal{V}_G$ is described by $\bar{\chi}_{G, k}(s)$ for $s = (v, \dots, v) \in \mathcal{V}_G^k$.
+Also note that, even though $k$-\acs{wl} refines $k$-tuple colors, lower-dimensional structures still get their own colors since a tuple does not have to consist of distinct vertices, i.e.\ in $k$-\acs{wl} the color of a single vertex $v \in \mathcal{V}_G$ is described by $\chi^{*}_{G, k}(s)$ for $s = (v, \dots, v) \in \mathcal{V}_G^k$.
 
 Let us now look at how the tuple colors are initialized.
-For this we use subgraph \textit{isomorphism types} which determine the initial color $\chi_{G,k}^0(v)$ of each $k$-tuple $v$.
+For this we use subgraph \textit{isomorphism types} which determine the initial color $\chi_{G,k}^{(0)}(v)$ of each $k$-tuple $v$.
 For $k = 1$ the isomorphism type of a vertex $v$ directly corresponds to its label $l_G(v)$.
@@
 \begin{align}
-	\chi_{G,k}^0(v_1, \dots, v_k) = \chi_{G,k}^0(w_1, \dots, w_k) \label{eq:related:wlk-init}
+	\chi_{G,k}^{(0)}(v_1, \dots, v_k) = \chi_{G,k}^{(0)}(w_1, \dots, w_k) \label{eq:related:wlk-init}
 	\iff &\ \forall i: l_G(v_i) = l_G(w_i) \\
@@
 	For each neighbor $(v[u/1], \dots,\allowbreak v[u/k])$ of $v$, there is a corresponding neighbor $(\varphi(v[u/1]), \dots \varphi(v[u/k]), \varphi(v[u/k]))$ of $\varphi(v)$.
-	Using \cref{eq:related:wlk-init} (for $i = 0$) and \cref{defn:related:wlk-refine} (for $i > 0$) it follows that $\forall v, w \in \mathcal{V}_G^k: \chi_{G, k}^i(v) \neq \chi_{G, k}^i(w) \rightarrow \chi_{G, k+1}^i(\varphi(v)) \neq \chi_{G, k+1}^i(\varphi(w))$.
+	Using \cref{eq:related:wlk-init} (for $i = 0$) and \cref{defn:related:wlk-refine} (for $i > 0$) it follows that $\forall v, w \in \mathcal{V}_G^k: \chi_{G, k}^{(i)}(v) \neq \chi_{G, k}^{(i)}(w) \rightarrow \chi_{G, k+1}^{(i)}(\varphi(v)) \neq \chi_{G, k+1}^{(i)}(\varphi(w))$.
 	The lemma then follows by \cref{defn:related:wl-distinguishable}.
 \end{hproof}
-\begin{fact}[see~\citet{Immerman1990}]\label{fact:related:wl-upper}
-	For all $k$ there are non-isomorphic graphs $G \not\simeq H$ of size $\mathcal{O}(k)$ with $G \mathrel{\simeq_k} H$.
-\end{fact}
+\begin{prop}[see~\citet{Immerman1990}]\label{prop:related:wl-upper}
+	For all $k \in \mathbb{N}$ there are non-isomorphic graphs $G \not\simeq H$ of size $\mathcal{O}(k)$ with $G \mathrel{\simeq_k} H$.
+\end{prop}
 \begin{figure}[ht]
@@
 \begin{cor}\label{cor:related:wl-monotonous}
-	The discriminative power of $k$-\acs{wl} grows monotonously with $k$ and never converges, i.e.\ for all $k \in \mathbb{N}$ there is an $l \in \mathbb{N}$ s.t.\ the set of $k$-\ac{wl} distinguishable graphs is a true subset of the $(k+l)$-\ac{wl} distinguishable graphs.
+	The discriminative power of $k$-\acs{wl} grows monotonously with $k$ and never converges, i.e.\ for all $k \in \mathbb{N}$ there is an $l \in \mathbb{N}$ s.t.\ the set of $k$-\ac{wl} distinguishable graphs is a proper subset of the $(k+l)$-\ac{wl} distinguishable graphs.
 \end{cor}
@@
 	Note that $k$-\acs{wl} trivially solves the graph isomorphism problem for all graphs of size $\leq k$ via the initial coloring (see~\cref{eq:related:wlk-init}).
-	Using \cref{lem:related:wl-lower} and \cref{fact:related:wl-upper} the corollary directly follows.
+	Using \cref{lem:related:wl-lower} and \cref{prop:related:wl-upper} the corollary directly follows.
 \end{hproof}
@@
 While in \texttt{word2vec} every word can be a context root as well as a context element, \texttt{graph2vec} uses the graphs $\mathbb{G}$ as context roots.
-The context $C_k(G_i)$ of a graph $G_i = (\mathcal{V}_i, \mathcal{E}_i)$ is defined as
+The context $C_k(G_i)$ of a graph $G_i$ is defined as
 \begin{equation}
-	C_k(G_i) \coloneqq \bigcup_{\mathclap{v_j \in \mathcal{V}_i}}\, {\left\{ {\chi_{G_i, 1}^l(v_j) \right\}}_{l = 0}^k % chktex 21
-	\text{ with $k \in \mathbb{N}$ and $\chi_{G_i, 1}^l$ as in \cref{defn:related:wl1-refine}.}
+	C_k(G_i) \coloneqq \bigcup_{\mathclap{v_j \in \mathcal{V}_{G_i}}}\, {\left\{ {\chi_{G_i, 1}^{(l)}(v_j) \right\}}_{l = 0}^k % chktex 21
+	\text{ with $k \in \mathbb{N}$ and $\chi_{G_i, 1}^{(l)}$ as in \cref{defn:related:wl1-refine}.}
 \end{equation}

### 2020-02-20 thesis/content/chapter-related-work.tex
continued wl section

@@
 	A graph $G \coloneqq (\mathcal{V}_G, \mathcal{E}_G)$ consists of a set of vertices $v_i \in \mathcal{V}_G$ and a set of edges $e_{ij} = (v_i, v_j) \in \mathcal{E}_G \subseteq \mathcal{V}_G^2$.
-	Optionally discrete vertex labels $l_i \in L_{\mathcal{V}}$ or edge labels $l_{ij} \in L_{\mathcal{E}}$ may be associated with all vertices $v_i \in \mathcal{V}_G$ and edges $e_{ij}\in \mathcal{E}_G$ respectively.
-	Also continuous feature vectors $x_{i} \in \mathcal{X}_{\mathcal{V}}$, $x_{ij} \in \mathcal{X}_{\mathcal{E}}$ may be given.
-	If $\mathcal{X}_{\mathcal{E}} = \mathbb{R}$, $x_{ij}$ can for example be interpreted as an edge weight of $e_{ij}$.
+	Optionally discrete vertex labels $l_G[v_i] \in L_{\mathcal{V}}$ or edge labels $l_G[e_{ij}] \in L_{\mathcal{E}}$ may be associated with all vertices $v_i \in \mathcal{V}_G$ and edges $e_{ij}\in \mathcal{E}_G$ respectively.
+	Also continuous feature vectors $x_G[v_{i}] \in \mathcal{X}_{\mathcal{V}}$, $x_G[e_{ij}] \in \mathcal{X}_{\mathcal{E}}$ may be given.
+	If $\mathcal{X}_{\mathcal{E}} = \mathbb{R}$, $x_G[e_{ij}]$ can for example be interpreted as an edge weight of $e_{ij}$.
 \end{defn}
-In this thesis only undirected graphs $G$ will be considered, i.e.\ ${e_{ij} \in \mathcal{E}_G} \leftrightarrow\allowbreak {e_{ji} \in \mathcal{E}_G} \allowbreak\land l_{ij} = l_{ji} \allowbreak\land x_{ij} = x_{ji}$.
+In this thesis all graphs $G$ are assumed to be undirected if not explicitly stated otherwise, i.e.\ ${e_{ij} \in \mathcal{E}_G} \leftrightarrow\allowbreak {e_{ji} \in \mathcal{E}_G} \allowbreak\land l_G[e_{ij}] = l_G[e_{ji}] \allowbreak\land x_G[e_{ij}] = x_G[e_{ji}]$.
 To classify or score such a graph, it first needs to be characterized by a set of relevant properties.
@@
 \label{sec:related:character:iso}
+
+In order to process a graph $G$ of size $n = |\mathcal{V}_G|$ , one is forced to choose some encoding, e.g.\ an adjacency matrix $A \in {\{0, 1\}}^{n \times n}$.
+Such an encoding introduces a vertex ordering $v_1, \dots, v_n$ that does not carry any semantic meaning.
+This means that there are $n!$ equivalent encodings of $G$. % chktex 40
+\begin{defn}
+	Two encoded graphs $G$ and $H$ with vertices $\mathcal{V}_G = \{ v_1, \dots, v_n \}$ and $\mathcal{V}_H = \{ w_1, \dots, w_n \}$ respectively are \textit{isomorphic} ($G \simeq H$) iff.\
+\end{defn}
 
@@
 \end{defn}
-It is easy to show that $G \mathrel{{\not\simeq}_k} H \implies G \not\simeq H$.
+As we will see, the \ac{wl} color refinement does not depend on vertex or edge encoding orderings, thus any difference in the final coloring of two graphs always implies the non-isomorphism of the colored graphs, i.e.\ $G \mathrel{{\not\simeq}_k} H \implies G \not\simeq H$.
 The opposite does however not necessarily hold;
-two $k$-\acs{wl} indistinguishable graphs are not always isomorphic, i.e.\ $G \mathrel{{\simeq}_k} H \centernot\implies G \simeq H$.
+two $k$-\acs{wl} indistinguishable graphs are not always isomorphic, i.e.\ $G \mathrel{{\simeq}_k} H \Longarrownot\implies G \simeq H$.
 To understand why this is the case and which types of graphs are $k$-\acs{wl} distinguishable, we will now look at the details of \ac{wl} color refinement step.
@@
 In the 1-dimensional \ac{wl} algorithm, a color is assigned to each vertex of a graph.
-If the vertices $\mathcal{V}_G$ of the input graph $G$ are already labeled via some labeling $l_G: \mathcal{V}_G \to L$, those labels can be used as the initial graph coloring $\chi_{G,1}^{(0)} = l_G$.
+If the vertices $v \in \mathcal{V}_G$ of the input graph $G$ are labeled, those labels $l_G[v] \in L_{\mathcal{V}} \subseteq \mathcal{C}$ can be used as the initial graph coloring $\chi_{G,1}^{(0)}(v) \coloneqq l_G[v]$.
 For unlabeled graphs a constant coloring is used instead, e.g.\ $\forall v \in \mathcal{V}_G: \chi_{G,1}^{(0)}(v) = \textcolor{t_blue}{\texttt{A}}$ for some initial color $\textcolor{t_blue}{\texttt{A}} \in \mathcal{C}$.
@@
 	with $\Gamma_G(v)$ denoting the set of adjecent vertices of $v \in \mathcal{V}_G$ and $h: \mathcal{C}^* \to \mathcal{C}$ denoting a perfect hash function that assigns a unique color to each finite combination of colors.
-	In practice $h$ is usually defined lazily by using $\mathcal{C} = \mathbb{N}$ and non-deterministically enumerating color combinations in the order in which they are hashed s.t.\ a new color is introduced every time a previously unseen color combination is queried.
 \end{defn}
+In practice the hash function $h$ is usually defined lazily by using $\mathcal{C} = \mathbb{N}$ and non-deterministically enumerating color combinations in the order in which they are hashed s.t.\ a new color is introduced every time a previously unseen color combination is queried.
+This is illustrated in \cref{fig:related:wl1-refine}.
 \begin{figure}[ht]
@@
 		After two iterations the coloring stabilizes.
-		Each vertex is labeled with its current color and has the aggregated neighbors, as in \cref{defn:related:wl1-refine}, written next to it.
+		Each vertex $v$ is labeled with its current color and has its previous color and the colors of the hashed neighbors $\Gamma_G(v)$ written next to it (see \cref{defn:related:wl1-refine}).
 	}\label{fig:related:wl1-refine}
@@
 For this we use subgraph \textit{isomorphism types} which determine the initial color $\chi_{G,k}^{(0)}(v)$ of each $k$-tuple $v$.
-For $k = 1$ the isomorphism type of a vertex $v$ directly corresponds to its label $l_G(v)$.
+For $k = 1$ the isomorphism type of a vertex $v$ directly corresponds to its label $l_G[v]$.
 For $k > 1$ the isomorphism types more generally correspond to the classes of isomorphic ordered subgraphs of $G$ induced by the $k$-tuples $\mathcal{V}_G^k$.
@@
 	\chi_{G,k}^{(0)}(v_1, \dots, v_k) = \chi_{G,k}^{(0)}(w_1, \dots, w_k) \label{eq:related:wlk-init}
-	\iff &\ \forall i: l_G(v_i) = l_G(w_i) \\
+	\iff &\ \forall i: l_G[v_i] = l_G[w_i] \\
 	\land &\ \forall i, j: v_i = v_j \leftrightarrow w_i = w_j \nonumber \\

### 2020-02-21 thesis/content/chapter-related-work.tex
added graph isomorphism section
+ updated wl section to incorporate the definitions from the new GI section

@@
 \end{align}
-where $\lambda \in \mathbb{R}^n$ is a weight vector with ${\|\lambda\|}_1 = 1$ and $\pi$ is a sorting permutation of the input scores with $y_i < y_j \Rightarrow \pi(i) < \pi(j)$. % chktex 21
+where $\lambda \in \mathbb{R}^n$ is a weight vector with ${\|\lambda\|}_1 = 1$ and $\pi: [n] \to [n]$ is a sorting permutation of the input scores with $[n] \coloneqq \{ 1, \dots, n \}$ and $y_i < y_j \Rightarrow \pi(i) < \pi(j)$.
+% chktex 21
 Depending on the choice of the vector $\lambda$, the \ac{owa} function $\mathcal{A}_\lambda$ can express common aggregation functions like
@@
 \begin{defn}
-	A graph $G \coloneqq (\mathcal{V}_G, \mathcal{E}_G)$ consists of a set of vertices $v_i \in \mathcal{V}_G$ and a set of edges $e_{ij} = (v_i, v_j) \in \mathcal{E}_G \subseteq \mathcal{V}_G^2$.
+	A graph $G \coloneqq (\mathcal{V}_G, \mathcal{E}_G)$ consists of a finite set of vertices $v_i \in \mathcal{V}_G$ and a set of edges $e_{ij} = (v_i, v_j) \in \mathcal{E}_G \subseteq \mathcal{V}_G^2$.
 	Optionally discrete vertex labels $l_G[v_i] \in L_{\mathcal{V}}$ or edge labels $l_G[e_{ij}] \in L_{\mathcal{E}}$ may be associated with all vertices $v_i \in \mathcal{V}_G$ and edges $e_{ij}\in \mathcal{E}_G$ respectively.
 	Also continuous feature vectors $x_G[v_{i}] \in \mathcal{X}_{\mathcal{V}}$, $x_G[e_{ij}] \in \mathcal{X}_{\mathcal{E}}$ may be given.
-	If $\mathcal{X}_{\mathcal{E}} = \mathbb{R}$, $x_G[e_{ij}]$ can for example be interpreted as an edge weight of $e_{ij}$.
+	If $\mathcal{X}_{\mathcal{E}} = \mathbb{R}$, $x_G[e_{ij}]$ can be interpreted as an edge weight of $e_{ij}$.
+
+	In this thesis all graphs $G$ are assumed to be undirected if not explicitly stated otherwise, i.e.\ ${e_{ij} \in \mathcal{E}_G} \leftrightarrow\allowbreak {e_{ji} \in \mathcal{E}_G} \allowbreak\land l_G[e_{ij}] = l_G[e_{ji}] \allowbreak\land x_G[e_{ij}] = x_G[e_{ji}]$.
+	We denote the set of all undirected graphs as $\mathcal{G}$.
+	Additionally we denote $G[S] \coloneqq (S, \mathcal{E}_G \cap S^2)$ as the subgraph of $G$ induced by $S \subseteq \mathcal{V}_G$.
 \end{defn}
-In this thesis all graphs $G$ are assumed to be undirected if not explicitly stated otherwise, i.e.\ ${e_{ij} \in \mathcal{E}_G} \leftrightarrow\allowbreak {e_{ji} \in \mathcal{E}_G} \allowbreak\land l_G[e_{ij}] = l_G[e_{ji}] \allowbreak\land x_G[e_{ij}] = x_G[e_{ji}]$.
-To classify or score such a graph, it first needs to be characterized by a set of relevant properties.
-The most strict characterization of a graph is its so-called \textit{isomorphism type}.
+To classify or score a graph, it first needs to be characterized by a set of relevant properties.
+The most strict characterization of a graph is its so-called \textit{isomorphism class}.
 It uniquely identifies a graph but lacks any notion of similarity between non-identical graphs.
@@
 
-In order to process a graph $G$ of size $n = |\mathcal{V}_G|$ , one is forced to choose some encoding, e.g.\ an adjacency matrix $A \in {\{0, 1\}}^{n \times n}$.
+In order to process a graph $G$ of size $n = |\mathcal{V}_G|$, one is forced to choose some encoding, e.g.\ an adjacency matrix $A \in {\{0, 1\}}^{n \times n}$.
 Such an encoding introduces a vertex ordering $v_1, \dots, v_n$ that does not carry any semantic meaning.
 This means that there are $n!$ equivalent encodings of $G$. % chktex 40
+To represent those encodings we introduce the notion of ordered induced subgraphs.
 \begin{defn}
-	Two encoded graphs $G$ and $H$ with vertices $\mathcal{V}_G = \{ v_1, \dots, v_n \}$ and $\mathcal{V}_H = \{ w_1, \dots, w_n \}$ respectively are \textit{isomorphic} ($G \simeq H$) iff.\
+	For all vertex $k$-tuples $v = (v_1, \dots, v_k) \in \mathcal{V}_G^k$ let $\hat{v} = \{ v_i\, |\, i \in [k] \}$.
+	Then $G[v] \coloneqq (\hat{v}, \mathcal{E}_G \cap \hat{v}^2, v)$ is called the ordered subgraph of $G$ induced by $v$.
+	If $\hat{v} = \mathcal{V}_G$ and $k = |\mathcal{V}_G|$, we call $v$ an ordering of $G$ and $G[v]$ an encoding of $G$.
 \end{defn}
+\begin{defn}\label{defn:related:ordered-subgraph-equivalence}
+	Two ordered induced subgraphs $G[v]$ and $H[w]$ with $(v_1, \dots, v_k) \in \mathcal{V}_G^k$ and $(w_1, \dots, w_k) \in \mathcal{V}_H^k$ are \textit{equivalent} ($G[v] \equiv H[w]$) iff.\
+	\begin{align*}
+		&\ \begin{aligned}
+			&\ \forall i, j: l_G[v_i, v_j] = l_H[w_i, w_j] &\land&\ \forall i: l_G[v_i] = l_H[w_i] \\
+			\land&\ \forall i, j: x_G[v_i, v_j] = x_H[w_i, w_j] &\land&\ \forall i: x_G[v_i] = x_H[w_i]
+		\end{aligned}\\
+		&\land\ \forall i, j: (v_i, v_j) \in \mathcal{E}_G \leftrightarrow (w_i, w_j) \in \mathcal{E}_H \text{.}
+	\end{align*}
+	Using this notion of equivalence, we call $[G[v]] \coloneqq \{ G[w] \equiv G[v]\, |\, w \in \mathcal{V}_G^{*} \}$ the ordered subgraph equivalence class of $G[v]$.
+\end{defn}
+\begin{defn}
+	Two graphs $G$ and $H$ are \textit{isomorphic} ($G \simeq H$) iff.\ there are equivalent encodings $G[v] \equiv H[w]$ of them.
+	Consequently $[G] \coloneqq \{ H \simeq G\, |\, H \in \mathcal{G} \}$ is called the isomorphism class of $G$.
+\end{defn}
+The goal of the \ac{gi} problem is to check whether $G \simeq H$ for two arbitrary graphs.
+Even though there is no known universal algorithm in $P$ that solves \ac{gi}, \citet{Babai1980} showed that almost all graphs can be trivially distinguished in linear time.
+More recently \citet{Babai2015} also presented a quasipolynomial upper time bound for the remaining hard \ac{gi} instances.
+For all practical purposes, \ac{gi} can therefore be solved efficiently; e.g.\ via the \texttt{nauty} program~\cite{McKay}\cite{McKay2013}.
+One important subroutine in most \ac{gi} checkers is the \acl{wl} algorithm which will be described next.
 
@@
 \begin{defn}
-	A coloring $\chi'$ \textit{refines} $\chi$ ($\chi' \succeq \chi$) iff.\ $\forall a, b \in \mathcal{V}_G^k: \chi(a) \neq \chi(b) \rightarrow \chi'(a) \neq \chi'(b)$, i.e.\ $\chi'$ distinguishes at least those tuples that are distinguished by $\chi$.
+	A coloring $\chi'$ \textit{refines} $\chi$ ($\chi' \preceq \chi$) iff.\ $\forall a, b \in \mathcal{V}_G^k: \chi(a) \neq \chi(b) \rightarrow \chi'(a) \neq \chi'(b)$, i.e.\ $\chi'$ distinguishes at least those tuples that are distinguished by $\chi$.
 \end{defn}
@@
 \end{defn}
-The $k$-dimensional \ac{wl} algorithm ($k$-\acs{wl}) works by iteratively refining $k$-colorings $\chi_{G, k}^{(0)} \preceq \chi_{G, k}^{(1)} \preceq \dots$ of a given graph $G$ until the convergence criterion $\chi_{G, k}^{(i)} \equiv \chi_{G, k}^{(i+1)}$ is satisfied.
-We denote the final, maximally refined, $k$-\acs{wl} coloring with $\chi^{*}_{G, k}$.
+The $k$-dimensional \ac{wl} algorithm ($k$-\acs{wl}) works by iteratively refining $k$-colorings $\chi_{G, k}^{(0)} \succeq \chi_{G, k}^{(1)} \succeq \dots$ of a given graph $G$ until the convergence criterion $\chi_{G, k}^{(i)} \equiv \chi_{G, k}^{(i+1)}$ is satisfied.
+We denote the final, maximally refined $k$-\acs{wl} coloring with $\chi^{*}_{G, k}$.
 \begin{defn}\label{defn:related:wl-distinguishable}
@@
 \end{defn}
-As we will see, the \ac{wl} color refinement does not depend on vertex or edge encoding orderings, thus any difference in the final coloring of two graphs always implies the non-isomorphism of the colored graphs, i.e.\ $G \mathrel{{\not\simeq}_k} H \implies G \not\simeq H$.
+As we will see, the \ac{wl} color refinement strategy is vertex order invariant, thus any difference in the final coloring of two graphs always implies the non-isomorphism of the colored graphs, i.e.\ $G \mathrel{{\not\simeq}_k} H \implies G \not\simeq H$.
 The opposite does however not necessarily hold;
 two $k$-\acs{wl} indistinguishable graphs are not always isomorphic, i.e.\ $G \mathrel{{\simeq}_k} H \Longarrownot\implies G \simeq H$.
-To understand why this is the case and which types of graphs are $k$-\acs{wl} distinguishable, we will now look at the details of \ac{wl} color refinement step.
-First the color refinement algorithm for the most simple case of $k = 1$ is described.
+
+In addition to the binary aspect of \ac{wl} distinguishability and its relation to the \ac{gi} problem, \ac{wl} colorings are also useful for more fuzzy graph similarity comparisons as we will see in \cref{sec:related:gcr:kernel} when we look at graph kernels.
+Before that however, the details of \ac{wl} color refinement strategy have to be described.
+We begin with the color refinement algorithm for the most simple case of $k = 1$.
 Then the definitions and intuitions from the 1-dimensional case are extended to its higher-dimensional generalization.
@@
 If the vertices $v \in \mathcal{V}_G$ of the input graph $G$ are labeled, those labels $l_G[v] \in L_{\mathcal{V}} \subseteq \mathcal{C}$ can be used as the initial graph coloring $\chi_{G,1}^{(0)}(v) \coloneqq l_G[v]$.
+Since \ac{wl} colors are inherently discrete, continuous vertex feature vectors $x_G[v]$ are not considered here.
 For unlabeled graphs a constant coloring is used instead, e.g.\ $\forall v \in \mathcal{V}_G: \chi_{G,1}^{(0)}(v) = \textcolor{t_blue}{\texttt{A}}$ for some initial color $\textcolor{t_blue}{\texttt{A}} \in \mathcal{C}$.
@@
 	$\chi_{G,1}^{(i+1)}(v) \coloneqq h\left(\chi_{G,1}^{(i)}(v), \ldblbrace \chi_{G,1}^{(i)}(u)\, |\, u \in \Gamma_{G}(v) \rdblbrace\right)$,
-	with $\Gamma_G(v)$ denoting the set of adjecent vertices of $v \in \mathcal{V}_G$ and $h: \mathcal{C}^* \to \mathcal{C}$ denoting a perfect hash function that assigns a unique color to each finite combination of colors.
+	with $\Gamma_G(v)$ denoting the set of adjacent vertices of $v \in \mathcal{V}_G$ and $h: \mathcal{C}^* \to \mathcal{C}$ denoting a perfect hash function that assigns a unique color to each finite combination of colors.
 \end{defn}
@@
 As we just saw, the 1-\acs{wl} algorithm iteratively refines colorings of single vertices.
-While the obtained colorings differ for most non-isomorphic graphs $G \not\simeq H$, 1-\acs{wl} does not generally solve the graph isomorphism problem as illustrated in \cref{fig:related:wl1-problem}.
+While the obtained colorings differ for most non-isomorphic graphs $G \not\simeq H$, 1-\acs{wl} does not generally solve the \ac{gi} problem as illustrated in \cref{fig:related:wl1-problem}.
 \begin{figure}[ht]
@@
 Let us now look at how the tuple colors are initialized.
-For this we use subgraph \textit{isomorphism types} which determine the initial color $\chi_{G,k}^{(0)}(v)$ of each $k$-tuple $v$.
-For $k = 1$ the isomorphism type of a vertex $v$ directly corresponds to its label $l_G[v]$.
-For $k > 1$ the isomorphism types more generally correspond to the classes of isomorphic ordered subgraphs of $G$ induced by the $k$-tuples $\mathcal{V}_G^k$.
-Formally this means that
+For this we use the ordered subgraph equivalence classes $[G[v]]$ (see~\cref{defn:related:ordered-subgraph-equivalence}) which determine the initial color $\chi_{G,k}^{(0)}(v)$ of each $k$-tuple $v$.
+For $k = 1$ the equivalence class of a vertex $v$ directly corresponds to its label $l_G[v]$.
+More generally for $k > 1$ this means that
 \begin{align}
-	\chi_{G,k}^{(0)}(v_1, \dots, v_k) = \chi_{G,k}^{(0)}(w_1, \dots, w_k) \label{eq:related:wlk-init}
-	\iff &\ \forall i: l_G[v_i] = l_G[w_i] \\
-	\land &\ \forall i, j: v_i = v_j \leftrightarrow w_i = w_j \nonumber \\
-	\land &\ \forall i, j: (v_i, v_j) \in \mathcal{E}_G \leftrightarrow (w_i, w_j) \in \mathcal{E}_G \text{.} \nonumber
+	\chi_{G,k}^{(0)}(v) = \chi_{G,k}^{(0)}(w) \iff G[v] \equiv G[w] \text{.} \label{eq:related:wlk-init}
 \end{align}
-If no explicit vertex labeling $l_G$ is given, a single constant label for all vertices is assumed.
-Also note that there is a fundamental difference in how the adjacency information encoded in $\mathcal{E}_G$ is used in 1-\acs{wl} vs.\ $k$-\acs{wl}:
+Note that there is a fundamental difference in how the adjacency information encoded in $\mathcal{E}_G$ is used in 1-\acs{wl} vs.\ $k$-\acs{wl}:
 In 1-\acs{wl} a vertex coloring by itself cannot encode adjacency which is why this information is explicitly incorporated in each refinement step via $\Gamma_G$ (see~\cref{defn:related:wl1-refine}).
 In $k$-\acs{wl} on the other hand each pair of vertices $(v, u) \in \mathcal{V}_G^2$ appears in at least one $k$-tuple (assuming $k \geq 2$) and therefore has at least one color which can implicitly encode the adjacency information.
-Edges and non-edges are colored differently in the initial coloring as defined in \cref{eq:related:wlk-init};
+Edges and non-edges are colored differently in the initial coloring since $G[(v, u)] \not\equiv G[(w, u)]$ if $(v, u) \in \mathcal{E}_G$ but $(w, u) \notin \mathcal{E}_G$;
 thus no explicit adjacency information is needed in the $k$-\acs{wl} color refinement step in \cref{defn:related:wlk-refine}.
@@
 \begin{hproof}
-	Note that $k$-\acs{wl} trivially solves the graph isomorphism problem for all graphs of size $\leq k$ via the initial coloring (see~\cref{eq:related:wlk-init}).
+	Note that $k$-\acs{wl} trivially solves the \ac{gi} problem for all graphs of size $\leq k$ via the initial coloring (see~\cref{eq:related:wlk-init}).
 	Using \cref{lem:related:wl-lower} and \cref{prop:related:wl-upper} the corollary directly follows.
@@
 \begin{equation}
-	C_k(G_i) \coloneqq \bigcup_{\mathclap{v_j \in \mathcal{V}_{G_i}}}\, {\left\{ {\chi_{G_i, 1}^{(l)}(v_j) \right\}}_{l = 0}^k % chktex 21
+	C_k(G_i) \coloneqq \bigcup_{\mathclap{v_j \in \mathcal{V}_{G_i}}}\, {\left\{ \chi_{G_i, 1}^{(l)}(v_j) \right\}}_{l = 0}^k % chktex 21
 	\text{ with $k \in \mathbb{N}$ and $\chi_{G_i, 1}^{(l)}$ as in \cref{defn:related:wl1-refine}.}

### 2020-02-22 thesis/content/chapter-related-work.tex
continued wl section

@@
 \begin{defn}
-	Two colorings $\chi$ and $\chi'$ are \textit{equivalent} ($\chi \equiv \chi'$) iff.\ $\forall a, b \in \mathcal{V}_G^k: {\chi(a) = \chi(b)} \leftrightarrow {\chi'(a) = \chi'(b)}$, i.e.\ $\chi$ is identical to $\chi'$ up to color substitutions.
+	Two colorings $\chi$ and $\chi'$ are \textit{equivalent} ($\chi \equiv \chi'$) iff.\ $\chi \preceq \chi' \land \chi' \preceq \chi$, i.e.\ $\chi$ is identical to $\chi'$ up to color substitutions.
 \end{defn}
@@
 \end{defn}
-As we will see, the \ac{wl} color refinement strategy is vertex order invariant, thus any difference in the final coloring of two graphs always implies the non-isomorphism of the colored graphs, i.e.\ $G \mathrel{{\not\simeq}_k} H \implies G \not\simeq H$.
+As we will see, the way in which \ac{wl} colorings are refined is vertex order invariant;
+thus any difference in the final coloring of two graphs always implies the non-isomorphism of the colored graphs, i.e.\ $G \mathrel{{\not\simeq}_k} H \implies G \not\simeq H$.
 The opposite does however not necessarily hold;
-two $k$-\acs{wl} indistinguishable graphs are not always isomorphic, i.e.\ $G \mathrel{{\simeq}_k} H \Longarrownot\implies G \simeq H$.
+two $k$-\acs{wl} indistinguishable graphs are not always isomorphic, i.e.\ $\exists\, G, H: G \mathrel{{\simeq}_k} H \land G \not\simeq H$. % chktex 21
 
@@
 Since \ac{wl} colors are inherently discrete, continuous vertex feature vectors $x_G[v]$ are not considered here.
-For unlabeled graphs a constant coloring is used instead, e.g.\ $\forall v \in \mathcal{V}_G: \chi_{G,1}^{(0)}(v) = \textcolor{t_blue}{\texttt{A}}$ for some initial color $\textcolor{t_blue}{\texttt{A}} \in \mathcal{C}$.
+For unlabeled graphs a constant coloring is used, e.g.\ $\forall v \in \mathcal{V}_G: \chi_{G,1}^{(0)}(v) = \textcolor{t_blue}{\texttt{A}}$ for some initial color $\textcolor{t_blue}{\texttt{A}} \in \mathcal{C}$.
 % chktex 25
@@
 \end{defn}
-In practice the hash function $h$ is usually defined lazily by using $\mathcal{C} = \mathbb{N}$ and non-deterministically enumerating color combinations in the order in which they are hashed s.t.\ a new color is introduced every time a previously unseen color combination is queried.
+In practice the hash function $h$ is usually defined lazily by using $\mathcal{C} = \mathbb{N}$ and enumerating color combinations in whichever order they are hashed s.t.\ a new color is introduced every time a previously unseen color combination appears at runtime.
 This is illustrated in \cref{fig:related:wl1-refine}.
@@
 	$\begin{aligned}[t]
-		\chi_{G,k}^{(i+1)}(v) \coloneqq &\ h\left(\chi_{G,k}^{(i)}(v), \ldblbrace (\chi_{G,k}^{(i)}(v[u/1]), \dots, \chi_{G,k}^{(i)}(v[u/k]))\, |\, u \in \mathcal{V}_G \rdblbrace\right)\\
-		\text{with } v = &\ (v_1, \dots, v_k) \in \mathcal{V}_G^k, v[u/j] \coloneqq (v_1, \dots, v_{j-1}, u, v_{j+1}, \dots, v_k) \text{.}
+		\chi_{G,k}^{(i+1)}(v) \coloneqq h\left(\chi_{G,k}^{(i)}(v), \ldblbrace (\chi_{G,k}^{(i)}(v[u/1]), \dots, \chi_{G,k}^{(i)}(v[u/k]))\, |\, u \in \mathcal{V}_G \rdblbrace\right)\\
+		\text{with } v = (v_1, \dots, v_k) \in \mathcal{V}_G^k,\quad v[u/j] \coloneqq (v_1, \dots, v_{j-1}, u, v_{j+1}, \dots, v_k) \text{.}
 	\end{aligned}$
@@
 For this we use the ordered subgraph equivalence classes $[G[v]]$ (see~\cref{defn:related:ordered-subgraph-equivalence}) which determine the initial color $\chi_{G,k}^{(0)}(v)$ of each $k$-tuple $v$.
-For $k = 1$ the equivalence class of a vertex $v$ directly corresponds to its label $l_G[v]$.
+For $k = 1$ the equivalence class of a single vertex $v$ directly corresponds to its label $l_G[v]$.
 More generally for $k > 1$ this means that
@@
 \end{prop}
+\begin{cor}\label{cor:related:wl-monotonous}
+	The discriminative power of $k$-\acs{wl} grows monotonously with $k$ and never converges, i.e.\ for all $k \in \mathbb{N}$ there is an $l \in \mathbb{N}$ s.t.\ the set of $k$-\ac{wl} distinguishable graphs is a proper subset of the $(k+l)$-\ac{wl} distinguishable graphs.
+\end{cor}
+\begin{hproof}
+	Note that $k$-\acs{wl} trivially solves the \ac{gi} problem for all graphs of size $\leq k$ via the initial coloring (see~\cref{eq:related:wlk-init}).
+	Using \cref{lem:related:wl-lower} and \cref{prop:related:wl-upper} the corollary directly follows.
+\end{hproof}
 \begin{figure}[ht]
@@
 \end{figure}
-\begin{cor}\label{cor:related:wl-monotonous}
-	The discriminative power of $k$-\acs{wl} grows monotonously with $k$ and never converges, i.e.\ for all $k \in \mathbb{N}$ there is an $l \in \mathbb{N}$ s.t.\ the set of $k$-\ac{wl} distinguishable graphs is a proper subset of the $(k+l)$-\ac{wl} distinguishable graphs.
-\end{cor}
-\begin{hproof}
-	Note that $k$-\acs{wl} trivially solves the \ac{gi} problem for all graphs of size $\leq k$ via the initial coloring (see~\cref{eq:related:wlk-init}).
-	Using \cref{lem:related:wl-lower} and \cref{prop:related:wl-upper} the corollary directly follows.
-\end{hproof}
 \Cref{fig:related:wl2-problem-solution} illustrates \cref{cor:related:wl-monotonous} by showing how 2-WL is able to distinguish the two 1-WL indistinguishable graphs from \cref{fig:related:wl1-problem}.
-Since the time complexity of $k$-\acs{wl} grows exponentially with $k$~\cite[corr.~1.9.7]{Immerman1990}, a trade-off between the runtime and the discriminative power has to be made.
+Since the time complexity of $k$-\acs{wl} grows exponentially with $k$~\cite[cor.~1.9.7]{Immerman1990}, it does not provide an efficient universal solution to \ac{gi}.
+However it turns out that almost all graphs are \ac{wl} distinguishable even for a small constant $k$.
+
+For $k = 1$ \citet{Babai1980} have shown that two randomly selected non-isomorphic graphs $G \not\simeq H$ of size $n$ are 1-\acs{wl} indistinguishable with probability $n^{-\nicefrac{1}{7}}$.
+Thus 1-\acs{wl} is already able to distinguish most graphs; it fails however to distinguish any pair of unlabeled $d$-regular graphs of the same size~\cite[cor.~1.8.5]{Immerman1990}.
+\begin{defn}
+	A graph $G$ is called $d$-regular ($\mathit{rg}_d(G)$) iff.\ $\forall v \in \mathcal{V}_G: |\Gamma_G(v)| = d$.
+\end{defn}
+We have already seen this in \cref{fig:related:wl1-problem} since the ``six-cycle'' and the ``two-triangle'' graphs are in fact both 2-regular and of size 6.
+This restriction alone is typically not an issue in the context of \ac{gcr} though as molecular structures for example are rarely perfectly regular.
+A more relevant restriction of 1-\ac{wl} is the fact that it is unable to detect cycles of size $l \geq 3$.
 

### 2020-02-23 thesis/content/chapter-related-work.tex
continued wl section

@@
 To represent those encodings we introduce the notion of ordered induced subgraphs.
-\begin{defn}
+\begin{defn}\label{defn:related:ordered-subgraph}
 	For all vertex $k$-tuples $v = (v_1, \dots, v_k) \in \mathcal{V}_G^k$ let $\hat{v} = \{ v_i\, |\, i \in [k] \}$.
@@
 \end{defn}
-We have already seen this in \cref{fig:related:wl1-problem} since the ``six-cycle'' and the ``two-triangle'' graphs are in fact both 2-regular and of size 6.
+We have already seen this in \cref{fig:related:wl1-problem} since the ``six-cycle'' and the ``two-triangles'' graphs are in fact both 2-regular and of size 6.
 This restriction alone is typically not an issue in the context of \ac{gcr} though as molecular structures for example are rarely perfectly regular.
-A more relevant restriction of 1-\ac{wl} is the fact that it is unable to detect cycles of size $l \geq 3$.
+A more relevant restriction of 1-\acs{wl} is the fact that it is unable to detect cycles of length $m \geq 3$.
+\begin{defn}
+	$k$-\acs{wl} computes a function $f: \mathcal{G} \to P$ iff.\ that function can be expressed as $f(G) = g(\chi_{G, k}^{*})$ via some function $g: (\mathcal{V}_G^k \to \mathcal{C}) \to P$.
+\end{defn}
+\begin{defn}
+	$k$-\acs{wl} counts a certain subgraph $S$ iff.\ it computes $\mathit{count}_S(G) \coloneqq \left|\{ \hat{v}\, |\, v \in \mathcal{V}_G^{*} \land G[v] \equiv S \}\right|$ (see~\cref{defn:related:ordered-subgraph}).
+	Similarly $k$-\acs{wl} detects a subgraph $S$ iff.\ it computes $\mathit{contain}_S(G) \coloneqq \mathbbm{1}[\mathit{count}_S(G) > 0]$ with $\mathbbm{1}$ denoting the indicator function.
+\end{defn}
+To see why 1-\acs{wl} is unable to detect $m$-cycles in graphs, note that \cref{fig:related:wl1-problem} already contradicts the positive statement for $m = 3$;
+this counterexample can be trivially generalized to all larger $m$ by replacing the ``six-cycle'' graph with a ``$(2m)$-cycle-graph'' graph and the ``two-triangles'' graph with a ``two-$m$-cycles'' graph, preserving the 2-regularity and therefore the 1-\acs{wl} indistinguishability.
 

### 2020-02-24 thesis/content/chapter-related-work.tex
completed first wl section draft

--- 2020-02-22 thesis/content/chapter-related-work.tex
@@
 Such an encoding introduces a vertex ordering $v_1, \dots, v_n$ that does not carry any semantic meaning.
-This means that there are $n!$ equivalent encodings of $G$. % chktex 40
+Consequently there are $n!$ equivalent encodings of $G$. % chktex 40
 To represent those encodings we introduce the notion of ordered induced subgraphs.
-\begin{defn}
+\begin{defn}\label{defn:related:ordered-subgraph}
 	For all vertex $k$-tuples $v = (v_1, \dots, v_k) \in \mathcal{V}_G^k$ let $\hat{v} = \{ v_i\, |\, i \in [k] \}$.
@@
 	\begin{align*}
-		&\ \begin{aligned}
-			&\ \forall i, j: l_G[v_i, v_j] = l_H[w_i, w_j] &\land&\ \forall i: l_G[v_i] = l_H[w_i] \\
-			\land&\ \forall i, j: x_G[v_i, v_j] = x_H[w_i, w_j] &\land&\ \forall i: x_G[v_i] = x_H[w_i]
-		\end{aligned}\\
-		&\land\ \forall i, j: (v_i, v_j) \in \mathcal{E}_G \leftrightarrow (w_i, w_j) \in \mathcal{E}_H \text{.}
+		&\quad\forall i, j \in [k]: (v_i, v_j) \in \mathcal{E}_G \leftrightarrow (w_i, w_j) \in \mathcal{E}_H \\
+		&\, \begin{aligned}
+			\land&\ \forall i, j \in [k]: l_G[v_i, v_j] = l_H[w_i, w_j] &\land&\ \forall i \in [k]: l_G[v_i] = l_H[w_i] \\
+			\land&\ \forall i, j \in [k]: x_G[v_i, v_j] = x_H[w_i, w_j] &\land&\ \forall i \in [k]: x_G[v_i] = x_H[w_i]\text{.}
+		\end{aligned}
 	\end{align*}
-	Using this notion of equivalence, we call $[G[v]] \coloneqq \{ G[w] \equiv G[v]\, |\, w \in \mathcal{V}_G^{*} \}$ the ordered subgraph equivalence class of $G[v]$.
+	Using this notion of equivalence, we call $[G[v]] \coloneqq \{ G[w]\, |\, G[w] \equiv G[v] \land w \in \mathcal{V}_G^{*} \}$ the ordered subgraph equivalence class of $G[v]$.
 \end{defn}
@@
 	Two graphs $G$ and $H$ are \textit{isomorphic} ($G \simeq H$) iff.\ there are equivalent encodings $G[v] \equiv H[w]$ of them.
-	Consequently $[G] \coloneqq \{ H \simeq G\, |\, H \in \mathcal{G} \}$ is called the isomorphism class of $G$.
+	Consequently $[G] \coloneqq \{ H\, |\, H \simeq G \}$ is called the isomorphism class of $G$.
 \end{defn}
@@
 We denote the final, maximally refined $k$-\acs{wl} coloring with $\chi^{*}_{G, k}$.
+\begin{defn}
+	The color distribution $\mathit{dist}_{\chi_{G, k}}: \mathcal{C} \to \mathbb{N}$ of a $k$-coloring $\chi_{G, k}$ counts each color $c \in \mathcal{C}$ in the coloring, i.e.\ $\mathit{dist}_{\chi_{G, k}}(c) \coloneqq \left|\left\{ v \in \mathcal{V}_G^k\, |\, \chi_{G, k}(v) = c \right\}\right|$.
+% chktex 21
+\end{defn}
 \begin{defn}\label{defn:related:wl-distinguishable}
-	Two graphs $G$ and $H$ are are $k$-\acs{wl} \textit{distinguishable} ($G \mathrel{{\not\simeq}_k} H$) iff.\ there exists a color $c \in \mathcal{C}$ s.t.\ $\left|\left\{ a \in \mathcal{V}_G^k\, |\, \chi^{*}_{G, k}(a) = c \right\}\right|\neq \left|\left\{ b \in \mathcal{V}_H^k\, |\, \chi^{*}_{H, k}(b) = c \right\}\right|$.
-% chktex 21
+	Two graphs $G$ and $H$ are are $k$-\acs{wl} \textit{distinguishable} ($G \mathrel{{\not\simeq}_k} H$) iff.\ there exists a color $c \in \mathcal{C}$ s.t.\ $\mathit{dist}_{\chi^{*}_{G, k}}(c) \neq \mathit{dist}_{\chi^{*}_{H, k}}(c)$.
 \end{defn}
@@
 	\centering
-	\includegraphics[width=0.75\linewidth]{gfx/related-work/wl1-refine.pdf}
+	\includegraphics[width=0.72\linewidth]{gfx/related-work/wl1-refine.pdf}
 	\caption[Example 1-WL color refinement steps.]{
@@
 \end{figure}
-By extending \ac{wl} to higher dimensions such 1-\acs{wl} indistinguishable cases can however be handled.
+By extending \ac{wl} to higher dimensions, such 1-\acs{wl} indistinguishable cases can however be handled.
 Analogous to the 1-dimensional \cref{defn:related:wl1-refine}, the $k$-dimensional color refinement step is defined by
@@
 \begin{defn}
-	A graph $G$ is called $d$-regular ($\mathit{rg}_d(G)$) iff.\ $\forall v \in \mathcal{V}_G: |\Gamma_G(v)| = d$.
+	A graph $G$ is called $d$-regular ($\mathit{rg}_d(G)$) iff.\ $\forall v \in \mathcal{V}_G: {|\Gamma_G(v)|} = d$.
 \end{defn}
-We have already seen this in \cref{fig:related:wl1-problem} since the ``six-cycle'' and the ``two-triangle'' graphs are in fact both 2-regular and of size 6.
-This restriction alone is typically not an issue in the context of \ac{gcr} though as molecular structures for example are rarely perfectly regular.
-A more relevant restriction of 1-\ac{wl} is the fact that it is unable to detect cycles of size $l \geq 3$.
+We have already seen this in \cref{fig:related:wl1-problem} since the ``six-cycle'' and the ``two-triangles'' graphs are in fact both 2-regular and of size 6.
+This restriction alone is typically not an issue in the context of \ac{gcr} though, as molecular structures or social interaction graphs for example are rarely perfectly regular.
+A more relevant restriction of 1-\acs{wl} is the fact that it is unable to detect cycles of length $m \geq 3$.
+\begin{defn}\label{defn:related:wl-compute}
+	$k$-\acs{wl} computes a function $f: \mathcal{G} \to Y$ iff.\ that function can be expressed as $f(G) = g(\mathit{dist}_{\chi_{G, k}^{*}})$ via some function $g: (\mathcal{C} \to \mathbb{N}) \to Y$.
+\end{defn}
+\begin{defn}\label{defn:related:wl-count-detect}
+	$k$-\acs{wl} counts a certain subgraph $S$ iff.\ it computes $\mathit{count}_S(G) \coloneqq \left|\{ \hat{v}\, |\, v \in \mathcal{V}_G^{*} \land G[v] \equiv S \}\right|$ (see~\cref{defn:related:ordered-subgraph}).
+	Similarly $k$-\acs{wl} detects a subgraph $S$ iff.\ it computes $\mathit{contain}_S(G) \coloneqq \mathbbm{1}[\mathit{count}_S(G) > 0]$, with $\mathbbm{1}$ denoting the indicator function.
+\end{defn}
+To see why 1-\acs{wl} is unable to detect $m$-cycles in graphs, note that \cref{fig:related:wl1-problem} already contradicts the positive statement for $m = 3$;
+this counterexample can be trivially generalized to all larger $m$ by replacing the ``six-cycle'' graph with a ``$(2m)$-cycle-graph'' graph and the ``two-triangles'' graph with a ``two-$m$-cycles'' graph which preserves the 2-regularity and therefore the 1-\acs{wl} indistinguishability.
+
+This cycle-detection restriction of 1-\acs{wl} is relevant in practice because cycle counts are used in many domains to analyze graphs, e.g.\ triangle counts are commonly used in social network analysis to find interaction clusters~\cite{Newman2003}\cite{Welser2007} and the detection of 4-, 5- or 6-cycles is required to determine important chemical properties like the aromaticity of a molecule~\cite{Adamson1973}\cite{Kekule1866}.
+
+If we increase the \acs{wl}-dimension to $k = 2$, the described 1-\acs{wl} restrictions no longer apply.
+2-\acs{wl} is able to distinguish more than $1 - \mathcal{O}(\nicefrac{1}{n})$ of the regular $n$-vertex graphs~\cite[cor.~1.8.6]{Immerman1990} and can even count cycles.
+\begin{prop}[see \citet{Fuerer2017}]\label{prop:related:wl2-cycle-count}
+	2-\acs{wl} is able to count $m$-cycles for $m \leq 6$ but it cannot detect 8-cycles.
+	There are no results regarding 7-cycles.
+\end{prop}
+\begin{hproof}
+	We refer to \citet{Fuerer2017} for the full proof.
+	Only the idea behind triangle and 4-cycle counting is outlined here to give an intuition for why \cref{prop:related:wl2-cycle-count} holds.
+	Note that the 2-\acs{wl} neighbors of each edge $(v, w) \in \mathcal{E}_G$ correspond to all possible paths $(v, u, w)$ of length 2, i.e.\ all possible triangles.
+	By \cref{defn:related:wlk-refine} there must thus be a color subset $C^{\triangle}_j \subseteq \mathcal{C}$ representing that an edge is involved in exactly $j$ triangles after one refinement step.
+	2-\acs{wl} then trivially counts triangles by setting $g(\mathit{dist}_\chi) = \frac{1}{3} \sum_j j \sum_{c \in C^{\triangle}_j} \mathit{dist}_\chi(c)$ to satisfy \cref{defn:related:wl-compute}.
+	Analogously for 4-cycles, let $C^{\square}_j \subseteq \mathcal{C}$ be the colors indicating a non-edge $(v, w) \notin \mathcal{E}_G$ with $j$ common vertex neighbors $u \in \Gamma_G(v) \cap \Gamma_G(w)$.
+	The number of 4-cycles is then determined by the colors of the diagonals through them via $g(\mathit{dist}_\chi) = \frac{1}{2} \sum_{j \geq 2} \binom{j}{2} \sum_{c \in C^{\square}_j} \mathit{dist}_\chi(c)$.
+	Using a similar but more involved combinatorial argument requiring multiple color refinement steps, 5- and 6-cycle counting can be shown.
+\end{hproof}
+As we just saw, 2-\acs{wl} is significantly more powerful than 1-\acs{wl}.
+By \cref{prop:related:wl-upper} there are of course still 2-\acs{wl} indistinguishable graphs;
+among others those are the strongly regular graphs $\mathit{srg}_{n, d, \lambda, \mu}(G)$.
+\begin{defn}
+	$\begin{aligned}[t]
+		\mathit{srg}_{n, d, \lambda, \mu}(G) \iff\ & {|\mathcal{V}_G|} = n \quad\land\quad \forall v \in \mathcal{V}_G: {|\Gamma_G(v)|} = d \\
+		&\land \forall (v, w) \in \mathcal{E}_G: {|\Gamma_G(v) \cap \Gamma_G(w)|} = \lambda \\
+		&\land \forall (v, w) \in \mathcal{V}_G^2 \setminus \mathcal{E}_G: {|\Gamma_G(v) \cap \Gamma_G(w)|} = \mu
+	\end{aligned}$
+\end{defn}
+Generally this restriction of 2-\acs{wl} is not an issue since strongly regular graphs do not appear in typical \ac{gcr} datasets.
+Lastly there are few results regarding the classes of distinguishable graphs and computable functions for higher \ac{wl}-dimensions.
+For $k = 3$, \citet{Kiefer2017} have shown that all planar graphs can be distinguished.
 
@@
 So-called \textit{fingerprint embeddings} try to match a fixed set of subgraphs $S_1, \dots, S_d$ to the input graph.
-The embedding of a graph $G$ is a binary vector $f(G) = x \in {\{0, 1\}}^d$ with $x_i = \mathds{1}[\text{$S_i$ is subgraph of $G$}]$, where $\mathds{1}$ denotes the indicator function.
+The embedding of a graph $G$ is a binary vector $f(G) = x \in {\{0, 1\}}^d$ with $x_i \coloneqq \mathit{contain}_{S_i}(G)$ (see~\cref{defn:related:wl-count-detect}).
 This simple approach usually requires a careful choice of subgraphs but can still be competitive with the other more recent approaches we will look at in the following sections.

### 2020-02-25 thesis/content/chapter-related-work.tex
minor additions

@@
 2-\acs{wl} is able to distinguish more than $1 - \mathcal{O}(\nicefrac{1}{n})$ of the regular $n$-vertex graphs~\cite[cor.~1.8.6]{Immerman1990} and can even count cycles.
-\begin{prop}[see \citet{Fuerer2017}]\label{prop:related:wl2-cycle-count}
-	2-\acs{wl} is able to count $m$-cycles for $m \leq 6$ but it cannot detect 8-cycles.
-	There are no results regarding 7-cycles.
+\begin{prop}[see \citet{Fuerer2017} and \citet{Arvind2019}]\label{prop:related:wl2-cycle-count}
+	2-\acs{wl} is able to count $m$-cycles for $m \leq 7$ but it cannot detect 8-cycles.
 \end{prop}
 \begin{hproof}
-	We refer to \citet{Fuerer2017} for the full proof.
 	Only the idea behind triangle and 4-cycle counting is outlined here to give an intuition for why \cref{prop:related:wl2-cycle-count} holds.
@@
 	The number of 4-cycles is then determined by the colors of the diagonals through them via $g(\mathit{dist}_\chi) = \frac{1}{2} \sum_{j \geq 2} \binom{j}{2} \sum_{c \in C^{\square}_j} \mathit{dist}_\chi(c)$.
-	Using a similar but more involved combinatorial argument requiring multiple color refinement steps, 5- and 6-cycle counting can be shown.
+	Using a similar but more involved combinatorial argument requiring multiple color refinement steps, 5-, 6- and 7-cycle counting can be shown.
 \end{hproof}
@@
 \label{sec:related:character:spectral}
+
+An alternative perspective on graph characterization is provided by \textit{spectral graph theory}.
 

### 2020-02-27 thesis/content/chapter-related-work.tex
started spectral graph theory section

@@
 	The number of 4-cycles is then determined by the colors of the diagonals through them via $g(\mathit{dist}_\chi) = \frac{1}{2} \sum_{j \geq 2} \binom{j}{2} \sum_{c \in C^{\square}_j} \mathit{dist}_\chi(c)$.
-	Using a similar but more involved combinatorial argument requiring multiple color refinement steps, 5-, 6- and 7-cycle counting can be shown.
+	Using a similar but more involved combinatorial argument requiring multiple color refinement steps, 5-, 6- \& 7-cycle counting can be shown.
 \end{hproof}
@@
 Generally this restriction of 2-\acs{wl} is not an issue since strongly regular graphs do not appear in typical \ac{gcr} datasets.
-Lastly there are few results regarding the classes of distinguishable graphs and computable functions for higher \ac{wl}-dimensions.
-For $k = 3$, \citet{Kiefer2017} have shown that all planar graphs can be distinguished.
+By going to $k = 3$, even some strongly regular graphs as well as all planar graphs can be distinguished though~\cite{Kiefer2017}.
+Apart from that there are currently few results regarding the classes of distinguishable graphs and computable functions for even higher \ac{wl}-dimensions.
 
@@
 An alternative perspective on graph characterization is provided by \textit{spectral graph theory}.
+While \ac{wl} characterizes a graph via its color distribution $\mathit{dist}_{\chi_{G,k}^{*}}$, the spectral approach uses its so-called \textit{spectrum} $\lambda_G = (\lambda_{G, 1}, \dots, \lambda_{G, n}) \in \mathbb{R}^n$ where $n = |\mathcal{V}_G|$.
+The key idea behind this is to interpret the adjacency matrix $A \in \mathbb{R}^{n \times n}$ of $G$ not simply as an encoding of the edges ($A_{i j} = \mathbbm{1}[(v_i, v_j) \in \mathcal{E}_G]$) but as a linear operator $A: (\mathcal{V}_G \to \mathbb{R}) \to (\mathcal{V}_G \to \mathbb{R})$ acting on the vector space of real-valued functions with a vertex domain.
+This means that $A$ transforms so-called \textit{graph signals} $x: \mathcal{V}_G \to \mathbb{R}$ which are functions assigning \textit{signal strengths} to vertices.
+By applying $Ax$ the signal strength $x[v_i]$ of each vertex is added to the signal strengths of its neighbors $\Gamma_G(v_i)$.
+Using this functional perspective, one can characterize graphs via the \acf{ft}.
+We will now see how this is done;
+however since a comprehensive description of spectral graph theory would exceed the scope of this thesis, only a brief overview will be given.
+For a more detailed introduction to the field we refer to \citet{Shuman2013}.
+
+\paragraph{\ac{ft} on complex domains}
+Let us begin by describing the \ac{ft} for the more common case of functions with complex domains\footnote{
+	This description of the \ac{ft} naturally also includes functions with real domains.
+	The complex case was chosen because it is more compact.
+}.
+All functions $f: \mathbb{C} \to \mathbb{C}$ can be interpreted as infinite-dimensional vectors $f = \int_{\mathbb{C}} f(t)\, b_t \dif t$ with $b_t: \mathbb{C} \to \mathbb{C}$ being a standard basis vector/function defined as $b_t(s) \coloneqq \begin{cases} 1 & \text{if } s = t \\[-6pt] 0 & \text{else} \end{cases}$.
+% chktex 21
+The \acl{ft} $\hat{f}$ of $f$ corresponds to a change of basis from the standard basis vectors $b_t$ to the Fourier basis vectors $u_\xi(t) \coloneqq e^{2 \pi i \xi t}$, i.e.\ $f = \int_{\mathbb{C}} \hat{f}(t)\, u_\xi \dif \xi$.
+The Fourier basis is characterized by the fact that it is an eigenbasis of the so-called \textit{Laplace operator} $\Delta$, i.e.\ $\Delta u_\xi = \lambda_\xi u_\xi$ with $\lambda_\xi$ being the eigenvalue corresponding to $u_\xi$.
+In the complex domain this Laplace operator $\Delta$ is effectively the same as the second-derivative $\frac{\dif^2}{\dif t^2}$.
+Thus it is easy to see that indeed $\Delta u_\xi = \frac{\dif^2}{\dif t^2} u_\xi = \lambda_\xi u_\xi$ for $\lambda_\xi = -{(2 \pi \xi)}^2$.
+Note that the complex exponentials $u_\xi$ are the only eigenbasis of $\Delta$ on the complex plane $\mathbb{C}$.
+This allows us to interpret the eigenvectors $u_\xi$ and their corresponding eigenvalues $\lambda_\xi$ as a characterization of the complex plane.
+
+\paragraph{\ac{ft} on vertex domains}
+Similar to how the eigenvectors and eigenvalues of the complex Laplacian $\frac{\dif^2}{\dif t^2}$ characterize $\mathbb{C}$, we can characterize a graph $G$ by the eigenvectors and eigenvalues of the so-called \textit{combinatorial graph Laplacian} $L$.
+\begin{defn}
+	The combinatorial graph Laplacian $L$ of a graph $G$ is defined as
+	\begin{align*}
+		L \coloneqq D - A\quad\text{with the degree matrix } D \coloneqq {
+			\renewcommand*{\arraystretch}{0.5}
+			\begin{pmatrix}
+				d_1 & & \\
+				& \ddots & \\
+				& & d_n
+			\end{pmatrix}
+		}, d_i \coloneqq |\Gamma_G(v_i)| \text{.}
+	\end{align*}
+\end{defn}
 
@@
 There is a large variety of so-called \acp{gk} to do this.
-\acp{gk} can be used in combination with any kernel method, typically a \ac{svm}, to solve the \ac{gcr} problem.
+\acp{gk} can be used in combination with any kernelized learner, typically \acp{svm}, to solve the \ac{gcr} problem.
 While there is a large variety of different \acp{gk}, we will focus on those that are based on the previously described family of \ac{wl} algorithms.

### 2020-02-28 thesis/content/chapter-related-work.tex
continued sgt section

--- 2020-02-25 thesis/content/chapter-related-work.tex
@@
 
-In order to process a graph $G$ of size $n = |\mathcal{V}_G|$, one is forced to choose some encoding, e.g.\ an adjacency matrix $A \in {\{0, 1\}}^{n \times n}$.
+In order to process a graph $G$ of size $n = |\mathcal{V}_G|$, one is forced to choose some encoding, e.g.\ an adjacency matrix $A_G \in {\{0, 1\}}^{n \times n}$.
 Such an encoding introduces a vertex ordering $v_1, \dots, v_n$ that does not carry any semantic meaning.
@@
 If we increase the \acs{wl}-dimension to $k = 2$, the described 1-\acs{wl} restrictions no longer apply.
-2-\acs{wl} is able to distinguish more than $1 - \mathcal{O}(\nicefrac{1}{n})$ of the regular $n$-vertex graphs~\cite[cor.~1.8.6]{Immerman1990} and can even count cycles.
+2-\acs{wl} is able to distinguish more than $1 - \mathcal{O}(\nicefrac{1}{n})$ of the regular $n$-vertex graphs~\cite[cor.~1.8.6]{Immerman1990} and can even count cycles:
 \begin{prop}[see \citet{Fuerer2017} and \citet{Arvind2019}]\label{prop:related:wl2-cycle-count}
@@
 	The number of 4-cycles is then determined by the colors of the diagonals through them via $g(\mathit{dist}_\chi) = \frac{1}{2} \sum_{j \geq 2} \binom{j}{2} \sum_{c \in C^{\square}_j} \mathit{dist}_\chi(c)$.
-	Using a similar but more involved combinatorial argument requiring multiple color refinement steps, 5-, 6- and 7-cycle counting can be shown.
+	Using a similar but more involved combinatorial argument requiring multiple color refinement steps, 5-, 6- \& 7-cycle counting can be shown.
 \end{hproof}
 As we just saw, 2-\acs{wl} is significantly more powerful than 1-\acs{wl}.
-By \cref{prop:related:wl-upper} there are of course still 2-\acs{wl} indistinguishable graphs;
+Though, by \cref{prop:related:wl-upper}, there are of course still 2-\acs{wl} indistinguishable graphs;
 among others those are the strongly regular graphs $\mathit{srg}_{n, d, \lambda, \mu}(G)$.
@@
 
-An alternative perspective on graph characterization is provided by \textit{spectral graph theory}.
+Let us now look at an alternative perspective on graph characterization which is provided by \textit{spectral graph theory}.
+While \ac{wl} characterizes a graph via its color distribution $\mathit{dist}_{\chi_{G,k}^{*}}$, the spectral approach uses its so-called \textit{spectrum} $\lambda_G = (\lambda_{G, 1}, \dots, \lambda_{G, n}) \in \mathbb{R}^n$ where $n = |\mathcal{V}_G|$.
+The key idea behind this is to interpret the adjacency matrix $A_G \in \mathbb{R}^{n \times n}$ of $G$ not simply as an encoding of the edges ($A_{G,i,j} = \mathbbm{1}[(v_i, v_j) \in \mathcal{E}_G]$) but as a linear operator $A_G: (\mathcal{V}_G \to \mathbb{R}) \to (\mathcal{V}_G \to \mathbb{R})$ acting on the vector space of real-valued functions with a vertex domain.
+This means that $A_G$ transforms so-called \textit{graph signals} $x: \mathcal{V}_G \to \mathbb{R}$ which are functions assigning \textit{signal strengths} to vertices.
+By applying $A_G x$ the signal strength $x[v_i]$ of each vertex is added to the signal strengths of its neighbors $\Gamma_G(v_i)$.
+Using this functional perspective, one can characterize graphs via the \acf{ft}.
+We will now see how this is done;
+however, since a comprehensive description of spectral graph theory would exceed the scope of this thesis, only a brief overview will be given.
+For a more detailed introduction to the field we refer to \citet{Shuman2013}.
+
+\paragraph{\ac{ft} on real domains}
+Let us begin by describing the \ac{ft} for the more common case of functions with real domains.
+All functions $f: \mathbb{R} \to \mathbb{R}$ can be interpreted as infinite-dimensional vectors $f = \int_{\mathbb{R}} f(t)\, b_t \dif t$ with $b_t: \mathbb{R} \to \mathbb{R}$ being a standard basis vector/function defined as $b_t(s) \coloneqq \begin{cases} 1 & \text{if } s = t \\[-6pt] 0 & \text{else} \end{cases}$.
+% chktex 21
+The values of $f$ are then described as its $b_t$ components $\langle b_t, f \rangle = f(t)$, where $\langle b_t, \cdot \rangle = \delta_t(\cdot)$ denotes the Dirac delta function translated by $t$.
+The \acl{ft} $\hat{f}$ of $f$ corresponds to a change of basis from the standard basis vectors $b_t$ to the Fourier basis vectors $u_\xi(t) \coloneqq e^{2 \pi i \xi t}$, i.e.\ $f = \int_{\mathbb{R}} \hat{f}(t)\, u_\xi \dif \xi$.
+The Fourier basis is characterized by the fact that it is an eigenbasis of the so-called \textit{Laplace operator} $\Delta$, i.e.\ $\Delta u_\xi = \lambda_\xi u_\xi$ with $\lambda_\xi$ being the eigenvalue corresponding to $u_\xi$.
+In the real domain this Laplace operator $\Delta$ is effectively the same as the second-derivative $\frac{\dif^2}{\dif t^2}$.
+Thus it is easy to see that indeed $\Delta u_\xi = \frac{\dif^2}{\dif t^2} u_\xi = \lambda_\xi u_\xi$ for $\lambda_\xi = -{(2 \pi \xi)}^2$.
+
+\paragraph{\ac{ft} on vertex domains}
+To extend the notion of the \ac{ft} from real functions $f: \mathbb{R} \to \mathbb{R}$ to graph signals $x: \mathcal{V}_G \to \mathbb{R}$ we need to define a graph variant of the Laplace operator $\Delta$.
+It turns out that there are multiple possible ways to do so, the most simple being the so-called \textit{combinatorial graph Laplacian}.
+\begin{defn}
+	The combinatorial graph Laplacian $L_G \in \mathbb{R}^{n \times n}$ is defined as
+	\begin{align*}
+		L_G \coloneqq D_G - A_G\quad\text{with the degree matrix } D_G\coloneqq {
+			\renewcommand*{\arraystretch}{0.5}
+			\begin{pmatrix}
+				d_1 & & \\
+				& \ddots & \\
+				& & d_n
+			\end{pmatrix}
+		}, d_i \coloneqq |\Gamma_G(v_i)| \text{.}
+	\end{align*}
+\end{defn}
+Using this definition, applying $L_G x$ is analogous to taking the second derivative $\frac{\dif^2}{\dif t^2} f$.
+Putting both Laplacian variants, $L_G$ and $\frac{\dif^2}{\dif t^2}$, side-by-side gives an intuition for why this is the case:
+\begin{equation*}
+	\begin{split}
+		-\frac{\dif^2}{\dif t^2} f(t) = \smashoperator[l]{\lim_{h \to 0}} \frac{1}{h^2} (\underbrace{f(t) - f(t-h)}_{\Delta_{t, t-h}} + \underbrace{f(t) -  f(t+h)}_{\Delta_{t, t+h}})
+	\end{split}
+	\begin{split}\ \left|\ %
+		L_G x[v] = \smashoperator[l]{\sum_{u \in \Gamma_G(v)}}\mkern-6mu \underbrace{(x[v] - x[u])}_{\Delta_{v, u}}
+	\right.\end{split}
+\end{equation*}
+The second derivative of a function $f$ essentially averages the value differences in the neighborhood of a point $t$.
+For real-valued functions this neighborhood only consists of the two infinitesimally close points to the left and to the right of $t$, i.e.\ $t - h$ and $t + h$.
+The combinatorial graph Laplacian represents the same operation, where each point/vertex $v$ might however have more than two neighbors $u \in \Gamma_G(v)$ that need to be averaged.
+
+The \ac{ft} of a graph signal $x$ then is $\hat{x}[i] = \langle u_{G,i}, x \rangle$ with $u_G = {\{ u_{G,i} \}}_{i=1}^{n}$ being the eigenvectors of $L_G$, s.t.\ $x = \sum_{i=1}^n \hat{x}[i] u_{G,i}$.
+Note that $L_G$ is guaranteed to only have real eigenvectors $u_G$ and eigenvalues $\lambda_G$ due to the fact that we only consider undirected graphs with symmetric adjacency matrices $A_G$.
+By convention we assume that the eigenvalues are ordered ascendingly: $\lambda_{G,0} \leq \cdots \leq \lambda_{G,n}$.
+Those eigenvalues are called the \textit{spectrum of $G$} and they are a vertex-permutation-invariant graph characterization.
+Intuitively those eigenvalues describe the connectivity between different parts of the graph.
+The initial eigenvalues represent connectivity at a general, coarse level while the last eigenvalues represent the connectivity between finer substructures.
+\Cref{fig:related:graph-fourier} illustrates this idea.
+A more concrete example for the relation between a graph's structure and its spectrum is that a graph with $m$ connected components has exactly $m$ zero eigenvalues, i.e.\ $\lambda_{G,1} \cdots = \lambda_{G,m} = 0$.
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=\linewidth]{gfx/related-work/graph-fourier.pdf}
+	\caption[Comparison between the basis vectors of the real domain \acl{ft} and the graph \acl{ft}.]{
+		Comparison between the basis functions/vectors of the real domain \ac{ft} and the graph \ac{ft}.
+		For the eigenfunctions on the upper half only the real cosine components of the complex exponentials are shown.
+	}\label{fig:related:graph-fourier}
+\end{figure}
 

### 2020-02-29 thesis/content/chapter-related-work.tex
fixed font issues

@@
 \begin{defn}\label{defn:related:wl-distinguishable}
-	Two graphs $G$ and $H$ are are $k$-\acs{wl} \textit{distinguishable} ($G \mathrel{{\not\simeq}_k} H$) iff.\ there exists a color $c \in \mathcal{C}$ s.t.\ $\mathit{dist}_{\chi^{*}_{G, k}}(c) \neq \mathit{dist}_{\chi^{*}_{H, k}}(c)$.
+	Two graphs $G$ and $H$ are $k$-\acs{wl} \textit{distinguishable} ($G \mathrel{{\not\simeq}_k} H$) iff.\ there exists a color $c \in \mathcal{C}$ s.t.\ $\mathit{dist}_{\chi^{*}_{G, k}}(c) \neq \mathit{dist}_{\chi^{*}_{H, k}}(c)$.
 \end{defn}
@@
 \begin{defn}
-	A graph $G$ is called $d$-regular ($\mathit{rg}_d(G)$) iff.\ $\forall v \in \mathcal{V}_G: {|\Gamma_G(v)|} = d$.
+	A graph $G$ is called \textit{$d$-regular} ($\mathit{rg}_d(G)$) iff.\ $\forall v \in \mathcal{V}_G: {|\Gamma_G(v)|} = d$.
 \end{defn}
@@
 \begin{defn}\label{defn:related:wl-compute}
-	$k$-\acs{wl} computes a function $f: \mathcal{G} \to Y$ iff.\ that function can be expressed as $f(G) = g(\mathit{dist}_{\chi_{G, k}^{*}})$ via some function $g: (\mathcal{C} \to \mathbb{N}) \to Y$.
+	$k$-\acs{wl} \textit{computes} a function $f: \mathcal{G} \to Y$ iff.\ that function can be expressed as $f(G) = g(\mathit{dist}_{\chi_{G, k}^{*}})$ via some function $g: (\mathcal{C} \to \mathbb{N}) \to Y$.
 \end{defn}
 \begin{defn}\label{defn:related:wl-count-detect}
-	$k$-\acs{wl} counts a certain subgraph $S$ iff.\ it computes $\mathit{count}_S(G) \coloneqq \left|\{ \hat{v}\, |\, v \in \mathcal{V}_G^{*} \land G[v] \equiv S \}\right|$ (see~\cref{defn:related:ordered-subgraph}).
-	Similarly $k$-\acs{wl} detects a subgraph $S$ iff.\ it computes $\mathit{contain}_S(G) \coloneqq \mathbbm{1}[\mathit{count}_S(G) > 0]$, with $\mathbbm{1}$ denoting the indicator function.
+	$k$-\acs{wl} \textit{counts} a certain subgraph $S$ iff.\ it computes $\mathit{count}_S(G) \coloneqq \left|\{ \hat{v}\, |\, v \in \mathcal{V}_G^{*} \land G[v] \equiv S \}\right|$ (see~\cref{defn:related:ordered-subgraph}).
+	Similarly $k$-\acs{wl} \textit{detects} a subgraph $S$ iff.\ it computes $\mathit{contain}_S(G) \coloneqq \mathbbm{1}[\mathit{count}_S(G) > 0]$, with $\mathbbm{1}$ denoting the indicator function.
 \end{defn}
@@
 		For the eigenfunctions on the upper half only the real cosine components of the complex exponentials are shown.
+		\source[based on]{Shuman2013}
 	}\label{fig:related:graph-fourier}

### 2020-02-29 thesis/content/chapter-related-work.tex
first sgt overview

@@
 	\centering
-	\includegraphics[width=0.75\linewidth]{gfx/related-work/bum.pdf}
+	\includegraphics[width=0.72\linewidth]{gfx/related-work/bum.pdf}
 	\caption[Illustration of how a \ac{bum} function is described as a linear spline and its relation to the \ac{owa} weights.]{
@@
 
-\paragraph{\ac{ft} on real domains}
+\paragraph{The classical \acl{ft}}
 Let us begin by describing the \ac{ft} for the more common case of functions with real domains.
@@
 In the real domain this Laplace operator $\Delta$ is effectively the same as the second-derivative $\frac{\dif^2}{\dif t^2}$.
-Thus it is easy to see that indeed $\Delta u_\xi = \frac{\dif^2}{\dif t^2} u_\xi = \lambda_\xi u_\xi$ for $\lambda_\xi = -{(2 \pi \xi)}^2$.
+Thus it is easy to see that indeed $\Delta u_\xi = -\frac{\dif^2}{\dif t^2} u_\xi = \lambda_\xi u_\xi$ for $\lambda_\xi = {(2 \pi \xi)}^2$.
 
-\paragraph{\ac{ft} on vertex domains}
+\paragraph{The graph \acl{ft}}
 To extend the notion of the \ac{ft} from real functions $f: \mathbb{R} \to \mathbb{R}$ to graph signals $x: \mathcal{V}_G \to \mathbb{R}$ we need to define a graph variant of the Laplace operator $\Delta$.
@@
 Intuitively those eigenvalues describe the connectivity between different parts of the graph.
-The initial eigenvalues represent connectivity at a general, coarse level while the last eigenvalues represent the connectivity between finer substructures.
+The first eigenvalues represent connectivity at a general, coarse level while the last eigenvalues represent the connectivity of finer substructures.
 \Cref{fig:related:graph-fourier} illustrates this idea.
-A more concrete example for the relation between a graph's structure and its spectrum is that a graph with $m$ connected components has exactly $m$ zero eigenvalues, i.e.\ $\lambda_{G,1} \cdots = \lambda_{G,m} = 0$.
+A concrete example for the relation between a graph's structure and its spectrum is the fact that a graph with $m$ connected components has exactly $m$ zero eigenvalues, i.e.\ $0 = \lambda_{G,1} = \cdots = \lambda_{G,m} < \lambda_{G,m+1}$.
+For a more detailed discussion of this relation we refer to \citet{Das2004}.
 \begin{figure}[ht]
@@
 \end{figure}
+\begin{defn}
+	Two graphs $G$ and $H$ are \textit{cospectral} ($G \mathrel{\simeq_\lambda} H$) iff.\ $\forall i: \lambda_{G,i} = \lambda_{H,i}$.
+\end{defn}
+\citet{Alzaga2010} have shown that the cospectrality test can be more powerful than 1-\acs{wl} but is always weaker than 2-\acs{wl}, i.e.\ $G \mathrel{\simeq_2} H \implies G \mathrel{\simeq_\lambda} H$.
+Despite this limit on the discriminative power of the graph spectrum, it is still useful for determining graph similarity, e.g.\ by defining a distance measure on the vector space of spectra~\cite{Gu2015}.
+
+As briefly mentioned before, the combinatorial graph Laplacian $L_G$ is not the only possible choice of graph Laplacian.
+Note that the eigenvalues of $L_G$ grow with the number of edges in $G$\footnote{
+	This is trivially shown by $\sum_{i=1}^n \lambda_{G,i} = \Tr(L_G) = \Tr(D_G) = |\mathcal{E}_G|$.
+}.
+Consequently a small graph $G$ will always have a large spectral distance to a large graph $H$ (assuming $|\mathcal{V}_G| \ll |\mathcal{V}_H|$) irrespective of their similarity when ignoring the scale difference.
+In domains where the absolute size of a graph should not influence its spectral characterization, it can therefore be useful to normalize the spectrum.
+One common way to do so is via the so-called \textit{symmetric normalized Laplacian} $L^{\text{sym}}_G$.
+\begin{defn}
+	$L^{\text{sym}}_G \coloneqq D^{-\frac{1}{2}} L_G D^{-\frac{1}{2}}$ is the symmetric normalized Laplacian of $G$.
+\end{defn}
+The eigenvalues $\lambda^{\text{sym}}_{G, i}$  of this Laplacian all lie in the range $[0, 1]$.
+There are also other ways to normalize the spectrum, e.g.\ by using the random walk Laplacian $L^{\text{rw}}_G \coloneqq D^{-1} L_G$, which we will however not cover here (see \citet{Shuman2013}).
+This concludes our overview of spectral graph characterization.
 
@@
 
-The basic idea of explicit graph embedding approaches is to map a graph $G \in \mathbb{G}$ to some vector in a finite vector space $\mathcal{X} = \mathbb{R}^d$.
-A function $f: \mathbb{G} \to \mathcal{X}$ is called a \textit{graph embedding function}.
+The basic idea of explicit graph embedding approaches is to map a graph $G \in \mathcal{G}$ to some vector in a finite vector space $\mathcal{X} = \mathbb{R}^d$.
+A function $f: \mathcal{G} \to \mathcal{X}$ is called a \textit{graph embedding function}.
 By embedding a graph into $\mathcal{X}$, any classification or regression algorithm that works with vectors can then be applied to solve the \ac{gcr} problem.
@@
 The so-called vertex embedding problem is closely related to the graph embedding problem.
-As the name suggests, a \textit{vertex embedding function} $f_G: V \to \mathcal{X}$ maps all vertices $v \in V$ of a graph $G$ into $\mathcal{X}$.
+As the name suggests, a \textit{vertex embedding function} $f_G$ maps all vertices $v \in \mathcal{V}_G$ to $\mathcal{X}$.
 The embedding vector $f_G(v)$ ideally encodes relevant information about a vertex and its structural position in $G$.
@@
 It is inspired by \texttt{doc2vec}~\cite{Le2014} which in turn is based on \texttt{word2vec}.
-\texttt{graph2vec} gets a set of graphs $\mathbb{G} = \{ G_1, \dots G_N \}$ as input and outputs graph embeddings $f(G_1), \dots, f(G_N)$.
-While in \texttt{word2vec} every word can be a context root as well as a context element, \texttt{graph2vec} uses the graphs $\mathbb{G}$ as context roots.
+\texttt{graph2vec} gets a set of graphs $\mathcal{G} = \{ G_1, \dots G_N \}$ as input and outputs graph embeddings $f(G_1), \dots, f(G_N)$.
+While in \texttt{word2vec} every word can be a context root as well as a context element, \texttt{graph2vec} uses the graphs $\mathcal{G}$ as context roots.
 The context $C_k(G_i)$ of a graph $G_i$ is defined as
@@
 	C_k(G_i) \coloneqq \bigcup_{\mathclap{v_j \in \mathcal{V}_{G_i}}}\, {\left\{ \chi_{G_i, 1}^{(l)}(v_j) \right\}}_{l = 0}^k % chktex 21
-	\text{ with $k \in \mathbb{N}$ and $\chi_{G_i, 1}^{(l)}$ as in \cref{defn:related:wl1-refine}.}
+	\quad\text{with $k \in \mathbb{N}$ and $\chi_{G_i, 1}^{(l)}$ as in \cref{defn:related:wl1-refine}.}
 \end{equation}
@@
 Since \ac{wl} is used to identify distinct subgraphs, \texttt{graph2vec} can only be applied to graphs with discrete vertex labels.
-Using the previous definitions, the context root embedding function has the signature $f: \mathbb{G} \to \mathbb{R}^d$ while the context element embedding function is of type $f_C: \bigcup_{G_i \in \mathbb{G}} C_k(G_i) \to \mathbb{R}^d$.
+Using the previous definitions, the context root embedding function has the signature $f: \mathcal{G} \to \mathbb{R}^d$ while the context element embedding function is of type $f_C: \bigcup_{G_i \in \mathcal{G}} C_k(G_i) \to \mathbb{R}^d$.
 To find those embeddings the \texttt{word2vec} objective from \cref{eq:related:gcr:embed:word2vec} is reused.

### 2020-02-29 implementation/src/ltag/evaluation/evaluate.py

@@ -39,7 +39,7 @@ def train(
   t = time_str()
   tb = keras.callbacks.TensorBoard(
     log_dir=log_dir_base / f"{t}{label}/",
-    histogram_freq=1,
+    histogram_freq=10,
     write_images=False)
   es = keras.callbacks.EarlyStopping(
     monitor="loss" if val_ds is None else "val_loss",

### 2020-03-01 thesis/content/chapter-related-work.tex
added sgt normalization illustration + started graph kernel description

@@
 		for simplicity neighbors with $u \in v$ are left out (see~\cref{defn:related:wlk-refine}).
-		Each neighbor is highlighted with a different color, except for 3-WL where the \textcolor{t_red}{red}, \textcolor{t_green}{green} and \textcolor{t_darkyellow}{yellow} triples actually form the single neighbor for $u = v_4$.
+		Each neighbor is highlighted with a different color, except for 3-WL where the \textcolor{t_red}{red}, \textcolor{t_darkgreen}{green} and \textcolor{t_darkyellow}{yellow} triples actually form the single neighbor for $u = v_4$.
 	}\label{fig:related:wl-neighbors}
@@
 \end{defn}
-\citet{Alzaga2010} have shown that the cospectrality test can be more powerful than 1-\acs{wl} but is always weaker than 2-\acs{wl}, i.e.\ $G \mathrel{\simeq_2} H \implies G \mathrel{\simeq_\lambda} H$.
-Despite this limit on the discriminative power of the graph spectrum, it is still useful for determining graph similarity, e.g.\ by defining a distance measure on the vector space of spectra~\cite{Gu2015}.
+\citet{Alzaga2010} have shown that, while the cospectrality test can be more powerful than 1-\acs{wl}, it is always weaker than 2-\acs{wl}, i.e.\ $G \mathrel{\simeq_2} H \implies G \mathrel{\simeq_\lambda} H$.
+Despite this limit on the discriminative power of the graph spectrum, it is still useful for determining graph similarity, e.g.\ by defining a distance measure on the vector space of spectra (see \citet{Gu2015}).
 
@@
 }.
-Consequently a small graph $G$ will always have a large spectral distance to a large graph $H$ (assuming $|\mathcal{V}_G| \ll |\mathcal{V}_H|$) irrespective of their similarity when ignoring the scale difference.
+Consequently a small graph $G$ will typically have a large spectral distance to a large graph $H$ (assuming ${|\mathcal{V}_G|} \ll {|\mathcal{V}_H|}$) irrespective of their similarity when ignoring the scale difference.
 In domains where the absolute size of a graph should not influence its spectral characterization, it can therefore be useful to normalize the spectrum.
@@
 \end{defn}
-The eigenvalues $\lambda^{\text{sym}}_{G, i}$  of this Laplacian all lie in the range $[0, 1]$.
-There are also other ways to normalize the spectrum, e.g.\ by using the random walk Laplacian $L^{\text{rw}}_G \coloneqq D^{-1} L_G$, which we will however not cover here (see \citet{Shuman2013}).
-This concludes our overview of spectral graph characterization.
+The eigenvalues $\lambda^{\text{sym}}_{G, i}$  of this Laplacian all lie in the range $[0, 2]$.
+\Cref{fig:related:spectral-normalization} illustrates how this makes it possible to compare the structure of graphs across varying vertex counts.
+Besides $L^{\text{sym}}_G$ there are also other ways to normalize the spectrum, e.g.\ by using the random walk Laplacian $L^{\text{rw}}_G \coloneqq D^{-1} L_G$, which will however not be covered here (see \citet{Shuman2013}).
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=\linewidth]{gfx/related-work/spectral-normalization.pdf}
+	\caption[Comparison of the unnormalized $L_G$ spectrum and the normalized $L^{\text{sym}}_G$ spectrum.]{
+		Comparison of the unnormalized $L_G$ spectrum and the normalized $L^{\text{sym}}_G$ spectrum.
+		Two structurally similar graphs of different size are compared.
+		The \textcolor{t_red}{large graph} is derived by replacing each vertex of the \textcolor{t_blue}{small 4-cycle graph} by a triangle.
+		While their unnormalized spectra have different absolute ranges (\textcolor{t_blue}{$[0, 4]$} vs.\ \textcolor{t_red}{$[0, 12]$}), the histogram on the right shows that their normalized eigenvalue distributions align.
+	}\label{fig:related:spectral-normalization}
+\end{figure}
 
@@
 \begin{equation}
-	C_k(G_i) \coloneqq \bigcup_{\mathclap{v_j \in \mathcal{V}_{G_i}}}\, {\left\{ \chi_{G_i, 1}^{(l)}(v_j) \right\}}_{l = 0}^k % chktex 21
+	C_k(G_i) \coloneqq \bigcup_{\mathclap{v_j \in \mathcal{V}_{G_i}}}\, {\left\{ \chi_{G_i, 1}^{(l)}(v_j) \right\}}_{l = 0}^L % chktex 21
 	\quad\text{with $k \in \mathbb{N}$ and $\chi_{G_i, 1}^{(l)}$ as in \cref{defn:related:wl1-refine}.}
 \end{equation}
-Intuitively this context can be understood as the set of 1-\acs{wl}-distinguishable subgraphs of $G_i$ with diameter $\leq 2k$.
+Intuitively this context can be understood as the set of 1-\acs{wl}-distinguishable subgraphs of $G_i$ with diameter $\leq 2L$.
 Since \ac{wl} is used to identify distinct subgraphs, \texttt{graph2vec} can only be applied to graphs with discrete vertex labels.
@@
 
-Instead of explicitly mapping a graph or vertices into a vector space, one can also do so implicitly by employing the kernel trick.
+Instead of mapping a graph into an explicitly defined vector space of fixed dimension, one can also do so implicitly by employing the kernel trick.
 There is a large variety of so-called \acp{gk} to do this.
@@
 \paragraph{\ac{wl} subtree kernel}
+One well-known \ac{gk} is based directly on the 1-\acs{wl} coloring algorithm, the so-called \textit{\ac{wl} subtree kernel}~\cite{Shervashidze2011}.
+It is uses the mapping
+\begin{align}
+	\varphi_{\text{ST}}(G) \coloneqq \bigoplus_{l=0}^{L} {\left( \mathit{dist}_{\chi_{G,1}^{(l)}}(c) \right)}_{c \in \mathcal{C}}
+	\quad\text{with $\oplus$ denoting vector concatenation.}
+\end{align}
+$\varphi_{\text{ST}}(G)$ encodes the color counts across a fixed number $L$ of 1-\acs{wl} refinement steps.
+Via this mapping, the \ac{wl} subtree kernel function $k_{\text{ST}}: \mathcal{G} \times \mathcal{G} \to \mathbb{R}$ can then be simply written as the standard inner vector product $k_{\text{ST}}(G, H) = \langle \varphi_{\text{ST}}(G), \varphi_{\text{ST}}(H) \rangle$.
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=0.7\linewidth]{gfx/related-work/wl-subtree.pdf}
+	\caption[Illustration of the correspondence between the \ac{wl} color $\chi_{G,1}^{(l)}(v)$ and the breadth-first subtree of depth $l$ rooted at $v$.]{
+		Illustration of the correspondence between the \ac{wl} color $\chi_{G,1}^{(l)}(v)$ and the breadth-first subtree of depth $l$ rooted at $v$ (for $l = 2$ and $v = \texttt{\textcolor{t_blue}{\circled{1}}}$).
+% chktex 25
+		\source[based~on]{Shervashidze2011}
+	}\label{fig:related:wl-subtree}
+\end{figure}
+
+\Cref{fig:related:wl-subtree} illustrates how this definition relates to subtrees.
+Since alls vertex colors $c \in \mathcal{C}$ correspond to a subtree isomorphism classes, $k_{\text{ST}}(G, H)$ effectively computes the similarity of $G$ and $H$ by comparing their local subtree structures of depth at most $L$.
+Note that the underlying mapping $\varphi_{\text{ST}}$ cannot be computed independently for each graph in a given training dataset since the set of colors $c \in \mathcal{C}$ introduced by the hashing function $h$ is conjointly determined by all graphs in the dataset (see \cref{defn:related:wl1-refine}).
+Consequently the dimensionality $d$ of $\varphi_{\text{ST}}(G) \in \mathbb{N}^d$ varies depending on the other training graphs.
 
 \paragraph{\ac{wl} shortest path kernel}
+Extending the idea of the \ac{wl} subtree kernel, the \textit{\ac{wl} shortest path kernel}~\cite{Shervashidze2011} adds global structural information to the local subtree comparisons.
 
@@
 \label{sec:related:gcr:nn}
-
-\subsubsection{Spatial GNNs}
-
-\subsubsection{Spectral GNNs}

### 2020-03-02 thesis/content/chapter-related-work.tex
completed gk section

@@
 \end{defn}
-In practice the hash function $h$ is usually defined lazily by using $\mathcal{C} = \mathbb{N}$ and enumerating color combinations in whichever order they are hashed s.t.\ a new color is introduced every time a previously unseen color combination appears at runtime.
-This is illustrated in \cref{fig:related:wl1-refine}.
 \begin{figure}[ht]
@@
 \end{figure}
+In practice the hash function $h$ is usually defined lazily by using $\mathcal{C} \subseteq \mathbb{N}$ and enumerating color combinations in whichever order they are hashed s.t.\ a new color is introduced every time a previously unseen color combination appears at runtime\footnote{
+	Note that, even though $\mathcal{C}$ is formally countably infinite, we will assume that it only contains the finite number of colors occurring in a given finite graph dataset.
+}.
 
@@
 	}\label{fig:related:wl1-problem}
-\end{figure}
+\end{figure}\\
 By extending \ac{wl} to higher dimensions, such 1-\acs{wl} indistinguishable cases can however be handled.
@@
 	$\begin{aligned}[t]
-		\chi_{G,k}^{(i+1)}(v) \coloneqq h\left(\chi_{G,k}^{(i)}(v), \ldblbrace (\chi_{G,k}^{(i)}(v[u/1]), \dots, \chi_{G,k}^{(i)}(v[u/k]))\, |\, u \in \mathcal{V}_G \rdblbrace\right)\\
-		\text{with } v = (v_1, \dots, v_k) \in \mathcal{V}_G^k,\quad v[u/j] \coloneqq (v_1, \dots, v_{j-1}, u, v_{j+1}, \dots, v_k) \text{.}
+		\chi_{G,k}^{(i+1)}(s) \coloneqq h\left(\chi_{G,k}^{(i)}(s), \ldblbrace (\chi_{G,k}^{(i)}(s[u/1]), \dots, \chi_{G,k}^{(i)}(s[u/k]))\, |\, u \in \mathcal{V}_G \rdblbrace\right)\\
+		\text{with } s = (v_1, \dots, v_k) \in \mathcal{V}_G^k,\quad s[u/j] \coloneqq (v_1, \dots, v_{j-1}, u, v_{j+1}, \dots, v_k) \text{.}
 	\end{aligned}$
@@
 In 1-\acs{wl} a vertex color is refined by combining the colors of neighboring vertices.
-In $k$-\acs{wl} the color of a $k$-tuple $v \in \mathcal{V}_G^k$ is refined by combining the colors of its neighborhood which is defined as the set of all $k$-tuples in which at most one vertex differs from $v$.
+In $k$-\acs{wl} the color of a $k$-tuple $s \in \mathcal{V}_G^k$ is refined by combining the colors of its neighborhood which is defined as the set of all $k$-tuples in which at most one vertex differs from $s$.
 Note that each vertex $k$-tuple has one neighbor for each $u \in \mathcal{V}_G$, each of which is a $k$-tuple of vertex $k$-tuples.
@@
 		Tuple neighborhoods for different values of $k$.
-		The vertices highlighted in \textcolor{t_blue}{blue} form the root tuple $v \in \mathcal{V}_G^k$ whose neighbors are shown;
-		for simplicity neighbors with $u \in v$ are left out (see~\cref{defn:related:wlk-refine}).
+		The vertices highlighted in \textcolor{t_blue}{blue} form the root tuple $s \in \mathcal{V}_G^k$ whose neighbors are shown;
+		for simplicity neighbors with $u \in s$ are left out (see~\cref{defn:related:wlk-refine}).
 		Each neighbor is highlighted with a different color, except for 3-WL where the \textcolor{t_red}{red}, \textcolor{t_darkgreen}{green} and \textcolor{t_darkyellow}{yellow} triples actually form the single neighbor for $u = v_4$.
@@
 Let us now look at how the tuple colors are initialized.
-For this we use the ordered subgraph equivalence classes $[G[v]]$ (see~\cref{defn:related:ordered-subgraph-equivalence}) which determine the initial color $\chi_{G,k}^{(0)}(v)$ of each $k$-tuple $v$.
+For this we use the ordered subgraph equivalence classes $[G[s]]$ (see~\cref{defn:related:ordered-subgraph-equivalence}) which determine the initial color $\chi_{G,k}^{(0)}(v)$ of each $k$-tuple $s$.
 For $k = 1$ the equivalence class of a single vertex $v$ directly corresponds to its label $l_G[v]$.
@@
 \begin{align}
-	\chi_{G,k}^{(0)}(v) = \chi_{G,k}^{(0)}(w) \iff G[v] \equiv G[w] \text{.} \label{eq:related:wlk-init}
+	\chi_{G,k}^{(0)}(s) = \chi_{G,k}^{(0)}(t) \iff G[s] \equiv G[t] \text{.} \label{eq:related:wlk-init}
 \end{align}
@@
 	All $k$-tuples can be mapped to $(k+1)$-tuples via $\varphi(v_1, \dots, v_k) \coloneqq (v_1, \dots,\allowbreak v_k, v_k)$.
-	For each neighbor $(v[u/1], \dots,\allowbreak v[u/k])$ of $v$, there is a corresponding neighbor $(\varphi(v[u/1]), \dots \varphi(v[u/k]), \varphi(v[u/k]))$ of $\varphi(v)$.
-	Using \cref{eq:related:wlk-init} (for $i = 0$) and \cref{defn:related:wlk-refine} (for $i > 0$) it follows that $\forall v, w \in \mathcal{V}_G^k: \chi_{G, k}^{(i)}(v) \neq \chi_{G, k}^{(i)}(w) \rightarrow \chi_{G, k+1}^{(i)}(\varphi(v)) \neq \chi_{G, k+1}^{(i)}(\varphi(w))$.
+	For each neighbor $(s[u/1], \dots,\allowbreak s[u/k])$ of $s \in \mathcal{V}_G^k$, there is a corresponding neighbor $(\varphi(s[u/1]), \dots \varphi(s[u/k]), \varphi(s[u/k]))$ of $\varphi(s)$.
+	Using \cref{eq:related:wlk-init} (for $i = 0$) and \cref{defn:related:wlk-refine} (for $i > 0$) it follows that $\forall s, t \in \mathcal{V}_G^k: \chi_{G, k}^{(i)}(s) \neq \chi_{G, k}^{(i)}(t) \rightarrow \chi_{G, k+1}^{(i)}(\varphi(s)) \neq \chi_{G, k+1}^{(i)}(\varphi(t))$.
 	The lemma then follows by \cref{defn:related:wl-distinguishable}.
 \end{hproof}
-\begin{prop}[see~\citet{Immerman1990}]\label{prop:related:wl-upper}
+\begin{prop}[see~\citet{Immerman1990} for the proof]\label{prop:related:wl-upper}
 	For all $k \in \mathbb{N}$ there are non-isomorphic graphs $G \not\simeq H$ of size $\mathcal{O}(k)$ with $G \mathrel{\simeq_k} H$.
@@
 
-This cycle-detection restriction of 1-\acs{wl} is relevant in practice because cycle counts are used in many domains to analyze graphs, e.g.\ triangle counts are commonly used in social network analysis to find interaction clusters~\cite{Newman2003}\cite{Welser2007} and the detection of 4-, 5- or 6-cycles is required to determine important chemical properties like the aromaticity of a molecule~\cite{Adamson1973}\cite{Kekule1866}.
+This cycle-detection restriction of 1-\acs{wl} is relevant in practice because cycle counts are used in many domains to analyze graphs, e.g.\ triangle counts are commonly used in social network analysis to find interaction clusters~\cite{Milo2002}\cite{Newman2003}\cite{Welser2007} and the detection of 4-, 5- or 6-cycles is required to determine important chemical properties like the aromaticity of a molecule~\cite{Adamson1973}\cite{Kekule1866}.
 
@@
 2-\acs{wl} is able to distinguish more than $1 - \mathcal{O}(\nicefrac{1}{n})$ of the regular $n$-vertex graphs~\cite[cor.~1.8.6]{Immerman1990} and can even count cycles:
-\begin{prop}[see \citet{Fuerer2017} and \citet{Arvind2019}]\label{prop:related:wl2-cycle-count}
+\begin{prop}[see \citet{Fuerer2017} and \citet{Arvind2019} for the full proof]\label{prop:related:wl2-cycle-count}
 	2-\acs{wl} is able to count $m$-cycles for $m \leq 7$ but it cannot detect 8-cycles.
@@
 As briefly mentioned before, the combinatorial graph Laplacian $L_G$ is not the only possible choice of graph Laplacian.
-Note that the eigenvalues of $L_G$ grow with the number of edges in $G$\footnote{
+Note that the eigenvalues of $L_G$ grow with the number of edges\footnote{
 	This is trivially shown by $\sum_{i=1}^n \lambda_{G,i} = \Tr(L_G) = \Tr(D_G) = |\mathcal{E}_G|$.
@@
 \begin{gather}
-	\max_{f, f_C} \sum_{i = 1}^n \log P(C_k(w_i) | w_i)\ =\ \max_{f, f_C} \sum_{i = 1}^n \smashoperator[r]{\sum_{w_j \in C_k(w_i)}}\, \left[ {f(w_i)}^\top f_C(w_j) - \log Z_{w_i} \right] \label{eq:related:gcr:embed:word2vec}\\
+	\max_{f, f_C} \sum_{i = 1}^n \log P(C_k(w_i) | w_i)\ =\ \max_{f, f_C} \sum_{i = 1}^n \smashoperator[r]{\sum_{w_j \in C_k(w_i)}}\, \left[ {f(w_i)}^\top f_C(w_j) - \log Z_{w_i} \right] \label{eq:related:word2vec}\\
 	\text{with } P(C_k(w_i) | w_i) \coloneqq \prod_{\mathclap{w_j \in C_k(w_i)}}\, \overbrace{\frac{\exp\left({f(w_i)}^\top f_C(w_j)\right)}{Z_{w_i}}}^{P(w_j | w_i)}
@@
 Using the previous definitions, the context root embedding function has the signature $f: \mathcal{G} \to \mathbb{R}^d$ while the context element embedding function is of type $f_C: \bigcup_{G_i \in \mathcal{G}} C_k(G_i) \to \mathbb{R}^d$.
-To find those embeddings the \texttt{word2vec} objective from \cref{eq:related:gcr:embed:word2vec} is reused.
+To find those embeddings the \texttt{word2vec} objective from \cref{eq:related:word2vec} is reused.
 Analogous to \texttt{word2vec}, \texttt{graph2vec} therefore embeds graphs that share subgraphs close to each other, whereas graphs that do not share substructures tend to be embedded further away from each other.
@@
 	\varphi_{\text{ST}}(G) \coloneqq \bigoplus_{l=0}^{L} {\left( \mathit{dist}_{\chi_{G,1}^{(l)}}(c) \right)}_{c \in \mathcal{C}}
-	\quad\text{with $\oplus$ denoting vector concatenation.}
+	\quad\text{with $\oplus$ denoting vector concatenation.} \label{eq:related:wl-subtree-kernel}
 \end{align}
 $\varphi_{\text{ST}}(G)$ encodes the color counts across a fixed number $L$ of 1-\acs{wl} refinement steps.
-Via this mapping, the \ac{wl} subtree kernel function $k_{\text{ST}}: \mathcal{G} \times \mathcal{G} \to \mathbb{R}$ can then be simply written as the standard inner vector product $k_{\text{ST}}(G, H) = \langle \varphi_{\text{ST}}(G), \varphi_{\text{ST}}(H) \rangle$.
+Via this mapping, the \ac{wl} subtree kernel function $k_{\text{ST}}: \mathcal{G} \times \mathcal{G} \to \mathbb{R}$ can then be simply written as the standard inner vector product $k_{\text{ST}}(G, H) \coloneqq \langle \varphi_{\text{ST}}(G), \varphi_{\text{ST}}(H) \rangle$.
 \begin{figure}[ht]
@@
 \Cref{fig:related:wl-subtree} illustrates how this definition relates to subtrees.
-Since alls vertex colors $c \in \mathcal{C}$ correspond to a subtree isomorphism classes, $k_{\text{ST}}(G, H)$ effectively computes the similarity of $G$ and $H$ by comparing their local subtree structures of depth at most $L$.
-Note that the underlying mapping $\varphi_{\text{ST}}$ cannot be computed independently for each graph in a given training dataset since the set of colors $c \in \mathcal{C}$ introduced by the hashing function $h$ is conjointly determined by all graphs in the dataset (see \cref{defn:related:wl1-refine}).
-Consequently the dimensionality $d$ of $\varphi_{\text{ST}}(G) \in \mathbb{N}^d$ varies depending on the other training graphs.
+Since alls vertex colors $c \in \mathcal{C}$ correspond to a subtree isomorphism class, $k_{\text{ST}}(G, H)$ effectively computes the similarity of $G$ and $H$ by comparing their local subtree structures of depth at most $L$.
+Note that the underlying mapping $\varphi_{\text{ST}}$ cannot be computed independently for each graph $G \in \mathcal{G}_{\mathcal{D}} \subseteq \mathcal{G}$ in a given training dataset $\mathcal{D}$ since the set of colors $c \in \mathcal{C}$ introduced by the hashing function $h$ is conjointly determined by all graphs $\mathcal{G}_{\mathcal{D}}$ (see \cref{defn:related:wl1-refine}).
+Consequently the dimensionality $d$ of $\varphi_{\text{ST}}(G) \in \mathbb{N}^d$ varies, depending on the entire dataset $\mathcal{D}$.
 
 \paragraph{\ac{wl} shortest path kernel}
-Extending the idea of the \ac{wl} subtree kernel, the \textit{\ac{wl} shortest path kernel}~\cite{Shervashidze2011} adds global structural information to the local subtree comparisons.
+Extending the idea of the \ac{wl} subtree kernel, the \textit{\ac{wl} shortest path kernel}~\cite{Shervashidze2011}\cite{Borgwardt2005} adds global structural information to the local subtree comparisons.
+The $i$-th component of a graph's shortest path vector $\varphi_{\text{SP}}(G)$ corresponds to the 4-tuple $(l_i, a_i, b_i, d_i) \in {\{ 0, \dots, L \}} \times \mathcal{C} \times \mathcal{C} \times \mathbb{N}$ and its value is described by
+\begin{align}
+	{\varphi_{\text{SP}}(G)}_i \coloneqq \left| \left\{ (v, u) \in \mathcal{V}_G^2 \,|\, {\chi_{G,1}^{(l_i)}(v) = a_i} \land {\chi_{G,1}^{(l_i)}(u) = b_i} \land {d_{\text{SP}}(v, u) = d_i} \right\} \right| \text{,} % chktex 21
+\end{align}
+with $d_{\text{SP}}(v, u)$ denoting the length of the shortest path from $v$ to $u$ in $G$.
+Analogous to $k_{\text{ST}}$, the shortest path kernel function is defined as $k_{\text{SP}}(G, H) \coloneqq \langle \varphi_{\text{SP}}(G), \varphi_{\text{SP}}(H) \rangle$.
+Since $\varphi_{\text{SP}}(G)$ has one component for each possible 4-tuple $(l_i, a_i, b_i, d_i)$, its dimensionality grows quadratically with the total number of introduced 1-\acs{wl} colors $\mathcal{C}$ and linearly with the length of the longest shortest path $\max_{G \in \mathcal{G}_{\mathcal{D}}, (v, u) \in \mathcal{V}_G^2} d_{\text{SP}}(v, u)$.
+While this approach is computationally significantly more expensive than the subtree kernel, it also has a higher discriminative power and allows the shortest path kernel to distinguish even some 1-\acs{wl} indistinguishable graphs.
+The subtree kernel only compares graphs by the presence of local subtrees;
+the shortest path kernel additionally checks how well the distances between those subtrees align.
+Looking back at the 1-\acs{wl} indistinguishable ``six-cycle''/``two-triangles'' example from \cref{fig:related:wl1-problem}, we see that this additional information distinguishes the two graphs due to the fact that the longest shortest path in a six-cycle has length 3 while the longest shortest path in a triangle is of length 1.
 
 \paragraph{Higher dimensional \ac{wl} kernels}
+Instead of incorporating shortest path information to increase the power of the \ac{wl} kernel, one can alternatively just increase the \acs{wl}-dimension $k$.
+A na{\"\i}ve generalization of the 1-\acs{wl} subtree kernel would simply use \cref{eq:related:wl-subtree-kernel} but with the k-\acs{wl} colorings $\chi_{G,k}^{(l)}$ instead of $\chi_{G,1}^{(l)}$.
+The problem with this approach is that the runtime of \ac{wl} increases exponentially with $k$;
+even for $k = 2$ the cost often becomes infeasibly high when working with large graphs.
+To tackle this problem \citet{Morris2017} proposed a combination of two optimizations:
+\begin{enumerate}[label=\textbf{\arabic*.}]
+	\item \textbf{$k$-multisets:}
+		The first optimization is to assign colors to vertex $k$-multisets instead of vertex $k$-tuples, where a $k$-multiset is any multiset $s \subseteq \mathcal{V}_G$ with ${|s|} = k$.
+		This reduces the amount of colors that have to be refined in each \ac{wl} iteration by a factor of $k!$. % chktex 40
+		Based on \cref{defn:related:wlk-refine}, the multiset color refinement step is defined as
+		\begin{align}
+			\chi_{G,k}^{(l+1)}(s) \coloneqq h\left(\chi_{G,k}^{(l)}(s), \ldblbrace \{ \chi_{G,k}^{(l)}(s \setminus \{v\} \cup \{u\})\, |\, v \in s \}\, |\, u \in \mathcal{V}_G \rdblbrace\right) \text{.} \label{eq:related:wlk-set-refine} % chktex 21
+		\end{align}
+		Even though this simplification generally reduces the discriminative power of the kernel, for $k = 2$ specifically no information and therefore no power is lost by replacing the tuples $(v, u)$ and $(u, v)$ with $\ldblbrace v, u \rdblbrace$ since the undirectedness of graphs implies that $\chi_{G,2}^{(l)}(v, u) = \chi_{G,2}^{(l)}(u, v)$ even when doing tuple-based color refinement.
+	\item \textbf{Neighborhood localization:}
+		The second optimization uses the fact that most real-world graphs tend to be sparse~\cite{Chung2010}, i.e.\ ${|\mathcal{E}_G|} = \mathcal{O}(|\mathcal{V}_G|)$.
+		The multiset-based $k$-\acs{wl} algorithm refines the color of a $k$-multiset $s$ by hashing the colors of its neighbors as defined in \cref{eq:related:wlk-set-refine} where there is one neighbor per vertex $u \in \mathcal{V}_G$.
+		By the sparsity assumption, most of those vertices $u$ are however not connected with any of the vertices in $s$, i.e.\ $u \notin \Gamma_G(s)$ with $\Gamma_G(s) \coloneqq \bigcup_{v \in s} \Gamma_G(v)$.
+		The refinement runtime can therefore often be reduced significantly by only considering ``local'' neighbors $u \in \Gamma_G(s)$ instead of the ``global'' neighborhood $u \in \mathcal{V}_G$ in \cref{eq:related:wlk-set-refine}.
+\end{enumerate}
+We call the kernel that only uses the first optimization \textit{$k$-GWL}~($k$-dim.\ global \ac{wl}) and the kernel that uses both optimizations \textit{$k$-LWL}~($k$-dim.\ local \ac{wl}).
+\citet{Morris2017} have empirically shown that focusing on local graph structures via LWL often actually performs better than the computationally more expensive GWL kernel.
 

### 2020-03-03 thesis/content/chapter-related-work.tex
minor fixes + start of gnn section

@@
 \acp{gk} can be used in combination with any kernelized learner, typically \acp{svm}, to solve the \ac{gcr} problem.
-While there is a large variety of different \acp{gk}, we will focus on those that are based on the previously described family of \ac{wl} algorithms.
+While there is a large variety of different \acp{gk}~\cite{Kriege2020}, we will focus on those that are based on the previously described family of \ac{wl} algorithms.
 
@@
 \label{sec:related:gcr:nn}
+
+The last family of \ac{gcr} approaches we will look at is that of \acp{gnn}.
+The idea to feed a graph into a \ac{nn} was first described by \citet{Gori2005}.
+Since then many variants and extensions of that idea have been proposed~\cite{Wu2019}.
+We will focus specifically on the so-called \acp{gcnn} which can be divided into two variants:
+Spectral and spatial \acp{gcnn}.
+
+\subsubsection{Spectral \acp{gcnn}}
+The class of spectral \acp{gcnn} is motivated by spectral graph theory (see \cref{sec:related:character:spectral}).
+
+\subsubsection{Spatial \acp{gcnn}}
+
+\subsubsection{Graph Pooling}

### 2020-03-04 thesis/content/chapter-related-work.tex
started with spectral gcnn section

@@
 The class of spectral \acp{gcnn} is motivated by spectral graph theory (see \cref{sec:related:character:spectral}).
+The spectral perspective on \ac{gcr} expects a graph $G$ with real vertex feature vectors $x_G[v_i] \in \mathbb{R}^f$ as its input.
+We call $X_G \coloneqq {(x_G[v_1], \dots, x_G[v_n])}^{\top} \in \mathbb{n \times f}$ the vertex feature matrix of $G$.
+Note that each of the $f$ columns of $X_G$ can be interpreted as a separate graph signal function $x_{G,j}: \mathcal{V}_G \to \mathbb{R}$.
+The core idea of spectral \acp{gcnn} is to learn a graph convolution kernel $g$, similar to the grid kernels in conventional \acp{cnn}~\cite{LeCun1998}.
+The problem with this idea is that the convolution operation $g * x$ requires some notion of distance in order to move the kernel $g$ over the function $x$.
+Graphs do not generally satisfy this requirement, i.e.\ the notion of a vertex distance $|v_i - v_j|$ is not clearly defined.
+Via the convolution theorem $\widehat{g * x} = \hat{g} \cdot \hat{f}$ we can however still define a graph convolution operator that works directly in the spectral domain:
+\begin{align}
+	g * x_{G,j} \coloneqq U \hat{g}(\Lambda) U^{\top} x_{G,j}
+	\text{ with the eigendecomposition $L_G = U \Lambda U^{\top}$}
+\end{align}
+To see the connection to the convolutional theorem, note that the matrix of eigenvector rows $U^{\top}$ describes the graph Fourier transform and conversely that $U$ describes the inverse graph Fourier transform.
+\citet{Bruna2013} proposed a \ac{nn} architecture that learns such a spectral filter $\hat{g}$.
 

### 2020-03-05 thesis/content/chapter-related-work.tex
completed first spectral gcnn section draft

--- 2020-03-03 thesis/content/chapter-related-work.tex
@@
 The values of $f$ are then described as its $b_t$ components $\langle b_t, f \rangle = f(t)$, where $\langle b_t, \cdot \rangle = \delta_t(\cdot)$ denotes the Dirac delta function translated by $t$.
-The \acl{ft} $\hat{f}$ of $f$ corresponds to a change of basis from the standard basis vectors $b_t$ to the Fourier basis vectors $u_\xi(t) \coloneqq e^{2 \pi i \xi t}$, i.e.\ $f = \int_{\mathbb{R}} \hat{f}(t)\, u_\xi \dif \xi$.
+The \acl{ft} $\hat{f} = \mathcal{F}(f)$ of $f$ corresponds to a change of basis from the standard basis vectors $b_t$ to the Fourier basis vectors $u_\xi(t) \coloneqq e^{2 \pi i \xi t}$, i.e.\ $f = \int_{\mathbb{R}} \hat{f}(t)\, u_\xi \dif \xi$.
 The Fourier basis is characterized by the fact that it is an eigenbasis of the so-called \textit{Laplace operator} $\Delta$, i.e.\ $\Delta u_\xi = \lambda_\xi u_\xi$ with $\lambda_\xi$ being the eigenvalue corresponding to $u_\xi$.
@@
 The \ac{ft} of a graph signal $x$ then is $\hat{x}[i] = \langle u_{G,i}, x \rangle$ with $u_G = {\{ u_{G,i} \}}_{i=1}^{n}$ being the eigenvectors of $L_G$, s.t.\ $x = \sum_{i=1}^n \hat{x}[i] u_{G,i}$.
-Note that $L_G$ is guaranteed to only have real eigenvectors $u_G$ and eigenvalues $\lambda_G$ due to the fact that we only consider undirected graphs with symmetric adjacency matrices $A_G$.
+Since the set of Laplacian eigenvectors is finite, we can express the graph \ac{ft} $\mathcal{F}_G$ as a change of basis matrix $U_G = \begin{pmatrix} u_{G,1,1} & \cdots & u_{G,1,n} \\ \vdots & \ddots & \vdots \\ u_{G,n,1} & \cdots & u_{G,n,n} \end{pmatrix} \in \mathbb{R}^{n \times n}$.
+Similarly the inverse \ac{ft} can be expressed as $\mathcal{F}^{-1} = U_G^{-1} = U_G^{\top}$ because $U_G$ is a real orthogonal matrix.
+To see why this is the case, note that $L_G$ is guaranteed to only have real eigenvectors $u_G$ and eigenvalues $\lambda_G$ due to the fact that we only consider undirected graphs with symmetric adjacency matrices $A_G$.
+
+\paragraph{Interpretation of the spectrum}
 By convention we assume that the eigenvalues are ordered ascendingly: $\lambda_{G,0} \leq \cdots \leq \lambda_{G,n}$.
@@
 For a more detailed discussion of this relation we refer to \citet{Das2004}.
+We will now instead look at the discriminative power of the spectrum.
 \begin{figure}[ht]
@@
 \end{defn}
+
+{\setlength{\parskip}{0}\paragraph{Spectral graph comparisons}
 \citet{Alzaga2010} have shown that, while the cospectrality test can be more powerful than 1-\acs{wl}, it is always weaker than 2-\acs{wl}, i.e.\ $G \mathrel{\simeq_2} H \implies G \mathrel{\simeq_\lambda} H$.
-Despite this limit on the discriminative power of the graph spectrum, it is still useful for determining graph similarity, e.g.\ by defining a distance measure on the vector space of spectra (see \citet{Gu2015}).
+Despite this limit on the discriminative power of the graph spectrum, it is still useful for determining graph similarity, e.g.\ by defining a distance measure on the vector space of spectra (see \citet{Gu2015}).}
 
@@
 \subsubsection{Spectral \acp{gcnn}}
+
 The class of spectral \acp{gcnn} is motivated by spectral graph theory (see \cref{sec:related:character:spectral}).
+A spectral \ac{gcnn} expects a graph $G$ with real vertex feature vectors $x_G[v_i] \in \mathbb{R}^d$ as its input.
+We call $X_G \coloneqq \begin{pmatrix} x_G[v_1] \\ \vdots \\ x_G[v_n] \end{pmatrix} \in \mathbb{R}^{n \times d}$ the vertex feature matrix of $G$.
+Note that each of the $d$ columns of $X_G$ can be interpreted as a separate graph signal function $x_{G,j}: \mathcal{V}_G \to \mathbb{R}$ for $j \in [d]$.
+The core idea of spectral \acp{gcnn} is to learn a graph convolution kernel $g$, similar to the grid kernels in conventional \acp{cnn}~\cite{LeCun1998}.
+The problem with this idea is that the convolution operation $g * x$ requires some notion of distance in order to ``move'' the kernel $g$ over the function $x$.
+Graphs do not generally satisfy this requirement, i.e.\ the notion of a vertex distance $\|v_i - v_j\|$ is not clearly defined.
+Via the convolution theorem $g * x = \mathcal{F}^{-1}(\hat{g} \odot \hat{x})$ we can however still define a graph convolution operator that works directly in the spectral domain:
+\begin{align}
+	g * x_{G,j} \coloneqq U_G^{\top} (\hat{g} \odot (U_G x_{G,j}))
+	\text{ with $\odot$ denoting element-wise multiplication.}
+\end{align}
+To see the connection to the convolution theorem, remember that the Laplacian eigenvector matrices $U_G$ and $U_G^{\top}$ correspond to the \ac{ft} $\mathcal{F}$ and inverse \ac{ft} $\mathcal{F}^{-1}$ respectively.
+In this formulation of convolution the Fourier transformed kernel $\hat{g}$ can be interpreted as a spectral filter, i.e.\ it dampens or amplifies certain eigenvector components of a graph.
+\citet{Bruna2013}\cite{Henaff2015} first described a \ac{gnn} architecture that learns such a spectral filter $\hat{g}$.
+Since a \ac{gnn} has to accept many different graphs of varying size $n$, the filter can however not be learned as a direct mapping $\hat{g}: [n] \to \mathbb{R}$.
+Therefore it is not expressed in terms of the indices $i$ of specific eigenvectors $u_{G,i}$ but instead in terms of the corresponding eigenvalues $\lambda_{G,i}$, i.e.\ $\hat{g}(\Lambda_G) = \begin{pmatrix} \hat{g}(\lambda_{G,1}) & & \\ & \ddots & \\ & & \hat{g}(\lambda_{G,n}) \end{pmatrix}$.
+In order to learn such an eigenvalue filter $\hat{g}: \mathbb{R} \to \mathbb{R}$ via gradient descent, it requires some differentiable parameterization.
+
+\paragraph{Chebyshev filters}
+One such parameterization is based on the family of recursively defined Chebyshev polynomials $T_k(x) = 2x T_{k-1}(x) - T_{k-2}(x)$ with $T_0 = 1$ and $T_1 = x$~\cite{Defferrard2016}.
+It describes a spectral filter as a linear combination $\hat{g}_{\theta}(\lambda) \coloneqq \sum_{k=0}^{K-1} \theta_k T_k(\lambda)$ with $\theta \in \mathbb{R}^K$.
+By restricting $\hat{g}_{\theta}$ to be a polynomial, the graph convolution can be rewritten as
+\begin{align}
+	g_{\theta} * x_{G,j} = U_G^{\top}\, \hat{g}_{\theta}(\tfrac{2}{\max \lambda_{G}} \Lambda_G - I)\, U_G\, x_{G,j} = \hat{g}_{\theta}(\tfrac{2}{\max \lambda_{G}} L_G - I)\, x_{G,j} % chktex 21
+\end{align}
+because $L_G = U_G^{\top} \Lambda_G U_G$ is an eigendecomposition of the Laplacian.
+The $\tfrac{2}{\max \lambda_{G}} \Lambda_G - I$ term normalizes the eigenvalues to $[-1, 1]$ which prevents vanishing and exploding gradients.
+The advantage of this formulation is that it can be evaluated without actually having to compute the expensive eigendecomposition.
+Instead, by interpreting $\hat{g}_{\theta}$ as a matrix polynomial, one only has to compute the powers $L^1, \dots, L^{K-1}$ which is generally much cheaper.
+
+\paragraph{Linear filters}
+\citet{Kipf2017} simplify the Chebyshev filter even further in the ambiguously named \ac{gcn} architecture\footnote{
+	We use ``\acs{gcn}'' to refer to their proposed specific method and ``\acs{gcnn}'' to refer to the broader family of \acl{gcnn} methods.
+}.
+By fixing $K = 2$ and by using a single filter parameter $\theta \in \mathbb{R}$ the convolution operation is reduced to
+\begin{align}
+	g_{\theta} * x_{G,j} = \theta \left( \tilde{D}_G^{-\frac{1}{2}} \tilde{A}_G \tilde{D}_G^{-\frac{1}{2}} \right) x_{G,j}
+	\quad\text{with $\tilde{A}_G = A_G + I$ and $\tilde{D}_G = D_G + I$.}
+\end{align}
+In this formulation the spectral filter is effectively just a linear function $\hat{g}_{\theta}(\lambda) = \theta \lambda$.
+Via this simplified notion of convolution for a single feature signal $x_{G,j}$, a convolutional layer over all features can analogously be defined as
+\begin{align}
+	Z^{(i)} \coloneqq \sigma\left( \hat{A} Z^{(i-1)} \Theta^{(i)} \right)
+	\quad\text{with } \hat{A} \coloneqq \tilde{D}_G^{-\frac{1}{2}} \tilde{A}_G \tilde{D}_G^{-\frac{1}{2}}\quad\text{and}\quad Z^{(0)} \coloneqq X_G \text{.}
+\end{align}
+Here $\sigma$ is some non-linearity, e.g.\ ReLU, and $\Theta^{(i)} \in \mathbb{R}^{d^{(i-1)} \times d^{(i)}}$ is a matrix of learned filter parameters.
+This type of convolutional layer takes a graph signal $Z^{(i-1)} \in \mathbb{R}^{n \times d^{(i-1)}}$ and the corresponding adjacency matrix $A_G \in \mathbb{R}^{n \times n}$ as input and outputs a convolved signal $Z^{(i)} \in \mathbb{R}^{n \times d^{(i)}}$.
 

### 2020-03-06 thesis/content/chapter-related-work.tex
started with spatial gcnn section + various minor improvements

@@
 	$\chi_{G,1}^{(i+1)}(v) \coloneqq h\left(\chi_{G,1}^{(i)}(v), \ldblbrace \chi_{G,1}^{(i)}(u)\, |\, u \in \Gamma_{G}(v) \rdblbrace\right)$,
-	with $\Gamma_G(v)$ denoting the set of adjacent vertices of $v \in \mathcal{V}_G$ and $h: \mathcal{C}^* \to \mathcal{C}$ denoting a perfect hash function that assigns a unique color to each finite combination of colors.
+	with $\Gamma_G(v)$ denoting the set of adjacent vertices of $v \in \mathcal{V}_G$ and $h: \mathcal{C}^* \to \mathcal{C}$ denoting an injective hash function that assigns a unique color to each finite combination of colors.
 \end{defn}
@@
 It turns out that there are multiple possible ways to do so, the most simple being the so-called \textit{combinatorial graph Laplacian}.
-\begin{defn}
+\begin{defn}\label{defn:related:laplacian-comb}
 	The combinatorial graph Laplacian $L_G \in \mathbb{R}^{n \times n}$ is defined as
@@
 \end{figure}
-\begin{defn}
+\begin{defn}\label{defn:related:cospectral}
 	Two graphs $G$ and $H$ are \textit{cospectral} ($G \mathrel{\simeq_\lambda} H$) iff.\ $\forall i: \lambda_{G,i} = \lambda_{H,i}$.
@@
 
-{\setlength{\parskip}{0}\paragraph{Spectral graph comparisons}
+{\setlength{\parskip}{0pt}\paragraph{Spectral graph comparisons}
 \citet{Alzaga2010} have shown that, while the cospectrality test can be more powerful than 1-\acs{wl}, it is always weaker than 2-\acs{wl}, i.e.\ $G \mathrel{\simeq_2} H \implies G \mathrel{\simeq_\lambda} H$.
@@
 One common way to do so is via the so-called \textit{symmetric normalized Laplacian} $L^{\text{sym}}_G$.
-\begin{defn}
+\begin{defn}\label{defn:related:laplacian-sym}
 	$L^{\text{sym}}_G \coloneqq D^{-\frac{1}{2}} L_G D^{-\frac{1}{2}}$ is the symmetric normalized Laplacian of $G$.
@@
 	g * x_{G,j} \coloneqq U_G^{\top} (\hat{g} \odot (U_G x_{G,j}))
-	\text{ with $\odot$ denoting element-wise multiplication.}
+	\text{ with $\odot$ denoting element-wise multiplication.} \label{eq:related:graph-conv-general}
 \end{align}
@@
 Since a \ac{gnn} has to accept many different graphs of varying size $n$, the filter can however not be learned as a direct mapping $\hat{g}: [n] \to \mathbb{R}$.
-Therefore it is not expressed in terms of the indices $i$ of specific eigenvectors $u_{G,i}$ but instead in terms of the corresponding eigenvalues $\lambda_{G,i}$, i.e.\ $\hat{g}(\Lambda_G) = \begin{pmatrix} \hat{g}(\lambda_{G,1}) & & \\ & \ddots & \\ & & \hat{g}(\lambda_{G,n}) \end{pmatrix}$.
+Therefore it is not expressed in terms of the indices $i$ of specific eigenvectors $u_{G,i}$ but instead in terms of the corresponding eigenvalues\footnote{
+	Here either the unnormalized eigenvalues of $L_G$ or the eigenvalues of some normalized Laplacian, e.g.\ $L_G^{\text{sym}}$, can be used.
+} via $\hat{g}(\Lambda_G) = \begin{pmatrix} \hat{g}(\lambda_{G,1}) & & \\ & \ddots & \\ & & \hat{g}(\lambda_{G,n}) \end{pmatrix}$.
 In order to learn such an eigenvalue filter $\hat{g}: \mathbb{R} \to \mathbb{R}$ via gradient descent, it requires some differentiable parameterization.
@@
 It describes a spectral filter as a linear combination $\hat{g}_{\theta}(\lambda) \coloneqq \sum_{k=0}^{K-1} \theta_k T_k(\lambda)$ with $\theta \in \mathbb{R}^K$.
-By restricting $\hat{g}_{\theta}$ to be a polynomial, the graph convolution can be rewritten as
+This restricts $\hat{g}_{\theta}$ to be a polynomial which allows us to rewrite the graph convolution from \cref{eq:related:graph-conv-general} as
 \begin{align}
-	g_{\theta} * x_{G,j} = U_G^{\top}\, \hat{g}_{\theta}(\tfrac{2}{\max \lambda_{G}} \Lambda_G - I)\, U_G\, x_{G,j} = \hat{g}_{\theta}(\tfrac{2}{\max \lambda_{G}} L_G - I)\, x_{G,j} % chktex 21
+	g_{\theta} * x_{G,j}
+	\enspace=\enspace U_G^{\top}\, \hat{g}_{\theta}(\tfrac{2}{\max{\lambda_{G}}} \Lambda_G - I)\, U_G\, x_{G,j}
+	\enspace=\enspace \hat{g}_{\theta}(\tfrac{2}{\max{\lambda_{G}}} L_G - I)\, x_{G,j} \label{eq:related:graph-conv-cheby} % chktex 21
 \end{align}
@@
 The advantage of this formulation is that it can be evaluated without actually having to compute the expensive eigendecomposition.
-Instead, by interpreting $\hat{g}_{\theta}$ as a matrix polynomial, one only has to compute the powers $L^1, \dots, L^{K-1}$ which is generally much cheaper.
+Instead, by interpreting $\hat{g}_{\theta}$ as a matrix polynomial, one only has to compute the powers $L^1, \dots, L^{K-1}$ and the largest eigenvalue $\max{\lambda_G}$ which is generally much cheaper than computing the full spectrum.
 
 \paragraph{Linear filters}
-\citet{Kipf2017} simplify the Chebyshev filter even further in the ambiguously named \ac{gcn} architecture\footnote{
-	We use ``\acs{gcn}'' to refer to their proposed specific method and ``\acs{gcnn}'' to refer to the broader family of \acl{gcnn} methods.
+\citet{Kipf2017} simplify the Chebyshev filter from \cref{eq:related:graph-conv-cheby} even further in the ambiguously named \ac{gcn} architecture\footnote{
+	We use ``\acs{gcn}'' to refer to their proposed specific method in order to be consistent with other literature.
+	We use ``\acs{gcnn}'' to refer to the broader family of \acl{gcnn} methods.
 }.
-By fixing $K = 2$ and by using a single filter parameter $\theta \in \mathbb{R}$ the convolution operation is reduced to
+By fixing $K = 2$, by using a single filter parameter $\theta \in \mathbb{R}$ and by assuming that $\max{\lambda_G^{\text{sym}}} \approx 2$ (see \cref{defn:related:laplacian-sym}, \cpageref{defn:related:laplacian-sym}) the convolution operation is reduced to
 \begin{align}
 	g_{\theta} * x_{G,j} = \theta \left( \tilde{D}_G^{-\frac{1}{2}} \tilde{A}_G \tilde{D}_G^{-\frac{1}{2}} \right) x_{G,j}
-	\quad\text{with $\tilde{A}_G = A_G + I$ and $\tilde{D}_G = D_G + I$.}
+	\quad\text{with $\tilde{A}_G = A_G + I$ and $\tilde{D}_G = D_G + I$.} \label{eq:related:graph-conv-gcn}
 \end{align}
 In this formulation the spectral filter is effectively just a linear function $\hat{g}_{\theta}(\lambda) = \theta \lambda$.
-Via this simplified notion of convolution for a single feature signal $x_{G,j}$, a convolutional layer over all features can analogously be defined as
+Via this simplified notion of convolution for a single feature signal $x_{G,j}$, a convolutional \acl{nn} layer over all features ${\{ x_{G,j} \}}_{j=1}^{d}$ can analogously be defined as
 \begin{align}
 	Z^{(i)} \coloneqq \sigma\left( \hat{A} Z^{(i-1)} \Theta^{(i)} \right)
-	\quad\text{with } \hat{A} \coloneqq \tilde{D}_G^{-\frac{1}{2}} \tilde{A}_G \tilde{D}_G^{-\frac{1}{2}}\quad\text{and}\quad Z^{(0)} \coloneqq X_G \text{.}
+	\quad\text{with } \hat{A} \coloneqq \tilde{D}_G^{-\frac{1}{2}} \tilde{A}_G \tilde{D}_G^{-\frac{1}{2}}\quad\text{and}\quad Z^{(0)} \coloneqq X_G \text{.} \label{eq:related:gcn-layer}
 \end{align}
@@
 This type of convolutional layer takes a graph signal $Z^{(i-1)} \in \mathbb{R}^{n \times d^{(i-1)}}$ and the corresponding adjacency matrix $A_G \in \mathbb{R}^{n \times n}$ as input and outputs a convolved signal $Z^{(i)} \in \mathbb{R}^{n \times d^{(i)}}$.
+By stacking multiple of those convolutional layers, complex non-linear signal filters can be learned despite the linearity of the underlying spectral filter.
 
 \subsubsection{Spatial \acp{gcnn}}
+Looking at the \ac{gcn} layer defined in \cref{eq:related:gcn-layer}, notice that, even though it is motivated by spectral graph theory, the spectrum $\lambda_G$ is not actually directly used.
+This allows for a very different perspective on \acp{gcn}:
+Instead of interpreting convolutions as applications of linear spectral filters, we can interpret them as vertex neighborhood aggregations, similar to 1-\acs{wl}.
+From this perspective a \ac{gcn} can be seen as a so-called spatial \ac{gcnn} because it operates directly in the vertex domain.
+To see why this is the case, we rewrite \cref{eq:related:gcn-layer} and compare it with the 1-\acs{wl} color refinement strategy (see \cref{defn:related:wl1-refine}, \cpageref{defn:related:wl1-refine}):
+\begin{align*}
+	\text{\ac{gcn}:}\quad&& Z^{(i)}[v] &= \sigma\left(\left( {c(v)}^2 Z^{(i-1)}[v] + \smashoperator[lr]{\sum_{u \in \Gamma_G(v)}} c(v) c(u) Z^{(i-1)}[u] \right) \Theta^{(i)} \right) \\
+	\text{1-\acs{wl}:}\quad&& \chi_{G,1}^{(i)}(v) &= h\left( \chi_{G,1}^{(i-1)}(v), \ldblbrace \chi_{G,1}^{(i-1)}(u)\, |\, u \in \Gamma_{G}(v) \rdblbrace\right)
+\end{align*}
+Here $c(v) \coloneqq {\left(\left| \Gamma_G(v) \right| + 1\right)}^{-\frac{1}{2}}$ are normalization factors introduced by $\tilde{D}_G^{-\frac{1}{2}}$ which can be ignored for now.
+The comparison shows that \ac{gcn} convolutions can be understood as iterative refinements of continuous ``color vectors'' $Z^{(i)}[v] \in \mathcal{R}^{d^{(i)}}$.
+The continuous analogue to the injective 1-\acs{wl} hash function $h: \mathcal{C}^{*} \to \mathcal{C}$ is the function $h_{\text{\acs{gcn}}}(z_1, \dots, z_m) = \sigma(\sum z_i \Theta^{(i)})$.
+
+The relation between this type of spatial convolution and 1-\acs{wl} was formally analyzed by \citet{Xu2018}.
+They show that the discriminative power of \acp{gnn} that use a vertex-aggregation scheme is in fact bounded by 1-\acs{wl}.
+Moreover they show that the \ac{gcn} architecture is strictly less powerful than 1-\acs{wl} due to the fact that $h_{\text{\acs{gcn}}}$ is not an injective hash function.
 

### 2020-03-06 thesis/content/chapter-related-work.tex
continued with spatial gcnn section

@@
 \end{align*}
-Here $c(v) \coloneqq {\left(\left| \Gamma_G(v) \right| + 1\right)}^{-\frac{1}{2}}$ are normalization factors introduced by $\tilde{D}_G^{-\frac{1}{2}}$ which can be ignored for now.
-The comparison shows that \ac{gcn} convolutions can be understood as iterative refinements of continuous ``color vectors'' $Z^{(i)}[v] \in \mathcal{R}^{d^{(i)}}$.
-The continuous analogue to the injective 1-\acs{wl} hash function $h: \mathcal{C}^{*} \to \mathcal{C}$ is the function $h_{\text{\acs{gcn}}}(z_1, \dots, z_m) = \sigma(\sum z_i \Theta^{(i)})$.
+Here $c(v) \coloneqq {\left(\left| \Gamma_G(v) \right| + 1\right)}^{-\frac{1}{2}}$ are normalization factors introduced by $\tilde{D}_G^{-\frac{1}{2}}$.
+The comparison shows that \ac{gcn} convolutions can be understood as iterative refinements of continuous ``color vectors'' $Z^{(i)}[v] \in \mathbb{R}^{d^{(i)}}$.
+Then the continuous analogue to the injective 1-\acs{wl} hash function $h: \mathcal{C}^{*} \to \mathcal{C}$ can be defined as $h_{\text{\acs{gcn}}}(z_0, z_1, \dots, z_m) = \sigma(\sum_{j=0}^m c(v_0) c(v_j) z_j \Theta^{(i)})$ which computes a single color vector from multiple color vectors $z_j$.
 
-The relation between this type of spatial convolution and 1-\acs{wl} was formally analyzed by \citet{Xu2018}.
-They show that the discriminative power of \acp{gnn} that use a vertex-aggregation scheme is in fact bounded by 1-\acs{wl}.
-Moreover they show that the \ac{gcn} architecture is strictly less powerful than 1-\acs{wl} due to the fact that $h_{\text{\acs{gcn}}}$ is not an injective hash function.
+\paragraph{A 1-\acs{wl} bound on \ac{gnn} power}
+The relation between \acp{gcnn} using neighborhood aggregation convolution and 1-\acs{wl} was formally analyzed by \citet{Xu2018}.
+They show that the discriminative power of such a \acp{gcnn} is upper-bounded by that of 1-\acs{wl}.
+In particular the \ac{gcn} architecture is strictly less powerful than 1-\acs{wl} due to the fact that $h_{\text{\acs{gcn}}}$ is not an injective hash function.
+To overcome this limitation an alternative neighborhood aggregation architecture called \ac{gin} is proposed:
+\begin{align}
+	Z^{(i)}[v] \coloneqq \text{\acs{mlp}}^{(i)}\left( Z^{(i-1)}[v] + \smashoperator[lr]{\sum_{u \in \Gamma_G(v)}} Z^{(i-1)}[u] \right)
+\end{align}
+By leaving out the normalization factors $c(v)$ and by using a \ac{mlp} instead of the single fully-connected layer $\sigma \circ \Theta^{(i)}$ the resulting hashing function $h_{\text{\acs{gin}}}(z_1, \dots, z_m) = \text{\acs{mlp}}^{(i)}(\sum_{j=1}^{m} z_j)$ becomes injective\footnote{
+	Under the assumption that the set of possible input feature matrices $X \in \mathbb{R}^{n \times d^{(0)}}$ is countable.
+	This is reasonable since real-world \ac{gcr} domains typically only have a finite set of possible vertex feature vectors.
+}, i.e.\ it assigns a unique color vector to each multiset of color vectors.
+This implies that \ac{gin} has the same discriminative power as 1-\ac{wl}.
+As expected, \ac{gin} therefore fits training data better than non-injective \acp{gcnn} like \ac{gcn}.
+Interestingly \ac{gin}'s additionally seem to generalize better on test data than non-injective methods.
 

### 2020-03-07 thesis/content/chapter-related-work.tex
added k-gnn description to spatial gcnn section

@@
 A spectral \ac{gcnn} expects a graph $G$ with real vertex feature vectors $x_G[v_i] \in \mathbb{R}^d$ as its input.
+Those feature vectors are typically one-hot encodings of labels or, if no such information is provided, vertex embedding vectors (see \cref{sec:related:gcr:embed}).
 We call $X_G \coloneqq \begin{pmatrix} x_G[v_1] \\ \vdots \\ x_G[v_n] \end{pmatrix} \in \mathbb{R}^{n \times d}$ the vertex feature matrix of $G$.
@@
 The relation between \acp{gcnn} using neighborhood aggregation convolution and 1-\acs{wl} was formally analyzed by \citet{Xu2018}.
-They show that the discriminative power of such a \acp{gcnn} is upper-bounded by that of 1-\acs{wl}.
+They show that the discriminative power of such a \ac{gcnn} is upper-bounded by that of 1-\acs{wl}.
 In particular the \ac{gcn} architecture is strictly less powerful than 1-\acs{wl} due to the fact that $h_{\text{\acs{gcn}}}$ is not an injective hash function.
-To overcome this limitation an alternative neighborhood aggregation architecture called \ac{gin} is proposed:
+To overcome this limitation an alternative neighborhood aggregation architecture called \ac{gin} was proposed:
 \begin{align}
@@
 	Under the assumption that the set of possible input feature matrices $X \in \mathbb{R}^{n \times d^{(0)}}$ is countable.
-	This is reasonable since real-world \ac{gcr} domains typically only have a finite set of possible vertex feature vectors.
+	This is reasonable since real-world \ac{gcr} domains typically only have a finite set of possible vertex feature vectors, e.g\ one-hot label encodings.
 }, i.e.\ it assigns a unique color vector to each multiset of color vectors.
-This implies that \ac{gin} has the same discriminative power as 1-\ac{wl}.
+It was shown that this implies that \ac{gin} has the same discriminative power as 1-\ac{wl}.
 As expected, \ac{gin} therefore fits training data better than non-injective \acp{gcnn} like \ac{gcn}.
-Interestingly \ac{gin}'s additionally seem to generalize better on test data than non-injective methods.
+Interestingly \ac{gin} additionally seems to generalize better on test data than non-injective methods~\cite{Xu2018}\cite{Errica2020}.
+It is not yet fully understood why this is the case.
+
+\paragraph{Higher dimensional \ac{wl} \acp{gnn}}
+The majority of \ac{gnn} methods uses a neighborhood aggregation scheme which limits their discriminative and, more importantly, computational power to that of 1-\acs{wl} (see \cref{defn:related:wl-count-detect}, \cpageref{defn:related:wl-count-detect}).
+To go beyond the limits of 1-\acs{wl}, \citet{Morris2019} proposed the first \ac{gcnn} architecture inspired by the higher dimensional $k$-\acs{wl} algorithms.
+Similar to the $k$-LWL kernel which we looked at in \cref{sec:related:gcr:kernel}, their so-called $k$-\acs{gnn} uses the $k$-multiset and neighborhood localization optimizations to keep runtime costs feasible.
+The $k$-\ac{gnn} convolution is described by
+\begin{align}
+	Z^{(i)}[s] \coloneqq \sigma\left( {Z^{(i-1)}[s]} W^{(i)} + \smashoperator[lr]{\sum_{v \in s, u \in \Gamma_G(v)}} {Z^{(i-1)}[s \setminus \{ v \} \cup \{ u \}]} W_{\Gamma}^{(i)} \right) \text{.} \label{eq:related:kgnn-layer}
+\end{align}
+It convolves the feature vectors of $k$-multisets $s \subseteq \mathcal{V}_G$ instead of single vertices $v \in \mathcal{V}_G$.
+Apart from this fundamental difference to 1-\acs{wl} bounded \acp{gnn}, $k$-\acp{gnn} additionally use two learnable parameter matrices $W^{(i)}, W_{\Gamma}^{(i)} \in \mathbb{R}^{d^{(i-1)} \times d^{(i)}}$ instead of the single $\Theta^{(i)}$ used in \acp{gcn}.
+This separation of parameter matrix effectively fulfills the same purpose as the \ac{mlp} in \ac{gin}; it makes the implicitly defined $k$-\acs{gnn} hash function injective.
+This can be seen by realizing that a stack of $k$-\acs{gnn} convolution layers with $W_{\Gamma}^{(i)} = 0$ for all but the first layer simulates a \ac{mlp}.
+
+As a final remark, note that the concept of ``neighborhood'' used in \cref{eq:related:kgnn-layer} is slightly different from that used by the $k$-\acs{wl} algorithm (see \cref{defn:related:wlk-refine}, \cpageref{defn:related:wlk-refine}):
+$k$-\acs{gnn} considers a single $k$-multiset as a neighbor while $k$-\acs{wl} calls a $k$-set of $k$-multisets a neighbor.
+We will get back to this difference later.
 

### 2020-03-08 thesis/content/chapter-related-work.tex
first draft of related works chapter completed

@@
 	\end{align*}
-	Using this notion of equivalence, we call $[G[v]] \coloneqq \{ G[w]\, |\, G[w] \equiv G[v] \land w \in \mathcal{V}_G^{*} \}$ the ordered subgraph equivalence class of $G[v]$.
+	Using this notion of equivalence, we call ${\left[ G[v] \right]} \coloneqq \{ G[w]\, |\, G[w] \equiv G[v] \land w \in \mathcal{V}_G^{*} \}$ the \textit{ordered subgraph equivalence class} of $G[v]$.
 \end{defn}
@@
 	Two graphs $G$ and $H$ are \textit{isomorphic} ($G \simeq H$) iff.\ there are equivalent encodings $G[v] \equiv H[w]$ of them.
-	Consequently $[G] \coloneqq \{ H\, |\, H \simeq G \}$ is called the isomorphism class of $G$.
+	Consequently $[G] \coloneqq \{ H\, |\, H \simeq G \}$ is called the \textit{isomorphism class} of $G$.
 \end{defn}
 The goal of the \ac{gi} problem is to check whether $G \simeq H$ for two arbitrary graphs.
-Even though there is no known universal algorithm in $P$ that solves \ac{gi}, \citet{Babai1980} showed that almost all graphs can be trivially distinguished in linear time.
+Even though there is no known universal polynomial algorithm that solves \ac{gi}, \citet{Babai1980} showed that almost all graphs can be trivially distinguished in linear time.
 More recently \citet{Babai2015} also presented a quasipolynomial upper time bound for the remaining hard \ac{gi} instances.
@@
 \subsubsection{The 1-dimensional \acs{wl} algorithm}
-In the 1-dimensional \ac{wl} algorithm, a color is assigned to each vertex of a graph.
+In the 1-dimensional \ac{wl} algorithm (1-WL), a color is assigned to each vertex of a graph.
 If the vertices $v \in \mathcal{V}_G$ of the input graph $G$ are labeled, those labels $l_G[v] \in L_{\mathcal{V}} \subseteq \mathcal{C}$ can be used as the initial graph coloring $\chi_{G,1}^{(0)}(v) \coloneqq l_G[v]$.
@@
 Note that the eigenvalues of $L_G$ grow with the number of edges\footnote{
-	This is trivially shown by $\sum_{i=1}^n \lambda_{G,i} = \Tr(L_G) = \Tr(D_G) = |\mathcal{E}_G|$.
+	This is trivially shown by $\sum_{i=1}^n \lambda_{G,i} = \Tr(t_G) = \Tr(D_G) = |\mathcal{E}_G|$.
 }.
@@
 \begin{equation}
-	C_k(G_i) \coloneqq \bigcup_{\mathclap{v_j \in \mathcal{V}_{G_i}}}\, {\left\{ \chi_{G_i, 1}^{(l)}(v_j) \right\}}_{l = 0}^L % chktex 21
-	\quad\text{with $k \in \mathbb{N}$ and $\chi_{G_i, 1}^{(l)}$ as in \cref{defn:related:wl1-refine}.}
+	C_k(G_i) \coloneqq \bigcup_{\mathclap{v_j \in \mathcal{V}_{G_i}}}\, {\left\{ \chi_{G_i, 1}^{(t)}(v_j) \right\}}_{l = 0}^L % chktex 21
+	\quad\text{with $k \in \mathbb{N}$ and $\chi_{G_i, 1}^{(t)}$ as in \cref{defn:related:wl1-refine}.}
 \end{equation}
@@
 \begin{align}
-	\varphi_{\text{ST}}(G) \coloneqq \bigoplus_{l=0}^{L} {\left( \mathit{dist}_{\chi_{G,1}^{(l)}}(c) \right)}_{c \in \mathcal{C}}
+	\varphi_{\text{ST}}(G) \coloneqq \bigoplus_{t=0}^{T} {\left( \mathit{dist}_{\chi_{G,1}^{(t)}}(c) \right)}_{c \in \mathcal{C}}
 	\quad\text{with $\oplus$ denoting vector concatenation.} \label{eq:related:wl-subtree-kernel}
 \end{align}
-$\varphi_{\text{ST}}(G)$ encodes the color counts across a fixed number $L$ of 1-\acs{wl} refinement steps.
+$\varphi_{\text{ST}}(G)$ encodes the color counts across a fixed number $T$ of 1-\acs{wl} refinement steps.
 Via this mapping, the \ac{wl} subtree kernel function $k_{\text{ST}}: \mathcal{G} \times \mathcal{G} \to \mathbb{R}$ can then be simply written as the standard inner vector product $k_{\text{ST}}(G, H) \coloneqq \langle \varphi_{\text{ST}}(G), \varphi_{\text{ST}}(H) \rangle$.
@@
 	\includegraphics[width=0.7\linewidth]{gfx/related-work/wl-subtree.pdf}
-	\caption[Illustration of the correspondence between the \ac{wl} color $\chi_{G,1}^{(l)}(v)$ and the breadth-first subtree of depth $l$ rooted at $v$.]{
-		Illustration of the correspondence between the \ac{wl} color $\chi_{G,1}^{(l)}(v)$ and the breadth-first subtree of depth $l$ rooted at $v$ (for $l = 2$ and $v = \texttt{\textcolor{t_blue}{\circled{1}}}$).
+	\caption[Illustration of the correspondence between the \ac{wl} color $\chi_{G,1}^{(t)}(v)$ and the breadth-first subtree of depth $l$ rooted at $v$.]{
+		Illustration of the correspondence between the \ac{wl} color $\chi_{G,1}^{(t)}(v)$ and the breadth-first subtree of depth $l$ rooted at $v$ (for $l = 2$ and $v = \texttt{\textcolor{t_blue}{\circled{1}}}$).
 % chktex 25
@@
 \Cref{fig:related:wl-subtree} illustrates how this definition relates to subtrees.
-Since alls vertex colors $c \in \mathcal{C}$ correspond to a subtree isomorphism class, $k_{\text{ST}}(G, H)$ effectively computes the similarity of $G$ and $H$ by comparing their local subtree structures of depth at most $L$.
+Since alls vertex colors $c \in \mathcal{C}$ correspond to a subtree isomorphism class, $k_{\text{ST}}(G, H)$ effectively computes the similarity of $G$ and $H$ by comparing their local subtree structures of depth at most $T$.
 Note that the underlying mapping $\varphi_{\text{ST}}$ cannot be computed independently for each graph $G \in \mathcal{G}_{\mathcal{D}} \subseteq \mathcal{G}$ in a given training dataset $\mathcal{D}$ since the set of colors $c \in \mathcal{C}$ introduced by the hashing function $h$ is conjointly determined by all graphs $\mathcal{G}_{\mathcal{D}}$ (see \cref{defn:related:wl1-refine}).
@@
 Extending the idea of the \ac{wl} subtree kernel, the \textit{\ac{wl} shortest path kernel}~\cite{Shervashidze2011}\cite{Borgwardt2005} adds global structural information to the local subtree comparisons.
-The $i$-th component of a graph's shortest path vector $\varphi_{\text{SP}}(G)$ corresponds to the 4-tuple $(l_i, a_i, b_i, d_i) \in {\{ 0, \dots, L \}} \times \mathcal{C} \times \mathcal{C} \times \mathbb{N}$ and its value is described by
+The $i$-th component of a graph's shortest path vector $\varphi_{\text{SP}}(G)$ corresponds to the 4-tuple $(t_i, a_i, b_i, d_i) \in {\{ 0, \dots, T \}} \times \mathcal{C} \times \mathcal{C} \times \mathbb{N}$ and its value is described by
 \begin{align}
-	{\varphi_{\text{SP}}(G)}_i \coloneqq \left| \left\{ (v, u) \in \mathcal{V}_G^2 \,|\, {\chi_{G,1}^{(l_i)}(v) = a_i} \land {\chi_{G,1}^{(l_i)}(u) = b_i} \land {d_{\text{SP}}(v, u) = d_i} \right\} \right| \text{,} % chktex 21
+	{\varphi_{\text{SP}}(G)}_i \coloneqq \left| \left\{ (v, u) \in \mathcal{V}_G^2 \,|\, {\chi_{G,1}^{(t_i)}(v) = a_i} \land {\chi_{G,1}^{(t_i)}(u) = b_i} \land {d_{\text{SP}}(v, u) = d_i} \right\} \right| \text{,} % chktex 21
 \end{align}
@@
 Analogous to $k_{\text{ST}}$, the shortest path kernel function is defined as $k_{\text{SP}}(G, H) \coloneqq \langle \varphi_{\text{SP}}(G), \varphi_{\text{SP}}(H) \rangle$.
-Since $\varphi_{\text{SP}}(G)$ has one component for each possible 4-tuple $(l_i, a_i, b_i, d_i)$, its dimensionality grows quadratically with the total number of introduced 1-\acs{wl} colors $\mathcal{C}$ and linearly with the length of the longest shortest path $\max_{G \in \mathcal{G}_{\mathcal{D}}, (v, u) \in \mathcal{V}_G^2} d_{\text{SP}}(v, u)$.
+Since $\varphi_{\text{SP}}(G)$ has one component for each possible 4-tuple $(t_i, a_i, b_i, d_i)$, its dimensionality grows quadratically with the total number of introduced 1-\acs{wl} colors $\mathcal{C}$ and linearly with the length of the longest shortest path $\max_{G \in \mathcal{G}_{\mathcal{D}}, (v, u) \in \mathcal{V}_G^2} d_{\text{SP}}(v, u)$.
 While this approach is computationally significantly more expensive than the subtree kernel, it also has a higher discriminative power and allows the shortest path kernel to distinguish even some 1-\acs{wl} indistinguishable graphs.
@@
 Instead of incorporating shortest path information to increase the power of the \ac{wl} kernel, one can alternatively just increase the \acs{wl}-dimension $k$.
-A na{\"\i}ve generalization of the 1-\acs{wl} subtree kernel would simply use \cref{eq:related:wl-subtree-kernel} but with the k-\acs{wl} colorings $\chi_{G,k}^{(l)}$ instead of $\chi_{G,1}^{(l)}$.
+A na{\"\i}ve generalization of the 1-\acs{wl} subtree kernel would simply use \cref{eq:related:wl-subtree-kernel} but with the k-\acs{wl} colorings $\chi_{G,k}^{(t)}$ instead of $\chi_{G,1}^{(t)}$.
 The problem with this approach is that the runtime of \ac{wl} increases exponentially with $k$;
@@
 		\begin{align}
-			\chi_{G,k}^{(l+1)}(s) \coloneqq h\left(\chi_{G,k}^{(l)}(s), \ldblbrace \{ \chi_{G,k}^{(l)}(s \setminus \{v\} \cup \{u\})\, |\, v \in s \}\, |\, u \in \mathcal{V}_G \rdblbrace\right) \text{.} \label{eq:related:wlk-set-refine} % chktex 21
+			\chi_{G,k}^{(t+1)}(s) \coloneqq h\left(\chi_{G,k}^{(t)}(s), \ldblbrace \{ \chi_{G,k}^{(t)}(s \setminus \{v\} \cup \{u\})\, |\, v \in s \}\, |\, u \in \mathcal{V}_G \rdblbrace\right) \text{.} \label{eq:related:wlk-set-refine} % chktex 21
 		\end{align}
-		Even though this simplification generally reduces the discriminative power of the kernel, for $k = 2$ specifically no information and therefore no power is lost by replacing the tuples $(v, u)$ and $(u, v)$ with $\ldblbrace v, u \rdblbrace$ since the undirectedness of graphs implies that $\chi_{G,2}^{(l)}(v, u) = \chi_{G,2}^{(l)}(u, v)$ even when doing tuple-based color refinement.
+		Even though this simplification generally reduces the discriminative power of the kernel, for $k = 2$ specifically no information and therefore no power is lost by replacing the tuples $(v, u)$ and $(u, v)$ with $\ldblbrace v, u \rdblbrace$ since the undirectedness of graphs implies that $\chi_{G,2}^{(t)}(v, u) = \chi_{G,2}^{(t)}(u, v)$ even when doing tuple-based color refinement.
 	\item \textbf{Neighborhood localization:}
@@
 The $\tfrac{2}{\max \lambda_{G}} \Lambda_G - I$ term normalizes the eigenvalues to $[-1, 1]$ which prevents vanishing and exploding gradients.
-The advantage of this formulation is that it can be evaluated without actually having to compute the expensive eigendecomposition.
-Instead, by interpreting $\hat{g}_{\theta}$ as a matrix polynomial, one only has to compute the powers $L^1, \dots, L^{K-1}$ and the largest eigenvalue $\max{\lambda_G}$ which is generally much cheaper than computing the full spectrum.
+The advantage of this formulation is that it can be evaluated without actually having to compute the expensive eigendecomposition of $L_G$.
+Instead, by interpreting $\hat{g}_{\theta}$ as a matrix polynomial, one only has to compute the powers $L_G^1, \dots, L_G^{K-1}$ and the largest eigenvalue $\max{\lambda_G}$ which is generally much cheaper than computing the full spectrum.
 
@@
 \begin{align}
-	Z^{(i)} \coloneqq \sigma\left( \hat{A} Z^{(i-1)} \Theta^{(i)} \right)
+	Z^{(t)} \coloneqq \sigma\left( \hat{A} Z^{(t-1)} \Theta^{(t)} \right)
 	\quad\text{with } \hat{A} \coloneqq \tilde{D}_G^{-\frac{1}{2}} \tilde{A}_G \tilde{D}_G^{-\frac{1}{2}}\quad\text{and}\quad Z^{(0)} \coloneqq X_G \text{.} \label{eq:related:gcn-layer}
 \end{align}
-Here $\sigma$ is some non-linearity, e.g.\ ReLU, and $\Theta^{(i)} \in \mathbb{R}^{d^{(i-1)} \times d^{(i)}}$ is a matrix of learned filter parameters.
-This type of convolutional layer takes a graph signal $Z^{(i-1)} \in \mathbb{R}^{n \times d^{(i-1)}}$ and the corresponding adjacency matrix $A_G \in \mathbb{R}^{n \times n}$ as input and outputs a convolved signal $Z^{(i)} \in \mathbb{R}^{n \times d^{(i)}}$.
+Here $\sigma$ is some non-linearity, e.g.\ ReLU, and $\Theta^{(t)} \in \mathbb{R}^{d^{(t-1)} \times d^{(t)}}$ is a matrix of learned filter parameters.
+This type of convolutional layer takes a graph signal $Z^{(t-1)} \in \mathbb{R}^{n \times d^{(t-1)}}$ and the corresponding adjacency matrix $A_G \in \mathbb{R}^{n \times n}$ as input and outputs a convolved signal $Z^{(t)} \in \mathbb{R}^{n \times d^{(t)}}$.
 By stacking multiple of those convolutional layers, complex non-linear signal filters can be learned despite the linearity of the underlying spectral filter.
@@
 \begin{align*}
-	\text{\ac{gcn}:}\quad&& Z^{(i)}[v] &= \sigma\left(\left( {c(v)}^2 Z^{(i-1)}[v] + \smashoperator[lr]{\sum_{u \in \Gamma_G(v)}} c(v) c(u) Z^{(i-1)}[u] \right) \Theta^{(i)} \right) \\
-	\text{1-\acs{wl}:}\quad&& \chi_{G,1}^{(i)}(v) &= h\left( \chi_{G,1}^{(i-1)}(v), \ldblbrace \chi_{G,1}^{(i-1)}(u)\, |\, u \in \Gamma_{G}(v) \rdblbrace\right)
+	\text{\ac{gcn}:}\quad&& Z^{(t)}[v] &= \sigma\left(\left( {c(v)}^2 Z^{(t-1)}[v] + \smashoperator[lr]{\sum_{u \in \Gamma_G(v)}} c(v) c(u) Z^{(t-1)}[u] \right) \Theta^{(t)} \right) \\
+	\text{1-\acs{wl}:}\quad&& \chi_{G,1}^{(t)}(v) &= h\left( \chi_{G,1}^{(t-1)}(v), \ldblbrace \chi_{G,1}^{(t-1)}(u)\, |\, u \in \Gamma_{G}(v) \rdblbrace\right)
 \end{align*}
 Here $c(v) \coloneqq {\left(\left| \Gamma_G(v) \right| + 1\right)}^{-\frac{1}{2}}$ are normalization factors introduced by $\tilde{D}_G^{-\frac{1}{2}}$.
-The comparison shows that \ac{gcn} convolutions can be understood as iterative refinements of continuous ``color vectors'' $Z^{(i)}[v] \in \mathbb{R}^{d^{(i)}}$.
-Then the continuous analogue to the injective 1-\acs{wl} hash function $h: \mathcal{C}^{*} \to \mathcal{C}$ can be defined as $h_{\text{\acs{gcn}}}(z_0, z_1, \dots, z_m) = \sigma(\sum_{j=0}^m c(v_0) c(v_j) z_j \Theta^{(i)})$ which computes a single color vector from multiple color vectors $z_j$.
+The comparison shows that \ac{gcn} convolutions can be understood as iterative refinements of continuous ``color vectors'' $Z^{(t)}[v] \in \mathbb{R}^{d^{(t)}}$.
+Then the continuous analogue to the injective 1-\acs{wl} hash function $h: \mathcal{C}^{*} \to \mathcal{C}$ can be defined as $h_{\text{\acs{gcn}}}(z_0, z_1, \dots, z_m) = \sigma(\sum_{j=0}^m c(v_0) c(v_j) z_j \Theta^{(t)})$ which computes a single color vector from multiple color vectors $z_j$.
 
@@
 \begin{align}
-	Z^{(i)}[v] \coloneqq \text{\acs{mlp}}^{(i)}\left( Z^{(i-1)}[v] + \smashoperator[lr]{\sum_{u \in \Gamma_G(v)}} Z^{(i-1)}[u] \right)
+	Z^{(t)}[v] \coloneqq \text{\acs{mlp}}^{(t)}\left( Z^{(t-1)}[v] + \smashoperator[lr]{\sum_{u \in \Gamma_G(v)}} Z^{(t-1)}[u] \right)
 \end{align}
-By leaving out the normalization factors $c(v)$ and by using a \ac{mlp} instead of the single fully-connected layer $\sigma \circ \Theta^{(i)}$ the resulting hashing function $h_{\text{\acs{gin}}}(z_1, \dots, z_m) = \text{\acs{mlp}}^{(i)}(\sum_{j=1}^{m} z_j)$ becomes injective\footnote{
+By leaving out the normalization factors $c(v)$ and by using a \ac{mlp} instead of the single fully-connected layer $\sigma \circ \Theta^{(t)}$ the resulting hashing function $h_{\text{\acs{gin}}}(z_1, \dots, z_m) = \text{\acs{mlp}}^{(t)}(\sum_{j=1}^{m} z_j)$ becomes injective\footnote{
 	Under the assumption that the set of possible input feature matrices $X \in \mathbb{R}^{n \times d^{(0)}}$ is countable.
@@
 \begin{align}
-	Z^{(i)}[s] \coloneqq \sigma\left( {Z^{(i-1)}[s]} W^{(i)} + \smashoperator[lr]{\sum_{v \in s, u \in \Gamma_G(v)}} {Z^{(i-1)}[s \setminus \{ v \} \cup \{ u \}]} W_{\Gamma}^{(i)} \right) \text{.} \label{eq:related:kgnn-layer}
+	Z^{(t)}[s] \coloneqq \sigma\left( {Z^{(t-1)}[s]} W^{(t)} + \smashoperator[lr]{\sum_{v \in s, u \in \Gamma_G(v)}} {Z^{(t-1)}[s \setminus \{ v \} \cup \{ u \}]} W_{\Gamma}^{(t)} \right) \text{.} \label{eq:related:kgnn-layer}
 \end{align}
 It convolves the feature vectors of $k$-multisets $s \subseteq \mathcal{V}_G$ instead of single vertices $v \in \mathcal{V}_G$.
-Apart from this fundamental difference to 1-\acs{wl} bounded \acp{gnn}, $k$-\acp{gnn} additionally use two learnable parameter matrices $W^{(i)}, W_{\Gamma}^{(i)} \in \mathbb{R}^{d^{(i-1)} \times d^{(i)}}$ instead of the single $\Theta^{(i)}$ used in \acp{gcn}.
+Apart from this fundamental difference to 1-\acs{wl} bounded \acp{gnn}, $k$-\acp{gnn} additionally use two learnable parameter matrices $W^{(t)}, W_{\Gamma}^{(t)} \in \mathbb{R}^{d^{(t-1)} \times d^{(t)}}$ instead of the single $\Theta^{(t)}$ used in \acp{gcn}.
 This separation of parameter matrix effectively fulfills the same purpose as the \ac{mlp} in \ac{gin}; it makes the implicitly defined $k$-\acs{gnn} hash function injective.
-This can be seen by realizing that a stack of $k$-\acs{gnn} convolution layers with $W_{\Gamma}^{(i)} = 0$ for all but the first layer simulates a \ac{mlp}.
+This can be seen by realizing that a stack of $k$-\acs{gnn} convolution layers with $W_{\Gamma}^{(t)} = 0$ for all but the first layer simulates a \ac{mlp}.
 
-As a final remark, note that the concept of ``neighborhood'' used in \cref{eq:related:kgnn-layer} is slightly different from that used by the $k$-\acs{wl} algorithm (see \cref{defn:related:wlk-refine}, \cpageref{defn:related:wlk-refine}):
-$k$-\acs{gnn} considers a single $k$-multiset as a neighbor while $k$-\acs{wl} calls a $k$-set of $k$-multisets a neighbor.
+As a final remark, note that the concept of ``neighborhood'' in \cref{eq:related:kgnn-layer} is slightly different from that used by the $k$-\acs{wl} algorithm (see \cref{defn:related:wlk-refine}, \cpageref{defn:related:wlk-refine}):
+$k$-\acs{gnn} considers a single $k$-multiset as a neighbor while $k$-\acs{wl} uses $k$-sets of $k$-multisets as neighbors.
 We will get back to this difference later.
@@
 \subsubsection{Graph Pooling}
+
+The graph convolution approaches that we just looked at all produce a set of convolved feature vectors ${\left\{ Z^{(T)}[s_i] \right\}}_{i=1}^{m}$ after $T$ convolutional layers. % chktex 21
+For 1-\acs{wl} bounded architectures like \ac{gcn} or \ac{gin}, there is one vector for each of the $m = {|\mathcal{V}_G|}$ vertices, while a $k$-\acs{wl} inspired architecture like $k$-\acs{gnn} produces $m \leq {|\mathcal{V}_G|}^{k}$ output vectors.
+In order to solve the \ac{gcr} problem, those vector sets need to be combined into a final predicted class or regression value.
+To do this, so-called \textit{graph pooling layers} are used.
+Generally speaking there are two types of pooling approaches:
+\begin{enumerate}[label=\textbf{\arabic*.}]
+	\item \textbf{Hierarchical Pooling:}
+		This type of pooling is similar to that found in \acp{cnn} on pictures where the resolution is iteratively reduced.
+		Similarly hierarchical graph pooling iteratively \textit{coarsens} a graph every couple of convolutional layers.
+		Graph coarsening works by merging vertices in a spectrum-preserving manner~\cite{Loukas2018}.
+		This is done until a single merged vertex is left whose feature vector then represents the entire graph.
+	\item \textbf{Global Pooling}:
+		Alternatively pooling can also be performed in a single step after the convolutional layers.
+		This type of pooling takes all $m$ feature vectors and directly maps them to a single graph feature vector.
+\end{enumerate}
+In this thesis we will focus on global pooling.
+The most simple global pooling approaches use static aggregation functions like component-wise $\min$, $\max$ or $\mean$ to produce a single graph feature vector $z_G \in \mathbb{R}^{d^{(L)}}$.
+This vector is then fed into a standard \ac{mlp} to produce the final prediction.
+The state-of-the-art graph pooling approaches go beyond static aggregation and try to incorporate structural information.
+We will now briefly look at two such approaches.
+
+\paragraph{SortPooling}
+One way to combine feature vectors is called \textit{SortPooling}~\cite{Zhang2018}.
+It interprets the convolved vectors as continuous \ac{wl} colors that encode different structural roles of vertices (or vertex sets in case of $k$-\acp{gnn}).
+By imposing a component-wise lexicographic ordering on the set of color vectors, it can be reduced to a vertex permutation invariant top-$p$ list of vectors $(z_1, \dots, z_p)$ for some fixed $p \in \mathbb{N}$.
+The final graph feature vector then is the concatenation $z_G = \left( \bigoplus_{i=1}^{p} z_i \right) \in \mathbb{R}^{d^{(L)} p}$.
+SortPooling is based on the idea that the convolutional layers will learn to use the most-significant lexicographic vector components to represent that vector's importance.
+
+\paragraph{\acs{sagpool}}
+Another global pooling approach uses self-attention to explicitly assign a structural importance score to each vertex (or vertex set in case of $k$-\acp{gnn}).
+The so-called \acf{sagpool}~\cite{Lee2019} learns those attention scores via a separate stack of convolutional layers.
+The set of convolved feature vectors is then filtered down from the size $m$ to $p = \lceil r m \rceil$ with $r \in [0, 1]$ by removing the vectors of the vertices with the lowest attention scores.
+Let $z_1, \dots, z_p$ be the remaining feature vectors.
+To obtain the final graph feature vector, \ac{sagpool} uses $z_G = {\left( \sum_{i=1}^{p} z_i \right)} \oplus {\max_{i=1}^p {z_i}}$.

### 2020-03-09 thesis/content/chapter-graph-lta.tex
started with gcr related work
continued related works chapter
started with ltag definition + various minor fixes

@@
-% !TEX root = ../main.tex
+%!TEX root = ../main.tex
 % chktex-file 46
@@
 
-\section{Formalization of LTA Methods}%
-\label{sec:ltag:formal}
+In the previous chapter an introduction to two separate fields of research was given:
+\begin{enumerate*}
+	\item \Acf{lta},
+	\item \Acf{gcr}
+\end{enumerate*}.
+In this chapter we will combine them and define an extension of \ac{lta} to the \ac{gcr} problem.
+This will be done in three steps:
+\begin{enumerate}
+	\item We begin with a formal definition of what actually constitutes an \ac{lta} method as opposed to non-\acs{lta} methods.
+	\item Using this definition, we will see that some of the previously described \ac{gcnn} methods can be interpreted as \ac{lta} variants under certain conditions.
+	\item Finally a new \acs{lta}-inspired \ac{gcnn} architecture will be described.
+\end{enumerate}
 
-\section{An LTA Interpretation of Graph Methods}%
+\section{A Generalized Definition of \acs*{lta}}%
+\label{sec:ltag:definition}
+
+In order to formally define \ac{lta}, we must first decide on its defining characteristic.
+We propose that this characteristic should be the \textit{localized explainability} of \ac{lta} predictions.
+As we saw in \cref{sec:related:lta}, an \ac{lta} prediction $y \in \mathcal{Y}$ for some multiset composition $\bm{c} = \ldblbrace c_1, \dots, c_n \rdblbrace$ can always be tracked back to a set of local constituent predictions $y_1, \dots, y_n \in \mathcal{Y}$.
+Under the assumption that each constituent $c_i$ represents some human interpretable object, a composition's prediction can therefore be explained by the presence of the constituents/objects whose local predictions are indicative of the global prediction.
+
+Based on this intuition we now give a generalized definition of \ac{lta} which applies to unstructured as well as structured input data.
+We assume that compositions are represented by graphs $G \in \mathcal{G}$;
+unstructured inputs are represented by graphs with one vertex per constituent ($\mathcal{V}_G = \{ v_{c_1}, \dots, v_{c_n} \}$) and no edges ($\mathcal{E}_G = \emptyset$).
+Given such graph inputs, an \ac{lta} method must satisfy three criteria:
+\begin{enumerate}[label=\textbf{\arabic*.}]
+	\item \textbf{Decomposition:}
+		A given graph must be decomposed into a set of constituents via a decomposition function $\varphi: \mathcal{G} \to \mathcal{P}(\mathcal{G})$ for which it holds that $\forall G \in \mathcal{G}: \forall c \in \varphi(G): \exists s \in \mathcal{V}_G^{*}: c \equiv G[s]$.
+		The decomposition function splits a given composition/graph $G$ into its constituents $c$ which must be subgraphs of $G$.
+
+		In the existing unstructured \ac{lta} approaches, the decomposition function is implicitly defined as $\varphi(G) \coloneqq {\{ G[v] \}}_{v \in \mathcal{V}_G}$ since each vertex corresponds to an interpretable constituent by definition.
+		For structured data however, a split into individual vertices is typically not appropriate.
+		Molecular graphs from chemical datasets for example are meaningfully characterized by the presence of so-called \textit{functional groups} consisting of multiple bonded atoms while a characterization on the level of individual atoms is generally less meaningful~\cite{McNaught1997}.
+	\item \textbf{Disaggregation:}
+		Each constituent $c \in \varphi(G)$ must be evaluated via some function $f: \mathcal{G} \to \mathcal{Y}$.
+		Learning this constituent evaluation function is called the \textit{disaggregation problem}.
+	\item \textbf{Aggregation:}
+		Lastly an aggregation function $\mathcal{A}_{\text{\acs*{sd}}}: \mathcal{Y}^{*} \to \mathcal{Y}$ must be learned.
+\end{enumerate}
+
+\section{An \acs*{lta} Interpretation of Existing \acsp*{gcnn}}%
 \label{sec:ltag:interpretation}
 
-\section{LTA on Dynamically Decomposed Graphs}%
-\label{sec:ltag:dyndecomp}
+\section{A Novel \acs*{lta}-Inspired \acs*{gcnn} Architecture}%
+\label{sec:ltag:wl2gnn}

### 2020-03-09 thesis/content/chapter-related-work.tex
started with ltag definition + various minor fixes

@@
 		To solve the \ac{gcr} problem, relevant graphs characteristics have to be defined in order to determine the similarity and dissimilarity of graphs.
-		We will look at two common approaches for graph characterization: The \acl{wl} algorithm and the notion of graph spectra.
-	\item Using the described graph characterization approaches, a brief overview of current \ac{gcr} methods will then be given.
+		We will look at three approaches for graph characterization: The graph isomorphism test, the so-called \acl{wl} coloring and lastly the notion of graph spectra.W
+	\item Based on the described graph characterization approaches, an overview of current \ac{gcr} methods will then be given.
 \end{enumerate}
@@
 In this simple model the \ac{lta} aggregation problem boils down to the optimization of $\lambda$.
-The \ac{lta} disaggregation problem is solved by jointly optimizing a \acf{lm}, i.e.\ the constituent scores ${\ldblbrace y_i \in [0, 1] \rdblbrace}_{c_i \in \bm{c}}$ are described by $y_i = {\left( 1 + \exp(-\theta^\top x_i) \right)}^{-1}$.
+The \ac{lta} disaggregation problem is solved by jointly optimizing a \acl{lm}, i.e.\ the constituent scores ${\ldblbrace y_i \in [0, 1] \rdblbrace}_{c_i \in \bm{c}}$ are described by $y_i = {\left( 1 + \exp(-\theta^\top x_i) \right)}^{-1}$.
 Overall an \ac{lta} model is therefore described by the uninorm parameter $\lambda$ and the regression coefficients $\theta$.
@@
 	For all vertex $k$-tuples $v = (v_1, \dots, v_k) \in \mathcal{V}_G^k$ let $\hat{v} = \{ v_i\, |\, i \in [k] \}$.
-	Then $G[v] \coloneqq (\hat{v}, \mathcal{E}_G \cap \hat{v}^2, v)$ is called the ordered subgraph of $G$ induced by $v$.
-	If $\hat{v} = \mathcal{V}_G$ and $k = |\mathcal{V}_G|$, we call $v$ an ordering of $G$ and $G[v]$ an encoding of $G$.
+	Then $G[v] \coloneqq (\hat{v}, \mathcal{E}_G \cap \hat{v}^2, v)$ is called the \textit{ordered subgraph} of $G$ induced by $v$.
+	If $\hat{v} = \mathcal{V}_G$ and $k = |\mathcal{V}_G|$, we call $v$ an \textit{ordering} of $G$ and $G[v]$ an \textit{encoding} of $G$.
 \end{defn}
@@
 		&\, \begin{aligned}
-			\land&\ \forall i, j \in [k]: l_G[v_i, v_j] = l_H[w_i, w_j] &\land&\ \forall i \in [k]: l_G[v_i] = l_H[w_i] \\
-			\land&\ \forall i, j \in [k]: x_G[v_i, v_j] = x_H[w_i, w_j] &\land&\ \forall i \in [k]: x_G[v_i] = x_H[w_i]\text{.}
+			\land&\ \forall i, j \in [k]: l_G[v_i, v_j] = l_H[w_i, w_j] &\land&\quad\forall i \in [k]: l_G[v_i] = l_H[w_i] \\
+			\land&\ \forall i, j \in [k]: x_G[v_i, v_j] = x_H[w_i, w_j] &\land&\quad\forall i \in [k]: x_G[v_i] = x_H[w_i]\text{.}
 		\end{aligned}
@@
 	\caption[Illustration of the correspondence between the \ac{wl} color $\chi_{G,1}^{(t)}(v)$ and the breadth-first subtree of depth $l$ rooted at $v$.]{
-		Illustration of the correspondence between the \ac{wl} color $\chi_{G,1}^{(t)}(v)$ and the breadth-first subtree of depth $l$ rooted at $v$ (for $l = 2$ and $v = \texttt{\textcolor{t_blue}{\circled{1}}}$).
+		Illustration of the correspondence between the \ac{wl} color $\chi_{G,1}^{(t)}(v)$ and the breadth-first subtree of depth $t$ rooted at $v$ (for $t = 2$ and $v = \texttt{\textcolor{t_blue}{\circled{1}}}$).
 % chktex 25
@@
 \begin{align*}
-	\text{\ac{gcn}:}\quad&& Z^{(t)}[v] &= \sigma\left(\left( {c(v)}^2 Z^{(t-1)}[v] + \smashoperator[lr]{\sum_{u \in \Gamma_G(v)}} c(v) c(u) Z^{(t-1)}[u] \right) \Theta^{(t)} \right) \\
+	\text{\ac*{gcn}:}\quad&& Z^{(t)}[v] &= \sigma\left(\left( {c(v)}^2 Z^{(t-1)}[v] + \smashoperator[lr]{\sum_{u \in \Gamma_G(v)}} c(v) c(u) Z^{(t-1)}[u] \right) \Theta^{(t)} \right) \\
 	\text{1-\acs{wl}:}\quad&& \chi_{G,1}^{(t)}(v) &= h\left( \chi_{G,1}^{(t-1)}(v), \ldblbrace \chi_{G,1}^{(t-1)}(u)\, |\, u \in \Gamma_{G}(v) \rdblbrace\right)
@@
 The comparison shows that \ac{gcn} convolutions can be understood as iterative refinements of continuous ``color vectors'' $Z^{(t)}[v] \in \mathbb{R}^{d^{(t)}}$.
-Then the continuous analogue to the injective 1-\acs{wl} hash function $h: \mathcal{C}^{*} \to \mathcal{C}$ can be defined as $h_{\text{\acs{gcn}}}(z_0, z_1, \dots, z_m) = \sigma(\sum_{j=0}^m c(v_0) c(v_j) z_j \Theta^{(t)})$ which computes a single color vector from multiple color vectors $z_j$.
+Then the continuous analogue to the injective 1-\acs{wl} hash function $h: \mathcal{C}^{*} \to \mathcal{C}$ can be defined as $h_{\text{\acs*{gcn}}}(z_0, z_1, \dots, z_m) = \sigma(\sum_{j=0}^m c(v_0) c(v_j) z_j \Theta^{(t)})$ which computes a single color vector from multiple color vectors $z_j$.
 
@@
 They show that the discriminative power of such a \ac{gcnn} is upper-bounded by that of 1-\acs{wl}.
-In particular the \ac{gcn} architecture is strictly less powerful than 1-\acs{wl} due to the fact that $h_{\text{\acs{gcn}}}$ is not an injective hash function.
+In particular the \ac{gcn} architecture is strictly less powerful than 1-\acs{wl} due to the fact that $h_{\text{\acs*{gcn}}}$ is not an injective hash function.
 To overcome this limitation an alternative neighborhood aggregation architecture called \ac{gin} was proposed:
 \begin{align}
-	Z^{(t)}[v] \coloneqq \text{\acs{mlp}}^{(t)}\left( Z^{(t-1)}[v] + \smashoperator[lr]{\sum_{u \in \Gamma_G(v)}} Z^{(t-1)}[u] \right)
+	Z^{(t)}[v] \coloneqq \text{\acs*{mlp}}^{(t)}\left( Z^{(t-1)}[v] + \smashoperator[lr]{\sum_{u \in \Gamma_G(v)}} Z^{(t-1)}[u] \right)
 \end{align}
-By leaving out the normalization factors $c(v)$ and by using a \ac{mlp} instead of the single fully-connected layer $\sigma \circ \Theta^{(t)}$ the resulting hashing function $h_{\text{\acs{gin}}}(z_1, \dots, z_m) = \text{\acs{mlp}}^{(t)}(\sum_{j=1}^{m} z_j)$ becomes injective\footnote{
+By leaving out the normalization factors $c(v)$ and by using a \ac{mlp} instead of the single fully-connected layer $\sigma \circ \Theta^{(t)}$ the resulting hashing function $h_{\text{\acs*{gin}}}(z_1, \dots, z_m) = \text{\acs*{mlp}}^{(t)}(\sum_{j=1}^{m} z_j)$ becomes injective\footnote{
 	Under the assumption that the set of possible input feature matrices $X \in \mathbb{R}^{n \times d^{(0)}}$ is countable.
@@
 
-\paragraph{\acs{sagpool}}
+\paragraph{\acs*{sagpool}}
 Another global pooling approach uses self-attention to explicitly assign a structural importance score to each vertex (or vertex set in case of $k$-\acp{gnn}).
-The so-called \acf{sagpool}~\cite{Lee2019} learns those attention scores via a separate stack of convolutional layers.
+The so-called \ac{sagpool}~\cite{Lee2019} learns those attention scores via a separate stack of convolutional layers.
 The set of convolved feature vectors is then filtered down from the size $m$ to $p = \lceil r m \rceil$ with $r \in [0, 1]$ by removing the vectors of the vertices with the lowest attention scores.

### 2020-03-10 thesis/content/chapter-graph-lta.tex
continued with ltag definition

@@
 We propose that this characteristic should be the \textit{localized explainability} of \ac{lta} predictions.
-As we saw in \cref{sec:related:lta}, an \ac{lta} prediction $y \in \mathcal{Y}$ for some multiset composition $\bm{c} = \ldblbrace c_1, \dots, c_n \rdblbrace$ can always be tracked back to a set of local constituent predictions $y_1, \dots, y_n \in \mathcal{Y}$.
-Under the assumption that each constituent $c_i$ represents some human interpretable object, a composition's prediction can therefore be explained by the presence of the constituents/objects whose local predictions are indicative of the global prediction.
+As seen in \cref{sec:related:lta}, an \ac{lta} prediction $\hat{y}_{C} \in \mathcal{Y}$ for some multiset composition $C = \ldblbrace c_1, \dots, c_n \rdblbrace$ can always be tracked back to a set of local constituent predictions $\hat{y}_1, \dots, \hat{y}_n \in \mathcal{Y}$.
+Under the assumption that each constituent $c_i$ represents some human interpretable object, a composition's prediction $\hat{y}_{C}$ can therefore be explained by the presence of the constituents/objects whose local predictions $\hat{y}_i$ are indicative of the global prediction.
 
@@
 unstructured inputs are represented by graphs with one vertex per constituent ($\mathcal{V}_G = \{ v_{c_1}, \dots, v_{c_n} \}$) and no edges ($\mathcal{E}_G = \emptyset$).
-Given such graph inputs, an \ac{lta} method must satisfy three criteria:
+Each composition has some target value $y_G \in \mathcal{Y}$ which could be a discrete class or continuous score.
+An \textit{\ac{lta} model} $h: \mathcal{G} \to \mathcal{Y}$ assigns predictions $\hat{y}_G \in \mathcal{Y}$ to compositions $G$.
+Such a model must satisfy three criteria:
 \begin{enumerate}[label=\textbf{\arabic*.}]
 	\item \textbf{Decomposition:}
-		A given graph must be decomposed into a set of constituents via a decomposition function $\varphi: \mathcal{G} \to \mathcal{P}(\mathcal{G})$ for which it holds that $\forall G \in \mathcal{G}: \forall c \in \varphi(G): \exists s \in \mathcal{V}_G^{*}: c \equiv G[s]$.
-		The decomposition function splits a given composition/graph $G$ into its constituents $c$ which must be subgraphs of $G$.
+		A given composition $G$ must be decomposed into a set of constituents $c_{G,i}$ via a decomposition function $\varphi: \mathcal{G} \to \mathcal{P}(\mathcal{G})$ for which it holds that $\forall G \in \mathcal{G}: \forall c_{G,i} \in \varphi(G): \exists s \in \mathcal{V}_G^{*}: c_{G,i} \equiv G[s]$.
+		The decomposition function splits $G$ into its constituents $c_{G,i}$ which must be subgraphs of $G$.
 
-		In the existing unstructured \ac{lta} approaches, the decomposition function is implicitly defined as $\varphi(G) \coloneqq {\{ G[v] \}}_{v \in \mathcal{V}_G}$ since each vertex corresponds to an interpretable constituent by definition.
+		In the existing unstructured \ac{lta} approaches the decomposition function is implicitly defined as $\varphi(G) \coloneqq {\{ G[v_{c_i}] \}}_{v_{c_i} \in \mathcal{V}_G}$ since each vertex $v_{c_i}$ corresponds to an interpretable constituent $c_i$ by definition.
 		For structured data however, a split into individual vertices is typically not appropriate.
@@
 	\item \textbf{Disaggregation:}
-		Each constituent $c \in \varphi(G)$ must be evaluated via some function $f: \mathcal{G} \to \mathcal{Y}$.
-		Learning this constituent evaluation function is called the \textit{disaggregation problem}.
+		The constituents $c_{G, i} \in \varphi(G)$ must be evaluated via some function $f: \mathcal{G} \to \mathcal{Y} \times \mathbb{R}$.
+		This evaluation function assigns a prediction $\hat{y}_{G, i} \in \mathcal{Y}$ and a weight $w_{G, i} \in \mathbb{R}$ to each constituent.
+		A constituent's weight $w_{G, i}$ can intuitively be interpreted as a measure of the confidence that the local prediction $\hat{y}_{G, i}$ is indicative of the composition's global target value $y_G$.
+		Learning local predictions and weights for all possible constituents is called the \textit{disaggregation problem}.
+
+		Note that there are no explicit constituent weights in the existing unstructured \ac{lta} approaches (i.e.\ all $w_{G, i} = 1$) because the constituents are explicitly given and assumed to be similarly indicative of $y_G$.
+		For structured data however, where the decomposition $\varphi(G)$ is not given as part of the input, this assumption does not necessarily hold.
 	\item \textbf{Aggregation:}
-		Lastly an aggregation function $\mathcal{A}_{\text{\acs*{sd}}}: \mathcal{Y}^{*} \to \mathcal{Y}$ must be learned.
+		Lastly a weighted aggregation function $\mathcal{A}: {(\mathcal{Y} \times \mathbb{R})}^{*} \to \mathcal{Y}$ must be learned.
+		It combines the multiset of local constituent predictions and weights into a single global composition prediction.
 \end{enumerate}

### 2020-03-10 thesis/content/chapter-related-work.tex
continued with ltag definition

@@
 The class of \ac{lta} problems was first described by \citet{Melnikov2016}.
-There an input instance is understood as a composition $\bm{c} = \ldblbrace c_1, \dots, c_n \rdblbrace$ of so-called constituents, i.e.\ as a variable-size multiset (denoted as $\ldblbrace \cdot \rdblbrace$).
-The assumption in \ac{lta} problems is that for all constituents $c_i \in \bm{c}$ a local score $y_i \in \mathcal{Y}$ is either given or computable.
-The set of those local scores should be indicative of the overall score $y \in \mathcal{Y}$ of the composition $\bm{c}$.
+There an input instance is understood as a composition $C = \ldblbrace c_1, \dots, c_n \rdblbrace$ of so-called constituents, i.e.\ as a variable-size multiset (denoted as $\ldblbrace \cdot \rdblbrace$).
+The assumption in \ac{lta} problems is that for all constituents $c_i \in C$ a local score $y_i \in \mathcal{Y}$ is either given or computable.
+The set of those local scores should be indicative of the overall score $y \in \mathcal{Y}$ of the composition $C$.
 \ac{lta} problems typically require two subproblems to be solved:
@@
 In this simple model the \ac{lta} aggregation problem boils down to the optimization of $\lambda$.
-The \ac{lta} disaggregation problem is solved by jointly optimizing a \acl{lm}, i.e.\ the constituent scores ${\ldblbrace y_i \in [0, 1] \rdblbrace}_{c_i \in \bm{c}}$ are described by $y_i = {\left( 1 + \exp(-\theta^\top x_i) \right)}^{-1}$.
+The \ac{lta} disaggregation problem is solved by jointly optimizing a \acl{lm}, i.e.\ the constituent scores ${\ldblbrace y_i \in [0, 1] \rdblbrace}_{c_i \in C}$ are described by $y_i = {\left( 1 + \exp(-\theta^\top x_i) \right)}^{-1}$.
 Overall an \ac{lta} model is therefore described by the uninorm parameter $\lambda$ and the regression coefficients $\theta$.

### 2020-03-11 thesis/content/chapter-graph-lta.tex
continued ltag definition

@@
 We propose that this characteristic should be the \textit{localized explainability} of \ac{lta} predictions.
-As seen in \cref{sec:related:lta}, an \ac{lta} prediction $\hat{y}_{C} \in \mathcal{Y}$ for some multiset composition $C = \ldblbrace c_1, \dots, c_n \rdblbrace$ can always be tracked back to a set of local constituent predictions $\hat{y}_1, \dots, \hat{y}_n \in \mathcal{Y}$.
-Under the assumption that each constituent $c_i$ represents some human interpretable object, a composition's prediction $\hat{y}_{C}$ can therefore be explained by the presence of the constituents/objects whose local predictions $\hat{y}_i$ are indicative of the global prediction.
+As seen in \cref{sec:related:lta}, an \ac{lta} score $y_{C} \in \mathcal{Y}$ for some multiset composition $C = \ldblbrace c_1, \dots, c_n \rdblbrace$ can always be tracked back to a set of local constituent scores $y_1, \dots, y_n \in \mathcal{Y}$.
+Under the assumption that each constituent $c_i$ represents some human interpretable object, a composition's score $y_{C}$ can therefore be explained by the presence of certain indicative constituents/objects $c_i$.
 
 Based on this intuition we now give a generalized definition of \ac{lta} which applies to unstructured as well as structured input data.
-We assume that compositions are represented by graphs $G \in \mathcal{G}$;
-unstructured inputs are represented by graphs with one vertex per constituent ($\mathcal{V}_G = \{ v_{c_1}, \dots, v_{c_n} \}$) and no edges ($\mathcal{E}_G = \emptyset$).
-Each composition has some target value $y_G \in \mathcal{Y}$ which could be a discrete class or continuous score.
-An \textit{\ac{lta} model} $h: \mathcal{G} \to \mathcal{Y}$ assigns predictions $\hat{y}_G \in \mathcal{Y}$ to compositions $G$.
+We assume that all compositions are represented by graphs $G \in \mathcal{G}$;
+an unstructured input is represented by a graph with one vertex per constituent ($\mathcal{V}_G = \{ v_{c_1}, \dots, v_{c_n} \}$) and no edges ($\mathcal{E}_G = \emptyset$).
+Each composition has some target score $y_G \in \mathcal{Y}$ which could be a discrete class or continuous value.
+An \textit{\ac{lta} model} $h: \mathcal{G} \to \mathcal{Y}$ assigns predictions $\hat{y}_G \in \mathcal{Y}$ to compositions $G$ which ideally correspond to the true score $y_G$.
 Such a model must satisfy three criteria:
@@
 	\item \textbf{Decomposition:}
-		A given composition $G$ must be decomposed into a set of constituents $c_{G,i}$ via a decomposition function $\varphi: \mathcal{G} \to \mathcal{P}(\mathcal{G})$ for which it holds that $\forall G \in \mathcal{G}: \forall c_{G,i} \in \varphi(G): \exists s \in \mathcal{V}_G^{*}: c_{G,i} \equiv G[s]$.
-		The decomposition function splits $G$ into its constituents $c_{G,i}$ which must be subgraphs of $G$.
+		A given composition $G$ must be decomposed into a set of constituents $c_{G,i}$ via a \textit{decomposition function} $\varphi$.
+		\begin{defn}
+			$\varphi: \mathcal{G} \to \mathcal{P}(\mathcal{G})$ is a \textit{decomposition function} iff.\ it splits a graph into a subset of its subgraphs, i.e.\ $\forall G \in \mathcal{G}: \forall c_{G,i} \in \varphi(G): \exists s \in \mathcal{V}_G^{*}: c_{G,i} \equiv G[s]$.
+		\end{defn}
 
@@
 		The constituents $c_{G, i} \in \varphi(G)$ must be evaluated via some function $f: \mathcal{G} \to \mathcal{Y} \times \mathbb{R}$.
-		This evaluation function assigns a prediction $\hat{y}_{G, i} \in \mathcal{Y}$ and a weight $w_{G, i} \in \mathbb{R}$ to each constituent.
-		A constituent's weight $w_{G, i}$ can intuitively be interpreted as a measure of the confidence that the local prediction $\hat{y}_{G, i}$ is indicative of the composition's global target value $y_G$.
+		This \textit{evaluation function} assigns a prediction $\hat{y}_{G, i} \in \mathcal{Y}$ and a weight $w_{G, i} \in \mathbb{R}$ to each constituent.
+		A constituent's weight $w_{G, i}$ can intuitively be interpreted as a measure of the confidence that the local prediction $\hat{y}_{G, i}$ is indicative of the composition's global target score $y_G$.
 		Learning local predictions and weights for all possible constituents is called the \textit{disaggregation problem}.
 
-		Note that there are no explicit constituent weights in the existing unstructured \ac{lta} approaches (i.e.\ all $w_{G, i} = 1$) because the constituents are explicitly given and assumed to be similarly indicative of $y_G$.
+		Note that there are no explicit constituent weights in the existing unstructured \ac{lta} approaches (i.e.\ implicitly all $w_{G, i} = 1$) because the explicitly given constituents are assumed to be equally indicative of $y_G$.
 		For structured data however, where the decomposition $\varphi(G)$ is not given as part of the input, this assumption does not necessarily hold.
+		By weighting the constituents, an \ac{lta} model can reduce the relevance or even ignore constituents that turn out to be irrelevant for the compositions target score $y_G$.
 	\item \textbf{Aggregation:}
-		Lastly a weighted aggregation function $\mathcal{A}: {(\mathcal{Y} \times \mathbb{R})}^{*} \to \mathcal{Y}$ must be learned.
+		Lastly a \textit{weighted aggregation function} $\mathcal{A}: {(\mathcal{Y} \times \mathbb{R})}^{*} \to \mathcal{Y}$ must be applied.
 		It combines the multiset of local constituent predictions and weights into a single global composition prediction.
+		\begin{defn}\label{defn:ltag:weighted-agg}
+			We call $\mathcal{A}$ a \textit{weighted aggregation function} iff.\ it satisfies
+			\begin{align*}
+				\text{idempotency: } & \forall y \in \mathcal{Y}, w \in \mathbb{R}_{> 0}^n: \mathcal{A}({\ldblbrace (y, w_i) \rdblbrace}_{i=1}^n) = y \\
+				\land\text{ zero invariance: } & \forall y_0 \in \mathcal{Y}, S = {\ldblbrace (y_i, w_i) \rdblbrace}_{i=1}^n: \mathcal{A}(S \cup \{ (y_0, 0) \}) = \mathcal{A}(S) \text{.}
+			\end{align*}
+		\end{defn}
+		One such weighted aggregator is the \textit{weighted mean} $\mean_w({\ldblbrace (y_i, w_i) \rdblbrace}_{i=1}^n) \coloneqq \sum_{i=1}^n w_i y_i$ which requires $\sum w_i = 1$ and $w_i \in [0, 1]$.
+		Alternatively an unweighted aggregation function like $\min$, $\max$ or \ac{owa} also trivially satisfies \cref{defn:ltag:weighted-agg} if all inputs $y_i$ with $w_i = 0$ are filtered out and the weights for all remaining inputs are ignored.
 \end{enumerate}
+Based on the notion of decomposition, disaggregation and aggregation we can now define the concept of \textit{\ac{lta} formulations}.
+\begin{defn}
+	A model $h: \mathcal{G} \to \mathcal{Y}$ is in an \textit{\ac{lta} formulation} iff.\ it is expressed as
+	\begin{align*}
+		h(G) \coloneqq \mathcal{A}(\ldblbrace f(c_{G,i})\, |\, c_{G,i} \in \varphi(G) \rdblbrace) \quad\text{with $\varphi$, $f$ and $\mathcal{A}$ as defined above.}
+	\end{align*}
+\end{defn}
+Note that every model $h: \mathcal{G} \to \mathcal{Y}$ has a trivial recursive \ac{lta} formulation by choosing $\varphi(G) = \{ G \}$, $f(G) = (h(G), 1)$ and an arbitrary weighted aggregation function $\mathcal{A}$.
+Those trivial \ac{lta} formulations do not split compositions into locally evaluated constituents and therefore do not fulfill the localized explainability characteristic of \ac{lta}.
 

### 2020-03-11 thesis/content/chapter-related-work.tex
continued ltag definition

@@
 \begin{defn}
-	A graph $G \coloneqq (\mathcal{V}_G, \mathcal{E}_G)$ consists of a finite set of vertices $v_i \in \mathcal{V}_G$ and a set of edges $e_{ij} = (v_i, v_j) \in \mathcal{E}_G \subseteq \mathcal{V}_G^2$.
+	A \textit{graph} $G \coloneqq (\mathcal{V}_G, \mathcal{E}_G)$ consists of a finite set of vertices $v_i \in \mathcal{V}_G$ and a set of edges $e_{ij} = (v_i, v_j) \in \mathcal{E}_G \subseteq \mathcal{V}_G^2$.
 	Optionally discrete vertex labels $l_G[v_i] \in L_{\mathcal{V}}$ or edge labels $l_G[e_{ij}] \in L_{\mathcal{E}}$ may be associated with all vertices $v_i \in \mathcal{V}_G$ and edges $e_{ij}\in \mathcal{E}_G$ respectively.

### 2020-03-12 thesis/content/chapter-related-work.tex
minor corrections

@@
 \begin{defn}
-	Two colorings $\chi$ and $\chi'$ are \textit{equivalent} ($\chi \equiv \chi'$) iff.\ $\chi \preceq \chi' \land \chi' \preceq \chi$, i.e.\ $\chi$ is identical to $\chi'$ up to color substitutions.
+	Two colorings $\chi$ and $\chi'$ are \textit{equivalent} ($\chi \equiv \chi'$) iff.\ $\chi \preceq \chi' \land \chi \succeq \chi'$, i.e.\ $\chi$ is identical to $\chi'$ up to color substitutions.
 \end{defn}
@@
 Note that the eigenvalues of $L_G$ grow with the number of edges\footnote{
-	This is trivially shown by $\sum_{i=1}^n \lambda_{G,i} = \Tr(t_G) = \Tr(D_G) = |\mathcal{E}_G|$.
+	This is trivially shown by $\sum_{i=1}^n \lambda_{G,i} = \Tr(L_G) = \Tr(D_G) = |\mathcal{E}_G|$.
 }.
@@
 Apart from this fundamental difference to 1-\acs{wl} bounded \acp{gnn}, $k$-\acp{gnn} additionally use two learnable parameter matrices $W^{(t)}, W_{\Gamma}^{(t)} \in \mathbb{R}^{d^{(t-1)} \times d^{(t)}}$ instead of the single $\Theta^{(t)}$ used in \acp{gcn}.
-This separation of parameter matrix effectively fulfills the same purpose as the \ac{mlp} in \ac{gin}; it makes the implicitly defined $k$-\acs{gnn} hash function injective.
+This separation of the parameter matrix effectively fulfills the same purpose as the \ac{mlp} in \ac{gin}; it makes the implicitly defined $k$-\acs{gnn} hash function injective.
 This can be seen by realizing that a stack of $k$-\acs{gnn} convolution layers with $W_{\Gamma}^{(t)} = 0$ for all but the first layer simulates a \ac{mlp}.

### 2020-03-12 thesis/content/chapter-graph-lta.tex
first ltag definition draft

@@
 Note that every model $h: \mathcal{G} \to \mathcal{Y}$ has a trivial recursive \ac{lta} formulation by choosing $\varphi(G) = \{ G \}$, $f(G) = (h(G), 1)$ and an arbitrary weighted aggregation function $\mathcal{A}$.
-Those trivial \ac{lta} formulations do not split compositions into locally evaluated constituents and therefore do not fulfill the localized explainability characteristic of \ac{lta}.
+Those trivial \ac{lta} formulations do not split compositions into locally evaluated constituents and therefore intuitively do not fulfill the postulated localized explainability characteristic of \ac{lta}.
+However since there is no commonly accepted formal criterion to decide whether a model's decisions are explainable~\cite{Lipton2018}, we do not attempt to strictly distinguish between \ac{lta} and non-\acs{lta} methods.
+Instead the notion of \ac{lta} formulations should be seen as way to identify how ``\acs{lta}-like'' a model is:
+\begin{itemize}
+	\item \textbf{Negative extreme:}
+		If a model $h$ only has trivial \ac{lta} formulations with the decomposition function $\varphi(G) = \{ G \}$, it is not considered to be an \ac{lta} model.
+	\item \textbf{Positive extreme:}
+		If a model has an \ac{lta} formulation with a decomposition function that returns interpretable constituents, it is considered to be an \ac{lta} model.
+		By definition this is true for the constituents $\varphi(G) \coloneqq {\{ G[v_{c_i}] \}}_{v_{c_i} \in \mathcal{V}_G}$ of \ac{lta} methods for unstructured data.
+	\item \textbf{In-between cases:}
+		An \ac{lta} method for structured data must produce models that lie somewhere in-between the two extremes.
+\end{itemize}
+
+The more ``\acs{lta}-like'' a given model is, the stronger its bias towards locally explainable predictions, which in turn reduces the potential expressive power of the model.
+\citet{Gilpin2018} describe this trade-off between explainability and expressive power in more detail.
+When considering problem domains in which the true composition scores $y_G$ are accurately described by an \acs{lta}-like generative process, a less expressive \ac{lta}-like model could however generalize better than a more expressive non-\ac{lta} model.
+This idea is captured by the so-called \textit{\ac{lta} assumption}.
+\begin{defn}
+	A problem domain $\mathcal{D}$ satisfies the \textit{\ac{lta} assumption} iff.\ there is an \ac{lta} method which produces models with an equal or lower out-of-sample error than the models produced by non-\acs{lta} methods for most training samples from $\mathcal{D}$.
+\end{defn}
+Due to the fuzziness of the class of \ac{lta} methods, the \ac{lta} assumption is of course also a fuzzy concept.
+Nonetheless evidence for its truthiness in a given domain can be empirically obtained by comparing candidate \ac{lta} methods with the best known non-\ac{lta} method, assuming that some cut-off condition for the required ``\ac{lta}-likeness'' of an \ac{lta} method is agreed upon.
 

### 2020-03-12 thesis/content/chapter-graph-lta.tex
various fixes

@@
 	\item We begin with a formal definition of what actually constitutes an \ac{lta} method as opposed to non-\acs{lta} methods.
-	\item Using this definition, we will see that some of the previously described \ac{gcnn} methods can be interpreted as \ac{lta} variants under certain conditions.
-	\item Finally a new \acs{lta}-inspired \ac{gcnn} architecture will be described.
+	\item Using this definition, we will see that some of the previously described \ac{gcr} methods can be interpreted as \ac{lta} variants under certain conditions.
+	\item Finally a novel \acs{lta}-inspired \ac{gcnn} architecture will be described.
 \end{enumerate}
@@
 	\item \textbf{Decomposition:}
-		A given composition $G$ must be decomposed into a set of constituents $c_{G,i}$ via a \textit{decomposition function} $\varphi$.
+		A given composition $G$ must be decomposed into a set of constituents $c_{G,i}$ via a \textit{decomposition function} $\psi: \mathcal{G} \to \mathcal{P}(\mathcal{G})$.
 		\begin{defn}
-			$\varphi: \mathcal{G} \to \mathcal{P}(\mathcal{G})$ is a \textit{decomposition function} iff.\ it splits a graph into a subset of its subgraphs, i.e.\ $\forall G \in \mathcal{G}: \forall c_{G,i} \in \varphi(G): \exists s \in \mathcal{V}_G^{*}: c_{G,i} \equiv G[s]$.
+			$\psi$ is a \textit{decomposition function} iff.\ it splits a graph into a subset of its subgraphs, i.e.\ $\forall G \in \mathcal{G}: \forall c_{G,i} \in \psi(G): \exists s \in \mathcal{V}_G^{*}: c_{G,i} \equiv G[s]$.
 		\end{defn}
 
-		In the existing unstructured \ac{lta} approaches the decomposition function is implicitly defined as $\varphi(G) \coloneqq {\{ G[v_{c_i}] \}}_{v_{c_i} \in \mathcal{V}_G}$ since each vertex $v_{c_i}$ corresponds to an interpretable constituent $c_i$ by definition.
+		In the existing unstructured \ac{lta} approaches the decomposition function is implicitly defined as $\psi(G) \coloneqq {\{ G[v_{c_i}] \}}_{v_{c_i} \in \mathcal{V}_G}$ since each vertex $v_{c_i}$ corresponds to an interpretable constituent $c_i$ by definition.
 		For structured data however, a split into individual vertices is typically not appropriate.
@@
 	\item \textbf{Disaggregation:}
-		The constituents $c_{G, i} \in \varphi(G)$ must be evaluated via some function $f: \mathcal{G} \to \mathcal{Y} \times \mathbb{R}$.
+		The constituents $c_{G, i} \in \psi(G)$ must be evaluated via some function $f: \mathcal{G} \to \mathcal{Y} \times \mathbb{R}$.
 		This \textit{evaluation function} assigns a prediction $\hat{y}_{G, i} \in \mathcal{Y}$ and a weight $w_{G, i} \in \mathbb{R}$ to each constituent.
@@
 		Note that there are no explicit constituent weights in the existing unstructured \ac{lta} approaches (i.e.\ implicitly all $w_{G, i} = 1$) because the explicitly given constituents are assumed to be equally indicative of $y_G$.
-		For structured data however, where the decomposition $\varphi(G)$ is not given as part of the input, this assumption does not necessarily hold.
+		For structured data however, where the decomposition $\psi(G)$ is not given as part of the input, this assumption does not necessarily hold.
 		By weighting the constituents, an \ac{lta} model can reduce the relevance or even ignore constituents that turn out to be irrelevant for the compositions target score $y_G$.
@@
 		\end{defn}
-		One such weighted aggregator is the \textit{weighted mean} $\mean_w({\ldblbrace (y_i, w_i) \rdblbrace}_{i=1}^n) \coloneqq \sum_{i=1}^n w_i y_i$ which requires $\sum w_i = 1$ and $w_i \in [0, 1]$.
+		One such weighted aggregator is the \textit{weighted mean} $\wmean({\ldblbrace (y_i, w_i) \rdblbrace}_{i=1}^n) \coloneqq \sum_{i=1}^n w_i y_i$ which requires $\sum w_i = 1$ and $w_i \in [0, 1]$.
 		Alternatively an unweighted aggregation function like $\min$, $\max$ or \ac{owa} also trivially satisfies \cref{defn:ltag:weighted-agg} if all inputs $y_i$ with $w_i = 0$ are filtered out and the weights for all remaining inputs are ignored.
@@
 	\begin{align*}
-		h(G) \coloneqq \mathcal{A}(\ldblbrace f(c_{G,i})\, |\, c_{G,i} \in \varphi(G) \rdblbrace) \quad\text{with $\varphi$, $f$ and $\mathcal{A}$ as defined above.}
+		h(G) \coloneqq \mathcal{A}(\ldblbrace f(c_{G,i})\, |\, c_{G,i} \in \psi(G) \rdblbrace) \quad\text{with $\psi$, $f$ and $\mathcal{A}$ as defined above.}
 	\end{align*}
 \end{defn}
-Note that every model $h: \mathcal{G} \to \mathcal{Y}$ has a trivial recursive \ac{lta} formulation by choosing $\varphi(G) = \{ G \}$, $f(G) = (h(G), 1)$ and an arbitrary weighted aggregation function $\mathcal{A}$.
+Note that every model $h: \mathcal{G} \to \mathcal{Y}$ has a trivial recursive \ac{lta} formulation by choosing $\psi(G) = \{ G \}$, $f(G) = (h(G), 1)$ and an arbitrary weighted aggregation function $\mathcal{A}$.
 Those trivial \ac{lta} formulations do not split compositions into locally evaluated constituents and therefore intuitively do not fulfill the postulated localized explainability characteristic of \ac{lta}.
@@
 	\item \textbf{Negative extreme:}
-		If a model $h$ only has trivial \ac{lta} formulations with the decomposition function $\varphi(G) = \{ G \}$, it is not considered to be an \ac{lta} model.
+		If a model $h$ only has trivial \ac{lta} formulations with the decomposition function $\psi(G) = \{ G \}$, it is not considered to be an \ac{lta} model.
 	\item \textbf{Positive extreme:}
 		If a model has an \ac{lta} formulation with a decomposition function that returns interpretable constituents, it is considered to be an \ac{lta} model.
-		By definition this is true for the constituents $\varphi(G) \coloneqq {\{ G[v_{c_i}] \}}_{v_{c_i} \in \mathcal{V}_G}$ of \ac{lta} methods for unstructured data.
+		By definition this is true for the constituents $\psi(G) \coloneqq {\{ G[v_{c_i}] \}}_{v_{c_i} \in \mathcal{V}_G}$ of \ac{lta} methods for unstructured data.
 	\item \textbf{In-between cases:}
@@
 \end{defn}
-Due to the fuzziness of the class of \ac{lta} methods, the \ac{lta} assumption is of course also a fuzzy concept.
-Nonetheless evidence for its truthiness in a given domain can be empirically obtained by comparing candidate \ac{lta} methods with the best known non-\ac{lta} method, assuming that some cut-off condition for the required ``\ac{lta}-likeness'' of an \ac{lta} method is agreed upon.
+Due to the fuzziness of the class of \ac{lta} methods, the \ac{lta} assumption is naturally also a fuzzy concept.
+Nonetheless evidence for its truthiness in a given domain $\mathcal{D}$ can be empirically obtained by comparing candidate \ac{lta} methods with the best known non-\ac{lta} method for $\mathcal{D}$, assuming that some cut-off condition for the required ``\ac{lta}-likeness'' of an \ac{lta} method is agreed upon.
 
-\section{An \acs*{lta} Interpretation of Existing \acsp*{gcnn}}%
-\label{sec:ltag:interpretation}
+\section{\acs*{lta} Formulations of Existing \acs*{gcr} Methods}%
+\label{sec:ltag:formulation}
+
+Based on the general definition of \ac{lta} from the last section, we will now see to what extent the \ac{gcr} methods described in \cref{sec:related:gcr} can be interpreted as \ac{lta} instances.
+First the relation between \acp{svm} on graph embeddings and \ac{lta} will be explored.
+Then we will provide an \ac{lta} perspective on \acp{gcnn}.
+
+\subsection{\acsp*{svm} on Graph Embeddings as \acs*{lta} Models}%
+\label{sec:ltag:formulation:svm}
+
+In \cref{sec:related:gcr:embed,sec:related:gcr:kernel} three different ways to map a given graph $G$ to a vector representation $\varphi(G) \in \mathbb{R}^d$ were described:
+\begin{enumerate*}
+	\item Fingerprint embeddings,
+	\item skip-gram inspired embeddings and
+	\item kernel embeddings
+\end{enumerate*}.
+
+\subsection{\acsp*{gcnn} as \acs*{lta} Models}%
+\label{sec:ltag:formulation:gcnn}
 

### 2020-03-12 thesis/content/chapter-related-work.tex
various fixes

@@
 The basic idea of explicit graph embedding approaches is to map a graph $G \in \mathcal{G}$ to some vector in a finite vector space $\mathcal{X} = \mathbb{R}^d$.
-A function $f: \mathcal{G} \to \mathcal{X}$ is called a \textit{graph embedding function}.
+A function $\varphi: \mathcal{G} \to \mathcal{X}$ is called a \textit{graph embedding function}.
 By embedding a graph into $\mathcal{X}$, any classification or regression algorithm that works with vectors can then be applied to solve the \ac{gcr} problem.
@@
 The so-called vertex embedding problem is closely related to the graph embedding problem.
-As the name suggests, a \textit{vertex embedding function} $f_G$ maps all vertices $v \in \mathcal{V}_G$ to $\mathcal{X}$.
-The embedding vector $f_G(v)$ ideally encodes relevant information about a vertex and its structural position in $G$.
+As the name suggests, a \textit{vertex embedding function} $\varphi_G$ maps all vertices $v \in \mathcal{V}_G$ to $\mathcal{X}$.
+The embedding vector $\varphi_G(v)$ ideally encodes relevant information about a vertex and its structural position in $G$.
 It can be used to solve the vertex classification and regression problem via arbitrary \ac{ml} methods for vectors.
@@
 So-called \textit{fingerprint embeddings} try to match a fixed set of subgraphs $S_1, \dots, S_d$ to the input graph.
-The embedding of a graph $G$ is a binary vector $f(G) = x \in {\{0, 1\}}^d$ with $x_i \coloneqq \mathit{contain}_{S_i}(G)$ (see~\cref{defn:related:wl-count-detect}).
+The embedding of a graph $G$ is a binary vector $\varphi(G) = x \in {\{0, 1\}}^d$ with $x_i \coloneqq \mathit{contain}_{S_i}(G)$ (see~\cref{defn:related:wl-count-detect}).
 This simple approach usually requires a careful choice of subgraphs but can still be competitive with the other more recent approaches we will look at in the following sections.
-Fingerprint embeddings are for example used in multiple state-of-the-art toxicity prediction tools like RASAR~\cite{Luechtefeld2018}\cite{ToxTrack}, ProTox~\cite{Drwal2014}\cite{Banerjee2018}\cite{ProTox} or the \ac{toxTest}~\cite{TEST}.
+Fingerprint embeddings are for example used in multiple state-of-the-art toxicity prediction tools like RASAR~\cite{Luechtefeld2018}\cite{ToxTrack}, ProTox~\cite{Drwal2014}\cite{ProTox}\cite{Banerjee2018} or the \ac{toxTest}~\cite{TEST}.
 
@@
 Let us first look at the \texttt{word2vec} skip-gram method.
-It gets a sequence of words $(w_0, \dots, w_n)$ as input and outputs embedding vectors $f(w_0), \dots, f(w_n) \in \mathbb{R}^d$.
-To do this the context $C_k(w_i) = \{ w_{i-k}, \dots, w_{i + k} \}$ is computed for all words where $w_i$ is the so-called \textit{context root}.
+It gets a sequence of words $(w_0, \dots, w_n)$ as input and outputs embedding vectors $\varphi(w_0), \dots, \varphi(w_n) \in \mathbb{R}^d$.
+To do this the context $\Gamma_k(w_i) = \{ w_{i-k}, \dots, w_{i + k} \}$ is computed for all words where $w_i$ is the so-called \textit{context root}.
 The word contexts are then used to optimize the following log-likelihood objective:
 \begin{gather}
-	\max_{f, f_C} \sum_{i = 1}^n \log P(C_k(w_i) | w_i)\ =\ \max_{f, f_C} \sum_{i = 1}^n \smashoperator[r]{\sum_{w_j \in C_k(w_i)}}\, \left[ {f(w_i)}^\top f_C(w_j) - \log Z_{w_i} \right] \label{eq:related:word2vec}\\
-	\text{with } P(C_k(w_i) | w_i) \coloneqq \prod_{\mathclap{w_j \in C_k(w_i)}}\, \overbrace{\frac{\exp\left({f(w_i)}^\top f_C(w_j)\right)}{Z_{w_i}}}^{P(w_j | w_i)}
-	\text{ and } Z_{w_i} \coloneqq \sum_{j = 1}^n \exp\left({f(w_i)}^\top f_C(w_j)\right) \nonumber
+	\max_{\varphi, \varphi_{\Gamma}} \sum_{i = 1}^n \log P(\Gamma_k(w_i) | w_i)\ =\ \max_{\varphi, \varphi_{\Gamma}} \sum_{i = 1}^n \smashoperator[r]{\sum_{w_j \in \Gamma_k(w_i)}}\, \left[ {\varphi(w_i)}^\top \varphi_{\Gamma}(w_j) - \log Z_{w_i} \right] \label{eq:related:word2vec}\\
+	\text{with } P(\Gamma_k(w_i) | w_i) \coloneqq \prod_{\mathclap{w_j \in \Gamma_k(w_i)}}\, \overbrace{\frac{\exp\left({\varphi(w_i)}^\top \varphi_{\Gamma}(w_j)\right)}{Z_{w_i}}}^{P(w_j | w_i)}
+	\text{ and } Z_{w_i} \coloneqq \sum_{j = 1}^n \exp\left({\varphi(w_i)}^\top \varphi_{\Gamma}(w_j)\right) \nonumber
 \end{gather}
-\texttt{word2vec} essentially uses an expectation maximization scheme to maximize the probabilities $P(w_j | w_i)$ of observing the context words $w_j \in C_k(w_i)$ of all words $w_i$.
-Those probabilities are described by the overlap of the embeddings $f(w_i)$ of words $w_i$ and the embeddings $f_C(w_j)$ of their context words $w_j$.
+\texttt{word2vec} essentially uses an expectation maximization scheme to maximize the probabilities $P(w_j | w_i)$ of observing the context words $w_j \in \Gamma_k(w_i)$ of all words $w_i$.
+Those probabilities are described by the overlap of the embeddings $\varphi(w_i)$ of words $w_i$ and the embeddings $\varphi_{\Gamma}(w_j)$ of their context words $w_j$.
 Intuitively this means that words with similar contexts will be mapped close to each other in the embedding space.
-Note that \texttt{word2vec} actually finds two embeddings $f(w)$ and $f_C(w)$ for each word of which only the first is returned.
-The two embeddings represent two different perspectives on words: $f$ describes a word $w_i$ as the root of a context $C_k(w_i)$, $f_C$ on the other hand describes a word $w_j$ as part of a context $C_k(w_i) \ni w_j$.
+Note that \texttt{word2vec} actually finds two embeddings $\varphi(w)$ and $\varphi_{\Gamma}(w)$ for each word of which only the first is returned.
+The two embeddings represent two different perspectives on words: $\varphi$ describes a word $w_i$ as the root of a context $\Gamma_k(w_i)$, $\varphi_{\Gamma}$ on the other hand describes a word $w_j$ as part of a context $\Gamma_k(w_i) \ni w_j$.
 
 \paragraph{Vertex Embeddings}
-Skip-gram embeddings can be  na{\"\i}vely extended to graphs by realizing that \texttt{word2vec} effectively already is a vertex embedding method for linear graphs in which $C_k(v)$ is simply the $k$-neighborhood of the vertex/word $v$.
+Skip-gram embeddings can be  na{\"\i}vely extended to graphs by realizing that \texttt{word2vec} effectively already is a vertex embedding method for linear graphs in which $\Gamma_k(v)$ is simply the $k$-neighborhood of the vertex/word $v$.
 The problem with this na{\"\i}ve extension is that the sizes of $k$-neighborhoods in arbitrary graphs can be much larger and often tend to grow exponentially with $k$.
@@
 Both methods only differ in the transition matrix that is used for the random walk.
-Another difference of DeepWalk and \texttt{node2vec} compared to \texttt{word2vec} is the so-called \textit{feature space symmetry} which states that the context root interpretation ($f$) of a vertex should be symmetric to its context element interpretation ($f_C$), i.e.\ $f = f_C$.
+Another difference of DeepWalk and \texttt{node2vec} compared to \texttt{word2vec} is the so-called \textit{feature space symmetry} which states that the context root interpretation ($\varphi$) of a vertex should be symmetric to its context element interpretation ($\varphi_{\Gamma}$), i.e.\ $\varphi = \varphi_{\Gamma}$.
 The combination of random walk context sampling and and the feature space symmetry assumption can be used to compute vertex embeddings even for very large graphs.
@@
 It is inspired by \texttt{doc2vec}~\cite{Le2014} which in turn is based on \texttt{word2vec}.
-\texttt{graph2vec} gets a set of graphs $\mathcal{G} = \{ G_1, \dots G_N \}$ as input and outputs graph embeddings $f(G_1), \dots, f(G_N)$.
+\texttt{graph2vec} gets a set of graphs $\mathcal{G} = \{ G_1, \dots G_N \}$ as input and outputs graph embeddings $\varphi(G_1), \dots, \varphi(G_N)$.
 While in \texttt{word2vec} every word can be a context root as well as a context element, \texttt{graph2vec} uses the graphs $\mathcal{G}$ as context roots.
-The context $C_k(G_i)$ of a graph $G_i$ is defined as
+The context $\Gamma_T(G_i)$ of a graph $G_i$ is defined as
 \begin{equation}
-	C_k(G_i) \coloneqq \bigcup_{\mathclap{v_j \in \mathcal{V}_{G_i}}}\, {\left\{ \chi_{G_i, 1}^{(t)}(v_j) \right\}}_{l = 0}^L % chktex 21
-	\quad\text{with $k \in \mathbb{N}$ and $\chi_{G_i, 1}^{(t)}$ as in \cref{defn:related:wl1-refine}.}
+	\Gamma_T(G_i) \coloneqq \bigcup_{\mathclap{v_j \in \mathcal{V}_{G_i}}}\, {\left\{ \chi_{G_i, 1}^{(t)}(v_j) \right\}}_{t = 0}^T % chktex 21
+	\quad\text{with $T \in \mathbb{N}$ and $\chi_{G_i, 1}^{(t)}$ as in \cref{defn:related:wl1-refine}.}
 \end{equation}
-Intuitively this context can be understood as the set of 1-\acs{wl}-distinguishable subgraphs of $G_i$ with diameter $\leq 2L$.
+Intuitively this context can be understood as the set of 1-\acs{wl}-distinguishable subgraphs of $G_i$ with diameter $\leq 2T$.
 Since \ac{wl} is used to identify distinct subgraphs, \texttt{graph2vec} can only be applied to graphs with discrete vertex labels.
-Using the previous definitions, the context root embedding function has the signature $f: \mathcal{G} \to \mathbb{R}^d$ while the context element embedding function is of type $f_C: \bigcup_{G_i \in \mathcal{G}} C_k(G_i) \to \mathbb{R}^d$.
+Using the previous definitions, the context root embedding function has the signature $\varphi: \mathcal{G} \to \mathbb{R}^d$ while the context element embedding function is of type $\varphi_{\Gamma}: \bigcup_{G_i \in \mathcal{G}} \Gamma_T(G_i) \to \mathbb{R}^d$.
 To find those embeddings the \texttt{word2vec} objective from \cref{eq:related:word2vec} is reused.
@@
 }.
-By fixing $K = 2$, by using a single filter parameter $\theta \in \mathbb{R}$ and by assuming that $\max{\lambda_G^{\text{sym}}} \approx 2$ (see \cref{defn:related:laplacian-sym}, \cpageref{defn:related:laplacian-sym}) the convolution operation is reduced to
+By fixing $K = 2$, by using a single filter parameter $\theta \in \mathbb{R}$ and by assuming that $\max{\lambda_G^{\text{sym}}} \approx 2$ (see \cfullref{defn:related:laplacian-sym}) the convolution operation is reduced to
 \begin{align}
@@
 From this perspective a \ac{gcn} can be seen as a so-called spatial \ac{gcnn} because it operates directly in the vertex domain.
-To see why this is the case, we rewrite \cref{eq:related:gcn-layer} and compare it with the 1-\acs{wl} color refinement strategy (see \cref{defn:related:wl1-refine}, \cpageref{defn:related:wl1-refine}):
+To see why this is the case, we rewrite \cref{eq:related:gcn-layer} and compare it with the 1-\acs{wl} color refinement strategy (see \cfullref{defn:related:wl1-refine}):
 \begin{align*}
@@
 \paragraph{Higher dimensional \ac{wl} \acp{gnn}}
-The majority of \ac{gnn} methods uses a neighborhood aggregation scheme which limits their discriminative and, more importantly, computational power to that of 1-\acs{wl} (see \cref{defn:related:wl-count-detect}, \cpageref{defn:related:wl-count-detect}).
+The majority of \ac{gnn} methods uses a neighborhood aggregation scheme which limits their discriminative and, more importantly, computational power to that of 1-\acs{wl} (see \cfullref{defn:related:wl-count-detect}).
 To go beyond the limits of 1-\acs{wl}, \citet{Morris2019} proposed the first \ac{gcnn} architecture inspired by the higher dimensional $k$-\acs{wl} algorithms.
@@
 
-As a final remark, note that the concept of ``neighborhood'' in \cref{eq:related:kgnn-layer} is slightly different from that used by the $k$-\acs{wl} algorithm (see \cref{defn:related:wlk-refine}, \cpageref{defn:related:wlk-refine}):
+As a final remark, note that the concept of ``neighborhood'' in \cref{eq:related:kgnn-layer} is slightly different from that used by the $k$-\acs{wl} algorithm (see \cfullref{defn:related:wlk-refine}):
 $k$-\acs{gnn} considers a single $k$-multiset as a neighbor while $k$-\acs{wl} uses $k$-sets of $k$-multisets as neighbors.

### 2020-03-13 thesis/content/chapter-graph-lta.tex
started with svm lta formulation proof

@@
 		A given composition $G$ must be decomposed into a set of constituents $c_{G,i}$ via a \textit{decomposition function} $\psi: \mathcal{G} \to \mathcal{P}(\mathcal{G})$.
-		\begin{defn}
-			$\psi$ is a \textit{decomposition function} iff.\ it splits a graph into a subset of its subgraphs, i.e.\ $\forall G \in \mathcal{G}: \forall c_{G,i} \in \psi(G): \exists s \in \mathcal{V}_G^{*}: c_{G,i} \equiv G[s]$.
+		\begin{defn}\label{defn:ltag:decomp}
+			$\psi$ is a \textit{decomposition function} iff.\ it splits a graph into a subset of its subgraphs, i.e.\ $\forall G \in \mathcal{G}: \forall c_{G,i} \in \psi(G): \exists s \in \mathcal{P}(\mathcal{V}_G): c_{G,i} = G[s]$.
 		\end{defn}
+		Note that the strict equality $c_{G,i} = G[s]$ is used in \cref{defn:ltag:decomp} instead of just requiring subgraph isomorphism ($c_{G,i} \simeq G[s]$) because a constituent $c_{G,i}$ represents a specific localized subgraph of a structured composition.
 
@@
 	\item \textbf{Aggregation:}
-		Lastly a \textit{weighted aggregation function} $\mathcal{A}: {(\mathcal{Y} \times \mathbb{R})}^{*} \to \mathcal{Y}$ must be applied.
-		It combines the multiset of local constituent predictions and weights into a single global composition prediction.
+		Lastly a \textit{weighted aggregation function} $\mathcal{A}: {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*} \to \mathcal{Y}$ must be applied.
+		It combines the multiset of local constituent predictions and non-negative weights into a single global composition prediction.
 		\begin{defn}\label{defn:ltag:weighted-agg}
@@
 			\begin{align*}
-				\text{idempotency: } & \forall y \in \mathcal{Y}, w \in \mathbb{R}_{> 0}^n: \mathcal{A}({\ldblbrace (y, w_i) \rdblbrace}_{i=1}^n) = y \\
-				\land\text{ zero invariance: } & \forall y_0 \in \mathcal{Y}, S = {\ldblbrace (y_i, w_i) \rdblbrace}_{i=1}^n: \mathcal{A}(S \cup \{ (y_0, 0) \}) = \mathcal{A}(S) \text{.}
+				\text{idempotency: } & \exists c \geq 0: \forall y \in \mathcal{Y}, w \in \mathbb{R}_{\geq 0}^n\text{ s.t. }{\max w} \geq c: \mathcal{A}({\ldblbrace (y, w_i) \rdblbrace}_{i=1}^n) = y \\
+				\land\text{ zero invar.: } & \forall y_0 \in \mathcal{Y}, S = {\ldblbrace (y_i, w_i) \rdblbrace}_{i=1}^n: \mathcal{A}(S \cup \{ (y_0, 0) \}) = \mathcal{A}(S) \text{.}
 			\end{align*}
 		\end{defn}
-		One such weighted aggregator is the \textit{weighted mean} $\wmean({\ldblbrace (y_i, w_i) \rdblbrace}_{i=1}^n) \coloneqq \sum_{i=1}^n w_i y_i$ which requires $\sum w_i = 1$ and $w_i \in [0, 1]$.
-		Alternatively an unweighted aggregation function like $\min$, $\max$ or \ac{owa} also trivially satisfies \cref{defn:ltag:weighted-agg} if all inputs $y_i$ with $w_i = 0$ are filtered out and the weights for all remaining inputs are ignored.
+		{\setlength{\parskip}{0pt}The idempotency constraint requires aggregation functions to agree with uniform input scores $y$ if at least one input is weighted above some threshold $c$.
+		The zero invariance constraint requires aggregation functions to ignore inputs with zero weight.}
+
+		One example for such a weighted aggregator is the \textit{weighted mean} function $\wmean({\ldblbrace (y_i, w_i) \rdblbrace}_{i=1}^n) \coloneqq \sum_{i=1}^n w_i y_i$ which requires $\sum w_i = 1$ and $w_i \in [0, 1]$.
+		Alternatively an unweighted aggregation function like $\min$, $\max$ or \ac{owa}\footnote{
+			Even though \ac{owa} uses weights, it does not get those weights as part of the input and is therefore considered to be unweighted here.
+		} also trivially satisfies \cref{defn:ltag:weighted-agg} if all inputs $y_i$ with $w_i = 0$ are filtered out and the weights for all remaining inputs are ignored.
 \end{enumerate}
@@
 Due to the fuzziness of the class of \ac{lta} methods, the \ac{lta} assumption is naturally also a fuzzy concept.
-Nonetheless evidence for its truthiness in a given domain $\mathcal{D}$ can be empirically obtained by comparing candidate \ac{lta} methods with the best known non-\ac{lta} method for $\mathcal{D}$, assuming that some cut-off condition for the required ``\ac{lta}-likeness'' of an \ac{lta} method is agreed upon.
+Nonetheless evidence for its truthiness in a given domain $\mathcal{D}$ can be empirically obtained by comparing candidate \ac{lta} methods with the best known non-\ac{lta} method for $\mathcal{D}$, assuming that some cut-off condition for the required ``\ac{lta}-ness'' of an \ac{lta} method is agreed upon.
 
@@
 
-In \cref{sec:related:gcr:embed,sec:related:gcr:kernel} three different ways to map a given graph $G$ to a vector representation $\varphi(G) \in \mathbb{R}^d$ were described:
+In \cref{sec:related:gcr:embed,sec:related:gcr:kernel} three different ways to map a given graph $G$ to a vector $\varphi(G) \in \mathbb{R}^d$ were described:
 \begin{enumerate*}
@@
 \end{enumerate*}.
+One common way to solve the \ac{gcr} problem via those embedding vectors is to train an \ac{svm} on them.
+We will now see that \acp{svm} can be interpreted as \ac{lta} models if they are trained on so-called \textit{substructure component embeddings}.
+\begin{defn}\label{defn:ltag:substruct-embedding}
+	A graph embedding $\varphi: \mathcal{G} \to \mathbb{N}^{d}$ is called a \textit{substructure component embedding} iff.\ %
+	$\forall i \in [d]: \exists S_i \in \mathcal{G}_{\mathit{ord}}, \gamma_i \in \mathbb{R}: \forall G \in \mathcal{G}: {{\varphi(G)}[i] = \mathit{contains}_{S_i}(G)} \lor {{\varphi(G)}[i] = \gamma_i \mathit{count}_{S_i}(G)}$,
+	with $\mathit{contains}$ and $\mathit{count}$ as in \cfullref[ on ]{defn:related:wl-count-detect}.
+\end{defn}
+Intuitively \cref{defn:ltag:substruct-embedding} states that every embedding dimension must represent the count or presence of some ordered subgraph $S_i$ in the embedded graph $G$.
+Based on this requirement we now proof the main theorem which shows the relation between \acp{svm} and \ac{lta}.
+\begin{thm}
+	A binary \ac{svm} classifier that takes substructure component embeddings $\varphi(G)$ as its input has an \ac{lta} formulation with a nontrivial decomposition function $\psi$.
+\end{thm}
+\begin{proof}
+	Let $h: \mathcal{G} \to {\{ -1, +1 \}}$ be a binary graph classifier expressed as $h = h_{\text{\acs*{svm}}} \circ \varphi$ where $\varphi: \mathcal{G} \to \mathbb{N}_0^{d}$ is a substructure component embedding and $h_{\text{\acs*{svm}}}: \mathbb{R}^{d} \to {\{ -1, +1 \}}$ an \ac{svm} classifier.
+	If $h$ is trained on a dataset $\mathcal{D}_{\text{train}} = {\{ (G_1, y_1), \dots, (G_N, y_N) \}}$, via the kernel trick it can be expressed as
+	{\allowdisplaybreaks\begin{align*}
+		h(G) &= \sgn\left( \sum_{j = 1}^{N} \alpha_j y_j {\langle \varphi(G), \varphi(G_j) \rangle} + b \right)
+		\quad\text{for some $\alpha \in \mathbb{R}^N$ and $b \in \mathbb{R}$} \\
+		&= \sgn\left( {\sum_{i=1}^{d} {\varphi(G)}[i] \underbrace{\left( \sum_{j = 1}^{N} \alpha_j y_j {\varphi(G_j)}[i] \right)}_{{} \eqqcolon z_i}} + b \right)
+		\text{.}
+	\end{align*}}
+
+	By \cref{defn:ltag:substruct-embedding} each $\varphi(G)[i] > 0$ counts or indicates the presence of some ordered subgraph $S_i$ in $G$;
+	more precisely:
+	\begin{itemize}
+		\item If $\varphi(G)[i] = \gamma_i \mathit{count}_{S_i}(G) > 0$, let $C_i \coloneqq {\{ G[s]\, |\, s \in \mathcal{P}(\mathcal{V}_G) \land G[s] \simeq S_i \}}$ be the $\mathit{count}_{S_i}(G)$ subgraphs counted by the $i$-th dimension.
+		\item If $\varphi(G)[i] = \mathit{contains}_{S_i}(G) = 1$, let $C_i \coloneqq {\{  \}}$.
+	\end{itemize}
+	A substructure component embedding vector can therefore be interpreted as an encoding of the graph decomposition \[
+		\psi(G) \coloneqq \bigcup_{i=1}^{d}
+		\text{.}
+	\]
+	\begin{align*}
+		h(G) &= \sgn\left( {\sum_{i=1}^{d} {\varphi(G)}[i] \underbrace{|z_i|}_{{} \eqqcolon \nu_i} \underbrace{\sgn(z_i)}_{{} \eqqcolon y_i}} + b \right)
+	\end{align*}
+\end{proof}
 

### 2020-03-13 thesis/content/chapter-related-work.tex
started with svm lta formulation proof

@@
 	\item \textbf{Aggregation:}
-		A variadic aggregation function $\mathcal{A}: \mathcal{Y}^{*} \to \mathcal{Y}$ that estimates composite scores has to be learned, i.e.\@ $y_i \approx \hat{y} = \mathcal{A}(y_{1}, \dots, y_{n})$.
+		A variadic aggregation function $\mathcal{A}: \mathcal{Y}^{*} \to \mathcal{Y}$ that estimates composite scores has to be learned, i.e.\@ $y \approx \hat{y} = \mathcal{A}(y_{1}, \dots, y_{n})$.
 		Typically the aggregation function $\mathcal{A}$ should be associative and commutative to fit with the multiset-structure of compositions.
@@
 	Then $G[v] \coloneqq (\hat{v}, \mathcal{E}_G \cap \hat{v}^2, v)$ is called the \textit{ordered subgraph} of $G$ induced by $v$.
+	We denote the set of all ordered subgraphs with $\mathcal{G}_{\mathit{ord}}$.
 	If $\hat{v} = \mathcal{V}_G$ and $k = |\mathcal{V}_G|$, we call $v$ an \textit{ordering} of $G$ and $G[v]$ an \textit{encoding} of $G$.
@@
 
-The \acf{wl} algorithm~\cite{Weisfeiler1968}\cite{Cai1992} characterizes a graph $G$ by assigning discrete labels $c \in \mathcal{C}$, called \textit{colors}, to vertex $k$-tuples $(v_1, \dots, v_k) \in \mathcal{V}_G^k$, where $k \in \mathbb{N}$ is the freely choosable \textit{\ac{wl}-dimension}.
+The \acf{wl} algorithm~\cite{Weisfeiler1968}\cite{Cai1992} characterizes a graph $G$ by assigning discrete labels $c \in \mathcal{C}$, called \textit{colors}, to vertex $k$-tuples $(v_1, \dots, v_k) \in \mathcal{V}_G^k$, where $k \in \mathbb{N}_0$ is the freely choosable \textit{\ac{wl}-dimension}.
 A mapping $\chi_{G, k}: \mathcal{V}_G^k \to \mathcal{C}$ is called a \textit{$k$-coloring} of $G$.
@@
 \begin{defn}
-	The color distribution $\mathit{dist}_{\chi_{G, k}}: \mathcal{C} \to \mathbb{N}$ of a $k$-coloring $\chi_{G, k}$ counts each color $c \in \mathcal{C}$ in the coloring, i.e.\ $\mathit{dist}_{\chi_{G, k}}(c) \coloneqq \left|\left\{ v \in \mathcal{V}_G^k\, |\, \chi_{G, k}(v) = c \right\}\right|$.
+	The color distribution $\mathit{dist}_{\chi_{G, k}}: \mathcal{C} \to \mathbb{N}_0$ of a $k$-coloring $\chi_{G, k}$ counts each color $c \in \mathcal{C}$ in the coloring, i.e.\ $\mathit{dist}_{\chi_{G, k}}(c) \coloneqq \left|\left\{ v \in \mathcal{V}_G^k\, |\, \chi_{G, k}(v) = c \right\}\right|$.
 % chktex 21
@@
 In practice the hash function $h$ is usually defined lazily by using $\mathcal{C} \subseteq \mathbb{N}$ and enumerating color combinations in whichever order they are hashed s.t.\ a new color is introduced every time a previously unseen color combination appears at runtime\footnote{
-	Note that, even though $\mathcal{C}$ is formally countably infinite, we will assume that it only contains the finite number of colors occurring in a given finite graph dataset.
+	Though $\mathcal{C}$ is formally countably infinite, we assume it only contains the finite no.\ of colors in a dataset.
 }.
@@
 In 1-\acs{wl} a vertex coloring by itself cannot encode adjacency which is why this information is explicitly incorporated in each refinement step via $\Gamma_G$ (see~\cref{defn:related:wl1-refine}).
-In $k$-\acs{wl} on the other hand each pair of vertices $(v, u) \in \mathcal{V}_G^2$ appears in at least one $k$-tuple (assuming $k \geq 2$) and therefore has at least one color which can implicitly encode the adjacency information.
+In $k$-\acs{wl} on the other hand, each pair of vertices $(v, u) \in \mathcal{V}_G^2$ appears in at least one $k$-tuple (assuming $k \geq 2$) and therefore has at least one color which can implicitly encode the adjacency information.
 Edges and non-edges are colored differently in the initial coloring since $G[(v, u)] \not\equiv G[(w, u)]$ if $(v, u) \in \mathcal{E}_G$ but $(w, u) \notin \mathcal{E}_G$;
@@
 \subsubsection{Discriminative Power of \acs{wl}}
-Now we will look at the types of graphs that can be distinguished by \acs{wl} in relation to the \acs{wl}-dimension $k$.
+Now we will look at the types of graphs that can be distinguished by \ac{wl} in relation to the \acs{wl}-dimension $k$.
+We begin by showing that the power of \ac{wl} grows with $k$.
 \begin{lem}\label{lem:related:wl-lower}
@@
 \begin{defn}\label{defn:related:wl-count-detect}
-	$k$-\acs{wl} \textit{counts} a certain subgraph $S$ iff.\ it computes $\mathit{count}_S(G) \coloneqq \left|\{ \hat{v}\, |\, v \in \mathcal{V}_G^{*} \land G[v] \equiv S \}\right|$ (see~\cref{defn:related:ordered-subgraph}).
-	Similarly $k$-\acs{wl} \textit{detects} a subgraph $S$ iff.\ it computes $\mathit{contain}_S(G) \coloneqq \mathbbm{1}[\mathit{count}_S(G) > 0]$, with $\mathbbm{1}$ denoting the indicator function.
+	$k$-\acs{wl} \textit{counts} a certain ordered subgraph $S \in \mathcal{G}_{\mathit{ord}}$ iff.\ it computes $\mathit{count}_S(G) \coloneqq \left|\{ \hat{v}\, |\, v \in \mathcal{V}_G^{*} \land G[v] \equiv S \}\right|$ (see~\cref{defn:related:ordered-subgraph}).
+	Similarly $k$-\acs{wl} \textit{detects} an ordered subgraph $S$ iff.\ it computes $\mathit{contain}_S(G) \coloneqq \mathbbm{1}[\mathit{count}_S(G) > 0]$, with $\mathbbm{1}$ denoting the indicator function.
 \end{defn}
@@
 It is inspired by \texttt{doc2vec}~\cite{Le2014} which in turn is based on \texttt{word2vec}.
-\texttt{graph2vec} gets a set of graphs $\mathcal{G} = \{ G_1, \dots G_N \}$ as input and outputs graph embeddings $\varphi(G_1), \dots, \varphi(G_N)$.
-While in \texttt{word2vec} every word can be a context root as well as a context element, \texttt{graph2vec} uses the graphs $\mathcal{G}$ as context roots.
+\texttt{graph2vec} gets a set of graphs $\mathcal{G}' = \{ G_1, \dots G_N \}$ as input and outputs graph embeddings $\varphi(G_1), \dots, \varphi(G_N)$.
+While in \texttt{word2vec} every word can be a context root as well as a context element, \texttt{graph2vec} uses the graphs $\mathcal{G}'$ as context roots.
 The context $\Gamma_T(G_i)$ of a graph $G_i$ is defined as
@@
 Extending the idea of the \ac{wl} subtree kernel, the \textit{\ac{wl} shortest path kernel}~\cite{Shervashidze2011}\cite{Borgwardt2005} adds global structural information to the local subtree comparisons.
-The $i$-th component of a graph's shortest path vector $\varphi_{\text{SP}}(G)$ corresponds to the 4-tuple $(t_i, a_i, b_i, d_i) \in {\{ 0, \dots, T \}} \times \mathcal{C} \times \mathcal{C} \times \mathbb{N}$ and its value is described by
+The $i$-th component of a graph's shortest path vector $\varphi_{\text{SP}}(G)$ corresponds to the 4-tuple $(t_i, a_i, b_i, d_i) \in {\{ 0, \dots, T \}} \times \mathcal{C} \times \mathcal{C} \times \mathbb{N}_0$ and its value is described by
 \begin{align}

### 2020-03-14 thesis/content/chapter-graph-lta.tex
continued svm lta proof + minor fixes

@@
 			\begin{align*}
-				\text{idempotency: } & \exists c \geq 0: \forall y \in \mathcal{Y}, w \in \mathbb{R}_{\geq 0}^n\text{ s.t. }{\max w} \geq c: \mathcal{A}({\ldblbrace (y, w_i) \rdblbrace}_{i=1}^n) = y \\
+				\text{idempotency: } & \exists c > 0: \forall y \in \mathcal{Y}, w \in \mathbb{R}_{\geq 0}^n\text{ s.t. }{\max w} \geq c: \mathcal{A}({\ldblbrace (y, w_i) \rdblbrace}_{i=1}^n) = y \\
 				\land\text{ zero invar.: } & \forall y_0 \in \mathcal{Y}, S = {\ldblbrace (y_i, w_i) \rdblbrace}_{i=1}^n: \mathcal{A}(S \cup \{ (y_0, 0) \}) = \mathcal{A}(S) \text{.}
@@
 		{\setlength{\parskip}{0pt}The idempotency constraint requires aggregation functions to agree with uniform input scores $y$ if at least one input is weighted above some threshold $c$.
-		The zero invariance constraint requires aggregation functions to ignore inputs with zero weight.}
-
-		One example for such a weighted aggregator is the \textit{weighted mean} function $\wmean({\ldblbrace (y_i, w_i) \rdblbrace}_{i=1}^n) \coloneqq \sum_{i=1}^n w_i y_i$ which requires $\sum w_i = 1$ and $w_i \in [0, 1]$.
-		Alternatively an unweighted aggregation function like $\min$, $\max$ or \ac{owa}\footnote{
-			Even though \ac{owa} uses weights, it does not get those weights as part of the input and is therefore considered to be unweighted here.
+		The zero invariance constraint requires aggregation functions to ignore inputs with zero weight.
+		Exemplary weighted aggregation function are:}
+		\begin{itemize}
+			\item The \textit{weighted mean} function $\wmean({\ldblbrace (y_i, w_i) \rdblbrace}_{i=1}^n) \coloneqq \sum_{i=1}^n w_i y_i$ which requires $\sum w_i = 1$, $w_i \in [0, 1]$ and the existence of some scalar multiplication operator $\cdot: \mathbb{R} \times \mathcal{Y} \to \mathcal{Y}$, typically from some subfield $\mathcal{Y} \subseteq \mathbb{R}$ or possibly also some vector subspace $\mathcal{Y} \subseteq \mathbb{R}^d$.
+			\item Another basic weighted aggregator is the \textit{weighted majority vote} function $\wmaj({\ldblbrace (y_i, w_i) \rdblbrace}_{i=1}^n) \coloneqq {\arg\max}_{y \in \mathcal{Y}} \sum_{y_i = y} w_i$ which returns the input with the highest total weight.
+				Unlike $\wmean$ it can also be applied to score domains $\mathcal{Y}$ without a multiplication operator, e.g.\ sets of discrete classes.
+			\item Alternatively an unweighted aggregation function like $\min$, $\max$, $\mean$ or \ac{owa}\footnote{
+				Even though \ac{owa} uses weights, it does not get those weights as part of the input and is therefore considered to be unweighted in this context.
 			} also trivially satisfies \cref{defn:ltag:weighted-agg} if all inputs $y_i$ with $w_i = 0$ are filtered out and the weights for all remaining inputs are ignored.
+		\end{itemize}
 \end{enumerate}
@@
 	A graph embedding $\varphi: \mathcal{G} \to \mathbb{N}^{d}$ is called a \textit{substructure component embedding} iff.\ %
-	$\forall i \in [d]: \exists S_i \in \mathcal{G}_{\mathit{ord}}, \gamma_i \in \mathbb{R}: \forall G \in \mathcal{G}: {{\varphi(G)}[i] = \mathit{contains}_{S_i}(G)} \lor {{\varphi(G)}[i] = \gamma_i \mathit{count}_{S_i}(G)}$,
+	$\forall i \in [d]: \exists S_i \in \mathcal{G}_{\mathit{ord}}, \gamma_i > 0: \forall G \in \mathcal{G}: {{\varphi(G)}[i] = \gamma_i \mathit{contains}_{S_i}(G)} \lor {{\varphi(G)}[i] = \gamma_i \mathit{count}_{S_i}(G)}$,
 	with $\mathit{contains}$ and $\mathit{count}$ as in \cfullref[ on ]{defn:related:wl-count-detect}.
@@
 Based on this requirement we now proof the main theorem which shows the relation between \acp{svm} and \ac{lta}.
-\begin{thm}
+\begin{thm}\label{thm:ltag:svm-ltag-formulation}
 	A binary \ac{svm} classifier that takes substructure component embeddings $\varphi(G)$ as its input has an \ac{lta} formulation with a nontrivial decomposition function $\psi$.
@@
 	Let $h: \mathcal{G} \to {\{ -1, +1 \}}$ be a binary graph classifier expressed as $h = h_{\text{\acs*{svm}}} \circ \varphi$ where $\varphi: \mathcal{G} \to \mathbb{N}_0^{d}$ is a substructure component embedding and $h_{\text{\acs*{svm}}}: \mathbb{R}^{d} \to {\{ -1, +1 \}}$ an \ac{svm} classifier.
+	By \cref{defn:ltag:substruct-embedding} each $\varphi(G)[i] > 0$ counts or indicates the presence of some ordered subgraph $S_i$ in $G$;
+	more precisely:
+	\begin{itemize}
+		\item If $\varphi(G)[i] = \gamma_i \mathit{count}_{S_i}(G) > 0$, let $C_i \coloneqq {\{ G[s]\, |\, s \in \mathcal{P}(\mathcal{V}_G) \land G[s] \simeq S_i \}}$ be the subgraphs counted by the $i$-th dimension.
+		\item If $\varphi(G)[i] = \gamma_i \mathit{contains}_{S_i}(G) > 0$, let $C_i \coloneqq {\left\{ G\left[ \bigcup \{ s \in \mathcal{P}(\mathcal{V}_G)\, |\, G[s] \simeq S_i \} \right] \right\}}$ be the union graph of all subgraphs of $G$ that are isomorphic to $S_i$.
+	\end{itemize}
+	By interpreting each $C_i$ as a set of isomorphic constituents, we can define the graph decomposition function $\psi_{\varphi}(G) \coloneqq \bigcup_{i=1}^{d} C_i$.
+
+	Based on this decomposition we now bring the \ac{svm} graph classifier $h$ into an \ac{lta} formulation.
 	If $h$ is trained on a dataset $\mathcal{D}_{\text{train}} = {\{ (G_1, y_1), \dots, (G_N, y_N) \}}$, via the kernel trick it can be expressed as
@@
 		\quad\text{for some $\alpha \in \mathbb{R}^N$ and $b \in \mathbb{R}$} \\
-		&= \sgn\left( {\sum_{i=1}^{d} {\varphi(G)}[i] \underbrace{\left( \sum_{j = 1}^{N} \alpha_j y_j {\varphi(G_j)}[i] \right)}_{{} \eqqcolon z_i}} + b \right)
+		&= \sgn\left( {\sum_{i=1}^{d} {
+				\underbrace{\varphi(G)[i]}_{{|C_i|} \gamma_i^{-1}}
+				\underbrace{\left( \sum_{j = 1}^{N} \alpha_j y_j {\varphi(G_j)}[i] \right)}_{{} \eqqcolon \beta_i}}
+			} + b \right)
+		 = \sgn\left( {\sum_{i=1}^{d} {{|C_i|} \gamma_i^{-1} \beta_i}} + b \right) \\
+		&= \sgn\left( {\sum_{c \in \psi_{\varphi}(G)} \underbrace{\sum_{i = 1}^d \mathbbm{1}[c \in C_i] \gamma_i^{-1} \beta_i}_{{} \eqqcolon z_i}} + b \right)
+		 = \sgn\left( {\sum_{c \in \psi_{\varphi}(G)}} \underbrace{|z_i|}_{{} \eqqcolon w_i} \underbrace{\sgn{z_i}}_{\eqqcolon y_i} + b \right)
 		\text{.}
 	\end{align*}}
-
-	By \cref{defn:ltag:substruct-embedding} each $\varphi(G)[i] > 0$ counts or indicates the presence of some ordered subgraph $S_i$ in $G$;
-	more precisely:
-	\begin{itemize}
-		\item If $\varphi(G)[i] = \gamma_i \mathit{count}_{S_i}(G) > 0$, let $C_i \coloneqq {\{ G[s]\, |\, s \in \mathcal{P}(\mathcal{V}_G) \land G[s] \simeq S_i \}}$ be the $\mathit{count}_{S_i}(G)$ subgraphs counted by the $i$-th dimension.
-		\item If $\varphi(G)[i] = \mathit{contains}_{S_i}(G) = 1$, let $C_i \coloneqq {\{  \}}$.
-	\end{itemize}
-	A substructure component embedding vector can therefore be interpreted as an encoding of the graph decomposition \[
-		\psi(G) \coloneqq \bigcup_{i=1}^{d}
-		\text{.}
-	\]
-	\begin{align*}
-		h(G) &= \sgn\left( {\sum_{i=1}^{d} {\varphi(G)}[i] \underbrace{|z_i|}_{{} \eqqcolon \nu_i} \underbrace{\sgn(z_i)}_{{} \eqqcolon y_i}} + b \right)
-	\end{align*}
 \end{proof}

### 2020-03-14 thesis/content/chapter-related-work.tex
continued svm lta proof + minor fixes

@@
 In practice the hash function $h$ is usually defined lazily by using $\mathcal{C} \subseteq \mathbb{N}$ and enumerating color combinations in whichever order they are hashed s.t.\ a new color is introduced every time a previously unseen color combination appears at runtime\footnote{
-	Though $\mathcal{C}$ is formally countably infinite, we assume it only contains the finite no.\ of colors in a dataset.
+	Though $\mathcal{C}$ is countably infinite, we assume it only contains the finite no.\ of colors in a dataset.
 }.

### 2020-03-15 implementation/src/ltag/datasets/manager.py
added evaluation code for FC models and varying WL2 neighborhood radii

@@ -25,6 +25,7 @@ class GraphDatasetManager:
     edge_one_labels=False,
     with_holdout=True,
     evaluation_args=None,
+    name_suffix="",
     **kwargs):
 
     self.kfold_class = kfold_class
@@ -38,6 +39,7 @@ class GraphDatasetManager:
     self.edge_one_labels = edge_one_labels
     self.with_holdout = with_holdout
     self.evaluation_args = evaluation_args
+    self.name_suffix = name_suffix
 
     self.outer_k = outer_k
     assert (outer_k is not None and outer_k > 0) or outer_k is None
@@ -92,6 +94,10 @@ class GraphDatasetManager:
     print(f"Encoded WL2c graphs.")
 
     return wl2_graphs, targets
+
+  @property
+  def full_name(self):
+    return self.name + self.name_suffix
 
   @property
   def dataset(self):

### 2020-03-15 implementation/src/ltag/evaluation/datasets.py

@@ -15,14 +15,31 @@ Mutag_8_mini = fy.partial(
     "batch_graph_count": 1
   })
 
+Mutag_1 = fy.partial(
+  tu.Mutag,
+  name_suffix="_n1",
+  wl2_neighborhood=1)
 Mutag_8 = fy.partial(
   tu.Mutag,
   wl2_neighborhood=8)
 
+NCI1_1 = fy.partial(
+  tu.NCI1,
+  name_suffix="_n1",
+  wl2_neighborhood=1)
 NCI1_8 = fy.partial(
   tu.NCI1,
   wl2_neighborhood=8)
 
+Proteins_1 = fy.partial(
+  tu.Proteins,
+  name_suffix="_n1",
+  wl2_neighborhood=1,
+  wl2_batch_size={
+    "fuzzy_batch_edge_count": 20000,
+    "upper_batch_edge_count": 50000,
+    "batch_graph_count": 20
+  })
 Proteins_5 = fy.partial(
   tu.Proteins,
   wl2_neighborhood=5,
@@ -31,16 +48,25 @@ Proteins_5 = fy.partial(
     "upper_batch_edge_count": 50000,
     "batch_graph_count": 20
   })
-
 Proteins_6 = fy.partial(
   tu.Proteins,
   wl2_neighborhood=6,
+  name_suffix="_n6",
   wl2_batch_size={
     "fuzzy_batch_edge_count": 20000,
     "upper_batch_edge_count": 50000,
     "batch_graph_count": 20
   })
 
+DD_1 = fy.partial(
+  tu.DD,
+  name_suffix="_n1",
+  wl2_neighborhood=1,
+  wl2_batch_size={
+    "fuzzy_batch_edge_count": 30000,
+    "upper_batch_edge_count": 60000,
+    "batch_graph_count": 10
+  })
 DD_2 = fy.partial(
   tu.DD,
   wl2_neighborhood=2,
@@ -50,10 +76,31 @@ DD_2 = fy.partial(
     "batch_graph_count": 10
   })
 
+IMDB_1 = fy.partial(
+  tu.IMDBBinary,
+  name_suffix="_n1",
+  wl2_neighborhood=1)
+IMDB_2 = fy.partial(
+  tu.IMDBBinary,
+  name_suffix="_n2",
+  wl2_neighborhood=2)
+IMDB_4 = fy.partial(
+  tu.IMDBBinary,
[…]

### 2020-03-15 implementation/src/ltag/evaluation/evaluate.py

@@ -113,7 +113,7 @@ def evaluation_step(
 def find_eval_dir(model_factory, ds_manager, label=None):
   label = "_" + label if label is not None else ""
   mf_name = model_factory.name
-  ds_name = ds_manager.name
+  ds_name = ds_manager.full_name
   return eval_dir_base / f"{ds_name}_{mf_name}{label}"
 
 def evaluate(
@@ -127,7 +127,7 @@ def evaluate(
 
   mf_name = model_factory.name
   ds_type = model_factory.input_type
-  ds_name = ds_manager.name
+  ds_name = ds_manager.full_name
   t = time_str()
   if eval_dir is None:
     eval_dir = find_eval_dir(model_factory, ds_manager, label)

### 2020-03-15 implementation/src/ltag/evaluation/models.py

@@ -43,7 +43,7 @@ def AvgWL2GCN_Binary_3x32(dsm):
   )
 
 @binary_classifier(gnn_models.with_fc(gnn_models.AvgWL2GCN))
-def AvgWL2GCN_FC_Binary(dsm):
+def AvgWL2GCN_FC_Binary_Small(dsm):
   "Small hyperparam space for averaging FC WL2GCNs + binary classification."
   in_dim = dsm.dim_wl2_features()
 
@@ -74,8 +74,8 @@ def AvgCWL2GCN_Binary(
   in_dim = dsm.dim_wl2_features()
 
   hidden = [
-   [b] * l
-   for b, l in cart(cwl2_layer_widths, cwl2_layer_depths)]
+    [b] * l
+    for b, l in cart(cwl2_layer_widths, cwl2_layer_depths)]
 
   return cart(
     conv_layer_dims=[[in_dim, *h, 1] for h in hidden],
@@ -87,6 +87,34 @@ def AvgCWL2GCN_Binary(
     squeeze_output=[True]
   )
 
+@binary_classifier(gnn_models.with_fc(gnn_models.AvgWL2GCN))
+def AvgCWL2GCN_FC_Binary(
+  dsm,
+  cwl2_local_act="sigmoid",
+  cwl2_layer_widths=[32, 64],
+  cwl2_layer_depths=[1, 3],
+  cwl2_stack_tfs=[None, "keep_input"]):
+  in_dim = dsm.dim_wl2_features()
+
+  hidden = [
+    ([b] * l, [b, b])
+    for b, l in cart(cwl2_layer_widths, cwl2_layer_depths)]
+
+  hidden_hp = [dict(
+    conv_layer_dims=[in_dim, *ch],
+    fc_layer_dims=[*fh, 1]
+  ) for ch, fh in hidden]
+
+  return cart_merge(cart(
+    conv_act=["sigmoid"],
+    conv_local_act=[cwl2_local_act],
+    conv_stack_tf=cwl2_stack_tfs,
+    conv_bias=[True],
+    fc_bias=[True],
+    learning_rate=[0.01, 0.001, 0.0001],
+    squeeze_output=[True]
+  ), hidden_hp)
+
 @binary_classifier(gnn_models.SagCWL2GCN)
 def SagCWL2GCN_Binary(
   dsm,
@@ -97,8 +125,8 @@ def SagCWL2GCN_Binary(
   in_dim = dsm.dim_wl2_features()
 
   hidden = [
-   [b] * l
[…]

### 2020-03-15 implementation/src/ltag/playground.py

@@ -10,6 +10,7 @@ from tensorflow import keras
 import numpy as np
 import sklearn as sk
 import grakel as gk
+import funcy as fy
 
 import ltag.models.gnn as gnn_models
 import ltag.models.kernel as kernel_models
@@ -17,6 +18,7 @@ import ltag.datasets.synthetic.datasets as synthetic
 import ltag.evaluation.datasets as eval_ds
 import ltag.evaluation.evaluate as evaluate
 import ltag.evaluate_datasets as eval_main
+from ltag.utils import cart, cart_merge, entry_duplicator
 
 # -%% codecell
 
@@ -149,6 +151,32 @@ def reddit_experient():
     label=f"{dsm.name}_{model.name}")
   print(model.evaluate(test))
 
+def imdb_experient():
+  model_class = gnn_models.SagCWL2GCN
+  dsm = eval_ds.IMDB_8()
+
+  model = model_class(
+    act="sigmoid", local_act="sigmoid", squeeze_output=True,
+    conv_layer_dims=[dsm.dim_wl2_features(), 40, 40, 40, 1],
+    att_conv_layer_dims=[dsm.dim_wl2_features(), 1],
+    conv_stack_tf="keep_input",
+    bias=True)
+
+  opt = keras.optimizers.Adam(0.0007)
+
+  model.compile(
+    optimizer=opt,
+    loss="binary_crossentropy",
+    metrics=["accuracy"])
+
+  train, val = dsm.get_train_fold(0, output_type=model_class.input_type)
+  test = dsm.get_test_fold(0, output_type=model_class.input_type)
+
+  evaluate.train(
+    model, train, test, verbose=2, epochs=1000,
+    label=f"{dsm.name}_{model.name}")
+  print(model.evaluate(test))
+
 def wl2_power_experiment():
   model_class = gnn_models.AvgCWL2GCN
   # model_class = gnn_models.with_fc(model_class)
@@ -247,8 +275,8 @@ def kernel_experiment():
     print(evaluate.train(model, ds, ds_test, label=f"{dsm.name}_{model.name}").history)
 
 
-synthetic_experiment2()
-# kernel_experiment()
+# synthetic_experiment2()
+# imdb_experient()
 
 #
 # list(dsm.get_all(output_type="grakel")[0])

### 2020-03-15 thesis/content/chapter-graph-lta.tex
completed first svm lta theorem

@@
 		Molecular graphs from chemical datasets for example are meaningfully characterized by the presence of so-called \textit{functional groups} consisting of multiple bonded atoms while a characterization on the level of individual atoms is generally less meaningful~\cite{McNaught1997}.
-	\item \textbf{Disaggregation:}
+	\item \textbf{Local evaluation:}
 		The constituents $c_{G, i} \in \psi(G)$ must be evaluated via some function $f: \mathcal{G} \to \mathcal{Y} \times \mathbb{R}$.
@@
 			\begin{align*}
-				\text{idempotency: } & \exists c > 0: \forall y \in \mathcal{Y}, w \in \mathbb{R}_{\geq 0}^n\text{ s.t. }{\max w} \geq c: \mathcal{A}({\ldblbrace (y, w_i) \rdblbrace}_{i=1}^n) = y \\
+				\text{idempotency: } & \exists \eta > 0: \forall y \in \mathcal{Y}, w \in \mathbb{R}_{\geq 0}^n\text{ s.t. }{\max w} \geq \eta: \mathcal{A}({\ldblbrace (y, w_i) \rdblbrace}_{i=1}^n) = y \\
 				\land\text{ zero invar.: } & \forall y_0 \in \mathcal{Y}, S = {\ldblbrace (y_i, w_i) \rdblbrace}_{i=1}^n: \mathcal{A}(S \cup \{ (y_0, 0) \}) = \mathcal{A}(S) \text{.}
@@
 		\end{defn}
-		{\setlength{\parskip}{0pt}The idempotency constraint requires aggregation functions to agree with uniform input scores $y$ if at least one input is weighted above some threshold $c$.
+		{\setlength{\parskip}{0pt}The idempotency constraint requires aggregation functions to agree with uniform input scores $y$ if at least one input is weighted above some threshold $\eta$.
 		The zero invariance constraint requires aggregation functions to ignore inputs with zero weight.
@@
 \end{enumerate}
-Based on the notion of decomposition, disaggregation and aggregation we can now define the concept of \textit{\ac{lta} formulations}.
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=\linewidth]{gfx/graph-lta/ltag-overview.pdf}
+	\caption{
+		Overview of the generalized \ac{lta} architecture for structured data.
+	}\label{fig:ltag:ltag-overview}
+\end{figure}
+Based on the notion of decomposition, local evaluation and aggregation we can now define the concept of \textit{\ac{lta} formulations}.
 \begin{defn}
@@
 \begin{defn}\label{defn:ltag:substruct-embedding}
-	A graph embedding $\varphi: \mathcal{G} \to \mathbb{N}^{d}$ is called a \textit{substructure component embedding} iff.\ %
-	$\forall i \in [d]: \exists S_i \in \mathcal{G}_{\mathit{ord}}, \gamma_i > 0: \forall G \in \mathcal{G}: {{\varphi(G)}[i] = \gamma_i \mathit{contains}_{S_i}(G)} \lor {{\varphi(G)}[i] = \gamma_i \mathit{count}_{S_i}(G)}$,
-	with $\mathit{contains}$ and $\mathit{count}$ as in \cfullref[ on ]{defn:related:wl-count-detect}.
+	A graph embedding $\varphi: \mathcal{G} \to \mathbb{N}_0^{d}$ is called a \textit{substructure component embedding} iff.\ there are decomposition functions $\psi_{\varphi, i}: \mathcal{G} \to \mathcal{P}(\mathcal{G})$ and normalization constants $\gamma_i \in \mathbb{R}$ for all embedding components $i \in [d]$ s.t.\ $\forall G \in \mathcal{G}, i \in [d]: \varphi(G)[i] = \gamma_i {| \psi_{\varphi, i}(G) |}$.
+	The joint decomposition $\psi_{\varphi}(G) \coloneqq \bigcup_{i=1}^d \psi_{\varphi,i}(G)$ is a so-called \textit{underlying decomposition} of the embedding $\varphi$.
 \end{defn}
-Intuitively \cref{defn:ltag:substruct-embedding} states that every embedding dimension must represent the count or presence of some ordered subgraph $S_i$ in the embedded graph $G$.
+Intuitively \cref{defn:ltag:substruct-embedding} states that every embedding dimension must represent the number of constituents produced by some decomposition function.
 Based on this requirement we now proof the main theorem which shows the relation between \acp{svm} and \ac{lta}.
+
 \begin{thm}\label{thm:ltag:svm-ltag-formulation}
-	A binary \ac{svm} classifier that takes substructure component embeddings $\varphi(G)$ as its input has an \ac{lta} formulation with a nontrivial decomposition function $\psi$.
+	A binary \ac{svm} graph classifier $h$ that applies a substructure component embedding $\varphi$ to its inputs has an \ac{lta} formulation with the decomposition function $\psi_{\varphi}$.
 \end{thm}
 \begin{proof}
-	Let $h: \mathcal{G} \to {\{ -1, +1 \}}$ be a binary graph classifier expressed as $h = h_{\text{\acs*{svm}}} \circ \varphi$ where $\varphi: \mathcal{G} \to \mathbb{N}_0^{d}$ is a substructure component embedding and $h_{\text{\acs*{svm}}}: \mathbb{R}^{d} \to {\{ -1, +1 \}}$ an \ac{svm} classifier.
-	By \cref{defn:ltag:substruct-embedding} each $\varphi(G)[i] > 0$ counts or indicates the presence of some ordered subgraph $S_i$ in $G$;
-	more precisely:
-	\begin{itemize}
-		\item If $\varphi(G)[i] = \gamma_i \mathit{count}_{S_i}(G) > 0$, let $C_i \coloneqq {\{ G[s]\, |\, s \in \mathcal{P}(\mathcal{V}_G) \land G[s] \simeq S_i \}}$ be the subgraphs counted by the $i$-th dimension.
-		\item If $\varphi(G)[i] = \gamma_i \mathit{contains}_{S_i}(G) > 0$, let $C_i \coloneqq {\left\{ G\left[ \bigcup \{ s \in \mathcal{P}(\mathcal{V}_G)\, |\, G[s] \simeq S_i \} \right] \right\}}$ be the union graph of all subgraphs of $G$ that are isomorphic to $S_i$.
-	\end{itemize}
-	By interpreting each $C_i$ as a set of isomorphic constituents, we can define the graph decomposition function $\psi_{\varphi}(G) \coloneqq \bigcup_{i=1}^{d} C_i$.
+	Let $h: \mathcal{G} \to {\{ -1, +1 \}}$ be a binary graph classifier expressed as $h = h_{\text{\acs*{svm}}} \circ \varphi$ where $\varphi: \mathcal{G} \to \mathbb{N}_0^{d}$ is a substructure component embedding and $h_{\text{\acs*{svm}}}: \mathbb{R}^{d} \to {\{ -1, +1 \}}$ a standard \ac{svm} classifier.
+	Additionally, let $\psi_{\varphi}$ be some underlying decomposition of $\varphi$.
 
@@
 	If $h$ is trained on a dataset $\mathcal{D}_{\text{train}} = {\{ (G_1, y_1), \dots, (G_N, y_N) \}}$, via the kernel trick it can be expressed as
-	{\allowdisplaybreaks\begin{align*}
+	\begin{align*}
 		h(G) &= \sgn\left( \sum_{j = 1}^{N} \alpha_j y_j {\langle \varphi(G), \varphi(G_j) \rangle} + b \right)
@@
 		&= \sgn\left( {\sum_{i=1}^{d} {
-				\underbrace{\varphi(G)[i]}_{{|C_i|} \gamma_i^{-1}}
-				\underbrace{\left( \sum_{j = 1}^{N} \alpha_j y_j {\varphi(G_j)}[i] \right)}_{{} \eqqcolon \beta_i}}
+				\underbrace{\varphi(G)[i]}_{{|\psi_{\varphi, i}(G)|} \gamma_i^{-1}}
+				\underbrace{\left( \sum_{j = 1}^{N} \alpha_j y_j {\varphi(G_j)}[i] \right)}_{\beta_i}}
 			} + b \right)
-		 = \sgn\left( {\sum_{i=1}^{d} {{|C_i|} \gamma_i^{-1} \beta_i}} + b \right) \\
-		&= \sgn\left( {\sum_{c \in \psi_{\varphi}(G)} \underbrace{\sum_{i = 1}^d \mathbbm{1}[c \in C_i] \gamma_i^{-1} \beta_i}_{{} \eqqcolon z_i}} + b \right)
-		 = \sgn\left( {\sum_{c \in \psi_{\varphi}(G)}} \underbrace{|z_i|}_{{} \eqqcolon w_i} \underbrace{\sgn{z_i}}_{\eqqcolon y_i} + b \right)
+		 = \sgn\left( {\sum_{i=1}^{d} {{|\psi_{\varphi, i}(G)|} \gamma_i^{-1} \beta_i}} + b \right) \\
+		&= \sgn\left( {\sum_{c_t \in \psi_{\varphi}(G)} \underbrace{\sum_{i = 1}^d \mathbbm{1}[c_t \in \psi_{\varphi, i}(G)] \gamma_i^{-1} \beta_i}_{z_t}} + b \right)
+		 = \sgn\left( {\sum_{c_t \in \psi_{\varphi}(G)}} \underbrace{|z_t|}_{w_t} \underbrace{\sgn{z_t}}_{y_t} + {\underbrace{|b|}_{w_b} \underbrace{\sgn{b}}_{y_b}} \right) \\
+		&= \wmaj\left( \ldblbrace (y_t, w_t)\, |\, c_t \in \psi_{\varphi}(G) \rdblbrace \cup \ldblbrace (y_b, w_b) \rdblbrace \right)
 		\text{.}
-	\end{align*}}
+	\end{align*}
+	By choosing $f_h(c_t) \coloneqq (y_t, w_t)$ and $\mathcal{A}_h(S) = \wmaj(S \cup \ldblbrace (y_b, w_b) \rdblbrace)$, the \ac{svm} model therefore has an \ac{lta} formulation with the decomposition function $\psi_{\varphi}$.
+	To see why $\mathcal{A}_h$ satisfies \cref{defn:ltag:weighted-agg}, note that it inherits the idempotency property from $\wmaj$ because $\mathcal{A}_h$ ignores the bias ``pseudo-constituent'' $(y_b, w_b)$ for all threshold weights $\max w_t \geq \eta > w_b$, similarly the zero invariance of $\wmaj$ is also directly inherited.
+	This concludes the proof.
 \end{proof}
+
+The central statement of \cref{thm:ltag:svm-ltag-formulation} is that the output of an \acs{svm}-based graph classifier can be interpreted as the weighted majority vote of constituent scores $y_t$.
+Those constituent scores are in turn also the result of a weighted majority vote as shown in the following proposition.
+\begin{prop}
+
+\end{prop}
 

### 2020-03-15 thesis/content/chapter-related-work.tex
completed first svm lta theorem

@@
 \begin{defn}\label{defn:related:ordered-subgraph}
-	For all vertex $k$-tuples $v = (v_1, \dots, v_k) \in \mathcal{V}_G^k$ let $\hat{v} = \{ v_i\, |\, i \in [k] \}$.
+	For all vertex $k$-tuples $v = (v_1, \dots, v_k) \in \mathcal{V}_G^k$ let $\hat{v} = \set(v) \coloneqq \{ v_i\, |\, i \in [k] \}$.
 	Then $G[v] \coloneqq (\hat{v}, \mathcal{E}_G \cap \hat{v}^2, v)$ is called the \textit{ordered subgraph} of $G$ induced by $v$.
@@
 \begin{defn}\label{defn:related:wl-count-detect}
-	$k$-\acs{wl} \textit{counts} a certain ordered subgraph $S \in \mathcal{G}_{\mathit{ord}}$ iff.\ it computes $\mathit{count}_S(G) \coloneqq \left|\{ \hat{v}\, |\, v \in \mathcal{V}_G^{*} \land G[v] \equiv S \}\right|$ (see~\cref{defn:related:ordered-subgraph}).
+	$k$-\acs{wl} \textit{counts} a certain ordered subgraph $S \in \mathcal{G}_{\mathit{ord}}$ iff.\ it computes $\mathit{count}_S(G) \coloneqq \left|\{ \set(v)\, |\, v \in \mathcal{V}_G^{*} \land G[v] \equiv S \}\right|$ (see~\cref{defn:related:ordered-subgraph}).
 	Similarly $k$-\acs{wl} \textit{detects} an ordered subgraph $S$ iff.\ it computes $\mathit{contain}_S(G) \coloneqq \mathbbm{1}[\mathit{count}_S(G) > 0]$, with $\mathbbm{1}$ denoting the indicator function.

### 2020-03-16 thesis/content/chapter-graph-lta.tex
started with lta formulations of specific embeddings
+ various fixes

@@
 		If a model has an \ac{lta} formulation with a decomposition function that returns interpretable constituents, it is considered to be an \ac{lta} model.
-		By definition this is true for the constituents $\psi(G) \coloneqq {\{ G[v_{c_i}] \}}_{v_{c_i} \in \mathcal{V}_G}$ of \ac{lta} methods for unstructured data.
+		By definition this is true for the single-vertex constituents $\psi(G) \coloneqq {\{ G[v_{c_i}] \}}_{v_{c_i} \in \mathcal{V}_G}$ of \ac{lta} methods for unstructured data.
 	\item \textbf{In-between cases:}
@@
 \citet{Gilpin2018} describe this trade-off between explainability and expressive power in more detail.
-When considering problem domains in which the true composition scores $y_G$ are accurately described by an \acs{lta}-like generative process, a less expressive \ac{lta}-like model could however generalize better than a more expressive non-\ac{lta} model.
+However, when considering problem domains in which the true composition scores $y_G$ are accurately described by an \acs{lta}-like generative process, a less expressive \ac{lta}-like model could generalize better than a more expressive non-\ac{lta} model.
 This idea is captured by the so-called \textit{\ac{lta} assumption}.
 \begin{defn}
-	A problem domain $\mathcal{D}$ satisfies the \textit{\ac{lta} assumption} iff.\ there is an \ac{lta} method which produces models with an equal or lower out-of-sample error than the models produced by non-\acs{lta} methods for most training samples from $\mathcal{D}$.
+	A problem domain $\mathcal{D}$ satisfies the \textit{\ac{lta} assumption} iff.\ there is an \ac{lta} method which produces models with an equal or lower out-of-sample error than the models produced by non-\acs{lta} methods for most training samples $\mathcal{D}_{\text{train}} \subseteq \mathcal{D}$.
 \end{defn}
@@
 One common way to solve the \ac{gcr} problem via those embedding vectors is to train an \ac{svm} on them.
-We will now see that \acp{svm} can be interpreted as \ac{lta} models if they are trained on so-called \textit{substructure component embeddings}.
+We will now see that \acp{svm} can be interpreted as \ac{lta} models if they are trained on so-called \acp*{sce}.
+\begin{defn}
+	Given a a multiset $A = \ldblbrace \underbrace{a_1, \dots, a_1}_{\gamma_A(a_1)\text{ times}},\allowbreak \dots, \underbrace{a_n, \dots, a_n}_{\gamma_A(a_n)\text{ times}} \rdblbrace \subseteq D$, the so-called \textit{multiplicity function} $\gamma_A: D \to \mathbb{N}_0$ maps each element of the domain $D$ to its multiplicity in $A$ with $\gamma_A(x) = 0 \Leftrightarrow x \notin A$.
+\end{defn}
 \begin{defn}\label{defn:ltag:substruct-embedding}
-	A graph embedding $\varphi: \mathcal{G} \to \mathbb{N}_0^{d}$ is called a \textit{substructure component embedding} iff.\ there are decomposition functions $\psi_{\varphi, i}: \mathcal{G} \to \mathcal{P}(\mathcal{G})$ and normalization constants $\gamma_i \in \mathbb{R}$ for all embedding components $i \in [d]$ s.t.\ $\forall G \in \mathcal{G}, i \in [d]: \varphi(G)[i] = \gamma_i {| \psi_{\varphi, i}(G) |}$.
-	The joint decomposition $\psi_{\varphi}(G) \coloneqq \bigcup_{i=1}^d \psi_{\varphi,i}(G)$ is a so-called \textit{underlying decomposition} of the embedding $\varphi$.
+	A graph embedding $\varphi: \mathcal{G} \to \mathbb{N}_0^{d}$ is called a \ac{sce} iff.\ %
+	there are decomposition functions $\psi_{\varphi, i}: \mathcal{G} \to \mathcal{P}(\mathcal{G})$ and functions $\gamma_{\varphi, i}: \mathcal{G} \to (\mathcal{G} \to \mathbb{N}_0)$ for all embedding components $i \in [d]$ s.t.\ %
+	$\forall G \in \mathcal{G}, i \in [d]: \varphi(G)[i] = \sum_{c \in \psi_{\varphi, i}(G)} \gamma_{\varphi, i}(G)(c)$ where $\gamma_{\varphi, i}(G): \mathcal{G} \to \mathbb{N}_0$ is the multiplicity function of a multiset of the constituents $\psi_{\varphi, i}(G)$.
+	We call $\psi_{\varphi, i}$ an \textit{underlying decomposition} of the $i$-th component of the embedding $\varphi$.
+	Similarly the joint decomposition $\psi_{\varphi}(G) \coloneqq \bigcup_{i=1}^d \psi_{\varphi,i}(G)$ is an \textit{underlying decomposition} of $\varphi$.
 \end{defn}
-Intuitively \cref{defn:ltag:substruct-embedding} states that every embedding dimension must represent the number of constituents produced by some decomposition function.
+Intuitively \cref{defn:ltag:substruct-embedding} states that the value of each \ac{sce} component must be derived from the number of constituents produced by some decomposition function where each constituent might be counted multiple times.
 Based on this requirement we now proof the main theorem which shows the relation between \acp{svm} and \ac{lta}.
@@
 \begin{thm}\label{thm:ltag:svm-ltag-formulation}
-	A binary \ac{svm} graph classifier $h$ that applies a substructure component embedding $\varphi$ to its inputs has an \ac{lta} formulation with the decomposition function $\psi_{\varphi}$.
+	A binary \ac{svm} graph classifier $h$ that applies a \ac{sce} $\varphi$ to its inputs has an \ac{lta} formulation that uses the underlying decomposition function $\psi_{\varphi}$.
 \end{thm}
 \begin{proof}
-	Let $h: \mathcal{G} \to {\{ -1, +1 \}}$ be a binary graph classifier expressed as $h = h_{\text{\acs*{svm}}} \circ \varphi$ where $\varphi: \mathcal{G} \to \mathbb{N}_0^{d}$ is a substructure component embedding and $h_{\text{\acs*{svm}}}: \mathbb{R}^{d} \to {\{ -1, +1 \}}$ a standard \ac{svm} classifier.
-	Additionally, let $\psi_{\varphi}$ be some underlying decomposition of $\varphi$.
+	Let $h: \mathcal{G} \to {\{ -1, +1 \}}$ be a binary graph classifier expressed as $h = h_{\text{\acs*{svm}}} \circ \varphi$ where $\varphi: \mathcal{G} \to \mathbb{N}_0^{d}$ is an \ac{sce} and $h_{\text{\acs*{svm}}}: \mathbb{R}^{d} \to {\{ -1, +1 \}}$ a standard \ac{svm} classifier.
+	Additionally, let $\psi_{\varphi}$ be some underlying decomposition of $\varphi$ with ${\{ \gamma_{\varphi, i} \}}_{i=1}^d$ being the corresponding multiplicity functions.
 
@@
 		h(G) &= \sgn\left( \sum_{j = 1}^{N} \alpha_j y_j {\langle \varphi(G), \varphi(G_j) \rangle} + b \right)
-		\quad\text{for some $\alpha \in \mathbb{R}^N$ and $b \in \mathbb{R}$} \\
+		\quad\text{for some $\alpha \in \mathbb{R}_{\geq 0}^N$ and $b \in \mathbb{R}$} \\
 		&= \sgn\left( {\sum_{i=1}^{d} {
-				\underbrace{\varphi(G)[i]}_{{|\psi_{\varphi, i}(G)|} \gamma_i^{-1}}
+				\varphi(G)[i]
 				\underbrace{\left( \sum_{j = 1}^{N} \alpha_j y_j {\varphi(G_j)}[i] \right)}_{\beta_i}}
 			} + b \right)
-		 = \sgn\left( {\sum_{i=1}^{d} {{|\psi_{\varphi, i}(G)|} \gamma_i^{-1} \beta_i}} + b \right) \\
-		&= \sgn\left( {\sum_{c_t \in \psi_{\varphi}(G)} \underbrace{\sum_{i = 1}^d \mathbbm{1}[c_t \in \psi_{\varphi, i}(G)] \gamma_i^{-1} \beta_i}_{z_t}} + b \right)
+		 = \sgn\left( {\sum_{i=1}^{d} \varphi(G)[i] \beta_i} + b \right) \\
+		&= \sgn\left( {\sum_{c_t \in \psi_{\varphi}(G)} \underbrace{\sum_{i = 1}^d \gamma_{\varphi, i}(G)(c_t) \beta_i}_{z_t}} + b \right)
 		 = \sgn\left( {\sum_{c_t \in \psi_{\varphi}(G)}} \underbrace{|z_t|}_{w_t} \underbrace{\sgn{z_t}}_{y_t} + {\underbrace{|b|}_{w_b} \underbrace{\sgn{b}}_{y_b}} \right) \\
@@
 
-The central statement of \cref{thm:ltag:svm-ltag-formulation} is that the output of an \acs{svm}-based graph classifier can be interpreted as the weighted majority vote of constituent scores $y_t$.
-Those constituent scores are in turn also the result of a weighted majority vote as shown in the following proposition.
-\begin{prop}
+The central statement of \cref{thm:ltag:svm-ltag-formulation} is that \acs{svm} graph classifiers that use an \ac{sce} $\varphi$ implicitly perform an \acs{lta}-like weighted majority vote aggregation of constituent scores $y_t$.
+Those constituent scores can in turn be expressed as a majority vote of the graph scores in the training dataset $\mathcal{D}_{\text{train}}$:
+\begin{align}
+	y_t \coloneqq \sgn z_t
+	 &= \sgn\left( \sum_{j=1}^N \underbrace{\left( \alpha_j \smashoperator[l]{\sum_{c_k \in \psi_{\varphi}(G_j)}} \sum_{i=1}^d \gamma_{\varphi,i}(G)(c_t) \gamma_{\varphi,i}(G_j)(c_k) \right)}_{w_j} y_j \right) \\
+	 &= \wmaj(\ldblbrace (y_j, w_j)\, |\, j \in [N] \rdblbrace) \nonumber
+\end{align}
+Note that, if the constituents of $G_j$ all come from different embedding components than those of $G$ ($\forall i \in [d]: {\psi_{\varphi, i}(G_j) \cap \psi_{\varphi, i}(G)} = \emptyset$), then $G_j$ has no influence on the score of $G$ ($w_j = 0$).
 
-\end{prop}
+To see what \cref{thm:ltag:svm-ltag-formulation} implies in practice, we will now apply it to the graph embedding approaches described in \cref{sec:related:gcr:embed,sec:related:gcr:kernel}:
+\begin{enumerate}[label=\textbf{\arabic*.}]
+	\item \textbf{Fingerprint embeddings:}
+		Here the $i$-th embedding component represent the number of occurrences of a substructure $S_i$, i.e.\ ${\varphi}_{\text{FP}}(G)[i] \coloneqq \mathit{count}_{S_i}(G)$ (see \cfullref{defn:related:wl-count-detect}).
+		Such a fingerprint embedding naturally is an \ac{sce} with the decomposition functions $\psi_{\text{FP}, i}(G) \coloneqq \{ G[\set(s)]\, |\, s \in \mathcal{V}_G^{*} \land G[s] \equiv S_i \}$ and the multiplicity functions $\gamma_{\text{FP}, i}(G)(c) \coloneqq \mathbbm{1}[c \in \psi_{\text{FP}, i}(G)]$.
+% chktex 21
+
+		Under the assumption that the substructure patterns $S_i$ are chosen s.t.\ the their instances $c \in \psi_{\text{FP}, i}(G)$ are nontrivial interpretable constituents, the underlying joint decomposition $\psi_{\text{FP}}(G) \coloneqq \bigcup_{i=1}^d \psi_{\text{FP}, i}(G)$ must also be nontrivial and interpretable.
+		Thus, by \cref{thm:ltag:svm-ltag-formulation}, \acp{svm} trained on fingerprint embeddings are \ac{lta} models that produce locally explainable predictions.
+		\badgebox[t_green]{\cmark\ LTA}
+	\item \textbf{\texttt{graph2vec}:}
+		The skip-gram inspired \texttt{graph2vec} embedding produces vectors $\text{FP}(G) \in \mathbb{R}^d$ whose individual components do not have a clear interpretation.
+		Since the component values can be arbitrary reals, this approach is not an \ac{sce} which in turn implies that it does not have an \ac{lta} formulation as in \cref{thm:ltag:svm-ltag-formulation}.
+		\badgebox[t_red]{\xmark\ non-LTA}
+	\item \textbf{\ac{wl} subtree kernel:}
+		The individual components of the \ac{wl} subtree embedding represent the number of occurrences of a color $c_i \in \mathcal{C}$, i.e.\ $\varphi_{\text{ST}}(G)[i] \coloneqq \mathit{dist}_{\chi_{G,1}^{(t)}}(c_i)$ as defined in \cfullref[ on ]{eq:related:wl-subtree-kernel}.
+		As illustrated in \cfullref[ on ]{fig:related:wl-subtree}, each occurrence of a color $c_i$ in the \ac{wl} coloring $\chi_{G,1}^{(t)}$ corresponds to the occurrence of a breadth-first subtree of depth $t$.
+		\badgebox[t_green]{\cmark\ LTA-like}
+	\item \textbf{\ac{wl} shortest path kernel:}
+		\badgebox[t_red]{\xmark\ non-LTA}
+	\item \textbf{$k$-LWL kernel:}
+		\badgebox[t_green]{\cmark\ LTA-like}
+	\item \textbf{$k$-GWL kernel:}
+		\badgebox[t_red]{\xmark\ non-LTA}
+\end{enumerate}
 

### 2020-03-16 thesis/content/chapter-related-work.tex
started with lta formulations of specific embeddings
+ various fixes

@@
 
+Now that we have looked at the \acs{wl}- and spectrum-based graph characterization approaches, we will see how those can be used to learn models that solve the \acf{gcr} problem.
 The existing approaches to tackle the \ac{gcr} problem can be categorized into three main families:
@@
 \end{enumerate*}.
-We will now look at the characteristics of those families and give a brief overview of specific methods.
+In the following sections the characteristics of those families are described and a brief overview of specific methods is given.
 
@@
 So-called \textit{fingerprint embeddings} try to match a fixed set of subgraphs $S_1, \dots, S_d$ to the input graph.
-The embedding of a graph $G$ is a binary vector $\varphi(G) = x \in {\{0, 1\}}^d$ with $x_i \coloneqq \mathit{contain}_{S_i}(G)$ (see~\cref{defn:related:wl-count-detect}).
+The embedding of a graph $G$ is a binary vector ${\varphi}_{\text{FP}}(G) \in {\{0, 1\}}^d$ with ${\varphi}_{\text{FP}}(G)[i] \coloneqq \mathit{contain}_{S_i}(G)$ (see~\cref{defn:related:wl-count-detect}).
+Alternatively a fingerprint embedding can additionally encode multiplicities via ${\varphi}_{\text{FP}}(G)[i] \coloneqq \mathit{count}_{S_i}(G) \in \mathbb{N}_0$.
+
 This simple approach usually requires a careful choice of subgraphs but can still be competitive with the other more recent approaches we will look at in the following sections.
@@
 \begin{align}
-	{\varphi_{\text{SP}}(G)}_i \coloneqq \left| \left\{ (v, u) \in \mathcal{V}_G^2 \,|\, {\chi_{G,1}^{(t_i)}(v) = a_i} \land {\chi_{G,1}^{(t_i)}(u) = b_i} \land {d_{\text{SP}}(v, u) = d_i} \right\} \right| \text{,} % chktex 21
+	{\varphi_{\text{SP}}(G)}[i] \coloneqq \left| \left\{ (v, u) \in \mathcal{V}_G^2 \,|\, {\chi_{G,1}^{(t_i)}(v) = a_i} \land {\chi_{G,1}^{(t_i)}(u) = b_i} \land {d_{\text{SP}}(v, u) = d_i} \right\} \right| \text{,} % chktex 21
 \end{align}

### 2020-03-16 implementation/src/ltag/evaluation/datasets.py

@@ -76,6 +76,15 @@ DD_2 = fy.partial(
     "batch_graph_count": 10
   })
 
+chemical_binary = [
+  "NCI1_8",
+  "Proteins_5",
+  "DD_2"
+]
+chemical = chemical_binary
+
+# Social:
+
 IMDB_1 = fy.partial(
   tu.IMDBBinary,
   name_suffix="_n1",
@@ -96,16 +105,6 @@ IMDB_8 = fy.partial(
   tu.IMDBBinary,
   wl2_neighborhood=8)
 
-chemical_binary = [
-  "NCI1_8",
-  "Proteins_5",
-  "DD_2",
-  "IMDB_8"
-]
-chemical = chemical_binary
-
-# Social:
-
 RedditBinary_1 = fy.partial(
   tu.RedditBinary,
   wl2_neighborhood=1,
@@ -117,6 +116,7 @@ RedditBinary_1 = fy.partial(
 )
 
 social_binary = [
+  "IMDB_8",
   "RedditBinary_1"
 ]
 social = social_binary

### 2020-03-17 thesis/content/chapter-graph-lta.tex
added the lta formulation for the wl_st kernel

@@
 	A graph embedding $\varphi: \mathcal{G} \to \mathbb{N}_0^{d}$ is called a \ac{sce} iff.\ %
-	there are decomposition functions $\psi_{\varphi, i}: \mathcal{G} \to \mathcal{P}(\mathcal{G})$ and functions $\gamma_{\varphi, i}: \mathcal{G} \to (\mathcal{G} \to \mathbb{N}_0)$ for all embedding components $i \in [d]$ s.t.\ %
-	$\forall G \in \mathcal{G}, i \in [d]: \varphi(G)[i] = \sum_{c \in \psi_{\varphi, i}(G)} \gamma_{\varphi, i}(G)(c)$ where $\gamma_{\varphi, i}(G): \mathcal{G} \to \mathbb{N}_0$ is the multiplicity function of a multiset of the constituents $\psi_{\varphi, i}(G)$.
+	there are decomposition functions $\psi_{\varphi, i}: \mathcal{G} \to \mathcal{P}(\mathcal{G})$ and multiplicity functions $\gamma_{\varphi, i}: \mathcal{G} \to \mathbb{N}_0$ for all embedding components $i \in [d]$ s.t.\ %
+	$\forall G \in \mathcal{G}, i \in [d]: \varphi(G)[i] = \sum_{c \in \psi_{\varphi, i}(G)} \gamma_{\varphi, i}(c)$ where $\gamma_{\varphi, i}$ is the multiplicity function of a multiset of the constituents $\psi_{\varphi, i}(G)$.
 	We call $\psi_{\varphi, i}$ an \textit{underlying decomposition} of the $i$-th component of the embedding $\varphi$.
@@
 \begin{thm}\label{thm:ltag:svm-ltag-formulation}
-	A binary \ac{svm} graph classifier $h$ that applies a \ac{sce} $\varphi$ to its inputs has an \ac{lta} formulation that uses the underlying decomposition function $\psi_{\varphi}$.
+	A binary \ac{svm} graph classifier $h$ that applies a \ac{sce} $\varphi$ to its inputs has an \ac{lta} formulation that uses an underlying decomposition $\psi_{\varphi}$ of $\varphi$.
 \end{thm}
@@
 		 = \sgn\left( {\sum_{i=1}^{d} \varphi(G)[i] \beta_i} + b \right) \\
-		&= \sgn\left( {\sum_{c_t \in \psi_{\varphi}(G)} \underbrace{\sum_{i = 1}^d \gamma_{\varphi, i}(G)(c_t) \beta_i}_{z_t}} + b \right)
+		&= \sgn\left( {\sum_{c_t \in \psi_{\varphi}(G)} \underbrace{\sum_{i = 1}^d \gamma_{\varphi, i}(c_t) \beta_i}_{z_t}} + b \right)
 		 = \sgn\left( {\sum_{c_t \in \psi_{\varphi}(G)}} \underbrace{|z_t|}_{w_t} \underbrace{\sgn{z_t}}_{y_t} + {\underbrace{|b|}_{w_b} \underbrace{\sgn{b}}_{y_b}} \right) \\
@@
 	By choosing $f_h(c_t) \coloneqq (y_t, w_t)$ and $\mathcal{A}_h(S) = \wmaj(S \cup \ldblbrace (y_b, w_b) \rdblbrace)$, the \ac{svm} model therefore has an \ac{lta} formulation with the decomposition function $\psi_{\varphi}$.
+
+	To complete the proof it now remains to show that $f_h$ and $\mathcal{A}_h$ are in fact a local evaluation function and a weighted aggregation function respectively.
+	To see that $f_h$ perform local evaluation, note that $f_h(c_t) \coloneqq (\sgn z_t, {|z_t|})$ with $z_t \coloneqq \sum_{i=1}^{d} \gamma_{\varphi,i}(c_t) \beta_i$ only depends on the shared multiplicity functions $\gamma_{\varphi,i}$, the constants $\beta_i$ and the constituent $c_t$;
+	apart from $c_t$ no other information from the input graph $G$ is required by $f_h$ which makes it a local evaluation function.
 	To see why $\mathcal{A}_h$ satisfies \cref{defn:ltag:weighted-agg}, note that it inherits the idempotency property from $\wmaj$ because $\mathcal{A}_h$ ignores the bias ``pseudo-constituent'' $(y_b, w_b)$ for all threshold weights $\max w_t \geq \eta > w_b$, similarly the zero invariance of $\wmaj$ is also directly inherited.
@@
 	y_t \coloneqq \sgn z_t
-	 &= \sgn\left( \sum_{j=1}^N \underbrace{\left( \alpha_j \smashoperator[l]{\sum_{c_k \in \psi_{\varphi}(G_j)}} \sum_{i=1}^d \gamma_{\varphi,i}(G)(c_t) \gamma_{\varphi,i}(G_j)(c_k) \right)}_{w_j} y_j \right) \\
+	 &= \sgn\left( \sum_{j=1}^N \underbrace{\left( \alpha_j \smashoperator[l]{\sum_{c_k \in \psi_{\varphi}(G_j)}} \sum_{i=1}^d \gamma_{\varphi,i}(c_t) \gamma_{\varphi,i}(c_k) \right)}_{w_j} y_j \right) \\
 	 &= \wmaj(\ldblbrace (y_j, w_j)\, |\, j \in [N] \rdblbrace) \nonumber
@@
 		Here the $i$-th embedding component represent the number of occurrences of a substructure $S_i$, i.e.\ ${\varphi}_{\text{FP}}(G)[i] \coloneqq \mathit{count}_{S_i}(G)$ (see \cfullref{defn:related:wl-count-detect}).
-		Such a fingerprint embedding naturally is an \ac{sce} with the decomposition functions $\psi_{\text{FP}, i}(G) \coloneqq \{ G[\set(s)]\, |\, s \in \mathcal{V}_G^{*} \land G[s] \equiv S_i \}$ and the multiplicity functions $\gamma_{\text{FP}, i}(G)(c) \coloneqq \mathbbm{1}[c \in \psi_{\text{FP}, i}(G)]$.
+		Such a fingerprint embedding naturally is an \ac{sce} with the decomposition functions $\psi_{\text{FP}, i}(G) \coloneqq \{ G[\set(s)]\, |\, s \in \mathcal{V}_G^{*} \land G[s] \equiv S_i \}$ and the multiplicity functions $\gamma_{\text{FP}, i}(c) \coloneqq \mathbbm{1}[c \simeq S_i]$ since those functions produce the subgraphs that are counted by ${\varphi}_{\text{FP}}(G)[i]$ with multiplicity $1$.
 % chktex 21
@@
 		Thus, by \cref{thm:ltag:svm-ltag-formulation}, \acp{svm} trained on fingerprint embeddings are \ac{lta} models that produce locally explainable predictions.
-		\badgebox[t_green]{\cmark\ LTA}
+		\badgebox[t_green]{LTA}
 	\item \textbf{\texttt{graph2vec}:}
-		The skip-gram inspired \texttt{graph2vec} embedding produces vectors $\text{FP}(G) \in \mathbb{R}^d$ whose individual components do not have a clear interpretation.
-		Since the component values can be arbitrary reals, this approach is not an \ac{sce} which in turn implies that it does not have an \ac{lta} formulation as in \cref{thm:ltag:svm-ltag-formulation}.
-		\badgebox[t_red]{\xmark\ non-LTA}
+		The skip-gram inspired \texttt{graph2vec} embedding produces vectors $\varphi(G) \in \mathbb{R}^d$ whose individual components do not have a clear interpretation.
+		\texttt{graph2vec} embeds graphs that have common 1-\acs{wl} colors closer to each other than graphs that do not share the same colors (see \cfullref{eq:related:graph2vec}).
+		The resulting component values can be arbitrary reals, therefore this approach is not an \ac{sce} which in turn implies that it does not have an \ac{lta} formulation as in \cref{thm:ltag:svm-ltag-formulation}.
+		\badgebox[t_red]{non-LTA}
 	\item \textbf{\ac{wl} subtree kernel:}
-		The individual components of the \ac{wl} subtree embedding represent the number of occurrences of a color $c_i \in \mathcal{C}$, i.e.\ $\varphi_{\text{ST}}(G)[i] \coloneqq \mathit{dist}_{\chi_{G,1}^{(t)}}(c_i)$ as defined in \cfullref[ on ]{eq:related:wl-subtree-kernel}.
-		As illustrated in \cfullref[ on ]{fig:related:wl-subtree}, each occurrence of a color $c_i$ in the \ac{wl} coloring $\chi_{G,1}^{(t)}$ corresponds to the occurrence of a breadth-first subtree of depth $t$.
-		\badgebox[t_green]{\cmark\ LTA-like}
+		The individual components of the \ac{wl} subtree kernel's embedding $\varphi_{\text{ST}}$ represent the number of occurrences of a color $\kappa_i \in \mathcal{C}$, i.e.\ $\varphi_{\text{ST}}(G)[i, t] \coloneqq \mathit{dist}_{\chi_{G,1}^{(t)}}(\kappa_i)$ as defined in \cfullref[ on ]{eq:related:wl-subtree-kernel}\footnote{
+			To avoid confusion $\kappa_i \in \mathcal{C}$ is used for colors and $c_j \in \mathcal{G}$ for constituents in this context.
+		}.
+		The occurrence of a color $\kappa_i$ at some vertex $v_j \in \mathcal{V}_G$ in the $t$-th refinement step (i.e.\ $\chi_{G,1}^{(t)}(v_j) = \kappa_i$) in turn implies that the \ac{bfs} tree $S_{j,t} \coloneqq \text{\acs*{bfs}}(G, v_j, t)$ of depth $t$ rooted at $v_j$ is isomorphic to some \ac{wl} color subtree $S_{\kappa_i}$ (i.e.\ $S_{j,t} \simeq S_{\kappa_i}$).
+		This relation between \ac{wl} colors $\kappa_i$ and subtrees $S_{\kappa_i}$ was illustrated in \cfullref[ on ]{fig:related:wl-subtree}.
+
+		To show that $\varphi_{\text{ST}}$ is an \ac{sce}, we have to define decomposition functions $\psi_{\varphi, i, t}$ and multiplicity functions $\gamma_{\varphi, i, t}$ s.t.\ $\varphi(G)[i, t] = \sum_{c \in \psi_{\varphi, i, t}(G)} \gamma_{\varphi, i, t}(c)$.
+		Since $\varphi(G)[i, t]$ counts the occurrences $S_{j,t}$ of a subtree $S_{\kappa_i}$ in $G$, the \ac{sce} requirement would be trivially satisfied if each $S_{j,t}$ were a constituent with multiplicity $1$.
+		This intuition is however not quite correct because the constituents of a graph $G$ must be induced subgraphs of $G$, not \ac{bfs} subtrees $S_{j, t}$ of $G$.
+		\begin{figure}[t]
+			\centering
+			\includegraphics[width=\linewidth]{gfx/graph-lta/wl1-constituents.pdf}
+			\caption[The constituents implied by the \ac{wl} subtree kernel embedding for an example graph.]{
+				The constituents $c_{j,t}$ implied by the \ac{wl} subtree embedding vector $\varphi(G) = (\underbrace{\textcolor{t_blue}{4}, \textcolor{t_red}{0}, \textcolor{t_green}{0}, \textcolor{t_darkyellow}{0}}_{t = 0}, \underbrace{\textcolor{t_blue}{0}, \textcolor{t_red}{2}, \textcolor{t_green}{1}, \textcolor{t_darkyellow}{1}}_{t = 1})$.
+% chktex 25
+				The subtrees $S_{j,t}$ that span each constituent $c_{j,t}$ are visualized by the colored arrows that originate from the colored root vertices.
+				Because $S_{1,1}$ and $S_{2,1}$ span the same constituent $c = c_{1,1} = c_{2,1} = G[\{ v_1, v_2, v_3 \}]$, its multiplicity is $\gamma_{\varphi, \textcolor{t_red}{\texttt{B}}, 1}(c) = 2$; % chktex 25
+				the multiplicity of all other constituents is $1$.
+			}\label{fig:ltag:wl1-constituents}
+		\end{figure}
+
+		To fix the previous intuition we ``convert'' the subtrees $S_{j, t}$ into proper subgraph constituents via $c_{j,t} \coloneqq G[\mathcal{V}_{S_{j,t}}]$, i.e.\ the subgraphs that are spanned by the subtrees.
+		The resulting decomposition functions are $\psi_{\varphi, i, t}(G) \coloneqq \{ c_{j,t}\, |\, {v_j \in \mathcal{V}_G}\,\land {S_{j,t} \simeq S_{\kappa_i}} \}$. % chktex 21
+		As illustrated in \cref{fig:ltag:wl1-constituents}, the number of distinct constituents $c_{j,t}$ might be smaller than the number of \ac{bfs} subtrees $S_{j,t}$ because two distinct subtrees might span the same set of vertices.
+		To fix this discrepancy between the subtree occurrence count (which equals $\varphi(G)[i, t]$) and the number of constituents, the multiplicities $\gamma_{\varphi, i, t}(c)$ of the constituents $c$ have to correspond to the number of \ac{bfs} subtrees they were spanned up by, i.e.\
+		\begin{align*}
+			\gamma_{\varphi, i, t}(c) &\coloneqq {|\{ S = \text{\acs*{bfs}}(c, v_{\text{root}}, t)\, |\, v_{\text{root}} \in \mathcal{V}_c \land S \simeq S_{\kappa_i} \land \mathit{complete}(S) \}|} \\
+			\text{with }\mathit{complete}(S) &\Leftrightarrow \forall\ \text{leaf nodes $v_{\text{leaf}}$ of the tree $S$}: {|\Gamma_c(v_{\text{leaf}})|} = {|\Gamma_G(v_{\text{leaf}})|}
+			\text{.}
+		\end{align*}
+		The purpose of the $\mathit{complete}(S)$ condition is to only count the \ac{bfs} subtrees of the constituent $c$ that are also \ac{bfs} subtrees of the complete graph $G$.
+		To see why this is required, note that in \cref{fig:ltag:wl1-constituents} the $c_{1,1}$/$c_{2,1}$ constituent contains the color subtree $S_{\textcolor{t_red}{\texttt{B}}}$ three times (rooted at $v_1$, $v_2$ and $v_3$) even though $G$ only contains it twice (rooted at $v_1$ and $v_2$).
+% chktex 25
+		Also note that, since $\gamma_{\varphi, i, t}(c)$ must only depend on a given constituent $c$ and not on the graph $G$ it was decomposed from, the degree information $|\Gamma_G(v_{\text{leaf}})|$ of the constituent vertices $v \in \mathcal{V}_c$ is assumed to be statically encoded in the labels $l_c[v]$ or feature vectors $x_c[v]$.
+
+		Via the decomposition and multiplicity functions that were just described, $\varphi_{\text{ST}}$ is in fact an \ac{sce} with a nontrivial, subtree-based decomposition function.
+		By \cref{thm:ltag:svm-ltag-formulation} \acp{svm} that use the \ac{wl} subtree kernel therefore have nontrivial \ac{lta} formulations, i.e.\ they can be considered to be ``\ac{lta}-like'' models.
+		However, unlike fingerprint embeddings, the \ac{wl} subtree constituents are not manually chosen to be interpretable.
+		This implies that the localized explainability characteristic of \ac{lta} is only partially satisfied since the \ac{svm} predictions are based on local constituent predictions that are not necessarily interpretable.
+		\badgebox[t_green]{LTA-like}
 	\item \textbf{\ac{wl} shortest path kernel:}
-		\badgebox[t_red]{\xmark\ non-LTA}
+		\badgebox[t_red]{non-LTA}
 	\item \textbf{$k$-LWL kernel:}
-		\badgebox[t_green]{\cmark\ LTA-like}
+		\badgebox[t_green]{LTA-like}
 	\item \textbf{$k$-GWL kernel:}
-		\badgebox[t_red]{\xmark\ non-LTA}
+		\badgebox[t_red]{non-LTA}
 \end{enumerate}

### 2020-03-17 thesis/content/chapter-related-work.tex
added the lta formulation for the wl_st kernel

@@
 The context $\Gamma_T(G_i)$ of a graph $G_i$ is defined as
-\begin{equation}
+\begin{equation}\label{eq:related:graph2vec}
 	\Gamma_T(G_i) \coloneqq \bigcup_{\mathclap{v_j \in \mathcal{V}_{G_i}}}\, {\left\{ \chi_{G_i, 1}^{(t)}(v_j) \right\}}_{t = 0}^T % chktex 21
@@
 \Cref{fig:related:wl-subtree} illustrates how this definition relates to subtrees.
-Since alls vertex colors $c \in \mathcal{C}$ correspond to a subtree isomorphism class, $k_{\text{ST}}(G, H)$ effectively computes the similarity of $G$ and $H$ by comparing their local subtree structures of depth at most $T$.
+Since all vertex colors $c \in \mathcal{C}$ correspond to a subtree isomorphism class, $k_{\text{ST}}(G, H)$ effectively computes the similarity of $G$ and $H$ by comparing their local subtree structures of depth at most $T$.
 Note that the underlying mapping $\varphi_{\text{ST}}$ cannot be computed independently for each graph $G \in \mathcal{G}_{\mathcal{D}} \subseteq \mathcal{G}$ in a given training dataset $\mathcal{D}$ since the set of colors $c \in \mathcal{C}$ introduced by the hashing function $h$ is conjointly determined by all graphs $\mathcal{G}_{\mathcal{D}}$ (see \cref{defn:related:wl1-refine}).
@@
 The comparison shows that \ac{gcn} convolutions can be understood as iterative refinements of continuous ``color vectors'' $Z^{(t)}[v] \in \mathbb{R}^{d^{(t)}}$.
-Then the continuous analogue to the injective 1-\acs{wl} hash function $h: \mathcal{C}^{*} \to \mathcal{C}$ can be defined as $h_{\text{\acs*{gcn}}}(z_0, z_1, \dots, z_m) = \sigma(\sum_{j=0}^m c(v_0) c(v_j) z_j \Theta^{(t)})$ which computes a single color vector from multiple color vectors $z_j$.
+Then the continuous analogue to the injective 1-\acs{wl} hash function $h: \mathcal{C}^{*} \to \mathcal{C}$ can be defined as $h_{\text{\acs*{gcn}}}(z_0, z_1, \dots, z_m) = \sigma(\sum_{j=0}^m c(v_0) c(v_j) z_j \Theta^{(t)})$ which computes a single color vector from multiple color vectors $z_j$ where $v_j$ are the corresponding vertices.
 

### 2020-03-19 thesis/content/chapter-graph-lta.tex
added lta formualtions for wl_sp, k-lwl and k-gwl

@@
 		Thus, by \cref{thm:ltag:svm-ltag-formulation}, \acp{svm} trained on fingerprint embeddings are \ac{lta} models that produce locally explainable predictions.
-		\badgebox[t_green]{LTA}
+		\badgebox[t_green]{\acs*{lta}}
 	\item \textbf{\texttt{graph2vec}:}
@@
 		The resulting component values can be arbitrary reals, therefore this approach is not an \ac{sce} which in turn implies that it does not have an \ac{lta} formulation as in \cref{thm:ltag:svm-ltag-formulation}.
-		\badgebox[t_red]{non-LTA}
+		\badgebox[t_red]{non-\acs*{lta}}
 	\item \textbf{\ac{wl} subtree kernel:}
@@
 		\begin{align*}
-			\gamma_{\varphi, i, t}(c) &\coloneqq {|\{ S = \text{\acs*{bfs}}(c, v_{\text{root}}, t)\, |\, v_{\text{root}} \in \mathcal{V}_c \land S \simeq S_{\kappa_i} \land \mathit{complete}(S) \}|} \\
+			\gamma_{\varphi, i, t}(c) &\coloneqq {\left|\{ S = \text{\acs*{bfs}}(c, v_{\text{root}}, t)\, |\, v_{\text{root}} \in \mathcal{V}_c \land S \simeq S_{\kappa_i} \land \mathit{complete}(S) \}\right|} \\
 			\text{with }\mathit{complete}(S) &\Leftrightarrow \forall\ \text{leaf nodes $v_{\text{leaf}}$ of the tree $S$}: {|\Gamma_c(v_{\text{leaf}})|} = {|\Gamma_G(v_{\text{leaf}})|}
@@
 		Via the decomposition and multiplicity functions that were just described, $\varphi_{\text{ST}}$ is in fact an \ac{sce} with a nontrivial, subtree-based decomposition function.
-		By \cref{thm:ltag:svm-ltag-formulation} \acp{svm} that use the \ac{wl} subtree kernel therefore have nontrivial \ac{lta} formulations, i.e.\ they can be considered to be ``\ac{lta}-like'' models.
+		Additionally each constituent's diameter is upper bounded by $2T$ (with $T$ being the maximum number of \ac{wl} iterations) which guarantees that constituents are localized within a neighborhood of bounded size.
+		By \cref{thm:ltag:svm-ltag-formulation}, \acp{svm} that use the \ac{wl} subtree kernel therefore have nontrivial \ac{lta} formulations with localized constituents, i.e.\ they can be considered to be ``\ac{lta}-like'' models.
 		However, unlike fingerprint embeddings, the \ac{wl} subtree constituents are not manually chosen to be interpretable.
 		This implies that the localized explainability characteristic of \ac{lta} is only partially satisfied since the \ac{svm} predictions are based on local constituent predictions that are not necessarily interpretable.
-		\badgebox[t_green]{LTA-like}
+		\badgebox[t_green]{\acs*{lta}-like}
 	\item \textbf{\ac{wl} shortest path kernel:}
-		\badgebox[t_red]{non-LTA}
+		As defined in \cfullref[ on ]{eq:related:wl-shortest-path-kernel}, the $i$-th component of the \ac{wl} shortest path embedding $\varphi_{\text{SP}}$ represents a 4-tuple $(t_i, a_i, b_i, d_i) \in {\{ 0, \dots, T \}} \times \mathcal{C} \times \mathcal{C} \times \mathbb{N}_0$.
+		The value $\varphi_{\text{SP}}(G)[i]$ is defined as the number of vertex pairs $v_a, v_b \in \mathcal{V}_G$ that have a shortest connecting path of length $d_i$ and the \ac{wl} color combination $\chi_{G,1}^{(t_i)}(v_a) = a_i$ and $\chi_{G,1}^{(t_i)}(v_b) = b_i$.
+
+		To determine the number of such vertex pairs via an \ac{sce} multiplicity function $\gamma_{\varphi, t_i, a_i, b_i, d_i}$, each connected pair of vertices $v_a, v_b$ and the shortest path between them must occur together in at least one constituent, otherwise a multiplicity function cannot compute whether $v_a$ and $v_b$ are in fact $d_i$ hops apart.
+		One simple decomposition which guarantees that all shortest paths are part of at least one constituent simply splits a given graph into its connected components.
+		Even though such a decomposition is non-trivial since it uses at least some structural information to determine the set of constituents, the fact that any pair of connected vertices must co-occur within a single constituent means that constituents must span arbitrarily large distances within a given graph.
+		Depending on ones domain-specific interpretation of \textit{localized explainability}, this restriction can be seen to be not ``\acs{lta}-like''.
+		Since we do not attempt to clearly separate \ac{lta} from non-\acs{lta} methods, $\varphi_{\text{SP}}$ is categorized as an in-between case here.
+		\badgebox[t_yellow]{partially \acs*{lta}-like}
 	\item \textbf{$k$-LWL kernel:}
-		\badgebox[t_green]{LTA-like}
+		The \ac{lta} interpretation of the $k$-LWL kernel is identical to that of the \ac{wl} subtree kernel.
+		The \ac{wl} color of a $k$-multiset of vertices $s$ after $t$ refinement steps is described by the joint $t$-neighborhood of all vertices $v \in s$, i.e.\ $\Gamma_{G}^t(s) \coloneqq \bigcup_{v \in s} \Gamma_{G}^t(v)$.
+		Analogous to \ac{wl} subtree embeddings, those $t$-neighborhoods form the localized constituents of the $k$-LWL kernel via a \ac{bfs} over neighboring $k$-multisets.
+		Consequently \acp{svm} using the $k$-LWL kernel can be interpreted as ``\acs{lta}-like'' methods with nontrivial localized decompositions.\\
+		\badgebox[t_green]{\acs*{lta}-like}
 	\item \textbf{$k$-GWL kernel:}
-		\badgebox[t_red]{non-LTA}
+		The only difference between the $k$-LWL and the $k$-GWL kernel is their definition of the $k$-multiset neighborhood.
+		Namely the refined color of a multiset $s \subseteq \mathcal{V}_G$ in $k$-GWL depends on the the colors of all vertices $\mathcal{V}_G$ since all multisets $s' = s \setminus \{ v \} \cup \{ u \}$ with $v \in s$ and $u \in \mathcal{V}_G$ are neighbors of $s$.
+		Because of those global multiset neighborhoods, all the color subtrees of a graph span the entire graph.
+		Therefore the $k$-GWL embedding is an \ac{sce} with only the trivial decomposition function $\psi_{\varphi}(G) = \{ G \}$.
+		Thus the \ac{lta} formulation of $k$-GWL \acp{svm} described in \cref{thm:ltag:svm-ltag-formulation} is also trivial which means that they are not \ac{lta} models.
+		\badgebox[t_red]{non-\acs*{lta}}
 \end{enumerate}
+
+This concludes our overview of \ac{lta} interpretations of \acp{svm} that use graph embeddings/kernels.
+Among the described approaches, fingerprint embeddings, the \ac{wl} subtree kernel and the $k$-LWL kernel were shown to have \ac{lta} formulations with nontrivial local decomposition functions.
+Approaches like \texttt{graph2vec}, the \ac{wl} shortest path kernel or $k$-GWL on the other hand, were shown to be less compatible with \ac{lta}.
 
@@
 
+Now we will look at the \ac{gcnn} approaches described in \cref{sec:related:gcr:nn} from an \ac{lta} perspective.
+While the \ac{svm} \ac{lta} formulation from \cref{thm:ltag:svm-ltag-formulation} uses a single fixed aggregation function, namely $\wmaj$, \ac{gcnn} architectures can use many other aggregation functions.
+As illustrated in x, this is due to the fact that the structure of \acp{gcnn} aligns more closely to that of \ac{lta}:
+The decomposition and local evaluation functions are implicitly defined by the convolutional layers, the aggregation function directly corresponds to the pooling layer in existing architectures.
+
 \section{A Novel \acs*{lta}-Inspired \acs*{gcnn} Architecture}%

### 2020-03-20 thesis/content/chapter-graph-lta.tex
started with description of gcnn lta relation

@@
 Now we will look at the \ac{gcnn} approaches described in \cref{sec:related:gcr:nn} from an \ac{lta} perspective.
-While the \ac{svm} \ac{lta} formulation from \cref{thm:ltag:svm-ltag-formulation} uses a single fixed aggregation function, namely $\wmaj$, \ac{gcnn} architectures can use many other aggregation functions.
-As illustrated in x, this is due to the fact that the structure of \acp{gcnn} aligns more closely to that of \ac{lta}:
-The decomposition and local evaluation functions are implicitly defined by the convolutional layers, the aggregation function directly corresponds to the pooling layer in existing architectures.
+As described in the last section, \acp{svm} can be interpreted as \ac{lta} methods if they use an \ac{sce} embedding;
+similarly \acp{gcnn} also have \ac{lta} formulations under certain conditions that are described in this section.
+
+While the \ac{svm} \ac{lta} formulation from \cref{thm:ltag:svm-ltag-formulation} uses a single fixed aggregation function, namely $\wmaj$, the aggregation function in \ac{gcnn} architectures can be chosen more freely.
+This is due to the fact that \acp{gcnn} have an \ac{lta} formulation in which the convolutional layers implicitly define the decomposition and local evaluation functions while the pooling layer directly corresponds to the \ac{lta} aggregation function.
+This correspondence is illustrated in \cref{fig:ltag:gcnn-structure}.
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=0.8\linewidth]{gfx/graph-lta/gcnn-structure.pdf}
+	\caption{
+		An \ac{lta} perspective on \acp{gcnn} that use vertex neighborhood convolutions.
+	}\label{fig:ltag:gcnn-structure}
+\end{figure}
+As described in \cref{sec:related:gcr:nn}, a stack of $T$ vertex convolution layers computes a feature vector $z_i \in \mathbb{R}^{(T)}$ for all vertices $v_i \in \mathcal{V}_G$ of an input graph $G$.
+Depending on the specific type of convolutional layer, each of those feature vectors $z_i$ is the result of some subset of the input feature vectors ${\{ x_i \}}_{v_i \in \mathcal{V}_G}$:
+\begin{enumerate}[label=\textbf{\arabic*.}]
+	\item \textbf{Spectral convolutions with arbitrary filters:}
+	\item \textbf{Spatial vertex convolutions:}
+		If the convolutional layers use a neighborhood aggregation scheme, the receptive field of vertices that have an influence on $z_i$ is equal to the 1-\acs{wl} subtree of $v_i$ after $T$ refinement steps as described by \citet{Xu2018}.
+		Similar to the \ac{wl} subtree kernel, each $z_i \in \mathbb{R}^{(T)}$ can therefore be interpreted as the evaluation of a localized subtree.
+\end{enumerate}
 

### 2020-03-21 thesis/content/chapter-graph-lta.tex
added spectral gcnn lta relation

--- 2020-03-19 thesis/content/chapter-graph-lta.tex
@@
 	y_t \coloneqq \sgn z_t
-	 &= \sgn\left( \sum_{j=1}^N \underbrace{\left( \alpha_j \smashoperator[l]{\sum_{c_k \in \psi_{\varphi}(G_j)}} \sum_{i=1}^d \gamma_{\varphi,i}(c_t) \gamma_{\varphi,i}(c_k) \right)}_{w_j} y_j \right) \\
+	 &= \sgn\left( \sum_{j=1}^N \underbrace{\left( \alpha_j \smashoperator[l]{\sum_{c_k \in \psi_{\varphi}(G_j)}} \sum_{i=1}^d \gamma_{\varphi,i}(c_t) \gamma_{\varphi,i}(c_k) \right)}_{w_j \geq 0} y_j \right) \\
 	 &= \wmaj(\ldblbrace (y_j, w_j)\, |\, j \in [N] \rdblbrace) \nonumber
@@
 
-Now we will look at the \ac{gcnn} approaches described in \cref{sec:related:gcr:nn} from an \ac{lta} perspective.
-While the \ac{svm} \ac{lta} formulation from \cref{thm:ltag:svm-ltag-formulation} uses a single fixed aggregation function, namely $\wmaj$, \ac{gcnn} architectures can use many other aggregation functions.
-As illustrated in x, this is due to the fact that the structure of \acp{gcnn} aligns more closely to that of \ac{lta}:
-The decomposition and local evaluation functions are implicitly defined by the convolutional layers, the aggregation function directly corresponds to the pooling layer in existing architectures.
+Now we will look at \ac{gcnn} methods from an \ac{lta} perspective.
+As shown in the last section, \acp{svm} can be interpreted as \ac{lta} methods if they use an \ac{sce} embedding;
+similarly \acp{gcnn} also have \ac{lta} formulations under certain conditions.
+In this section we will describe those conditions and the \ac{lta} formulations of the \acp{gcnn} that satisfy them.
+
+In \cref{sec:related:gcr:nn} an overview of graph convolution and graph pooling layers was given.
+While there are many ways to combine those layers, we focus on global pooling \ac{gcnn} architectures that use the following three steps:
+\begin{enumerate}[label=\textbf{\arabic*.}]
+	\item \textbf{Convolution:}
+		First a stack of $T$ convolutional layers is applied to the input feature matrix $X \in \mathbb{R}^{n \times d^{(0)}}$ where each row is a feature vector $x_i \in \mathbb{R}^d$ for the vertex $v_i \in \mathcal{V}_G$ or the vertex $k$-multiset $s_i \subseteq \mathcal{V}_G$ in case of $k$-\acsp{gnn}.
+		The convolutional layers produce a convolved feature matrix $Z \in \mathbb{R}^{n \times d^{(T)}}$.
+	\item \textbf{Pooling:}
+		Then the convolved feature matrix $Z$ is reduced to a single graph feature vector $z_G \in \mathbb{R}^{d^{(T)}}$ via a pooling layer.
+	\item \textbf{\ac{mlp}:}
+		Lastly the final output $y_G \in \mathcal{Y}$ is computed by applying a standard \ac{mlp} to the graph feature vector $z_G$.
+\end{enumerate}
+\Cref{fig:ltag:gcnn-structure} illustrates this three step architecture for vertex neighborhood convolutions like those used by \acp{gcn}~\cite{Kipf2017} or \acp{gin}~\cite{Xu2018}.
+Intuitively the convolutional layers implicitly define the decomposition and local evaluation functions while the pooling layer corresponds to the \ac{lta} aggregation function.
+This intuition is now formalized.
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=\linewidth]{gfx/graph-lta/gcnn-structure.pdf}
+	\caption{
+		Computational steps of \acp{gcnn} that use vertex neighborhood convolutions.
+	}\label{fig:ltag:gcnn-structure}
+\end{figure}
+
+\subsubsection{Graph Convolutions as \ac*{lta} Decomposition and Evaluation Functions}
+
+In order to interpret a stack of convolutional layers as a representation of an \ac{lta} decomposition function $\psi: \mathcal{G} \to \mathcal{P}(\mathcal{G})$ and evaluation function $f: \mathcal{G} \to \mathcal{Y} \times \mathbb{R}_{\geq 0}$, there must be a mapping $\tau: \mathbb{R}^{* \times d^{(T)}} \to {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*}$ s.t.\ $\tau(Z) = {\ldblbrace f(c_i) \rdblbrace}_{c_i \in \psi(G)}$.
+Whether such a translation map $\tau$ from convolved feature matrices to constituent evaluations exists, depends on the specific type of convolutional layer that is used.
+
+\paragraph{Spectral convolutions with arbitrary filters}
+Let us begin with general spectral graph convolutions~\cite{Bruna2013}\cite{Henaff2015}.
+Following \cfullref[ on ]{eq:related:graph-conv-general}, after a single convolution the convolved feature vector $z_i$ of a vertex $v_i$ is defined as
+\begin{align}\label{eq:ltag:general-spectral-conv}
+	z_i = {\left( U^{\top} \hat{g}(\Lambda) U X \right)}_i = \sum_{k=1}^n \underbrace{\left( \sum_{l=1}^n \hat{g}(\lambda_l) u_{l,k} u_{l,i} \right)}_{\alpha_k} x_k
+\end{align}
+where $L = U^{\top} \Lambda U$ is the eigendecomposition of the input graph's Laplacian and $\hat{g}$ is some eigenvalue filter.
+\begin{lem}\label{lem:ltag:general-spectral-conv-nonlocality}
+	$\exists \eta \in \mathbb{R}$ s.t.\ if $\hat{g}(0) = \eta$, every $z_i$ depends on all input vertex features $x_k$.
+\end{lem}
+\begin{proof}
+	Note that by \cfullref[ on ]{defn:related:laplacian-comb} it follows that $u_1 = \frac{1}{\sqrt{n}}{(1, \dots, 1)}^{\top}$ is an eigenvector with eigenvalue $\lambda_1 = 0$ of the combinatorial Laplacian as well as the symmetrically normalized Laplacian.
+	Consequently in \cref{eq:ltag:general-spectral-conv} the first summand of each $\alpha_k$ equals $\hat{g}(0) u_{1,k} u_{1,i} = \frac{\eta}{n}$ for all $i \in [n]$.
+	The lemma then follows by choosing $\eta$ s.t.\ all $\alpha_k \neq 0$.
+\end{proof}
+Consequently because a \ac{gcnn} with an arbitrary spectral filter $\hat{g}$ can use $\hat{g}(0) = \eta$, the convolved vertex feature vectors ${\{ z_i \}}_{v_i \in \mathcal{V}_G}$ all depend on the entire set of input features ${\{ x_i \}}_{v_i \in \mathcal{V}_G}$.
+This in turn implies that no entry in the convolved vertex feature matrix $Z$ is based on a proper subset of the input vertices.
+If we want to interpret any entry of $Z$ as the score or weight of a constituent, this constituent has to span the entire input graph since all vertex features potentially had an influence on them.
+In other words, spectral \acp{gcnn} with arbitrary filters only have trivial \ac{lta} formulations which makes them non-\acs{lta} methods.
+
+\paragraph{Vertex neighborhood convolutions}
+
+\paragraph{Vertex $k$-multiset neighborhood convolutions}
+
+\subsubsection{Graph Pooling as \ac*{lta} Aggregation Functions}
 

### 2020-03-21 thesis/content/chapter-related-work.tex
added lta formualtions for wl_sp, k-lwl and k-gwl
started with description of gcnn lta relation
added spectral gcnn lta relation

@@
 In practice the hash function $h$ is usually defined lazily by using $\mathcal{C} \subseteq \mathbb{N}$ and enumerating color combinations in whichever order they are hashed s.t.\ a new color is introduced every time a previously unseen color combination appears at runtime\footnote{
-	Though $\mathcal{C}$ is countably infinite, we assume it only contains the finite no.\ of colors in a dataset.
+	$\mathcal{C}$ is countably infinite but we assume it only contains the finite no.\ of colors occurring in a dataset.
 }.
@@
 \paragraph{Interpretation of the spectrum}
-By convention we assume that the eigenvalues are ordered ascendingly: $\lambda_{G,0} \leq \cdots \leq \lambda_{G,n}$.
+By convention we assume that the eigenvalues are ordered ascendingly: $\lambda_{G,1} \leq \cdots \leq \lambda_{G,n}$.
 Those eigenvalues are called the \textit{spectrum of $G$} and they are a vertex-permutation-invariant graph characterization.
@@
 % chktex 25
+		An arrow $A \to B$ indicates that the refined color of $A$ depends on the color of $B$.
 		\source[based~on]{Shervashidze2011}
@@
 \begin{align}
-	{\varphi_{\text{SP}}(G)}[i] \coloneqq \left| \left\{ (v, u) \in \mathcal{V}_G^2 \,|\, {\chi_{G,1}^{(t_i)}(v) = a_i} \land {\chi_{G,1}^{(t_i)}(u) = b_i} \land {d_{\text{SP}}(v, u) = d_i} \right\} \right| \text{,} % chktex 21
+	{\varphi_{\text{SP}}(G)}[i] \coloneqq \left| \left\{ (v, u) \in \mathcal{V}_G^2 \,|\, {\chi_{G,1}^{(t_i)}(v) = a_i} \land {\chi_{G,1}^{(t_i)}(u) = b_i} \land {d_{\text{SP}}(v, u) = d_i} \right\} \right| \text{,} \label{eq:related:wl-shortest-path-kernel} % chktex 21
 \end{align}
@@
 \begin{align*}
-	\text{\ac*{gcn}:}\quad&& Z^{(t)}[v] &= \sigma\left(\left( {c(v)}^2 Z^{(t-1)}[v] + \smashoperator[lr]{\sum_{u \in \Gamma_G(v)}} c(v) c(u) Z^{(t-1)}[u] \right) \Theta^{(t)} \right) \\
+	\text{\ac*{gcn}:}\quad&& Z^{(t)}[v] &= \sigma\left(\left( {\mu(v)}^2 Z^{(t-1)}[v] + \smashoperator[lr]{\sum_{u \in \Gamma_G(v)}} \mu(v) \mu(u) Z^{(t-1)}[u] \right) \Theta^{(t)} \right) \\
 	\text{1-\acs{wl}:}\quad&& \chi_{G,1}^{(t)}(v) &= h\left( \chi_{G,1}^{(t-1)}(v), \ldblbrace \chi_{G,1}^{(t-1)}(u)\, |\, u \in \Gamma_{G}(v) \rdblbrace\right)
 \end{align*}
-Here $c(v) \coloneqq {\left(\left| \Gamma_G(v) \right| + 1\right)}^{-\frac{1}{2}}$ are normalization factors introduced by $\tilde{D}_G^{-\frac{1}{2}}$.
+Here $\mu(v) \coloneqq {\left(\left| \Gamma_G(v) \right| + 1\right)}^{-\frac{1}{2}}$ are normalization factors introduced by $\tilde{D}_G^{-\frac{1}{2}}$.
 The comparison shows that \ac{gcn} convolutions can be understood as iterative refinements of continuous ``color vectors'' $Z^{(t)}[v] \in \mathbb{R}^{d^{(t)}}$.
-Then the continuous analogue to the injective 1-\acs{wl} hash function $h: \mathcal{C}^{*} \to \mathcal{C}$ can be defined as $h_{\text{\acs*{gcn}}}(z_0, z_1, \dots, z_m) = \sigma(\sum_{j=0}^m c(v_0) c(v_j) z_j \Theta^{(t)})$ which computes a single color vector from multiple color vectors $z_j$ where $v_j$ are the corresponding vertices.
+Then the continuous analogue to the injective 1-\acs{wl} hash function $h: \mathcal{C}^{*} \to \mathcal{C}$ can be defined as $h_{\text{\acs*{gcn}}}(z_0, z_1, \dots, z_m) = \sigma(\sum_{j=0}^m \mu(v_0) \mu(v_j) z_j \Theta^{(t)})$ which computes a single color vector from multiple color vectors $z_j$ where $v_j$ are the corresponding vertices.
 
@@
 \end{align}
-By leaving out the normalization factors $c(v)$ and by using a \ac{mlp} instead of the single fully-connected layer $\sigma \circ \Theta^{(t)}$ the resulting hashing function $h_{\text{\acs*{gin}}}(z_1, \dots, z_m) = \text{\acs*{mlp}}^{(t)}(\sum_{j=1}^{m} z_j)$ becomes injective\footnote{
+By leaving out the normalization factors $\mu(v)$ and by using a \ac{mlp} instead of the single fully-connected layer $\sigma \circ \Theta^{(t)}$ the resulting hashing function $h_{\text{\acs*{gin}}}(z_1, \dots, z_m) = \text{\acs*{mlp}}^{(t)}(\sum_{j=1}^{m} z_j)$ becomes injective\footnote{
 	Under the assumption that the set of possible input feature matrices $X \in \mathbb{R}^{n \times d^{(0)}}$ is countable.

### 2020-03-22 thesis/content/chapter-graph-lta.tex
started with spatial gcnn lta relation

@@
 To see what \cref{thm:ltag:svm-ltag-formulation} implies in practice, we will now apply it to the graph embedding approaches described in \cref{sec:related:gcr:embed,sec:related:gcr:kernel}:
-\begin{enumerate}[label=\textbf{\arabic*.}]
-	\item \textbf{Fingerprint embeddings:}
+\begin{enumerate}[label=\textbf{\arabic*.},ref={\arabic*}]
+	\item \textbf{Fingerprint embeddings:}\label{itm:ltag:fingerprint-lta-formulation}
 		Here the $i$-th embedding component represent the number of occurrences of a substructure $S_i$, i.e.\ ${\varphi}_{\text{FP}}(G)[i] \coloneqq \mathit{count}_{S_i}(G)$ (see \cfullref{defn:related:wl-count-detect}).
@@
 		\badgebox[t_red]{non-\acs*{lta}}
-	\item \textbf{\ac{wl} subtree kernel:}
+	\item \textbf{\ac{wl} subtree kernel:}\label{itm:ltag:wlst-lta-formulation}
 		The individual components of the \ac{wl} subtree kernel's embedding $\varphi_{\text{ST}}$ represent the number of occurrences of a color $\kappa_i \in \mathcal{C}$, i.e.\ $\varphi_{\text{ST}}(G)[i, t] \coloneqq \mathit{dist}_{\chi_{G,1}^{(t)}}(\kappa_i)$ as defined in \cfullref[ on ]{eq:related:wl-subtree-kernel}\footnote{
@@
 		\badgebox[t_green]{\acs*{lta}-like}
-	\item \textbf{\ac{wl} shortest path kernel:}
+	\item \textbf{\ac{wl} shortest path kernel:}\label{itm:ltag:wlsp-lta-formulation}
 		As defined in \cfullref[ on ]{eq:related:wl-shortest-path-kernel}, the $i$-th component of the \ac{wl} shortest path embedding $\varphi_{\text{SP}}$ represents a 4-tuple $(t_i, a_i, b_i, d_i) \in {\{ 0, \dots, T \}} \times \mathcal{C} \times \mathcal{C} \times \mathbb{N}_0$.
@@
 		\badgebox[t_yellow]{partially \acs*{lta}-like}
-	\item \textbf{$k$-LWL kernel:}
+	\item \textbf{$k$-LWL kernel:}\label{itm:ltag:klwl-lta-formulation}
 		The \ac{lta} interpretation of the $k$-LWL kernel is identical to that of the \ac{wl} subtree kernel.
-		The \ac{wl} color of a $k$-multiset of vertices $s$ after $t$ refinement steps is described by the joint $t$-neighborhood of all vertices $v \in s$, i.e.\ $\Gamma_{G}^t(s) \coloneqq \bigcup_{v \in s} \Gamma_{G}^t(v)$.
+		The \ac{wl} color of a $k$-multiset of vertices $s$ after $t$ refinement steps is described by the joint $t$-neighborhood of all vertices $v \in s$, i.e.\ by $\Gamma_{G}^t(s) \coloneqq \bigcup_{v \in s} \Gamma_{G}^t(v)$ where $\Gamma_{G}^t(v) \coloneqq \Gamma_{G}^{t-1}(v) \cup \bigcup_{u \in \Gamma_{G}(v)} \Gamma_{G}^{t-1}(u)$ and $\Gamma_{G}^0(v) \coloneqq \{ v \}$.
 		Analogous to \ac{wl} subtree embeddings, those $t$-neighborhoods form the localized constituents of the $k$-LWL kernel via a \ac{bfs} over neighboring $k$-multisets.
-		Consequently \acp{svm} using the $k$-LWL kernel can be interpreted as ``\acs{lta}-like'' methods with nontrivial localized decompositions.\\
+		Consequently \acp{svm} using the $k$-LWL kernel can be interpreted as ``\acs{lta}-like'' methods with nontrivial localized decompositions.
 		\badgebox[t_green]{\acs*{lta}-like}
-	\item \textbf{$k$-GWL kernel:}
+	\item \textbf{$k$-GWL kernel:}\label{itm:ltag:kgwl-lta-formulation}
 		The only difference between the $k$-LWL and the $k$-GWL kernel is their definition of the $k$-multiset neighborhood.
@@
 \end{enumerate}
+\[
+	\text{\acs*{gcnn}}(G) \coloneqq \text{\acs*{mlp}}(\text{Pooling}(\text{Convolution}(G)))
+\]
 \Cref{fig:ltag:gcnn-structure} illustrates this three step architecture for vertex neighborhood convolutions like those used by \acp{gcn}~\cite{Kipf2017} or \acp{gin}~\cite{Xu2018}.
-Intuitively the convolutional layers implicitly define the decomposition and local evaluation functions while the pooling layer corresponds to the \ac{lta} aggregation function.
+Intuitively the convolutional layers implicitly define an \ac{lta} decomposition function $\psi$ and a local evaluation function $f$ while the pooling layer corresponds to an \ac{lta} aggregation function $\mathcal{A}$.
 This intuition is now formalized.
@@
 	\includegraphics[width=\linewidth]{gfx/graph-lta/gcnn-structure.pdf}
-	\caption{
-		Computational steps of \acp{gcnn} that use vertex neighborhood convolutions.
+	\caption[Computational steps of a \ac{gcnn} that uses a single vertex neighborhood convolution layer.]{
+		Computational steps of a \ac{gcnn} that uses a single vertex neighborhood convolution layer.
+		The colored \ac{bfs} trees next to each convolved feature vector $z_i$ show the vertices $v_j$ whose input feature vectors $x_j$ were used to compute $z_i$.
 	}\label{fig:ltag:gcnn-structure}
@@
 In order to interpret a stack of convolutional layers as a representation of an \ac{lta} decomposition function $\psi: \mathcal{G} \to \mathcal{P}(\mathcal{G})$ and evaluation function $f: \mathcal{G} \to \mathcal{Y} \times \mathbb{R}_{\geq 0}$, there must be a mapping $\tau: \mathbb{R}^{* \times d^{(T)}} \to {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*}$ s.t.\ $\tau(Z) = {\ldblbrace f(c_i) \rdblbrace}_{c_i \in \psi(G)}$.
-Whether such a translation map $\tau$ from convolved feature matrices to constituent evaluations exists, depends on the specific type of convolutional layer that is used.
+Whether such a translation map $\tau$ from convolved feature matrices $Z$ to constituent evaluations exists, depends on the specific type of convolutional layer that is used.
 
@@
 \begin{align}\label{eq:ltag:general-spectral-conv}
-	z_i = {\left( U^{\top} \hat{g}(\Lambda) U X \right)}_i = \sum_{k=1}^n \underbrace{\left( \sum_{l=1}^n \hat{g}(\lambda_l) u_{l,k} u_{l,i} \right)}_{\alpha_k} x_k
+	z_i = {\left( U^{\top} \hat{g}(\Lambda) U X \right)}_i
+	= \sum_{j=1}^n \underbrace{\left( \sum_{l=1}^n \hat{g}(\lambda_l) u_{l,i} u_{l,j} \right)}_{\alpha_j} x_j
 \end{align}
@@
 \begin{lem}\label{lem:ltag:general-spectral-conv-nonlocality}
-	$\exists \eta \in \mathbb{R}$ s.t.\ if $\hat{g}(0) = \eta$, every $z_i$ depends on all input vertex features $x_k$.
+	$\exists \eta \in \mathbb{R}$ s.t.\ if $\hat{g}(0) = \eta$, every $z_i$ depends on all input vertex features $x_j$.
 \end{lem}
 \begin{proof}
-	Note that by \cfullref[ on ]{defn:related:laplacian-comb} it follows that $u_1 = \frac{1}{\sqrt{n}}{(1, \dots, 1)}^{\top}$ is an eigenvector with eigenvalue $\lambda_1 = 0$ of the combinatorial Laplacian as well as the symmetrically normalized Laplacian.
-	Consequently in \cref{eq:ltag:general-spectral-conv} the first summand of each $\alpha_k$ equals $\hat{g}(0) u_{1,k} u_{1,i} = \frac{\eta}{n}$ for all $i \in [n]$.
-	The lemma then follows by choosing $\eta$ s.t.\ all $\alpha_k \neq 0$.
+	Note that by \cfullref[ on ]{defn:related:laplacian-comb} it follows that $u_1 = \frac{1}{\sqrt{n}}{(1, \dots, 1)}^{\top}$ is an eigenvector with eigenvalue $\lambda_1 = 0$ of the combinatorial Laplacian as well as the symmetric normalized Laplacian\footnote{
+		This is illustrated in \cfullref[ on ]{fig:related:graph-fourier}. We refer to \citet[section II.C.]{Shuman2013} for a more in-depth discussion of this fact.
+	}.
+	Consequently in \cref{eq:ltag:general-spectral-conv} the first summand of each $\alpha_j$ equals $\hat{g}(0) u_{1,j} u_{1,i} = \frac{\eta}{n}$ for all $i \in [n]$.
+	The lemma then follows by choosing $\eta$ s.t.\ all $\alpha_j \neq 0$.
 \end{proof}
-Consequently because a \ac{gcnn} with an arbitrary spectral filter $\hat{g}$ can use $\hat{g}(0) = \eta$, the convolved vertex feature vectors ${\{ z_i \}}_{v_i \in \mathcal{V}_G}$ all depend on the entire set of input features ${\{ x_i \}}_{v_i \in \mathcal{V}_G}$.
-This in turn implies that no entry in the convolved vertex feature matrix $Z$ is based on a proper subset of the input vertices.
-If we want to interpret any entry of $Z$ as the score or weight of a constituent, this constituent has to span the entire input graph since all vertex features potentially had an influence on them.
-In other words, spectral \acp{gcnn} with arbitrary filters only have trivial \ac{lta} formulations which makes them non-\acs{lta} methods.
+Consequently because a \ac{gcnn} with an arbitrary learned spectral filter $\hat{g}$ can learn $\hat{g}(0) = \eta$, the convolved vertex feature vectors ${\{ z_i \}}_{v_i \in \mathcal{V}_G}$ potentially all depend on the complete set of input feature vectors ${\{ x_i \}}_{v_i \in \mathcal{V}_G}$.
+Therefore all constituent scores and weights computed by a mapping $\tau(Z)$ can also depend on all input feature vectors, i.e.\ the corresponding constituents must span the entire input graph.
+In other words, spectral \acp{gcnn} with arbitrary filters do not generally have nontrivial \ac{lta} formulations which makes them non-\acs{lta} methods.
 
 \paragraph{Vertex neighborhood convolutions}
+Unlike spectral convolutions, spatial approaches like \acp{gcn} or \acp{gin} compute convolved feature vectors $z_i$ which only depend on the $T$-neighborhood of each vertex $v_i \in \mathcal{V}_G$.
+Just like the \ac{bfs} subtree evaluations of \ac{wl} subtree kernel \acp{svm} (see \cfullref{itm:ltag:wlst-lta-formulation}), each $z_i$ can therefore be interpreted as an evaluation of a local constituent $c_i$.
+As previously described, this means that two evaluations $z_i, z_j$ with $i \neq j$ are computed for a single subgraph constituent $c_i = c_j$ iff.\ the \ac{bfs} trees of $v_i$ and $v_j$ span the same vertices.
+This is illustrated in \cref{fig:ltag:gcnn-structure} by \textcolor{t_red}{$z_1$} and \textcolor{t_red}{$z_2$} which are both evaluations of the constituent
+$\textcolor{t_red}{c_1} = \textcolor{t_red}{c_2} = G[\{ v_1, v_2, v_3 \}]$. % chktex 25
 
 \paragraph{Vertex $k$-multiset neighborhood convolutions}
+Analogously to single vertex convolutions, the $k$-multiset convolutions of $k$-\acp{gnn} produce local evaluation vectors $z_i$ for the $T$-neighborhood of each $k$-multiset $s_i \subseteq \mathcal{V}_G$.
+By \cfullref[ on ]{eq:related:kgnn-layer} the constituent $c_i$ spanned by the vertices that have an influence on $z_i$ can thus be written as $c_i = G\left[ \bigcup_{v \in s_i} \Gamma_G^T(v) \right]$.
+This constituent definition is identical to that used in the \ac{lta} formulation of $k$-LWL \acp{svm} (see \cfullref{itm:ltag:klwl-lta-formulation}).
+
+Both, the single vertex neighborhood convolution as well as the vertex $k$-multiset neighborhood convolution produce vectors $z_i \in \mathbb{R}^{d^{(T)}}$, each of which is based on a local constituent $c_i$ that might not be unique, i.e.\ depending on the input graph there can be $j \neq i$ with $c_j = c_i$.
 

### 2020-03-22 thesis/content/chapter-related-work.tex
started with spatial gcnn lta relation

@@
 \end{lem}
-\begin{hproof}
+\begin{proof}
 	All $k$-tuples can be mapped to $(k+1)$-tuples via $\varphi(v_1, \dots, v_k) \coloneqq (v_1, \dots,\allowbreak v_k, v_k)$.
@@
 	The lemma then follows by \cref{defn:related:wl-distinguishable}.
-\end{hproof}
+\end{proof}
 \begin{prop}[see~\citet{Immerman1990} for the proof]\label{prop:related:wl-upper}
@@
 \end{cor}
-\begin{hproof}
+\begin{proof}
 	Note that $k$-\acs{wl} trivially solves the \ac{gi} problem for all graphs of size $\leq k$ via the initial coloring (see~\cref{eq:related:wlk-init}).
 	Using \cref{lem:related:wl-lower} and \cref{prop:related:wl-upper} the corollary directly follows.
-\end{hproof}
+\end{proof}
 \begin{figure}[ht]

### 2020-03-23 thesis/content/chapter-graph-lta.tex
continued with spatial gcnn lta relation

@@
 		To fix this discrepancy between the subtree occurrence count (which equals $\varphi(G)[i, t]$) and the number of constituents, the multiplicities $\gamma_{\varphi, i, t}(c)$ of the constituents $c$ have to correspond to the number of \ac{bfs} subtrees they were spanned up by, i.e.\
-		\begin{align*}
-			\gamma_{\varphi, i, t}(c) &\coloneqq {\left|\{ S = \text{\acs*{bfs}}(c, v_{\text{root}}, t)\, |\, v_{\text{root}} \in \mathcal{V}_c \land S \simeq S_{\kappa_i} \land \mathit{complete}(S) \}\right|} \\
-			\text{with }\mathit{complete}(S) &\Leftrightarrow \forall\ \text{leaf nodes $v_{\text{leaf}}$ of the tree $S$}: {|\Gamma_c(v_{\text{leaf}})|} = {|\Gamma_G(v_{\text{leaf}})|}
-			\text{.}
-		\end{align*}
+		\begin{align}
+			&\gamma_{\varphi, i, t}(c) \coloneqq {\left|\{ S = \text{\acs*{bfs}}(c, v_{\text{root}}, t)\, |\, v_{\text{root}} \in \mathcal{V}_c \land S \simeq S_{\kappa_i} \land \mathit{complete}(S) \}\right|}\label{eq:ltag:wlst-constituent-subtree-count} \\
+			&\text{with }\mathit{complete}(S) \Leftrightarrow \forall\ \text{leaf nodes $v_{\text{leaf}}$ of the tree $S$}: {|\Gamma_c(v_{\text{leaf}})|} = {|\Gamma_G(v_{\text{leaf}})|}
+			\text{.} \nonumber
+		\end{align}
 		The purpose of the $\mathit{complete}(S)$ condition is to only count the \ac{bfs} subtrees of the constituent $c$ that are also \ac{bfs} subtrees of the complete graph $G$.
@@
 
-In order to interpret a stack of convolutional layers as a representation of an \ac{lta} decomposition function $\psi: \mathcal{G} \to \mathcal{P}(\mathcal{G})$ and evaluation function $f: \mathcal{G} \to \mathcal{Y} \times \mathbb{R}_{\geq 0}$, there must be a mapping $\tau: \mathbb{R}^{* \times d^{(T)}} \to {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*}$ s.t.\ $\tau(Z) = {\ldblbrace f(c_i) \rdblbrace}_{c_i \in \psi(G)}$.
-Whether such a translation map $\tau$ from convolved feature matrices $Z$ to constituent evaluations exists, depends on the specific type of convolutional layer that is used.
+To define an \ac{lta} formulation for \acp{gcnn}, we define two functions that are entirely based on the stack of convolutional layers:
+\begin{enumerate*}
+	\item The decomposition function $\psi: \mathcal{G} \to \mathcal{P}(\mathcal{G})$.
+	\item The so-called \textit{multi-score evaluation function} $f^{*}: \mathcal{G} \to {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*}$ which is similar to a regular \ac{lta} evaluation function $f$ but assigns a multiset of weighted scores to each constituent instead of just a single one.
+\end{enumerate*}
+Those two functions make up the first half of a \ac{gcnn} graph formulation.
+How exactly they are defined depends on the type convolutional layer that is used.
 
@@
 \begin{align}\label{eq:ltag:general-spectral-conv}
-	z_i = {\left( U^{\top} \hat{g}(\Lambda) U X \right)}_i
-	= \sum_{j=1}^n \underbrace{\left( \sum_{l=1}^n \hat{g}(\lambda_l) u_{l,i} u_{l,j} \right)}_{\alpha_j} x_j
+	z_{\textcolor{t_red}{i}} = {\left( U^{\top} \hat{g}(\Lambda) U X \right)}_{\textcolor{t_red}{i}} % chktex 25
+	= \sum_{l=1}^n u_{l,\textcolor{t_red}{i}} \hat{g}(\lambda_l) \underbrace{\sum_{j=1}^n u_{l,j} x_j}_{\mathclap{\text{$d^{(T)}$-dim.\ vector of inner products}}} % chktex 25
+	= \sum_{j=1}^n \underbrace{\left( \sum_{l=1}^n \hat{g}(\lambda_l) u_{l,\textcolor{t_red}{i}} u_{l,j} \right)}_{\alpha_j} x_j % chktex 25
 \end{align}
-where $L = U^{\top} \Lambda U$ is the eigendecomposition of the input graph's Laplacian and $\hat{g}$ is some eigenvalue filter.
+where $L = U^{\top} \Lambda U = \sum_{l=1}^n u_l \lambda_l u_l^{\top}$ is the eigendecomposition of the input graph's Laplacian and $\hat{g}$ is some (learned) eigenvalue filter.
 \begin{lem}\label{lem:ltag:general-spectral-conv-nonlocality}
-	$\exists \eta \in \mathbb{R}$ s.t.\ if $\hat{g}(0) = \eta$, every $z_i$ depends on all input vertex features $x_j$.
+	For all graphs $G$ consisting of the connected components $C_1, \dots, C_m$ there is an $\eta \in \mathbb{R}$ s.t.\ if $\hat{g}(0) = \eta$, the convolved feature vector $z_i$ of each $v_i \in \mathcal{V}_{C_l}$ with $l \in [m]$ depends on all input features $x_j$ of the vertices $v_j \in \mathcal{V}_{C_l}$ from the same component.
 \end{lem}
-\begin{proof}
-	Note that by \cfullref[ on ]{defn:related:laplacian-comb} it follows that $u_1 = \frac{1}{\sqrt{n}}{(1, \dots, 1)}^{\top}$ is an eigenvector with eigenvalue $\lambda_1 = 0$ of the combinatorial Laplacian as well as the symmetric normalized Laplacian\footnote{
-		This is illustrated in \cfullref[ on ]{fig:related:graph-fourier}. We refer to \citet[section II.C.]{Shuman2013} for a more in-depth discussion of this fact.
+\begin{hproof}
+	Note that by \cfullref[ on ]{defn:related:laplacian-comb} it follows that the first $m$ eigenvalues of a graph with the connected components $C_1, \dots, C_m$ are $\lambda_1 = \cdots = \lambda_m = 0$ with the corresponding nullspace-spanning unnormalized eigenvectors $u_l[v_j] = \mathbbm{1}[v_j \in \mathcal{V}_{C_l}]$\daggerfootnote{
+		We refer to \citet{Das2004} for a more in-depth discussion of graph Laplacian eigenvectors.
 	}.
-	Consequently in \cref{eq:ltag:general-spectral-conv} the first summand of each $\alpha_j$ equals $\hat{g}(0) u_{1,j} u_{1,i} = \frac{\eta}{n}$ for all $i \in [n]$.
-	The lemma then follows by choosing $\eta$ s.t.\ all $\alpha_j \neq 0$.
-\end{proof}
-Consequently because a \ac{gcnn} with an arbitrary learned spectral filter $\hat{g}$ can learn $\hat{g}(0) = \eta$, the convolved vertex feature vectors ${\{ z_i \}}_{v_i \in \mathcal{V}_G}$ potentially all depend on the complete set of input feature vectors ${\{ x_i \}}_{v_i \in \mathcal{V}_G}$.
-Therefore all constituent scores and weights computed by a mapping $\tau(Z)$ can also depend on all input feature vectors, i.e.\ the corresponding constituents must span the entire input graph.
-In other words, spectral \acp{gcnn} with arbitrary filters do not generally have nontrivial \ac{lta} formulations which makes them non-\acs{lta} methods.
+	Consequently in \cref{eq:ltag:general-spectral-conv} the first $m$ summands of each $\alpha_j$ sum up to $\hat{g}(0) u_{l,j} u_{l,i} = \eta$ for all $v_i \in C_l$ with $l \in [m]$.
+	The lemma then follows by choosing $\eta$ s.t.\ the last $n - m$ summands of every $\alpha_j$ do not sum up to $\eta$.
+\end{hproof}
+Because a \ac{gcnn} with an arbitrary learned spectral filter $\hat{g}$ can learn $\hat{g}(0) = \eta$, each convolved vertex feature vector $z_i$ potentially depends on the entire connected component $C_l$ with $v_i \in \mathcal{V}_{C_l}$.
+This means that any constituent scores computed based on $z_i$ can only be computed for constituents that span entire connected components.
+In other words, spectral \acp{gcnn} with arbitrary filters generally only have \ac{lta} formulations with the decomposition function $\psi(G) = \{ C_1, \dots, C_m \}$.
+This restriction makes them an in-between case which is only partially \acs{lta}-like.
 
@@
 
-Both, the single vertex neighborhood convolution as well as the vertex $k$-multiset neighborhood convolution produce vectors $z_i \in \mathbb{R}^{d^{(T)}}$, each of which is based on a local constituent $c_i$ that might not be unique, i.e.\ depending on the input graph there can be $j \neq i$ with $c_j = c_i$.
+\paragraph{\ac{gcnn} \ac{lta} decomposition}
+Both, the single vertex neighborhood convolution as well as the vertex $k$-multiset neighborhood convolution produce vectors $z_i \in \mathbb{R}^{d^{(T)}}$, which are each based on the constituent $c_i$ spanned by the $T$-neighborhood of some vertex $v_i$ or vertex $k$-multiset $s_i$.
+Consequently we define the \ac{lta} decomposition function of such neighborhood convolution \acp{gcnn} as $\psi(G) \coloneqq {\{ c_i \}}_{i=1}^n$.
+
+\paragraph{\ac{gcnn} \ac{lta} multi-score evaluation}
+In addition to this decomposition function, the convolved vectors $z_i \in \mathbb{R}^{d^{(T)}}$ can be interpreted as local evaluations if there is a translation function $\tau: \mathbb{R}^{d^{(T)}} \to \mathcal{Y} \times \mathbb{R}_{\geq 0}$.
+One trivial example for such a translation function would be the identity $\tau(z) = z$ for a graph regression problem with the score domain $\mathcal{Y} = [0, 1]$ and where the last convolutional layer uses a sigmoid activation function with two output dimensions s.t.\ $z \in {[0,1]}^2$.
+Assuming that there is a translation function $\tau$ we can define a multi-score evaluation function $f^{*}(c) \coloneqq {\ldblbrace \tau(z_i)\, |\, {i \in [n]} \land {c = c_i} \rdblbrace} = {\ldblbrace \tau(\text{Convolution}(X_c)[v_i])\, |\, v_i \in \mathit{root}(c) \rdblbrace}$ with $X_c$ being the input feature matrix of the subgraph/constituent $c$.
+The function $\mathit{root}(c) = \{ v_i\, |\, {i \in [n]} \land {c = c_i} \}$ returns the set of root vertices $v_i$ (or root vertex $k$-multisets in case of $k$-\acs{gnn}) whose \ac{bfs} subtree of depth $T$ in $G$ spans exactly the vertices in $c$.
+To compute $\mathit{root}(c)$ given only a single constituent $c$ and not the entire graph $G$, it has to be determined which \ac{bfs} subtrees of $c$ are also \ac{bfs} subtrees of $G$.
+We already saw how to do this in the \ac{lta} formulation of \ac{wl} subtree kernel \acp{svm} (see $\mathit{complete}$ in \cfullref{eq:ltag:wlst-constituent-subtree-count}).
 
@@
 
+Based on the decomposition function $\psi: \mathcal{G} \to \mathcal{P}(\mathcal{G})$ and the multi-score evaluation function $f^{*}: \mathcal{G} \to {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*}$ which are implicitly defined by the stack of convolutional layers of a \ac{gcnn}, we will now define a proper \ac{lta} constituent evaluation function $f: \mathcal{G} \to \mathcal{Y} \times \mathbb{R}_{\geq 0}$ as well as an aggregation function $\mathcal{A}: {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*} \to \mathcal{Y}$.
+
 \section{A Novel \acs*{lta}-Inspired \acs*{gcnn} Architecture}%

### 2020-03-23 implementation/src/ltag/evaluation/models.py

@@ -87,7 +87,7 @@ def AvgCWL2GCN_Binary(
     squeeze_output=[True]
   )
 
-@binary_classifier(gnn_models.with_fc(gnn_models.AvgWL2GCN))
+@binary_classifier(gnn_models.with_fc(gnn_models.AvgCWL2GCN))
 def AvgCWL2GCN_FC_Binary(
   dsm,
   cwl2_local_act="sigmoid",

### 2020-03-24 thesis/content/chapter-graph-lta.tex
continued with spatial gcnn lta relation

@@
 
-\subsection{\acsp*{svm} on Graph Embeddings as \acs*{lta} Models}%
+\subsection{\acsp*{svm} using Graph Embeddings as \acs*{lta} Models}%
 \label{sec:ltag:formulation:svm}
@@
 	\item \textbf{Convolution:}
-		First a stack of $T$ convolutional layers is applied to the input feature matrix $X \in \mathbb{R}^{n \times d^{(0)}}$ where each row is a feature vector $x_i \in \mathbb{R}^d$ for the vertex $v_i \in \mathcal{V}_G$ or the vertex $k$-multiset $s_i \subseteq \mathcal{V}_G$ in case of $k$-\acsp{gnn}.
-		The convolutional layers produce a convolved feature matrix $Z \in \mathbb{R}^{n \times d^{(T)}}$.
+		First a stack of $T$ convolutional layers is applied to the input feature matrix $X \in \mathbb{R}^{n \times d^{(0)}}$ where each row is a feature vector $x_i \in \mathbb{R}^{d^{(0)}}$ for the vertex $v_i \in \mathcal{V}_G$ or the vertex $k$-multiset $s_i \subseteq \mathcal{V}_G$ in case of $k$-\acsp{gnn}.
+		The convolutional layers produce a convolved feature matrix $Z \in \mathbb{R}^{n \times d^{(T)}}$ where each row is a convolved feature vector $z_i \in \mathbb{R}^{d^{(\text{pool})}}$.
 	\item \textbf{Pooling:}
-		Then the convolved feature matrix $Z$ is reduced to a single graph feature vector $z_G \in \mathbb{R}^{d^{(T)}}$ via a pooling layer.
+		Then the convolved feature matrix $Z$ is reduced to a single graph feature vector $z_G \in \mathbb{R}^{d^{(\text{pool})}}$ via a pooling layer.
 	\item \textbf{\ac{mlp}:}
@@
 \[
-	\text{\acs*{gcnn}}(G) \coloneqq \text{\acs*{mlp}}(\text{Pooling}(\text{Convolution}(G)))
+	\text{\acs*{gcnn}}(G) \coloneqq \mathit{\acs*{mlp}}(\mathit{Pool}(\mathit{Conv}(G)))
 \]
@@
 \paragraph{\ac{gcnn} \ac{lta} multi-score evaluation}
-In addition to this decomposition function, the convolved vectors $z_i \in \mathbb{R}^{d^{(T)}}$ can be interpreted as local evaluations if there is a translation function $\tau: \mathbb{R}^{d^{(T)}} \to \mathcal{Y} \times \mathbb{R}_{\geq 0}$.
-One trivial example for such a translation function would be the identity $\tau(z) = z$ for a graph regression problem with the score domain $\mathcal{Y} = [0, 1]$ and where the last convolutional layer uses a sigmoid activation function with two output dimensions s.t.\ $z \in {[0,1]}^2$.
-Assuming that there is a translation function $\tau$ we can define a multi-score evaluation function $f^{*}(c) \coloneqq {\ldblbrace \tau(z_i)\, |\, {i \in [n]} \land {c = c_i} \rdblbrace} = {\ldblbrace \tau(\text{Convolution}(X_c)[v_i])\, |\, v_i \in \mathit{root}(c) \rdblbrace}$ with $X_c$ being the input feature matrix of the subgraph/constituent $c$.
+In addition to this decomposition function, the convolved vectors $z_i \in \mathbb{R}^{d^{(T)}}$ can be interpreted as local evaluations if there is a translation function $\tau_{1}: \mathbb{R}^{d^{(T)}} \to \mathcal{Y} \times \mathbb{R}_{\geq 0}$.
+One trivial example for such a translation function would be the identity $\tau_{1}(z) = z$ for a graph regression problem with the score domain $\mathcal{Y} = [0, 1]$ and where the last convolutional layer uses a sigmoid activation function with two output dimensions s.t.\ $z \in {[0,1]}^2$.
+Assuming that there is a translation function $\tau_{1}$ we can define a multi-score evaluation function $f^{*}(c) \coloneqq {\ldblbrace \tau_{1}(z_i)\, |\, {i \in [n]} \land {c = c_i} \rdblbrace} = {\ldblbrace \tau_{1}(\mathit{Conv}(X_c)[v_i])\, |\, v_i \in \mathit{root}(c) \rdblbrace}$ with $X_c$ being the input feature matrix of the subgraph/constituent $c$.
 The function $\mathit{root}(c) = \{ v_i\, |\, {i \in [n]} \land {c = c_i} \}$ returns the set of root vertices $v_i$ (or root vertex $k$-multisets in case of $k$-\acs{gnn}) whose \ac{bfs} subtree of depth $T$ in $G$ spans exactly the vertices in $c$.
@@
 
-Based on the decomposition function $\psi: \mathcal{G} \to \mathcal{P}(\mathcal{G})$ and the multi-score evaluation function $f^{*}: \mathcal{G} \to {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*}$ which are implicitly defined by the stack of convolutional layers of a \ac{gcnn}, we will now define a proper \ac{lta} constituent evaluation function $f: \mathcal{G} \to \mathcal{Y} \times \mathbb{R}_{\geq 0}$ as well as an aggregation function $\mathcal{A}: {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*} \to \mathcal{Y}$.
+Based on the decomposition function $\psi: \mathcal{G} \to \mathcal{P}(\mathcal{G})$ and the multi-score evaluation function $f^{*}: \mathcal{G} \to {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*}$ which are implicitly defined by the stack of convolutional layers $\mathit{Conv}$ of a given \ac{gcnn}, we now complete our \ac{gcnn} \ac{lta} formulation by defining an evaluation function $f: \mathcal{G} \to \mathcal{Y} \times \mathbb{R}_{\geq 0}$ as well as an aggregation function $\mathcal{A}: {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*} \to \mathcal{Y}$ based on the pooling layer $\mathit{Pool}: \mathbb{R}^{n \times d^{(T)}} \to \mathbb{R}^{d^{(\text{pool})}}$.
+\begin{defn}\label{defn:ltag:assoc-algebra}
+	A variadic function $\xi: A^{*} \to B$ is called \textit{variadic associative} iff.\ there is an operator $\rho: A^{*} \to A$ s.t.\ $\forall n \in \mathbb{N}: \forall (a_1, \dots, a_n) \in A^n: \forall {1 \leq i \leq j \leq n}: \xi(a_1, \dots, a_n) = \xi(a_1, \dots, a_{i-1}, \rho(a_i, \dots, a_j), a_{j+1}, \dots, a_n)$.
+	The heterogeneous algebra $((A, B), (\xi, \rho))$ is called an \textit{associative algebra} of $\xi$~\cite{Birkhoff1970}.
+\end{defn}
+\begin{defn}[see \citet{Novotny2002}]\label{defn:ltag:assoc-algebra-homomorphism}
+	A function tuple $(\tau_1: A \to C, \tau_2: B \to D)$ is a \textit{homomorphism} between two associative algebras $((A, B), (\xi, \rho))$ and $((C, D), (\xi', \rho'))$
+	\begin{align*}
+		\text{iff.\ }
+		\forall n \in \mathbb{N}: \forall (a_1, \dots, a_n) \in A^n:
+		\, &\tau_2(\xi(a_1, \dots, a_n)) = \xi'(\tau_1(a_1), \dots, \tau_1(a_n)) \\
+		\land\, &\tau_1(\rho(a_1, \dots, a_n)) = \rho'(\tau_1(a_1), \dots, \tau_1(a_n)) % chktex 21
+		\text{.}
+	\end{align*}
+\end{defn}
+\begin{defn}\label{defn:ltag:assoc-aggregation-pooling}
+	A pooling layer $\mathit{Pool}: \mathbb{R}^{n \times d^{(T)}} \to \mathbb{R}^{d^{(\text{pool})}}$ is called a \textit{weighted associative aggregation layer} iff.
+	\begin{enumerate*}
+		\item there is an associative algebra of $\mathit{Pool}$,
+		\item there is a weighted aggregation function $\mathcal{A}$ with an associative algebra and
+		\item there is a homomorphism $(\tau_{1}: \mathbb{R}^{d^{(T)}} \to (\mathcal{Y} \times \mathbb{R}_{\geq 0}), \tau_{2}: \mathbb{R}^{d^{(\text{pool})}} \to \mathcal{Y})$ from the associative algebra of $\mathit{Pool}$ to that of $\mathcal{A}$
+	\end{enumerate*}.
+\end{defn}
+Using those definitions we can complete the connection between \acp{gcnn} and \ac{lta}:
+\begin{thm}\label{thm:ltag:gcnn-ltag-formulation}
+	A \ac{gcnn} of the form $\mathit{Pool}(\mathit{Conv}(X_G))$ has a nontrivial \ac{lta} formulation if $\mathit{Conv}$ is a stack of neighborhood convolutions and $\mathit{Pool}$ is a weighted associative aggregation layer.
+\end{thm}
+\begin{proof}
+	We already saw that the stack of neighborhood convolutions computes the values of a multi-score evaluation function $f^{*}$ for the subtree constituents returned by a decomposition function $\psi$.
+	Since $\mathit{Pool}$ is assumed to be a weighted associative aggregation layer, there must be an operator $\rho_{\mathit{Pool}}: \mathbb{R}^{n \times d^{(T)}} \to \mathbb{R}^{n \times d^{(T)}}$ and a homomorphism $(\tau_{1}, \tau_{2})$ to an associative algebra with the operators $\mathcal{A}: {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*} \to \mathcal{Y}$ and $\rho_{\mathcal{A}}: {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*} \to {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}$.
+	This allows us to rewrite the \ac{gcnn} as
+	\begin{align*}
+		\tau_2(\mathit{Pool}(\mathit{Conv}(X_G)))
+		&= \tau_2(\mathit{Pool}(\ldblbrace z_i\, |\, z_i = \mathit{Conv}(X_G)[v_i] \land v_i \in \mathcal{V}_G \rdblbrace)) \\
+		&= \tau_2(\mathit{Pool}(\ldblbrace \mathit{Conv}(X_c)[v_i]\, |\, c \in \psi(G) \land v_i \in \mathit{root}(c) \rdblbrace)) \\
+		&= \tau_2(\mathit{Pool}(\ldblbrace \rho_{\mathit{Pool}}(\ldblbrace \mathit{Conv}(X_c)[v_i]\, |\, v_i \in \mathit{root}(c) \rdblbrace)\, |\, c \in \psi(G) \rdblbrace)) \\
+		&= \mathcal{A}(\ldblbrace \rho_{\mathcal{A}}(\ldblbrace \tau_1(\mathit{Conv}(X_c)[v_i])\, |\, v_i \in \mathit{root}(c) \rdblbrace)\, |\, c \in \psi(G) \rdblbrace) \\
+		&= \mathcal{A}(\ldblbrace \underbrace{\rho_{\mathcal{A}}(f^{*}(c))}_{f(c)}\, |\, c \in \psi(G) \rdblbrace)
+		= \mathcal{A}(\ldblbrace f(c)\, |\, c \in \psi(G) \rdblbrace)
+		\text{.}
+	\end{align*}
+	The \ac{lta} formulation above is for vertex neighborhood convolutions;
+	a formulation for vertex $k$-multiset neighborhood convolutions can be obtained analogously which concludes the proof\footnote{
+		Note that the \ac{lta} formulation in this proof assumes that the \ac{gcnn}'s output is translated into the target score space via $\tau_{2}: \mathbb{R}^{d} \to \mathcal{Y}$, e.g.\ if the \ac{gcnn} returns one-hot vector encodings of discrete classes $y \in \mathcal{Y}$ those must be decoded.
+	}.
+\end{proof}
+Note that \cref{thm:ltag:gcnn-ltag-formulation} requires \acp{gcnn} without the final \ac{mlp} proposed in the \ac{gcn}~\cite{Kipf2017} and \gc{gin}~\cite{Xu2018} papers (see \cref{fig:ltag:gcnn-structure}).
+The reason for this is that the composition $\mathit{\acs*{mlp}} \circ \mathit{Pool}$ cannot be guaranteed to be homomorphic to a weighted aggregation function $\mathcal{A}$ due to the universal approximation theorem.
+
+Lastly we will see which graph pooling layers satisfy \cref{defn:ltag:assoc-aggregation-pooling} and are therefore allowed in \ac{lta} \acp{gcnn}.
 
@@
 \label{sec:ltag:wl2gnn}
+
+\subsection{Shortcomings of $k$-\acsp*{gnn}}%
+\label{sec:ltag:wl2gnn:kgnn-problems}
+
+\subsection{The 2-\acs*{wl} Convolution Operator}%
+\label{sec:ltag:wl2gnn:definition}
+
+\subsection{The Expressive Power of 2-\acs*{wl}-\acsp*{gcnn}}%
+\label{sec:ltag:wl2gnn:properties}
+
+\subsection{Efficient Implementation on \acsp*{gpgpu}}%
+\label{sec:ltag:wl2gnn:implementation}

### 2020-03-25 thesis/content/chapter-graph-lta.tex
complted gcnn lta relation section

--- 2020-03-23 thesis/content/chapter-graph-lta.tex
@@
 similarly \acp{gcnn} also have \ac{lta} formulations under certain conditions.
-In this section we will describe those conditions and the \ac{lta} formulations of the \acp{gcnn} that satisfy them.
+In this section those conditions and the \ac{lta} formulations of the \acp{gcnn} that satisfy them are described.
 
-In \cref{sec:related:gcr:nn} an overview of graph convolution and graph pooling layers was given.
-While there are many ways to combine those layers, we focus on global pooling \ac{gcnn} architectures that use the following three steps:
+Looking back, an overview of graph convolution and graph pooling layers was given in \cref{sec:related:gcr:nn}.
+There are many ways to combine those layers in a concrete \ac{gnn} but we focus on \textit{global pooling \ac{gcnn} architectures} that use the following three steps:
 \begin{enumerate}[label=\textbf{\arabic*.}]
 	\item \textbf{Convolution:}
-		First a stack of $T$ convolutional layers is applied to the input feature matrix $X \in \mathbb{R}^{n \times d^{(0)}}$ where each row is a feature vector $x_i \in \mathbb{R}^d$ for the vertex $v_i \in \mathcal{V}_G$ or the vertex $k$-multiset $s_i \subseteq \mathcal{V}_G$ in case of $k$-\acsp{gnn}.
-		The convolutional layers produce a convolved feature matrix $Z \in \mathbb{R}^{n \times d^{(T)}}$.
+		First a stack of $T$ convolution layers is applied to the input feature matrix $X \in \mathbb{R}^{n \times d^{(0)}}$ where each row is a feature vector $x_i \in \mathbb{R}^{d^{(0)}}$ for the vertex $v_i \in \mathcal{V}_G$ or the vertex $k$-multiset $s_i \subseteq \mathcal{V}_G$ in case of $k$-\acsp{gnn}.
+		The convolution layers produce a convolved feature matrix $Z \in \mathbb{R}^{n \times d^{(T)}}$ where each row is a convolved feature vector $z_i \in \mathbb{R}^{d^{(\text{pool})}}$.
 	\item \textbf{Pooling:}
-		Then the convolved feature matrix $Z$ is reduced to a single graph feature vector $z_G \in \mathbb{R}^{d^{(T)}}$ via a pooling layer.
+		Then the convolved feature matrix $Z$ is reduced to a single graph feature vector $z_G \in \mathbb{R}^{d^{(\text{pool})}}$ via a pooling layer.
 	\item \textbf{\ac{mlp}:}
@@
 \[
-	\text{\acs*{gcnn}}(G) \coloneqq \text{\acs*{mlp}}(\text{Pooling}(\text{Convolution}(G)))
+	\text{\acs*{gcnn}}(G) \coloneqq \mathit{\acs*{mlp}}(\mathit{Pool}(\mathit{Conv}(G)))
 \]
 \Cref{fig:ltag:gcnn-structure} illustrates this three step architecture for vertex neighborhood convolutions like those used by \acp{gcn}~\cite{Kipf2017} or \acp{gin}~\cite{Xu2018}.
-Intuitively the convolutional layers implicitly define an \ac{lta} decomposition function $\psi$ and a local evaluation function $f$ while the pooling layer corresponds to an \ac{lta} aggregation function $\mathcal{A}$.
+Intuitively the convolution layers implicitly define an \ac{lta} decomposition function $\psi$ and a local evaluation function $f$ while the pooling layer roughly corresponds to an \ac{lta} aggregation function $\mathcal{A}$.
 This intuition is now formalized.
@@
 
-To define an \ac{lta} formulation for \acp{gcnn}, we define two functions that are entirely based on the stack of convolutional layers:
+To define an \ac{lta} formulation for \acp{gcnn}, we define two functions that are entirely based on the stack of convolution layers ($\mathit{Conv}$):
 \begin{enumerate*}
@@
 Those two functions make up the first half of a \ac{gcnn} graph formulation.
-How exactly they are defined depends on the type convolutional layer that is used.
+How exactly they are defined depends on the type convolution layer that is used.
 
 \paragraph{Spectral convolutions with arbitrary filters}
-Let us begin with general spectral graph convolutions~\cite{Bruna2013}\cite{Henaff2015}.
+Let us begin with general spectral graph convolution layers~\cite{Bruna2013}\cite{Henaff2015}.
 Following \cfullref[ on ]{eq:related:graph-conv-general}, after a single convolution the convolved feature vector $z_i$ of a vertex $v_i$ is defined as
@@
 \paragraph{\ac{gcnn} \ac{lta} multi-score evaluation}
-In addition to this decomposition function, the convolved vectors $z_i \in \mathbb{R}^{d^{(T)}}$ can be interpreted as local evaluations if there is a translation function $\tau: \mathbb{R}^{d^{(T)}} \to \mathcal{Y} \times \mathbb{R}_{\geq 0}$.
-One trivial example for such a translation function would be the identity $\tau(z) = z$ for a graph regression problem with the score domain $\mathcal{Y} = [0, 1]$ and where the last convolutional layer uses a sigmoid activation function with two output dimensions s.t.\ $z \in {[0,1]}^2$.
-Assuming that there is a translation function $\tau$ we can define a multi-score evaluation function $f^{*}(c) \coloneqq {\ldblbrace \tau(z_i)\, |\, {i \in [n]} \land {c = c_i} \rdblbrace} = {\ldblbrace \tau(\text{Convolution}(X_c)[v_i])\, |\, v_i \in \mathit{root}(c) \rdblbrace}$ with $X_c$ being the input feature matrix of the subgraph/constituent $c$.
+In addition to this decomposition function, $\mathit{Conv}$ also implicitly computes local constituent evaluations.
+If there is a translation function $\tau_{1}: \mathbb{R}^{d^{(T)}} \to \mathcal{Y} \times \mathbb{R}_{\geq 0}$, the convolved vectors $z_i \in \mathbb{R}^{d^{(T)}}$ can be interpreted as constituent evaluations.
+One trivial example for such a translation function would be the identity $\tau_{1}(z) = z$ for a graph regression problem with the score domain $\mathcal{Y} = [0, 1]$ and where the last convolution layer uses a sigmoid activation function with two output dimensions s.t.\ $z \in {[0,1]}^2$.
+Assuming that there is a translation function $\tau_{1}$ we can define a multi-score evaluation function $f^{*}(c) \coloneqq {\ldblbrace \tau_{1}(z_i)\, |\, {i \in [n]} \land {c = c_i} \rdblbrace} = {\ldblbrace \tau_{1}(\mathit{Conv}(c)[v_i])\, |\, v_i \in \mathit{root}(c) \rdblbrace}$.
 The function $\mathit{root}(c) = \{ v_i\, |\, {i \in [n]} \land {c = c_i} \}$ returns the set of root vertices $v_i$ (or root vertex $k$-multisets in case of $k$-\acs{gnn}) whose \ac{bfs} subtree of depth $T$ in $G$ spans exactly the vertices in $c$.
@@
 
-Based on the decomposition function $\psi: \mathcal{G} \to \mathcal{P}(\mathcal{G})$ and the multi-score evaluation function $f^{*}: \mathcal{G} \to {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*}$ which are implicitly defined by the stack of convolutional layers of a \ac{gcnn}, we will now define a proper \ac{lta} constituent evaluation function $f: \mathcal{G} \to \mathcal{Y} \times \mathbb{R}_{\geq 0}$ as well as an aggregation function $\mathcal{A}: {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*} \to \mathcal{Y}$.
+Based on the decomposition function $\psi: \mathcal{G} \to \mathcal{P}(\mathcal{G})$ and the multi-score evaluation function $f^{*}: \mathcal{G} \to {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*}$ which are implicitly defined by the stack of convolution layers $\mathit{Conv}$ of a given \ac{gcnn}, we now complete our \ac{gcnn} \ac{lta} formulation by defining an evaluation function $f: \mathcal{G} \to \mathcal{Y} \times \mathbb{R}_{\geq 0}$ as well as an aggregation function $\mathcal{A}: {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*} \to \mathcal{Y}$ based on the pooling layer $\mathit{Pool}: \mathbb{R}^{n \times d^{(T)}} \to \mathbb{R}^{d^{(\text{pool})}}$.
+\begin{defn}\label{defn:ltag:assoc-algebra}
+	A variadic function $\xi: A^{*} \to B$ is called \textit{associative} iff.\ there is an operator $\rho: A^{*} \to A$ s.t.\ $\forall n \in \mathbb{N}: \forall (a_1, \dots, a_n) \in A^n: \forall {1 \leq i \leq j \leq n}: \xi(a_1, \dots, a_n) = \xi(a_1, \dots, a_{i-1}, \rho(a_i, \dots, a_j), a_{j+1}, \dots, a_n)$.
+	The heterogeneous algebra\footnote{
+		See \citet{Birkhoff1970} for a more detailed description of heterogeneous algebras.
+	} $((A, B), (\xi, \rho))$ is called an \textit{associative algebra} of $\xi$.
+\end{defn}
+\begin{defn}[see \citet{Novotny2002}]\label{defn:ltag:assoc-algebra-homomorphism}
+	A function tuple $(\tau_1: A \to C, \tau_2: B \to D)$ is a \textit{homomorphism} between two associative algebras $((A, B), (\xi, \rho))$ and $((C, D), (\xi', \rho'))$
+	\begin{align*}
+		\text{iff.\ }
+		\forall n \in \mathbb{N}: \forall (a_1, \dots, a_n) \in A^n:
+		\, &\tau_2(\xi(a_1, \dots, a_n)) = \xi'(\tau_1(a_1), \dots, \tau_1(a_n)) \\
+		\land\, &\tau_1(\rho(a_1, \dots, a_n)) = \rho'(\tau_1(a_1), \dots, \tau_1(a_n)) % chktex 21
+		\text{.}
+	\end{align*}
+\end{defn}
+\begin{defn}\label{defn:ltag:assoc-aggregation-pooling}
+	A pooling layer $\mathit{Pool}: \mathbb{R}^{n \times d^{(T)}} \to \mathbb{R}^{d^{(\text{pool})}}$ is called an \textit{associative weighted aggregation layer} iff.
+	\begin{enumerate*}
+		\item there is an associative algebra of $\mathit{Pool}$,
+		\item there is a weighted aggregation function $\mathcal{A}: {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*} \to \mathcal{Y}$ with an associative algebra and
+		\item there is a homomorphism $(\tau_{1}: \mathbb{R}^{d^{(T)}} \to (\mathcal{Y} \times \mathbb{R}_{\geq 0}), \tau_{2}: \mathbb{R}^{d^{(\text{pool})}} \to \mathcal{Y})$ from the associative algebra of $\mathit{Pool}$ to that of $\mathcal{A}$
+	\end{enumerate*}.
+\end{defn}
+Using those definitions we can complete the connection between \acp{gcnn} and \ac{lta}:
+\begin{thm}\label{thm:ltag:gcnn-ltag-formulation}
+	A \ac{gcnn} of the form $\mathit{Pool}(\mathit{Conv}(G))$ has a nontrivial \ac{lta} formulation if $\mathit{Conv}$ is a stack of neighborhood convolution layers and $\mathit{Pool}$ is an associative weighted aggregation layer.
+\end{thm}
+\begin{proof}
+	We already saw that a stack of neighborhood convolutions computes the values of a multi-score evaluation function $f^{*}$ for the subtree constituents returned by a decomposition function $\psi$.
+	Since $\mathit{Pool}$ is assumed to be an associative weighted aggregation layer, there must be an operator $\rho_{\mathit{Pool}}: \mathbb{R}^{n \times d^{(T)}} \to \mathbb{R}^{n \times d^{(T)}}$ and a homomorphism $(\tau_{1}, \tau_{2})$ to an associative algebra with the operators $\mathcal{A}: {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*} \to \mathcal{Y}$ and $\rho_{\mathcal{A}}: {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*} \to {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}$.
+	This allows us to rewrite the \ac{gcnn} as
+	\begin{align*}
+		\tau_2(\mathit{Pool}(\mathit{Conv}(G)))
+		&= \tau_2(\mathit{Pool}(\ldblbrace z_i\, |\, z_i = \mathit{Conv}(G)[v_i] \land v_i \in \mathcal{V}_G \rdblbrace)) \\
+		&= \tau_2(\mathit{Pool}(\ldblbrace \mathit{Conv}(c)[v_i]\, |\, c \in \psi(G) \land v_i \in \mathit{root}(c) \rdblbrace)) \\
+		&= \tau_2(\mathit{Pool}(\ldblbrace \rho_{\mathit{Pool}}(\ldblbrace \mathit{Conv}(c)[v_i]\, |\, v_i \in \mathit{root}(c) \rdblbrace)\, |\, c \in \psi(G) \rdblbrace)) \\
+		&= \mathcal{A}(\ldblbrace \rho_{\mathcal{A}}(\ldblbrace \tau_1(\mathit{Conv}(c)[v_i])\, |\, v_i \in \mathit{root}(c) \rdblbrace)\, |\, c \in \psi(G) \rdblbrace) \\
+		&= \mathcal{A}(\ldblbrace \underbrace{\rho_{\mathcal{A}}(f^{*}(c))}_{f(c)}\, |\, c \in \psi(G) \rdblbrace)
+		= \mathcal{A}(\ldblbrace f(c)\, |\, c \in \psi(G) \rdblbrace)
+		\text{.}
+	\end{align*}
+	The \ac{lta} formulation above is for vertex neighborhood convolutions;
+	a formulation for vertex $k$-multiset neighborhood convolutions can be obtained analogously which concludes the proof\footnote{
+		The \ac{lta} formulation in this proof assumes that the \ac{gcnn}'s output is translated into the target score space via $\tau_{2}: \mathbb{R}^{d} \to \mathcal{Y}$.
+		All this means is that an \acs{lta}-like \ac{gcnn} must produce outputs $y \in \mathcal{Y}$;
+		for example if a \ac{gcnn} returns one-hot vector encodings of discrete classes those must be decoded.
+	}.
+\end{proof}
+Note that \cref{thm:ltag:gcnn-ltag-formulation} requires \acp{gcnn} without a final \ac{mlp} which is unlike the architecture proposed in the \ac{gcn}~\cite{Kipf2017} and \ac{gin}~\cite{Xu2018} papers.
+The reason for this is that the composition $\mathit{\acs*{mlp}} \circ \mathit{Pool}$ cannot be guaranteed to be homomorphic to a weighted aggregation function $\mathcal{A}$ due to the universal approximation theorem.
+
+To conclude our analysis of the relation between \ac{lta} and \acp{gcnn}, we check which of the graph pooling layers described in \cref{sec:related:gcr:nn} satisfy \cref{defn:ltag:assoc-aggregation-pooling} are therefore allowed in \acs{lta}-like \acp{gcnn} by \cref{thm:ltag:gcnn-ltag-formulation}.
+For simplicity we assume that the target score domain is $\mathcal{Y} = [0, 1]$.
+\begin{enumerate}[label={\textbf{\arabic*.}}]
+	\item \textbf{Unweighted mean, minimum and maximum pooling:}
+		If the last convolution layer has a single output dimension and an activation function with the signature $\sigma: \mathbb{R} \to [0,1]$ (e.g.\ the logistic function), the simple unweighted pooling layers $\mean$, $\min$ and $\max$ trivially satisfy \cref{defn:ltag:assoc-aggregation-pooling} due to their associativity.
+	\item \textbf{SortPooling \& \ac{sagpool}:}
+		Apart from such unweighted pooling layers we also looked at two state-of-the-art graph pooling layers:
+		\begin{enumerate*}[label={\arabic*.}]
+			\item SortPooling~\cite{Zhang2018} works by aggregating the top-$p$ convolved vectors $z_i \in \mathbb{R}^{d^{(T)}}$ based on a lexicographic ordering,
+			\item \ac{sagpool}~\cite{Lee2019} also aggregates the top-$p$ vectors $z_i$ but determines their order based on attention weights which are encoded as one component of each $z_i$
+		\end{enumerate*}.
+		Both of those pooling approaches are not compatible with the \ac{lta} formulation from \cref{thm:ltag:gcnn-ltag-formulation} because top-$p$ aggregation is generally not associative.
+		This is shown by the fact that there is no $\rho: \mathbb{R}^{* \times d^{(T)}} \to \mathbb{R}^{ d^{(T)}}$ s.t.\ for all vectors $z_1 < z_2$ and and aggregation functions $\mathcal{A}$ it holds that%
+		\begin{alignat*}{2}
+			&& \mathcal{A}(\mathit{top}_{2}(\ldblbrace z_1, z_2, z_2 \rdblbrace)) &= \mathcal{A}(\mathit{top}_{2}(\ldblbrace z_1, \rho(\ldblbrace z_2, z_2 \rdblbrace) \rdblbrace)) \\
+			\Leftrightarrow&& \mathcal{A}(\ldblbrace z_2, z_2 \rdblbrace) &= \mathcal{A}(\ldblbrace z_1, \rho(\ldblbrace z_2, z_2 \rdblbrace) \rdblbrace)\quad\lightning
+		\end{alignat*}
+		From the \ac{lta} perspective the non-associativity of SortPooling and \ac{sagpool} implies that, if they were used in a \ac{gcnn}, it could happen that some subtree evaluations $(y_i, w_i) \in f^{*}(c)$ of a single constituent $c$ are considered by the aggregation function $\mathcal{A}$ while some are ignored depending on the subtree evaluations of other constituents, i.e.\ the constituent evaluation $f(c)$ could then depend on other constituents.
+		This is forbidden in our definition of \ac{lta}.
+\end{enumerate}
+
+Since neither SortPooling nor \ac{sagpool} are compatible with \ac{lta}, we propose an associative weighted aggregation layer based on \ac{sagpool} which uses a softmax-based attention filter instead of a top-$p$ filter.
+\begin{defn}
+	We define \ac{sampool} as
+	\[\mathit{SAM}({\ldblbrace (y_i, w_i) \rdblbrace}_{i=1}^n) \coloneqq\frac{1}{\sum_{i=1}^n e^{w_i}} \sum_{i=1}^n e^{w_i} y_i \text{.}\]
+\end{defn}
+If the stack of convolution layers of a \ac{gcnn} produces vectors $z_i = (y_i, w_i) \in [0, 1] \times \mathbb{R}$, \ac{sampool} satisfies \cref{defn:ltag:assoc-aggregation-pooling} and is therefore compatible with \acs{lta}.
+To see why this is the case, the associativity and homomorphicity to a weighted aggregation function $\mathcal{A}_{\text{SAM}}$ must be shown.
+The associativity of $\mathit{SAM}$ follows directly from its commutativity and the fact that
+\begin{gather*}
+	\mathit{SAM}({\ldblbrace (y_i, w_i) \rdblbrace}_{i=1}^n) = \frac{1}{e^{w'} + \sum_{i=j+1}^n e^{w_i}} \left( e^{w'} \underbrace{\mathit{SAM}({\ldblbrace (y_i, w_i) \rdblbrace}_{i=1}^j)}_{y'} + \smashoperator{\sum_{i=j+1}^n} e^{w_i} y_i \right) \\
+	\text{for all } j \in [n] \text{ with } w' \coloneqq \ln\sum_{i=1}^{j} e^{w_i} \text{, i.e.\ } \rho_{\mathit{SAM}}({\ldblbrace (y_i, w_i) \rdblbrace}_{i=1}^j) = (y', w')\text{.}
+\end{gather*}
+The homomorphicity of $\mathit{SAM}$ to a weighted aggregation function $\mathcal{A}_{\text{SAM}}$ follows by choosing $\mathcal{A}_{\text{SAM}}({\ldblbrace (y_i, w_i) \rdblbrace}_{i=1}^n) \coloneqq {\left(\sum_{i=1}^n w_i\right)}^{-1} \sum_{i=1}^n w_i y_i$ and the trivial homomorphism $\tau_1(y, w) \coloneqq (y, e^w)$, $\tau_2(y) \coloneqq y$ between the algebras of $\mathit{SAM}$ and $\mathcal{A}_{\text{SAM}}$.
+
+This concludes our analysis of the relation between \ac{lta} and \acp{gcnn}.
+We saw that spectral convolutions with arbitrary filters are only partially \acs{lta}-like since their constituents generally span entire connected components.
+\Acp{gcnn} using vertex neighborhood convolutions on the other hand were shown in \cref{thm:ltag:gcnn-ltag-formulation} to have an \ac{lta} formulation with localized subtree constituents.
+The only requirement for this formulation to exist is that an associative pooling layer like $\mean$, $\min$, $\max$ or $\mathit{SAM}$ is used to combine the convolved feature vectors into the final output.
 

### 2020-03-25 thesis/content/chapter-related-work.tex
complted gcnn lta relation section

@@
 To see why this is the case, we rewrite \cref{eq:related:gcn-layer} and compare it with the 1-\acs{wl} color refinement strategy (see \cfullref{defn:related:wl1-refine}):
-\begin{align*}
+\begin{alignat*}{2}
 	\text{\ac*{gcn}:}\quad&& Z^{(t)}[v] &= \sigma\left(\left( {\mu(v)}^2 Z^{(t-1)}[v] + \smashoperator[lr]{\sum_{u \in \Gamma_G(v)}} \mu(v) \mu(u) Z^{(t-1)}[u] \right) \Theta^{(t)} \right) \\
 	\text{1-\acs{wl}:}\quad&& \chi_{G,1}^{(t)}(v) &= h\left( \chi_{G,1}^{(t-1)}(v), \ldblbrace \chi_{G,1}^{(t-1)}(u)\, |\, u \in \Gamma_{G}(v) \rdblbrace\right)
-\end{align*}
+\end{alignat*}
 Here $\mu(v) \coloneqq {\left(\left| \Gamma_G(v) \right| + 1\right)}^{-\frac{1}{2}}$ are normalization factors introduced by $\tilde{D}_G^{-\frac{1}{2}}$.
@@
 In this thesis we will focus on global pooling.
-The most simple global pooling approaches use static aggregation functions like component-wise $\min$, $\max$ or $\mean$ to produce a single graph feature vector $z_G \in \mathbb{R}^{d^{(L)}}$.
+The most simple global pooling approaches use static aggregation functions like component-wise $\min$, $\max$ or $\mean$ to produce a single graph feature vector $z_G \in \mathbb{R}^{d^{(T)}}$.
 This vector is then fed into a standard \ac{mlp} to produce the final prediction.
@@
 By imposing a component-wise lexicographic ordering on the set of color vectors, it can be reduced to a vertex permutation invariant top-$p$ list of vectors $(z_1, \dots, z_p)$ for some fixed $p \in \mathbb{N}$.
-The final graph feature vector then is the concatenation $z_G = \left( \bigoplus_{i=1}^{p} z_i \right) \in \mathbb{R}^{d^{(L)} p}$.
+The final graph feature vector then is the concatenation $z_G = \left( \bigoplus_{i=1}^{p} z_i \right) \in \mathbb{R}^{d^{(T)} p}$.
 SortPooling is based on the idea that the convolutional layers will learn to use the most-significant lexicographic vector components to represent that vector's importance.

### 2020-03-26 implementation/backup_evaluations.sh
added backup script for evaluation results

@@ -1,0 +1,16 @@
+#!/usr/bin/env bash
+
+cd "${BASH_SOURCE%/*}" || exit
+
+TO=${1:-"$HOME/ltag_backup"}
+
+TO_TF="$TO/tf/"
+TO_TORCH="$TO/torch/"
+
+echo "Backing up TF evaluation results to $TO_TF..."
+rsync -zarvm --info=progress2 --include "*/" --include "*.json" --include "*.txt" --exclude "*" "./evaluations/" "$TO_TF"
+
+echo "Backing up Torch evaluation results to $TO_TORCH..."
+rsync -zarvm --info=progress2 --include "*/" --include "*.json" --include "*.log" --exclude "*" "./libs/gnn-comparison/RESULTS/" "$TO_TORCH"
+
+echo "Done."

### 2020-03-26 thesis/content/chapter-graph-lta.tex
reorganized thesis: wl2gnn gets its own chapter
includes the intro for the wl2gcnn chapter
+ various improvements

@@
 	\item We begin with a formal definition of what actually constitutes an \ac{lta} method as opposed to non-\acs{lta} methods.
-	\item Using this definition, we will see that some of the previously described \ac{gcr} methods can be interpreted as \ac{lta} variants under certain conditions.
-	\item Finally a novel \acs{lta}-inspired \ac{gcnn} architecture will be described.
+	\item Using this definition, we will see that an \ac{svm} that uses one of the previously described graph embedding approaches can be interpreted as an \ac{lta} variant under certain conditions.
+	\item In the last step we look at the previously described \ac{gcnn} architectures and formalize their relation to \ac{lta}.
 \end{enumerate}
@@
 
-\section{\acs*{lta} Formulations of Existing \acs*{gcr} Methods}%
-\label{sec:ltag:formulation}
+\section{\acsp*{svm} with Graph Embeddings as \acs*{lta} Models}%
+\label{sec:ltag:formulation:svm}
 
 Based on the general definition of \ac{lta} from the last section, we will now see to what extent the \ac{gcr} methods described in \cref{sec:related:gcr} can be interpreted as \ac{lta} instances.
-First the relation between \acp{svm} on graph embeddings and \ac{lta} will be explored.
-Then we will provide an \ac{lta} perspective on \acp{gcnn}.
-
-\subsection{\acsp*{svm} using Graph Embeddings as \acs*{lta} Models}%
-\label{sec:ltag:formulation:svm}
-
+This section explores the relation between \acp{svm} that use graph embeddings and \ac{lta}.
 In \cref{sec:related:gcr:embed,sec:related:gcr:kernel} three different ways to map a given graph $G$ to a vector $\varphi(G) \in \mathbb{R}^d$ were described:
@@
 
-\subsection{\acsp*{gcnn} as \acs*{lta} Models}%
+\section{\acsp*{gcnn} as \acs*{lta} Models}%
 \label{sec:ltag:formulation:gcnn}
@@
 		First a stack of $T$ convolution layers is applied to the input feature matrix $X \in \mathbb{R}^{n \times d^{(0)}}$ where each row is a feature vector $x_i \in \mathbb{R}^{d^{(0)}}$ for the vertex $v_i \in \mathcal{V}_G$ or the vertex $k$-multiset $s_i \subseteq \mathcal{V}_G$ in case of $k$-\acsp{gnn}.
-		The convolution layers produce a convolved feature matrix $Z \in \mathbb{R}^{n \times d^{(T)}}$ where each row is a convolved feature vector $z_i \in \mathbb{R}^{d^{(\text{pool})}}$.
+		The convolution layers produce a convolved feature matrix $Z \in \mathbb{R}^{n \times d^{(T)}}$ where each row is a convolved feature vector $z_i \in \mathbb{R}^{d^{(\mathrm{pool})}}$.
 	\item \textbf{Pooling:}
-		Then the convolved feature matrix $Z$ is reduced to a single graph feature vector $z_G \in \mathbb{R}^{d^{(\text{pool})}}$ via a pooling layer.
+		Then the convolved feature matrix $Z$ is reduced to a single graph feature vector $z_G \in \mathbb{R}^{d^{(\mathrm{pool})}}$ via a pooling layer.
 	\item \textbf{\ac{mlp}:}
@@
 \[
-	\text{\acs*{gcnn}}(G) \coloneqq \mathit{\acs*{mlp}}(\mathit{Pool}(\mathit{Conv}(G)))
+	h_{\mathrm{\acs*{gcnn}}}(G) \coloneqq \mathrm{\acs*{mlp}}(\mathit{Pool}(\mathit{Conv}(G)))
 \]
@@
 
-Based on the decomposition function $\psi: \mathcal{G} \to \mathcal{P}(\mathcal{G})$ and the multi-score evaluation function $f^{*}: \mathcal{G} \to {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*}$ which are implicitly defined by the stack of convolution layers $\mathit{Conv}$ of a given \ac{gcnn}, we now complete our \ac{gcnn} \ac{lta} formulation by defining an evaluation function $f: \mathcal{G} \to \mathcal{Y} \times \mathbb{R}_{\geq 0}$ as well as an aggregation function $\mathcal{A}: {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*} \to \mathcal{Y}$ based on the pooling layer $\mathit{Pool}: \mathbb{R}^{n \times d^{(T)}} \to \mathbb{R}^{d^{(\text{pool})}}$.
+Based on the decomposition function $\psi: \mathcal{G} \to \mathcal{P}(\mathcal{G})$ and the multi-score evaluation function $f^{*}: \mathcal{G} \to {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*}$ which are implicitly defined by the stack of convolution layers $\mathit{Conv}$ of a given \ac{gcnn}, we now complete our \ac{gcnn} \ac{lta} formulation by defining an evaluation function $f: \mathcal{G} \to \mathcal{Y} \times \mathbb{R}_{\geq 0}$ as well as an aggregation function $\mathcal{A}: {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*} \to \mathcal{Y}$ based on the pooling layer $\mathit{Pool}: \mathbb{R}^{n \times d^{(T)}} \to \mathbb{R}^{d^{(\mathrm{pool})}}$.
 \begin{defn}\label{defn:ltag:assoc-algebra}
@@
 \begin{defn}\label{defn:ltag:assoc-aggregation-pooling}
-	A pooling layer $\mathit{Pool}: \mathbb{R}^{n \times d^{(T)}} \to \mathbb{R}^{d^{(\text{pool})}}$ is called an \textit{associative weighted aggregation layer} iff.
+	A pooling layer $\mathit{Pool}: \mathbb{R}^{n \times d^{(T)}} \to \mathbb{R}^{d^{(\mathrm{pool})}}$ is called an \textit{associative weighted aggregation layer} iff.
 	\begin{enumerate*}
@@
 		\item there is a weighted aggregation function $\mathcal{A}: {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*} \to \mathcal{Y}$ with an associative algebra and
-		\item there is a homomorphism $(\tau_{1}: \mathbb{R}^{d^{(T)}} \to (\mathcal{Y} \times \mathbb{R}_{\geq 0}), \tau_{2}: \mathbb{R}^{d^{(\text{pool})}} \to \mathcal{Y})$ from the associative algebra of $\mathit{Pool}$ to that of $\mathcal{A}$
+		\item there is a homomorphism $(\tau_{1}: \mathbb{R}^{d^{(T)}} \to (\mathcal{Y} \times \mathbb{R}_{\geq 0}), \tau_{2}: \mathbb{R}^{d^{(\mathrm{pool})}} \to \mathcal{Y})$ from the associative algebra of $\mathit{Pool}$ to that of $\mathcal{A}$
 	\end{enumerate*}.
@@
 \begin{thm}\label{thm:ltag:gcnn-ltag-formulation}
-	A \ac{gcnn} of the form $\mathit{Pool}(\mathit{Conv}(G))$ has a nontrivial \ac{lta} formulation if $\mathit{Conv}$ is a stack of neighborhood convolution layers and $\mathit{Pool}$ is an associative weighted aggregation layer.
+	A \ac{gcnn} of the form $h(G) = \mathit{Pool}(\mathit{Conv}(G))$ has a nontrivial \ac{lta} formulation if $\mathit{Conv}$ is a stack of neighborhood convolution layers and $\mathit{Pool}$ is an associative weighted aggregation layer.
 \end{thm}
@@
 	a formulation for vertex $k$-multiset neighborhood convolutions can be obtained analogously which concludes the proof\footnote{
-		The \ac{lta} formulation in this proof assumes that the \ac{gcnn}'s output is translated into the target score space via $\tau_{2}: \mathbb{R}^{d} \to \mathcal{Y}$.
-		All this means is that an \acs{lta}-like \ac{gcnn} must produce outputs $y \in \mathcal{Y}$;
-		for example if a \ac{gcnn} returns one-hot vector encodings of discrete classes those must be decoded.
+		The \ac{lta} formulation in this proof assumes that the \ac{gcnn}'s output is translated into the target score space via $\tau_{2}: \mathbb{R}^{d^{(\mathrm{pool})}} \to \mathcal{Y}$.
+		All this means is that an \acs{lta}-like \ac{gcnn} must produce outputs that are in $\mathcal{Y}$;
+		for example if $\mathit{Pool}$ returns one-hot vector encodings of discrete classes, those must be decoded.
 	}.
@@
 Note that \cref{thm:ltag:gcnn-ltag-formulation} requires \acp{gcnn} without a final \ac{mlp} which is unlike the architecture proposed in the \ac{gcn}~\cite{Kipf2017} and \ac{gin}~\cite{Xu2018} papers.
-The reason for this is that the composition $\mathit{\acs*{mlp}} \circ \mathit{Pool}$ cannot be guaranteed to be homomorphic to a weighted aggregation function $\mathcal{A}$ due to the universal approximation theorem.
+The reason for this is that the composition $\mathrm{\acs*{mlp}} \circ \mathit{Pool}$ cannot be guaranteed to be homomorphic to a weighted aggregation function $\mathcal{A}$ due to the universal approximation theorem.
 
@@
 		\begin{enumerate*}[label={\arabic*.}]
-			\item SortPooling~\cite{Zhang2018} works by aggregating the top-$p$ convolved vectors $z_i \in \mathbb{R}^{d^{(T)}}$ based on a lexicographic ordering,
-			\item \ac{sagpool}~\cite{Lee2019} also aggregates the top-$p$ vectors $z_i$ but determines their order based on attention weights which are encoded as one component of each $z_i$
+			\item SortPooling~\cite{Zhang2018} works by aggregating the top-$p$ convolved vectors $z_i \in \mathbb{R}^{d^{(T)}}$ based on a lexicographic ordering $\prec$,
+			\item \ac{sagpool}~\cite{Lee2019} also aggregates the top-$p$ vectors $z_i$ but determines their ordering $\prec$ based on attention weights which are encoded as one component of each $z_i$
 		\end{enumerate*}.
 		Both of those pooling approaches are not compatible with the \ac{lta} formulation from \cref{thm:ltag:gcnn-ltag-formulation} because top-$p$ aggregation is generally not associative.
-		This is shown by the fact that there is no $\rho: \mathbb{R}^{* \times d^{(T)}} \to \mathbb{R}^{ d^{(T)}}$ s.t.\ for all vectors $z_1 < z_2$ and and aggregation functions $\mathcal{A}$ it holds that%
-		\begin{alignat*}{2}
-			&& \mathcal{A}(\mathit{top}_{2}(\ldblbrace z_1, z_2, z_2 \rdblbrace)) &= \mathcal{A}(\mathit{top}_{2}(\ldblbrace z_1, \rho(\ldblbrace z_2, z_2 \rdblbrace) \rdblbrace)) \\
-			\Leftrightarrow&& \mathcal{A}(\ldblbrace z_2, z_2 \rdblbrace) &= \mathcal{A}(\ldblbrace z_1, \rho(\ldblbrace z_2, z_2 \rdblbrace) \rdblbrace)\quad\lightning
+		This is shown by the fact that there is no $\rho: \mathbb{R}^{* \times d^{(T)}} \to \mathbb{R}^{ d^{(T)}}$ s.t.\ for all aggregation functions $\mathcal{A}$ and vectors $z_1 \prec z_2$ with $\mathcal{A}(\ldblbrace z_2, z_2 \rdblbrace) = z_2$ (see \cref{defn:ltag:weighted-agg}) it holds that
+		\begin{alignat*}{3}
+			&& \mathcal{A}(\mathit{top}_{2}(\ldblbrace z_1, z_2, z_2 \rdblbrace)) &= \mathcal{A}(\mathit{top}_{2}(\ldblbrace z_1, \rho(\ldblbrace z_2, z_2 \rdblbrace) \rdblbrace))&& \\
+			\Leftrightarrow&& z_2 &= \mathcal{A}(\ldblbrace z_1, \rho(\ldblbrace z_2, z_2 \rdblbrace) \rdblbrace)&&\lightning\text{ if $\mathcal{A} \neq \max$.}
 		\end{alignat*}
@@
 
-Since neither SortPooling nor \ac{sagpool} are compatible with \ac{lta}, we propose an associative weighted aggregation layer based on \ac{sagpool} which uses a softmax-based attention filter instead of a top-$p$ filter.
+Since neither SortPooling nor \ac{sagpool} are compatible with \ac{lta}, we propose an associative weighted aggregation layer inspired by \ac{sagpool} which uses a softmax-based attention filter instead of a top-$p$ filter.
 \begin{defn}
 	We define \ac{sampool} as
-	\[\mathit{SAM}({\ldblbrace (y_i, w_i) \rdblbrace}_{i=1}^n) \coloneqq\frac{1}{\sum_{i=1}^n e^{w_i}} \sum_{i=1}^n e^{w_i} y_i \text{.}\]
+	\[\mathrm{SAM}({\ldblbrace (y_i, w_i) \rdblbrace}_{i=1}^n) \coloneqq\frac{1}{\sum_{i=1}^n e^{w_i}} \sum_{i=1}^n e^{w_i} y_i \text{.}\]
 \end{defn}
 If the stack of convolution layers of a \ac{gcnn} produces vectors $z_i = (y_i, w_i) \in [0, 1] \times \mathbb{R}$, \ac{sampool} satisfies \cref{defn:ltag:assoc-aggregation-pooling} and is therefore compatible with \acs{lta}.
-To see why this is the case, the associativity and homomorphicity to a weighted aggregation function $\mathcal{A}_{\text{SAM}}$ must be shown.
-The associativity of $\mathit{SAM}$ follows directly from its commutativity and the fact that
+To see why this is the case, the associativity and homomorphicity to a weighted aggregation function $\mathcal{A}_{\mathrm{SAM}}$ must be shown.
+The associativity of $\mathrm{SAM}$ follows directly from its commutativity and the fact that
 \begin{gather*}
-	\mathit{SAM}({\ldblbrace (y_i, w_i) \rdblbrace}_{i=1}^n) = \frac{1}{e^{w'} + \sum_{i=j+1}^n e^{w_i}} \left( e^{w'} \underbrace{\mathit{SAM}({\ldblbrace (y_i, w_i) \rdblbrace}_{i=1}^j)}_{y'} + \smashoperator{\sum_{i=j+1}^n} e^{w_i} y_i \right) \\
-	\text{for all } j \in [n] \text{ with } w' \coloneqq \ln\sum_{i=1}^{j} e^{w_i} \text{, i.e.\ } \rho_{\mathit{SAM}}({\ldblbrace (y_i, w_i) \rdblbrace}_{i=1}^j) = (y', w')\text{.}
+	\mathrm{SAM}({\ldblbrace (y_i, w_i) \rdblbrace}_{i=1}^n) = \frac{1}{e^{w'} + \sum_{i=j+1}^n e^{w_i}} \left( e^{w'} \underbrace{\mathrm{SAM}({\ldblbrace (y_i, w_i) \rdblbrace}_{i=1}^j)}_{y'} + \smashoperator{\sum_{i=j+1}^n} e^{w_i} y_i \right) \\
+	\text{for all } j \in [n] \text{ with } w' \coloneqq \ln\sum_{i=1}^{j} e^{w_i} \text{, i.e.\ } \rho_{\mathrm{SAM}}({\ldblbrace (y_i, w_i) \rdblbrace}_{i=1}^j) = (y', w')\text{.}
 \end{gather*}
-The homomorphicity of $\mathit{SAM}$ to a weighted aggregation function $\mathcal{A}_{\text{SAM}}$ follows by choosing $\mathcal{A}_{\text{SAM}}({\ldblbrace (y_i, w_i) \rdblbrace}_{i=1}^n) \coloneqq {\left(\sum_{i=1}^n w_i\right)}^{-1} \sum_{i=1}^n w_i y_i$ and the trivial homomorphism $\tau_1(y, w) \coloneqq (y, e^w)$, $\tau_2(y) \coloneqq y$ between the algebras of $\mathit{SAM}$ and $\mathcal{A}_{\text{SAM}}$.
+The homomorphicity of $\mathrm{SAM}$ to a weighted aggregation function $\mathcal{A}_{\mathrm{SAM}}$ follows by choosing $\mathcal{A}_{\mathrm{SAM}}({\ldblbrace (y_i, w_i) \rdblbrace}_{i=1}^n) \coloneqq {\left(\sum_{i=1}^n w_i\right)}^{-1} \sum_{i=1}^n w_i y_i$ and the trivial homomorphism $\tau_1(y, w) \coloneqq (y, e^w)$, $\tau_2(y) \coloneqq y$ between the algebras of $\mathrm{SAM}$ and $\mathcal{A}_{\mathrm{SAM}}$.
 
@@
 \Acp{gcnn} using vertex neighborhood convolutions on the other hand were shown in \cref{thm:ltag:gcnn-ltag-formulation} to have an \ac{lta} formulation with localized subtree constituents.
-The only requirement for this formulation to exist is that an associative pooling layer like $\mean$, $\min$, $\max$ or $\mathit{SAM}$ is used to combine the convolved feature vectors into the final output.
-
-\section{A Novel \acs*{lta}-Inspired \acs*{gcnn} Architecture}%
-\label{sec:ltag:wl2gnn}
-
-\subsection{Shortcomings of $k$-\acsp*{gnn}}%
-\label{sec:ltag:wl2gnn:kgnn-problems}
-
-\subsection{The 2-\acs*{wl} Convolution Operator}%
-\label{sec:ltag:wl2gnn:definition}
-
-\subsection{The Expressive Power of 2-\acs*{wl}-\acsp*{gcnn}}%
-\label{sec:ltag:wl2gnn:properties}
-
-\subsection{Efficient Implementation on \acsp*{gpgpu}}%
-\label{sec:ltag:wl2gnn:implementation}
+The only requirement for this formulation to exist is that an associative pooling layer like $\mean$, $\min$, $\max$ or $\mathrm{SAM}$ is used to combine the convolved feature vectors into the final output.

### 2020-03-26 thesis/content/chapter-ltd.tex
reorganized thesis: wl2gnn gets its own chapter
includes the intro for the wl2gcnn chapter
+ various improvements

%!TEX root = ../main.tex
% chktex-file 46
\chapter{Learning to Decompose Graphs}%
\label{sec:ltd}

In the last chapter we saw how existing \ac{gcr} approaches relate to \ac{lta}.
Despite their differences, all described \ac{lta} formulations have one thing in common:
Their decomposition functions $\psi$ all split a given graph $G$ into constituents spanned by \ac{bfs} subtrees of $G$ or simply into connected components of $G$.
The only exception to this is the \ac{lta} formulation of \acp{svm} that use fingerprint embeddings; there the constituents are all isomorphic to handpicked substructures.

Apart from the fingerprint embedding approach, which requires domain knowledge in order to pick meaningful substructure patterns, current \ac{gcr} approaches use constituents that are at-best \textit{localized} but not necessarily \textit{interpretable}.
The defining characteristic of \ac{lta} proposed in \cref{sec:ltag:definition}, \textit{localized explainability}, is therefore only partially satisfied by existing approaches.
This shortcoming of the existing \ac{lta} formulations for structured input data gives rise to a new problem:
\begin{defn}
	The \ac{ltd} problem is solved by finding a graph decomposition function $\psi: \mathcal{G}_{\mathcal{D}} \to \mathcal{P}(\mathcal{G}_{\mathcal{D}})$ which splits all graphs $G \in \mathcal{G}_{\mathcal{D}}$ from a given domain $\mathcal{D}$ into constituent subgraphs which are individually ``meaningful'' in $\mathcal{D}$.
\end{defn}

Since a comprehensive analysis of the \ac{ltd} problem would be beyond the scope of this thesis, we focus on two simpler questions:
\begin{enumerate}
	\item \textit{How can the \ac{ltd} problem be solved by a stack of graph convolution layers?}
	\item \textit{What could be the foundation for such an ``\acs{ltd}-convolution'' layer?}
\end{enumerate}

We provide an answer to the first question in \cref{sec:ltd:edge-filter} by showing how in principle a graph convolution layer can learn decomposition functions $\psi$.
This is motivated by the fact that neighborhood convolutions already implicitly define a static subtree decomposition function as we saw in \cref{sec:ltag:formulation:gcnn}.

After establishing the connection between the \acs{ltd} problem and graph convolutions on a high level, we then tackle the second question in \crefrange{sec:ltd:kgnn-problems}{sec:ltd:wl2gnn-implementation}.
There a novel graph convolution approach is proposed which could serve as a starting point for a true ``\ac{ltd}-\ac{gcnn}''.

\section{Learning Constituents via Edge-Filters}%
\label{sec:ltd:edge-filter}

\section{Shortcomings of the Existing $k$-\acs*{gnn}}%
\label{sec:ltd:kgnn-problems}

\section{The 2-\acs*{wl} Convolution Operator}%
\label{sec:ltd:wl2gnn-definition}

\section{The Expressive Power of 2-\acs*{wl}-\acsp*{gnn}}%
\label{sec:ltd:wl2gnn-properties}

\section{Implementing 2-\acs*{wl}-\acsp*{gnn} on \acsp*{gpgpu}}%
\label{sec:ltd:wl2gnn-implementation}

### 2020-03-26 thesis/content/chapter-related-work.tex
reorganized thesis: wl2gnn gets its own chapter
includes the intro for the wl2gcnn chapter
+ various improvements

@@
 }.
-Consequently a small graph $G$ will typically have a large spectral distance to a large graph $H$ (assuming ${|\mathcal{V}_G|} \ll {|\mathcal{V}_H|}$) irrespective of their similarity when ignoring the scale difference.
+Consequently a small graph $G$ will typically have a large spectral distance to a large graph $H$ (assuming ${|\mathcal{E}_G|} \ll {|\mathcal{E}_H|}$) irrespective of their similarity when ignoring the scale difference.
 In domains where the absolute size of a graph should not influence its spectral characterization, it can therefore be useful to normalize the spectrum.
@@
 \begin{defn}\label{defn:related:laplacian-sym}
-	$L^{\text{sym}}_G \coloneqq D^{-\frac{1}{2}} L_G D^{-\frac{1}{2}}$ is the symmetric normalized Laplacian of $G$.
+	$L^{\text{sym}}_G \coloneqq D^{-\frac{1}{2}} L_G D^{-\frac{1}{2}}$ is the \textit{symmetric normalized Laplacian} of $G$.
 \end{defn}
@@
 The comparison shows that \ac{gcn} convolutions can be understood as iterative refinements of continuous ``color vectors'' $Z^{(t)}[v] \in \mathbb{R}^{d^{(t)}}$.
-Then the continuous analogue to the injective 1-\acs{wl} hash function $h: \mathcal{C}^{*} \to \mathcal{C}$ can be defined as $h_{\text{\acs*{gcn}}}(z_0, z_1, \dots, z_m) = \sigma(\sum_{j=0}^m \mu(v_0) \mu(v_j) z_j \Theta^{(t)})$ which computes a single color vector from multiple color vectors $z_j$ where $v_j$ are the corresponding vertices.
+Then the continuous analogue to the injective 1-\acs{wl} hash function $h: \mathcal{C}^{*} \to \mathcal{C}$ can be defined as $h_{\mathrm{\acs*{gcn}}}(z_0, z_1, \dots, z_m) = \sigma(\sum_{j=0}^m \mu(v_0) \mu(v_j) z_j \Theta^{(t)})$ which computes a single color vector from multiple color vectors $z_j$ where $v_j$ are the corresponding vertices.
 
@@
 They show that the discriminative power of such a \ac{gcnn} is upper-bounded by that of 1-\acs{wl}.
-In particular the \ac{gcn} architecture is strictly less powerful than 1-\acs{wl} due to the fact that $h_{\text{\acs*{gcn}}}$ is not an injective hash function.
+In particular the \ac{gcn} architecture is strictly less powerful than 1-\acs{wl} due to the fact that $h_{\mathrm{\acs*{gcn}}}$ is not an injective hash function.
 To overcome this limitation an alternative neighborhood aggregation architecture called \ac{gin} was proposed:
 \begin{align}
-	Z^{(t)}[v] \coloneqq \text{\acs*{mlp}}^{(t)}\left( Z^{(t-1)}[v] + \smashoperator[lr]{\sum_{u \in \Gamma_G(v)}} Z^{(t-1)}[u] \right)
+	Z^{(t)}[v] \coloneqq \mathrm{\acs*{mlp}}^{(t)}\left( Z^{(t-1)}[v] + \smashoperator[lr]{\sum_{u \in \Gamma_G(v)}} Z^{(t-1)}[u] \right)
 \end{align}
-By leaving out the normalization factors $\mu(v)$ and by using a \ac{mlp} instead of the single fully-connected layer $\sigma \circ \Theta^{(t)}$ the resulting hashing function $h_{\text{\acs*{gin}}}(z_1, \dots, z_m) = \text{\acs*{mlp}}^{(t)}(\sum_{j=1}^{m} z_j)$ becomes injective\footnote{
+By leaving out the normalization factors $\mu(v)$ and by using a \ac{mlp} instead of the single fully-connected layer $\sigma \circ \Theta^{(t)}$ the resulting hashing function $h_{\mathrm{\acs*{gin}}}(z_1, \dots, z_m) = \mathrm{\acs*{mlp}}^{(t)}(\sum_{j=1}^{m} z_j)$ becomes injective\footnote{
 	Under the assumption that the set of possible input feature matrices $X \in \mathbb{R}^{n \times d^{(0)}}$ is countable.
@@
 One way to combine feature vectors is called \textit{SortPooling}~\cite{Zhang2018}.
-It interprets the convolved vectors as continuous \ac{wl} colors that encode different structural roles of vertices (or vertex sets in case of $k$-\acp{gnn}).
+It interprets the convolved vectors as continuous \ac{wl} colors that encode different structural roles of vertices (or vertex $k$-multisets in case of $k$-\acp{gnn}).
 By imposing a component-wise lexicographic ordering on the set of color vectors, it can be reduced to a vertex permutation invariant top-$p$ list of vectors $(z_1, \dots, z_p)$ for some fixed $p \in \mathbb{N}$.

### 2020-03-27 thesis/content/chapter-graph-lta.tex
added edge filtering section

@@
 \begin{defn}
-	A problem domain $\mathcal{D}$ satisfies the \textit{\ac{lta} assumption} iff.\ there is an \ac{lta} method which produces models with an equal or lower out-of-sample error than the models produced by non-\acs{lta} methods for most training samples $\mathcal{D}_{\text{train}} \subseteq \mathcal{D}$.
+	A problem domain $\mathcal{D}$ satisfies the \textit{\ac{lta} assumption} iff.\ there is an \ac{lta} method which produces models with an equal or lower out-of-sample error than the models produced by non-\acs{lta} methods for most training samples $\mathcal{D}_{\mathrm{train}} \subseteq \mathcal{D}$.
 \end{defn}
@@
 \begin{proof}
-	Let $h: \mathcal{G} \to {\{ -1, +1 \}}$ be a binary graph classifier expressed as $h = h_{\text{\acs*{svm}}} \circ \varphi$ where $\varphi: \mathcal{G} \to \mathbb{N}_0^{d}$ is an \ac{sce} and $h_{\text{\acs*{svm}}}: \mathbb{R}^{d} \to {\{ -1, +1 \}}$ a standard \ac{svm} classifier.
+	Let $h: \mathcal{G} \to {\{ -1, +1 \}}$ be a binary graph classifier expressed as $h = h_{\mathrm{\acs*{svm}}} \circ \varphi$ where $\varphi: \mathcal{G} \to \mathbb{N}_0^{d}$ is an \ac{sce} and $h_{\mathrm{\acs*{svm}}}: \mathbb{R}^{d} \to {\{ -1, +1 \}}$ a standard \ac{svm} classifier.
 	Additionally, let $\psi_{\varphi}$ be some underlying decomposition of $\varphi$ with ${\{ \gamma_{\varphi, i} \}}_{i=1}^d$ being the corresponding multiplicity functions.
@@
 	Based on this decomposition we now bring the \ac{svm} graph classifier $h$ into an \ac{lta} formulation.
-	If $h$ is trained on a dataset $\mathcal{D}_{\text{train}} = {\{ (G_1, y_1), \dots, (G_N, y_N) \}}$, via the kernel trick it can be expressed as
+	If $h$ is trained on a dataset $\mathcal{D}_{\mathrm{train}} = {\{ (G_1, y_1), \dots, (G_N, y_N) \}}$, via the kernel trick it can be expressed as
 	\begin{align*}
@@
 The central statement of \cref{thm:ltag:svm-ltag-formulation} is that \acs{svm} graph classifiers that use an \ac{sce} $\varphi$ implicitly perform an \acs{lta}-like weighted majority vote aggregation of constituent scores $y_t$.
-Those constituent scores can in turn be expressed as a majority vote of the graph scores in the training dataset $\mathcal{D}_{\text{train}}$:
+Those constituent scores can in turn be expressed as a majority vote of the graph scores in the training dataset $\mathcal{D}_{\mathrm{train}}$:
 \begin{align}
@@
 	\item \textbf{Fingerprint embeddings:}\label{itm:ltag:fingerprint-lta-formulation}
-		Here the $i$-th embedding component represent the number of occurrences of a substructure $S_i$, i.e.\ ${\varphi}_{\text{FP}}(G)[i] \coloneqq \mathit{count}_{S_i}(G)$ (see \cfullref{defn:related:wl-count-detect}).
-		Such a fingerprint embedding naturally is an \ac{sce} with the decomposition functions $\psi_{\text{FP}, i}(G) \coloneqq \{ G[\set(s)]\, |\, s \in \mathcal{V}_G^{*} \land G[s] \equiv S_i \}$ and the multiplicity functions $\gamma_{\text{FP}, i}(c) \coloneqq \mathbbm{1}[c \simeq S_i]$ since those functions produce the subgraphs that are counted by ${\varphi}_{\text{FP}}(G)[i]$ with multiplicity $1$.
+		Here the $i$-th embedding component represent the number of occurrences of a substructure $S_i$, i.e.\ ${\varphi}_{\mathrm{FP}}(G)[i] \coloneqq \mathit{count}_{S_i}(G)$ (see \cfullref{defn:related:wl-count-detect}).
+		Such a fingerprint embedding naturally is an \ac{sce} with the decomposition functions $\psi_{\mathrm{FP}, i}(G) \coloneqq \{ G[\set(s)]\, |\, s \in \mathcal{V}_G^{*} \land G[s] \equiv S_i \}$ and the multiplicity functions $\gamma_{\mathrm{FP}, i}(c) \coloneqq \mathbbm{1}[c \simeq S_i]$ since those functions produce the subgraphs that are counted by ${\varphi}_{\mathrm{FP}}(G)[i]$ with multiplicity $1$.
 % chktex 21
 
-		Under the assumption that the substructure patterns $S_i$ are chosen s.t.\ the their instances $c \in \psi_{\text{FP}, i}(G)$ are nontrivial interpretable constituents, the underlying joint decomposition $\psi_{\text{FP}}(G) \coloneqq \bigcup_{i=1}^d \psi_{\text{FP}, i}(G)$ must also be nontrivial and interpretable.
+		Under the assumption that the substructure patterns $S_i$ are chosen s.t.\ the their instances $c \in \psi_{\mathrm{FP}, i}(G)$ are nontrivial interpretable constituents, the underlying joint decomposition $\psi_{\mathrm{FP}}(G) \coloneqq \bigcup_{i=1}^d \psi_{\mathrm{FP}, i}(G)$ must also be nontrivial and interpretable.
 		Thus, by \cref{thm:ltag:svm-ltag-formulation}, \acp{svm} trained on fingerprint embeddings are \ac{lta} models that produce locally explainable predictions.
@@
 	\item \textbf{\ac{wl} subtree kernel:}\label{itm:ltag:wlst-lta-formulation}
-		The individual components of the \ac{wl} subtree kernel's embedding $\varphi_{\text{ST}}$ represent the number of occurrences of a color $\kappa_i \in \mathcal{C}$, i.e.\ $\varphi_{\text{ST}}(G)[i, t] \coloneqq \mathit{dist}_{\chi_{G,1}^{(t)}}(\kappa_i)$ as defined in \cfullref[ on ]{eq:related:wl-subtree-kernel}\footnote{
+		The individual components of the \ac{wl} subtree kernel's embedding $\varphi_{\mathrm{ST}}$ represent the number of occurrences of a color $\kappa_i \in \mathcal{C}$, i.e.\ $\varphi_{\mathrm{ST}}(G)[i, t] \coloneqq \mathit{dist}_{\chi_{G,1}^{(t)}}(\kappa_i)$ as defined in \cfullref[ on ]{eq:related:wl-subtree-kernel}\footnote{
 			To avoid confusion $\kappa_i \in \mathcal{C}$ is used for colors and $c_j \in \mathcal{G}$ for constituents in this context.
@@
 
-		To show that $\varphi_{\text{ST}}$ is an \ac{sce}, we have to define decomposition functions $\psi_{\varphi, i, t}$ and multiplicity functions $\gamma_{\varphi, i, t}$ s.t.\ $\varphi(G)[i, t] = \sum_{c \in \psi_{\varphi, i, t}(G)} \gamma_{\varphi, i, t}(c)$.
+		To show that $\varphi_{\mathrm{ST}}$ is an \ac{sce}, we have to define decomposition functions $\psi_{\varphi, i, t}$ and multiplicity functions $\gamma_{\varphi, i, t}$ s.t.\ $\varphi(G)[i, t] = \sum_{c \in \psi_{\varphi, i, t}(G)} \gamma_{\varphi, i, t}(c)$.
 		Since $\varphi(G)[i, t]$ counts the occurrences $S_{j,t}$ of a subtree $S_{\kappa_i}$ in $G$, the \ac{sce} requirement would be trivially satisfied if each $S_{j,t}$ were a constituent with multiplicity $1$.
@@
 		\begin{align}
-			&\gamma_{\varphi, i, t}(c) \coloneqq {\left|\{ S = \text{\acs*{bfs}}(c, v_{\text{root}}, t)\, |\, v_{\text{root}} \in \mathcal{V}_c \land S \simeq S_{\kappa_i} \land \mathit{complete}(S) \}\right|}\label{eq:ltag:wlst-constituent-subtree-count} \\
-			&\text{with }\mathit{complete}(S) \Leftrightarrow \forall\ \text{leaf nodes $v_{\text{leaf}}$ of the tree $S$}: {|\Gamma_c(v_{\text{leaf}})|} = {|\Gamma_G(v_{\text{leaf}})|}
+			&\gamma_{\varphi, i, t}(c) \coloneqq {\left|\{ S = \text{\acs*{bfs}}(c, v_{\mathrm{root}}, t)\, |\, v_{\mathrm{root}} \in \mathcal{V}_c \land S \simeq S_{\kappa_i} \land \mathit{complete}(S) \}\right|}\label{eq:ltag:wlst-constituent-subtree-count} \\
+			&\text{with }\mathit{complete}(S) \Leftrightarrow \forall\ \text{leaf nodes $v_{\mathrm{leaf}}$ of the tree $S$}: {|\Gamma_c(v_{\mathrm{leaf}})|} = {|\Gamma_G(v_{\mathrm{leaf}})|}
 			\text{.} \nonumber
@@
 % chktex 25
-		Also note that, since $\gamma_{\varphi, i, t}(c)$ must only depend on a given constituent $c$ and not on the graph $G$ it was decomposed from, the degree information $|\Gamma_G(v_{\text{leaf}})|$ of the constituent vertices $v \in \mathcal{V}_c$ is assumed to be statically encoded in the labels $l_c[v]$ or feature vectors $x_c[v]$.
+		Also note that, since $\gamma_{\varphi, i, t}(c)$ must only depend on a given constituent $c$ and not on the graph $G$ it was decomposed from, the degree information $|\Gamma_G(v_{\mathrm{leaf}})|$ of the constituent vertices $v \in \mathcal{V}_c$ is assumed to be statically encoded in the labels $l_c[v]$ or feature vectors $x_c[v]$.
 
-		Via the decomposition and multiplicity functions that were just described, $\varphi_{\text{ST}}$ is in fact an \ac{sce} with a nontrivial, subtree-based decomposition function.
+		Via the decomposition and multiplicity functions that were just described, $\varphi_{\mathrm{ST}}$ is in fact an \ac{sce} with a nontrivial, subtree-based decomposition function.
 		Additionally each constituent's diameter is upper bounded by $2T$ (with $T$ being the maximum number of \ac{wl} iterations) which guarantees that constituents are localized within a neighborhood of bounded size.
@@
 	\item \textbf{\ac{wl} shortest path kernel:}\label{itm:ltag:wlsp-lta-formulation}
-		As defined in \cfullref[ on ]{eq:related:wl-shortest-path-kernel}, the $i$-th component of the \ac{wl} shortest path embedding $\varphi_{\text{SP}}$ represents a 4-tuple $(t_i, a_i, b_i, d_i) \in {\{ 0, \dots, T \}} \times \mathcal{C} \times \mathcal{C} \times \mathbb{N}_0$.
-		The value $\varphi_{\text{SP}}(G)[i]$ is defined as the number of vertex pairs $v_a, v_b \in \mathcal{V}_G$ that have a shortest connecting path of length $d_i$ and the \ac{wl} color combination $\chi_{G,1}^{(t_i)}(v_a) = a_i$ and $\chi_{G,1}^{(t_i)}(v_b) = b_i$.
+		As defined in \cfullref[ on ]{eq:related:wl-shortest-path-kernel}, the $i$-th component of the \ac{wl} shortest path embedding $\varphi_{\mathrm{SP}}$ represents a 4-tuple $(t_i, a_i, b_i, d_i) \in {\{ 0, \dots, T \}} \times \mathcal{C} \times \mathcal{C} \times \mathbb{N}_0$.
+		The value $\varphi_{\mathrm{SP}}(G)[i]$ is defined as the number of vertex pairs $v_a, v_b \in \mathcal{V}_G$ that have a shortest connecting path of length $d_i$ and the \ac{wl} color combination $\chi_{G,1}^{(t_i)}(v_a) = a_i$ and $\chi_{G,1}^{(t_i)}(v_b) = b_i$.
 
@@
 		Depending on ones domain-specific interpretation of \textit{localized explainability}, this restriction can be seen to be not ``\acs{lta}-like''.
-		Since we do not attempt to clearly separate \ac{lta} from non-\acs{lta} methods, $\varphi_{\text{SP}}$ is categorized as an in-between case here.
+		Since we do not attempt to clearly separate \ac{lta} from non-\acs{lta} methods, $\varphi_{\mathrm{SP}}$ is categorized as an in-between case here.
 		\badgebox[t_yellow]{partially \acs*{lta}-like}
@@
 
-\subsubsection{Graph Convolutions as \ac*{lta} Decomposition and Evaluation Functions}
+\subsection{Graph Convolutions as Decomposition and Evaluation}%
+\label{sec:ltag:formulation:gcnn:conv}
 
@@
 If there is a translation function $\tau_{1}: \mathbb{R}^{d^{(T)}} \to \mathcal{Y} \times \mathbb{R}_{\geq 0}$, the convolved vectors $z_i \in \mathbb{R}^{d^{(T)}}$ can be interpreted as constituent evaluations.
-One trivial example for such a translation function would be the identity $\tau_{1}(z) = z$ for a graph regression problem with the score domain $\mathcal{Y} = [0, 1]$ and where the last convolution layer uses a sigmoid activation function with two output dimensions s.t.\ $z \in {[0,1]}^2$.
+One trivial example for such a translation function would be the identity $\tau_{1}(z) = z$ for a graph regression problem with the score domain $\mathcal{Y} = [0, 1]$ and where the last convolution layer uses a logistic activation function with two output dimensions s.t.\ $z \in {[0,1]}^2$.
 Assuming that there is a translation function $\tau_{1}$ we can define a multi-score evaluation function $f^{*}(c) \coloneqq {\ldblbrace \tau_{1}(z_i)\, |\, {i \in [n]} \land {c = c_i} \rdblbrace} = {\ldblbrace \tau_{1}(\mathit{Conv}(c)[v_i])\, |\, v_i \in \mathit{root}(c) \rdblbrace}$.
@@
 
-\subsubsection{Graph Pooling as \ac*{lta} Aggregation Functions}
+\subsection{Graph Pooling as Aggregation}%
+\label{sec:ltag:formulation:gcnn:pool}
 
@@
 \begin{defn}\label{defn:ltag:assoc-algebra}
-	A variadic function $\xi: A^{*} \to B$ is called \textit{associative} iff.\ there is an operator $\rho: A^{*} \to A$ s.t.\ $\forall n \in \mathbb{N}: \forall (a_1, \dots, a_n) \in A^n: \forall {1 \leq i \leq j \leq n}: \xi(a_1, \dots, a_n) = \xi(a_1, \dots, a_{i-1}, \rho(a_i, \dots, a_j), a_{j+1}, \dots, a_n)$.
+	A variadic function $\xi: A^{*} \to B$ is called \textit{associative} iff.\ there is a replacement operator $\rho: A^{*} \to A$ s.t.\ $\forall n \in \mathbb{N}: \forall (a_1, \dots, a_n) \in A^n: \forall {1 \leq i \leq j \leq n}: \xi(a_1, \dots, a_n) = \xi(a_1, \dots, a_{i-1}, \rho(a_i, \dots, a_j), a_{j+1}, \dots, a_n)$.
 	The heterogeneous algebra\footnote{
@@
 	We already saw that a stack of neighborhood convolutions computes the values of a multi-score evaluation function $f^{*}$ for the subtree constituents returned by a decomposition function $\psi$.
-	Since $\mathit{Pool}$ is assumed to be an associative weighted aggregation layer, there must be an operator $\rho_{\mathit{Pool}}: \mathbb{R}^{n \times d^{(T)}} \to \mathbb{R}^{n \times d^{(T)}}$ and a homomorphism $(\tau_{1}, \tau_{2})$ to an associative algebra with the operators $\mathcal{A}: {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*} \to \mathcal{Y}$ and $\rho_{\mathcal{A}}: {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*} \to {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}$.
+	Since $\mathit{Pool}$ is assumed to be an associative weighted aggregation layer, there must be a replacement operator $\rho_{\mathit{Pool}}: \mathbb{R}^{n \times d^{(T)}} \to \mathbb{R}^{n \times d^{(T)}}$ and a homomorphism $(\tau_{1}, \tau_{2})$ to an associative algebra with the operators $\mathcal{A}: {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*} \to \mathcal{Y}$ and $\rho_{\mathcal{A}}: {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*} \to {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}$.
 	This allows us to rewrite the \ac{gcnn} as

### 2020-03-27 thesis/content/chapter-ltd.tex
added edge filtering section

@@
 \begin{defn}
-	The \ac{ltd} problem is solved by finding a graph decomposition function $\psi: \mathcal{G}_{\mathcal{D}} \to \mathcal{P}(\mathcal{G}_{\mathcal{D}})$ which splits all graphs $G \in \mathcal{G}_{\mathcal{D}}$ from a given domain $\mathcal{D}$ into constituent subgraphs which are individually ``meaningful'' in $\mathcal{D}$.
+	The \ac{ltd} problem is solved by finding a graph decomposition function $\psi: \mathcal{G}_{\mathcal{D}} \to \mathcal{P}(\mathcal{G}_{\mathcal{D}})$ which splits all graphs $G \in \mathcal{G}_{\mathcal{D}}$ from a given domain $\mathcal{D}$ into constituent subgraphs which are individually ``meaningful'' in the domain $\mathcal{D}$.
 \end{defn}
 
-Since a comprehensive analysis of the \ac{ltd} problem would be beyond the scope of this thesis, we focus on two simpler questions:
+As per our definition of \ac{lta} from \cref{sec:ltag:definition}, the quality of an \ac{lta} formulation is determined by its chosen solution for the \ac{ltd} problem.
+Since a comprehensive analysis of this problem would be beyond the scope of this thesis, we focus on the relation between \ac{ltd} and \acp{gcnn}.
+The goal of this chapter is to answer the following two questions:
 \begin{enumerate}
-	\item \textit{How can the \ac{ltd} problem be solved by a stack of graph convolution layers?}
+	\item \textit{Can a stack of graph convolution layers learn a decomposition function dynamically instead of using a static subtree decomposition?}
 	\item \textit{What could be the foundation for such an ``\acs{ltd}-convolution'' layer?}
@@
 
-We provide an answer to the first question in \cref{sec:ltd:edge-filter} by showing how in principle a graph convolution layer can learn decomposition functions $\psi$.
-This is motivated by the fact that neighborhood convolutions already implicitly define a static subtree decomposition function as we saw in \cref{sec:ltag:formulation:gcnn}.
+We provide an answer to the first question in \cref{sec:ltd:edge-filter} by showing how decomposition functions $\psi$ can be learned via so-called \textit{edge filters} as part of a convolutional \ac{gnn} architecture.
+This establishes the connection between the \acs{ltd} problem and graph convolutions on a high level.
+Then the second question is tackled in \crefrange{sec:ltd:kgnn-problems}{sec:ltd:wl2gnn-implementation}.
+There a novel graph convolution approach is proposed which could serve as a starting point for an \ac{ltd} convolution layer.
 
-After establishing the connection between the \acs{ltd} problem and graph convolutions on a high level, we then tackle the second question in \crefrange{sec:ltd:kgnn-problems}{sec:ltd:wl2gnn-implementation}.
-There a novel graph convolution approach is proposed which could serve as a starting point for a true ``\ac{ltd}-\ac{gcnn}''.
+\section{Learning Constituents via Edge Filters}%
+\label{sec:ltd:edge-filter}
 
-\section{Learning Constituents via Edge-Filters}%
-\label{sec:ltd:edge-filter}
+As we saw in \cref{sec:ltag:formulation:gcnn:conv}, the constituents of neighborhood convolutions are spanned by \ac{bfs} trees of depth $T$, where $T$ is the number of convolutional layers.
+From this perspective the problem of learning constituents corresponds to learning a pruning operator on the branches of \ac{bfs} trees.
+Such a pruning operator filters the edges that are traversed in each \ac{bfs} step.
+There are two general edge filtering strategies:
+\begin{enumerate}[label={\textbf{\arabic*.}}]
+	\item \textbf{Edge prefiltering:}
+		Here the edge filtering and convolution operations are performed in independent steps:
+		First the edges of a given graph are filtered, then the convolution layers are applied to the filtered graph.
+		This so-called \textit{edge prefiltering} strategy is illustrated in \cref{fig:ltd:subtree-filtering}.
+		The main advantage of prefiltering is that it allows arbitrary combinations of edge filtering and convolution approaches.
+		The main disadvantage is however that the same edges are removed in all \ac{bfs} subtrees.
+		This restricts the expressive power of the learned decomposition function as shown in \cref{fig:ltd:dynamic-subtree-filtering}.
+	\item \textbf{Dynamic edge filtering:}
+		By filtering edges as part of the convolution operation itself, more flexible decompositions can be obtained.
+		In the \textit{dynamic edge filtering} strategy the edge filter is part of the convolution operation and decides which neighbors of a given root node should be aggregated.
+		Using this strategy a decomposition such as that shown in \cref{fig:ltd:dynamic-subtree-filtering} is learnable.
+\end{enumerate}
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=0.8\linewidth]{gfx/ltd/subtree-filtering.pdf}
+	\caption{Illustration of a pruned \ac{bfs} subtree when two edges are removed via prefiltering.}\label{fig:ltd:subtree-filtering}
+\end{figure}
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=0.75\linewidth]{gfx/ltd/dynamic-subtree-filtering.pdf}
+	\caption[A graph decomposition obtained via dynamic edge filtering which cannot be modeled by edge prefiltering.]{
+		A graph decomposition obtained via dynamic edge filtering which cannot be modeled by edge prefiltering.
+		Depending on the \ac{bfs} root node $v_{\mathrm{root}}$ different edges are removed.
+		For $v_{\mathrm{root}} \in \textcolor{t_blue}{\{ v_3, v_4 \}}$ the edges $\{ (v_3, v_1), (v_3, v_2), (v_4, v_5), (v_4, v_6) \}$ are filtered out while for $v_{\mathrm{root}} \in \textcolor{t_red}{\{ v_1, v_2, v_5, v_6 \}}$ the edge $(v_3, v_4)$ is removed.
+% chktex 25
+	}\label{fig:ltd:dynamic-subtree-filtering}
+\end{figure}
+Irrespective of the chosen edge filtering strategy, the decision whether to remove a given edge $e_{ij} = (v_i, v_j)$ or not has to be made based on relevant information about $e_{ij}$ and its surroundings.
+One natural way to encode the information about $e_{ij}$ is via an edge feature vector $z_{ij} \in \mathbb{R}^d$ that is fed into the edge filter.
+We propose that those edge feature vectors could be obtained by a 2-\acs{wl} inspired graph convolution layer since 2-\acs{wl} colors represent the structural roles of edges as described in \cref{sec:related:character:wl}.
 

### 2020-03-27 thesis/content/chapter-related-work.tex
added edge filtering section

@@
 	\centering
-	\begin{subfigure}{0.33\textwidth}
+	\begin{subfigure}{0.32\textwidth}
 		\centering
-		\includegraphics[width=0.8\linewidth]{gfx/related-work/wl1-neighbors.pdf}
-		\subcaption{1-WL $v_1$ neighbors}\label{fig:related:wl-neighbors:1}
+		\includegraphics[width=0.8\textwidth]{gfx/related-work/wl1-neighbors.pdf}
+		\subcaption{$v_1$'s 1-WL neighbors}\label{fig:related:wl-neighbors:1}
 	\end{subfigure}%
-	\begin{subfigure}{0.33\textwidth}
+	\begin{subfigure}{0.32\textwidth}
 		\centering
-		\includegraphics[width=0.8\linewidth]{gfx/related-work/wl2-neighbors.pdf}
-		\subcaption{2-WL $(v_1, v_2)$ neighbors}\label{fig:related:wl-neighbors:2}
+		\includegraphics[width=0.8\textwidth]{gfx/related-work/wl2-neighbors.pdf}
+		\subcaption{$(v_1, v_2)$'s 2-WL neighbors}\label{fig:related:wl-neighbors:2}
 	\end{subfigure}%
-	\begin{subfigure}{0.33\textwidth}
+	\begin{subfigure}{0.36\textwidth}
 		\centering
-		\includegraphics[width=0.8\linewidth]{gfx/related-work/wl3-neighbors.pdf}
-		\subcaption{3-WL $(v_1, v_2, v_3)$ neighbors}\label{fig:related:wl-neighbors:3}
+		\includegraphics[width=0.711\textwidth]{gfx/related-work/wl3-neighbors.pdf}
+		\subcaption{$(v_1, v_2, v_3)$'s 3-WL neighbors}\label{fig:related:wl-neighbors:3}
 	\end{subfigure}
@@
 So-called \textit{fingerprint embeddings} try to match a fixed set of subgraphs $S_1, \dots, S_d$ to the input graph.
-The embedding of a graph $G$ is a binary vector ${\varphi}_{\text{FP}}(G) \in {\{0, 1\}}^d$ with ${\varphi}_{\text{FP}}(G)[i] \coloneqq \mathit{contain}_{S_i}(G)$ (see~\cref{defn:related:wl-count-detect}).
-Alternatively a fingerprint embedding can additionally encode multiplicities via ${\varphi}_{\text{FP}}(G)[i] \coloneqq \mathit{count}_{S_i}(G) \in \mathbb{N}_0$.
+The embedding of a graph $G$ is a binary vector ${\varphi}_{\mathrm{FP}}(G) \in {\{0, 1\}}^d$ with ${\varphi}_{\mathrm{FP}}(G)[i] \coloneqq \mathit{contain}_{S_i}(G)$ (see~\cref{defn:related:wl-count-detect}).
+Alternatively a fingerprint embedding can additionally encode multiplicities via ${\varphi}_{\mathrm{FP}}(G)[i] \coloneqq \mathit{count}_{S_i}(G) \in \mathbb{N}_0$.
 
 This simple approach usually requires a careful choice of subgraphs but can still be competitive with the other more recent approaches we will look at in the following sections.
-Fingerprint embeddings are for example used in multiple state-of-the-art toxicity prediction tools like RASAR~\cite{Luechtefeld2018}\cite{ToxTrack}, ProTox~\cite{Drwal2014}\cite{ProTox}\cite{Banerjee2018} or the \ac{toxTest}~\cite{TEST}.
+Fingerprint embeddings are for example used in multiple state-of-the-art toxicity prediction tools like RASAR~\cite{Luechtefeld2018}\cite{ToxTrack}, the \ac{toxTest}~\cite{TEST} or ProTox~\cite{Drwal2014}\cite{ProTox}\cite{Banerjee2018}.
 
@@
 \begin{align}
-	\varphi_{\text{ST}}(G) \coloneqq \bigoplus_{t=0}^{T} {\left( \mathit{dist}_{\chi_{G,1}^{(t)}}(c) \right)}_{c \in \mathcal{C}}
+	\varphi_{\mathrm{ST}}(G) \coloneqq \bigoplus_{t=0}^{T} {\left( \mathit{dist}_{\chi_{G,1}^{(t)}}(c) \right)}_{c \in \mathcal{C}}
 	\quad\text{with $\oplus$ denoting vector concatenation.} \label{eq:related:wl-subtree-kernel}
 \end{align}
-$\varphi_{\text{ST}}(G)$ encodes the color counts across a fixed number $T$ of 1-\acs{wl} refinement steps.
-Via this mapping, the \ac{wl} subtree kernel function $k_{\text{ST}}: \mathcal{G} \times \mathcal{G} \to \mathbb{R}$ can then be simply written as the standard inner vector product $k_{\text{ST}}(G, H) \coloneqq \langle \varphi_{\text{ST}}(G), \varphi_{\text{ST}}(H) \rangle$.
+$\varphi_{\mathrm{ST}}(G)$ encodes the color counts across a fixed number $T$ of 1-\acs{wl} refinement steps.
+Via this mapping, the \ac{wl} subtree kernel function $k_{\mathrm{ST}}: \mathcal{G} \times \mathcal{G} \to \mathbb{R}$ can then be simply written as the standard inner vector product $k_{\mathrm{ST}}(G, H) \coloneqq \langle \varphi_{\mathrm{ST}}(G), \varphi_{\mathrm{ST}}(H) \rangle$.
 \begin{figure}[ht]
 	\centering
-	\includegraphics[width=0.7\linewidth]{gfx/related-work/wl-subtree.pdf}
+	\includegraphics[width=0.75\linewidth]{gfx/related-work/wl-subtree.pdf}
 	\caption[Illustration of the correspondence between the \ac{wl} color $\chi_{G,1}^{(t)}(v)$ and the breadth-first subtree of depth $l$ rooted at $v$.]{
-		Illustration of the correspondence between the \ac{wl} color $\chi_{G,1}^{(t)}(v)$ and the breadth-first subtree of depth $t$ rooted at $v$ (for $t = 2$ and $v = \texttt{\textcolor{t_blue}{\circled{1}}}$).
+		Illustration of the correspondence between the \ac{wl} color $\chi_{G,1}^{(t)}(v)$ and the breadth-first subtree of depth $t$ rooted at $v$ (for $t = 2$ and $v = \textcolor{t_red}{v_1}$).
 % chktex 25
-		An arrow $A \to B$ indicates that the refined color of $A$ depends on the color of $B$.
+		An arrow $\displaystyle v \rightarrowRHD u$ indicates that the refined color of $v$ depends on the color of $u$.
 		\source[based~on]{Shervashidze2011}
@@
 \Cref{fig:related:wl-subtree} illustrates how this definition relates to subtrees.
-Since all vertex colors $c \in \mathcal{C}$ correspond to a subtree isomorphism class, $k_{\text{ST}}(G, H)$ effectively computes the similarity of $G$ and $H$ by comparing their local subtree structures of depth at most $T$.
-Note that the underlying mapping $\varphi_{\text{ST}}$ cannot be computed independently for each graph $G \in \mathcal{G}_{\mathcal{D}} \subseteq \mathcal{G}$ in a given training dataset $\mathcal{D}$ since the set of colors $c \in \mathcal{C}$ introduced by the hashing function $h$ is conjointly determined by all graphs $\mathcal{G}_{\mathcal{D}}$ (see \cref{defn:related:wl1-refine}).
-Consequently the dimensionality $d$ of $\varphi_{\text{ST}}(G) \in \mathbb{N}^d$ varies, depending on the entire dataset $\mathcal{D}$.
+Since all vertex colors $c \in \mathcal{C}$ correspond to a subtree isomorphism class, $k_{\mathrm{ST}}(G, H)$ effectively computes the similarity of $G$ and $H$ by comparing their local subtree structures of depth at most $T$.
+Note that the underlying mapping $\varphi_{\mathrm{ST}}$ cannot be computed independently for each graph $G \in \mathcal{G}_{\mathcal{D}} \subseteq \mathcal{G}$ in a given training dataset $\mathcal{D}$ since the set of colors $c \in \mathcal{C}$ introduced by the hashing function $h$ is conjointly determined by all graphs $\mathcal{G}_{\mathcal{D}}$ (see \cref{defn:related:wl1-refine}).
+Consequently the dimensionality $d$ of $\varphi_{\mathrm{ST}}(G) \in \mathbb{N}^d$ varies, depending on the entire dataset $\mathcal{D}$.
 
@@
 Extending the idea of the \ac{wl} subtree kernel, the \textit{\ac{wl} shortest path kernel}~\cite{Shervashidze2011}\cite{Borgwardt2005} adds global structural information to the local subtree comparisons.
-The $i$-th component of a graph's shortest path vector $\varphi_{\text{SP}}(G)$ corresponds to the 4-tuple $(t_i, a_i, b_i, d_i) \in {\{ 0, \dots, T \}} \times \mathcal{C} \times \mathcal{C} \times \mathbb{N}_0$ and its value is described by
+The $i$-th component of a graph's shortest path vector $\varphi_{\mathrm{SP}}(G)$ corresponds to the 4-tuple $(t_i, a_i, b_i, d_i) \in {\{ 0, \dots, T \}} \times \mathcal{C} \times \mathcal{C} \times \mathbb{N}_0$ and its value is described by
 \begin{align}
-	{\varphi_{\text{SP}}(G)}[i] \coloneqq \left| \left\{ (v, u) \in \mathcal{V}_G^2 \,|\, {\chi_{G,1}^{(t_i)}(v) = a_i} \land {\chi_{G,1}^{(t_i)}(u) = b_i} \land {d_{\text{SP}}(v, u) = d_i} \right\} \right| \text{,} \label{eq:related:wl-shortest-path-kernel} % chktex 21
+	{\varphi_{\mathrm{SP}}(G)}[i] \coloneqq \left| \left\{ (v, u) \in \mathcal{V}_G^2 \,|\, {\chi_{G,1}^{(t_i)}(v) = a_i} \land {\chi_{G,1}^{(t_i)}(u) = b_i} \land {d_{\mathrm{SP}}(v, u) = d_i} \right\} \right| \text{,} \label{eq:related:wl-shortest-path-kernel} % chktex 21
 \end{align}
-with $d_{\text{SP}}(v, u)$ denoting the length of the shortest path from $v$ to $u$ in $G$.
-Analogous to $k_{\text{ST}}$, the shortest path kernel function is defined as $k_{\text{SP}}(G, H) \coloneqq \langle \varphi_{\text{SP}}(G), \varphi_{\text{SP}}(H) \rangle$.
-Since $\varphi_{\text{SP}}(G)$ has one component for each possible 4-tuple $(t_i, a_i, b_i, d_i)$, its dimensionality grows quadratically with the total number of introduced 1-\acs{wl} colors $\mathcal{C}$ and linearly with the length of the longest shortest path $\max_{G \in \mathcal{G}_{\mathcal{D}}, (v, u) \in \mathcal{V}_G^2} d_{\text{SP}}(v, u)$.
+with $d_{\mathrm{SP}}(v, u)$ denoting the length of the shortest path from $v$ to $u$ in $G$.
+Analogous to $k_{\mathrm{ST}}$, the shortest path kernel function is defined as $k_{\mathrm{SP}}(G, H) \coloneqq \langle \varphi_{\mathrm{SP}}(G), \varphi_{\mathrm{SP}}(H) \rangle$.
+Since $\varphi_{\mathrm{SP}}(G)$ has one component for each possible 4-tuple $(t_i, a_i, b_i, d_i)$, its dimensionality grows quadratically with the total number of introduced 1-\acs{wl} colors $\mathcal{C}$ and linearly with the length of the longest shortest path $\max_{G \in \mathcal{G}_{\mathcal{D}}, (v, u) \in \mathcal{V}_G^2} d_{\mathrm{SP}}(v, u)$.
 While this approach is computationally significantly more expensive than the subtree kernel, it also has a higher discriminative power and allows the shortest path kernel to distinguish even some 1-\acs{wl} indistinguishable graphs.

### 2020-03-28 thesis/content/chapter-ltd.tex
started with k-gnn limits section

@@
 Irrespective of the chosen edge filtering strategy, the decision whether to remove a given edge $e_{ij} = (v_i, v_j)$ or not has to be made based on relevant information about $e_{ij}$ and its surroundings.
-One natural way to encode the information about $e_{ij}$ is via an edge feature vector $z_{ij} \in \mathbb{R}^d$ that is fed into the edge filter.
-We propose that those edge feature vectors could be obtained by a 2-\acs{wl} inspired graph convolution layer since 2-\acs{wl} colors represent the structural roles of edges as described in \cref{sec:related:character:wl}.
+One way to encode the information about $e_{ij}$ is via an edge feature vector $z_{ij} \in \mathbb{R}^d$ that is fed into the edge filter.
 
-\section{Shortcomings of the Existing $k$-\acs*{gnn}}%
-\label{sec:ltd:kgnn-problems}
+We propose that those edge feature vectors can be obtained via a 2-\acs{wl} inspired graph convolution layer since 2-\acs{wl} colors naturally represent the structural roles of edges as described in \cref{sec:related:character:wl}.
+We already saw a \ac{gcnn} architecture in \cref{sec:related:gcr:nn} which could potentially compute such 2-\acs{wl} inspired edge feature vectors, the $k$-\acs{gnn}~\cite{Morris2019}.
+Its variant for $k = 2$ produces edge feature vectors that can be used as the input for an edge filter.
+However, as we will see in the next section, there are significant limitations to the discriminative and computational power of 2-\acsp{gnn}.
+
+\section{Limitations of the Existing 2-\acs*{gnn}}%
+\label{sec:ltd:kgnn-limits}
+
+The $k$-\acs{gnn} is a \ac{gcnn} inspired by the $k$-\ac{wl} algorithm, it convolves feature vectors of vertex $k$-multisets.
+In this section we will compare its $k = 2$ variant with the 2-\ac{wl} algorithm.
+The main difference between the two boils down to their notion of ``neighborhood''.
+As already briefly mentioned in \cref{sec:related:gcr:nn}, 2-\acsp{gnn} define the neighbors of an edge $\textcolor{t_blue}{e_{ij}} = (\textcolor{t_red}{v_i}, \textcolor{t_darkgreen}{v_j})$ to be the edges that are incident to either $\textcolor{t_red}{v_i}$ or $\textcolor{t_darkgreen}{v_j}$.
+% chktex 25
+In 2-\acs{wl} on the other hand, the neighbors of $\textcolor{t_blue}{e_{ij}}$ are the edge pairs ${\left\{(\textcolor{t_red}{e_{il}}, \textcolor{t_darkgreen}{e_{lj}})\right\}}_{v_l \in \mathcal{V}_G}$, i.e.\ all possible paths of length two that start at $\textcolor{t_red}{v_i}$ and end at $\textcolor{t_darkgreen}{v_j}$.
+% chktex 21 chktex 25
+This difference becomes clear when comparing the definition of convolution in 2-\acsp{gnn} with that of color refinement in 2-\acs{wl} (see \cfullref[ on ]{eq:related:kgnn-layer} and \cfullref[ on ]{defn:related:wlk-refine}):
+\begin{alignat*}{4}
+	\text{2-\acs{gnn}: } && Z^{(t)}[e_{ij}] &= \sigma\Biggl( \textcolor{t_blue}{Z^{(t-1)}[e_{ij}]} W^{(t)} + &&\left( \smashoperator[r]{\sum_{v_l \in \Gamma_G(v_j)}} \textcolor{t_red}{Z^{(t-1)}[e_{il}]} + \smashoperator[lr]{\sum_{v_l \in \Gamma_G(v_i)}} \textcolor{t_darkgreen}{Z^{(t-1)}[e_{lj}]} \right) W_{\Gamma}^{(t)} \Biggr) \\ % chktex 25
+	\text{2-\acs{wl}: } && \chi_{G,2}^{(t)}(e_{ij}) &= h\Bigl(\textcolor{t_blue}{\chi_{G,2}^{(t-1)}(e_{ij})}, &&\ldblbrace (\textcolor{t_red}{\chi_{G,2}^{(t-1)}(e_{il})}, \textcolor{t_darkgreen}{\chi_{G,2}^{(t-1)}(e_{lj})})\, |\, v_l \in \mathcal{V}_G \rdblbrace\Bigr) % chktex 25
+\end{alignat*}
+We will now analyze what those different notions of neighborhood imply for the discriminative and computational power of 2-\acsp{gnn}.
+In the first step we show that the discriminative power of 2-\acsp{gnn} on all graphs $G \in \mathcal{G}$ is upper bounded by that of 1-\acs{wl} on the so-called \textit{edge neighborhood graphs} $G_{\mathcal{E}} \in \mathcal{G}_{\mathcal{E}}$.
+\begin{defn}
+	The \textit{edge neighborhood graph} of a given graph $G = (\mathcal{V}_G, \mathcal{E}_G)$ is defined as $G_{\mathcal{E}} \coloneqq (\mathcal{V}_G \cup \mathcal{E}_G, \mathcal{E}_{G_{\mathcal{E}}})$.
+Its \textit{edge neighborhood edges} are defined as
+	\begin{align*}
+		\mathcal{E}_{G_{\mathcal{E}}} \coloneqq
+		\{ s \in \{ (v, e), (e, v) \}\, |\, v \in \mathcal{V}_G \land e \in \Gamma_G(v) \} \cup \{ (e, e') \in \mathcal{E}_G^2\, |\, e \cap e' \neq \emptyset \} % chktex 21
+		\text{.}
+	\end{align*}
+\end{defn}
+\begin{prop}
+	For all 2-\acsp{gnn} $h_2$ there is a \ac{gin} $h_1$ s.t.\ $\forall G \in \mathcal{G}: h_2(G) = h_1(G_{\mathcal{E}})$.
+\end{prop}
+\begin{proof}
+	Note that the 2-\acs{gnn} convolution operation defined in \cfullref[ on ]{eq:related:kgnn-layer} can be rewritten as
+	\begin{align*}
+		Z^{(t)}[e] = \sigma \left( Z^{(t-1)}[e] W^{(t)} + \smashoperator[lr]{\sum_{e' \in \Gamma_{G_{\mathcal{E}}}(e)}} Z^{(t-1)}[e'] W_{\Gamma}^{(t)} \right)
+	\end{align*}
+\end{proof}
 

### 2020-03-28 thesis/content/chapter-related-work.tex
started with k-gnn limits section

@@
 \begin{alignat*}{2}
-	\text{\ac*{gcn}:}\quad&& Z^{(t)}[v] &= \sigma\left(\left( {\mu(v)}^2 Z^{(t-1)}[v] + \smashoperator[lr]{\sum_{u \in \Gamma_G(v)}} \mu(v) \mu(u) Z^{(t-1)}[u] \right) \Theta^{(t)} \right) \\
-	\text{1-\acs{wl}:}\quad&& \chi_{G,1}^{(t)}(v) &= h\left( \chi_{G,1}^{(t-1)}(v), \ldblbrace \chi_{G,1}^{(t-1)}(u)\, |\, u \in \Gamma_{G}(v) \rdblbrace\right)
+	\text{\ac*{gcn}:}\quad&& Z^{(t)}[v] &= \sigma\left(\left( {\mu(v)}^2 \textcolor{t_blue}{Z^{(t-1)}[v]} + \smashoperator[lr]{\sum_{u \in \Gamma_G(v)}} \mu(v) \mu(u) \textcolor{t_red}{Z^{(t-1)}[u]} \right) \Theta^{(t)} \right) \\ % chktex 25
+	\text{1-\acs{wl}:}\quad&& \chi_{G,1}^{(t)}(v) &= h\left( \mkern60mu\textcolor{t_blue}{\chi_{G,1}^{(t-1)}(v)}, \ldblbrace \textcolor{t_red}{\chi_{G,1}^{(t-1)}(u)}\, |\, u \in \Gamma_{G}(v) \rdblbrace\right) % chktex 25
 \end{alignat*}
@@
 \begin{align}
-	Z^{(t)}[s] \coloneqq \sigma\left( {Z^{(t-1)}[s]} W^{(t)} + \smashoperator[lr]{\sum_{v \in s, u \in \Gamma_G(v)}} {Z^{(t-1)}[s \setminus \{ v \} \cup \{ u \}]} W_{\Gamma}^{(t)} \right) \text{.} \label{eq:related:kgnn-layer}
+	Z^{(t)}[s] \coloneqq \sigma\left( {Z^{(t-1)}[s]} W^{(t)} + \smashoperator[lr]{\sum_{{v \in s}, {u \in \Gamma_G(v)}}} {Z^{(t-1)}[s \setminus \{ v \} \cup \{ u \}]} W_{\Gamma}^{(t)} \right) \text{.} \label{eq:related:kgnn-layer}
 \end{align}
@@
 This separation of the parameter matrix effectively fulfills the same purpose as the \ac{mlp} in \ac{gin}; it makes the implicitly defined $k$-\acs{gnn} hash function injective.
-This can be seen by realizing that a stack of $k$-\acs{gnn} convolution layers with $W_{\Gamma}^{(t)} = 0$ for all but the first layer simulates a \ac{mlp}.
+This can be seen by realizing that a stack of $k$-\acs{gnn} convolution layers with $W_{\Gamma}^{(t)} = \mathbf{0}$ for all but the first layer simulates a \ac{mlp}.
 
@@
 $k$-\acs{gnn} considers a single $k$-multiset as a neighbor while $k$-\acs{wl} uses $k$-sets of $k$-multisets as neighbors.
-We will get back to this difference later.
+We will get back to this difference later in \cref{sec:ltd:kgnn-limits}.
 
@@
 		Graph coarsening works by merging vertices in a spectrum-preserving manner~\cite{Loukas2018}.
-		This is done until a single merged vertex is left whose feature vector then represents the entire graph.
+		This is done until a single merged vertex is left whose feature vector then represents the entire graph~\cite{Ying2018}.
 	\item \textbf{Global Pooling}:

### 2020-03-29 thesis/content/chapter-ltd.tex
continued with k-gnn limits section

--- 2020-03-27 thesis/content/chapter-ltd.tex
@@
 This establishes the connection between the \acs{ltd} problem and graph convolutions on a high level.
-Then the second question is tackled in \crefrange{sec:ltd:kgnn-problems}{sec:ltd:wl2gnn-implementation}.
-There a novel graph convolution approach is proposed which could serve as a starting point for an \ac{ltd} convolution layer.
+Then the second question is tackled in \cref{sec:ltd:kgnn-limits,sec:ltd:wl2gnn}.
+There a novel graph convolution approach is proposed which could serve as a starting point for a convolution layer which solves the \ac{ltd} problem.
 
@@
 Irrespective of the chosen edge filtering strategy, the decision whether to remove a given edge $e_{ij} = (v_i, v_j)$ or not has to be made based on relevant information about $e_{ij}$ and its surroundings.
-One natural way to encode the information about $e_{ij}$ is via an edge feature vector $z_{ij} \in \mathbb{R}^d$ that is fed into the edge filter.
-We propose that those edge feature vectors could be obtained by a 2-\acs{wl} inspired graph convolution layer since 2-\acs{wl} colors represent the structural roles of edges as described in \cref{sec:related:character:wl}.
+One way to encode the information about $e_{ij}$ is via an edge feature vector $z_{ij} \in \mathbb{R}^d$ that is fed into the edge filter.
 
-\section{Shortcomings of the Existing $k$-\acs*{gnn}}%
-\label{sec:ltd:kgnn-problems}
+We propose that those edge feature vectors can be obtained via a 2-\acs{wl} inspired graph convolution layer since 2-\acs{wl} colors naturally represent the structural roles of edges as described in \cref{sec:related:character:wl}.
+We already saw a \ac{gcnn} architecture in \cref{sec:related:gcr:nn} which could potentially compute such 2-\acs{wl} inspired edge feature vectors, the $k$-\acs{gnn}~\cite{Morris2019}.
+Its variant for $k = 2$ produces edge feature vectors that can be used as the input for an edge filter.
+However, as we will see in the next section, there are significant limitations to the discriminative and computational power of 2-\acsp{gnn}.
 
-\section{The 2-\acs*{wl} Convolution Operator}%
-\label{sec:ltd:wl2gnn-definition}
+\section{Limitations of the Existing 2-\acs*{gnn}}%
+\label{sec:ltd:kgnn-limits}
 
-\section{The Expressive Power of 2-\acs*{wl}-\acsp*{gnn}}%
-\label{sec:ltd:wl2gnn-properties}
+The $k$-\acs{gnn} is a \ac{gcnn} inspired by the $k$-\ac{wl} algorithm, it convolves feature vectors of vertex $k$-multisets.
+In this section we will compare its $k = 2$ variant with the 2-\ac{wl} algorithm.
+The main difference between the two boils down to their notion of ``neighborhood''.
+As already briefly mentioned in \cref{sec:related:gcr:nn}, 2-\acsp{gnn} define the neighbors of an edge $\textcolor{t_blue}{e_{ij}} = (v_i, v_j)$ to be the edges that are incident to either $v_i$ or $v_j$.
+% chktex 25
+In 2-\acs{wl} on the other hand, the neighbors of $\textcolor{t_blue}{e_{ij}}$ are the edge pairs ${\left\{(\textcolor{t_red}{e_{il}}, \textcolor{t_darkgreen}{e_{lj}})\right\}}_{v_l \in \mathcal{V}_G}$, i.e.\ all possible paths of length two that start at $v_i$ and end at $v_j$.
+% chktex 21 chktex 25
+This difference becomes clear when comparing the definition of convolution in 2-\acsp{gnn} with that of color refinement in 2-\acs{wl} (see \cfullref[ on ]{eq:related:kgnn-layer} and \cfullref[ on ]{defn:related:wlk-refine}):
+\begin{alignat*}{4}
+	\text{2-\acs{gnn}\footnotemark: } && Z^{(t)}[e_{ij}] &= \sigma\Biggl( \textcolor{t_blue}{Z^{(t-1)}[e_{ij}]} W^{(t)} + &&\left( \smashoperator[r]{\sum_{v_l \in \Gamma_G(v_j)}} \textcolor{t_red}{Z^{(t-1)}[e_{il}]} + \smashoperator[lr]{\sum_{v_l \in \Gamma_G(v_i)}} \textcolor{t_darkgreen}{Z^{(t-1)}[e_{lj}]} \right) W_{\Gamma}^{(t)} \Biggr) \\ % chktex 25
+	\text{2-\acs{wl}: } && \chi_{G,2}^{(t)}(e_{ij}) &= h\Bigl(\textcolor{t_blue}{\chi_{G,2}^{(t-1)}(e_{ij})}, &&\ldblbrace (\textcolor{t_red}{\chi_{G,2}^{(t-1)}(e_{il})}, \textcolor{t_darkgreen}{\chi_{G,2}^{(t-1)}(e_{lj})})\, |\, v_l \in \mathcal{V}_G \rdblbrace\Bigr) % chktex 25
+\end{alignat*}\footnotetext{
+	To highlight the relation between 2-\acsp{gnn} and 2-\acs{wl}, a 2-\acs{gnn} definition that is not generally correct is shown here;
+	for self-loops with $i = j$ it incorrectly sums the feature vectors of neighboring edges twice.
+	The correct general formula uses a single sum over $v_l \in {\Gamma_G(v_j) \cup \Gamma_G(v_i)}$.
+}\begin{figure}[ht]
+	\centering
+	\includegraphics[width=\linewidth]{gfx/ltd/2gnn-2wl-diff.pdf}
+	\caption[Edge colorings on which 2-\acsp{gnn} and 2-\acs{wl} behave differently.]{
+		Two edge colorings on which 2-\acsp{gnn} and 2-\acs{wl} behave differently.
+		A 2-\acs{gnn} will refine the ``color vector'' of $e_{ij}$ to \textsf{\textbf{\small\textcolor{t_blue}{D}}} for both initial colorings.
+		2-\acs{wl} on the other hand differentiates both colorings by preserving the color tuple information.
+	}\label{fig:ltd:2gnn-2wl-diff}
+\end{figure}
 
-\section{Implementing 2-\acs*{wl}-\acsp*{gnn} on \acsp*{gpgpu}}%
-\label{sec:ltd:wl2gnn-implementation}
+We will now analyze what those different notions of neighborhood imply for the discriminative and computational power of 2-\acsp{gnn} in comparison to 2-\acs{wl}.
+In the first step we show that the discriminative power of 2-\acsp{gnn} on all graphs $G \in \mathcal{G}$ is upper bounded by that of 1-\acs{wl} on the so-called \textit{edge neighborhood graphs} $G^{\mathcal{E}} \in \mathcal{G}^{\mathcal{E}}$.
+\begin{defn}\label{defn:ltd:edge-graph}
+	The \textit{edge neighborhood graph} of a given graph $G = (\mathcal{V}_G, \mathcal{E}_G)$ is defined as $G^{\mathcal{E}} \coloneqq (\mathcal{V}_{G^{\mathcal{E}}}, \mathcal{E}_{G^{\mathcal{E}}})$ with the vertices $\mathcal{V}_{G^{\mathcal{E}}} \coloneqq \{ \ldblbrace v, u \rdblbrace\, |\, (v, u) \in \mathcal{E}_G \lor v = u \}$ and the edges $\mathcal{E}_{G^{\mathcal{E}}} \coloneqq \left\{ (e, e') \in \mathcal{V}_{\mathcal{G}^{\mathcal{E}}}^2 |\, \left|e \cap e'\right| = 1 \right\}$.
+% chktex 21
+\end{defn}
+\begin{prop}\label{prop:ltd:2gnn-wl1-limit}
+	The discriminative power of all 2-\acsp{gnn} $h_2$ is bounded by that of 1-\ac{wl} on edge neighborhood graphs, i.e.\ $\forall\, G, H \in \mathcal{G}: G^{\mathcal{E}} \mathrel{\simeq_1} H^{\mathcal{E}} \rightarrow h_2(G) = h_2(H)$.
+% chktex 21
+\end{prop}
+\begin{proof}
+	By \cref{defn:ltd:edge-graph} it holds that $\forall e_{ij} \in \mathcal{E}_G: {\Gamma_G(v_i) \cup \Gamma_G(v_j)} = \Gamma_{G^{\mathcal{E}}}(e_{ij})$.
+	Therefore the 2-\acs{gnn} convolution operation defined in \cfullref[ on ]{eq:related:kgnn-layer} can be rewritten as a vertex neighborhood convolution operator
+	\begin{align*}
+		Z^{(t)}[e] = \sigma\left( Z^{(t-1)}[e] W^{(t)} + \smashoperator[lr]{\sum_{e' \in \Gamma_{G^{\mathcal{E}}}(e)}} Z^{(t-1)}[e'] W_{\Gamma}^{(t)} \right)
+		\text{.}
+	\end{align*}
+	\Cref{prop:ltd:2gnn-wl1-limit} then follows from \cref{prop:related:gcnn-wl1-limit}~(\cpageref{prop:related:gcnn-wl1-limit}).
+\end{proof}
+\begin{lem}\label{lem:wl1-regular-edge-neighbor-limit}
+	1-\ac{wl} cannot distinguish the edge neighborhood graphs $G^{\mathcal{E}}$ and $H^{\mathcal{E}}$ of any pair of $d$-regular graphs $G$ and $H$ with $n$ vertices.
+\end{lem}
+\begin{proof}
+	Let $G$ and $H$ be two $d$-regular graphs of size $n$.
+	Their corresponding edge neighborhood graphs $G^{\mathcal{E}}$ and $H^{\mathcal{E}}$ both have $n^{\mathcal{E}} = n + \frac{nd}{2}$ vertices.
+	$n$ of those edge neighborhood vertices correspond to the vertices of $G$ and $H$ respectively, we will refer to them as \textit{loop vertices} $L_G$/$L_H$.
+	The remaining $\frac{nd}{2}$ edge neighborhood vertices correspond to the edges of $G$ and $H$, we will refer to them as \textit{edge vertices} $E_G$/$E_H$.
+
+	W.l.o.g.\ we define the initial colors of the loop vertices as $\chi^{(0)}(v) = \textcolor{t_blue}{\texttt{A}}$ for all $v \in L_G \cup L_H$. % chktex 25
+	The initial colors of the edge vertices are defined as $\chi^{(0)}(e) = \textcolor{t_red}{\texttt{B}}$ for all $e \in E_G \cup E_H$. % chktex 25
+	Note that each loop vertex $\ldblbrace v_i, v_i \rdblbrace$ with $v_i \in \mathcal{V}_G \cup \mathcal{V}_H$ has $d$ neighbors, the edges incident to $v_i$.
+	Similarly each edge vertex $\ldblbrace v_i, v_j \rdblbrace$ has $2d$ neighbors, two of which are the loop vertices $\ldblbrace v_i, v_i \rdblbrace$ and $\ldblbrace v_j, v_j \rdblbrace$ with the remaining $2d - 2$ neighbors corresponding to the edges that are incident to $e_{ij}$.
+
+	After one color refinement step we then get $\chi^{(1)}(v) = h(\textcolor{t_blue}{\texttt{A}}, \ldblbrace \underbrace{\textcolor{t_red}{\texttt{B}}, \dots, \textcolor{t_red}{\texttt{B}}}_{d\text{ times}} \rdblbrace) \eqqcolon \textcolor{t_blue}{\texttt{C}}$ for all loop vertices $v \in  L_G \cup L_H$ and $\chi^{(1)}(e) = h(\textcolor{t_red}{\texttt{B}}, \ldblbrace \textcolor{t_blue}{\texttt{A}}, \textcolor{t_blue}{\texttt{A}}, \underbrace{\textcolor{t_red}{\texttt{B}}, \dots, \textcolor{t_red}{\texttt{B}}}_{\mathclap{2d - 2 \text{ times}}} \rdblbrace) \eqqcolon \textcolor{t_red}{\texttt{D}}$.
+% chktex 25
+	This means that $\chi^{(0)}$ and $\chi^{(1)}$ are identical up to the color substitutions $\textcolor{t_blue}{\texttt{A}} \to \textcolor{t_blue}{\texttt{C}}$ and $\textcolor{t_red}{\texttt{B}} \to \textcolor{t_red}{\texttt{D}}$, i.e.\ $\chi^{(0)} \equiv \chi^{(1)}$, which in turn implies that 1-\ac{wl} terminates after one iteration.
+% chktex 25
+	\Cref{lem:wl1-regular-edge-neighbor-limit} then directly follows since both $G^{\mathcal{E}}$ and $H^{\mathcal{E}}$ have $n$ vertices with the final color $\textcolor{t_blue}{\texttt{C}}$ and $\frac{nd}{2}$ vertices with the final color $\textcolor{t_red}{\texttt{D}}$, i.e.\ $G^{\mathcal{E}} \mathrel{\simeq_1} H^{\mathcal{E}}$.
+% chktex 25
+\end{proof}
+\begin{cor}
+	A 2-\acs{gnn} cannot distinguish $d$-regular graphs and therefore has a lower discriminative power than 2-\ac{wl}.
+\end{cor}
+\begin{proof}
+	The corollary directly follows from \cref{prop:ltd:2gnn-wl1-limit}, \cref{lem:wl1-regular-edge-neighbor-limit} and the fact that 2-\acs{wl} is able to distinguish most regular graphs~\cite[cor.~1.8.6]{Immerman1990}.
+\end{proof}
+
+\section{A Novel 2-\acs*{wl}-inspired \acs*{gnn}}%
+\label{sec:ltd:wl2gnn}
+
+\subsection{Definition of the 2-\acs*{wl} Convolution Operator}%
+\label{sec:ltd:wl2gnn:definition}
+
+\subsection{Expressive Power of 2-\acs*{wl}-\acsp*{gnn}}%
+\label{sec:ltd:wl2gnn:properties}
+
+\subsection{Implementation of 2-\acs*{wl}-\acsp*{gnn} on \acsp*{gpgpu}}%
+\label{sec:ltd:wl2gnn:implementation}

### 2020-03-29 thesis/content/chapter-related-work.tex
continued with k-gnn limits section

@@
 \paragraph{A 1-\acs{wl} bound on \ac{gnn} power}
-The relation between \acp{gcnn} using neighborhood aggregation convolution and 1-\acs{wl} was formally analyzed by \citet{Xu2018}.
-They show that the discriminative power of such a \ac{gcnn} is upper-bounded by that of 1-\acs{wl}.
-In particular the \ac{gcn} architecture is strictly less powerful than 1-\acs{wl} due to the fact that $h_{\mathrm{\acs*{gcn}}}$ is not an injective hash function.
-To overcome this limitation an alternative neighborhood aggregation architecture called \ac{gin} was proposed:
+In the next step we will take a closer look at the relation between \acp{gcnn} that use vertex neighborhood convolution and 1-\acs{wl} by comparing their respective discriminative power.
+\begin{prop}[see {\citet[lemma 2]{Xu2018}} for the proof]\label{prop:related:gcnn-wl1-limit}
+	The discriminative power of a \ac{gcnn} that convolves vertex feature vectors via a convolution operator of the form $Z^{(t)}[v] = h^{(t)}\left(Z^{(t-1)}[v], \ldblbrace Z^{(t-1)}[u]\, |\, u \in \Gamma_G(v) \rdblbrace\right)$ is upper-bounded by that of 1-\acs{wl}, where $h^{(t)}: {(\mathbb{R}^{d^{(t-1)}})}^{*} \to \mathbb{R}^{d^{(t)}}$ is an arbitrary vertex neighborhood hashing function.
+	Moreover, if $h^{(t)}$ is not injective, the \ac{gcnn} has a strictly lower discriminative power than 1-\acs{wl}.
+\end{prop}
+By \cref{prop:related:gcnn-wl1-limit} the \ac{gcn} architecture is strictly less powerful than 1-\acs{wl} due to the fact that $h_{\mathrm{\acs*{gcn}}}$ is not an injective hash function.
+To overcome this limitation \citet{Xu2018} proposed an alternative vertex neighborhood convolution architecture called \ac{gin}:
 \begin{align}
@@
 }, i.e.\ it assigns a unique color vector to each multiset of color vectors.
-It was shown that this implies that \ac{gin} has the same discriminative power as 1-\ac{wl}.
+By \cref{prop:related:gcnn-wl1-limit} this implies that \ac{gin} has the same discriminative power as 1-\ac{wl}.
 As expected, \ac{gin} therefore fits training data better than non-injective \acp{gcnn} like \ac{gcn}.
@@
 \paragraph{Higher dimensional \ac{wl} \acp{gnn}}
-The majority of \ac{gnn} methods uses a neighborhood aggregation scheme which limits their discriminative and, more importantly, computational power to that of 1-\acs{wl} (see \cfullref{defn:related:wl-count-detect}).
+The majority of \ac{gnn} methods uses a vertex neighborhood convolution scheme which limits their discriminative and, more importantly, computational power to that of 1-\acs{wl} (see \cfullref{defn:related:wl-count-detect}).
 To go beyond the limits of 1-\acs{wl}, \citet{Morris2019} proposed the first \ac{gcnn} architecture inspired by the higher dimensional $k$-\acs{wl} algorithms.

### 2020-03-30 implementation/src/ltag/datasets/disk/manager.py
added 2-GNN implementation and evaluation config

@@ -44,12 +44,12 @@ class StoredGraphDatasetManager(GraphDatasetManager):
       if not wl2_dir.exists():
         os.makedirs(wl2_dir)
 
-        wl2_dataset = super()._load_wl2_dataset(neighborhood)
+      wl2_dataset = super()._load_wl2_dataset(neighborhood)
 
-        with open(wl2_dir / f"{self.name}.pickle", "wb") as f:
-          pickle.dump(wl2_dataset, f)
+      with open(wl2_dir / f"{self.name}.pickle", "wb") as f:
+        pickle.dump(wl2_dataset, f)
 
-          return wl2_dataset
+      return wl2_dataset
     else:
       with open(wl2_dir / f"{self.name}.pickle", "rb") as f:
         return pickle.load(f)
@@ -61,14 +61,31 @@ class StoredGraphDatasetManager(GraphDatasetManager):
       if not wl2_dir.exists():
         os.makedirs(wl2_dir)
 
-        wl2_dataset = super()._load_wl2c_dataset(neighborhood)
+      wl2_dataset = super()._load_wl2c_dataset(neighborhood)
 
-        with open(wl2_dir / f"{self.name}.pickle", "wb") as f:
-          pickle.dump(wl2_dataset, f)
+      with open(wl2_dir / f"{self.name}.pickle", "wb") as f:
+        pickle.dump(wl2_dataset, f)
 
-          return wl2_dataset
+      return wl2_dataset
     else:
       with open(wl2_dir / f"{self.name}.pickle", "rb") as f:
+        return pickle.load(f)
+
+  def _load_en_dataset(self):
+    en_dir = self.root_dir / "en"
+
+    if not (en_dir / f"{self.name}.pickle").exists():
+      if not en_dir.exists():
+        os.makedirs(en_dir)
+
+      en_dataset = super()._load_en_dataset()
+
+      with open(en_dir / f"{self.name}.pickle", "wb") as f:
+        pickle.dump(en_dataset, f)
+
+      return en_dataset
+    else:
+      with open(en_dir / f"{self.name}.pickle", "rb") as f:
         return pickle.load(f)
 
   def _make_splits(self):

### 2020-03-30 implementation/src/ltag/datasets/manager.py

@@ -52,6 +52,7 @@ class GraphDatasetManager:
     self._dataset = None
     self._wl2_dataset = None
     self._wl2c_dataset = None
+    self._en_dataset = None
     self._splits = None
 
     self._setup(**kwargs)
@@ -94,6 +95,22 @@ class GraphDatasetManager:
     print(f"Encoded WL2c graphs.")
 
     return wl2_graphs, targets
+
+  def _load_en_dataset(self):
+    graphs, targets = self.dataset
+
+    print(f"Encoding {self.name} as edge neighborhood graphs...")
+
+    en_graphs = np.array([
+      ds_utils.edge_neighborhood_encode(
+        g,
+        self._dim_node_features, self._dim_edge_features,
+        self._num_node_labels, self._num_edge_labels)
+      for g in graphs])
+
+    print(f"Encoded graphs as edge neighborhood graphs.")
+
+    return en_graphs, targets
 
   @property
   def full_name(self):
@@ -138,6 +155,13 @@ class GraphDatasetManager:
       self._wl2c_dataset = self._load_wl2c_dataset(self.wl2_neighborhood)
 
     return self._wl2c_dataset
+
+  @property
+  def en_dataset(self):
+    if self._en_dataset is None:
+      self._en_dataset = self._load_en_dataset()
+
+    return self._en_dataset
 
   @property
   def splits(self):
@@ -318,6 +342,30 @@ class GraphDatasetManager:
 
     return batches
 
+  def _get_en_batches(self, name, idxs=None):
+    if self.wl2_cache:
+      graphs, targets = self.en_dataset
+    else:
+      graphs, targets = self.dataset
+
+    if idxs is not None:
+      graphs = graphs[idxs]
+      targets = targets[idxs]
+
+    batches = ds_utils.to_wl2_ds(
+      graphs, targets,
+      dim_node_features=self._dim_node_features,
+      dim_edge_features=self._dim_edge_features,
+      num_node_labels=self._num_node_labels,
+      num_edge_labels=self._num_edge_labels,
+      with_indices=self.wl2_indices,
+      lazy=self.wl2_cache,  # en preencoded graphs can be batched quickly
+      en_encode=True,
+      preencoded=self.wl2_cache,
+      **self.wl2_batch_size)
+
+    return batches
+
   def _compute_gram_matrix(self, output_fn):
     return output_fn(self)
 
@@ -404,6 +452,17 @@ class GraphDatasetManager:
         return batches
 
[…]

### 2020-03-30 implementation/src/ltag/datasets/utils.py

@@ -60,6 +60,98 @@ def eid_lookup(e_ids, i, j):
 
   return e_ids[(i, j)]
 
+@cp.tolerant
+def edge_neighborhood_encode(
+  g,
+  dim_node_features=None, dim_edge_features=None,
+  num_node_labels=None, num_edge_labels=None,
+  with_indices=False):
+  """
+    Takes a graph with node and edge features and converts it
+    into an sparse edge neighborhood representation for 2-GNNs.
+  """
+
+  if dim_node_features is None:
+    dim_node_features = 0
+
+  if dim_edge_features is None:
+    dim_edge_features = 0
+
+  if num_node_labels is None:
+    dim_edge_features = 0
+
+  if num_edge_labels is None:
+    num_edge_labels = 0
+
+  dim_node_wl2 = dim_node_features + num_node_labels
+  dim_edge_wl2 = dim_edge_features + num_edge_labels
+  n_zero_f = np.zeros(dim_node_features)
+  n_zero_l = np.zeros(num_node_labels)
+  e_zero_f = np.zeros(dim_edge_features)
+  e_zero_l = np.zeros(num_edge_labels)
+  I_n = np.eye(num_node_labels)
+  I_e = np.eye(num_edge_labels)
+
+  for node, data in g.nodes(data=True):
+    g.add_edge(
+      node, node,
+      features=data.get("features", n_zero_f),
+      label=data.get("label"))
+
+  x = []
+  ref_a = []
+  backref = []
+  e_ids = {}
+
+  for i, edge in enumerate(g.edges):
+    e_ids[edge] = i
+
+  for i, edge in enumerate(g.edges):
+    a, b = edge
+    if (
+      (a != b and dim_edge_wl2 == 0)
+      or (a == b and dim_node_wl2 == 0)):
+      f = []
+      lab = []
+    else:
+      d = g.get_edge_data(a, b)
+      f = d.get("features", e_zero_f)
+      if a == b and num_node_labels > 0:
+        lab = I_n[d["label"] - 1]
+      elif a != b and num_edge_labels > 0:
+        lab = I_e[d["label"] - 1]
+      else:
+        lab = []
+
+    n_ab = {eid_lookup(e_ids, a, k) for k in nx.neighbors(g, a) if k != b}
+    if a != b:
+      n_ab |= {
+        eid_lookup(e_ids, b, k)
+        for k in nx.neighbors(g, b) if k != a}
+
+    n_ab = list(n_ab)
[…]

### 2020-03-30 implementation/src/ltag/evaluate_datasets.py

@@ -84,7 +84,8 @@ if __name__ == "__main__":
     else:
       ms = [
         "WL_st", "WL_sp", "LWL2", "GWL2",
-        "AvgCWL2GCN_Binary", "SagCWL2GCN_Binary"]
+        "AvgCWL2GCN_Binary", "SagCWL2GCN_Binary",
+        "AvgK2GNN_Binary", "SagK2GNN_Binary"]
   else:
     ms = args.model
 

### 2020-03-30 implementation/src/ltag/evaluation/datasets.py

@@ -140,24 +140,35 @@ noisy_triangle_classification_2 = fy.partial(
     }
   })
 
+balanced_eval_args = {
+  "epochs": 5000,
+  "patience": 1000,
+  "winner_repeat": 5,
+  "hp_args": {
+    # other widths ignored based on previous experiments:
+    "cwl2_layer_widths": [32],
+    "cwl2_layer_depths": [3],
+    "cwl2_local_act": "relu"
+  },
+  "ignore_worst": 2
+}
+
+balanced_triangle_classification_1 = fy.partial(
+  syn.balanced_triangle_classification_dataset(stored=True),
+  name_suffix="_n1",
+  wl2_neighborhood=1,
+  wl2_batch_size={
+    "batch_graph_count": 228
+  },
+  evaluation_args=balanced_eval_args)
+
 balanced_triangle_classification_2 = fy.partial(
   syn.balanced_triangle_classification_dataset(stored=True),
   wl2_neighborhood=2,
   wl2_batch_size={
     "batch_graph_count": 228
   },
-  evaluation_args={
-    "epochs": 5000,
-    "patience": 1000,
-    "winner_repeat": 5,
-    "hp_args": {
-      # other widths ignored based on previous experiments:
-      "cwl2_layer_widths": [32],
-      "cwl2_layer_depths": [3],
-      "cwl2_local_act": "relu"
-    },
-    "ignore_worst": 2
-  })
+  evaluation_args=balanced_eval_args)
 
 synthetic_binary = [
   "balanced_triangle_classification_2"

### 2020-03-30 implementation/src/ltag/evaluation/models.py

@@ -64,6 +64,9 @@ def AvgWL2GCN_FC_Binary_Small(dsm):
 
   return cart_merge(base, hidden_hp)
 
+
+# CWL2GCN:
+
 @binary_classifier(gnn_models.AvgCWL2GCN)
 def AvgCWL2GCN_Binary(
   dsm,
@@ -210,6 +213,77 @@ def SagCWL2GCN_Binary_quick_max(
     learning_rate=[0.001],
     squeeze_output=[True]
   )
+
+  duplicate_settings = {
+    "conv_layer_dims": ["att_conv_layer_dims"],
+    "conv_act": ["att_conv_act"],
+    "conv_local_act": ["att_conv_local_act"],
+    "conv_stack_tf": ["att_conv_stack_tf"],
+    "conv_bias": ["att_conv_bias"],
+  }
+
+  return fy.map(entry_duplicator(duplicate_settings), hps)
+
+
+# K2GNN:
+
+@binary_classifier(gnn_models.with_fc(gnn_models.AvgK2GNN))
+def AvgK2GNN_FC_Binary(
+  dsm,
+  cwl2_local_act="sigmoid",
+  cwl2_layer_widths=[32, 64],
+  cwl2_layer_depths=[1, 3],
+  cwl2_stack_tfs=[None, "keep_input"]):
+  in_dim = dsm.dim_wl2_features()
+
+  hidden = [
+    ([b] * l, [b, b])
+    for b, l in cart(cwl2_layer_widths, cwl2_layer_depths)]
+
+  hidden_hp = [dict(
+    conv_layer_dims=[in_dim, *ch],
+    fc_layer_dims=[*fh, 1]
+  ) for ch, fh in hidden]
+
+  return cart_merge(cart(
+    conv_act=["sigmoid"],
+    conv_local_act=[cwl2_local_act],
+    conv_stack_tf=cwl2_stack_tfs,
+    conv_bias=[True],
+    fc_bias=[True],
+    learning_rate=[0.01, 0.001, 0.0001],
+    squeeze_output=[True]
+  ), hidden_hp)
+
+@binary_classifier(gnn_models.with_fc(gnn_models.SagK2GNN))
+def SagK2GNN_FC_Binary(
+  dsm,
+  cwl2_local_act="sigmoid",
+  cwl2_layer_widths=[32, 64],
+  cwl2_layer_depths=[2, 4],
+  cwl2_stack_tfs=[None, "keep_input"]):
[…]

### 2020-03-30 implementation/src/ltag/layers/wl2/K2GNNLayer.py

@@ -8,44 +8,32 @@ from tensorflow.keras import regularizers as regs
 import ltag.ops as ops
 from ltag.layers.DenseLayer import DenseLayer
 
-class CWL2GCNLayer(keras.layers.Layer):
+class K2GNNLayer(keras.layers.Layer):
   def __init__(
     self, out_dim, in_dim=None,
-    act="relu", bias=True, shared_W_back=False,
+    act="relu", bias=True,
     W_regularizer=None, W_prop_regularizer=None,
-    W_back_regularizer=None,
-    b_regularizer=None, b_prop_regularizer=None,
-    local_act="relu", intersperse_dense=False, no_local_hash=False):
+    b_regularizer=None, intersperse_dense=False):
     super().__init__()
     self.out_dim = out_dim
     self.in_dim = in_dim
     self.act = keras.activations.get(act)
-    self.local_act = keras.activations.get(local_act)
     self.bias = bias
-    self.shared_W_back = shared_W_back
     self.intersperse_dense = intersperse_dense
-    self.no_local_hash = no_local_hash
     self.W_regularizer = regs.get(W_regularizer)
     self.W_prop_regularizer = regs.get(W_prop_regularizer)
-    self.W_back_regularizer = regs.get(W_back_regularizer)
     self.b_regularizer = regs.get(b_regularizer)
-    self.b_prop_regularizer = regs.get(b_prop_regularizer)
 
   def get_config(self):
     base_config = super().get_config()
     base_config["out_dim"] = self.out_dim
     base_config["in_dim"] = self.in_dim
     base_config["act"] = keras.activations.serialize(self.act)
-    base_config["local_act"] = keras.activations.serialize(self.local_act)
     base_config["bias"] = self.bias
-    base_config["shared_W_back"] = self.shared_W_back
     base_config["intersperse_dense"] = self.intersperse_dense
-    base_config["no_local_hash"] = self.no_local_hash
     base_config["W_regularizer"] = regs.serialize(self.W_regularizer)
     base_config["W_prop_regularizer"] = regs.serialize(self.W_prop_regularizer)
-    base_config["W_back_regularizer"] = regs.serialize(self.W_back_regularizer)
     base_config["b_regularizer"] = regs.serialize(self.b_regularizer)
[…]

### 2020-03-30 implementation/src/ltag/layers/wl2/__init__.py

@@ -6,6 +6,7 @@ from __future__ import absolute_import, division, print_function,\
 # Convolution:
 from ltag.layers.wl2.WL2GCNLayer import WL2GCNLayer
 from ltag.layers.wl2.CWL2GCNLayer import CWL2GCNLayer
+from ltag.layers.wl2.K2GNNLayer import K2GNNLayer
 
 # Pooling:
 from ltag.layers.wl2.AvgPooling import AvgPooling
@@ -13,6 +14,6 @@ from ltag.layers.wl2.SortPooling import SortPooling
 from ltag.layers.wl2.SagPooling import SagPooling
 
 __all__ = [
-  WL2GCNLayer, CWL2GCNLayer,
+  WL2GCNLayer, CWL2GCNLayer, K2GNNLayer,
   AvgPooling, SortPooling, SagPooling
 ]

### 2020-03-30 implementation/src/ltag/models/gnn.py

@@ -40,11 +40,22 @@ def gnn_wl2c_inputs():
 
   return X, ref_a, ref_b, backref, e_map, v_count
 
+@cm.model_inputs
+def gnn_en_inputs():
+  X = keras.Input(shape=(None,), dtype=tf.float32, name="X")
+  ref_a = keras.Input(shape=(), dtype=tf.int32, name="ref_a")
+  backref = keras.Input(shape=(), dtype=tf.int32, name="backref")
+  e_map = keras.Input(shape=(), dtype=tf.int32, name="e_map")
+  v_count = keras.Input(shape=(), dtype=tf.int32, name="v_count")
+
+  return X, ref_a, backref, e_map, v_count
+
 
 input_types = {
   "dense": gnn_dense_inputs,
   "wl2": gnn_wl2_inputs,
-  "wl2c": gnn_wl2c_inputs
+  "wl2c": gnn_wl2c_inputs,
+  "en": gnn_en_inputs
 }
 
 @cm.model_step
@@ -112,6 +123,11 @@ CWL2GCN = gnn_model("CWL2GCN", [
   select_features],
   input_type="wl2c")
 
+K2GNN = gnn_model("K2GNN", [
+  with_layers(lwl2.K2GNNLayer, prefix="conv"),
+  select_features],
+  input_type="en")
+
 # Averaging GNNs:
 
 AvgDenseGCN = DenseGCN.extend("AvgDenseGCN", [
@@ -122,6 +138,8 @@ AvgDenseWL2GCN = DenseWL2GCN.extend("AvgDenseWL2GCN", [
 AvgWL2GCN = WL2GCN.extend("AvgWL2GCN", [
   cm.with_layer(lwl2.AvgPooling, with_inputs=True)])
 AvgCWL2GCN = CWL2GCN.extend("AvgCWL2GCN", [
+  cm.with_layer(lwl2.AvgPooling, with_inputs=True)])
+AvgK2GNN = K2GNN.extend("AvgK2GNN", [
   cm.with_layer(lwl2.AvgPooling, with_inputs=True)])
 
 # Max GNNs:
@@ -144,3 +162,14 @@ SagCWL2GCN = gnn_model("SagCWL2GCN", [
   cm.merge_ios,
   cm.with_layer(lwl2.SagPooling, with_inputs=True)],
   input_type="wl2c")
+SagK2GNN = gnn_model("SagK2GNN", [
+  ([
[…]

### 2020-03-30 implementation/src/ltag/ops.py

@@ -84,6 +84,16 @@ def wl2_convolution_compact(X, ref_a, ref_b, backref, agg, with_back=False):
 
   return X_agg
 
+@tf.function
+def wl1_convolution_compact(X, ref_a, backref):
+  X_a = tf.gather(X, ref_a, axis=0)
+
+  X_shape = tf.shape(X)
+  backref = tf.expand_dims(backref, axis=-1)
+  X_agg = tf.scatter_nd(backref, X_a, shape=X_shape)
+
+  return X_agg
+
 
 # import ltag.datasets.synthetic.datasets as synthetic
 #

### 2020-03-30 implementation/src/ltag/playground.py

@@ -178,15 +178,15 @@ def imdb_experient():
   print(model.evaluate(test))
 
 def wl2_power_experiment():
-  model_class = gnn_models.AvgCWL2GCN
+  model_class = gnn_models.SagK2GNN
   # model_class = gnn_models.with_fc(model_class)
   dsm = synthetic.threesix_dataset(stored=True)(
     wl2_neighborhood=1)  # ok(3, 2, 1)
 
-  if model_class.input_type == "wl2c":
-    in_dim = dsm.dim_wl2_features()
-  else:
+  if model_class.input_type == "dense":
     in_dim = dsm.dim_dense_features()
+  else:
+    in_dim = dsm.dim_wl2_features()
 
   if in_dim == 0:
     in_dim = 1
@@ -198,7 +198,7 @@ def wl2_power_experiment():
     layer_dims=[in_dim, 4, 1],
     fc_layer_dims=[1, 2, 1],
     neighborhood_mask=1,  # ok(3, 2), nok(-1, 1)
-    bias=False)
+    bias=False, no_local_hash=True)
 
   model.compile(
     optimizer=opt,
@@ -219,16 +219,16 @@ def wl2_power_experiment():
     model.predict(dsm.get_all(output_type=model_class.input_type)))
 
 def synthetic_experiment2():
-  model_class = gnn_models.SagCWL2GCN
+  model_class = gnn_models.AvgCWL2GCN
   dsm = synthetic.balanced_triangle_classification_dataset(stored=True)(
     with_holdout=False,
-    wl2_neighborhood=2,
+    wl2_neighborhood=1,
     wl2_batch_size=dict(batch_graph_count=228))
 
-  if model_class.input_type == "wl2c":
-    in_dim = dsm.dim_wl2_features()
-  else:
+  if model_class.input_type == "dense":
     in_dim = dsm.dim_dense_features()
+  else:
+    in_dim = dsm.dim_wl2_features()
 
   if in_dim == 0:
     in_dim = 1
@@ -240,7 +240,7 @@ def synthetic_experiment2():
     squeeze_output=True,
     layer_dims=[in_dim, 32, 32, 32, 1],
     att_conv_layer_dims=[in_dim, 32, 32, 32, 1],
-    bias=True)
+    bias=True, no_local_hash=True)
 
   model.compile(
     optimizer=opt,
@@ -275,7 +275,8 @@ def kernel_experiment():
[…]

### 2020-03-30 thesis/content/chapter-ltd.tex
completed k-gnn limits section

@@
 \end{proof}
-\begin{lem}\label{lem:wl1-regular-edge-neighbor-limit}
+\begin{lem}\label{lem:ltd:wl1-regular-edge-neighbor-limit}
 	1-\ac{wl} cannot distinguish the edge neighborhood graphs $G^{\mathcal{E}}$ and $H^{\mathcal{E}}$ of any pair of $d$-regular graphs $G$ and $H$ with $n$ vertices.
@@
 	Similarly each edge vertex $\ldblbrace v_i, v_j \rdblbrace$ has $2d$ neighbors, two of which are the loop vertices $\ldblbrace v_i, v_i \rdblbrace$ and $\ldblbrace v_j, v_j \rdblbrace$ with the remaining $2d - 2$ neighbors corresponding to the edges that are incident to $e_{ij}$.
+	This is illustrated in \cref{fig:ltd:edge-neighborhood}.
 
-	After one color refinement step we then get $\chi^{(1)}(v) = h(\textcolor{t_blue}{\texttt{A}}, \ldblbrace \underbrace{\textcolor{t_red}{\texttt{B}}, \dots, \textcolor{t_red}{\texttt{B}}}_{d\text{ times}} \rdblbrace) \eqqcolon \textcolor{t_blue}{\texttt{C}}$ for all loop vertices $v \in  L_G \cup L_H$ and $\chi^{(1)}(e) = h(\textcolor{t_red}{\texttt{B}}, \ldblbrace \textcolor{t_blue}{\texttt{A}}, \textcolor{t_blue}{\texttt{A}}, \underbrace{\textcolor{t_red}{\texttt{B}}, \dots, \textcolor{t_red}{\texttt{B}}}_{\mathclap{2d - 2 \text{ times}}} \rdblbrace) \eqqcolon \textcolor{t_red}{\texttt{D}}$.
+	After one color refinement step we get $\chi^{(1)}(v) = h(\textcolor{t_blue}{\texttt{A}}, \ldblbrace \underbrace{\textcolor{t_red}{\texttt{B}}, \dots, \textcolor{t_red}{\texttt{B}}}_{d\text{ times}} \rdblbrace) \eqqcolon \textcolor{t_blue}{\texttt{C}}$ for all loop vertices $v \in  L_G \cup L_H$ and $\chi^{(1)}(e) = h(\textcolor{t_red}{\texttt{B}}, \ldblbrace \textcolor{t_blue}{\texttt{A}}, \textcolor{t_blue}{\texttt{A}}, \underbrace{\textcolor{t_red}{\texttt{B}}, \dots, \textcolor{t_red}{\texttt{B}}}_{\mathclap{2d - 2 \text{ times}}} \rdblbrace) \eqqcolon \textcolor{t_red}{\texttt{D}}$.
 % chktex 25
@@
 % chktex 25
-	\Cref{lem:wl1-regular-edge-neighbor-limit} then directly follows since both $G^{\mathcal{E}}$ and $H^{\mathcal{E}}$ have $n$ vertices with the final color $\textcolor{t_blue}{\texttt{C}}$ and $\frac{nd}{2}$ vertices with the final color $\textcolor{t_red}{\texttt{D}}$, i.e.\ $G^{\mathcal{E}} \mathrel{\simeq_1} H^{\mathcal{E}}$.
+	\Cref{lem:ltd:wl1-regular-edge-neighbor-limit} then directly follows since both $G^{\mathcal{E}}$ and $H^{\mathcal{E}}$ have $n$ vertices with the final color $\textcolor{t_blue}{\texttt{C}}$ and $\frac{nd}{2}$ vertices with the final color $\textcolor{t_red}{\texttt{D}}$, i.e.\ $G^{\mathcal{E}} \mathrel{\simeq_1} H^{\mathcal{E}}$.
 % chktex 25
 \end{proof}
-\begin{cor}
-	A 2-\acs{gnn} cannot distinguish $d$-regular graphs and therefore has a lower discriminative power than 2-\ac{wl}.
-\end{cor}
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=\linewidth]{gfx/ltd/edge-neighborhood.pdf}
+	\caption{
+		Illustration of the edge neighborhood graphs of two 2-regular graphs of size 6.
+	}\label{fig:ltd:edge-neighborhood}
+\end{figure}
+\begin{prop}\label{prop:ltd:2gnn-regular-limit}
+	A 2-\acs{gnn} cannot distinguish regular graphs of the same size and therefore has a lower discriminative power than 2-\ac{wl}.
+\end{prop}
 \begin{proof}
-	The corollary directly follows from \cref{prop:ltd:2gnn-wl1-limit}, \cref{lem:wl1-regular-edge-neighbor-limit} and the fact that 2-\acs{wl} is able to distinguish most regular graphs~\cite[cor.~1.8.6]{Immerman1990}.
+	The proposition directly follows from \cref{prop:ltd:2gnn-wl1-limit}, \cref{lem:ltd:wl1-regular-edge-neighbor-limit} and the fact that 2-\acs{wl} is able to distinguish most regular graphs~\cite[cor.~1.8.6]{Immerman1990}.
 \end{proof}
 
-\section{A Novel 2-\acs*{wl}-inspired \acs*{gnn}}%
+As described in \cref{sec:related:character:wl} the discriminative power of a model is, by itself, not necessarily relevant for real-world graph datasets.
+In \cfullref[ on ]{prop:related:wl2-cycle-count} we saw however that 2-\acs{wl} not only has a higher discriminative power than 1-\acs{wl} but also that it is able to count the number of $m$-cycles in a given graph for all $m \leq 7$.
+This is relevant because cycle counts are a commonly used metric in real-world domains such as social network and molecular structure analysis~\cite{Milo2002}\cite{Newman2003}\cite{Welser2007}\cite{Adamson1973}\cite{Kekule1866}.
+To conclude this section we now show that 2-\acsp{gnn} not only have a lower discriminative power than 2-\acs{wl} but are also unable to detect cycles.
+\begin{prop}
+	2-\acsp{gnn} cannot detect $m$-cycles for all $m \geq 3$.
+\end{prop}
+\begin{proof}
+	Let $n$ be the \ac{lcm} of $3$ and some $m > 3$.
+	We define $c_3 \coloneqq \frac{n}{3}$ and $c_m \coloneqq \frac{n}{m}$.
+	Based on that we define the following two graphs:
+	Let $G_3$ be a graph consisting of $c_3$ disconnected cycles of length $3$, analogously let $G_m$ be a graph consisting of $c_m$ disconnected cycles of length $m$.
+	Since both $G_3$ and $G_m$ are 2-regular and have the size $n$, any 2-\ac{gnn} $h_2: \mathcal{G} \to \mathcal{Y}$ must map both of them to the same $y \in \mathcal{Y}$ by \cref{prop:ltd:2gnn-regular-limit}.
+
+	Let us assume that $h_2$ is able to detect cycles of length $3$, i.e.\ triangles.
+	By \cfullref[ on ]{defn:related:wl-count-detect} this would imply that there is a function $g: \mathcal{Y} \to \{ 0, 1\}$ s.t.\ $g(h_2(G_3)) = g(y) = 1$ and $g(h_2(G_m)) = g(y) = 0\ \lightning$.
+	Conversely, assuming that $h_2$ is able to detect cycles of length $m > 3$, we obtain the contradiction $g(h_2(G_3)) = g(y) = 0$  and $g(h_2(G_m)) = g(y) = 1\ \lightning$.
+	This concludes the proof.
+\end{proof}
+
+In this section we compared 2-\acsp{gnn} with the 2-\acs{wl} algorithm and found that they have a significantly lower discriminative and computational power than 2-\acs{wl}.
+Motivated by those limitations we describe a novel graph convolution operator which is closer to 2-\acs{wl} in the following section.
+
+\section{A Novel 2-\acs*{wl} Inspired \acs*{gnn}}%
 \label{sec:ltd:wl2gnn}

### 2020-03-30 thesis/content/chapter-related-work.tex
completed k-gnn limits section

@@
 We have already seen this in \cref{fig:related:wl1-problem} since the ``six-cycle'' and the ``two-triangles'' graphs are in fact both 2-regular and of size 6.
-This restriction alone is typically not an issue in the context of \ac{gcr} though, as molecular structures or social interaction graphs for example are rarely perfectly regular.
+This restriction alone is typically not an issue in the context of \ac{gcr} though, as graphs in many real-world domains are rarely perfectly regular~\cite{Abedijaberi2018}.
 A more relevant restriction of 1-\acs{wl} is the fact that it is unable to detect cycles of length $m \geq 3$.
@@
 \begin{prop}[see \citet{Fuerer2017} and \citet{Arvind2019} for the full proof]\label{prop:related:wl2-cycle-count}
-	2-\acs{wl} is able to count $m$-cycles for $m \leq 7$ but it cannot detect 8-cycles.
+	2-\acs{wl} is able to count $m$-cycles for all $m \leq 7$ but it cannot even detect them for $m > 7$.
 \end{prop}

### 2020-03-31 thesis/content/chapter-ltd.tex
started with 2-WL-GNN definition

@@
 \begin{alignat*}{4}
-	\text{2-\acs{gnn}\footnotemark: } && Z^{(t)}[e_{ij}] &= \sigma\Biggl( \textcolor{t_blue}{Z^{(t-1)}[e_{ij}]} W^{(t)} + &&\left( \smashoperator[r]{\sum_{v_l \in \Gamma_G(v_j)}} \textcolor{t_red}{Z^{(t-1)}[e_{il}]} + \smashoperator[lr]{\sum_{v_l \in \Gamma_G(v_i)}} \textcolor{t_darkgreen}{Z^{(t-1)}[e_{lj}]} \right) W_{\Gamma}^{(t)} \Biggr) \\ % chktex 25
+	\text{2-\acs{gnn}\footnotemark: } && Z_G^{(t)}[e_{ij}] &= \sigma\Biggl( \textcolor{t_blue}{Z_G^{(t-1)}[e_{ij}]} W^{(t)} + &&\left( \smashoperator[r]{\sum_{v_l \in \Gamma_G(v_j)}} \textcolor{t_red}{Z_G^{(t-1)}[e_{il}]} + \smashoperator[lr]{\sum_{v_l \in \Gamma_G(v_i)}} \textcolor{t_darkgreen}{Z_G^{(t-1)}[e_{lj}]} \right) W_{\Gamma}^{(t)} \Biggr) \\ % chktex 25
 	\text{2-\acs{wl}: } && \chi_{G,2}^{(t)}(e_{ij}) &= h\Bigl(\textcolor{t_blue}{\chi_{G,2}^{(t-1)}(e_{ij})}, &&\ldblbrace (\textcolor{t_red}{\chi_{G,2}^{(t-1)}(e_{il})}, \textcolor{t_darkgreen}{\chi_{G,2}^{(t-1)}(e_{lj})})\, |\, v_l \in \mathcal{V}_G \rdblbrace\Bigr) % chktex 25
@@
 	\begin{align*}
-		Z^{(t)}[e] = \sigma\left( Z^{(t-1)}[e] W^{(t)} + \smashoperator[lr]{\sum_{e' \in \Gamma_{G^{\mathcal{E}}}(e)}} Z^{(t-1)}[e'] W_{\Gamma}^{(t)} \right)
+		Z_G^{(t)}[e] = \sigma\left( Z_G^{(t-1)}[e] W^{(t)} + \smashoperator[lr]{\sum_{e' \in \Gamma_{G^{\mathcal{E}}}(e)}} Z_G^{(t-1)}[e'] W_{\Gamma}^{(t)} \right)
 		\text{.}
@@
 
+We have seen that the main difference between 2-\acsp{gnn} and 2-\acs{wl} is their notion of ``neighborhood''.
+In this section we describe a convolution operator which uses edge tuple neighbors just like 2-\acs{wl} to overcome the limitations of 2-\acsp{gnn}.
+This will be done in three steps:
+\begin{enumerate*}
+	\item We begin by formally defining the 2-\acs{wl} convolution operator.
+	\item Then the discriminative and computational power of this operator are analyzed.
+	\item Lastly we describe how 2-\acs{wl} convolutions can be implemented efficiently on a modern \ac{gpgpu}.
+\end{enumerate*}
+
 \subsection{Definition of the 2-\acs*{wl} Convolution Operator}%
 \label{sec:ltd:wl2gnn:definition}
+
+Similar to the 2-LWL kernel described in \cref{sec:related:gcr:kernel}, the 2-\acs{wl} convolution operator reduces the computational cost of 2-\acs{wl} via two simplifications:
+\begin{enumerate}[label=\textbf{\arabic*.}]
+	\item \textbf{2-multisets:}
+		Since we assume that graphs are undirected, $e_{ij}$ and $e_{ji}$ have identical feature vectors $x[e_{ij}] = x[e_{ji}] \in \mathcal{X}_{\mathcal{E}}$ and the same 2-\acs{wl} neighborhood.
+		Instead of refining/convolving the feature vectors of 2-tuples $(v_i, v_j)$ we can therefore refine/convolve the feature vectors of 2-multisets $\ldblbrace v_i, v_j \rdblbrace$.
+		This halves the number of feature vectors without affecting the discriminative or computational power of the 2-\acs{wl} convolution operator.
+	\item \textbf{Neighborhood localization:}
+		2-\acs{wl} refines the color of all tuples $(v_i, v_j) \in \mathcal{V}_G^2$ by hashing its current color and the colors of all pairs ${\{ ((v_i, v_l), (v_l, v_j)) \}}_{v_l \in \mathcal{V}_G}$.
+		This means that the time complexity of a single refinement step is $\mathcal{O}(n^3)$ for $n \coloneqq \left| \mathcal{V}_G \right|$ which quickly becomes infeasible for large graphs.
+		We address this issue by reducing both the number of colored tuples as well as the number of neighbors of each tuple.
+		Through this reduction the time complexity of a refinement step becomes $\mathcal{O}(n d^k)$ where $d \coloneqq \max_{v \in \mathcal{V}_G} \left|\Gamma_G(v)\right|$ is the maximum vertex degree and $k \in \mathbb{N}$ is the freely choosable so-called \textit{neighborhood degree}.
+		We will now look at the details of this simplification.
+\end{enumerate}
+\begin{defn}
+	The \textit{$k$-th power of a graph $G$} is defined as
+	\begin{align*}
+		G^k \coloneqq \left(\mathcal{V}_G, \left\{ e_{ij} \in \mathcal{V}_G^2\, |\, d_{\mathrm{SP}, G}(v_i, v_j) \leq k \right\}\right) % chktex 21
+	\end{align*}
+	where $d_{\mathrm{SP}, G}(v_i, v_j)$ is the length of the shortest path between $v_i$ and $v_j$ in $G$.
+	The distance of a vertex $v_i \in \mathcal{V}_G$ to itself is defined as $d_{\mathrm{SP}, G}(v_i, v_i) \coloneqq 0$.
+	Note that $G^1$ does not generally equal $G$ because $G^1$ has self-loop edges $(v_i, v_i) \in \mathcal{E}_{G^1}$ for all vertices.
+\end{defn}
+\begin{defn}
+	The \textit{initial feature matrix $Z_G^{(0)}$} of the 2-\acs{wl} convolution operator with the neighborhood degree $k \in \mathbb{N}$ combines the vertex features $x[v_i] \in \mathcal{X}_{\mathcal{V}} = \mathbb{R}^{d_{\mathcal{V}}}$ and the edge features $x[e_{ij}] \in \mathcal{X}_{\mathcal{E}} = \mathbb{R}^{d_{\mathcal{E}}}$ of a given graph $G$.
+	More specifically $Z_G^{(0)} \in \mathbb{R}^{{|\mathcal{E}_{G^k}|} \times (d_{\mathcal{V}} + d_{\mathcal{E}})}$ assigns a row vector $Z_G^{(0)}[e_{ij}]$ to all edges $e_{ij} \in \mathcal{E}_{G^k}$ of $G^k$.
+	Those edge feature vectors are defined by the following vector concatenation:
+	\begin{align*}
+		Z_G^{(0)}[e_{ij}] \coloneqq \left( \begin{cases}
+			x[v_i] & \text{if $i = j$} \\
+			\mathbf{0} & \text{else}
+		\end{cases} \right) \oplus \left( \begin{cases}
+			x[e_{ij}] & \text{if $e_{ij} \in \mathcal{E}_G$} \\
+			\mathbf{0} & \text{else}
+		\end{cases} \right)
+	\end{align*}
+\end{defn}
+Using this definition, the number of convolved feature vectors can be controlled via the neighborhood degree $k$.
+For $k = 1$ only the vertices and edges of a given graph $G$ have a feature vector.
+For $k > 1$ nonexistent edges $e_{ij} \notin \mathcal{E}_G$ also get feature vectors.
+\begin{defn}
+	We define the \textit{2-\acs{wl} graph convolution operator} with the neighborhood degree $k \in \mathbb{N}$ as
+	\begin{gather*}
+		Z_G^{(t)}[e_{ij}] \coloneqq \sigma\left( Z_G^{(t-1)}[e_{ij}] W_{\mathrm{L}}^{(t)} + \smashoperator[lr]{\sum_{v_l \in {\Gamma_{G^k}(v_i) \cap \Gamma_{G^k}(v_j)}}} \kappa^{(t)}\left( Z_G^{(t-1)}[e_{ij}], \ldblbrace Z_G^{(t-1)}[e_{il}], Z_G^{(t-1)}[e_{lj}] \rdblbrace \right) \right) \\
+		\text{with } \kappa^{(t)}(z_{ij}, \ldblbrace z_{il}, z_{lj} \rdblbrace) \coloneqq \left( z_{ij} W_{\mathrm{F}}^{(t)} \right) \odot \sigma_{\Gamma}\left(\left( z_{il} + z_{lj} \right) W_{\Gamma}^{(t)} \right)
+		\text{.}
+	\end{gather*}
+	The operator is parameterized by the three matrices $W_{\mathrm{L}}^{(t)}, W_{\mathrm{F}}^{(t)},  W_{\Gamma}^{(t)} \in \mathbb{R}^{d^{(t-1)} \times d^{(t)}}$.
+\end{defn}
 

### 2020-03-31 thesis/content/chapter-related-work.tex
started with 2-WL-GNN definition

@@
 
-\begin{defn}
+\begin{defn}\label{defn:related:graph}
 	A \textit{graph} $G \coloneqq (\mathcal{V}_G, \mathcal{E}_G)$ consists of a finite set of vertices $v_i \in \mathcal{V}_G$ and a set of edges $e_{ij} = (v_i, v_j) \in \mathcal{E}_G \subseteq \mathcal{V}_G^2$.
@@
 		\begin{align}
-			\chi_{G,k}^{(t+1)}(s) \coloneqq h\left(\chi_{G,k}^{(t)}(s), \ldblbrace \{ \chi_{G,k}^{(t)}(s \setminus \{v\} \cup \{u\})\, |\, v \in s \}\, |\, u \in \mathcal{V}_G \rdblbrace\right) \text{.} \label{eq:related:wlk-set-refine} % chktex 21
+			\chi_{G,k}^{(t+1)}(s) \coloneqq h\left(\chi_{G,k}^{(t)}(s), \ldblbrace \ldblbrace \chi_{G,k}^{(t)}(s \setminus \{v\} \cup \{u\})\, |\, v \in s \rdblbrace\, |\, u \in \mathcal{V}_G \rdblbrace\right) \text{.} \label{eq:related:wlk-set-refine} % chktex 21
 		\end{align}
@@
 A spectral \ac{gcnn} expects a graph $G$ with real vertex feature vectors $x_G[v_i] \in \mathbb{R}^d$ as its input.
-Those feature vectors are typically one-hot encodings of labels or, if no such information is provided, vertex embedding vectors (see \cref{sec:related:gcr:embed}).
-We call $X_G \coloneqq \begin{pmatrix} x_G[v_1] \\ \vdots \\ x_G[v_n] \end{pmatrix} \in \mathbb{R}^{n \times d}$ the vertex feature matrix of $G$.
+Those feature vectors often are one-hot encodings of the labels $l_G[v_i]$ or, if no such information is provided, vertex embedding vectors (see \cref{sec:related:gcr:embed}).
+\begin{defn}\label{defn:related:vertex-feature-matrix}
+		We call $X_G \coloneqq \begin{pmatrix} x_G[v_1] \\ \vdots \\ x_G[v_n] \end{pmatrix} \in \mathbb{R}^{n \times d}$ the \textit{vertex feature matrix} of $G$.
+\end{defn}
 Note that each of the $d$ columns of $X_G$ can be interpreted as a separate graph signal function $x_{G,j}: \mathcal{V}_G \to \mathbb{R}$ for $j \in [d]$.
@@
 \begin{align}
-	Z^{(t)} \coloneqq \sigma\left( \hat{A} Z^{(t-1)} \Theta^{(t)} \right)
-	\quad\text{with } \hat{A} \coloneqq \tilde{D}_G^{-\frac{1}{2}} \tilde{A}_G \tilde{D}_G^{-\frac{1}{2}}\quad\text{and}\quad Z^{(0)} \coloneqq X_G \text{.} \label{eq:related:gcn-layer}
+	Z_G^{(t)} \coloneqq \sigma\left( \hat{A} Z_G^{(t-1)} \Theta^{(t)} \right)
+	\quad\text{with } \hat{A} \coloneqq \tilde{D}_G^{-\frac{1}{2}} \tilde{A}_G \tilde{D}_G^{-\frac{1}{2}}\quad\text{and}\quad Z_G^{(0)} \coloneqq X_G \text{.} \label{eq:related:gcn-layer}
 \end{align}
 Here $\sigma$ is some non-linearity, e.g.\ ReLU, and $\Theta^{(t)} \in \mathbb{R}^{d^{(t-1)} \times d^{(t)}}$ is a matrix of learned filter parameters.
-This type of convolutional layer takes a graph signal $Z^{(t-1)} \in \mathbb{R}^{n \times d^{(t-1)}}$ and the corresponding adjacency matrix $A_G \in \mathbb{R}^{n \times n}$ as input and outputs a convolved signal $Z^{(t)} \in \mathbb{R}^{n \times d^{(t)}}$.
+This type of convolutional layer takes a graph signal $Z_G^{(t-1)} \in \mathbb{R}^{n \times d^{(t-1)}}$ and the corresponding adjacency matrix $A_G \in \mathbb{R}^{n \times n}$ as input and outputs a convolved signal $Z_G^{(t)} \in \mathbb{R}^{n \times d^{(t)}}$.
 By stacking multiple of those convolutional layers, complex non-linear signal filters can be learned despite the linearity of the underlying spectral filter.
@@
 \begin{alignat*}{2}
-	\text{\ac*{gcn}:}\quad&& Z^{(t)}[v] &= \sigma\left(\left( {\mu(v)}^2 \textcolor{t_blue}{Z^{(t-1)}[v]} + \smashoperator[lr]{\sum_{u \in \Gamma_G(v)}} \mu(v) \mu(u) \textcolor{t_red}{Z^{(t-1)}[u]} \right) \Theta^{(t)} \right) \\ % chktex 25
+	\text{\ac*{gcn}:}\quad&& Z_G^{(t)}[v] &= \sigma\left(\left( {\mu(v)}^2 \textcolor{t_blue}{Z_G^{(t-1)}[v]} + \smashoperator[lr]{\sum_{u \in \Gamma_G(v)}} \mu(v) \mu(u) \textcolor{t_red}{Z_G^{(t-1)}[u]} \right) \Theta^{(t)} \right) \\ % chktex 25
 	\text{1-\acs{wl}:}\quad&& \chi_{G,1}^{(t)}(v) &= h\left( \mkern60mu\textcolor{t_blue}{\chi_{G,1}^{(t-1)}(v)}, \ldblbrace \textcolor{t_red}{\chi_{G,1}^{(t-1)}(u)}\, |\, u \in \Gamma_{G}(v) \rdblbrace\right) % chktex 25
@@
 Here $\mu(v) \coloneqq {\left(\left| \Gamma_G(v) \right| + 1\right)}^{-\frac{1}{2}}$ are normalization factors introduced by $\tilde{D}_G^{-\frac{1}{2}}$.
-The comparison shows that \ac{gcn} convolutions can be understood as iterative refinements of continuous ``color vectors'' $Z^{(t)}[v] \in \mathbb{R}^{d^{(t)}}$.
+The comparison shows that \ac{gcn} convolutions can be understood as iterative refinements of continuous ``color vectors'' $Z_G^{(t)}[v] \in \mathbb{R}^{d^{(t)}}$.
 Then the continuous analogue to the injective 1-\acs{wl} hash function $h: \mathcal{C}^{*} \to \mathcal{C}$ can be defined as $h_{\mathrm{\acs*{gcn}}}(z_0, z_1, \dots, z_m) = \sigma(\sum_{j=0}^m \mu(v_0) \mu(v_j) z_j \Theta^{(t)})$ which computes a single color vector from multiple color vectors $z_j$ where $v_j$ are the corresponding vertices.
@@
 \begin{prop}[see {\citet[lemma 2]{Xu2018}} for the proof]\label{prop:related:gcnn-wl1-limit}
-	The discriminative power of a \ac{gcnn} that convolves vertex feature vectors via a convolution operator of the form $Z^{(t)}[v] = h^{(t)}\left(Z^{(t-1)}[v], \ldblbrace Z^{(t-1)}[u]\, |\, u \in \Gamma_G(v) \rdblbrace\right)$ is upper-bounded by that of 1-\acs{wl}, where $h^{(t)}: {(\mathbb{R}^{d^{(t-1)}})}^{*} \to \mathbb{R}^{d^{(t)}}$ is an arbitrary vertex neighborhood hashing function.
+	The discriminative power of a \ac{gcnn} that convolves vertex feature vectors via a convolution operator of the form $Z_G^{(t)}[v] = h^{(t)}\left(Z_G^{(t-1)}[v], \ldblbrace Z_G^{(t-1)}[u]\, |\, u \in \Gamma_G(v) \rdblbrace\right)$ is upper-bounded by that of 1-\acs{wl}, where $h^{(t)}: {(\mathbb{R}^{d^{(t-1)}})}^{*} \to \mathbb{R}^{d^{(t)}}$ is an arbitrary vertex neighborhood hashing function.
 	Moreover, if $h^{(t)}$ is not injective, the \ac{gcnn} has a strictly lower discriminative power than 1-\acs{wl}.
@@
 \begin{align}
-	Z^{(t)}[v] \coloneqq \mathrm{\acs*{mlp}}^{(t)}\left( Z^{(t-1)}[v] + \smashoperator[lr]{\sum_{u \in \Gamma_G(v)}} Z^{(t-1)}[u] \right)
+	Z_G^{(t)}[v] \coloneqq \mathrm{\acs*{mlp}}^{(t)}\left( Z_G^{(t-1)}[v] + \smashoperator[lr]{\sum_{u \in \Gamma_G(v)}} Z_G^{(t-1)}[u] \right)
 \end{align}
@@
 \begin{align}
-	Z^{(t)}[s] \coloneqq \sigma\left( {Z^{(t-1)}[s]} W^{(t)} + \smashoperator[lr]{\sum_{{v \in s}, {u \in \Gamma_G(v)}}} {Z^{(t-1)}[s \setminus \{ v \} \cup \{ u \}]} W_{\Gamma}^{(t)} \right) \text{.} \label{eq:related:kgnn-layer}
+	Z_G^{(t)}[s] \coloneqq \sigma\left( {Z_G^{(t-1)}[s]} W^{(t)} + \smashoperator[lr]{\sum_{{v \in s}, {u \in \Gamma_G(v)}}} {Z_G^{(t-1)}[s \setminus \{ v \} \cup \{ u \}]} W_{\Gamma}^{(t)} \right) \text{.} \label{eq:related:kgnn-layer}
 \end{align}
@@
 
-The graph convolution approaches that we just looked at all produce a set of convolved feature vectors ${\left\{ Z^{(T)}[s_i] \right\}}_{i=1}^{m}$ after $T$ convolutional layers. % chktex 21
+The graph convolution approaches that we just looked at all produce a set of convolved feature vectors ${\left\{ Z_G^{(T)}[s_i] \right\}}_{i=1}^{m}$ after $T$ convolutional layers. % chktex 21
 For 1-\acs{wl} bounded architectures like \ac{gcn} or \ac{gin}, there is one vector for each of the $m = {|\mathcal{V}_G|}$ vertices, while a $k$-\acs{wl} inspired architecture like $k$-\acs{gnn} produces $m \leq {|\mathcal{V}_G|}^{k}$ output vectors.

### 2020-04-01 thesis/content/chapter-ltd.tex
extended 2-wl-gnn definition

--- 2020-03-30 thesis/content/chapter-ltd.tex
@@
 
+We have seen that the main difference between 2-\acsp{gnn} and 2-\acs{wl} is their notion of ``neighborhood''.
+In this section we describe a novel convolution operator which uses edge tuple neighbors just like 2-\acs{wl} to overcome the limitations of 2-\acsp{gnn}.
+This will be done in three steps:
+\begin{enumerate*}
+	\item We begin by formally defining the 2-\acs{wl} convolution operator.
+	\item Then the discriminative and computational power of this operator are analyzed.
+	\item Lastly we describe how 2-\acs{wl} convolutions can be implemented efficiently on a modern \ac{gpgpu}.
+\end{enumerate*}
+
 \subsection{Definition of the 2-\acs*{wl} Convolution Operator}%
 \label{sec:ltd:wl2gnn:definition}
+
+Similar to the 2-LWL kernel described in \cref{sec:related:gcr:kernel}, our 2-\acs{wl} convolution operator reduces the computational cost of 2-\acs{wl} via two simplifications:
+\begin{enumerate}[label=\textbf{\arabic*.}]
+	\item \textbf{2-multisets:}
+		Since we assume that graphs are undirected, $e_{ij}$ and $e_{ji}$ have identical feature vectors $x[e_{ij}] = x[e_{ji}] \in \mathcal{X}_{\mathcal{E}}$ and the same 2-\acs{wl} neighborhood.
+		Instead of refining/convolving the feature vectors of 2-tuples $(v_i, v_j)$ we can therefore refine/convolve the feature vectors of 2-multisets $\ldblbrace v_i, v_j \rdblbrace$.
+		This halves the number of feature vectors without affecting the discriminative or computational power of the 2-\acs{wl} convolution operator.
+		To simplify the notation we assume that $e_{ij} = e_{ji} = \ldblbrace v_i, v_j \rdblbrace$ in the rest of this section.
+	\item \textbf{Neighborhood localization:}
+		Using the 2-multiset simplification, the original 2-\acs{wl} algorithm refines the color of all multisets $e_{ij} \in \mathcal{V}_G^2$ by hashing its current color and the colors of all neighbors ${\{ \ldblbrace e_{il}, e_{lj} \rdblbrace \}}_{v_l \in \mathcal{V}_G}$.
+		This means that the time complexity of a single refinement step is $\mathcal{O}(n^3)$ for $n \coloneqq \left| \mathcal{V}_G \right|$ which quickly becomes infeasible for large graphs.
+
+		To address this issue we reduce both the number of colored edges as well as the number of neighbors of each edge.
+		This is achieved by only considering the edges that are part of the so-called \textit{$r$-th power of a given graph $G$} where $r \in \mathbb{N}$ is the freely choosable \textit{neighborhood radius}.
+		\begin{defn}
+			The \textit{$r$-th power of a graph $G$} is defined as
+			\begin{align*}
+				G^r \coloneqq \left(\mathcal{V}_G, \left\{ e_{ij} \in \mathcal{V}_G^2\, |\, d_{\mathrm{SP}, G}(v_i, v_j) \leq r \right\}\right) % chktex 21
+			\end{align*}
+			where $d_{\mathrm{SP}, G}(v_i, v_j)$ is the length of the shortest path between $v_i$ and $v_j$ in $G$.
+			The distance of a vertex $v_i \in \mathcal{V}_G$ to itself is defined as $d_{\mathrm{SP}, G}(v_i, v_i) \coloneqq 0$.
+			Note that $G^1$ does not generally equal $G$ because $G^1$ has self-loop edges $e_{ii} \in \mathcal{E}_{G^1}$ at all vertices.
+		\end{defn}
+		For the neighborhood radius $r = 1$ only the self-loop edges ${\{ e_{ii} \}}_{v_i \in \mathcal{V}_G}$ and the edges $\mathcal{E}_{G}$ are considered;
+		for $r > 1$ edges between indirectly connected vertices are considered as well.
+		\Cref{fig:ltd:neighborhood-radius} illustrates this for varying neighborhood radii.
+		Through the reduction of the considered edges, the neighbors of each $e_{ij} \in \mathcal{E}_{G^r}$ are in turn reduced to the common $r$-neighbors of $v_i$ and $v_j$, i.e.\ $\{ \ldblbrace e_{il}, e_{lj} \rdblbrace\, |\, v_l \in {\Gamma_{G^r}(v_i) \cap \Gamma_{G^r}(v_j)} \}$.
+% chktex 21
+		\begin{figure}[t]
+			\centering
+			\includegraphics[width=0.9\linewidth]{gfx/ltd/neighborhood-radius.pdf}
+			\caption{
+				Illustration of the powers of the six-cycle graph for varying $r$.
+				The self-loop edge at each of the vertices is not explicitly shown.
+				For $r = 3$ all possible edges between the six vertices will be considered just as in the original 2-\acs{wl} algorithm.
+			}\label{fig:ltd:neighborhood-radius}
+		\end{figure}
+
+		Let us now consider what the reduced number of considered edges and the reduced number of edge neighbors implies for the runtime of a refinement step.
+		If $G$ is a sparse graph with the maximum vertex degree $d \coloneqq \max_{v \in \mathcal{V}_G} \left|\Gamma_G(v)\right|$, the number of considered edges is bounded by $\mathcal{O}(n d^r)$ where each edge has at most $\mathcal{O}(d^r)$ neighbors.
+		Consequently the time complexity of a refinement step becomes $\mathcal{O}(n d^{2r})$ which is a significant improvement over the $\mathcal{O}(n^3)$ bound of a full 2-\acs{wl} refinement step (assuming $d \ll n$).
+\end{enumerate}
+Based on the 2-multiset and the neighborhood localization simplifications we can now define the 2-\ac{wl} convolution operator.
+\begin{defn}\label{defn:ltd:wl2-conv-init}
+	The \textit{initial feature matrix $Z_G^{(0)}$} of the 2-\acs{wl} convolution operator with the neighborhood radius $r \in \mathbb{N}$ contains both the vertex features $x[v_i] \in \mathcal{X}_{\mathcal{V}} = \mathbb{R}^{d_{\mathcal{V}}}$ as well as the edge features $x[e_{ij}] \in \mathcal{X}_{\mathcal{E}} = \mathbb{R}^{d_{\mathcal{E}}}$ of a given graph $G$.
+	More specifically $Z_G^{(0)} \in \mathbb{R}^{{|\mathcal{E}_{G^r}|} \times (d_{\mathcal{V}} + d_{\mathcal{E}})}$ assigns a row vector $Z_G^{(0)}[e_{ij}]$ to all edges $e_{ij} \in \mathcal{E}_{G^r}$.
+	Those initial edge feature vectors are defined by the following vector concatenation:
+	\begin{align*}
+		Z_G^{(0)}[e_{ij}] \coloneqq \left( \begin{cases}
+			x[v_i] & \text{if $i = j$} \\
+			\mathbf{0} & \text{else}
+		\end{cases} \right) \oplus \left( \begin{cases}
+			x[e_{ij}] & \text{if $e_{ij} \in \mathcal{E}_G$} \\
+			\mathbf{0} & \text{else}
+		\end{cases} \right)
+	\end{align*}
+\end{defn}
+\begin{defn}\label{defn:ltd:wl2-conv-step}
+	We define the \textit{2-\acs{wl} graph convolution operator} as
+	\begin{gather*}
+		Z_G^{(t)}[e_{ij}] \coloneqq \sigma\left( Z_G^{(t-1)}[e_{ij}] W_{\mathrm{L}}^{(t)} + \smashoperator[lr]{\sum_{v_l \in {\Gamma_{G^r}(v_i) \cap \Gamma_{G^r}(v_j)}}} \kappa^{(t)}\left( Z_G^{(t-1)}[e_{ij}], \ldblbrace Z_G^{(t-1)}[e_{il}], Z_G^{(t-1)}[e_{lj}] \rdblbrace \right) \right) \\
+		\text{with } \kappa^{(t)}(z_{ij}, \ldblbrace z_{il}, z_{lj} \rdblbrace) \coloneqq \left( z_{ij} W_{\mathrm{F}}^{(t)} \right) \odot \sigma_{\Gamma}\left(\left( z_{il} + z_{lj} \right) W_{\Gamma}^{(t)} \right)
+		\text{.}
+	\end{gather*}
+\end{defn}
+The convolution operator from \cref{defn:ltd:wl2-conv-step} is parameterized by the three matrices $W_{\mathrm{L}}^{(t)}, W_{\mathrm{F}}^{(t)},  W_{\Gamma}^{(t)} \in \mathbb{R}^{d^{(t-1)} \times d^{(t)}}$ and uses two freely choosable nonlinearities $\sigma$ and $\sigma_{\Gamma}$.
+There are three properties which motivate this particular choice of convolution layer:
+\begin{enumerate}[label=\textbf{\arabic*.}]
+	\item \textbf{Simulation of \acp{mlp}:}
+		By choosing $W_{F}^{(t)} = \mathbf{0}$ the convolution behaves like a regular fully connected layer.
+		A stack of 2-\acs{wl} convolution layers can therefore simulate arbitrary \acp{mlp}.
+	\item \textbf{Perservation of edge pair information:}
+		The 2-\acs{wl} convolution layer computes a feature vector for each neighbor $\ldblbrace e_{il}, e_{lj} \rdblbrace$ of $e_{ij}$ via $\sigma_{\Gamma}\left(\left( z_{il} + z_{lj} \right) W_{\Gamma}^{(t)} \right)$.
+		A 2-\acs{gnn} use a similar formulation but leaves out the inner nonlinearity $\sigma_{\Gamma}$;
+		as we saw in \cref{fig:ltd:2gnn-2wl-diff} this causes 2-\acsp{gnn} to lose the edge pair information due to the commutativity and associativity of $+$.
+	\item \textbf{Context-dependent neighborhood filtering:}
+		Instead of filtering out all neighbors, $W_{F}^{(t)}$ also allows to filter the neighborhood of $e_{ij}$ more selectively.
+		We can interpret $z_{ij} W_{\mathrm{F}}^{(t)} \in \mathbb{R}^{d^{(t)}}$ as a row vector of feature dimension weights.
+		The feature dimensions of neighbors are rescaled via those weights which allows the model to aggregate different types of neighbors depending on $z_{ij}$.
+\end{enumerate}
 
@@
 
+In this section we will analyze the discriminative and computational power of the 2-\ac{wl} convolution operator that was just defined.
+We begin by showing that 2-\acs{wl}-\acsp{gnn} are at least as powerful as 1-\acs{wl}.
+
+\begin{prop}
+	2-\acs{wl}-\acsp{gnn} can simulate arbitrary \acp{gin}.
+\end{prop}
+
 \subsection{Implementation of 2-\acs*{wl}-\acsp*{gnn} on \acsp*{gpgpu}}%

### 2020-04-02 thesis/content/chapter-ltd.tex
started with 2-wl-gnn expressive power

@@
 
-\begin{prop}
-	2-\acs{wl}-\acsp{gnn} can simulate arbitrary \acp{gin}.
-\end{prop}
+\begin{thm}
+	A stack of 2-\acs{wl} convolution layers can simulate a stack of \ac{gin} convolution layers as defined in \cfullref[ on ]{eq:related:gin-layer}, i.e.\ all vertices $v_i \in \mathcal{V}_G$ of a given graph $G$ have the same feature vectors after applying the \ac{gin} convolutions as the self-loops $e_{ii}$ after applying the 2-\acs{wl} convolutions.
+\end{thm}
+\begin{proof}
+	A stack of \ac{gin} convolution layers can be described as a sequence of neighborhood aggregations of the form $Z_G^{(t)}[v_i] = (1 + \varepsilon) Z_G^{(t-1)}[v_i] + \sum_{v_j \in \Gamma_G(v_i)} Z_G^{(t-1)}[v_j]$ with fully connected layers interspersed between them\footnote{
+		The $(1 + \varepsilon) \in \mathbb{R}$ factor is required to guarantee the injectivity and therefore 1-\acs{wl} equivalence of \acp{gin}~\cite{Xu2018}.
+		In our previous description of \ac{gin} we assumed that $\varepsilon = 0$ for simplicity.
+		However for the formal correctness of this proof we also have to account for the case of $\varepsilon > 0$.
+	}.
+	As already mentioned, the 2-\acs{wl} convolution layer is reduced to a fully connected layer if $W_{F}^{(t)} = \mathbf{0}$.
+	To show that 2-\acs{wl} convolutions can simulate \ac{gin}, it is therefore sufficient to show that for the $t$-th \ac{gin} layer there is a stack $S^{(t)}$ of 2-\ac{wl} convolution layers s.t.\ %
+	\begin{align*}
+		\exists P^{(t)}: \forall v_i \in \mathcal{V}_G: S^{(t)}\left(Z_{G}^{(t-1)}\right)[e_{ii}] P^{(t)} = (1 + \varepsilon) Z_{G}^{(t-1)}[e_{ii}] P^{(t)} + \smashoperator[lr]{\sum_{v_l \in \Gamma_G(v_i)}} Z_{G}^{(t-1)}[e_{ll}] P^{(t)}
+		\text{.}
+	\end{align*}
+	Because \ac{gin}'s neighborhood aggregation is described by a sum of vertex feature vectors, the individual vertex feature dimensions can be treated independently.
+	W.l.o.g.\ we will therefore assume that each vertex only has a single scalar feature to simplify notation.
+
+	We now inductively construct the set of 2-\acs{wl} stacks ${\{ S^{(t)} \}}_{t=1}^T$. % chktex 21
+	A neighborhood radius of $r = 1$ will be sufficient to simulate the \ac{gin} aggregation.
+Therefore each convolved feature matrix $Z_G^{(t)}$ contains $m \coloneqq \left|\mathcal{E}_{G^1}\right| = \left|\mathcal{V}_G\right| + \left|\mathcal{E}_G\right|$ rows, one for each vertex/self-loop and one for each edge of the input graph $G$.
+	Based on the assumption that each vertex has one-dimensional features, the initial feature matrix $Z_G^{(0)}$ can be described by the following row vectors:
+	\begin{align*}
+		\forall v_i \in \mathcal{V}_G: Z_G^{(0)}[e_{ii}] \coloneqq (1, x[v_i], 0)
+		\quad\text{and}\quad
+		\forall e_{ij} \in \mathcal{E}_G: Z_G^{(0)}[e_{ij}] \coloneqq (0, 0, 1)
+		\text{.}
+	\end{align*}
+	The first stack $S^{(1)} \coloneqq S^{(1, 2)} \circ S^{(1, 1)}$ takes those initial feature vector and performs a neighborhood aggregation by applying two 2-\acs{wl} convolution layers.
+	The first 2-\acs{wl} layer $S^{(1, 1)}: \mathbb{R}^{m \times 3} \to \mathbb{R}^{m \times 4}$ is defined as
+	\begin{gather*}
+		\begin{aligned}
+			S^{(1, 1)}\left(Z^{(0)}\right)[e_{ij}] &= \textcolor{t_blue}{Z^{(0)}[e_{ij}] W_L^{(1,1)}} + \smashoperator[lr]{\sum_{v_l \in {\Gamma_{G^1}(v_i) \cap \Gamma_{G^1}(v_j)}}} \textcolor{t_red}{\left( Z^{(0)}[e_{ij}] W_F^{(1,1)}\right)} \odot \textcolor{t_darkgreen}{\left( \left( Z^{(0)}[e_{il}] + Z^{(0)}[e_{lj}] \right) W_{\Gamma}^{(1,1)} \right)} \\ % chktex 25
+			&= \begin{cases}
+				\textcolor{t_blue}{(1, \varepsilon x[v_i], 0, 0)} - \left|\Gamma_{G^1}(v_i)\right| \textcolor{t_red}{(0, x[v_i], 0, 0)} & \text{if $i = j$} \\ % chktex 25
+				\textcolor{t_blue}{(0, 0, 1, 0)} + \textcolor{t_darkgreen}{(0, 0, 0, x[v_i])} + \textcolor{t_darkgreen}{(0, 0, 0, x[v_j])} & \text{else} % chktex 25
+			\end{cases} \\
+			&= \begin{cases}
+				\left(1, (\varepsilon - \left|\Gamma_{G^1}(v_i)\right|) x[v_i], 0, 0\right) & \text{if $i = j$} \\
+				(0, 0, 1, x[v_i] + x[v_j]) & \text{else}
+			\end{cases}
+		\end{aligned}\\
+		\text{with }
+		W_L^{(1,1)} \coloneqq \begin{pmatrix}
+			1 & 0 & 0 & 0 \\
+			0 & \varepsilon & 0 & 0 \\
+			0 & 0 & 1 & 0
+		\end{pmatrix},
+		W_F^{(1,1)} \coloneqq \begin{pmatrix}
+			0 & 0 & 0 & 0 \\
+			0 & -\frac{1}{2} & 0 & 0 \\
+			0 & 0 & 0 & 1
+		\end{pmatrix},
+		W_{\Gamma}^{(1,1)} \coloneqq \begin{pmatrix}
+			0 & 1 & 0 & 0 \\
+			0 & 0 & 0 & 1 \\
+			0 & 1 & 0 & 0
+		\end{pmatrix}
+		\text{.}
+	\end{gather*}
+	The aggregation is then completed in the second layer $S^{(1, 2)}: \mathbb{R}^{m \times 4} \to \mathbb{R}^{m \times 3}$:
+	\begin{gather*}
+		\begin{aligned}
+			S^{(1)}\left(Z^{(0)}\right)[e_{ij}] &= \begin{cases}
+				\textcolor{t_blue}{\left(1, (\varepsilon - \left|\Gamma_{G^1}(v_i)\right|) x[v_i], 0\right)} + \displaystyle\smashoperator[lr]{\sum_{v_l \in \Gamma_{G^1}(v_i)}} \textcolor{t_darkgreen}{(0, x[v_i] + x[v_l], 0)} & \text{if $i = j$} \\ % chktex 25
+				\textcolor{t_blue}{(0, 0, 1)} & \text{else} % chktex 25
+			\end{cases} \\
+			&= \begin{cases}
+				\left(1, (1 + \varepsilon) x[v_i] + \displaystyle\smashoperator[lr]{\sum_{v_l \in \Gamma_{G}(v_i)}} x[v_l], 0\right) & \text{if $i = j$} \\
+				(0, 0, 1) & \text{else}
+			\end{cases}
+		\end{aligned}\\
+		\text{with }
+		W_L^{(1,2)} \coloneqq \begin{pmatrix}
+			1 & 0 & 0 \\
+			0 & 1 & 0 \\
+			0 & 0 & 1 \\
+			0 & 0 & 0
+		\end{pmatrix},\quad
+		W_F^{(1,2)} \coloneqq \begin{pmatrix}
+			0 & \frac{1}{2} & 0 \\
+			0 & 0 & 0 \\
+			0 & 0 & 0 \\
+			0 & 0 & 0
+		\end{pmatrix},\quad
+		W_{\Gamma}^{(1,2)} \coloneqq \begin{pmatrix}
+			0 & 0 & 0 \\
+			0 & 0 & 0 \\
+			0 & 0 & 0 \\
+			0 & 1 & 0
+		\end{pmatrix}
+		\text{.}
+	\end{gather*}
+\end{proof}
 

### 2020-04-03 thesis/content/chapter-graph-lta.tex
completed proof that 2-wl-gnns are at least as powerful as 1-wl

@@
 	\begin{enumerate*}
-		\item there is an associative algebra of $\mathit{Pool}$,
+		\item there is an associative algebra of $\mathit{Pool}$ and
 		\item there is a weighted aggregation function $\mathcal{A}: {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*} \to \mathcal{Y}$ with an associative algebra and
@@
 Note that \cref{thm:ltag:gcnn-ltag-formulation} requires \acp{gcnn} without a final \ac{mlp} which is unlike the architecture proposed in the \ac{gcn}~\cite{Kipf2017} and \ac{gin}~\cite{Xu2018} papers.
-The reason for this is that the composition $\mathrm{\acs*{mlp}} \circ \mathit{Pool}$ cannot be guaranteed to be homomorphic to a weighted aggregation function $\mathcal{A}$ due to the universal approximation theorem.
+The reason for this is that the composition $\mathrm{\acs*{mlp}} \circ \mathit{Pool}$ cannot be guaranteed to be homomorphic to a weighted aggregation function $\mathcal{A}$ due to the well-known universal approximation theorem~\cite{Hornik1991}.
 

### 2020-04-03 thesis/content/chapter-ltd.tex
completed proof that 2-wl-gnns are at least as powerful as 1-wl

--- 2020-04-01 thesis/content/chapter-ltd.tex
@@
 			\includegraphics[width=0.9\linewidth]{gfx/ltd/neighborhood-radius.pdf}
-			\caption{
+			\caption[Illustration of the powers of the six-cycle graph.]{
 				Illustration of the powers of the six-cycle graph for varying $r$.
@@
 
-\begin{prop}
-	2-\acs{wl}-\acsp{gnn} can simulate arbitrary \acp{gin}.
-\end{prop}
+\begin{defn}\label{defn:ltd:vert-conv}
+	A \ac{gnn} $h_1: \mathcal{G} \to \mathcal{Y}$ uses \textit{weighted vertex neighborhood sums} if its convolutional layers can be described by
+	\begin{align*}
+		Z_{G}^{(t)}[v_i] = \mathrm{\acs*{mlp}}^{(t)}\left( w_{ii} Z_{G}^{(t-1)}[v_i] + \smashoperator[lr]{\sum_{v_j \in \Gamma_G(v_i)}} w_{ij} Z_{G}^{(t-1)}[v_l] \right)
+		\text{.}
+	\end{align*}
+\end{defn}
+\Cref{defn:ltd:vert-conv} includes \acp{gcn}~\cite{Kipf2017} where the \ac{mlp} only consists of a single layer and the weights are $w_{ij} = {\left(\left| \Gamma_G(v_i) \right| + 1\right)}^{-\frac{1}{2}}{\left(\left| \Gamma_G(v_j) \right| + 1\right)}^{-\frac{1}{2}}$ (see \cfullref{eq:related:gcn-wl1}).
+\Acp{gin}~\cite{Xu2018} also trivially satisfy the definition (see \cfullref{eq:related:gin-layer}).
+\begin{thm}\label{thm:ltd:wl2-simulation}
+	For each \ac{gnn} $h_1$ that uses weighted vertex neighborhood sums, there is a 2-\acs{wl}-\acs{gnn} $h_2$ which simulates $h_1$, i.e.\ $\forall G \in \mathcal{G}: h_1(G) = h_2(G)$.
+\end{thm}
+\begin{proof}
+	We prove the theorem by construction.
+	Let $G \in \mathcal{G}$ be an arbitrary input graph with $n \coloneqq \left|\mathcal{V}_G\right|$ and $m \coloneqq n + \left|\mathcal{E}_G\right|$.
+	By definition $h_1$ is a \ac{gnn} of the form $\mathit{Pool}_1(\mathit{Conv}_1(G))$ where $\mathit{Conv}_1$ is a stack of $T$ weighted vertex neighborhood sum convolutions ${\left\{ S^{(t)}: \mathbb{R}^{n \times d^{(t-1)}} \to \mathbb{R}^{n \times d^{(t)}} \right\}}_{t=1}^T$ with each corresponding $\mathrm{\acs*{mlp}}^{(t)}$ having $K$ layers.
+% chktex 21
+	$\mathit{Pool}_1$ combines the vertex feature vectors produced by $\mathit{Conv}_1$.
+
+	Let $h_2$ be a \ac{gnn} of the form $\mathit{Pool}_2(\mathit{Conv}_2(G))$ where $\mathit{Conv}_2$ is a stack of $(2 + K) T$ 2-\acs{wl} convolution layers ${\left\{ S^{(t, k)}: \mathbb{R}^{m \times d^{(t,k-1)}} \to \mathbb{R}^{m \times d^{(t,k)}} \right\}}_{(t, k) \in [T] \times [2+K]}$ with the neighborhood radius $r \coloneqq 1$.
+% chktex 21
+	Thus $\mathit{Conv}_2$ produces a feature vector for each of the $m$ edges of $G^1$, i.e.\ one for each vertex and each edge of $G$.
+	We denote the initial 2-\acs{wl} feature matrix with $Z^{(1,0)} \in \mathbb{R}^{m \times d^{(0,0)}}$.
+	The layers ${\left\{ S^{(t, 2 + K)} \right\}}_{t=1}^T$ produce the feature matrices $Z^{(t, 2 + K)} = Z^{(t + 1, 0)}$ which are then fed as input into the successor layer $S^{(t + 1, 1)}$.
+% chktex 21
+	Intuitively $\mathit{Conv}_2$ simulates each layer of $\mathit{Conv}_1$ via a stack of $2 + K$ 2-\acs{wl} convolution layers.
+	This is illustrated in \cref{fig:ltd:wl2-simulation}.
+	\begin{figure}[ht]
+		\centering
+		\includegraphics[width=\linewidth]{gfx/ltd/wl2-simulation.pdf}
+		\caption[Illustration of a 2-\acs{wl}-\acs{gnn} architecture that simulates vertex neighborhood convolutions.]{
+			Illustration of the correspondence between $\mathit{Conv}_1$ and $\mathit{Conv}_2$.
+		}\label{fig:ltd:wl2-simulation}
+	\end{figure}
+
+	Let $\varphi: \mathbb{R}^{d^{(T, 2 + K)}} \to \mathbb{R}^{d^{(T)}} \cup \{ \texttt{nil} \}$ be a function which maps the final 2-\acs{wl} feature vectors produced by $\mathit{Conv}_2$ to the output space of $\mathit{Conv}_1$ or the constant \texttt{nil}.
+	Let $\mathit{Pool}_2(Z_G^{(T,2+K)}) \coloneqq \mathit{Pool}_1(\ldblbrace z_{ij}\, |\, z_{ij} = \varphi(Z_G^{(T,2+K)}[e_{ij}]) \land e_{ij} \in \mathcal{E}_{G^1} \land z_{ij} \neq \texttt{nil} \rdblbrace)$.
+	\Cref{thm:ltd:wl2-simulation} then follows if there is a function $\varphi$ s.t.\ $\forall v_i \in \mathcal{V}_G: \mathit{Conv}_1(G)[v_i] = \varphi(\mathit{Conv}_2(G)[e_{ii}])$ and $\forall e_{ij} \in \mathcal{E}_G: \varphi(\mathit{Conv}_2(G)[e_{ij}]) = \texttt{nil}$.
+	To guarantee that there is such a function $\varphi$ we now inductively prove the following three invariants which have to hold for all $t \in \{ 0, \dots, T \}$:
+	\begin{enumerate}[label={(P\arabic*)}]
+		\item\label{inv:ltd:wl2-simulation:indicator}
+			${Z_G^{(t,2+K)}[e_{ij}]}_1 = \mathbbm{1}[i = j]$, i.e.\ the first component of each 2-\acs{wl} feature vector allows $\varphi$ to decide whether it should be mapped to \texttt{nil}.
+		\item\label{inv:ltd:wl2-simulation:feature}
+			${Z_G^{(t,2+K)}[e_{ii}]}_{2, \dots, (d^{(t)} + 1)} = Z_G^{(t)}[v_{i}]$, i.e.\ the second to $(1 + d^{(t)})$-th components of each self-loop feature vector contain the corresponding convolved vertex feature vector at layer $t$.
+		\item\label{inv:ltd:wl2-simulation:weight}
+			${Z_G^{(t,2+K)}[e_{ij}]}_{d^{(t)} + 2} = w_{ij}$, i.e.\ the weights for the vertex neighborhood sums are encoded in the edge and self-loop feature vectors.
+	\end{enumerate}
+	For $t = 0$ all three invariants hold by \cref{defn:ltd:wl2-conv-init}:
+	\begin{align*}
+		\forall v_i \in \mathcal{V}_G: Z_G^{(1, 0)}[e_{ii}] \coloneqq (1) \oplus x[v_i] \oplus (w_{ii})
+		\text{ and }
+		\forall e_{ij} \in \mathcal{E}_G: Z_G^{(1, 0)}[e_{ij}] \coloneqq (0) \oplus \mathbf{0} \oplus (w_{ij})
+		\text{.}
+	\end{align*}
+	Assuming the invariants hold for $t - 1$ we now show that they also hold for $t$.
+	As illustrated in \cref{fig:ltd:wl2-simulation} the layers $S^{(t, 1)}$ and $S^{(t, 2)}$ should compute the weighted vertex neighborhood sums
+	\begin{align*}
+		{Z^{(t, 2)}[e_{ii}]}_{2, \dots, (1+d^{(t-1)})} = w_{ii} {Z^{(t, 0)}[e_{ii}]}_{2, \dots, (1+d^{(t-1)})} + \smashoperator[lr]{\sum_{v_j \in \Gamma_G(v_i)}} w_{ij} {Z^{(t, 0)}[e_{jj}]}_{2, \dots, (1+d^{(t-1)})}
+		\text{.}
+	\end{align*}
+	We now explicitly define parameter matrices for $S^{(t, 1)}$ and $S^{(t, 2)}$ s.t.\ this weighted sum is produced.
+	Note that the weighted vertex neighborhood sum only requires scalar multiplication and vector addition, i.e.\ the $d^{(t-1)}$ vertex feature dimensions are mutually independent.
+	W.l.o.g.\ this allows us to simplify notation by treating the vertex feature vectors as if they were scalars in the following definitions, i.e.\ we can assume $d^{(t-1)} = 1$ and $Z^{(t,0)}[e_{ii}] = (1, Z^{(t-1)}[v_i], w_{ii}) \in \mathbb{R}^{3}$.
+	The layer $S^{(t, 1)}$ is defined by
+	\begin{gather*}
+		\begin{aligned}
+			Z^{(t,1)}[e_{ij}] &= \textcolor{t_blue}{Z^{(t,0)}[e_{ij}] W_{\mathrm{L}}^{(t,1)}} + \smashoperator[lr]{\sum_{v_l \in {\Gamma_{G^1}(v_i) \cap \Gamma_{G^1}(v_j)}}} \textcolor{t_red}{\left( Z^{(t,0)}[e_{ij}] W_{\mathrm{F}}^{(t,1)}\right)} \odot \textcolor{t_darkgreen}{\left( \left( Z^{(t,0)}[e_{il}] + Z^{(t,0)}[e_{lj}] \right) W_{\Gamma}^{(t,1)} \right)} \\ % chktex 25
+			&= \begin{cases}
+				\textcolor{t_blue}{(1, 0, w_{ii}, 0)} + \textcolor{t_darkgreen}{\left(0, 0, 0, 2\textcolor{t_red}{w_{ii}} Z^{(t-1)}[v_i]\right)} + \smashoperator[lr]{\sum_{v_l \in \Gamma_{G^1}(v_i)}} \textcolor{t_darkgreen}{\left(0, w_{il} \textcolor{t_red}{Z^{(t-1)}[v_i]}, 0, 0\right)} & \text{if $i = j$} \\ % chktex 25
+				\textcolor{t_blue}{(0, 0, w_{ij}, 0)} + \textcolor{t_darkgreen}{\left(0, 0, 0, \textcolor{t_red}{w_{ij}} Z^{(t-1)}[v_i]\right)} + \textcolor{t_darkgreen}{\left(0, 0, 0, \textcolor{t_red}{w_{ij}} Z^{(t-1)}[v_j]\right)} & \text{else} % chktex 25
+			\end{cases} \\
+			&= \begin{cases}
+				\left(1, \mkern6mu\smashoperator[lr]{\sum_{v_l \in \Gamma_{G^1}(v_i)}} w_{il} Z^{(t-1)}[v_i], w_{ii}, 2 w_{ii} Z^{(t-1)}[v_i] \right) & \text{if $i = j$} \\
+				\left(0, 0, w_{ij}, w_{ij} (Z^{(t-1)}[v_i] + Z^{(t-1)}[v_j])\right) & \text{else}
+			\end{cases}
+		\end{aligned}\\
+		\text{with }
+		W_{\mathrm{L}}^{(t,1)} \coloneqq \begin{pmatrix}
+			\textcolor{t_blue}{1} & 0 & 0 & 0 \\ % chktex 25
+			0 & 0 & 0 & 0 \\
+			0 & 0 & \textcolor{t_blue}{1} & 0 % chktex 25
+		\end{pmatrix},
+		W_{\mathrm{F}}^{(t,1)} \coloneqq \begin{pmatrix}
+			0 & 0 & 0 & 0 \\
+			0 & \textcolor{t_red}{\frac{1}{2}} & 0 & 0 \\ % chktex 25
+			0 & 0 & 0 & \textcolor{t_red}{1} % chktex 25
+		\end{pmatrix},
+		W_{\Gamma}^{(t,1)} \coloneqq \begin{pmatrix}
+			0 & 0 & 0 & 0 \\
+			0 & 0 & 0 & \textcolor{t_darkgreen}{1} \\ % chktex 25
+			0 & \textcolor{t_darkgreen}{1} & 0 & 0 % chktex 25
+		\end{pmatrix}
+		\text{.}
+	\end{gather*}
+	The vertex neighborhood summation is completed via $S^{(t, 2)}$ which is defined by
+	\begin{gather*}
+		\begin{aligned}
+			Z^{(t,2)}[e_{ij}] &= \begin{cases}
+				\textcolor{t_blue}{\left(1, -\smashoperator[lr]{\sum_{v_l \in \Gamma_{G^1}(v_i)}} w_{il} Z^{(t-1)}[v_i], w_{ii} \right)} + \smashoperator[lr]{\sum_{v_l \in \Gamma_{G^1}(v_i)}} \textcolor{t_darkgreen}{\left(0, w_{il}\left(Z^{(t-1)}[v_i] + Z^{(t-1)}[v_l]\right), 0\right)} & \text{if $i = j$} \\ % chktex 25
+				\textcolor{t_blue}{(0, 0, w_{ij})} & \text{else} % chktex 25
+			\end{cases} \\
+			&= \begin{cases}
+				\left(1, w_{ii} Z^{(t-1)}[v_i] + \smashoperator[lr]{\sum_{v_l \in \Gamma_{G}(v_i)}} w_{il} Z^{(t-1)}[v_l], w_{ii}\right) & \text{if $i = j$} \\
+				(0, 0, w_{ij}) & \text{else}
+			\end{cases}
+		\end{aligned}\\
+		\text{with }
+		W_{\mathrm{L}}^{(t,2)} \coloneqq \begin{pmatrix}
+			\textcolor{t_blue}{1} & 0 & 0 \\ % chktex 25
+			0 & \textcolor{t_blue}{-1} & 0 \\ % chktex 25
+			0 & 0 & \textcolor{t_blue}{1} \\ % chktex 25
+			0 & 0 & 0
+		\end{pmatrix},\quad
+		W_{\mathrm{F}}^{(t,2)} \coloneqq \begin{pmatrix}
+			0 & \textcolor{t_red}{\frac{1}{2}} & 0 \\ % chktex 25
+			0 & 0 & 0 \\
+			0 & 0 & 0 \\
+			0 & 0 & 0
+		\end{pmatrix},\quad
+		W_{\Gamma}^{(t,2)} \coloneqq \begin{pmatrix}
+			0 & 0 & 0 \\
+			0 & 0 & 0 \\
+			0 & 0 & 0 \\
+			0 & \textcolor{t_darkgreen}{1} & 0 % chktex 25
+		\end{pmatrix}
+		\text{.}
+	\end{gather*}
+	Using the two layers $S^{(t, 1)}$ and $S^{(t, 2)}$ that we just defined, the weighted vertex neighborhood sum for all $v_i \in \mathcal{V}_G$ is contained in $Z^{(t,2)}[e_{ii}]$.
+	Additionally for all $e_{ij} \in \mathcal{E}_{G^1}$ the indicators ${Z^{(t,2)}[e_{ij}]}_1 = \mathbbm{1}[i = j]$ and the weights ${Z^{(t,2)}[e_{ij}]}_{d^{(t)} + 2} = w_{ij}$ are preserved.
+	This means that the invariants~\ref{inv:ltd:wl2-simulation:indicator} and~\ref{inv:ltd:wl2-simulation:weight} are satisfied after $S^{(t, 2)}$.
+	To complete the induction step it now remains to show that all three invariants hold after applying the layers $S^{(t, 2+1)}, \dots, S^{(t, 2+K)}$.
+	As previously mentioned, a 2-\acs{wl} convolution layer is reduced to a fully connected layer if $W_{F}^{(t)} = \mathbf{0}$.
+	Via the universal approximation theorem~\cite{Hornik1991} we can therefore use $S^{(t, 2+1)}, \dots, S^{(t, 2+K)}$ to simulate the $K$ layers of $\mathrm{\acs*{mlp}}^{(t)}$ without changing the first and last dimension of each feature vector to preserve the invariants~\ref{inv:ltd:wl2-simulation:indicator} and~\ref{inv:ltd:wl2-simulation:weight}.
+	The resulting feature matrix $Z^{(t,2+K)}$ then satisfies all three invariants which completes the induction.
+
+	Using invariants~\ref{inv:ltd:wl2-simulation:indicator} and~\ref{inv:ltd:wl2-simulation:feature} for $t = T$ we can therefore set
+	\begin{align*}
+		\varphi(Z_G^{(T,2+K)}[e_{ij}]) \coloneqq \begin{cases}
+			{Z_G^{(T,2+K)}[e_{ij}]}_{2, \dots, (d^{T} + 1)} & \text{if ${Z_G^{(T,2+K)}[e_{ij}]}_1 = 1$} \\
+			\texttt{nil} & \text{else}
+		\end{cases}
+		\text{.}
+	\end{align*}
+	By our previous definition of $\mathit{Pool}_2$, this in turn implies that $\mathit{Pool}_2(Z_G^{(T,2+K)}) = \mathit{Pool}_1(Z_G^{(T)}) \iff h_2(G) = h_1(G)$ which concludes the proof.
+\end{proof}
 

### 2020-04-03 thesis/content/chapter-related-work.tex
started with 2-wl-gnn expressive power
completed proof that 2-wl-gnns are at least as powerful as 1-wl

@@
 To see why this is the case, we rewrite \cref{eq:related:gcn-layer} and compare it with the 1-\acs{wl} color refinement strategy (see \cfullref{defn:related:wl1-refine}):
-\begin{alignat*}{2}
+\begin{equation}
+	\begin{alignedat}{2}
 		\text{\ac*{gcn}:}\quad&& Z_G^{(t)}[v] &= \sigma\left(\left( {\mu(v)}^2 \textcolor{t_blue}{Z_G^{(t-1)}[v]} + \smashoperator[lr]{\sum_{u \in \Gamma_G(v)}} \mu(v) \mu(u) \textcolor{t_red}{Z_G^{(t-1)}[u]} \right) \Theta^{(t)} \right) \\ % chktex 25
 		\text{1-\acs{wl}:}\quad&& \chi_{G,1}^{(t)}(v) &= h\left( \mkern60mu\textcolor{t_blue}{\chi_{G,1}^{(t-1)}(v)}, \ldblbrace \textcolor{t_red}{\chi_{G,1}^{(t-1)}(u)}\, |\, u \in \Gamma_{G}(v) \rdblbrace\right) % chktex 25
-\end{alignat*}
+	\end{alignedat}\label{eq:related:gcn-wl1}
+\end{equation}
 Here $\mu(v) \coloneqq {\left(\left| \Gamma_G(v) \right| + 1\right)}^{-\frac{1}{2}}$ are normalization factors introduced by $\tilde{D}_G^{-\frac{1}{2}}$.
@@
 \begin{align}
-	Z_G^{(t)}[v] \coloneqq \mathrm{\acs*{mlp}}^{(t)}\left( Z_G^{(t-1)}[v] + \smashoperator[lr]{\sum_{u \in \Gamma_G(v)}} Z_G^{(t-1)}[u] \right)
+	Z_G^{(t)}[v] \coloneqq \mathrm{\acs*{mlp}}^{(t)}\left( Z_G^{(t-1)}[v] + \smashoperator[lr]{\sum_{u \in \Gamma_G(v)}} Z_G^{(t-1)}[u] \right) \label{eq:related:gin-layer}
 \end{align}

### 2020-04-04 thesis/content/chapter-ltd.tex
completed 2-wl-gnn expressive power section

@@
 \end{defn}
-The convolution operator from \cref{defn:ltd:wl2-conv-step} is parameterized by the three matrices $W_{\mathrm{L}}^{(t)}, W_{\mathrm{F}}^{(t)},  W_{\Gamma}^{(t)} \in \mathbb{R}^{d^{(t-1)} \times d^{(t)}}$ and uses two freely choosable nonlinearities $\sigma$ and $\sigma_{\Gamma}$.
+The convolution operator from \cref{defn:ltd:wl2-conv-step} is parameterized by the three matrices $W_{\mathrm{L}}^{(t)}, W_{\mathrm{F}}^{(t)},  W_{\Gamma}^{(t)} \in \mathbb{R}^{d^{(t-1)} \times d^{(t)}}$ and uses two freely choosable activation functions $\sigma$ and $\sigma_{\Gamma}$.
 There are three properties which motivate this particular choice of convolution layer:
@@
 	\item \textbf{Simulation of \acp{mlp}:}
-		By choosing $W_{F}^{(t)} = \mathbf{0}$ the convolution behaves like a regular fully connected layer.
+		By choosing $W_{\mathrm{F}}^{(t)} = \mathbf{0}$ the convolution behaves like a regular fully connected layer.
 		A stack of 2-\acs{wl} convolution layers can therefore simulate arbitrary \acp{mlp}.
@@
 	\item \textbf{Context-dependent neighborhood filtering:}
-		Instead of filtering out all neighbors, $W_{F}^{(t)}$ also allows to filter the neighborhood of $e_{ij}$ more selectively.
+		Instead of filtering out all neighbors, $W_{\mathrm{F}}^{(t)}$ also allows to filter the neighborhood of $e_{ij}$ more selectively.
 		We can interpret $z_{ij} W_{\mathrm{F}}^{(t)} \in \mathbb{R}^{d^{(t)}}$ as a row vector of feature dimension weights.
@@
 	Let $\varphi: \mathbb{R}^{d^{(T, 2 + K)}} \to \mathbb{R}^{d^{(T)}} \cup \{ \texttt{nil} \}$ be a function which maps the final 2-\acs{wl} feature vectors produced by $\mathit{Conv}_2$ to the output space of $\mathit{Conv}_1$ or the constant \texttt{nil}.
-	Let $\mathit{Pool}_2(Z_G^{(T,2+K)}) \coloneqq \mathit{Pool}_1(\ldblbrace z_{ij}\, |\, z_{ij} = \varphi(Z_G^{(T,2+K)}[e_{ij}]) \land e_{ij} \in \mathcal{E}_{G^1} \land z_{ij} \neq \texttt{nil} \rdblbrace)$.
+	Let $\mathit{Pool}_2\left(Z_G^{(T,2+K)}\right) \coloneqq \mathit{Pool}_1(\ldblbrace z_{ij}\, |\, z_{ij} = \varphi\left(Z_G^{(T,2+K)}[e_{ij}]\right) \land e_{ij} \in \mathcal{E}_{G^1} \land z_{ij} \neq \texttt{nil} \rdblbrace)$.
 	\Cref{thm:ltd:wl2-simulation} then follows if there is a function $\varphi$ s.t.\ $\forall v_i \in \mathcal{V}_G: \mathit{Conv}_1(G)[v_i] = \varphi(\mathit{Conv}_2(G)[e_{ii}])$ and $\forall e_{ij} \in \mathcal{E}_G: \varphi(\mathit{Conv}_2(G)[e_{ij}]) = \texttt{nil}$.
@@
 		\item\label{inv:ltd:wl2-simulation:indicator}
-			${Z_G^{(t,2+K)}[e_{ij}]}_1 = \mathbbm{1}[i = j]$, i.e.\ the first component of each 2-\acs{wl} feature vector allows $\varphi$ to decide whether it should be mapped to \texttt{nil}.
+			${Z_G^{(t,2+K)}[e_{ij}]}_1 = \mathbbm{1}[i = j]$, i.e.\ the first component of each 2-\acs{wl} feature vector allows $\varphi$ to decide whether that vector should be mapped to \texttt{nil}.
 		\item\label{inv:ltd:wl2-simulation:feature}
-			${Z_G^{(t,2+K)}[e_{ii}]}_{2, \dots, (d^{(t)} + 1)} = Z_G^{(t)}[v_{i}]$, i.e.\ the second to $(1 + d^{(t)})$-th components of each self-loop feature vector contain the corresponding convolved vertex feature vector at layer $t$.
+			${Z_G^{(t,2+K)}[e_{ii}]}_{2, \dots, (d^{(t)} + 1)} = Z_G^{(t)}[v_{i}]$, i.e.\ the second to $(1 + d^{(t)})$-th components of each self-loop feature vector in $h_2$ contain the corresponding convolved vertex feature vector at layer $t$ in $h_1$.
 		\item\label{inv:ltd:wl2-simulation:weight}
@@
 	To complete the induction step it now remains to show that all three invariants hold after applying the layers $S^{(t, 2+1)}, \dots, S^{(t, 2+K)}$.
-	As previously mentioned, a 2-\acs{wl} convolution layer is reduced to a fully connected layer if $W_{F}^{(t)} = \mathbf{0}$.
+	As previously mentioned, a 2-\acs{wl} convolution layer is reduced to a fully connected layer if $W_{\mathrm{F}}^{(t)} = \mathbf{0}$.
 	Via the universal approximation theorem~\cite{Hornik1991} we can therefore use $S^{(t, 2+1)}, \dots, S^{(t, 2+K)}$ to simulate the $K$ layers of $\mathrm{\acs*{mlp}}^{(t)}$ without changing the first and last dimension of each feature vector to preserve the invariants~\ref{inv:ltd:wl2-simulation:indicator} and~\ref{inv:ltd:wl2-simulation:weight}.
@@
 	\begin{align*}
-		\varphi(Z_G^{(T,2+K)}[e_{ij}]) \coloneqq \begin{cases}
+		\varphi\left(Z_G^{(T,2+K)}[e_{ij}]\right) \coloneqq \begin{cases}
 			{Z_G^{(T,2+K)}[e_{ij}]}_{2, \dots, (d^{T} + 1)} & \text{if ${Z_G^{(T,2+K)}[e_{ij}]}_1 = 1$} \\
@@
 \end{proof}
+\begin{cor}\label{cor:ltd:wl2-gnn-wl1-power}
+	2-\acs{wl}-\acsp{gnn} have at least the same discriminative power as 1-\acs{wl}.
+\end{cor}
+\begin{proof}
+	The corollary directly follows from the fact that 2-\acs{wl}-\acsp{gnn} can simulate \acp{gin} by \cref{thm:ltd:wl2-simulation} and the fact that \acp{gin} have the same discriminative power as 1-\acs{wl} by \cref{prop:related:gcnn-wl1-limit} (see \cpageref{prop:related:gcnn-wl1-limit}) because they use injective vertex neighborhood hashing functions~\cite{Xu2018}.
+\end{proof}
+
+To complete our analysis of the expressive power of 2-\acs{wl}-\acsp{gnn} we now show that they are not just as powerful as 1-\acs{wl} but in fact more powerful than 1-\acs{wl}.
+\begin{prop}\label{cor:ltd:wl2-gnn-regular}
+	There are $d$-regular graphs $G$ and $H$ of size $n$ which can be distinguished by 2-\acs{wl}-\acsp{gnn}.
+\end{prop}
+\begin{proof}
+	The proposition follows if we choose the six-cycle graph for $G$ and the two three-cycles graph for $H$, both of which should be well-known at this point.
+	Let $h_2 = \mathit{Pool} \circ S$ be a 2-\acs{wl}-\acs{gnn} with the neighborhood radius $r = 1$ which consists of a single 2-\acs{wl} convolution layer $S: \mathbb{R}^{* \times 2} \to \mathbb{R}^{* \times 1}$ and the pooling layer $\mathit{Pool} = \min$.
+	In accordance with \cref{defn:ltd:wl2-conv-init} we set the initial feature vectors of the vertices $v_i$ of $G$ and $H$ to $Z^{(0)}[e_{ii}] \coloneqq (1, 0)$ and the initial feature vectors of their edges $e_{ij}$ to $Z^{(0)}[e_{ij}] \coloneqq (0, 1)$.
+	Let the weight matrices of $S$ be $W_{\mathrm{L}} \coloneqq \mathbf{0}$ and $W_{\mathrm{F}} = W_{\Gamma} \coloneqq \begin{pmatrix} 1\\1 \end{pmatrix}$.
+	For simplicity we choose the identity activation functions $\sigma = \sigma_{\Gamma} = \mathrm{id}$.
+
+	By \cref{defn:ltd:wl2-conv-step} all self-loops $e_{ii}$ of $G^1$ and $H^1$ have the three neighbors $\{ \textcolor{t_blue}{\ldblbrace e_{ii}, e_{ii} \rdblbrace},\allowbreak \textcolor{t_red}{\ldblbrace e_{ij}, e_{ji} \rdblbrace},\allowbreak \textcolor{t_darkgreen}{\ldblbrace e_{il}, e_{li} \rdblbrace} \}$, i.e.\ the length-two walk along $e_{ii}$ itself and the length-two walks to and from the two neighboring vertices $\Gamma(v_i) = \{ v_j, v_l \}$.
+% chktex 25
+	Therefore the convolved feature vector of all self-loops are $Z^{(1)}[e_{ii}] = (1) \odot (\textcolor{t_blue}{(1 + 1)} + \textcolor{t_red}{(1 + 1)} + \textcolor{t_darkgreen}{(1 + 1)}) = 6$.
+% chktex 25
+	However for the non-self-loops of $G^1$ and $H^1$, i.e.\ the edges of $G$ and $H$, we get differing convolved feature vectors.
+	The 2-\acs{wl} neighbors of $e_{ij} \in \mathcal{E}_G$ are $\{ \ldblbrace e_{ii},  e_{ij} \rdblbrace,\allowbreak \ldblbrace e_{ij}, e_{jj} \rdblbrace \}$.
+	The 2-\acs{wl} neighbors of $e_{ij}' \in \mathcal{E}_H$ are $\{ \ldblbrace e_{ii}',  e_{ij}' \rdblbrace,\allowbreak \ldblbrace e_{ij}',  e_{jj}' \rdblbrace,\allowbreak \ldblbrace e_{il}',  e_{lj}' \rdblbrace \}$ where $v_l' \in \mathcal{V}_H$ is the common neighbor of $v_i'$ and $v_j'$.
+	The different neighborhood sizes of the edges of $G$ and $H$ imply that $\forall e_{ij} \in \mathcal{E}_G: Z^{(1)}[e_{ij}] = 4$ while $\forall e_{ij}' \in \mathcal{E}_H: Z^{(1)}[e_{ij}] = 6$.
+	Thus $h_2(G) = \min\{ 4, 6 \} \neq \min\{ 6, 6 \} = h_2(H)$ which concludes the proof. % chktex 21
+\end{proof}
+\begin{cor}\label{cor:ltd:wl2-gnn-more-wl1-power}
+	The discriminative power of 2-\acs{wl}-\acsp{gnn} is strictly greater than that of the 1-\acs{wl} algorithm.
+\end{cor}
+\begin{proof}
+	The corollary directly follows from \cref{cor:ltd:wl2-gnn-wl1-power} and \cref{cor:ltd:wl2-gnn-regular} since 1-\acs{wl} cannot distinguish regular graphs~\cite[cor.~1.8.5]{Immerman1990}.
+\end{proof}
+
+This concludes our analysis of the expressive power of 2-\acs{wl}-\acsp{gnn}.
+The key takeaway from this section is that 2-\acs{wl}-\acsp{gnn} are more powerful than all vertex neighborhood aggregation \acp{gcnn} because the power of the latter is upper bounded by 1-\acs{wl}.
+Additionally we can conclude that 2-\acs{wl}-\acsp{gnn} are in fact also more powerful than 2-\acsp{gnn} due to \cref{prop:ltd:2gnn-regular-limit}.
+A more thorough discussion of the relation between 2-\acs{wl}-\acsp{gnn} and 2-\acs{wl} is left for future work.
 

### 2020-04-04 thesis/content/chapter-related-work.tex
completed 2-wl-gnn expressive power section

@@
 \end{figure}
-For $k = 2$ this means that each potential edge $(v, w) \in \mathcal{V}_G^2$ has all possible paths of length $2$ from $v$ to $w$ as its neighbors (see~\cref{fig:related:wl-neighbors:2}).
+For $k = 2$ this means that each potential edge $(v, w) \in \mathcal{V}_G^2$ has all possible walks of length $2$ from $v$ to $w$ as its neighbors (see~\cref{fig:related:wl-neighbors:2}).
 Also note that, even though $k$-\acs{wl} refines $k$-tuple colors, lower-dimensional structures still get their own colors since a tuple does not have to consist of distinct vertices, i.e.\ in $k$-\acs{wl} the color of a single vertex $v \in \mathcal{V}_G$ is described by $\chi^{*}_{G, k}(s)$ for $s = (v, \dots, v) \in \mathcal{V}_G^k$.
@@
 	Only the idea behind triangle and 4-cycle counting is outlined here to give an intuition for why \cref{prop:related:wl2-cycle-count} holds.
-	Note that the 2-\acs{wl} neighbors of each edge $(v, w) \in \mathcal{E}_G$ correspond to all possible paths $(v, u, w)$ of length 2, i.e.\ all possible triangles.
+	Note that the 2-\acs{wl} neighborhood of each edge $(v, w) \in \mathcal{E}_G$ includes all possible paths $(v, u, w)$ of length $2$, i.e.\ all possible triangles.
 	By \cref{defn:related:wlk-refine} there must thus be a color subset $C^{\triangle}_j \subseteq \mathcal{C}$ representing that an edge is involved in exactly $j$ triangles after one refinement step.
@@
 In the next step we will take a closer look at the relation between \acp{gcnn} that use vertex neighborhood convolution and 1-\acs{wl} by comparing their respective discriminative power.
-\begin{prop}[see {\citet[lemma 2]{Xu2018}} for the proof]\label{prop:related:gcnn-wl1-limit}
-	The discriminative power of a \ac{gcnn} that convolves vertex feature vectors via a convolution operator of the form $Z_G^{(t)}[v] = h^{(t)}\left(Z_G^{(t-1)}[v], \ldblbrace Z_G^{(t-1)}[u]\, |\, u \in \Gamma_G(v) \rdblbrace\right)$ is upper-bounded by that of 1-\acs{wl}, where $h^{(t)}: {(\mathbb{R}^{d^{(t-1)}})}^{*} \to \mathbb{R}^{d^{(t)}}$ is an arbitrary vertex neighborhood hashing function.
-	Moreover, if $h^{(t)}$ is not injective, the \ac{gcnn} has a strictly lower discriminative power than 1-\acs{wl}.
+\begin{prop}[see {\citet[lemma 2 and theorem 3]{Xu2018}} for the proof]\label{prop:related:gcnn-wl1-limit}
+	The discriminative power of a \ac{gcnn} that convolves vertex feature vectors via a convolution operator of the form $Z_G^{(t)}[v] = h^{(t)}\left(Z_G^{(t-1)}[v], \ldblbrace Z_G^{(t-1)}[u]\, |\, u \in \Gamma_G(v) \rdblbrace\right)$ is upper-bounded by that of 1-\acs{wl}, where $h^{(t)}: {\left(\mathbb{R}^{d^{(t-1)}}\right)}^{*} \to \mathbb{R}^{d^{(t)}}$ is an arbitrary vertex neighborhood hashing function.
+	Moreover, iff.\ $h^{(t)}$ is injective, the \ac{gcnn} has the same discriminative power as 1-\acs{wl}.
 \end{prop}

### 2020-04-04 thesis/content/chapter-ltd.tex
various fixes + started with 2-wl-gnn implementation section

@@
 		Two edge colorings on which 2-\acsp{gnn} and 2-\acs{wl} behave differently.
-		A 2-\acs{gnn} will refine the ``color vector'' of $e_{ij}$ to \textsf{\textbf{\small\textcolor{t_blue}{D}}} for both initial colorings.
+		A 2-\acs{gnn} will refine the ``color vector'' of $e_{ij}$ to \colorlabel{t_blue}{D} for both initial colorings.
 		2-\acs{wl} on the other hand differentiates both colorings by preserving the color tuple information.
@@
 
-	W.l.o.g.\ we define the initial colors of the loop vertices as $\chi^{(0)}(v) = \textcolor{t_blue}{\texttt{A}}$ for all $v \in L_G \cup L_H$. % chktex 25
+	W.l.o.g.\ we define the initial colors of the loop vertices as $\chi^{(0)}(v) = \colorlabel{t_blue}{A}$ for all $v \in L_G \cup L_H$. % chktex 25
 	The initial colors of the edge vertices are defined as $\chi^{(0)}(e) = \textcolor{t_red}{\texttt{B}}$ for all $e \in E_G \cup E_H$. % chktex 25
@@
 
-	After one color refinement step we get $\chi^{(1)}(v) = h(\textcolor{t_blue}{\texttt{A}}, \ldblbrace \underbrace{\textcolor{t_red}{\texttt{B}}, \dots, \textcolor{t_red}{\texttt{B}}}_{d\text{ times}} \rdblbrace) \eqqcolon \textcolor{t_blue}{\texttt{C}}$ for all loop vertices $v \in  L_G \cup L_H$ and $\chi^{(1)}(e) = h(\textcolor{t_red}{\texttt{B}}, \ldblbrace \textcolor{t_blue}{\texttt{A}}, \textcolor{t_blue}{\texttt{A}}, \underbrace{\textcolor{t_red}{\texttt{B}}, \dots, \textcolor{t_red}{\texttt{B}}}_{\mathclap{2d - 2 \text{ times}}} \rdblbrace) \eqqcolon \textcolor{t_red}{\texttt{D}}$.
+	After one color refinement step we get $\chi^{(1)}(v) = h(\colorlabel{t_blue}{A}, \ldblbrace \underbrace{\textcolor{t_red}{\texttt{B}}, \dots, \textcolor{t_red}{\texttt{B}}}_{d\text{ times}} \rdblbrace) \eqqcolon \textcolor{t_blue}{\texttt{C}}$ for all loop vertices $v \in  L_G \cup L_H$ and $\chi^{(1)}(e) = h(\textcolor{t_red}{\texttt{B}}, \ldblbrace \colorlabel{t_blue}{A}, \colorlabel{t_blue}{A}, \underbrace{\textcolor{t_red}{\texttt{B}}, \dots, \textcolor{t_red}{\texttt{B}}}_{\mathclap{2d - 2 \text{ times}}} \rdblbrace) \eqqcolon \textcolor{t_red}{\texttt{D}}$.
 % chktex 25
-	This means that $\chi^{(0)}$ and $\chi^{(1)}$ are identical up to the color substitutions $\textcolor{t_blue}{\texttt{A}} \to \textcolor{t_blue}{\texttt{C}}$ and $\textcolor{t_red}{\texttt{B}} \to \textcolor{t_red}{\texttt{D}}$, i.e.\ $\chi^{(0)} \equiv \chi^{(1)}$, which in turn implies that 1-\ac{wl} terminates after one iteration.
+	This means that $\chi^{(0)}$ and $\chi^{(1)}$ are identical up to the color substitutions $\colorlabel{t_blue}{A} \to \textcolor{t_blue}{\texttt{C}}$ and $\textcolor{t_red}{\texttt{B}} \to \textcolor{t_red}{\texttt{D}}$, i.e.\ $\chi^{(0)} \equiv \chi^{(1)}$, which in turn implies that 1-\ac{wl} terminates after one iteration.
 % chktex 25
@@
 \end{enumerate}
-Based on the 2-multiset and the neighborhood localization simplifications we can now define the 2-\ac{wl} convolution operator.
+Based on the 2-multiset and the neighborhood localization simplifications we can now define the 2-\ac{wl} convolution operator and the corresponding so-called \textit{2-\acs{wl}-\acs{gnn}}.
 \begin{defn}\label{defn:ltd:wl2-conv-init}
@@
 		as we saw in \cref{fig:ltd:2gnn-2wl-diff} this causes 2-\acsp{gnn} to lose the edge pair information due to the commutativity and associativity of $+$.
+		2-\acs{wl}-\acsp{gnn} do not generally have this problem because $\kappa^{(t)}(\colorlabel{t_blue}{A}, \ldblbrace \colorlabel{t_red}{B}, \colorlabel{t_red}{B} \rdblbrace) + \kappa^{(t)}(\colorlabel{t_blue}{A}, \ldblbrace \colorlabel{t_darkgreen}{C}, \colorlabel{t_darkgreen}{C} \rdblbrace) \neq \kappa^{(t)}(\colorlabel{t_blue}{A}, \ldblbrace \colorlabel{t_red}{B}, \colorlabel{t_darkgreen}{C} \rdblbrace) + \kappa^{(t)}(\colorlabel{t_blue}{A}, \ldblbrace \colorlabel{t_red}{B}, \colorlabel{t_darkgreen}{C} \rdblbrace)$ if $\sigma_{\Gamma}$ is chosen to be a nonlinear activation function.
+% chktex 25
 	\item \textbf{Context-dependent neighborhood filtering:}
@@
 
-In this section we will analyze the discriminative and computational power of the 2-\ac{wl} convolution operator that was just defined.
-We begin by showing that 2-\acs{wl}-\acsp{gnn} are at least as powerful as 1-\acs{wl}.
+In this section we will analyze the power of \acp{gnn} which use the 2-\acs{wl} convolution operator that was just defined.
+Our goal is to show that such 2-\acs{wl}-\acsp{gnn} have a strictly larger discriminative power than 1-\acs{wl}.
+We begin by proving that 2-\acs{wl}-\acsp{gnn} are at least as powerful as 1-\acs{wl}.
 
@@
 	The layer $S^{(t, 1)}$ is defined by
-	\begin{gather*}
+	\begin{gather*}\allowdisplaybreaks
 		\begin{aligned}
@@
 			\end{cases}
-		\end{aligned}\\
+		\end{aligned}\\[10pt]
 		\text{with }
@@
 	The vertex neighborhood summation is completed via $S^{(t, 2)}$ which is defined by
-	\begin{gather*}
+	\begin{gather*}\allowdisplaybreaks
 		\begin{aligned}
@@
 Additionally we can conclude that 2-\acs{wl}-\acsp{gnn} are in fact also more powerful than 2-\acsp{gnn} due to \cref{prop:ltd:2gnn-regular-limit}.
-A more thorough discussion of the relation between 2-\acs{wl}-\acsp{gnn} and 2-\acs{wl} is left for future work.
+
+Note that no statement regarding the discriminative or computational power of 2-\acs{wl}-\acsp{gnn} compared to 2-\acs{wl} was made.
+It is easy to see that 2-\acs{wl}-\acsp{gnn} \textit{generally} cannot have the same power as 2-\acs{wl} due to the neighborhood localization simplification;
+e.g.\ for a small neighborhood radius of $r = 1$, nonexistent edges $e_{ij} \notin \mathcal{E}_G$ do not have a feature vector which implies that the proof of 2-\acs{wl}'s $m$-cycle counting ability no longer holds for $m \geq 4$ (see \cfullref{prop:related:wl2-cycle-count}).
+We leave a more thorough discussion of the relation between 2-\acs{wl}-\acsp{gnn} and 2-\acs{wl} for future work.
 
@@
 \label{sec:ltd:wl2gnn:implementation}
+
+Apart from the theoretical expressive power of a model, it also has to be computable efficiently in order to be useful in practice.
+In this section we will therefore describe how 2-\acs{wl} convolutions can be implemented on \acp{gpgpu}.
+
+Efficient high-level software libraries for the implementation of vertex neighborhood convolution approaches such as \ac{gcn} or \ac{gin} already exist.
+They describe convolutions via a message-passing abstraction in which vertex feature vectors are passed along their neighboring edges (see \citet{Battaglia2018}).
+A few notable implementations of this abstraction are the \citetitle{GN} library~\cite{GN}, \citetitle{PyG}~\cite{Fey2019}\cite{PyG}, \citetitle{DGL}~\cite{Wang2019}\cite{DGL} and \citetitle{Spektral}~\cite{Spektral}.

### 2020-04-04 thesis/content/chapter-related-work.tex
various fixes + started with 2-wl-gnn implementation section

@@
 More recently \citet{Babai2015} also presented a quasipolynomial upper time bound for the remaining hard \ac{gi} instances.
-For all practical purposes, \ac{gi} can therefore be solved efficiently; e.g.\ via the \texttt{nauty} program~\cite{McKay}\cite{McKay2013}.
+For all practical purposes, \ac{gi} can therefore be solved efficiently; e.g.\ via the \texttt{nauty} program~\cite{McKay2013}\cite{McKay}.
 One important subroutine in most \ac{gi} checkers is the \acl{wl} algorithm which will be described next.
@@
 Since \ac{wl} colors are inherently discrete, continuous vertex feature vectors $x_G[v]$ are not considered here.
-For unlabeled graphs a constant coloring is used, e.g.\ $\forall v \in \mathcal{V}_G: \chi_{G,1}^{(0)}(v) = \textcolor{t_blue}{\texttt{A}}$ for some initial color $\textcolor{t_blue}{\texttt{A}} \in \mathcal{C}$.
+For unlabeled graphs a constant coloring is used, e.g.\ $\forall v \in \mathcal{V}_G: \chi_{G,1}^{(0)}(v) = \colorlabel{t_blue}{A}$ for some initial color $\colorlabel{t_blue}{A} \in \mathcal{C}$.
 % chktex 25
@@
 This simple approach usually requires a careful choice of subgraphs but can still be competitive with the other more recent approaches we will look at in the following sections.
-Fingerprint embeddings are for example used in multiple state-of-the-art toxicity prediction tools like RASAR~\cite{Luechtefeld2018}\cite{ToxTrack}, the \ac{toxTest}~\cite{TEST} or ProTox~\cite{Drwal2014}\cite{ProTox}\cite{Banerjee2018}.
+Fingerprint embeddings are for example used in multiple state-of-the-art toxicity prediction tools like RASAR~\cite{Luechtefeld2018}\cite{ToxTrack}, the \ac{toxTest}~\cite{TEST} or ProTox~\cite{Drwal2014}\cite{Banerjee2018}\cite{ProTox}.
 

### 2020-04-05 implementation/src/ltag/evaluate_datasets.py
added wl_st kernel models with fewer iterations

@@ -83,7 +83,7 @@ if __name__ == "__main__":
       ms = ["SagCWL2GCN_Binary_quick_max"]
     else:
       ms = [
-        "WL_st", "WL_sp", "LWL2", "GWL2",
+        "WL_st", "WL_st_1", "WL_st_2", "WL_sp", "LWL2", "GWL2",
         "AvgCWL2GCN_Binary", "SagCWL2GCN_Binary",
         "AvgK2GNN_Binary", "SagK2GNN_Binary"]
   else:

### 2020-04-05 implementation/src/ltag/evaluation/kernel_models.py

@@ -7,6 +7,8 @@ from ltag.utils import cart
 
 kernel_hps = lambda dsm: cart(C=[1, 0.1, 0.01, 0.001, 0.0001])
 WL_st = kernel_classifier(kernel_models.WL_st)(kernel_hps)
+WL_st_1 = kernel_classifier(kernel_models.WL_st_1)(kernel_hps)
+WL_st_2 = kernel_classifier(kernel_models.WL_st_2)(kernel_hps)
 WL_sp = kernel_classifier(kernel_models.WL_sp)(kernel_hps)
 LWL2 = kernel_classifier(kernel_models.LWL2)(kernel_hps)
 GWL2 = kernel_classifier(kernel_models.GWL2)(kernel_hps)

### 2020-04-05 implementation/src/ltag/evaluation/models.py

@@ -7,9 +7,9 @@ import ltag.models.gnn as gnn_models
 from ltag.utils import cart, cart_merge, entry_duplicator
 from ltag.evaluation.model_factories import binary_classifier
 from ltag.evaluation.kernel_models import (
-  WL_st, WL_sp, LWL2, GWL2)
+  WL_st, WL_st_1, WL_st_2, WL_sp, LWL2, GWL2)
 
-kernel_models = [WL_st, WL_sp, LWL2, GWL2]
+kernel_models = [WL_st, WL_st_1, WL_st_2, WL_sp, LWL2, GWL2]
 
 @binary_classifier(gnn_models.AvgWL2GCN)
 def AvgWL2GCN_Binary(dsm):

### 2020-04-05 implementation/src/ltag/models/kernel.py

@@ -80,6 +80,10 @@ def fs_kernel(name):
 # Models:
 WL_st = kernel_model("WL_st", grakel_kernel(
   "WL_st", kernel=[{"name": "WL"}, "VH"]))
+WL_st_1 = kernel_model("WL_st_1", grakel_kernel(
+  "WL_st_1", kernel=[{"name": "WL", "n_iter": 1}, "VH"]))
+WL_st_2 = kernel_model("WL_st_2", grakel_kernel(
+  "WL_st_2", kernel=[{"name": "WL", "n_iter": 2}, "VH"]))
 WL_sp = kernel_model("WL_sp", grakel_kernel(
   "WL_sp", kernel=[{"name": "WL"}, "SP"]))
 LWL2 = kernel_model("LWL2", fs_kernel("LWL2"))

### 2020-04-05 thesis/content/chapter-ltd.tex
completed 2-wl-gnn implementation section

@@
 Apart from the theoretical expressive power of a model, it also has to be computable efficiently in order to be useful in practice.
-In this section we will therefore describe how 2-\acs{wl} convolutions can be implemented on \acp{gpgpu}.
+In this section we will therefore describe how 2-\acs{wl} convolutions can be implemented on \acfp{gpgpu}.
 
@@
 A few notable implementations of this abstraction are the \citetitle{GN} library~\cite{GN}, \citetitle{PyG}~\cite{Fey2019}\cite{PyG}, \citetitle{DGL}~\cite{Wang2019}\cite{DGL} and \citetitle{Spektral}~\cite{Spektral}.
+Since a message-passing model along edges is incompatible with the edge-pair neighborhoods of 2-\acs{wl}, a custom convolution implementation is required for 2-\acs{wl}-\acsp{gnn}.
+
+For this purpose we propose a sparse 2-\acs{wl} graph representation which is inspired by the coordinate list adjacency format described by \citet{Fey2019}.
+Given a neighborhood radius $r$ we encode a graph $G$ using the following two matrices:
+\begin{enumerate}
+	\item $Z_G^{(0)} \in \mathbb{R}^{m \times d^{(0)}}$:
+		The initial feature matrix is represented directly as a dense floating point matrix with $m \coloneqq \left| \mathcal{E}_{G^r} \right|$ rows, each of which encodes the feature vector of an edge $e_{ij} \in \mathcal{E}_{G^r}$.
+		Edge feature duplicates are prevented by only encoding edges with $i \leq j$ for some arbitrary vertex ordering of $G$.
+	\item $R_G \in {[m]}^{\gamma \times 3}$:
+		The reference matrix $R_G$ encodes the edge neighborhood information.
+		It consists of $\gamma \coloneqq \sum_{e_{ij} \in \mathcal{E}_{G^r}} \left| \Gamma_{G^r}(v_i) \cap \Gamma_{G^r}(v_j) \right|$ rows, one for each 2-\acs{wl} neighbor $\ldblbrace e_{il}, e_{lj} \rdblbrace$ of each edge $e_{ij}$.
+		Each neighbor row is a vector $(r_{\mathrm{L}}, r_{\Gamma,1}, r_{\Gamma,2}) \in {[m]}^3$ of three index pointers to rows in $Z_G^{(0)}$.
+		$r_{\mathrm{L}}$ points to the row index of the feature vector of $e_{ij}$ while $r_{\Gamma,1}$ and $r_{\Gamma,2}$ point to the indices of $e_{il}$ and $e_{lj}$ respectively.
+		We will refer to the three column vectors of $R_G$ as $R_{G, \mathrm{L}}$, $R_{G, \Gamma, 1}$ and $R_{G, \Gamma, 2}$.
+\end{enumerate}
+This encoding can also be used to represent graph batches by simply concatenating the rows of each graph's feature and reference matrices while shifting the index pointers to account for the concatenation offsets.
+\Cref{fig:ltd:wl2-encoding} illustrates how such a batch encoding might look like.
+\begin{figure}[t]
+	\centering
+	\def\vertvec{$(1, 0)$}
+	\def\edgevec{$(0, 1)$}
+	\raisebox{-.5\height}{\includegraphics[height=11em]{gfx/ltd/wl2-encoding.pdf}}\hspace{3em}%
+	\begin{tabu}{@{\makebox[2.5em][c]{$\rownumber$\space}} c c}
+		\multicolumn{1}{@{\makebox[2.5em][c]{idx.}} c}{edge} & $Z^{(0)}$ \\
+		\toprule
+		\rowfont{\color{t_blue}}$e_{11}$ & \vertvec\\ % chktex 21
+		\rowfont{\color{t_blue}}$e_{22}$ & \vertvec\\ % chktex 21
+		\rowfont{\color{t_blue}}$e_{33}$ & \vertvec\\ % chktex 21
+		\rowfont{\color{t_red}}$e_{12}$ & \edgevec\\ % chktex 21
+		\rowfont{\color{t_red}}$e_{13}$ & \edgevec\\ % chktex 21
+		\rowfont{\color{t_red}}$e_{23}$ & \edgevec\\ % chktex 21
+		\midrule
+		\rowfont{\color{t_blue}}$e_{11}'$ & \vertvec\\ % chktex 21
+		\rowfont{\color{t_blue}}$e_{22}'$ & \vertvec\\ % chktex 21
+		\rowfont{\color{t_red}}$e_{12}'$ & \edgevec% chktex 21
+	\end{tabu}\hspace{3em}{\tiny\begin{tabu}{c c c}
+		\rowfont{\footnotesize}$R_{\mathrm{L}}$ & $R_{\Gamma, 1}$ & $R_{\Gamma, 2}$ \\
+		\toprule
+		\rowfont{\color{t_blue}}$1$ & $1$ & $1$ \\
+		\rowfont{\color{t_blue}}$1$ & $4$ & $4$ \\
+		\rowfont{\color{t_blue}}$1$ & $5$ & $5$ \\
+		\rowfont{\color{t_blue}}$2$ & $2$ & $2$ \\
+		\rowfont{\color{t_blue}}$2$ & $4$ & $4$ \\
+		\rowfont{\color{t_blue}}$2$ & $6$ & $6$ \\
+		\rowfont{\color{t_blue}}$3$ & $3$ & $3$ \\
+		\rowfont{\color{t_blue}}$3$ & $5$ & $5$ \\
+		\rowfont{\color{t_blue}}$3$ & $6$ & $6$ \\
+		\rowfont{\color{t_red}}$4$ & $1$ & $4$ \\
+		\rowfont{\color{t_red}}$4$ & $4$ & $2$ \\
+		\rowfont{\color{t_red}}$4$ & $5$ & $6$ \\
+		\rowfont{\color{t_red}}$5$ & $1$ & $5$ \\
+		\rowfont{\color{t_red}}$5$ & $5$ & $3$ \\
+		\rowfont{\color{t_red}}$5$ & $4$ & $6$ \\
+		\rowfont{\color{t_red}}$6$ & $2$ & $6$ \\
+		\rowfont{\color{t_red}}$6$ & $6$ & $3$ \\
+		\rowfont{\color{t_red}}$6$ & $4$ & $5$ \\
+		\midrule
+		\rowfont{\color{t_blue}}$7$ & $7$ & $7$ \\
+		\rowfont{\color{t_blue}}$7$ & $9$ & $9$ \\
+		\rowfont{\color{t_blue}}$8$ & $8$ & $8$ \\
+		\rowfont{\color{t_blue}}$8$ & $9$ & $9$ \\
+		\rowfont{\color{t_red}}$9$ & $7$ & $9$ \\
+		\rowfont{\color{t_red}}$9$ & $9$ & $8$
+	\end{tabu}%
+	}\caption{
+		Exemplary 2-\acs{wl} encoding of a batch of two small graphs.
+	}\label{fig:ltd:wl2-encoding}
+\end{figure}
+After encoding a graph dataset as 2-\acs{wl} matrices, convolutions can be computed efficiently on \acp{gpgpu} via the common \textit{gather-scatter} pattern from parallel programming~\cite{He2007}.
+The so-called $\mathit{gather}$ operator takes two inputs: A list $Z$ of $m$ row vectors and a list $R$ of $\gamma$ pointers into $Z$.
+It returns a list $X$ of $\gamma$ row vectors $X[i] = Z[R[i]]$ for $i \in [\gamma]$.
+The $\mathit{scatter}_{\Sigma}$ operator can be understood as the opposite of $\mathit{gather}$.
+$\mathit{scatter}_{\Sigma}$ takes a list $X$ of $\gamma$ row vectors and a list $R$ of $\gamma$ pointers from the range $[m]$.
+It returns a list $Z$ of $m$ row vectors $Z[i] = \sum_{j \in [\gamma] \land R[j] = i} X[j]$ for $i \in [m]$.
+
+Using the $\mathit{gather}$ and $\mathit{scatter}_{\Sigma}$ operator, the 2-\acs{wl} convolution operator from \cref{defn:ltd:wl2-conv-step} can be computed via the following algorithm:
+\begin{algorithm}[H]
+	\caption{Parallel Implementation of a 2-\acs{wl} Convolution Layer $S^{(t)}$}\label{algo:ltd:wl2-conv}
+	\begin{algorithmic}[1]
+		\Function{$S^{(t)}$}{$Z^{(t-1)} \in \mathbb{R}^{m \times d^{(t-1)}}, R \in {[m]}^{\gamma \times 3}$}
+			\State{$Z_{\mathrm{L}} \coloneqq Z^{(t-1)} W_{\mathrm{L}}^{(t)}$}
+			\Comment{Matrix multiply: $\mathbb{R}^{m \times d^{(t-1)}} \to \mathbb{R}^{m \times d^{(t)}}$}
+			\State{$Z_{\mathrm{F}} \coloneqq Z^{(t-1)} W_{\mathrm{F}}^{(t)}$}
+			\State{$Z_{\Gamma} \coloneqq Z^{(t-1)} W_{\Gamma}^{(t)}$}
+			\State{$X_{\Gamma, 1} \coloneqq \mathit{gather}(Z_{\Gamma}, R_{\Gamma, 1})$}
+			\Comment{Gather: $\mathbb{R}^{m \times d^{(t)}} \times {[m]}^{\gamma} \to \mathbb{R}^{\gamma \times d^{(t)}}$}
+			\State{$X_{\Gamma, 2} \coloneqq \mathit{gather}(Z_{\Gamma}, R_{\Gamma, 2})$}
+			\State{$X_{\Gamma} \coloneqq \sigma_{\Gamma}\left(X_{\Gamma, 1} + X_{\Gamma, 2}\right)$}
+			\Comment{Element-wise operations}
+			\State{$Z_{\Sigma\Gamma} \coloneqq \mathit{scatter}_{\Sigma}(X_{\Gamma}, R_{\mathrm{L}})$}
+			\Comment{Scatter: $\mathbb{R}^{\gamma \times d^{(t)}} \times {[m]}^{\gamma} \to \mathbb{R}^{m \times d^{(t)}}$}
+			\State{$Z^{(t)} \coloneqq \sigma\left(Z_{\mathrm{L}} + Z_{\mathrm{F}} \odot Z_{\Sigma\Gamma} \right)$}
+			\Comment{Element-wise operations}
+			\State{\Return{$Z^{(t)}$}}
+		\EndFunction{}
+	\end{algorithmic}
+\end{algorithm}
+All operations in this implementation scale well on parallel computing hardware and are supported by common \ac{ml} libraries like \citetitle{TF}~\cite{Abadi2015}\cite{TF} or \citetitle{PyT}~\cite{Paszke2019}\cite{PyT}.
+2-\acs{wl} convolutions can therefore be easily integrated into existing \ac{ml} pipelines.
+This concludes our description of 2-\acs{wl}-\acsp{gnn}.

### 2020-04-05 thesis/content/chapter-evaluation.tex
added thesis structure
started with ltag definition + various minor fixes
continued with spatial gcnn lta relation
continued with k-gnn limits section
various fixes + started with evaluation chapter

--- 2019-10-29 thesis/content/chapter-introduction.tex
@@
-% !TEX root = ../main.tex
+%!TEX root = ../main.tex
 % chktex-file 46
-\chapter{Introduction}%
-\label{sec:intro}
+\chapter{Evaluation}%
+\label{sec:eval}
 
-\pagenumbering{arabic}			% arabic page numbering
-\setcounter{page}{1}			% set page counter
+We will now empirically evaluate the ideas presented in the previous chapters.
+In \cref{sec:ltag} the relation between \ac{lta} and existing \ac{gcr} approaches was formally analyzed.
+There we saw that the \acs{lta}-formulations of existing approaches mostly use static decomposition functions, e.g.\ \acs{bfs}-subtree decompositions.
+Motivated by the idea of dynamically learning decompositions via edge filters, we proposed the novel 2-\acs{wl}-\acs{gnn} in the last chapter.
+Note however that a 2-\acs{wl}-\acs{gnn} by itself is not more ``\acs{lta}-like'' than other \acp{gnn}.
 
-\section{Motivation}%
-\label{sec:intro:motivation}
+In order to empirically evaluate the theoretical results of this thesis, we therefore differentiate between two independent evaluation goals:
+\begin{enumerate}[label={\textbf{\arabic*.}}]
+	\item \textbf{Comparison of 2-\acs{wl}-\acsp{gnn} with other \ac{gnn} methods:}
+		Even though it was motivated by \ac{lta}, the 2-\acs{wl} convolution layer should be compared with other convolution approaches in a non-\acs{lta} configuration.
+	\item \textbf{Comparison of \acs{lta}-like \ac{gcr} methods with non-\acs{lta} methods:}
+\end{enumerate}
 
-\section{Goals}%
-\label{sec:intro:goals}
+\section{Experimental Setup}%
+\label{sec:eval:setup}
 
-\section{Structure}%
-\label{sec:intro:structure}
+\section{Synthetic Datasets}%
+\label{sec:eval:synthetic}
 
-\paragraph{\Cref{sec:conclusion}: \nameref{sec:conclusion}}
+\section{Real-World Datasets}%
+\label{sec:eval:real}

### 2020-04-05 thesis/content/chapter-ltd.tex
various fixes + started with evaluation chapter

@@
 	\begin{enumerate}[label={(P\arabic*)}]
-		\item\label{inv:ltd:wl2-simulation:indicator}
+		\item\label[inv]{inv:ltd:wl2-simulation:indicator}
 			${Z_G^{(t,2+K)}[e_{ij}]}_1 = \mathbbm{1}[i = j]$, i.e.\ the first component of each 2-\acs{wl} feature vector allows $\varphi$ to decide whether that vector should be mapped to \texttt{nil}.
-		\item\label{inv:ltd:wl2-simulation:feature}
+		\item\label[inv]{inv:ltd:wl2-simulation:feature}
 			${Z_G^{(t,2+K)}[e_{ii}]}_{2, \dots, (d^{(t)} + 1)} = Z_G^{(t)}[v_{i}]$, i.e.\ the second to $(1 + d^{(t)})$-th components of each self-loop feature vector in $h_2$ contain the corresponding convolved vertex feature vector at layer $t$ in $h_1$.
-		\item\label{inv:ltd:wl2-simulation:weight}
+		\item\label[inv]{inv:ltd:wl2-simulation:weight}
 			${Z_G^{(t,2+K)}[e_{ij}]}_{d^{(t)} + 2} = w_{ij}$, i.e.\ the weights for the vertex neighborhood sums are encoded in the edge and self-loop feature vectors.
@@
 	Additionally for all $e_{ij} \in \mathcal{E}_{G^1}$ the indicators ${Z^{(t,2)}[e_{ij}]}_1 = \mathbbm{1}[i = j]$ and the weights ${Z^{(t,2)}[e_{ij}]}_{d^{(t)} + 2} = w_{ij}$ are preserved.
-	This means that the invariants~\ref{inv:ltd:wl2-simulation:indicator} and~\ref{inv:ltd:wl2-simulation:weight} are satisfied after $S^{(t, 2)}$.
+	This means that \cref{inv:ltd:wl2-simulation:indicator,inv:ltd:wl2-simulation:weight} are satisfied after $S^{(t, 2)}$.
 	To complete the induction step it now remains to show that all three invariants hold after applying the layers $S^{(t, 2+1)}, \dots, S^{(t, 2+K)}$.
 	As previously mentioned, a 2-\acs{wl} convolution layer is reduced to a fully connected layer if $W_{\mathrm{F}}^{(t)} = \mathbf{0}$.
-	Via the universal approximation theorem~\cite{Hornik1991} we can therefore use $S^{(t, 2+1)}, \dots, S^{(t, 2+K)}$ to simulate the $K$ layers of $\mathrm{\acs*{mlp}}^{(t)}$ without changing the first and last dimension of each feature vector to preserve the invariants~\ref{inv:ltd:wl2-simulation:indicator} and~\ref{inv:ltd:wl2-simulation:weight}.
+	Via the universal approximation theorem~\cite{Hornik1991} we can therefore use $S^{(t, 2+1)}, \dots, S^{(t, 2+K)}$ to simulate the $K$ layers of $\mathrm{\acs*{mlp}}^{(t)}$ without changing the first and last dimension of each feature vector to preserve \cref{inv:ltd:wl2-simulation:indicator,inv:ltd:wl2-simulation:weight}.
 	The resulting feature matrix $Z^{(t,2+K)}$ then satisfies all three invariants which completes the induction.
 
-	Using invariants~\ref{inv:ltd:wl2-simulation:indicator} and~\ref{inv:ltd:wl2-simulation:feature} for $t = T$ we can therefore set
+	Using \cref{inv:ltd:wl2-simulation:indicator,inv:ltd:wl2-simulation:feature} for $t = T$ we can therefore set
 	\begin{align*}
@@
 		\toprule
-		\rowfont{\color{t_blue}}$1$ & $1$ & $1$ \\
-		\rowfont{\color{t_blue}}$1$ & $4$ & $4$ \\
-		\rowfont{\color{t_blue}}$1$ & $5$ & $5$ \\
-		\rowfont{\color{t_blue}}$2$ & $2$ & $2$ \\
-		\rowfont{\color{t_blue}}$2$ & $4$ & $4$ \\
-		\rowfont{\color{t_blue}}$2$ & $6$ & $6$ \\
-		\rowfont{\color{t_blue}}$3$ & $3$ & $3$ \\
-		\rowfont{\color{t_blue}}$3$ & $5$ & $5$ \\
-		\rowfont{\color{t_blue}}$3$ & $6$ & $6$ \\
-		\rowfont{\color{t_red}}$4$ & $1$ & $4$ \\
-		\rowfont{\color{t_red}}$4$ & $4$ & $2$ \\
-		\rowfont{\color{t_red}}$4$ & $5$ & $6$ \\
-		\rowfont{\color{t_red}}$5$ & $1$ & $5$ \\
-		\rowfont{\color{t_red}}$5$ & $5$ & $3$ \\
-		\rowfont{\color{t_red}}$5$ & $4$ & $6$ \\
-		\rowfont{\color{t_red}}$6$ & $2$ & $6$ \\
-		\rowfont{\color{t_red}}$6$ & $6$ & $3$ \\
-		\rowfont{\color{t_red}}$6$ & $4$ & $5$ \\
+		\rowfont{\color{t_blue}}$(1,$ & $1,$ & $1)$ \\ % chktex 40
+		\rowfont{\color{t_blue}}$(1,$ & $4,$ & $4)$ \\ % chktex 40
+		\rowfont{\color{t_blue}}$(1,$ & $5,$ & $5)$ \\ % chktex 40
+		\rowfont{\color{t_blue}}$(2,$ & $2,$ & $2)$ \\ % chktex 40
+		\rowfont{\color{t_blue}}$(2,$ & $4,$ & $4)$ \\ % chktex 40
+		\rowfont{\color{t_blue}}$(2,$ & $6,$ & $6)$ \\ % chktex 40
+		\rowfont{\color{t_blue}}$(3,$ & $3,$ & $3)$ \\ % chktex 40
+		\rowfont{\color{t_blue}}$(3,$ & $5,$ & $5)$ \\ % chktex 40
+		\rowfont{\color{t_blue}}$(3,$ & $6,$ & $6)$ \\ % chktex 40
+		\rowfont{\color{t_red}}$(4,$ & $1,$ & $4)$ \\ % chktex 40
+		\rowfont{\color{t_red}}$(4,$ & $4,$ & $2)$ \\ % chktex 40
+		\rowfont{\color{t_red}}$(4,$ & $5,$ & $6)$ \\ % chktex 40
+		\rowfont{\color{t_red}}$(5,$ & $1,$ & $5)$ \\ % chktex 40
+		\rowfont{\color{t_red}}$(5,$ & $5,$ & $3)$ \\ % chktex 40
+		\rowfont{\color{t_red}}$(5,$ & $4,$ & $6)$ \\ % chktex 40
+		\rowfont{\color{t_red}}$(6,$ & $2,$ & $6)$ \\ % chktex 40
+		\rowfont{\color{t_red}}$(6,$ & $6,$ & $3)$ \\ % chktex 40
+		\rowfont{\color{t_red}}$(6,$ & $4,$ & $5)$ \\ % chktex 40
 		\midrule
-		\rowfont{\color{t_blue}}$7$ & $7$ & $7$ \\
-		\rowfont{\color{t_blue}}$7$ & $9$ & $9$ \\
-		\rowfont{\color{t_blue}}$8$ & $8$ & $8$ \\
-		\rowfont{\color{t_blue}}$8$ & $9$ & $9$ \\
-		\rowfont{\color{t_red}}$9$ & $7$ & $9$ \\
-		\rowfont{\color{t_red}}$9$ & $9$ & $8$
+		\rowfont{\color{t_blue}}$(7,$ & $7,$ & $7)$ \\ % chktex 40
+		\rowfont{\color{t_blue}}$(7,$ & $9,$ & $9)$ \\ % chktex 40
+		\rowfont{\color{t_blue}}$(8,$ & $8,$ & $8)$ \\ % chktex 40
+		\rowfont{\color{t_blue}}$(8,$ & $9,$ & $9)$ \\ % chktex 40
+		\rowfont{\color{t_red}}$(9,$ & $7,$ & $9)$ \\ % chktex 40
+		\rowfont{\color{t_red}}$(9,$ & $9,$ & $8)$ % chktex 40
 	\end{tabu}%
@@
 \end{algorithm}
-All operations in this implementation scale well on parallel computing hardware and are supported by common \ac{ml} libraries like \citetitle{TF}~\cite{Abadi2015}\cite{TF} or \citetitle{PyT}~\cite{Paszke2019}\cite{PyT}.
-2-\acs{wl} convolutions can therefore be easily integrated into existing \ac{ml} pipelines.
+All operations in this implementation scale well on parallel computing hardware, they are differentiable and are supported by common \ac{ml} libraries like \citetitle{TF}~\cite{Abadi2015}\cite{TF} and \citetitle{PyT}~\cite{Paszke2019}\cite{PyT}.
+2-\acs{wl} convolutions can therefore be easily integrated into existing \ac{ml} tooling and be optimized via well-known gradient-based methods like \textit{Adam}~\cite{Kingma2015}.
 This concludes our description of 2-\acs{wl}-\acsp{gnn}.

### 2020-04-06 implementation/evaluate_datasets.sh
allow running experiments on CPU (even though this is too slow to be useful for almost all dataset)

@@ -1,11 +1,16 @@
 #!/usr/bin/env bash
 
 USER="-u $(id -u):$(id -g)"
+CUDA_ENV=""
+
+if [ ! -z "$CUDA_VISIBLE_DEVICES" ]; then
+	CUDA_ENV="-e CUDA_VISIBLE_DEVICES=$CUDA_VISIBLE_DEVICES"
+fi
 
 if [ -z "$LTAG_CONTAINER_NAME" ]; then
 	LTAG_CONTAINER_NAME="ltag"
 fi
 
-docker exec -it $USER $(docker ps -aqf "name=^$LTAG_CONTAINER_NAME\$") python3 ./ltag/evaluate_datasets.py $@ \
+docker exec -it $USER $CUDA_ENV $(docker ps -aqf "name=^$LTAG_CONTAINER_NAME\$") python3 ./ltag/evaluate_datasets.py $@ \
 	| grep --line-buffered -vE \
 	"BaseCollectiveExecutor::StartAbort|IteratorGetNext|Shape/|Shape_[0-9]+/"

### 2020-04-06 thesis/content/chapter-evaluation.tex
added experimental setup section

--- 2019-10-29 thesis/content/chapter-introduction.tex
@@
-% !TEX root = ../main.tex
+%!TEX root = ../main.tex
 % chktex-file 46
-\chapter{Introduction}%
-\label{sec:intro}
+\chapter{Evaluation}%
+\label{sec:eval}
 
-\pagenumbering{arabic}			% arabic page numbering
-\setcounter{page}{1}			% set page counter
+In \cref{sec:ltag} the relation between \ac{lta} and existing \ac{gcr} approaches was formally analyzed.
+There we saw that the \ac{lta} formulations of existing approaches mostly use static decomposition functions, e.g.\ \ac{bfs} subtree decompositions.
+Motivated by the idea of dynamically learning decompositions via edge filters, we then proposed the novel 2-\acs{wl}-\acs{gnn} in \cref{sec:ltd}.
+The ides presented in both chapters will now be empirically evaluated.
+Do to so we differentiate between two mostly independent evaluation aspects:
+\begin{enumerate}[label={\textbf{\arabic*.}}]
+	\item \textbf{Evaluation of 2-\acs{wl}-\acsp{gnn}:}
+		Even though it was motivated by \ac{lta}, a 2-\acs{wl}-\acs{gnn} is not generally more ``\acs{lta}-like'' than other approaches.
+		Nonetheless, due to the theoretical advantages described in \cref{sec:ltd:wl2gnn:properties} it is an interesting approach independently from its potential applications in \ac{lta} (see \cref{sec:ltd:edge-filter}).
+		Thus the first aspect of our evaluation is to compare 2-\acs{wl}-\acsp{gnn} with the other previously described \ac{gcr} methods in a general non-\acs{lta} fashion, i.e.\ with an added \ac{mlp} after the pooling layer since this is how \acp{gnn} are typically evaluated in other works.
+	\item \textbf{Evaluation of the \ac{lta} assumption:}
+		We previously described that a given domain problem satisfies the \ac{lta} assumption if its solutions can be described by an \ac{lta} formulation (see \cfullref{defn:ltag:lta-assumption}).
+		The inherent bias of an \acs{lta}-like model towards such \ac{lta} formulations could potentially increase its generalization performance compared to more general non-\acs{lta} models.
+		Therefore the second aspect of our evaluation is to compare the performance of the previously described \acs{lta}-like methods with that of non-\acs{lta} approaches on datasets from multiple problem domains.
+\end{enumerate}
+This chapter will tackle those two aspects in three steps:
+\begin{enumerate*}[label={\circled{\small\arabic*}}]
+	\item We begin by describing the experimental setup used to obtain the evaluation results in \cref{sec:eval:setup}.
+	\item We then present results on synthetically generated data in \cref{sec:eval:synthetic}.
+	 	There we will illustrate the higher expressive power of 2-\acs{wl}-\acsp{gnn} when compared to other \ac{gcr} approaches which confirms the theoretical results from \cref{sec:ltd:wl2gnn:properties}.
+	\item Finally the evaluation results on real-world datasets are described in \cref{sec:eval:real}.
+		There we will see how 2-\acs{wl}-\acsp{gnn} compare to other \acp{gnn} in practice as well as how \acs{lta}-like models compare to non-\acs{lta} models.
+\end{enumerate*}
 
-\section{Motivation}%
-\label{sec:intro:motivation}
+\section{Experimental Setup}%
+\label{sec:eval:setup}
 
-\section{Goals}%
-\label{sec:intro:goals}
+In our experimental evaluation we focus on two types of learners:
+\acp{svm} using graph kernels and \acp{gcnn}.
+We evaluate those learners by comparing their test accuracies on multiple binary classification problems.
+To obtain those accuracies we follow the graph classification benchmarking framework recently proposed by \citet{Errica2020}.
+Their benchmarking framework is motivated by the observation that most recent publications in the field of \acp{gnn} do not provide reproducible results.
+To tackle this issue they evaluated multiple state-of-the-art methods using a unified model selection procedure:\\
+{\setlength{\intextsep}{0pt}
+\begin{minipage}[t]{0.55\linewidth-1em}
+	\begin{algorithm}[H]
+		\caption{$k$-fold Model Assessment}\label{algo:eval:assessment}
+		\begin{algorithmic}[1]
+			\State{\textbf{Input:} Dataset $\mathcal{D}$, configurations $\Theta$}
+			\State{Split $\mathcal{D}$ into $k$ folds $F_1, \dots, F_{k}$}
+			\For{$i \leftarrow 1, \dots, k$}
+				\State{$\mathcal{D}_{\mathrm{train/val}}, \mathcal{D}_{\mathrm{test}} \leftarrow \left( \bigcup_{j \neq i} F_{j} \right), F_i$}
+				\State{$\theta_{\mathrm{best}} \leftarrow \Call{Select}{\mathcal{D}_{\mathrm{train/val}}, \Theta}$}
+				\For{$r \leftarrow 1, \dots, R$}
+					\State{$h_{i,r} \leftarrow \Call{Train}{\mathcal{D}_{\mathrm{train/val}}, \theta_{\mathrm{best}}}$}
+					\State{$\mathit{acc}_{i,r} \leftarrow \Call{Eval}{h_{i,r}, \mathcal{D}_{\mathrm{test}}}$}
+				\EndFor{}
+				\State{$\mathit{acc}_{i} \leftarrow \mean_{r \in [R]}{\mathit{acc}_{i,r}}$}
+			\EndFor{}
+			\State{\Return{$\mean_{i \in [k]} \mathit{acc}_i, \mathrm{stddev}_{i \in [k]}\, \mathit{acc}_i$}}
+		\end{algorithmic}
+	\end{algorithm}
+\end{minipage}\hspace*{1em}%
+\begin{minipage}[t]{0.45\linewidth}
+	\begin{algorithm}[H]
+		\caption{Model Selection}\label{algo:eval:selection}
+		\begin{algorithmic}[1]
+			\Function{Select}{$\mathcal{D}, \Theta$}
+			\State{Split $\mathcal{D}$ into $\mathcal{D}_{\mathrm{train}}, \mathcal{D}_{\mathrm{val}}$}
+			\ForAll{$\theta \in \Theta$}
+				\State{$h_{\theta} \leftarrow \Call{Train}{\mathcal{D}_{\mathrm{train}}, \theta}$}
+				\State{$\mathit{acc}_{\theta} \leftarrow \Call{Eval}{h_{\theta}, \mathcal{D}_{\mathrm{val}}}$}
+			\EndFor{}
+			\State{$\theta_{\mathrm{best}} \leftarrow \arg\max_{\theta \in \Theta}{\mathit{acc}_{\theta}}$}
+			\State{\Return{$\theta_{\mathrm{best}}$}}
+			\EndFunction{}
+		\end{algorithmic}
+	\end{algorithm}
+\end{minipage}}
 
-\section{Structure}%
-\label{sec:intro:structure}
+We use this assessment strategy with $k = 10$ folds and $r = 3$ repeats per fold to smooth out differences caused by random weight initializations.
+For each dataset the same folds are used across the evaluated models; class proportions are preserved within each fold by using stratified splits.
+To keep the total runtime of the experiments feasible, a single 90\%/10\% holdout split into training and validation data is used instead of cross-validation.
+In each experiment training is performed with an early stopping condition which cancels the optimization if there is no improvement to the validation loss for more than $p$ epochs.
+The patience period $p$ is part of the hyperparameter configurations $\theta \in \Theta$.
 
-\paragraph{\Cref{sec:conclusion}: \nameref{sec:conclusion}}
+Using this assessment strategy we evaluate \acp{svm} with the following graph kernels:
+\begin{enumerate}[label={\textbf{\arabic*.}},itemsep=2pt,parsep=2pt]
+	\item \textbf{\ac{wl} subtree kernel (\acs{wl}\textsubscript{ST})} with the iteration counts $T \in \{ 1, 2, 5 \}$ to evaluate the influence of the depth of \ac{bfs} subtrees which span \ac{lta} constituents.
+	\item \textbf{\ac{wl} shortest path kernel (\acs{wl}\textsubscript{SP})} with the iteration count $T = 5$.
+	\item \textbf{2-LWL kernel} with the iteration count $T = 3$.
+	\item \textbf{2-GWL kernel} with the iteration count $T = 3$.
+\end{enumerate}
+The gram matrices of the \acs{wl}\textsubscript{ST} and \acs{wl}\textsubscript{SP} kernels are computed via the \citetitle{GK} library~\cite{Siglidis2018}\cite{GK}.
+For the gram matrices of the two dimensional \ac{wl} kernels we use a modified version\footnote{\url{https://github.com/Cortys/glocalwl}} of the reference implementation provided by \citet{Morris2017}.
+To train \acp{svm} with those kernels \citetitle{SKL}~\cite{Pedregosa2011}\cite{SKL} is used.
+
+For the evaluation of \acp{gcnn} we selected the following methods:
+\begin{enumerate}[label={\textbf{\arabic*.}},itemsep=2pt,parsep=2pt]
+	\item \textbf{Structure unaware baseline:}
+		\citet{Errica2020} describe a simple model which simply applies a standard \ac{mlp} to each individual vertex feature vector, then sums up the resulting feature vectors and applies another \ac{mlp} to the vector sum.
+		This approach does not use any structural information and therefore serves as a baseline to detect whether a \ac{gnn} is able to exploit graph structure.
+	\item \textbf{\ac{gin}} is evaluated as described by \citet{Xu2018}, i.e.\ with a sum pooling layer and an appended \ac{mlp} to produce the final prediction.
+	\item \textbf{2-\acs{gnn}} is evaluated with both a static $\mean$ pooling layer and with \ac{sampool} (see \cfullref{defn:ltag:sam-pool}).
+		After the pooling layer a \ac{mlp} is used to produce the final prediction.
+	\item \textbf{2-\acs{wl}-\ac{gnn}} \textit{(our method)} is evaluated using the same configurations as 2-\acs{gnn}.
+		To test the \ac{lta} assumption we additionally evaluate it in \acs{lta}-like configurations, i.e.\ with a stack of 2-\acs{wl} convolutions that produce a local prediction $y_{ij} \in [0, 1]$ for each edge $e_{ij}$ and without the final \ac{mlp}.
+\end{enumerate}
+The baseline and \ac{gin} results are obtained using the PyTorch-based implementation provided by \citet{Errica2020}.
+For both 2-\acs{gnn} and 2-\acs{wl}-\ac{gnn} a custom TensorFlow-based implementation is used.
+The code for all conducted experiments as well as the used dataset splits can be found on GitHub\footnote{\url{https://github.com/Cortys/master-thesis}}.
+
+\section{Evaluation on Synthetic Data}%
+\label{sec:eval:synthetic}
+
+\section{Evaluation on Real-World Data}%
+\label{sec:eval:real}

### 2020-04-06 thesis/content/chapter-graph-lta.tex
added experimental setup section

@@
 This idea is captured by the so-called \textit{\ac{lta} assumption}.
-\begin{defn}
+\begin{defn}\label{defn:ltag:lta-assumption}
 	A problem domain $\mathcal{D}$ satisfies the \textit{\ac{lta} assumption} iff.\ there is an \ac{lta} method which produces models with an equal or lower out-of-sample error than the models produced by non-\acs{lta} methods for most training samples $\mathcal{D}_{\mathrm{train}} \subseteq \mathcal{D}$.
@@
 Since neither SortPooling nor \ac{sagpool} are compatible with \ac{lta}, we propose an associative weighted aggregation layer inspired by \ac{sagpool} which uses a softmax-based attention filter instead of a top-$p$ filter.
-\begin{defn}
+\begin{defn}\label{defn:ltag:sam-pool}
 	We define \ac{sampool} as

### 2020-04-06 implementation/src/ltag/datasets/disk/manager.py
added dot export support to graph dataset managers

@@ -10,6 +10,7 @@ import zipfile
 from pathlib import Path
 import pickle
 import numpy as np
+import networkx as nx
 
 from ltag.utils import NumpyEncoder
 from ltag.datasets.manager import GraphDatasetManager
@@ -232,6 +233,18 @@ class StoredGraphDatasetManager(GraphDatasetManager):
 
   def _download(self):
     raise NotImplementedError
+
+  def export_dot(self, idx):
+    dot_dir = self.root_dir / "dot"
+
+    if not dot_dir.exists():
+      os.makedirs(dot_dir)
+
+    gs, ys = self.dataset
+    g = gs[idx]
+    y = ys[idx]
+    name = f"g{idx}_{y}.dot"
+    nx.drawing.nx_pydot.write_dot(g, dot_dir / name)
 
 
 class TUDatasetManager(StoredGraphDatasetManager):

### 2020-04-06 implementation/src/ltag/datasets/manager.py

@@ -512,8 +512,9 @@ class GraphDatasetManager:
 
     return train_ds, val_ds
 
-  def draw(self, idx, with_features=False):
+  def draw(self, idx, with_features=False, label_colors=False):
     gs, ys = self.dataset
     g = gs[idx]
     y = ys[idx]
-    return ds_utils.draw_graph(g, y, with_features=with_features)
+    return ds_utils.draw_graph(
+      g, y, with_features=with_features, label_colors=label_colors)

### 2020-04-06 implementation/src/ltag/datasets/utils.py

@@ -530,14 +530,15 @@ def vec_to_unit(feat):
 
   return u
 
-def draw_graph(g, y, with_features=False, with_feature_colors=True):
+def draw_graph(
+  g, y, with_features=False, with_colors=True, label_colors=False):
   plt.figure()
   plt.title('Label: {}'.format(y))
 
   cmap = plt.get_cmap("hsv")
   node_color = [
-    vec_to_unit(d.get("features", []))
-    for n, d in g.nodes(data=True)] if with_feature_colors else "#1f78b4"
+    vec_to_unit([d.get("label", 0)] if label_colors else d.get("features", []))
+    for n, d in g.nodes(data=True)] if with_colors else "#1f78b4"
 
   if with_features:
     labels = {

### 2020-04-07 thesis/content/chapter-appendix.tex
started with ltag definition + various minor fixes
started with synthetic evaluation section + more kernel evals

@@
 \label{sec:appendix}
+
+\section{Dataset Statistics and Descriptions}%
+\label{sec:appendix:ds-stats}
+
+\begin{table}[ht]
+	\caption{Sizes of the evaluated binary classification datasets and their graphs.}\label{tbl:appendix:ds-stats}
+	\csvreader[
+		tabular={lrrrrrrrr},
+		separator=comma,
+		table head={%
+			\multicolumn{1}{c}{} & \multicolumn{1}{c}{} & &\multicolumn{3}{c}{vertex count $\left|\mathcal{V}_G\right|$} & \multicolumn{3}{c}{edge count $\left|\mathcal{E}_G\right|$} \\%
+			& \multirow{-2}{*}[-0.2em]{\shortstack[r]{no.\ of\\ graphs}} & \multirow{-2}{*}[-0.2em]{\shortstack[r]{no.\ of vertex\\ feature dims.}} & $\min$ & $\mean$ & $\max$ & $\min$ & $\mean$ & $\max$ \\\toprule%
+		},
+		table foot=\bottomrule,
+		late after line=\\
+	]{data/ds_stats.csv}%
+	{name=\name,graph_count=\gcount,%
+	node_count_min=\ncountmin,node_count_mean=\ncountmean,node_count_max=\ncountmax,%
+	edge_count_min=\ecountmin,edge_count_mean=\ecountmean,edge_count_max=\ecountmax,%
+	node_degree_min=\ndegmin,node_degree_mean=\ndegmean,node_degree_max=\ndegmax,%
+	dim_node_features=\nfdim,dim_edge_features=\efdim%
+	}%
+	{\textbf{\name}&\gcount&%
+	\nfdim&%
+	\ncountmin&\ncountmean&\ncountmax&%
+	\ecountmin&\ecountmean&\ecountmax%
+	}
+\end{table}
+
+\paragraph{TRIANGLE}
+The triangle detection dataset was generated by sampling three graphs with exactly one unicolored triangle uniformly at random for each possible combination of the following parameters:
+The number of vertices (between 6 and 32), the vertex color proportions (either 50/50\%, 75/25\% or 25/75\% vertices with the color \colorlabel{t_blue}{A}/\colorlabel{t_red}{B}), the graph density (${\left| \mathcal{V}_{G} \right|}^{-2} \left| \mathcal{E}_{G} \right| \in \{ \nicefrac{1}{4}, \nicefrac{1}{2} \}$) the graph class (add a triangle with either the color \colorlabel{t_blue}{A} or \colorlabel{t_red}{B}).

### 2020-04-07 thesis/content/chapter-evaluation.tex
started with synthetic evaluation section + more kernel evals

@@
 	 	There we will illustrate the higher expressive power of 2-\acs{wl}-\acsp{gnn} when compared to other \ac{gcr} approaches which confirms the theoretical results from \cref{sec:ltd:wl2gnn:properties}.
-	\item Finally the evaluation results on real-world datasets are described in \cref{sec:eval:real}.
+	\item Finally evaluation results on real-world datasets are described in \cref{sec:eval:real}.
 		There we will see how 2-\acs{wl}-\acsp{gnn} compare to other \acp{gnn} in practice as well as how \acs{lta}-like models compare to non-\acs{lta} models.
@@
 				\State{$\mathcal{D}_{\mathrm{train/val}}, \mathcal{D}_{\mathrm{test}} \leftarrow \left( \bigcup_{j \neq i} F_{j} \right), F_i$}
-				\State{$\theta_{\mathrm{best}} \leftarrow \Call{Select}{\mathcal{D}_{\mathrm{train/val}}, \Theta}$}
+				\State{Split $\mathcal{D}_{\mathrm{train/val}}$ into $\mathcal{D}_{\mathrm{train}}, \mathcal{D}_{\mathrm{val}}$}
+				\State{$\theta_{\mathrm{best}} \leftarrow \Call{Select}{\mathcal{D}_{\mathrm{train}}, \mathcal{D}_{\mathrm{val}}, \Theta}$}
 				\For{$r \leftarrow 1, \dots, R$}
-					\State{$h_{i,r} \leftarrow \Call{Train}{\mathcal{D}_{\mathrm{train/val}}, \theta_{\mathrm{best}}}$}
+					\State{$h_{i,r} \leftarrow \Call{Train}{\mathcal{D}_{\mathrm{train}}, \theta_{\mathrm{best}}}$}
 					\State{$\mathit{acc}_{i,r} \leftarrow \Call{Eval}{h_{i,r}, \mathcal{D}_{\mathrm{test}}}$}
@@
 		\begin{algorithmic}[1]
-			\Function{Select}{$\mathcal{D}, \Theta$}
-			\State{Split $\mathcal{D}$ into $\mathcal{D}_{\mathrm{train}}, \mathcal{D}_{\mathrm{val}}$}
+			\Function{Select}{$\mathcal{D}_{\mathrm{train}}, \mathcal{D}_{\mathrm{val}}, \Theta$}
 			\ForAll{$\theta \in \Theta$}
@@
 
-We use this assessment strategy with $k = 10$ folds and $r = 3$ repeats per fold to smooth out differences caused by random weight initializations.
+We base our evaluations on this assessment strategy with $k = 10$ folds and $r = 3$ repeats per fold to smooth out differences caused by random weight initializations.
 For each dataset the same folds are used across the evaluated models; class proportions are preserved within each fold by using stratified splits.
 To keep the total runtime of the experiments feasible, a single 90\%/10\% holdout split into training and validation data is used instead of cross-validation.
-In each experiment training is performed with an early stopping condition which cancels the optimization if there is no improvement to the validation loss for more than $p$ epochs.
+In each experiment, training is performed with an early stopping condition which cancels the optimization if there is no improvement to the validation loss for more than $p$ epochs.
 The patience period $p$ is part of the hyperparameter configurations $\theta \in \Theta$.
@@
 For the gram matrices of the two dimensional \ac{wl} kernels we use a modified version\footnote{\url{https://github.com/Cortys/glocalwl}} of the reference implementation provided by \citet{Morris2017}.
-To train \acp{svm} with those kernels \citetitle{SKL}~\cite{Pedregosa2011}\cite{SKL} is used.
-
+To train \acp{svm} with those kernels, \citetitle{SKL}~\cite{Pedregosa2011}\cite{SKL} is used.
 For the evaluation of \acp{gcnn} we selected the following methods:
@@
 For both 2-\acs{gnn} and 2-\acs{wl}-\ac{gnn} a custom TensorFlow-based implementation is used.
-The code for all conducted experiments as well as the used dataset splits can be found on GitHub\footnote{\url{https://github.com/Cortys/master-thesis}}.
+The code for all conducted experiments as well as the used dataset splits are available GitHub\footnote{\url{https://github.com/Cortys/master-thesis}}.
 
@@
 
+We begin with an evaluation of graph kernels and \acp{gnn} on a synthetic binary classification dataset which demonstrates the potential advantages of a higher dimensional \ac{wl} method such as the proposed 2-\acs{wl}-\acs{gnn}.
+To determine the classes of the graphs in this dataset, a learner has to solve the so-called \textit{unicolored triangle detection} problem:
+Given a graph $G$ with vertices that are colored as either $l_G[v] = \colorlabel{t_blue}{A}$ or as $l_G[v] = \colorlabel{t_red}{B}$, the learner has to find the unique triangle $(v_i, v_j, v_k)$ in $G$ for which $l_G[v_i] = l_G[v_j] = l_G[v_k]$.
+% chktex 25
+The class of $G$ is then determined by the color of the vertices $(v_i, v_j, v_k)$.
+
+Based on this problem we generate a synthetic triangle detection dataset.
+It contains randomly generated graphs with varying vertex counts and vertex color proportions (see \cref{sec:appendix:ds-stats} for a more detailed description).
+This dataset allows us to test whether a learner is able to ignore varying amounts of noisy random structure and focus on relevant local substructures, in this case unicolored triangles.
+\Cref{fig:evaluation:triangle-problem} shows two graphs with different classes which both have the same size and equal vertex color proportions.
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=0.75\linewidth]{gfx/evaluation/triangle-problem.pdf}
+	\caption[Two example graphs from the triangle detection dataset.]{
+		Two examples from the triangle detection dataset.
+		The unique unicolored triangle which has to be detected by the learners is highlighted in both graphs.
+	}\label{fig:evaluation:triangle-problem}
+\end{figure}
+
 \section{Evaluation on Real-World Data}%

### 2020-04-07 implementation/src/ltag/datasets/disk/manager.py

@@ -121,11 +121,15 @@ class StoredGraphDatasetManager(GraphDatasetManager):
 
     try:
       gram = super()._compute_gram_matrix(output_fn)
-    except MemoryError:
+    except MemoryError as err:
+      print("Memory error during gram matrix computation:", err)
       gram = "OOM"
 
     with open(gram_filename, "wb") as f:
       pickle.dump(gram, f)
+
+    if gram == "OOM":
+      return None
 
     return gram
 

### 2020-04-07 implementation/src/ltag/datasets/manager.py

@@ -8,6 +8,7 @@ import tensorflow as tf
 import grakel as gk
 from sklearn.model_selection import train_test_split, StratifiedKFold
 
+import ltag.utils as utils
 import ltag.chaining.pipeline as cp
 import ltag.datasets.utils as ds_utils
 
@@ -518,3 +519,25 @@ class GraphDatasetManager:
     y = ys[idx]
     return ds_utils.draw_graph(
       g, y, with_features=with_features, label_colors=label_colors)
+
+  def stats(self):
+    gs, ys = self.dataset
+    node_counts = []
+    edge_counts = []
+    radii = []
+    diameters = []
+    degrees = []
+
+    for g in gs:
+      node_counts.append(g.order())
+      edge_counts.append(g.size())
+      degrees += [d for n, d in g.degree()]
+
+    return dict(
+      node_counts=utils.statistics(node_counts),
+      edge_counts=utils.statistics(edge_counts),
+      node_degrees=utils.statistics(degrees),
+      dim_node_features=self.dim_node_features,
+      dim_edge_features=self.dim_edge_features,
+      num_node_labels=self.num_node_labels,
+      num_edge_labels=self.num_edge_labels)

### 2020-04-07 implementation/src/ltag/evaluate_datasets.py

@@ -83,7 +83,8 @@ if __name__ == "__main__":
       ms = ["SagCWL2GCN_Binary_quick_max"]
     else:
       ms = [
-        "WL_st", "WL_st_1", "WL_st_2", "WL_sp", "LWL2", "GWL2",
+        "WL_st", "WL_st_1", "WL_st_2", "WL_st_3", "WL_st_4",
+        "WL_sp", "WL_sp_3", "LWL2", "GWL2",
         "AvgCWL2GCN_Binary", "SagCWL2GCN_Binary",
         "AvgK2GNN_Binary", "SagK2GNN_Binary"]
   else:

### 2020-04-07 implementation/src/ltag/evaluation/kernel_models.py

@@ -9,6 +9,9 @@ kernel_hps = lambda dsm: cart(C=[1, 0.1, 0.01, 0.001, 0.0001])
 WL_st = kernel_classifier(kernel_models.WL_st)(kernel_hps)
 WL_st_1 = kernel_classifier(kernel_models.WL_st_1)(kernel_hps)
 WL_st_2 = kernel_classifier(kernel_models.WL_st_2)(kernel_hps)
+WL_st_3 = kernel_classifier(kernel_models.WL_st_3)(kernel_hps)
+WL_st_4 = kernel_classifier(kernel_models.WL_st_4)(kernel_hps)
 WL_sp = kernel_classifier(kernel_models.WL_sp)(kernel_hps)
+WL_sp_3 = kernel_classifier(kernel_models.WL_sp_3)(kernel_hps)
 LWL2 = kernel_classifier(kernel_models.LWL2)(kernel_hps)
 GWL2 = kernel_classifier(kernel_models.GWL2)(kernel_hps)

### 2020-04-07 implementation/src/ltag/evaluation/models.py

@@ -7,9 +7,12 @@ import ltag.models.gnn as gnn_models
 from ltag.utils import cart, cart_merge, entry_duplicator
 from ltag.evaluation.model_factories import binary_classifier
 from ltag.evaluation.kernel_models import (
-  WL_st, WL_st_1, WL_st_2, WL_sp, LWL2, GWL2)
+  WL_st, WL_st_1, WL_st_2, WL_st_3, WL_st_4, WL_sp, WL_sp_3, LWL2, GWL2)
 
-kernel_models = [WL_st, WL_st_1, WL_st_2, WL_sp, LWL2, GWL2]
+kernel_models = [
+  WL_st, WL_st_1, WL_st_2, WL_st_3, WL_st_4,
+  WL_sp, WL_sp_3,
+  LWL2, GWL2]
 
 @binary_classifier(gnn_models.AvgWL2GCN)
 def AvgWL2GCN_Binary(dsm):

### 2020-04-07 implementation/src/ltag/evaluation/summary.py

@@ -3,11 +3,11 @@ from __future__ import absolute_import, division, print_function,\
 
 import os
 import json
-from collections import defaultdict, Sized
+from collections import defaultdict
 import funcy as fy
 import numpy as np
 
-from ltag.utils import NumpyEncoder
+from ltag.utils import NumpyEncoder, statistics
 
 selection_metrics = {
   "accuracy": "max",
@@ -20,15 +20,6 @@ selection_metrics = {
 
 def dict_map(f, d):
   return {k: f(v) for k, v in d.items()}
-
-def statistics(vals):
-  return {
-    "mean": np.mean(vals),
-    "std": np.std(vals),
-    "min": np.min(vals),
-    "max": np.max(vals),
-    "count": len(vals) if isinstance(vals, Sized) else 1
-  }
 
 def summarize_evaluation(
   eval_dir, selection_metric="val_accuracy", ignore_worst=0):

### 2020-04-07 implementation/src/ltag/models/kernel.py

@@ -84,7 +84,15 @@ WL_st_1 = kernel_model("WL_st_1", grakel_kernel(
   "WL_st_1", kernel=[{"name": "WL", "n_iter": 1}, "VH"]))
 WL_st_2 = kernel_model("WL_st_2", grakel_kernel(
   "WL_st_2", kernel=[{"name": "WL", "n_iter": 2}, "VH"]))
+WL_st_3 = kernel_model("WL_st_3", grakel_kernel(
+  "WL_st_3", kernel=[{"name": "WL", "n_iter": 3}, "VH"]))
+WL_st_4 = kernel_model("WL_st_4", grakel_kernel(
+  "WL_st_4", kernel=[{"name": "WL", "n_iter": 4}, "VH"]))
+
 WL_sp = kernel_model("WL_sp", grakel_kernel(
   "WL_sp", kernel=[{"name": "WL"}, "SP"]))
+WL_sp_3 = kernel_model("WL_sp_3", grakel_kernel(
+  "WL_sp_3", kernel=[{"name": "WL", "n_iter": 3}, "SP"]))
+
 LWL2 = kernel_model("LWL2", fs_kernel("LWL2"))
 GWL2 = kernel_model("GWL2", fs_kernel("GWL2"))

### 2020-04-07 implementation/src/ltag/playground.py

@@ -11,6 +11,7 @@ import numpy as np
 import sklearn as sk
 import grakel as gk
 import funcy as fy
+import json
 
 import ltag.models.gnn as gnn_models
 import ltag.models.kernel as kernel_models
@@ -18,7 +19,7 @@ import ltag.datasets.synthetic.datasets as synthetic
 import ltag.evaluation.datasets as eval_ds
 import ltag.evaluation.evaluate as evaluate
 import ltag.evaluate_datasets as eval_main
-from ltag.utils import cart, cart_merge, entry_duplicator
+from ltag.utils import cart, cart_merge, entry_duplicator, NumpyEncoder
 
 # -%% codecell
 
@@ -274,10 +275,34 @@ def kernel_experiment():
     print(i)
     print(evaluate.train(model, ds, ds_test, label=f"{dsm.name}_{model.name}").history)
 
+def ds_stats():
+  datasets = [
+    synthetic.balanced_triangle_classification_dataset(stored=True)(),
+    eval_ds.NCI1_8(), eval_ds.Proteins_5(), eval_ds.DD_2(),
+    eval_ds.RedditBinary_1(), eval_ds.IMDB_8()
+    ]
+  s = {}
+
+  for ds in datasets:
+    print(f"DS stats for {ds.name}:")
+    stats = ds.stats()
+    print(stats)
+    s[ds.name] = stats
+    print("-------------------------")
+
+  with open("/data/stats.json", "w") as f:
+    json.dump(s, f, cls=NumpyEncoder, indent="\t")
+  print("Done.")
 
 # wl2_power_experiment()
-synthetic_experiment2()
+# synthetic_experiment2()
 # imdb_experient()
+
+# synthetic.balanced_triangle_classification_dataset(stored=True)().draw(152, label_colors=True)
+# synthetic.balanced_triangle_classification_dataset(stored=True)().draw(153, label_colors=True)
+# synthetic.balanced_triangle_classification_dataset(stored=True)().export_dot(152)
+# synthetic.balanced_triangle_classification_dataset(stored=True)().export_dot(153)
+ds_stats()
 
 #
 # list(dsm.get_all(output_type="grakel")[0])

### 2020-04-07 implementation/src/ltag/utils.py

@@ -6,6 +6,7 @@ import itertools
 import contextlib
 import numpy as np
 import funcy as fy
+from collections import Sized
 
 class NumpyEncoder(json.JSONEncoder):
   def default(self, obj):
@@ -18,6 +19,15 @@ class NumpyEncoder(json.JSONEncoder):
       or isinstance(obj, np.int64)):
       return np.asscalar(obj)
     return json.JSONEncoder.default(self, obj)
+
+def statistics(vals, mask_invalid=False):
+  return {
+    "mean": np.mean(np.ma.masked_invalid(vals) if mask_invalid else vals),
+    "std": np.std(np.ma.masked_invalid(vals) if mask_invalid else vals),
+    "min": np.min(vals),
+    "max": np.max(vals),
+    "count": len(vals) if isinstance(vals, Sized) else 1
+  }
 
 @contextlib.contextmanager
 def local_seed(seed):

### 2020-04-08 implementation/src/ltag/evaluation/kernel_models.py
added single iteration 2-LWL kernel

@@ -14,4 +14,5 @@ WL_st_4 = kernel_classifier(kernel_models.WL_st_4)(kernel_hps)
 WL_sp = kernel_classifier(kernel_models.WL_sp)(kernel_hps)
 WL_sp_3 = kernel_classifier(kernel_models.WL_sp_3)(kernel_hps)
 LWL2 = kernel_classifier(kernel_models.LWL2)(kernel_hps)
+LWL2_1 = kernel_classifier(kernel_models.LWL2_1)(kernel_hps)
 GWL2 = kernel_classifier(kernel_models.GWL2)(kernel_hps)

### 2020-04-08 implementation/src/ltag/evaluation/models.py

@@ -7,12 +7,12 @@ import ltag.models.gnn as gnn_models
 from ltag.utils import cart, cart_merge, entry_duplicator
 from ltag.evaluation.model_factories import binary_classifier
 from ltag.evaluation.kernel_models import (
-  WL_st, WL_st_1, WL_st_2, WL_st_3, WL_st_4, WL_sp, WL_sp_3, LWL2, GWL2)
+  WL_st, WL_st_1, WL_st_2, WL_st_3, WL_st_4, WL_sp, WL_sp_3, LWL2, LWL2_1, GWL2)
 
 kernel_models = [
   WL_st, WL_st_1, WL_st_2, WL_st_3, WL_st_4,
   WL_sp, WL_sp_3,
-  LWL2, GWL2]
+  LWL2, LWL2_1, GWL2]
 
 @binary_classifier(gnn_models.AvgWL2GCN)
 def AvgWL2GCN_Binary(dsm):

### 2020-04-08 implementation/src/ltag/models/kernel.py

@@ -95,4 +95,5 @@ WL_sp_3 = kernel_model("WL_sp_3", grakel_kernel(
   "WL_sp_3", kernel=[{"name": "WL", "n_iter": 3}, "SP"]))
 
 LWL2 = kernel_model("LWL2", fs_kernel("LWL2"))
+LWL2_1 = kernel_model("LWL2_1", fs_kernel("LWL2_1"))
 GWL2 = kernel_model("GWL2", fs_kernel("GWL2"))

### 2020-04-08 thesis/content/chapter-evaluation.tex
added synthetic eval section

@@
 \begin{enumerate}[label={\textbf{\arabic*.}},itemsep=2pt,parsep=2pt]
-	\item \textbf{\ac{wl} subtree kernel (\acs{wl}\textsubscript{ST})} with the iteration counts $T \in \{ 1, 2, 5 \}$ to evaluate the influence of the depth of \ac{bfs} subtrees which span \ac{lta} constituents.
+	\item \textbf{\ac{wl} subtree kernel (\acs{wl}\textsubscript{ST})} with the iteration counts $T \in [5]$ to evaluate the influence of the depth of \ac{bfs} subtrees which span \ac{lta} constituents.
 	\item \textbf{\ac{wl} shortest path kernel (\acs{wl}\textsubscript{SP})} with the iteration count $T = 5$.
@@
 
-Based on this problem we generate a synthetic triangle detection dataset.
+Based on this problem we generated a synthetic triangle detection dataset.
 It contains randomly generated graphs with varying vertex counts and vertex color proportions (see \cref{sec:appendix:ds-stats} for a more detailed description).
-This dataset allows us to test whether a learner is able to ignore varying amounts of noisy random structure and focus on relevant local substructures, in this case unicolored triangles.
-\Cref{fig:evaluation:triangle-problem} shows two graphs with different classes which both have the same size and equal vertex color proportions.
-\begin{figure}[ht]
+We use this dataset to evaluate whether a learner is able to ignore varying amounts of noisy random structure and focus on relevant local substructures, in this case unicolored triangles.
+\begin{table}[t]
+	\caption{Mean accuracies and standard deviations on the triangle detection dataset.}\label{tbl:eval:synthetic}
 	\centering
-	\includegraphics[width=0.75\linewidth]{gfx/evaluation/triangle-problem.pdf}
+	\csvreader[
+		tabular={clrrr},
+		separator=semicolon,
+		table head={%
+			& \multicolumn{2}{l}{Model (Iterations/Pooling)} & \multicolumn{1}{c}{Train} & \multicolumn{1}{c}{Test} \\\toprule%
+		},
+		table foot=\bottomrule,
+		late after line=\ifthenelse{\equal{\id}{9}}{\\\midrule}{\\},
+		head to column names,
+		filter=\equal{\isDefault}{1}\or\equal{\id}{0}
+	]{data/results.csv}{}{%
+		\ifthenelse{\equal{\id}{0}}{\multirow{5}{*}[0em]{\rotatebox[origin=c]{90}{\small\textsc{Kernel}}}}{}%
+		\ifthenelse{\equal{\id}{9}}{\multirow{8}{*}[0em]{\rotatebox[origin=c]{90}{\small\textsc{\ac{gnn}}}}}{} &%
+		\textbf{\model} ($\params$) & \ifthenelse{\equal{\isLta}{1}}{\badgeboxinline[t_green]{\scriptsize \acs*{lta}-like}}{} &%
+		$\condbold{\triangleBestTrain}{\triangleTrainMean} \pm \triangleTrainStd$ & $\condbold{\triangleBestTest}{\triangleTestMean} \pm \triangleTestStd$%
+	}
+\end{table}
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=0.7\linewidth]{gfx/evaluation/triangle-problem.pdf}
 	\caption[Two example graphs from the triangle detection dataset.]{
@@
 \end{figure}
+Looking at the results in \cref{tbl:eval:synthetic}, it can be seen that the structure unaware baseline method is completely unable to do so, as expected.
+The structure aware learners on the other hand all perform better than random guessing and are in fact mostly able to fit the training data perfectly.
+This shows that all generated graphs are 1-\acs{wl} distinguishable; the \ac{wl} subtree kernel for example can simply memorize the training graphs via their unique 1-\acs{wl} color distribution after $T = 5$ refinement steps.
+
+However, the ability to distinguish training graphs is not sufficient to also classify previously unseen graphs correctly.
+Since 1-\acs{wl} cannot detect triangles, all 1-\acs{wl} bounded approaches (\acs{wl}\textsubscript{ST}, \acs{wl}\textsubscript{SP}, Baseline, \acs{gin}) are therefore unable to generalize, as can be seen in their test accuracies.
+The fact that they perform better than random guessing can be explained by the following proxy indicator:
+The presence of an \colorlabel{t_blue}{A}-colored triangle in a graph $G$ implies that there is a local region with a slightly higher density of \colorlabel{t_blue}{A}-colored vertices than in a \colorlabel{t_red}{B}-colored graph $H$ with the same vertex color proportions.
+This local color density difference is already detectable in the depth-1 \acs{bfs} subtrees used by 1-\acs{wl} after a single refinement step; this explains why \acs{wl}\textsubscript{ST} performs similarly for $T = 1$ and $T = 5$.
+
+Let us now take a look at the 2-\acs{wl} inspired kernels: 2-LWL and 2-GWL.\@
+Interestingly both kernels do not appear to generalize better the 1-\acs{wl} bounded methods;
+we explain this by the fairly small size of the triangle detection dataset (228 graphs).
+Even though both kernels embed graphs into a space with dimensions that indicate the presence of a unicolored triangle (see \cfullref{prop:related:wl2-cycle-count}), there are many of those triangle-indicating embedding dimensions s.t.\ the indicating dimensions found in a given training split might not overlap with those in the test split.
+
+Looking at the 2-\acs{wl} inspired \acp{gnn} (2-\acs{gnn}, 2-\acs{wl}-\acs{gnn}), we find that the proposed 2-\acs{wl}-\acs{gnn} significantly outperforms all other evaluated methods which confirms our theoretical results from \cref{sec:ltd:wl2gnn:properties}.
+We evaluated 2-\acs{wl}-\acsp{gnn} in an \acs{lta}-like configuration without a final \ac{mlp} and in a non-\acs{lta} configuration with an \ac{mlp} after the pooling layer.
+Comparing both configurations we clearly see that the \acs{lta}-variant has a worse generalization performance despite the fact that the triangle detection problem fits well into the \ac{lta} framework:
+Find the unicolored triangle constituent, classify it locally then use that local class to classify the graph.
+Further investigations are required to determine to which extent the worse performance of the \acs{lta}-like configuration is caused by the static subtree-based constituents and whether a more dynamic solution to the \acf{ltd} problem could improve the performance (e.g.\ via the edge filters described in \cref{sec:ltd:edge-filter}).
+
+Finally, looking at the two evaluated \acs{lta}-compatible pooling layers, static $\mean$ pooling and \ac{sampool}, we see that the attention mechanism significantly improves the generalization performance of both 2-\acs{gnn} and 2-\acs{wl}-\acs{gnn}.
+This indicates that \ac{sampool} is successful at filtering out the randomly generated noisy parts of a given graph and put most attention to the relevant unicolored triangle.
 

### 2020-04-08 thesis/content/chapter-related-work.tex
added synthetic eval section

@@
 The problem with this approach is that the runtime of \ac{wl} increases exponentially with $k$;
-even for $k = 2$ the cost often becomes infeasibly high when working with large graphs.
+even for $k = 2$ the cost can become infeasibly high when working with large graphs.
 To tackle this problem \citet{Morris2017} proposed a combination of two optimizations:

### 2020-04-08 thesis/content/chapter-evaluation.tex
started with real-world eval section + various improvements

@@
 To tackle this issue they evaluated multiple state-of-the-art methods using a unified model selection procedure:\\
-{\setlength{\intextsep}{0pt}
+{\setlength{\intextsep}{0pt}%
 \begin{minipage}[t]{0.55\linewidth-1em}
@@
 To keep the total runtime of the experiments feasible, a single 90\%/10\% holdout split into training and validation data is used instead of cross-validation.
-In each experiment, training is performed with an early stopping condition which cancels the optimization if there is no improvement to the validation loss for more than $p$ epochs.
+To train models that require gradient-based optimization, we use the well-known Adam optimizer~\cite{Kingma2015}.
+In all experiments, training is performed with an early stopping condition which cancels the optimization if there is no improvement to the validation loss for more than $p$ epochs.
 The patience period $p$ is part of the hyperparameter configurations $\theta \in \Theta$.
@@
 		After the pooling layer a \ac{mlp} is used to produce the final prediction.
-	\item \textbf{2-\acs{wl}-\ac{gnn}} \textit{(our method)} is evaluated using the same configurations as 2-\acs{gnn}.
-		To test the \ac{lta} assumption we additionally evaluate it in \acs{lta}-like configurations, i.e.\ with a stack of 2-\acs{wl} convolutions that produce a local prediction $y_{ij} \in [0, 1]$ for each edge $e_{ij}$ and without the final \ac{mlp}.
+	\item \textbf{2-\acs{wl}-\ac{gnn}} \textit{(our method)} is also evaluated with $\mean$ pooling and \ac{sampool}, just like 2-\ac{gnn}.
+		To test the \ac{lta} assumption we evaluate 2-\acs{wl}-\acs{gnn} in an \acs{lta}-like configuration and in a non-\acs{lta} configuration.
+		The \acs{lta}-like configuration uses a stack of 2-\acs{wl} convolutions that produce a local prediction $y_{ij} \in [0, 1]$ for each edge $e_{ij}$ instead of a local feature vector $z_{ij} \in \mathbb{R}^{d^{(T)}}$ and it does not have an \ac{mlp} after the pooling/aggregation layer.
 \end{enumerate}
@@
 		head to column names,
-		filter=\equal{\isDefault}{1}\or\equal{\id}{0}
+		filter=\equal{\isDefault}{1}
 	]{data/results.csv}{}{%
@@
 		\ifthenelse{\equal{\id}{9}}{\multirow{8}{*}[0em]{\rotatebox[origin=c]{90}{\small\textsc{\ac{gnn}}}}}{} &%
-		\textbf{\model} ($\params$) & \ifthenelse{\equal{\isLta}{1}}{\badgeboxinline[t_green]{\scriptsize \acs*{lta}-like}}{} &%
-		$\condbold{\triangleBestTrain}{\triangleTrainMean} \pm \triangleTrainStd$ & $\condbold{\triangleBestTest}{\triangleTestMean} \pm \triangleTestStd$%
+		\textbf{\ifthenelse{\equal{\isLta}{1}}{\textcolor{t_darkgreen}{\model*}}{\model\ifthenelse{\equal{\model}{2-WL-GNN}}{\phantom{*}}{}}} ($\params$) & \ifthenelse{\equal{\isLta}{1}}{\badgeboxinline[t_green]{\scriptsize \acs*{lta}-like}}{} &%
+		\evalres{\triangleBestTrain}{\triangleTrainMean}{\triangleTrainStd} & \evalres{\triangleBestTest}{\triangleTestMean}{\triangleTestStd}%
 	}
@@
 We evaluated 2-\acs{wl}-\acsp{gnn} in an \acs{lta}-like configuration without a final \ac{mlp} and in a non-\acs{lta} configuration with an \ac{mlp} after the pooling layer.
-Comparing both configurations we clearly see that the \acs{lta}-variant has a worse generalization performance despite the fact that the triangle detection problem fits well into the \ac{lta} framework:
-Find the unicolored triangle constituent, classify it locally then use that local class to classify the graph.
-Further investigations are required to determine to which extent the worse performance of the \acs{lta}-like configuration is caused by the static subtree-based constituents and whether a more dynamic solution to the \acf{ltd} problem could improve the performance (e.g.\ via the edge filters described in \cref{sec:ltd:edge-filter}).
+Comparing both configurations, we see that the \acs{lta}-variant has a worse generalization performance despite the fact that the triangle detection problem fits well into the \ac{lta} framework:
+\begin{enumerate*}[label={\circled{\small\arabic*}}]
+	\item Decomposition corresponds to finding a single unicolored triangle constituent,
+	\item local evaluation then corresponds to checking the color of the triangle constituent and
+	\item aggregation is trivial since there is only one constituent.
+\end{enumerate*}
 
-Finally, looking at the two evaluated \acs{lta}-compatible pooling layers, static $\mean$ pooling and \ac{sampool}, we see that the attention mechanism significantly improves the generalization performance of both 2-\acs{gnn} and 2-\acs{wl}-\acs{gnn}.
-This indicates that \ac{sampool} is successful at filtering out the randomly generated noisy parts of a given graph and put most attention to the relevant unicolored triangle.
+However, as we saw in the \ac{lta} formulation of \acp{gcnn} (see \cfullref{thm:ltag:gcnn-ltag-formulation}), a 2-\acs{wl}-\acs{gnn} does not actually learn to decompose a graph into unicolored triangle constituents but uses subtree constituents instead.
+Further investigations are required to determine to which extent the worse performance of the \acs{lta}-like configuration is caused by this static decomposition strategy and whether a more dynamic solution to the \acf{ltd} problem could improve the performance (e.g.\ via the edge filtering idea described in \cref{sec:ltd:edge-filter}).
+
+Finally, if we look at the two evaluated pooling layers, static $\mean$ pooling and \ac{sampool}, we see that the attention mechanism of the latter significantly improves the generalization performance of both 2-\acs{gnn} and 2-\acs{wl}-\acs{gnn}.
+This indicates that \ac{sampool} is successful at filtering out the randomly generated noisy parts of a given graph and putting most attention to the relevant unicolored triangle.
 
@@
 \label{sec:eval:real}
+
+Since the synthetic triangle detection problem was designed specifically to highlight the advantages of a higher dimensional \ac{wl} method such as 2-\ac{wl}-\acsp{gnn}, we now evaluate the approaches more fairly on five common graph classification benchmark datasets from three different real-world domains.
+
+\begin{table}[ht]
+	\caption[Mean test accuracies and standard deviations on real-world data.]{
+		Mean test accuracies and standard deviations on real-world data.
+		\hfill\textcolor{t_darkgreen}{*}\,\badgeboxinline[t_green]{\scriptsize\acs*{lta}-like}
+	}\label{tbl:eval:real}
+	\centering
+	\csvreader[
+		tabular={lrrrrr},
+		separator=semicolon,
+		table head={%
+			& %
+			\multicolumn{1}{c}{\textbf{NCI1}} &%
+			\multicolumn{1}{c}{\textbf{PROTEINS}} &%
+			\multicolumn{1}{c}{\textbf{D\&D}} &%
+			\multicolumn{1}{c}{\textbf{REDDIT}} &%
+			\multicolumn{1}{c}{\textbf{IMDB}}%
+			\\\toprule%
+		},
+		table foot=\bottomrule,
+		late after line=\ifthenelse{\equal{\id}{9}}{\\\midrule}{\\},
+		head to column names,
+		filter={\equal{\isDefault}{1}}
+	]{data/results.csv}{}{%
+		% \ifthenelse{\equal{\id}{0}}{\multirow{5}{*}[0em]{\rotatebox[origin=c]{90}{\small\textsc{Kernel}}}}{}%
+		% \ifthenelse{\equal{\id}{9}}{\multirow{8}{*}[0em]{\rotatebox[origin=c]{90}{\small\textsc{\ac{gnn}}}}}{} &%
+		\textbf{\ifthenelse{\equal{\isLta}{1}}{\textcolor{t_darkgreen}{\model*}}{\model\ifthenelse{\equal{\model}{2-WL-GNN}}{\phantom{*}}{}}} {\footnotesize($\params$)} &%
+		{\small\evalres{\nciBestTest}{\nciTestMean}{\nciTestStd}}&%
+		{\small\evalres{\proteinsBestTest}{\proteinsTestMean}{\proteinsTestStd}}&%
+		{\small\evalres{\ddBestTest}{\ddTestMean}{\ddTestStd}}&%
+		{\small\evalres{\redditBestTest}{\redditTestMean}{\redditTestStd}}&%
+		{\small\evalres{\imdbBestTest}{\imdbTestMean}{\imdbTestStd}}%
+	}
+\end{table}

### 2020-04-08 thesis/content/chapter-ltd.tex
started with real-world eval section + various improvements

@@
 	Via the universal approximation theorem~\cite{Hornik1991} we can therefore use $S^{(t, 2+1)}, \dots, S^{(t, 2+K)}$ to simulate the $K$ layers of $\mathrm{\acs*{mlp}}^{(t)}$ without changing the first and last dimension of each feature vector to preserve \cref{inv:ltd:wl2-simulation:indicator,inv:ltd:wl2-simulation:weight}.
-	The resulting feature matrix $Z^{(t,2+K)}$ then satisfies all three invariants which completes the induction.
+	The resulting feature matrix $Z^{(t,2+K)}$ then satisfies all three invariants, which completes the induction.
 
@@
 \begin{proof}
-	The proposition follows if we choose the six-cycle graph for $G$ and the two three-cycles graph for $H$, both of which should be well-known at this point.
+	The proposition follows if we choose the six-cycle graph for $G$ and the two three-cycles graph for $H$, both of which should be familiar at this point.
 	Let $h_2 = \mathit{Pool} \circ S$ be a 2-\acs{wl}-\acs{gnn} with the neighborhood radius $r = 1$ which consists of a single 2-\acs{wl} convolution layer $S: \mathbb{R}^{* \times 2} \to \mathbb{R}^{* \times 1}$ and the pooling layer $\mathit{Pool} = \min$.
@@
 
-	By \cref{defn:ltd:wl2-conv-step} all self-loops $e_{ii}$ of $G^1$ and $H^1$ have the three neighbors $\{ \textcolor{t_blue}{\ldblbrace e_{ii}, e_{ii} \rdblbrace},\allowbreak \textcolor{t_red}{\ldblbrace e_{ij}, e_{ji} \rdblbrace},\allowbreak \textcolor{t_darkgreen}{\ldblbrace e_{il}, e_{li} \rdblbrace} \}$, i.e.\ the length-two walk along $e_{ii}$ itself and the length-two walks to and from the two neighboring vertices $\Gamma(v_i) = \{ v_j, v_l \}$.
+	By \cref{defn:ltd:wl2-conv-step}, all self-loops $e_{ii}$ of $G^1$ and $H^1$ have the three neighbors $\{ \textcolor{t_blue}{\ldblbrace e_{ii}, e_{ii} \rdblbrace},\allowbreak \textcolor{t_red}{\ldblbrace e_{ij}, e_{ji} \rdblbrace},\allowbreak \textcolor{t_darkgreen}{\ldblbrace e_{il}, e_{li} \rdblbrace} \}$, i.e.\ the length-two walk along $e_{ii}$ itself and the length-two walks to and from the two neighboring vertices $\Gamma(v_i) = \{ v_j, v_l \}$.
 % chktex 25
@@
 	The different neighborhood sizes of the edges of $G$ and $H$ imply that $\forall e_{ij} \in \mathcal{E}_G: Z^{(1)}[e_{ij}] = 4$ while $\forall e_{ij}' \in \mathcal{E}_H: Z^{(1)}[e_{ij}] = 6$.
-	Thus $h_2(G) = \min\{ 4, 6 \} \neq \min\{ 6, 6 \} = h_2(H)$ which concludes the proof. % chktex 21
+	Thus $h_2(G) = \min\{ 4, 6 \} \neq \min\{ 6, 6 \} = h_2(H)$, which concludes the proof. % chktex 21
 \end{proof}
@@
 
-Using the $\mathit{gather}$ and $\mathit{scatter}_{\Sigma}$ operator, the 2-\acs{wl} convolution operator from \cref{defn:ltd:wl2-conv-step} can be computed via the following algorithm:
+Using the $\mathit{gather}$ and $\mathit{scatter}_{\Sigma}$ operators, the 2-\acs{wl} convolution operator from \cref{defn:ltd:wl2-conv-step} can be computed via the following algorithm:
 \begin{algorithm}[H]

### 2020-04-09 thesis/content/chapter-appendix.tex
added synthetic eval section
continued with real-world section + various improvements

--- 2019-10-29 thesis/content/chapter-appendix.tex
@@
 \label{sec:appendix}
+
+\section{Evaluated Hyperparameter Grids}%
+\label{sec:appendix:config-grid}
+
+\section{Dataset Statistics and Descriptions}%
+\label{sec:appendix:ds-stats}
+
+\begin{table}[ht]
+	\caption{Sizes of the evaluated binary classification datasets and their graphs.}\label{tbl:appendix:ds-stats}
+	\centering
+	\csvreader[
+		tabular={lrrrrrrrr},
+		separator=comma,
+		table head={%
+			\multicolumn{1}{c}{} & \multicolumn{1}{c}{} & &\multicolumn{3}{c}{vertex count $\left|\mathcal{V}_G\right|$} & \multicolumn{3}{c}{edge count $\left|\mathcal{E}_G\right|$} \\%
+			& \multirow{-2}{*}[-0.2em]{\shortstack[c]{no.\ of\\ graphs}} & \multirow{-2}{*}[-0.2em]{\shortstack[c]{vertex data\\(feat.\ + lab.)}} & $\min$ & $\mean$ & $\max$ & $\min$ & $\mean$ & $\max$ \\\toprule%
+		},
+		table foot=\bottomrule,
+		late after line=\\
+	]{data/ds_stats.csv}%
+	{name=\name,graph_count=\gcount,%
+	node_count_min=\ncountmin,node_count_mean=\ncountmean,node_count_max=\ncountmax,%
+	edge_count_min=\ecountmin,edge_count_mean=\ecountmean,edge_count_max=\ecountmax,%
+	node_degree_min=\ndegmin,node_degree_mean=\ndegmean,node_degree_max=\ndegmax,%
+	dim_node_features=\nfdim,dim_edge_features=\efdim%
+	}%
+	{\textbf{\name}&%
+	$\gcount$&%
+	$\nfdim$&%
+	$\ncountmin$&$\ncountmean$&$\ncountmax$&%
+	$\ecountmin$&$\ecountmean$&$\ecountmax$%
+	}
+\end{table}
+
+\paragraph{TRIANGLE}
+The triangle detection dataset was generated by sampling three graphs with exactly one unicolored triangle uniformly at random for each possible combination of the following parameters:
+The number of vertices (between 6 and 32), the vertex color proportions (either 50/50\%, 75/25\% or 25/75\% vertices with the colors \colorlabel{t_blue}{A}/\colorlabel{t_red}{B}), the graph density (${\left| \mathcal{V}_{G} \right|}^{-2} \left| \mathcal{E}_{G} \right| \in \{ \nicefrac{1}{4}, \nicefrac{1}{2} \}$) the graph class (add a triangle with either the color \colorlabel{t_blue}{A} or \colorlabel{t_red}{B}).
+
+\paragraph{NCI1}
+This dataset was made available by \citet{Shervashidze2011}.
+It contains a balanced subset of molecule graphs that were originally published by the US \ac{nci}~\cite{Wale2007}.
+In each molecule graph, vertices correspond to atoms and edges to bonds between them.
+The binary classes in this dataset describe whether a molecule is able to suppress or inhibit the growth of certain lung cancer and ovarian cancer cell lines in humans.
+
+\paragraph{PROTEINS and D\&D}
+The graphs in the PROTEINS dataset represent proteins~\cite{Borgwardt2005a}.
+Each vertex corresponds to a so-called \ac{sse}, i.e.\ a certain molecular substructure.
+An edge encodes either that two \acp{sse} are neighbors in the protein's amino-acid sequence or that those \acp{sse} are close to each other in 3D space.
+Each protein graph is classified by whether it is an enzyme or not.
+Very similarly, the D\&D dataset~\cite{Dobson2003} is also about classifying proteins as enzymes or non-enzymes but it uses a different selection of vertex features.
+
+\paragraph{REDDIT}
+This balanced dataset contains graphs that represent online discussion threads on the website Reddit~\cite{Yanardag2015}.
+Each vertex corresponds to a user; an edge is drawn between two users iff.\ at least one of them replied to a comment of another.
+Such social interaction graphs were sampled from two types of subreddits:
+Question/answer-based and discussion-based.
+The classification goal is to predict from which type of subreddit a given graph was sampled.
+
+\paragraph{IMDB}
+This dataset contains so-called \textit{ego-networks} of movie actors~\cite{Yanardag2015}.
+Vertices in such networks represent actors and edges encode whether two actors starred in the same movie.
+The graphs in the dataset are derived from the actors starring in either action or romance movies.
+The classification goal for each graph is to predict the movie genre it was derived from.
+
+\section{Fold-wise Accuracy Deltas}%
+\label{sec:appendix:fold-diffs}

### 2020-04-09 thesis/content/chapter-evaluation.tex
continued with real-world section + various improvements

@@
 To keep the total runtime of the experiments feasible, a single 90\%/10\% holdout split into training and validation data is used instead of cross-validation.
-To train models that require gradient-based optimization, we use the well-known Adam optimizer~\cite{Kingma2015}.
+To train models that require gradient-based optimization, we use the well-known Adam optimizer~\cite{Kingma2015} and the standard binary crossentropy loss.
 In all experiments, training is performed with an early stopping condition which cancels the optimization if there is no improvement to the validation loss for more than $p$ epochs.
@@
 \begin{enumerate}[label={\textbf{\arabic*.}},itemsep=2pt,parsep=2pt]
-	\item \textbf{\ac{wl} subtree kernel (\acs{wl}\textsubscript{ST})} with the iteration counts $T \in [5]$ to evaluate the influence of the depth of \ac{bfs} subtrees which span \ac{lta} constituents.
+	\item \textbf{\ac{wl} subtree kernel (\acs{wl}\textsubscript{ST})} with the iteration counts $T \in \{ 1,2,3,4,5 \}$ to evaluate the influence of the depth of \ac{bfs} subtrees which span \ac{lta} constituents.
 	\item \textbf{\ac{wl} shortest path kernel (\acs{wl}\textsubscript{SP})} with the iteration count $T = 5$.
@@
 	\item \textbf{2-\acs{wl}-\ac{gnn}} \textit{(our method)} is also evaluated with $\mean$ pooling and \ac{sampool}, just like 2-\ac{gnn}.
-		To test the \ac{lta} assumption we evaluate 2-\acs{wl}-\acs{gnn} in an \acs{lta}-like configuration and in a non-\acs{lta} configuration.
+		To test the \ac{lta} assumption we evaluate 2-\acs{wl}-\acs{gnn} in an \acs{lta}-like configuration and in the standard non-\acs{lta} configuration.
 		The \acs{lta}-like configuration uses a stack of 2-\acs{wl} convolutions that produce a local prediction $y_{ij} \in [0, 1]$ for each edge $e_{ij}$ instead of a local feature vector $z_{ij} \in \mathbb{R}^{d^{(T)}}$ and it does not have an \ac{mlp} after the pooling/aggregation layer.
@@
 The code for all conducted experiments as well as the used dataset splits are available GitHub\footnote{\url{https://github.com/Cortys/master-thesis}}.
+The evaluated hyperparameter configurations $\Theta$ are described in \cref{sec:appendix:config-grid}.
 
@@
 We use this dataset to evaluate whether a learner is able to ignore varying amounts of noisy random structure and focus on relevant local substructures, in this case unicolored triangles.
+For the evaluation of 2-\acs{wl}-\acsp{gnn}, the neighborhood radius $r = 2$ is used.
 \begin{table}[t]
@@
 		Two examples from the triangle detection dataset.
-		The unique unicolored triangle which has to be detected by the learners is highlighted in both graphs.
+		The unique unicolored triangle, which has to be detected by the learners, is highlighted in both graphs.
 	}\label{fig:evaluation:triangle-problem}
 \end{figure}
-Looking at the results in \cref{tbl:eval:synthetic}, it can be seen that the structure unaware baseline method is completely unable to do so, as expected.
+
+Looking at the results in \cref{tbl:eval:synthetic}, it can be seen that the structure unaware baseline method is completely unable to detect triangles, as expected.
 The structure aware learners on the other hand all perform better than random guessing and are in fact mostly able to fit the training data perfectly.
-This shows that all generated graphs are 1-\acs{wl} distinguishable; the \ac{wl} subtree kernel for example can simply memorize the training graphs via their unique 1-\acs{wl} color distribution after $T = 5$ refinement steps.
+This shows that all generated graphs are 1-\acs{wl} distinguishable; the \ac{wl} subtree kernel, for example, can simply memorize the training graphs via their unique 1-\acs{wl} color distribution after $T = 5$ refinement steps.
 
@@
 
-Finally, if we look at the two evaluated pooling layers, static $\mean$ pooling and \ac{sampool}, we see that the attention mechanism of the latter significantly improves the generalization performance of both 2-\acs{gnn} and 2-\acs{wl}-\acs{gnn}.
+Finally, if we look at the two evaluated pooling layers, static $\mean$ pooling and \ac{sampool}, we see that the attention mechanism of the latter significantly improves the generalization performance of both 2-\acsp{gnn} and 2-\acs{wl}-\acsp{gnn}.
 This indicates that \ac{sampool} is successful at filtering out the randomly generated noisy parts of a given graph and putting most attention to the relevant unicolored triangle.
@@
 \label{sec:eval:real}
-
-Since the synthetic triangle detection problem was designed specifically to highlight the advantages of a higher dimensional \ac{wl} method such as 2-\ac{wl}-\acsp{gnn}, we now evaluate the approaches more fairly on five common graph classification benchmark datasets from three different real-world domains.
 
@@
 \end{table}
+Since the synthetic triangle detection problem was designed specifically to highlight the advantages of a higher dimensional \ac{wl} method such as 2-\ac{wl}-\acsp{gnn}, we now evaluate the approaches more fairly on five common graph classification benchmark datasets from two different real-world domains:
+\begin{enumerate}[label={\textbf{\arabic*.}}]
+	\item \textbf{Bioinformatics:}
+		The NCI1~\cite{Shervashidze2011}, PROTEINS~\cite{Borgwardt2005a} and D\&D~\cite{Dobson2003} datasets all contain molecular graphs.
+		For NCI1 the goal is to predict whether a molecule is effective against certain types of cancer.
+		For PROTEINS and D\&D, a learner has to determine whether a given protein molecule is an enzyme.
+	\item \textbf{Social networks:}
+		The REDDIT and IMDB datasets~\cite{Yanardag2015} contain graphs representing the relations between users in online discussions and movie actors respectively.
+		For REDDIT the type of a discussion thread has to be predicted; for IMDB the goal is to determine movie genres.
+\end{enumerate}
+A more detailed description of the evaluated datasets can be found in \cref{sec:appendix:ds-stats}.
+\Cref{tbl:eval:real} shows our evaluation results\footnote{
+	Note that, for the NCI1, PROTEINS and D\&D datasets we used different training/test splits than those originally used by \citet{Errica2020}.
+	The reason for this is that the original splits were only published after our evaluations of those datasets had already been completed.
+	For REDDIT and IMDB however, we used the same splits.
+}.
+For the evaluation of 2-\acs{wl}-\acsp{gnn}, different neighborhood radii were used for each dataset.
+In the order of the columns in the table above, the results were obtained with the radii $r = 8$, $5$, $2$, $1$ and $4$ respectively.
+
+If we compare the real-world results with those for the synthetic triangle detection dataset, we do no longer observe the clear advantage of 2-\acs{wl}-\acsp{gnn} over the other approaches.
+This indicates that the theoretical advantages of 2-\acs{wl} over 1-\acs{wl} are not necessarily relevant for the five evaluated problems.
+Nonetheless, the test performance of 2-\acs{wl}-\acsp{gnn} is generally comparable to that of the other state-of-the-art learners with the exception of NCI1.
+If we look at the enzyme detection problem (PROTEINS and D\&D), we observe that all evaluated approaches appear to be unable to leverage structural information for a significant improvement over the baseline learner.
+In the social network datasets (REDDIT and IMDB) on the other hand, the structure aware methods clearly outperform the baseline.
+This confirms the very similar results of \citet{Errica2020}.

### 2020-04-09 thesis/content/chapter-graph-lta.tex
started with real-world eval section + various improvements
continued with real-world section + various improvements

@@
 To define an \ac{lta} formulation for \acp{gcnn}, we define two functions that are entirely based on the stack of convolution layers ($\mathit{Conv}$):
-\begin{enumerate*}
+\begin{enumerate*}[label={\circled{\small\arabic*}}]
 	\item The decomposition function $\psi: \mathcal{G} \to \mathcal{P}(\mathcal{G})$.
@@
 	A pooling layer $\mathit{Pool}: \mathbb{R}^{n \times d^{(T)}} \to \mathbb{R}^{d^{(\mathrm{pool})}}$ is called an \textit{associative weighted aggregation layer} iff.
-	\begin{enumerate*}
+	\begin{enumerate*}[label={\circled{\small\arabic*}}]
 		\item there is an associative algebra of $\mathit{Pool}$ and
@@
 		Apart from such unweighted pooling layers we also looked at two state-of-the-art graph pooling layers:
-		\begin{enumerate*}[label={\arabic*.}]
-			\item SortPooling~\cite{Zhang2018} works by aggregating the top-$p$ convolved vectors $z_i \in \mathbb{R}^{d^{(T)}}$ based on a lexicographic ordering $\prec$,
-			\item \ac{sagpool}~\cite{Lee2019} also aggregates the top-$p$ vectors $z_i$ but determines their ordering $\prec$ based on attention weights which are encoded as one component of each $z_i$
+		\begin{enumerate*}[label={\circled{\small\arabic*}}]
+			\item \textit{SortPooling}~\cite{Zhang2018} works by aggregating the top-$p$ convolved vectors $z_i \in \mathbb{R}^{d^{(T)}}$ based on a lexicographic ordering $\prec$,
+			\item \textit{\ac{sagpool}}~\cite{Lee2019} also aggregates the top-$p$ vectors $z_i$ but determines their ordering $\prec$ based on attention weights which are encoded as one component of each $z_i$
 		\end{enumerate*}.
@@
 			&& \mathcal{A}(\mathit{top}_{2}(\ldblbrace z_1, z_2, z_2 \rdblbrace)) &= \mathcal{A}(\mathit{top}_{2}(\ldblbrace z_1, \rho(\ldblbrace z_2, z_2 \rdblbrace) \rdblbrace))&& \\
-			\Leftrightarrow&& z_2 &= \mathcal{A}(\ldblbrace z_1, \rho(\ldblbrace z_2, z_2 \rdblbrace) \rdblbrace)&&\lightning\text{ if $\mathcal{A} \neq \max$.}
+			\Leftrightarrow&& z_2 &= \mathcal{A}(\ldblbrace z_1, \rho(\ldblbrace z_2, z_2 \rdblbrace) \rdblbrace)&&\lightning\text{ for $\mathcal{A} = \min$.}
 		\end{alignat*}
-		From the \ac{lta} perspective the non-associativity of SortPooling and \ac{sagpool} implies that, if they were used in a \ac{gcnn}, it could happen that some subtree evaluations $(y_i, w_i) \in f^{*}(c)$ of a single constituent $c$ are considered by the aggregation function $\mathcal{A}$ while some are ignored depending on the subtree evaluations of other constituents, i.e.\ the constituent evaluation $f(c)$ could then depend on other constituents.
+		From an \ac{lta} perspective the non-associativity of SortPooling and \ac{sagpool} implies that, if they were used in a \ac{gcnn}, it could happen that some subtree evaluations $(y_i, w_i) \in f^{*}(c)$ of a single constituent $c$ are considered by the aggregation function $\mathcal{A}$ while some are ignored depending on the subtree evaluations of other constituents, i.e.\ the constituent evaluation $f(c)$ could then depend on other constituents.
 		This is forbidden in our definition of \ac{lta}.

### 2020-04-09 thesis/content/chapter-ltd.tex
continued with real-world section + various improvements

@@
 	Using the two layers $S^{(t, 1)}$ and $S^{(t, 2)}$ that we just defined, the weighted vertex neighborhood sum for all $v_i \in \mathcal{V}_G$ is contained in $Z^{(t,2)}[e_{ii}]$.
-	Additionally for all $e_{ij} \in \mathcal{E}_{G^1}$ the indicators ${Z^{(t,2)}[e_{ij}]}_1 = \mathbbm{1}[i = j]$ and the weights ${Z^{(t,2)}[e_{ij}]}_{d^{(t)} + 2} = w_{ij}$ are preserved.
+	Additionally for all $e_{ij} \in \mathcal{E}_{G^1}$, the indicators ${Z^{(t,2)}[e_{ij}]}_1 = \mathbbm{1}[i = j]$ and the weights ${Z^{(t,2)}[e_{ij}]}_{d^{(t)} + 2} = w_{ij}$ are preserved.
 	This means that \cref{inv:ltd:wl2-simulation:indicator,inv:ltd:wl2-simulation:weight} are satisfied after $S^{(t, 2)}$.
-	To complete the induction step it now remains to show that all three invariants hold after applying the layers $S^{(t, 2+1)}, \dots, S^{(t, 2+K)}$.
+	To complete the induction step, it now remains to show that all three invariants hold after applying the layers $S^{(t, 2+1)}, \dots, S^{(t, 2+K)}$.
 	As previously mentioned, a 2-\acs{wl} convolution layer is reduced to a fully connected layer if $W_{\mathrm{F}}^{(t)} = \mathbf{0}$.

### 2020-04-09 implementation/src/ltag/evaluate_datasets.py

@@ -30,7 +30,7 @@ def resume(mf, dsm, **kwargs):
 
   return evaluate.resume_evaluation(mf, dsm, **kwargs)
 
-def summarize(mf, dsm):
+def summarize(mf, dsm, **kwargs):
   if dsm.evaluation_args is not None:
     kwargs = fy.project(dsm.evaluation_args, ["ignore_worst"])
   else:
@@ -86,7 +86,8 @@ if __name__ == "__main__":
         "WL_st", "WL_st_1", "WL_st_2", "WL_st_3", "WL_st_4",
         "WL_sp", "WL_sp_3", "LWL2", "GWL2",
         "AvgCWL2GCN_Binary", "SagCWL2GCN_Binary",
-        "AvgK2GNN_Binary", "SagK2GNN_Binary"]
+        "AvgCWL2GCN_FC_Binary", "SagCWL2GCN_FC_Binary",
+        "AvgK2GNN_FC_Binary", "SagK2GNN_FC_Binary"]
   else:
     ms = args.model
 
@@ -117,10 +118,19 @@ if __name__ == "__main__":
   print("----------------------------------------------------------\n")
 
   for m in ms:
+    split_m = m.split("?", 1)
+
+    if len(split_m) == 2:
+      m, hp_i = split_m
+      hp_i = int(hp_i)  # Evaluate only a single hp.
+    else:
+      hp_i = None  # Evaluate all hps.
+
     model = getattr(models, m)
     for d in ds:
       dsm = getattr(datasets, d)
-      f(model, dsm())
+      print(f"- Model: {m}, Dataset: {d}. -")
+      f(model, dsm(), single_hp=hp_i)
       print("\n----------------------------------------------------------\n")
 
   print(f"Grid evaluation (# datasets = {dsl}, # models = {msl}) completed.")

### 2020-04-09 implementation/src/ltag/evaluation/evaluate.py

@@ -121,7 +121,7 @@ def evaluate(
   outer_k=None, repeat=1, winner_repeat=3, epochs=1000,
   patience=100, stopping_min_delta=0.0001,
   restore_best=False, hp_args=None, label=None,
-  eval_dir=None, verbose=2, dry=False, ignore_worst=0):
+  eval_dir=None, verbose=2, dry=False, ignore_worst=0, single_hp=None):
   outer_k = outer_k or ds_manager.outer_k
   inner_k = None
 
@@ -190,6 +190,10 @@ def evaluate(
   i_start = -1
   print(t, f"- Evaluating {ds_name} using {mf_name}...")
 
+  if single_hp is not None:
+    print(f"!!! USING SINGLE HP MODE FOR HYPERPARAM {single_hp} !!!")
+    print("The resulting mean accuracy might be negatively affected.")
+
   if resume and pos_file.exists():
     pos = pos_file.read_text().split(",")
     if len(pos) == 3:
@@ -235,6 +239,11 @@ def evaluate(
       for hp_i, hp in enumerate(hps):
         hp_str = f"{hp_i+1}/{hpc}"
         curr_i_start = 0
+
+        if single_hp != hp_i:
+          print(f"Skipping {fold_str} with hp {hp_str} due to single hp mode.")
+          continue
+
         if k == k_start:
           if hp_i < hp_start:
             print(f"Already evaluated {fold_str} with hyperparams {hp_str}.")
@@ -252,19 +261,21 @@ def evaluate(
             model_ctr, train_ds, val_ds, test_ds, k, hp_i, i, hp,
             res_dir, fold_str, hp_str, verbose, log_dir_base,
             **config)
-          pos_file.write_text(f"{k},{hp_i},{i}")
+          if single_hp is None:
+            pos_file.write_text(f"{k},{hp_i},{i}")
 
       t_end_fold = timer()
       dur_fold = t_end_fold - t_start_fold
+      summ = summary.summarize_evaluation(
+        eval_dir, ignore_worst=ignore_worst)
       print(time_str(), f"- Evaluated hps of fold {fold_str} in {dur_fold}s.")
 
       if winner_repeat > repeat:
-        if hp_start == hpc and i_start + 1 == winner_repeat:
+        if single_hp is not None:
+          print(f"No repeats because of single hp mode for hp={single_hp}.")
+        elif hp_start == hpc and i_start + 1 == winner_repeat:
           print(f"Already did winner evaluations of fold {fold_str}.")
[…]

### 2020-04-09 implementation/src/ltag/evaluation/summary.py

@@ -23,7 +23,9 @@ def dict_map(f, d):
 
 def summarize_evaluation(
   eval_dir, selection_metric="val_accuracy", ignore_worst=0):
-  assert eval_dir.exists(), f"No evalutation '{eval_dir}' found."
+  if not eval_dir.exists():
+    print(f"No evalutation '{eval_dir}' found.")
+    return
 
   with open(eval_dir / "config.json") as f:
     config = json.load(f)
@@ -52,11 +54,15 @@ def summarize_evaluation(
   best_goal = selection_metrics[selection_metric]
 
   results = []
+  all_hps = True
 
   for fold_i, param_files in folds:
     best_res = None
+    param_file_items = list(param_files.items())
 
-    for hp_i, files in param_files.items():
+    all_hps = all_hps and len(param_files) == len(hps)
+
+    for hp_i, files in param_file_items:
       hp_train_results = defaultdict(list)
       hp_test_results = defaultdict(list)
       selection_vals = []
@@ -126,7 +132,8 @@ def summarize_evaluation(
     "combined_test": combined_test,
     "args": {
       "ignore_worst": ignore_worst
-    }
+    },
+    "done": all_hps and len(folds) == 10
   }
 
   with open(summary_dir / "results.json", "w") as f:

### 2020-04-10 thesis/content/chapter-appendix.tex
continued with real-world section

@@
 \label{sec:appendix:config-grid}
+
+To tune the hyperparameters of the evaluated models, we used a regular grid search.
+Depending on the type of model, different sets of hyperparameter configurations $\Theta$ were used.
+
+\paragraph{Graph Kernels}
+As described in \cref{sec:eval:setup}, we used the \ac{svm} classifier from \citetitle{SKL} to evaluate the graph kernel approaches.
+We tuned only the regularization parameter \texttt{C} of this classifier;
+the evaluated values are $\mathtt{C} \in \{ 1, \num{1e-1},\allowbreak \num{1e-2},\allowbreak \num{1e-3},\allowbreak \num{1e-4} \}$.
+All other parameters were left at the default setting (using \texttt{scikit-learn 0.22.1}).
+
+\paragraph{Baseline and \ac{gin}}
+For the evaluation of the structure unaware baseline learner and \ac{gin}, we used the same hyperparameter configurations as \citet{Errica2020}.
+We therefore refer to their work for a complete list of the tuned hyperparameters for those models.
+
+\paragraph{2-\acs{gnn} and 2-\acs{wl}-\acs{gnn}}
+We evaluated our implementations of 2-\acsp{gnn} as well as 2-\acs{wl}-\acsp{gnn} on the grid spanned by the following hyperparameter values:
+\begin{itemize}[itemsep=2pt,parsep=2pt]
+	\item \textbf{Number of convolutional layers $T \in \{ 3, 5 \}$:}
+		This parameter describes only the depth of the stack of convolutional layers.
+		The \ac{mlp} after the pooling layer is always configured with a single hidden layer.
+	\item \textbf{Layer width $d \in \{ 32, 64 \}$:}
+		This parameter describes the output dimensionalities $d = d^{(1)} = \cdots = d^{(T)}$ of the convolutional layers and (if applicable) also the hidden layer width of the final \ac{mlp} after the pooling layer.
+	\item \textbf{Jumping knowledge $\mathrm{JK} \in \{ \mathtt{true}, \mathtt{false} \}$:}
+		\citet{Xu2018a} have demonstrated that it can be advantageous for graph classification to pass the outputs of convolutional layers to their successors.
+		We incorporate this idea by passing the input feature vectors not only to the first convolution layer but to all convolution layers through concatenation iff.\ $\mathrm{JK} = \mathtt{true}$.
+	\item \textbf{Learning rate $\eta \in \{ \num{1e-2}, \num{1e-3}, \num{1e-4} \}$} of the Adam optimizer.
+	\item \textbf{Activation functions $\sigma$ and $\sigma_{\Gamma}$} are set to the standard logistic function.
+\end{itemize}
 
@@
 \paragraph{PROTEINS and D\&D}
-The graphs in the PROTEINS dataset represent proteins~\cite{Borgwardt2005a}.
+The graphs in both the PROTEINS dataset~\cite{Borgwardt2005a} as well as the D\&D dataset~\cite{Dobson2003} represent proteins.
 Each vertex corresponds to a so-called \ac{sse}, i.e.\ a certain molecular substructure.
@@
 Each protein graph is classified by whether it is an enzyme or not.
-Very similarly, the D\&D dataset~\cite{Dobson2003} is also about classifying proteins as enzymes or non-enzymes but it uses a different selection of vertex features.
+The main difference between the two datasets is their selection of vertex features/labels.
 
@@
 \label{sec:appendix:fold-diffs}
+
+Due to the relatively small sizes of the evaluated benchmark datasets, the variance of the test accuracies across different folds is quite large.
+When directly comparing the mean accuracies of two learners, it is therefore often impossible to tell whether one consistently outperforms the other.
+We therefore now list the mean and standard deviations of the fold-wise test accuracy differences of all pairs of learners for all datasets.
+This effectively removes the variance introduced by ``easy'' and ``hard'' folds on which all learners might tend to perform consistently better/worse.
+
+In the following \crefrange{tbl:appendix:diff-triangle}{tbl:appendix:diff-imdb} (see \cpagerefrange{tbl:appendix:diff-triangle}{tbl:appendix:diff-imdb}), the differences are computed as $\textit{row accuracy} - \textit{column accuracy}$.
+For each row $i$ and column $j$ we highlight the corresponding cell $(i,j)$ in \textcolor{t_red}{red} or \textcolor{t_darkgreen}{green} iff.\ the learner~$i$ performs consistently \textcolor{t_red}{worse} (or \textcolor{t_darkgreen}{better} respectively) than $j$ with a significance level of $2\sigma$.
+To compute the deltas for 2-\acs{wl}-\acsp{gnn}, we use the same neighborhood radii as in \cref{tbl:eval:synthetic,tbl:eval:real}, i.e.\ $r = 2$ for the synthetic triangle detection dataset and $r = 8$, $5$, $2$, $1$ and $4$ for NCI1, PROTEINS, D\&D, REDDIT and IMDB respectively.
+
+{\captionsetup[table]{name=Matrix}%
+\begin{table}[ht]
+	\caption{Fold-wise accuracy delta means and standard deviations on the triangle dataset.}\label[mat]{tbl:appendix:diff-triangle}
+	\centering
+	\input{data/diffs/triangle.tex}
+\end{table}
+\begin{table}[ht]
+	\caption{Fold-wise accuracy delta means and standard deviations on NCI1.}\label[mat]{tbl:appendix:diff-nci}
+	\centering
+	\input{data/diffs/nci.tex}
+\end{table}
+\begin{table}[ht]
+	\caption{Fold-wise accuracy delta means and standard deviations on PROTEINS.}\label[mat]{tbl:appendix:diff-proteins}
+	\centering
+	\input{data/diffs/proteins.tex}
+\end{table}
+\begin{table}[ht]
+	\caption{Fold-wise accuracy delta means and standard deviations on D\&D.}\label[mat]{tbl:appendix:diff-dd}
+	\centering
+	\input{data/diffs/dd.tex}
+\end{table}
+\begin{table}[ht]
+	\caption{Fold-wise accuracy delta means and standard deviations on REDDIT.}\label[mat]{tbl:appendix:diff-reddit}
+	\centering
+	\input{data/diffs/reddit.tex}
+\end{table}
+\begin{table}[ht]
+	\caption{Fold-wise accuracy delta means and standard deviations on IMDB.}\label[mat]{tbl:appendix:diff-imdb}
+	\centering
+	\input{data/diffs/imdb.tex}
+\end{table}%
+}

### 2020-04-10 thesis/content/chapter-evaluation.tex
continued with real-world section

@@
 This indicates that the theoretical advantages of 2-\acs{wl} over 1-\acs{wl} are not necessarily relevant for the five evaluated problems.
-Nonetheless, the test performance of 2-\acs{wl}-\acsp{gnn} is generally comparable to that of the other state-of-the-art learners with the exception of NCI1.
-If we look at the enzyme detection problem (PROTEINS and D\&D), we observe that all evaluated approaches appear to be unable to leverage structural information for a significant improvement over the baseline learner.
-In the social network datasets (REDDIT and IMDB) on the other hand, the structure aware methods clearly outperform the baseline.
+Nonetheless, the test performance of 2-\acs{wl}-\acsp{gnn} is generally comparable to that of the other state-of-the-art learners with the exception of NCI1,
+where by comparable we mean that the performance of the evaluated 2-\acs{wl}-\acs{gnn} models is within the $2\sigma$ confidence interval of the best evaluated model even when compared fold-by-fold (see \cref{sec:appendix:fold-diffs}).
+
+If we look at the enzyme detection problem (PROTEINS and D\&D), we observe that all evaluated approaches appear to be unable to leverage structural information for a significant improvement over the baseline learner (see \cref{tbl:appendix:diff-proteins,{tbl:appendix:diff-dd}}).
+In the social network datasets (REDDIT and IMDB) on the other hand, the structure aware methods clearly outperform the baseline (see \cref{tbl:appendix:diff-imdb}).
 This confirms the very similar results of \citet{Errica2020}.
+
+\begin{figure}[ht]
+	\centering
+	\begin{tikzpicture}
+		\begin{axis}[
+			width=0.33\linewidth,
+			height=0.3\linewidth,
+			try min ticks=6,
+			xmin=1, xmax=5,
+			ymin=70, ymax=100,
+			label style={font=\tiny},
+			tick label style={font=\tiny},
+			ymajorgrids,
+			ytick style={draw=none},
+			axis line style={gray},
+			mark size={1.5pt}
+		]
+			\addplot [color=t_blue, only marks, mark=*]
+			plot[error bars/.cd, y dir=both, y explicit]
+			table [x=T, y=nciTrainMean, y error=nciTrainStd, col sep=semicolon] {data/wlst_depths.csv};
+			\addplot [color=t_red, only marks, mark=square*]
+			plot[error bars/.cd, y dir=both, y explicit]
+			table [x=T, y=nciTestMean, y error=nciTestStd, col sep=semicolon] {data/wlst_depths.csv};
+		\end{axis}
+	\end{tikzpicture}
+	\caption{test}\label{fig:eval:wlst-depths}
+\end{figure}

### 2020-04-10 thesis/content/chapter-ltd.tex
continued with real-world section

@@
 	W.l.o.g.\ we define the initial colors of the loop vertices as $\chi^{(0)}(v) = \colorlabel{t_blue}{A}$ for all $v \in L_G \cup L_H$. % chktex 25
-	The initial colors of the edge vertices are defined as $\chi^{(0)}(e) = \textcolor{t_red}{\texttt{B}}$ for all $e \in E_G \cup E_H$. % chktex 25
+	The initial colors of the edge vertices are defined as $\chi^{(0)}(e) = \colorlabel{t_red}{B}$ for all $e \in E_G \cup E_H$. % chktex 25
 	Note that each loop vertex $\ldblbrace v_i, v_i \rdblbrace$ with $v_i \in \mathcal{V}_G \cup \mathcal{V}_H$ has $d$ neighbors, the edges incident to $v_i$.
@@
 
-	After one color refinement step we get $\chi^{(1)}(v) = h(\colorlabel{t_blue}{A}, \ldblbrace \underbrace{\textcolor{t_red}{\texttt{B}}, \dots, \textcolor{t_red}{\texttt{B}}}_{d\text{ times}} \rdblbrace) \eqqcolon \textcolor{t_blue}{\texttt{C}}$ for all loop vertices $v \in  L_G \cup L_H$ and $\chi^{(1)}(e) = h(\textcolor{t_red}{\texttt{B}}, \ldblbrace \colorlabel{t_blue}{A}, \colorlabel{t_blue}{A}, \underbrace{\textcolor{t_red}{\texttt{B}}, \dots, \textcolor{t_red}{\texttt{B}}}_{\mathclap{2d - 2 \text{ times}}} \rdblbrace) \eqqcolon \textcolor{t_red}{\texttt{D}}$.
+	After one color refinement step we get $\chi^{(1)}(v) = h(\colorlabel{t_blue}{A}, \ldblbrace \underbrace{\colorlabel{t_red}{B}, \dots, \colorlabel{t_red}{B}}_{d\text{ times}} \rdblbrace) \eqqcolon \colorlabel{t_blue}{C}$ for all loop vertices $v \in  L_G \cup L_H$ and $\chi^{(1)}(e) = h(\colorlabel{t_red}{B}, \ldblbrace \colorlabel{t_blue}{A}, \colorlabel{t_blue}{A}, \underbrace{\colorlabel{t_red}{B}, \dots, \colorlabel{t_red}{B}}_{\mathclap{2d - 2 \text{ times}}} \rdblbrace) \eqqcolon \colorlabel{t_red}{D}$.
 % chktex 25
-	This means that $\chi^{(0)}$ and $\chi^{(1)}$ are identical up to the color substitutions $\colorlabel{t_blue}{A} \to \textcolor{t_blue}{\texttt{C}}$ and $\textcolor{t_red}{\texttt{B}} \to \textcolor{t_red}{\texttt{D}}$, i.e.\ $\chi^{(0)} \equiv \chi^{(1)}$, which in turn implies that 1-\ac{wl} terminates after one iteration.
+	This means that $\chi^{(0)}$ and $\chi^{(1)}$ are identical up to the color substitutions $\colorlabel{t_blue}{A} \to \colorlabel{t_blue}{C}$ and $\colorlabel{t_red}{B} \to \colorlabel{t_red}{D}$, i.e.\ $\chi^{(0)} \equiv \chi^{(1)}$, which in turn implies that 1-\ac{wl} terminates after one iteration.
 % chktex 25
-	\Cref{lem:ltd:wl1-regular-edge-neighbor-limit} then directly follows since both $G^{\mathcal{E}}$ and $H^{\mathcal{E}}$ have $n$ vertices with the final color $\textcolor{t_blue}{\texttt{C}}$ and $\frac{nd}{2}$ vertices with the final color $\textcolor{t_red}{\texttt{D}}$, i.e.\ $G^{\mathcal{E}} \mathrel{\simeq_1} H^{\mathcal{E}}$.
+	\Cref{lem:ltd:wl1-regular-edge-neighbor-limit} then directly follows since both $G^{\mathcal{E}}$ and $H^{\mathcal{E}}$ have $n$ vertices with the final color $\colorlabel{t_blue}{C}$ and $\frac{nd}{2}$ vertices with the final color $\colorlabel{t_red}{D}$, i.e.\ $G^{\mathcal{E}} \mathrel{\simeq_1} H^{\mathcal{E}}$.
 % chktex 25
@@
 
-	Let $\varphi: \mathbb{R}^{d^{(T, 2 + K)}} \to \mathbb{R}^{d^{(T)}} \cup \{ \texttt{nil} \}$ be a function which maps the final 2-\acs{wl} feature vectors produced by $\mathit{Conv}_2$ to the output space of $\mathit{Conv}_1$ or the constant \texttt{nil}.
-	Let $\mathit{Pool}_2\left(Z_G^{(T,2+K)}\right) \coloneqq \mathit{Pool}_1(\ldblbrace z_{ij}\, |\, z_{ij} = \varphi\left(Z_G^{(T,2+K)}[e_{ij}]\right) \land e_{ij} \in \mathcal{E}_{G^1} \land z_{ij} \neq \texttt{nil} \rdblbrace)$.
-	\Cref{thm:ltd:wl2-simulation} then follows if there is a function $\varphi$ s.t.\ $\forall v_i \in \mathcal{V}_G: \mathit{Conv}_1(G)[v_i] = \varphi(\mathit{Conv}_2(G)[e_{ii}])$ and $\forall e_{ij} \in \mathcal{E}_G: \varphi(\mathit{Conv}_2(G)[e_{ij}]) = \texttt{nil}$.
+	Let $\varphi: \mathbb{R}^{d^{(T, 2 + K)}} \to \mathbb{R}^{d^{(T)}} \cup \{ \mathtt{nil} \}$ be a function which maps the final 2-\acs{wl} feature vectors produced by $\mathit{Conv}_2$ to the output space of $\mathit{Conv}_1$ or the constant \texttt{nil}.
+	Let $\mathit{Pool}_2\left(Z_G^{(T,2+K)}\right) \coloneqq \mathit{Pool}_1(\ldblbrace z_{ij}\, |\, z_{ij} = \varphi\left(Z_G^{(T,2+K)}[e_{ij}]\right) \land e_{ij} \in \mathcal{E}_{G^1} \land z_{ij} \neq \mathtt{nil} \rdblbrace)$.
+	\Cref{thm:ltd:wl2-simulation} then follows if there is a function $\varphi$ s.t.\ $\forall v_i \in \mathcal{V}_G: \mathit{Conv}_1(G)[v_i] = \varphi(\mathit{Conv}_2(G)[e_{ii}])$ and $\forall e_{ij} \in \mathcal{E}_G: \varphi(\mathit{Conv}_2(G)[e_{ij}]) = \mathtt{nil}$.
 	To guarantee that there is such a function $\varphi$ we now inductively prove the following three invariants which have to hold for all $t \in \{ 0, \dots, T \}$:
@@
 			{Z_G^{(T,2+K)}[e_{ij}]}_{2, \dots, (d^{T} + 1)} & \text{if ${Z_G^{(T,2+K)}[e_{ij}]}_1 = 1$} \\
-			\texttt{nil} & \text{else}
+			\mathtt{nil} & \text{else}
 		\end{cases}

### 2020-04-10 thesis/data/diffs/imdb.tex
continued with real-world section

% This file was generated by the LTAG results postprocessor. Do not edit manually.
{\setlength\tabcolsep{2.5pt}\setlength{\extrarowheight}{2pt}%
\begin{tabular}{lccccccccccccc}
& \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=5$)} & \rotatebox[origin=l]{90}{\textbf{WL\textsubscript{SP}} ($T=5$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-GWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{Baseline} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{GIN} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)} \\
\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)& & {\tiny\Vectorstack{-1.7\\ \pm 3.2}} & {\tiny\Vectorstack{-2.6\\ \pm 3.5}} & {\tiny\Vectorstack{-1.2\\ \pm 2.6}} & {\tiny\Vectorstack{+0.6\\ \pm 3.5}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+20\\ \pm 4.6}}} & {\tiny\Vectorstack{+4.2\\ \pm 6.1}} & {\tiny\Vectorstack{-0.4\\ \pm 3.1}} & {\tiny\Vectorstack{+0.1\\ \pm 2.4}} & {\tiny\Vectorstack{+0.1\\ \pm 3.5}} & {\tiny\Vectorstack{-0.2\\ \pm 4.4}} & {\tiny\Vectorstack{-1.2\\ \pm 3.1}} & {\tiny\Vectorstack{-0.1\\ \pm 3.8}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=5$)&{\tiny\Vectorstack{+1.7\\ \pm 3.2}} &  & {\tiny\Vectorstack{-0.9\\ \pm 3.7}} & {\tiny\Vectorstack{+0.5\\ \pm 3.3}} & {\tiny\Vectorstack{+2.3\\ \pm 3.2}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+22\\ \pm 4.6}}} & {\tiny\Vectorstack{+5.9\\ \pm 6.1}} & {\tiny\Vectorstack{+1.3\\ \pm 3.8}} & {\tiny\Vectorstack{+1.8\\ \pm 4.3}} & {\tiny\Vectorstack{+1.8\\ \pm 4.1}} & {\tiny\Vectorstack{+1.5\\ \pm 4.6}} & {\tiny\Vectorstack{+0.5\\ \pm 4.1}} & {\tiny\Vectorstack{+1.6\\ \pm 5.1}} \\[2pt]
\textbf{WL\textsubscript{SP}} ($T=5$)&{\tiny\Vectorstack{+2.6\\ \pm 3.5}} & {\tiny\Vectorstack{+0.9\\ \pm 3.7}} &  & {\tiny\Vectorstack{+1.4\\ \pm 4.9}} & {\tiny\Vectorstack{+3.2\\ \pm 5.0}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+23\\ \pm 5.0}}} & {\tiny\Vectorstack{+6.8\\ \pm 6.5}} & {\tiny\Vectorstack{+2.2\\ \pm 4.8}} & {\tiny\Vectorstack{+2.7\\ \pm 4.4}} & {\tiny\Vectorstack{+2.7\\ \pm 5.0}} & {\tiny\Vectorstack{+2.4\\ \pm 5.7}} & {\tiny\Vectorstack{+1.4\\ \pm 5.1}} & {\tiny\Vectorstack{+2.5\\ \pm 5.6}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)&{\tiny\Vectorstack{+1.2\\ \pm 2.6}} & {\tiny\Vectorstack{-0.5\\ \pm 3.3}} & {\tiny\Vectorstack{-1.4\\ \pm 4.9}} &  & {\tiny\Vectorstack{+1.8\\ \pm 2.6}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+22\\ \pm 5.7}}} & {\tiny\Vectorstack{+5.4\\ \pm 7.2}} & {\tiny\Vectorstack{+0.8\\ \pm 2.0}} & {\tiny\Vectorstack{+1.3\\ \pm 3.2}} & {\tiny\Vectorstack{+1.3\\ \pm 3.2}} & {\tiny\Vectorstack{+1.0\\ \pm 3.5}} & {\tiny\Vectorstack{+0.0\\ \pm 3.0}} & {\tiny\Vectorstack{+1.1\\ \pm 3.7}} \\[2pt]
\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-0.6\\ \pm 3.5}} & {\tiny\Vectorstack{-2.3\\ \pm 3.2}} & {\tiny\Vectorstack{-3.2\\ \pm 5.0}} & {\tiny\Vectorstack{-1.8\\ \pm 2.6}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+20\\ \pm 5.6}}} & {\tiny\Vectorstack{+3.6\\ \pm 7.1}} & {\tiny\Vectorstack{-1.0\\ \pm 2.6}} & {\tiny\Vectorstack{-0.5\\ \pm 4.1}} & {\tiny\Vectorstack{-0.5\\ \pm 3.3}} & {\tiny\Vectorstack{-0.8\\ \pm 3.1}} & {\tiny\Vectorstack{-1.8\\ \pm 3.0}} & {\tiny\Vectorstack{-0.7\\ \pm 4.3}} \\[2pt]
\textbf{Baseline} ($\mathrm{sum}$)&\cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-20\\ \pm 4.6}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-22\\ \pm 4.6}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-23\\ \pm 5.0}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 5.7}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-20\\ \pm 5.6}}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-16\\ \pm 6.3}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 6.0}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-20\\ \pm 5.6}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-20\\ \pm 5.8}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 6.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 5.5}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-20\\ \pm 6.9}}} \\[2pt]
\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{-4.2\\ \pm 6.1}} & {\tiny\Vectorstack{-5.9\\ \pm 6.1}} & {\tiny\Vectorstack{-6.8\\ \pm 6.5}} & {\tiny\Vectorstack{-5.4\\ \pm 7.2}} & {\tiny\Vectorstack{-3.6\\ \pm 7.1}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+16\\ \pm 6.3}}} &  & {\tiny\Vectorstack{-4.6\\ \pm 7.5}} & {\tiny\Vectorstack{-4.1\\ \pm 7.1}} & {\tiny\Vectorstack{-4.1\\ \pm 7.3}} & {\tiny\Vectorstack{-4.4\\ \pm 7.9}} & {\tiny\Vectorstack{-5.4\\ \pm 7.0}} & {\tiny\Vectorstack{-4.3\\ \pm 8.4}} \\[2pt]
\textbf{2-GNN} ($\mean$)&{\tiny\Vectorstack{+0.4\\ \pm 3.1}} & {\tiny\Vectorstack{-1.3\\ \pm 3.8}} & {\tiny\Vectorstack{-2.2\\ \pm 4.8}} & {\tiny\Vectorstack{-0.8\\ \pm 2.0}} & {\tiny\Vectorstack{+1.0\\ \pm 2.6}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+21\\ \pm 6.0}}} & {\tiny\Vectorstack{+4.6\\ \pm 7.5}} &  & {\tiny\Vectorstack{+0.5\\ \pm 2.8}} & {\tiny\Vectorstack{+0.5\\ \pm 2.5}} & {\tiny\Vectorstack{+0.2\\ \pm 2.8}} & {\tiny\Vectorstack{-0.8\\ \pm 2.6}} & {\tiny\Vectorstack{+0.2\\ \pm 2.4}} \\[2pt]
\textbf{2-GNN} ($\mathrm{SAM}$)&{\tiny\Vectorstack{-0.1\\ \pm 2.4}} & {\tiny\Vectorstack{-1.8\\ \pm 4.3}} & {\tiny\Vectorstack{-2.7\\ \pm 4.4}} & {\tiny\Vectorstack{-1.3\\ \pm 3.2}} & {\tiny\Vectorstack{+0.5\\ \pm 4.1}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+20\\ \pm 5.6}}} & {\tiny\Vectorstack{+4.1\\ \pm 7.1}} & {\tiny\Vectorstack{-0.5\\ \pm 2.8}} &  & {\tiny\Vectorstack{+0.0\\ \pm 2.8}} & {\tiny\Vectorstack{-0.3\\ \pm 3.5}} & {\tiny\Vectorstack{-1.3\\ \pm 2.5}} & {\tiny\Vectorstack{-0.2\\ \pm 2.7}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)&{\tiny\Vectorstack{-0.1\\ \pm 3.5}} & {\tiny\Vectorstack{-1.8\\ \pm 4.1}} & {\tiny\Vectorstack{-2.7\\ \pm 5.0}} & {\tiny\Vectorstack{-1.3\\ \pm 3.2}} & {\tiny\Vectorstack{+0.5\\ \pm 3.3}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+20\\ \pm 5.8}}} & {\tiny\Vectorstack{+4.1\\ \pm 7.3}} & {\tiny\Vectorstack{-0.5\\ \pm 2.5}} & {\tiny\Vectorstack{-0.0\\ \pm 2.8}} &  & {\tiny\Vectorstack{-0.3\\ \pm 1.7}} & {\tiny\Vectorstack{-1.3\\ \pm 1.1}} & {\tiny\Vectorstack{-0.3\\ \pm 2.2}} \\[2pt]
\textbf{2-WL-GNN\phantom{*}} ($\mean$)&{\tiny\Vectorstack{+0.2\\ \pm 4.4}} & {\tiny\Vectorstack{-1.5\\ \pm 4.6}} & {\tiny\Vectorstack{-2.4\\ \pm 5.7}} & {\tiny\Vectorstack{-1.0\\ \pm 3.5}} & {\tiny\Vectorstack{+0.8\\ \pm 3.1}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+21\\ \pm 6.4}}} & {\tiny\Vectorstack{+4.4\\ \pm 7.9}} & {\tiny\Vectorstack{-0.2\\ \pm 2.8}} & {\tiny\Vectorstack{+0.3\\ \pm 3.5}} & {\tiny\Vectorstack{+0.3\\ \pm 1.7}} &  & {\tiny\Vectorstack{-1.0\\ \pm 2.0}} & {\tiny\Vectorstack{+0.1\\ \pm 3.4}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{+1.2\\ \pm 3.1}} & {\tiny\Vectorstack{-0.5\\ \pm 4.1}} & {\tiny\Vectorstack{-1.4\\ \pm 5.1}} & {\tiny\Vectorstack{-0.0\\ \pm 3.0}} & {\tiny\Vectorstack{+1.8\\ \pm 3.0}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+21\\ \pm 5.5}}} & {\tiny\Vectorstack{+5.4\\ \pm 7.0}} & {\tiny\Vectorstack{+0.8\\ \pm 2.6}} & {\tiny\Vectorstack{+1.3\\ \pm 2.5}} & {\tiny\Vectorstack{+1.3\\ \pm 1.1}} & {\tiny\Vectorstack{+1.0\\ \pm 2.0}} &  & {\tiny\Vectorstack{+1.0\\ \pm 2.6}} \\[2pt]
\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{+0.1\\ \pm 3.8}} & {\tiny\Vectorstack{-1.6\\ \pm 5.1}} & {\tiny\Vectorstack{-2.5\\ \pm 5.6}} & {\tiny\Vectorstack{-1.1\\ \pm 3.7}} & {\tiny\Vectorstack{+0.7\\ \pm 4.3}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+20\\ \pm 6.9}}} & {\tiny\Vectorstack{+4.3\\ \pm 8.4}} & {\tiny\Vectorstack{-0.2\\ \pm 2.4}} & {\tiny\Vectorstack{+0.2\\ \pm 2.7}} & {\tiny\Vectorstack{+0.3\\ \pm 2.2}} & {\tiny\Vectorstack{-0.1\\ \pm 3.4}} & {\tiny\Vectorstack{-1.0\\ \pm 2.6}} & 
\end{tabular}}

### 2020-04-10 thesis/data/diffs/nci.tex
continued with real-world section

% This file was generated by the LTAG results postprocessor. Do not edit manually.
{\setlength\tabcolsep{2.5pt}\setlength{\extrarowheight}{2pt}%
\begin{tabular}{lccccccccccc}
& \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=5$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-GWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{Baseline} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{GIN} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)} \\
\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)& & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-12\\ \pm 1.3}}} & {\tiny\Vectorstack{-2.8\\ \pm 1.5}} & {\tiny\Vectorstack{+2.3\\ \pm 2.4}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.2\\ \pm 2.9}}} & {\tiny\Vectorstack{-3.5\\ \pm 2.2}} & {\tiny\Vectorstack{-1.9\\ \pm 2.9}} & {\tiny\Vectorstack{-4.4\\ \pm 2.3}} & {\tiny\Vectorstack{+3.1\\ \pm 2.5}} & {\tiny\Vectorstack{+1.6\\ \pm 2.6}} & {\tiny\Vectorstack{+3.7\\ \pm 3.6}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=5$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+12\\ \pm 1.3}}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.7\\ \pm 1.7}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+14\\ \pm 1.9}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+18\\ \pm 2.7}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.0\\ \pm 2.7}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+9.6\\ \pm 2.2}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+7.1\\ \pm 2.0}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+15\\ \pm 2.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+13\\ \pm 2.0}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+15\\ \pm 3.8}}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)&{\tiny\Vectorstack{+2.8\\ \pm 1.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.7\\ \pm 1.7}}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+5.1\\ \pm 2.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+9.0\\ \pm 3.0}}} & {\tiny\Vectorstack{-0.7\\ \pm 2.7}} & {\tiny\Vectorstack{+0.9\\ \pm 2.0}} & {\tiny\Vectorstack{-1.6\\ \pm 2.3}} & {\tiny\Vectorstack{+5.9\\ \pm 3.2}} & {\tiny\Vectorstack{+4.4\\ \pm 2.5}} & {\tiny\Vectorstack{+6.5\\ \pm 4.2}} \\[2pt]
\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-2.3\\ \pm 2.4}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-14\\ \pm 1.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-5.1\\ \pm 2.3}}} &  & {\tiny\Vectorstack{+3.9\\ \pm 4.1}} & {\tiny\Vectorstack{-5.8\\ \pm 3.6}} & {\tiny\Vectorstack{-4.3\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.7\\ \pm 2.5}}} & {\tiny\Vectorstack{+0.8\\ \pm 3.7}} & {\tiny\Vectorstack{-0.8\\ \pm 2.7}} & {\tiny\Vectorstack{+1.4\\ \pm 4.0}} \\[2pt]
\textbf{Baseline} ($\mathrm{sum}$)&\cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.2\\ \pm 2.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-18\\ \pm 2.7}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-9.0\\ \pm 3.0}}} & {\tiny\Vectorstack{-3.9\\ \pm 4.1}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-9.8\\ \pm 2.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.2\\ \pm 2.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 2.6}}} & {\tiny\Vectorstack{-3.1\\ \pm 3.0}} & {\tiny\Vectorstack{-4.7\\ \pm 3.1}} & {\tiny\Vectorstack{-2.6\\ \pm 5.0}} \\[2pt]
\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{+3.5\\ \pm 2.2}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.0\\ \pm 2.7}}} & {\tiny\Vectorstack{+0.7\\ \pm 2.7}} & {\tiny\Vectorstack{+5.8\\ \pm 3.6}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+9.8\\ \pm 2.4}}} &  & {\tiny\Vectorstack{+1.6\\ \pm 3.5}} & {\tiny\Vectorstack{-0.9\\ \pm 2.2}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.6\\ \pm 2.5}}} & {\tiny\Vectorstack{+5.1\\ \pm 3.2}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+7.2\\ \pm 3.3}}} \\[2pt]
\textbf{2-GNN} ($\mean$)&{\tiny\Vectorstack{+1.9\\ \pm 2.9}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-9.6\\ \pm 2.2}}} & {\tiny\Vectorstack{-0.9\\ \pm 2.0}} & {\tiny\Vectorstack{+4.3\\ \pm 2.5}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.2\\ \pm 2.9}}} & {\tiny\Vectorstack{-1.6\\ \pm 3.5}} &  & {\tiny\Vectorstack{-2.4\\ \pm 2.6}} & {\tiny\Vectorstack{+5.1\\ \pm 3.5}} & {\tiny\Vectorstack{+3.5\\ \pm 2.2}} & {\tiny\Vectorstack{+5.6\\ \pm 4.8}} \\[2pt]
\textbf{2-GNN} ($\mathrm{SAM}$)&{\tiny\Vectorstack{+4.4\\ \pm 2.3}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-7.1\\ \pm 2.0}}} & {\tiny\Vectorstack{+1.6\\ \pm 2.3}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.7\\ \pm 2.5}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 2.6}}} & {\tiny\Vectorstack{+0.9\\ \pm 2.2}} & {\tiny\Vectorstack{+2.4\\ \pm 2.6}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+7.5\\ \pm 3.1}}} & {\tiny\Vectorstack{+5.9\\ \pm 3.2}} & {\tiny\Vectorstack{+8.1\\ \pm 4.2}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)&{\tiny\Vectorstack{-3.1\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-15\\ \pm 2.3}}} & {\tiny\Vectorstack{-5.9\\ \pm 3.2}} & {\tiny\Vectorstack{-0.8\\ \pm 3.7}} & {\tiny\Vectorstack{+3.1\\ \pm 3.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.6\\ \pm 2.5}}} & {\tiny\Vectorstack{-5.1\\ \pm 3.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-7.5\\ \pm 3.1}}} &  & {\tiny\Vectorstack{-1.5\\ \pm 2.2}} & {\tiny\Vectorstack{+0.6\\ \pm 2.9}} \\[2pt]
\textbf{2-WL-GNN\phantom{*}} ($\mean$)&{\tiny\Vectorstack{-1.6\\ \pm 2.6}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-13\\ \pm 2.0}}} & {\tiny\Vectorstack{-4.4\\ \pm 2.5}} & {\tiny\Vectorstack{+0.8\\ \pm 2.7}} & {\tiny\Vectorstack{+4.7\\ \pm 3.1}} & {\tiny\Vectorstack{-5.1\\ \pm 3.2}} & {\tiny\Vectorstack{-3.5\\ \pm 2.2}} & {\tiny\Vectorstack{-5.9\\ \pm 3.2}} & {\tiny\Vectorstack{+1.5\\ \pm 2.2}} &  & {\tiny\Vectorstack{+2.1\\ \pm 3.4}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{-3.7\\ \pm 3.6}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-15\\ \pm 3.8}}} & {\tiny\Vectorstack{-6.5\\ \pm 4.2}} & {\tiny\Vectorstack{-1.4\\ \pm 4.0}} & {\tiny\Vectorstack{+2.6\\ \pm 5.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-7.2\\ \pm 3.3}}} & {\tiny\Vectorstack{-5.6\\ \pm 4.8}} & {\tiny\Vectorstack{-8.1\\ \pm 4.2}} & {\tiny\Vectorstack{-0.6\\ \pm 2.9}} & {\tiny\Vectorstack{-2.1\\ \pm 3.4}} & 
\end{tabular}}

### 2020-04-10 thesis/data/diffs/reddit.tex
continued with real-world section

% This file was generated by the LTAG results postprocessor. Do not edit manually.
{\setlength\tabcolsep{2.5pt}\setlength{\extrarowheight}{2pt}%
\begin{tabular}{lcccccccccc}
& \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=5$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-GWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{Baseline} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{GIN} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)} \\
\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)& & {\tiny\Vectorstack{-1.1\\ \pm 5.3}} & {\tiny\Vectorstack{+0.5\\ \pm 4.9}} & {\tiny\Vectorstack{+0.8\\ \pm 3.6}} & {\tiny\Vectorstack{+4.2\\ \pm 10}} & {\tiny\Vectorstack{-11\\ \pm 6.9}} & {\tiny\Vectorstack{+2.9\\ \pm 3.5}} & {\tiny\Vectorstack{-7.4\\ \pm 6.2}} & {\tiny\Vectorstack{-0.9\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-12\\ \pm 3.4}}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=5$)&{\tiny\Vectorstack{+1.1\\ \pm 5.3}} &  & {\tiny\Vectorstack{+1.5\\ \pm 3.5}} & {\tiny\Vectorstack{+1.9\\ \pm 5.0}} & {\tiny\Vectorstack{+5.2\\ \pm 11}} & {\tiny\Vectorstack{-9.7\\ \pm 8.1}} & {\tiny\Vectorstack{+4.0\\ \pm 6.1}} & {\tiny\Vectorstack{-6.4\\ \pm 8.1}} & {\tiny\Vectorstack{+0.2\\ \pm 5.1}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 3.7}}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)&{\tiny\Vectorstack{-0.5\\ \pm 4.9}} & {\tiny\Vectorstack{-1.5\\ \pm 3.5}} &  & {\tiny\Vectorstack{+0.3\\ \pm 4.2}} & {\tiny\Vectorstack{+3.7\\ \pm 11}} & {\tiny\Vectorstack{-11\\ \pm 7.3}} & {\tiny\Vectorstack{+2.4\\ \pm 5.5}} & {\tiny\Vectorstack{-7.9\\ \pm 5.5}} & {\tiny\Vectorstack{-1.4\\ \pm 5.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-13\\ \pm 3.5}}} \\[2pt]
\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-0.8\\ \pm 3.6}} & {\tiny\Vectorstack{-1.9\\ \pm 5.0}} & {\tiny\Vectorstack{-0.3\\ \pm 4.2}} &  & {\tiny\Vectorstack{+3.3\\ \pm 11}} & {\tiny\Vectorstack{-12\\ \pm 7.6}} & {\tiny\Vectorstack{+2.1\\ \pm 2.4}} & {\tiny\Vectorstack{-8.2\\ \pm 6.2}} & {\tiny\Vectorstack{-1.7\\ \pm 3.3}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-13\\ \pm 3.6}}} \\[2pt]
\textbf{Baseline} ($\mathrm{sum}$)&{\tiny\Vectorstack{-4.2\\ \pm 10}} & {\tiny\Vectorstack{-5.2\\ \pm 11}} & {\tiny\Vectorstack{-3.7\\ \pm 11}} & {\tiny\Vectorstack{-3.3\\ \pm 11}} &  & {\tiny\Vectorstack{-15\\ \pm 12}} & {\tiny\Vectorstack{-1.2\\ \pm 12}} & {\tiny\Vectorstack{-12\\ \pm 13}} & {\tiny\Vectorstack{-5.0\\ \pm 11}} & {\tiny\Vectorstack{-16\\ \pm 10}} \\[2pt]
\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{+11\\ \pm 6.9}} & {\tiny\Vectorstack{+9.7\\ \pm 8.1}} & {\tiny\Vectorstack{+11\\ \pm 7.3}} & {\tiny\Vectorstack{+12\\ \pm 7.6}} & {\tiny\Vectorstack{+15\\ \pm 12}} &  & {\tiny\Vectorstack{+14\\ \pm 8.3}} & {\tiny\Vectorstack{+3.3\\ \pm 9.6}} & {\tiny\Vectorstack{+9.9\\ \pm 7.4}} & {\tiny\Vectorstack{-1.3\\ \pm 7.0}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)&{\tiny\Vectorstack{-2.9\\ \pm 3.5}} & {\tiny\Vectorstack{-4.0\\ \pm 6.1}} & {\tiny\Vectorstack{-2.4\\ \pm 5.5}} & {\tiny\Vectorstack{-2.1\\ \pm 2.4}} & {\tiny\Vectorstack{+1.2\\ \pm 12}} & {\tiny\Vectorstack{-14\\ \pm 8.3}} &  & {\tiny\Vectorstack{-10\\ \pm 7.1}} & {\tiny\Vectorstack{-3.8\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-15\\ \pm 3.9}}} \\[2pt]
\textbf{2-WL-GNN\phantom{*}} ($\mean$)&{\tiny\Vectorstack{+7.4\\ \pm 6.2}} & {\tiny\Vectorstack{+6.4\\ \pm 8.1}} & {\tiny\Vectorstack{+7.9\\ \pm 5.5}} & {\tiny\Vectorstack{+8.2\\ \pm 6.2}} & {\tiny\Vectorstack{+12\\ \pm 13}} & {\tiny\Vectorstack{-3.3\\ \pm 9.6}} & {\tiny\Vectorstack{+10\\ \pm 7.1}} &  & {\tiny\Vectorstack{+6.6\\ \pm 7.2}} & {\tiny\Vectorstack{-4.6\\ \pm 7.1}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{+0.9\\ \pm 2.5}} & {\tiny\Vectorstack{-0.2\\ \pm 5.1}} & {\tiny\Vectorstack{+1.4\\ \pm 5.0}} & {\tiny\Vectorstack{+1.7\\ \pm 3.3}} & {\tiny\Vectorstack{+5.0\\ \pm 11}} & {\tiny\Vectorstack{-9.9\\ \pm 7.4}} & {\tiny\Vectorstack{+3.8\\ \pm 2.5}} & {\tiny\Vectorstack{-6.6\\ \pm 7.2}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 2.6}}} \\[2pt]
\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+12\\ \pm 3.4}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 3.7}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+13\\ \pm 3.5}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+13\\ \pm 3.6}}} & {\tiny\Vectorstack{+16\\ \pm 10}} & {\tiny\Vectorstack{+1.3\\ \pm 7.0}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+15\\ \pm 3.9}}} & {\tiny\Vectorstack{+4.6\\ \pm 7.1}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 2.6}}} & 
\end{tabular}}

### 2020-04-10 thesis/data/diffs/triangle.tex
continued with real-world section

% This file was generated by the LTAG results postprocessor. Do not edit manually.
{\setlength\tabcolsep{2.5pt}\setlength{\extrarowheight}{2pt}%
\begin{tabular}{lccccccccccccc}
& \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=5$)} & \rotatebox[origin=l]{90}{\textbf{WL\textsubscript{SP}} ($T=5$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-GWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{Baseline} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{GIN} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)} \\
\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)& & {\tiny\Vectorstack{+2.2\\ \pm 9.2}} & {\tiny\Vectorstack{-4.9\\ \pm 12}} & {\tiny\Vectorstack{+8.2\\ \pm 15}} & {\tiny\Vectorstack{+2.9\\ \pm 12}} & {\tiny\Vectorstack{+20\\ \pm 19}} & {\tiny\Vectorstack{-5.2\\ \pm 16}} & {\tiny\Vectorstack{-12\\ \pm 18}} & {\tiny\Vectorstack{-17\\ \pm 19}} & {\tiny\Vectorstack{-26\\ \pm 15}} & {\tiny\Vectorstack{-28\\ \pm 17}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-26\\ \pm 13}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-35\\ \pm 13}}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=5$)&{\tiny\Vectorstack{-2.2\\ \pm 9.2}} &  & {\tiny\Vectorstack{-7.1\\ \pm 14}} & {\tiny\Vectorstack{+6.1\\ \pm 9.8}} & {\tiny\Vectorstack{+0.8\\ \pm 13}} & {\tiny\Vectorstack{+18\\ \pm 14}} & {\tiny\Vectorstack{-7.4\\ \pm 11}} & {\tiny\Vectorstack{-14\\ \pm 16}} & {\tiny\Vectorstack{-19\\ \pm 16}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-28\\ \pm 14}}} & {\tiny\Vectorstack{-30\\ \pm 17}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-28\\ \pm 9.6}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-37\\ \pm 11}}} \\[2pt]
\textbf{WL\textsubscript{SP}} ($T=5$)&{\tiny\Vectorstack{+4.9\\ \pm 12}} & {\tiny\Vectorstack{+7.1\\ \pm 14}} &  & {\tiny\Vectorstack{+13\\ \pm 13}} & {\tiny\Vectorstack{+7.9\\ \pm 10}} & {\tiny\Vectorstack{+25\\ \pm 16}} & {\tiny\Vectorstack{-0.3\\ \pm 13}} & {\tiny\Vectorstack{-7.1\\ \pm 10}} & {\tiny\Vectorstack{-12\\ \pm 12}} & {\tiny\Vectorstack{-21\\ \pm 11}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-23\\ \pm 11}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 10}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-30\\ \pm 8.9}}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)&{\tiny\Vectorstack{-8.2\\ \pm 15}} & {\tiny\Vectorstack{-6.1\\ \pm 9.8}} & {\tiny\Vectorstack{-13\\ \pm 13}} &  & {\tiny\Vectorstack{-5.3\\ \pm 11}} & {\tiny\Vectorstack{+12\\ \pm 7.5}} & {\tiny\Vectorstack{-13\\ \pm 6.7}} & {\tiny\Vectorstack{-20\\ \pm 13}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-25\\ \pm 9.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-34\\ \pm 8.6}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-36\\ \pm 10}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-34\\ \pm 7.0}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-43\\ \pm 6.8}}} \\[2pt]
\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-2.9\\ \pm 12}} & {\tiny\Vectorstack{-0.8\\ \pm 13}} & {\tiny\Vectorstack{-7.9\\ \pm 10}} & {\tiny\Vectorstack{+5.3\\ \pm 11}} &  & {\tiny\Vectorstack{+17\\ \pm 15}} & {\tiny\Vectorstack{-8.2\\ \pm 13}} & {\tiny\Vectorstack{-15\\ \pm 14}} & {\tiny\Vectorstack{-20\\ \pm 13}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-29\\ \pm 8.8}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-31\\ \pm 9.0}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-29\\ \pm 10}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-38\\ \pm 9.0}}} \\[2pt]
\textbf{Baseline} ($\mathrm{sum}$)&{\tiny\Vectorstack{-20\\ \pm 19}} & {\tiny\Vectorstack{-18\\ \pm 14}} & {\tiny\Vectorstack{-25\\ \pm 16}} & {\tiny\Vectorstack{-12\\ \pm 7.5}} & {\tiny\Vectorstack{-17\\ \pm 15}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-25\\ \pm 8.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-32\\ \pm 14}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-37\\ \pm 9.2}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-46\\ \pm 9.8}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-48\\ \pm 14}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-46\\ \pm 9.2}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-55\\ \pm 8.5}}} \\[2pt]
\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{+5.2\\ \pm 16}} & {\tiny\Vectorstack{+7.4\\ \pm 11}} & {\tiny\Vectorstack{+0.3\\ \pm 13}} & {\tiny\Vectorstack{+13\\ \pm 6.7}} & {\tiny\Vectorstack{+8.2\\ \pm 13}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+25\\ \pm 8.4}}} &  & {\tiny\Vectorstack{-6.8\\ \pm 13}} & {\tiny\Vectorstack{-12\\ \pm 9.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 9.9}}} & {\tiny\Vectorstack{-23\\ \pm 14}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 8.3}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-29\\ \pm 7.3}}} \\[2pt]
\textbf{2-GNN} ($\mean$)&{\tiny\Vectorstack{+12\\ \pm 18}} & {\tiny\Vectorstack{+14\\ \pm 16}} & {\tiny\Vectorstack{+7.1\\ \pm 10}} & {\tiny\Vectorstack{+20\\ \pm 13}} & {\tiny\Vectorstack{+15\\ \pm 14}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+32\\ \pm 14}}} & {\tiny\Vectorstack{+6.8\\ \pm 13}} &  & {\tiny\Vectorstack{-5.0\\ \pm 6.6}} & {\tiny\Vectorstack{-14\\ \pm 9.6}} & {\tiny\Vectorstack{-16\\ \pm 11}} & {\tiny\Vectorstack{-14\\ \pm 8.8}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-23\\ \pm 9.8}}} \\[2pt]
\textbf{2-GNN} ($\mathrm{SAM}$)&{\tiny\Vectorstack{+17\\ \pm 19}} & {\tiny\Vectorstack{+19\\ \pm 16}} & {\tiny\Vectorstack{+12\\ \pm 12}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+25\\ \pm 9.4}}} & {\tiny\Vectorstack{+20\\ \pm 13}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+37\\ \pm 9.2}}} & {\tiny\Vectorstack{+12\\ \pm 9.5}} & {\tiny\Vectorstack{+5.0\\ \pm 6.6}} &  & {\tiny\Vectorstack{-8.9\\ \pm 6.3}} & {\tiny\Vectorstack{-11\\ \pm 8.8}} & {\tiny\Vectorstack{-9.2\\ \pm 7.1}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-18\\ \pm 7.2}}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)&{\tiny\Vectorstack{+26\\ \pm 15}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+28\\ \pm 14}}} & {\tiny\Vectorstack{+21\\ \pm 11}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+34\\ \pm 8.6}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+29\\ \pm 8.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+46\\ \pm 9.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+21\\ \pm 9.9}}} & {\tiny\Vectorstack{+14\\ \pm 9.6}} & {\tiny\Vectorstack{+8.9\\ \pm 6.3}} &  & {\tiny\Vectorstack{-2.2\\ \pm 7.2}} & {\tiny\Vectorstack{-0.3\\ \pm 6.0}} & {\tiny\Vectorstack{-8.7\\ \pm 4.6}} \\[2pt]
\textbf{2-WL-GNN\phantom{*}} ($\mean$)&{\tiny\Vectorstack{+28\\ \pm 17}} & {\tiny\Vectorstack{+30\\ \pm 17}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+23\\ \pm 11}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+36\\ \pm 10}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+31\\ \pm 9.0}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+48\\ \pm 14}}} & {\tiny\Vectorstack{+23\\ \pm 14}} & {\tiny\Vectorstack{+16\\ \pm 11}} & {\tiny\Vectorstack{+11\\ \pm 8.8}} & {\tiny\Vectorstack{+2.2\\ \pm 7.2}} &  & {\tiny\Vectorstack{+1.9\\ \pm 9.7}} & {\tiny\Vectorstack{-6.5\\ \pm 8.7}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+26\\ \pm 13}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+28\\ \pm 9.6}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+21\\ \pm 10}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+34\\ \pm 7.0}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+29\\ \pm 10}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+46\\ \pm 9.2}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+21\\ \pm 8.3}}} & {\tiny\Vectorstack{+14\\ \pm 8.8}} & {\tiny\Vectorstack{+9.2\\ \pm 7.1}} & {\tiny\Vectorstack{+0.3\\ \pm 6.0}} & {\tiny\Vectorstack{-1.9\\ \pm 9.7}} &  & {\tiny\Vectorstack{-8.5\\ \pm 4.7}} \\[2pt]
\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+35\\ \pm 13}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+37\\ \pm 11}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+30\\ \pm 8.9}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+43\\ \pm 6.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+38\\ \pm 9.0}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+55\\ \pm 8.5}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+29\\ \pm 7.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+23\\ \pm 9.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+18\\ \pm 7.2}}} & {\tiny\Vectorstack{+8.7\\ \pm 4.6}} & {\tiny\Vectorstack{+6.5\\ \pm 8.7}} & {\tiny\Vectorstack{+8.5\\ \pm 4.7}} & 
\end{tabular}}

### 2020-04-11 thesis/content/chapter-appendix.tex
completed real-world section + started with lta locality eval section

@@
 		The \ac{mlp} after the pooling layer is always configured with a single hidden layer.
+		The \acs{lta}-like configurations of 2-\acs{wl}-\acsp{gnn} are evaluated with the depths $T \in \{ 4, 5 \}$ to compensate for the missing \ac{mlp} after the pooling layer.
 	\item \textbf{Layer width $d \in \{ 32, 64 \}$:}
 		This parameter describes the output dimensionalities $d = d^{(1)} = \cdots = d^{(T)}$ of the convolutional layers and (if applicable) also the hidden layer width of the final \ac{mlp} after the pooling layer.
+		The \acs{lta}-like configurations of 2-\acs{wl}-\acsp{gnn} are evaluated with the same widths but use $d^{(T)} = 1$ in the final layer.
 	\item \textbf{Jumping knowledge $\mathrm{JK} \in \{ \mathtt{true}, \mathtt{false} \}$:}
@@
 	\item \textbf{Activation functions $\sigma$ and $\sigma_{\Gamma}$} are set to the standard logistic function.
+	\item \textbf{Number of epochs $E$ and early stopping patience $p$} are set to $E = 1000$ and $p = 100$, except for the evaluation of the synthetic TRIANGLE dataset for which we used $E = 5000$ and $p = 1000$ to ensure model convergence.
 \end{itemize}
@@
 	\caption{Sizes of the evaluated binary classification datasets and their graphs.}\label{tbl:appendix:ds-stats}
-	\centering
+	\centering\small
 	\csvreader[
-		tabular={lrrrrrrrr},
+		tabular={lrrrrrrrrr},
 		separator=comma,
+		before reading=\setlength{\tabcolsep}{5pt},
 		table head={%
-			\multicolumn{1}{c}{} & \multicolumn{1}{c}{} & &\multicolumn{3}{c}{vertex count $\left|\mathcal{V}_G\right|$} & \multicolumn{3}{c}{edge count $\left|\mathcal{E}_G\right|$} \\%
-			& \multirow{-2}{*}[-0.2em]{\shortstack[c]{no.\ of\\ graphs}} & \multirow{-2}{*}[-0.2em]{\shortstack[c]{vertex data\\(feat.\ + lab.)}} & $\min$ & $\mean$ & $\max$ & $\min$ & $\mean$ & $\max$ \\\toprule%
+			\multicolumn{1}{c}{} & \multicolumn{1}{c}{} & &\multicolumn{3}{c}{vertex count $\left|\mathcal{V}_G\right|$} & \multicolumn{3}{c}{edge count $\left|\mathcal{E}_G\right|$} & \multicolumn{1}{c}{radius} \\%
+			& \multirow{-2}{*}[-0.2em]{\shortstack[c]{no.\ of\\ graphs}} & \multirow{-2}{*}[-0.2em]{\shortstack[c]{vertex data\\{\scriptsize (feat.\ + lab.)}}} & $\min$ & $\mean$ & $\max$ & $\min$ & $\mean$ & $\max$ & $\mean \pm\, \sigma$ \\\toprule% chktex 21
 		},
@@
 	node_degree_min=\ndegmin,node_degree_mean=\ndegmean,node_degree_max=\ndegmax,%
-	dim_node_features=\nfdim,dim_edge_features=\efdim%
+	dim_node_features=\nfdim,dim_edge_features=\efdim, radius_mean=\radiusmean, radius_std=\radiusstd%
 	}%
@@
 	$\ncountmin$&$\ncountmean$&$\ncountmax$&%
-	$\ecountmin$&$\ecountmean$&$\ecountmax$%
+	$\ecountmin$&$\ecountmean$&$\ecountmax$&$\radiusmean \pm \radiusstd$%
 	}
@@
 
-In the following \crefrange{tbl:appendix:diff-triangle}{tbl:appendix:diff-imdb} (see \cpagerefrange{tbl:appendix:diff-triangle}{tbl:appendix:diff-imdb}), the differences are computed as $\textit{row accuracy} - \textit{column accuracy}$.
+In the following \crefrange{tbl:appendix:diff-triangle}{tbl:appendix:diff-imdb} (\cpagerefrange{tbl:appendix:diff-triangle}{tbl:appendix:diff-imdb}), the differences are computed as $\textit{row accuracy} - \textit{column accuracy}$.
 For each row $i$ and column $j$ we highlight the corresponding cell $(i,j)$ in \textcolor{t_red}{red} or \textcolor{t_darkgreen}{green} iff.\ the learner~$i$ performs consistently \textcolor{t_red}{worse} (or \textcolor{t_darkgreen}{better} respectively) than $j$ with a significance level of $2\sigma$.

### 2020-04-11 thesis/content/chapter-evaluation.tex
completed real-world section + started with lta locality eval section

@@
 \end{enumerate}
-This chapter will tackle those two aspects in three steps:
+This chapter will tackle those two aspects in four steps:
 \begin{enumerate*}[label={\circled{\small\arabic*}}]
@@
 	 	There we will illustrate the higher expressive power of 2-\acs{wl}-\acsp{gnn} when compared to other \ac{gcr} approaches which confirms the theoretical results from \cref{sec:ltd:wl2gnn:properties}.
-	\item Finally evaluation results on real-world datasets are described in \cref{sec:eval:real}.
-		There we will see how 2-\acs{wl}-\acsp{gnn} compare to other \acp{gnn} in practice as well as how \acs{lta}-like models compare to non-\acs{lta} models.
+	\item Then evaluation results on real-world datasets are described in \cref{sec:eval:real}.
+		There we will see how 2-\acs{wl}-\acsp{gnn} compare to other \acp{gnn} in practice as well as how \acs{lta}-like models compare to non-\acs{lta} model.
+	\item Finally, \cref{sec:eval:lta} looks at the synthetic and real-world results from an \ac{lta} perspective.
+		There we will see how the predictive performance of an \acs{lta}-like model relates to the size and locality of its constituents.
 \end{enumerate*}
@@
 	\item \textbf{\ac{wl} subtree kernel (\acs{wl}\textsubscript{ST})} with the iteration counts $T \in \{ 1,2,3,4,5 \}$ to evaluate the influence of the depth of \ac{bfs} subtrees which span \ac{lta} constituents.
-	\item \textbf{\ac{wl} shortest path kernel (\acs{wl}\textsubscript{SP})} with the iteration count $T = 5$.
+	\item \textbf{\ac{wl} shortest path kernel (\acs{wl}\textsubscript{SP})} with the iteration count $T = 3$.
 	\item \textbf{2-LWL kernel} with the iteration count $T = 3$.
@@
 		head to column names,
-		filter=\equal{\isDefault}{1}
+		filter=\equal{\isDefault}{1}\or\equal{\id}{4}
 	]{data/results.csv}{}{%
-		\ifthenelse{\equal{\id}{0}}{\multirow{5}{*}[0em]{\rotatebox[origin=c]{90}{\small\textsc{Kernel}}}}{}%
+		\ifthenelse{\equal{\id}{0}}{\multirow{6}{*}[0em]{\rotatebox[origin=c]{90}{\small\textsc{Kernel}}}}{}%
 		\ifthenelse{\equal{\id}{9}}{\multirow{8}{*}[0em]{\rotatebox[origin=c]{90}{\small\textsc{\ac{gnn}}}}}{} &%
 		\textbf{\ifthenelse{\equal{\isLta}{1}}{\textcolor{t_darkgreen}{\model*}}{\model\ifthenelse{\equal{\model}{2-WL-GNN}}{\phantom{*}}{}}} ($\params$) & \ifthenelse{\equal{\isLta}{1}}{\badgeboxinline[t_green]{\scriptsize \acs*{lta}-like}}{} &%
-		\evalres{\triangleBestTrain}{\triangleTrainMean}{\triangleTrainStd} & \evalres{\triangleBestTest}{\triangleTestMean}{\triangleTestStd}%
+		\evalres[\and\not\equal{\id}{8}]{\triangleBestTrain}{\triangleTrainMean}{\triangleTrainStd} & \evalres{\triangleBestTest}{\triangleTestMean}{\triangleTestStd}%
 	}
@@
 			& %
-			\multicolumn{1}{c}{\textbf{NCI1}} &%
-			\multicolumn{1}{c}{\textbf{PROTEINS}} &%
-			\multicolumn{1}{c}{\textbf{D\&D}} &%
-			\multicolumn{1}{c}{\textbf{REDDIT}} &%
-			\multicolumn{1}{c}{\textbf{IMDB}}%
+			\multicolumn{1}{c}{\hyperref[tbl:appendix:diff-nci]{\textbf{NCI1}}} &%
+			\multicolumn{1}{c}{\hyperref[tbl:appendix:diff-proteins]{\textbf{PROTEINS}}} &%
+			\multicolumn{1}{c}{\hyperref[tbl:appendix:diff-dd]{\textbf{D\&D}}} &%
+			\multicolumn{1}{c}{\hyperref[tbl:appendix:diff-reddit]{\textbf{REDDIT}}} &%
+			\multicolumn{1}{c}{\hyperref[tbl:appendix:diff-imdb]{\textbf{IMDB}}}%
 			\\\toprule%
@@
 		The REDDIT and IMDB datasets~\cite{Yanardag2015} contain graphs representing the relations between users in online discussions and movie actors respectively.
-		For REDDIT the type of a discussion thread has to be predicted; for IMDB the goal is to determine movie genres.
+		For REDDIT the type of a community has to be predicted based on a discussion thread; for IMDB the goal is to determine movie genres.
 \end{enumerate}
@@
 If we look at the enzyme detection problem (PROTEINS and D\&D), we observe that all evaluated approaches appear to be unable to leverage structural information for a significant improvement over the baseline learner (see \cref{tbl:appendix:diff-proteins,{tbl:appendix:diff-dd}}).
-In the social network datasets (REDDIT and IMDB) on the other hand, the structure aware methods clearly outperform the baseline (see \cref{tbl:appendix:diff-imdb}).
+In the social network datasets (REDDIT and IMDB) on the other hand, the structure aware methods clearly outperform the baseline (see \cref{tbl:appendix:diff-reddit,tbl:appendix:diff-imdb}).
 This confirms the very similar results of \citet{Errica2020}.
 
-\begin{figure}[ht]
-	\centering
+Lastly, note that the \acs{lta}-like 2-\acs{wl}-\acs{gnn} configurations either perform significantly worse or roughly similar to their non-\acs{lta} counterparts.
+This mirrors our result on the synthetic triangle detection dataset.
+Regarding the \acs{lta} assumption (see \cfullref{defn:ltag:lta-assumption}), this is evidence that the decomposition function $\psi$ of 2-\acs{wl}-\acsp{gnn} (see \cfullref{thm:ltag:gcnn-ltag-formulation}) does not produce constituents $c_i \in \psi(G)$ whose local evaluations $f(c_i) \in \mathcal{Y}$ are as indicative of the graph $G$'s class $y \in \mathcal{Y}$ as an arbitrary feature vector $z_i \in \mathbb{R}^{d^{(T)}}$.
+This does however not imply, that the evaluated real-world problem domains are generally incompatible with the \ac{lta} assumption.
+When considering the results of the \ac{lta}-like \acs{wl}\textsubscript{ST} model, we find that it is mostly comparable with the alternative non-\acs{lta} approaches and, in the case of NCI1, even significantly better.
+This shows that local constituent evaluations are in principle suitable for graph classification tasks if the right decomposition, evaluation and aggregation functions are chosen.
+
+\section{Evaluation of \acs*{lta} Constituent Locality}%
+\label{sec:eval:lta}
+
+To conclude our evaluation, we now look at the influence of constituent sizes on the test accuracy of \acs{lta}-like models, such as \acs{wl}\textsubscript{ST} and 2-\acs{wl}-\acsp{gnn}.
+As described in \cref{sec:ltag:formulation:svm,sec:ltag:formulation:gcnn:conv}, the constituents in the \ac{lta} formulations of both of these models are spanned by \ac{bfs} subtrees of some depth $T$.
+Thus, for large $T$, decompositions consist of just the connected components of a graph, i.e.\ they are only partially \acs{lta}-like.
+
+We now analyze for which degree of constituent locality \acs{wl}\textsubscript{ST} and 2-\acs{wl}-\acsp{gnn} perform best by evaluating them using varying tree depths $T$.
+For \acs{wl}\textsubscript{ST} we can set $T$ directly via the number of refinement steps.
+\Cref{fig:eval:wlst-depths} shows the the accuracies of \acs{wl}\textsubscript{ST} for $T \in \{ 1, \dots, 5 \}$.
+On all evaluated datasets, the test accuracy tends to be optimal for some $T \in \{ 1, 2, 3 \}$.%
+\newcommand{\wlstDepthPlot}[4][ymin=70, ymax=100, try min ticks=6]{%
 \begin{tikzpicture}
 	\begin{axis}[
-			width=0.33\linewidth,
+		width=0.35\linewidth,
 		height=0.3\linewidth,
-			try min ticks=6,
-			xmin=1, xmax=5,
-			ymin=70, ymax=100,
+		xmin=0.7, xmax=5.3,
+		#1,
 		label style={font=\tiny},
@@
 		ytick style={draw=none},
+		ylabel shift=-6pt,
+		xlabel shift=-5pt,
 		axis line style={gray},
-			mark size={1.5pt}
+		mark size={1.5pt},
+		title={\small #3},
+		title style={yshift=-0.5em}
 	]
+		\begin{scope}[gray]
+			\draw[gray, dashed] ({axis cs:#4,0}|-{rel axis cs:0,1}) -- ({axis cs:#4,0}|-{rel axis cs:0,0});
+		\end{scope}
 		\addplot [color=t_blue, only marks, mark=*]
 		plot[error bars/.cd, y dir=both, y explicit]
-			table [x=T, y=nciTrainMean, y error=nciTrainStd, col sep=semicolon] {data/wlst_depths.csv};
+		table [x=T, y=#2TrainMean, y error=#2TrainStd, col sep=semicolon] {data/wlst_depths.csv};
+		\label{pgfplots:eval:wlst-depth:#2-train}
 		\addplot [color=t_red, only marks, mark=square*]
 		plot[error bars/.cd, y dir=both, y explicit]
-			table [x=T, y=nciTestMean, y error=nciTestStd, col sep=semicolon] {data/wlst_depths.csv};
+		table [x=T, y=#2TestMean, y error=#2TestStd, col sep=semicolon] {data/wlst_depths.csv};
+		\label{pgfplots:eval:wlst-depth:#2-test}
 	\end{axis}
-	\end{tikzpicture}
-	\caption{test}\label{fig:eval:wlst-depths}
+\end{tikzpicture}} % chktex 31
+\begin{figure}[t]
+	\centering
+	\begin{tabular}{cccc}
+		\wlstDepthPlot[ymin=40, ymax=100, try min ticks=6, ylabel={accuracy}]{triangle}{TRIANGLE}{2.3} &
+		\wlstDepthPlot{nci}{NCI1}{7} &
+		\wlstDepthPlot{proteins}{PROTEINS}{6.1} &
+		\multirow{2}{*}[1.5em]{\rotatebox[origin=c]{90}{\scriptsize%
+			\ref{pgfplots:eval:wlst-depth:triangle-test}~Test\quad % chktex 2
+			\ref{pgfplots:eval:wlst-depth:triangle-train}~Train % chktex 2
+		}}\\
+		\wlstDepthPlot[ymin=70, ymax=100, try min ticks=6, ylabel={accuracy}, xlabel={no.\ of iterations}]{dd}{D\&D}{10.8} &
+		\wlstDepthPlot[ymin=70, ymax=100, try min ticks=6, xlabel={no.\ of iterations}]{reddit}{REDDIT}{4.6} &
+		\wlstDepthPlot[ymin=70, ymax=100, try min ticks=6, xlabel={no.\ of iterations}]{imdb}{IMDB}{1.0} &
+	\end{tabular}
+	\caption[Training and test accuracy of \acs{wl}\textsubscript{ST} for varying iteration counts.]{
+		Accuracy of \acs{wl}\textsubscript{ST} for varying iteration counts on the six evaluated datasets from the previous sections.
+		The error bars indicate standard deviations.
+		The dashed lines indicate the mean graph radius in each dataset (if it is less than $5$).
+	}\label{fig:eval:wlst-depths}
 \end{figure}
+If we compute the mean radii of the graphs in each dataset\footnote{
+	The radius of a graph $G$ is defined as $\min_{v \in \mathcal{V}_G} \max_{u \in \mathcal{V}_G} d_{\mathrm{SP}, G}(v, u)$, i.e.\ the distance of the furthest vertex $u$ from the most central vertex $v$ of $G$.
+}, we find that they are all larger than $3$, with the exception of the triangle detection and the IMDB dataset (see \cref{tbl:appendix:ds-stats}).
+This implies that the best-performing subtree constituents do not span entire connected components but smaller substructures on average.
+The REDDIT dataset shows this very clearly; there a subtree must have a depth of at least $5$ in order to span an average graph while the best-performing subtree constituents only have a depth of $T = 2$.
+This indicates that the REDDIT community detection problem is better described by small localized constituents than by large non-localized ones.
+
+Let us now look at the constituent locality of 2-\acs{wl}-\acsp{gnn}.
+There the depth of constituent subtrees is determined by both the number of convolutional layers and the neighborhood radius $r$.
+Just like the iteration count in \acs{wl}\textsubscript{ST}, the number of layers directly determines the number of neighborhood aggregation steps and therefore the subtree sizes.
+The neighborhood radius $r$ on the other hand determines how many new edges are present in $G^r$ in comparison to $G$;
+when $r$ is increased, the \ac{bfs} subtrees can span larger constituents due to the additional connections.
+
+In addition to its influence on constituent sizes, a higher neighborhood radius also introduces additional feature vectors into the convolved feature matrices $Z^{(t)} \in \mathbb{R}^{\left|\mathcal{E}_{G^r}\right| \times d^{(t)}}$.
+As we saw in the proof sketch of 2-\acs{wl}'s cycle counting ability (see \cfullref{prop:related:wl2-cycle-count}), those additional features/colors can carry important structural information about a graph.

### 2020-04-11 thesis/data/diffs/nci.tex
completed real-world section + started with lta locality eval section

@@
 {\setlength\tabcolsep{2.5pt}\setlength{\extrarowheight}{2pt}%
-\begin{tabular}{lccccccccccc}
-& \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=5$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-GWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{Baseline} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{GIN} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)} \\
-\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)& & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-12\\ \pm 1.3}}} & {\tiny\Vectorstack{-2.8\\ \pm 1.5}} & {\tiny\Vectorstack{+2.3\\ \pm 2.4}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.2\\ \pm 2.9}}} & {\tiny\Vectorstack{-3.5\\ \pm 2.2}} & {\tiny\Vectorstack{-1.9\\ \pm 2.9}} & {\tiny\Vectorstack{-4.4\\ \pm 2.3}} & {\tiny\Vectorstack{+3.1\\ \pm 2.5}} & {\tiny\Vectorstack{+1.6\\ \pm 2.6}} & {\tiny\Vectorstack{+3.7\\ \pm 3.6}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=5$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+12\\ \pm 1.3}}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.7\\ \pm 1.7}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+14\\ \pm 1.9}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+18\\ \pm 2.7}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.0\\ \pm 2.7}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+9.6\\ \pm 2.2}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+7.1\\ \pm 2.0}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+15\\ \pm 2.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+13\\ \pm 2.0}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+15\\ \pm 3.8}}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)&{\tiny\Vectorstack{+2.8\\ \pm 1.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.7\\ \pm 1.7}}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+5.1\\ \pm 2.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+9.0\\ \pm 3.0}}} & {\tiny\Vectorstack{-0.7\\ \pm 2.7}} & {\tiny\Vectorstack{+0.9\\ \pm 2.0}} & {\tiny\Vectorstack{-1.6\\ \pm 2.3}} & {\tiny\Vectorstack{+5.9\\ \pm 3.2}} & {\tiny\Vectorstack{+4.4\\ \pm 2.5}} & {\tiny\Vectorstack{+6.5\\ \pm 4.2}} \\[2pt]
-\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-2.3\\ \pm 2.4}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-14\\ \pm 1.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-5.1\\ \pm 2.3}}} &  & {\tiny\Vectorstack{+3.9\\ \pm 4.1}} & {\tiny\Vectorstack{-5.8\\ \pm 3.6}} & {\tiny\Vectorstack{-4.3\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.7\\ \pm 2.5}}} & {\tiny\Vectorstack{+0.8\\ \pm 3.7}} & {\tiny\Vectorstack{-0.8\\ \pm 2.7}} & {\tiny\Vectorstack{+1.4\\ \pm 4.0}} \\[2pt]
-\textbf{Baseline} ($\mathrm{sum}$)&\cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.2\\ \pm 2.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-18\\ \pm 2.7}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-9.0\\ \pm 3.0}}} & {\tiny\Vectorstack{-3.9\\ \pm 4.1}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-9.8\\ \pm 2.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.2\\ \pm 2.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 2.6}}} & {\tiny\Vectorstack{-3.1\\ \pm 3.0}} & {\tiny\Vectorstack{-4.7\\ \pm 3.1}} & {\tiny\Vectorstack{-2.6\\ \pm 5.0}} \\[2pt]
-\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{+3.5\\ \pm 2.2}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.0\\ \pm 2.7}}} & {\tiny\Vectorstack{+0.7\\ \pm 2.7}} & {\tiny\Vectorstack{+5.8\\ \pm 3.6}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+9.8\\ \pm 2.4}}} &  & {\tiny\Vectorstack{+1.6\\ \pm 3.5}} & {\tiny\Vectorstack{-0.9\\ \pm 2.2}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.6\\ \pm 2.5}}} & {\tiny\Vectorstack{+5.1\\ \pm 3.2}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+7.2\\ \pm 3.3}}} \\[2pt]
-\textbf{2-GNN} ($\mean$)&{\tiny\Vectorstack{+1.9\\ \pm 2.9}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-9.6\\ \pm 2.2}}} & {\tiny\Vectorstack{-0.9\\ \pm 2.0}} & {\tiny\Vectorstack{+4.3\\ \pm 2.5}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.2\\ \pm 2.9}}} & {\tiny\Vectorstack{-1.6\\ \pm 3.5}} &  & {\tiny\Vectorstack{-2.4\\ \pm 2.6}} & {\tiny\Vectorstack{+5.1\\ \pm 3.5}} & {\tiny\Vectorstack{+3.5\\ \pm 2.2}} & {\tiny\Vectorstack{+5.6\\ \pm 4.8}} \\[2pt]
-\textbf{2-GNN} ($\mathrm{SAM}$)&{\tiny\Vectorstack{+4.4\\ \pm 2.3}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-7.1\\ \pm 2.0}}} & {\tiny\Vectorstack{+1.6\\ \pm 2.3}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.7\\ \pm 2.5}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 2.6}}} & {\tiny\Vectorstack{+0.9\\ \pm 2.2}} & {\tiny\Vectorstack{+2.4\\ \pm 2.6}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+7.5\\ \pm 3.1}}} & {\tiny\Vectorstack{+5.9\\ \pm 3.2}} & {\tiny\Vectorstack{+8.1\\ \pm 4.2}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)&{\tiny\Vectorstack{-3.1\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-15\\ \pm 2.3}}} & {\tiny\Vectorstack{-5.9\\ \pm 3.2}} & {\tiny\Vectorstack{-0.8\\ \pm 3.7}} & {\tiny\Vectorstack{+3.1\\ \pm 3.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.6\\ \pm 2.5}}} & {\tiny\Vectorstack{-5.1\\ \pm 3.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-7.5\\ \pm 3.1}}} &  & {\tiny\Vectorstack{-1.5\\ \pm 2.2}} & {\tiny\Vectorstack{+0.6\\ \pm 2.9}} \\[2pt]
-\textbf{2-WL-GNN\phantom{*}} ($\mean$)&{\tiny\Vectorstack{-1.6\\ \pm 2.6}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-13\\ \pm 2.0}}} & {\tiny\Vectorstack{-4.4\\ \pm 2.5}} & {\tiny\Vectorstack{+0.8\\ \pm 2.7}} & {\tiny\Vectorstack{+4.7\\ \pm 3.1}} & {\tiny\Vectorstack{-5.1\\ \pm 3.2}} & {\tiny\Vectorstack{-3.5\\ \pm 2.2}} & {\tiny\Vectorstack{-5.9\\ \pm 3.2}} & {\tiny\Vectorstack{+1.5\\ \pm 2.2}} &  & {\tiny\Vectorstack{+2.1\\ \pm 3.4}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{-3.7\\ \pm 3.6}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-15\\ \pm 3.8}}} & {\tiny\Vectorstack{-6.5\\ \pm 4.2}} & {\tiny\Vectorstack{-1.4\\ \pm 4.0}} & {\tiny\Vectorstack{+2.6\\ \pm 5.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-7.2\\ \pm 3.3}}} & {\tiny\Vectorstack{-5.6\\ \pm 4.8}} & {\tiny\Vectorstack{-8.1\\ \pm 4.2}} & {\tiny\Vectorstack{-0.6\\ \pm 2.9}} & {\tiny\Vectorstack{-2.1\\ \pm 3.4}} &
+\begin{tabular}{lcccccccccccc}
+& \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-GWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{Baseline} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{GIN} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)} \\
+\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)& & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 1.3}}} & {\tiny\Vectorstack{-2.8\\ \pm 1.5}} & {\tiny\Vectorstack{+2.3\\ \pm 2.4}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.2\\ \pm 2.9}}} & {\tiny\Vectorstack{-3.5\\ \pm 2.2}} & {\tiny\Vectorstack{-1.9\\ \pm 2.9}} & {\tiny\Vectorstack{-4.4\\ \pm 2.3}} & {\tiny\Vectorstack{+3.1\\ \pm 2.5}} & {\tiny\Vectorstack{+1.6\\ \pm 2.6}} & {\tiny\Vectorstack{+3.7\\ \pm 3.6}} & {\tiny\Vectorstack{+1.5\\ \pm 4.0}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=3$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 1.3}}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.1\\ \pm 1.6}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+13\\ \pm 1.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+17\\ \pm 2.9}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+7.3\\ \pm 2.7}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.9\\ \pm 2.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.5\\ \pm 2.0}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+14\\ \pm 2.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+12\\ \pm 2.5}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+15\\ \pm 4.2}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+12\\ \pm 4.1}}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)&{\tiny\Vectorstack{+2.8\\ \pm 1.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.1\\ \pm 1.6}}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+5.1\\ \pm 2.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+9.0\\ \pm 3.0}}} & {\tiny\Vectorstack{-0.7\\ \pm 2.7}} & {\tiny\Vectorstack{+0.9\\ \pm 2.0}} & {\tiny\Vectorstack{-1.6\\ \pm 2.3}} & {\tiny\Vectorstack{+5.9\\ \pm 3.2}} & {\tiny\Vectorstack{+4.4\\ \pm 2.5}} & {\tiny\Vectorstack{+6.5\\ \pm 4.2}} & {\tiny\Vectorstack{+4.3\\ \pm 4.3}} \\[2pt]
+\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-2.3\\ \pm 2.4}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-13\\ \pm 1.8}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-5.1\\ \pm 2.3}}} &  & {\tiny\Vectorstack{+3.9\\ \pm 4.1}} & {\tiny\Vectorstack{-5.8\\ \pm 3.6}} & {\tiny\Vectorstack{-4.3\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.7\\ \pm 2.5}}} & {\tiny\Vectorstack{+0.8\\ \pm 3.7}} & {\tiny\Vectorstack{-0.8\\ \pm 2.7}} & {\tiny\Vectorstack{+1.4\\ \pm 4.0}} & {\tiny\Vectorstack{-0.8\\ \pm 5.3}} \\[2pt]
+\textbf{Baseline} ($\mathrm{sum}$)&\cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.2\\ \pm 2.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-17\\ \pm 2.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-9.0\\ \pm 3.0}}} & {\tiny\Vectorstack{-3.9\\ \pm 4.1}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-9.8\\ \pm 2.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.2\\ \pm 2.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 2.6}}} & {\tiny\Vectorstack{-3.1\\ \pm 3.0}} & {\tiny\Vectorstack{-4.7\\ \pm 3.1}} & {\tiny\Vectorstack{-2.6\\ \pm 5.0}} & {\tiny\Vectorstack{-4.7\\ \pm 3.3}} \\[2pt]
+\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{+3.5\\ \pm 2.2}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-7.3\\ \pm 2.7}}} & {\tiny\Vectorstack{+0.7\\ \pm 2.7}} & {\tiny\Vectorstack{+5.8\\ \pm 3.6}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+9.8\\ \pm 2.4}}} &  & {\tiny\Vectorstack{+1.6\\ \pm 3.5}} & {\tiny\Vectorstack{-0.9\\ \pm 2.2}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.6\\ \pm 2.5}}} & {\tiny\Vectorstack{+5.1\\ \pm 3.2}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+7.2\\ \pm 3.3}}} & {\tiny\Vectorstack{+5.0\\ \pm 4.2}} \\[2pt]
+\textbf{2-GNN} ($\mean$)&{\tiny\Vectorstack{+1.9\\ \pm 2.9}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.9\\ \pm 2.3}}} & {\tiny\Vectorstack{-0.9\\ \pm 2.0}} & {\tiny\Vectorstack{+4.3\\ \pm 2.5}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.2\\ \pm 2.9}}} & {\tiny\Vectorstack{-1.6\\ \pm 3.5}} &  & {\tiny\Vectorstack{-2.4\\ \pm 2.6}} & {\tiny\Vectorstack{+5.1\\ \pm 3.5}} & {\tiny\Vectorstack{+3.5\\ \pm 2.2}} & {\tiny\Vectorstack{+5.6\\ \pm 4.8}} & {\tiny\Vectorstack{+3.4\\ \pm 4.2}} \\[2pt]
+\textbf{2-GNN} ($\mathrm{SAM}$)&{\tiny\Vectorstack{+4.4\\ \pm 2.3}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.5\\ \pm 2.0}}} & {\tiny\Vectorstack{+1.6\\ \pm 2.3}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.7\\ \pm 2.5}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 2.6}}} & {\tiny\Vectorstack{+0.9\\ \pm 2.2}} & {\tiny\Vectorstack{+2.4\\ \pm 2.6}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+7.5\\ \pm 3.1}}} & {\tiny\Vectorstack{+5.9\\ \pm 3.2}} & {\tiny\Vectorstack{+8.1\\ \pm 4.2}} & {\tiny\Vectorstack{+5.9\\ \pm 5.0}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)&{\tiny\Vectorstack{-3.1\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-14\\ \pm 2.8}}} & {\tiny\Vectorstack{-5.9\\ \pm 3.2}} & {\tiny\Vectorstack{-0.8\\ \pm 3.7}} & {\tiny\Vectorstack{+3.1\\ \pm 3.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.6\\ \pm 2.5}}} & {\tiny\Vectorstack{-5.1\\ \pm 3.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-7.5\\ \pm 3.1}}} &  & {\tiny\Vectorstack{-1.5\\ \pm 2.2}} & {\tiny\Vectorstack{+0.6\\ \pm 2.9}} & {\tiny\Vectorstack{-1.6\\ \pm 3.9}} \\[2pt]
+\textbf{2-WL-GNN\phantom{*}} ($\mean$)&{\tiny\Vectorstack{-1.6\\ \pm 2.6}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-12\\ \pm 2.5}}} & {\tiny\Vectorstack{-4.4\\ \pm 2.5}} & {\tiny\Vectorstack{+0.8\\ \pm 2.7}} & {\tiny\Vectorstack{+4.7\\ \pm 3.1}} & {\tiny\Vectorstack{-5.1\\ \pm 3.2}} & {\tiny\Vectorstack{-3.5\\ \pm 2.2}} & {\tiny\Vectorstack{-5.9\\ \pm 3.2}} & {\tiny\Vectorstack{+1.5\\ \pm 2.2}} &  & {\tiny\Vectorstack{+2.1\\ \pm 3.4}} & {\tiny\Vectorstack{-0.1\\ \pm 3.6}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{-3.7\\ \pm 3.6}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-15\\ \pm 4.2}}} & {\tiny\Vectorstack{-6.5\\ \pm 4.2}} & {\tiny\Vectorstack{-1.4\\ \pm 4.0}} & {\tiny\Vectorstack{+2.6\\ \pm 5.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-7.2\\ \pm 3.3}}} & {\tiny\Vectorstack{-5.6\\ \pm 4.8}} & {\tiny\Vectorstack{-8.1\\ \pm 4.2}} & {\tiny\Vectorstack{-0.6\\ \pm 2.9}} & {\tiny\Vectorstack{-2.1\\ \pm 3.4}} &  & {\tiny\Vectorstack{-2.2\\ \pm 5.6}} \\[2pt]
+\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{-1.5\\ \pm 4.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-12\\ \pm 4.1}}} & {\tiny\Vectorstack{-4.3\\ \pm 4.3}} & {\tiny\Vectorstack{+0.8\\ \pm 5.3}} & {\tiny\Vectorstack{+4.7\\ \pm 3.3}} & {\tiny\Vectorstack{-5.0\\ \pm 4.2}} & {\tiny\Vectorstack{-3.4\\ \pm 4.2}} & {\tiny\Vectorstack{-5.9\\ \pm 5.0}} & {\tiny\Vectorstack{+1.6\\ \pm 3.9}} & {\tiny\Vectorstack{+0.1\\ \pm 3.6}} & {\tiny\Vectorstack{+2.2\\ \pm 5.6}} &
 \end{tabular}}

### 2020-04-11 thesis/data/diffs/reddit.tex
completed real-world section + started with lta locality eval section

@@
 \begin{tabular}{lcccccccccc}
-& \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=5$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-GWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{Baseline} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{GIN} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)} \\
-\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)& & {\tiny\Vectorstack{-1.1\\ \pm 5.3}} & {\tiny\Vectorstack{+0.5\\ \pm 4.9}} & {\tiny\Vectorstack{+0.8\\ \pm 3.6}} & {\tiny\Vectorstack{+4.2\\ \pm 10}} & {\tiny\Vectorstack{-11\\ \pm 6.9}} & {\tiny\Vectorstack{+2.9\\ \pm 3.5}} & {\tiny\Vectorstack{-7.4\\ \pm 6.2}} & {\tiny\Vectorstack{-0.9\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-12\\ \pm 3.4}}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=5$)&{\tiny\Vectorstack{+1.1\\ \pm 5.3}} &  & {\tiny\Vectorstack{+1.5\\ \pm 3.5}} & {\tiny\Vectorstack{+1.9\\ \pm 5.0}} & {\tiny\Vectorstack{+5.2\\ \pm 11}} & {\tiny\Vectorstack{-9.7\\ \pm 8.1}} & {\tiny\Vectorstack{+4.0\\ \pm 6.1}} & {\tiny\Vectorstack{-6.4\\ \pm 8.1}} & {\tiny\Vectorstack{+0.2\\ \pm 5.1}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 3.7}}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)&{\tiny\Vectorstack{-0.5\\ \pm 4.9}} & {\tiny\Vectorstack{-1.5\\ \pm 3.5}} &  & {\tiny\Vectorstack{+0.3\\ \pm 4.2}} & {\tiny\Vectorstack{+3.7\\ \pm 11}} & {\tiny\Vectorstack{-11\\ \pm 7.3}} & {\tiny\Vectorstack{+2.4\\ \pm 5.5}} & {\tiny\Vectorstack{-7.9\\ \pm 5.5}} & {\tiny\Vectorstack{-1.4\\ \pm 5.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-13\\ \pm 3.5}}} \\[2pt]
-\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-0.8\\ \pm 3.6}} & {\tiny\Vectorstack{-1.9\\ \pm 5.0}} & {\tiny\Vectorstack{-0.3\\ \pm 4.2}} &  & {\tiny\Vectorstack{+3.3\\ \pm 11}} & {\tiny\Vectorstack{-12\\ \pm 7.6}} & {\tiny\Vectorstack{+2.1\\ \pm 2.4}} & {\tiny\Vectorstack{-8.2\\ \pm 6.2}} & {\tiny\Vectorstack{-1.7\\ \pm 3.3}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-13\\ \pm 3.6}}} \\[2pt]
-\textbf{Baseline} ($\mathrm{sum}$)&{\tiny\Vectorstack{-4.2\\ \pm 10}} & {\tiny\Vectorstack{-5.2\\ \pm 11}} & {\tiny\Vectorstack{-3.7\\ \pm 11}} & {\tiny\Vectorstack{-3.3\\ \pm 11}} &  & {\tiny\Vectorstack{-15\\ \pm 12}} & {\tiny\Vectorstack{-1.2\\ \pm 12}} & {\tiny\Vectorstack{-12\\ \pm 13}} & {\tiny\Vectorstack{-5.0\\ \pm 11}} & {\tiny\Vectorstack{-16\\ \pm 10}} \\[2pt]
-\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{+11\\ \pm 6.9}} & {\tiny\Vectorstack{+9.7\\ \pm 8.1}} & {\tiny\Vectorstack{+11\\ \pm 7.3}} & {\tiny\Vectorstack{+12\\ \pm 7.6}} & {\tiny\Vectorstack{+15\\ \pm 12}} &  & {\tiny\Vectorstack{+14\\ \pm 8.3}} & {\tiny\Vectorstack{+3.3\\ \pm 9.6}} & {\tiny\Vectorstack{+9.9\\ \pm 7.4}} & {\tiny\Vectorstack{-1.3\\ \pm 7.0}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)&{\tiny\Vectorstack{-2.9\\ \pm 3.5}} & {\tiny\Vectorstack{-4.0\\ \pm 6.1}} & {\tiny\Vectorstack{-2.4\\ \pm 5.5}} & {\tiny\Vectorstack{-2.1\\ \pm 2.4}} & {\tiny\Vectorstack{+1.2\\ \pm 12}} & {\tiny\Vectorstack{-14\\ \pm 8.3}} &  & {\tiny\Vectorstack{-10\\ \pm 7.1}} & {\tiny\Vectorstack{-3.8\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-15\\ \pm 3.9}}} \\[2pt]
-\textbf{2-WL-GNN\phantom{*}} ($\mean$)&{\tiny\Vectorstack{+7.4\\ \pm 6.2}} & {\tiny\Vectorstack{+6.4\\ \pm 8.1}} & {\tiny\Vectorstack{+7.9\\ \pm 5.5}} & {\tiny\Vectorstack{+8.2\\ \pm 6.2}} & {\tiny\Vectorstack{+12\\ \pm 13}} & {\tiny\Vectorstack{-3.3\\ \pm 9.6}} & {\tiny\Vectorstack{+10\\ \pm 7.1}} &  & {\tiny\Vectorstack{+6.6\\ \pm 7.2}} & {\tiny\Vectorstack{-4.6\\ \pm 7.1}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{+0.9\\ \pm 2.5}} & {\tiny\Vectorstack{-0.2\\ \pm 5.1}} & {\tiny\Vectorstack{+1.4\\ \pm 5.0}} & {\tiny\Vectorstack{+1.7\\ \pm 3.3}} & {\tiny\Vectorstack{+5.0\\ \pm 11}} & {\tiny\Vectorstack{-9.9\\ \pm 7.4}} & {\tiny\Vectorstack{+3.8\\ \pm 2.5}} & {\tiny\Vectorstack{-6.6\\ \pm 7.2}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 2.6}}} \\[2pt]
-\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+12\\ \pm 3.4}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 3.7}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+13\\ \pm 3.5}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+13\\ \pm 3.6}}} & {\tiny\Vectorstack{+16\\ \pm 10}} & {\tiny\Vectorstack{+1.3\\ \pm 7.0}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+15\\ \pm 3.9}}} & {\tiny\Vectorstack{+4.6\\ \pm 7.1}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 2.6}}} &
+& \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-GWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{Baseline} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{GIN} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)} \\
+\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)& & {\tiny\Vectorstack{-1.8\\ \pm 1.8}} & {\tiny\Vectorstack{+0.5\\ \pm 4.9}} & {\tiny\Vectorstack{+0.8\\ \pm 3.6}} & {\tiny\Vectorstack{+4.2\\ \pm 10}} & {\tiny\Vectorstack{-11\\ \pm 6.9}} & {\tiny\Vectorstack{+2.9\\ \pm 3.5}} & {\tiny\Vectorstack{-7.4\\ \pm 6.2}} & {\tiny\Vectorstack{-0.9\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-13\\ \pm 3.6}}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=3$)&{\tiny\Vectorstack{+1.8\\ \pm 1.8}} &  & {\tiny\Vectorstack{+2.3\\ \pm 4.7}} & {\tiny\Vectorstack{+2.6\\ \pm 3.5}} & {\tiny\Vectorstack{+5.9\\ \pm 11}} & {\tiny\Vectorstack{-9.0\\ \pm 7.1}} & {\tiny\Vectorstack{+4.7\\ \pm 3.3}} & {\tiny\Vectorstack{-5.7\\ \pm 6.8}} & {\tiny\Vectorstack{+0.9\\ \pm 1.6}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 2.8}}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)&{\tiny\Vectorstack{-0.5\\ \pm 4.9}} & {\tiny\Vectorstack{-2.3\\ \pm 4.7}} &  & {\tiny\Vectorstack{+0.3\\ \pm 4.2}} & {\tiny\Vectorstack{+3.7\\ \pm 11}} & {\tiny\Vectorstack{-11\\ \pm 7.3}} & {\tiny\Vectorstack{+2.4\\ \pm 5.5}} & {\tiny\Vectorstack{-7.9\\ \pm 5.5}} & {\tiny\Vectorstack{-1.4\\ \pm 5.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-13\\ \pm 3.6}}} \\[2pt]
+\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-0.8\\ \pm 3.6}} & {\tiny\Vectorstack{-2.6\\ \pm 3.5}} & {\tiny\Vectorstack{-0.3\\ \pm 4.2}} &  & {\tiny\Vectorstack{+3.3\\ \pm 11}} & {\tiny\Vectorstack{-12\\ \pm 7.6}} & {\tiny\Vectorstack{+2.1\\ \pm 2.4}} & {\tiny\Vectorstack{-8.2\\ \pm 6.2}} & {\tiny\Vectorstack{-1.7\\ \pm 3.3}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-14\\ \pm 3.1}}} \\[2pt]
+\textbf{Baseline} ($\mathrm{sum}$)&{\tiny\Vectorstack{-4.2\\ \pm 10}} & {\tiny\Vectorstack{-5.9\\ \pm 11}} & {\tiny\Vectorstack{-3.7\\ \pm 11}} & {\tiny\Vectorstack{-3.3\\ \pm 11}} &  & {\tiny\Vectorstack{-15\\ \pm 12}} & {\tiny\Vectorstack{-1.2\\ \pm 12}} & {\tiny\Vectorstack{-12\\ \pm 13}} & {\tiny\Vectorstack{-5.0\\ \pm 11}} & {\tiny\Vectorstack{-17\\ \pm 11}} \\[2pt]
+\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{+11\\ \pm 6.9}} & {\tiny\Vectorstack{+9.0\\ \pm 7.1}} & {\tiny\Vectorstack{+11\\ \pm 7.3}} & {\tiny\Vectorstack{+12\\ \pm 7.6}} & {\tiny\Vectorstack{+15\\ \pm 12}} &  & {\tiny\Vectorstack{+14\\ \pm 8.3}} & {\tiny\Vectorstack{+3.3\\ \pm 9.6}} & {\tiny\Vectorstack{+9.9\\ \pm 7.4}} & {\tiny\Vectorstack{-2.2\\ \pm 7.2}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)&{\tiny\Vectorstack{-2.9\\ \pm 3.5}} & {\tiny\Vectorstack{-4.7\\ \pm 3.3}} & {\tiny\Vectorstack{-2.4\\ \pm 5.5}} & {\tiny\Vectorstack{-2.1\\ \pm 2.4}} & {\tiny\Vectorstack{+1.2\\ \pm 12}} & {\tiny\Vectorstack{-14\\ \pm 8.3}} &  & {\tiny\Vectorstack{-10\\ \pm 7.1}} & {\tiny\Vectorstack{-3.8\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-16\\ \pm 3.3}}} \\[2pt]
+\textbf{2-WL-GNN\phantom{*}} ($\mean$)&{\tiny\Vectorstack{+7.4\\ \pm 6.2}} & {\tiny\Vectorstack{+5.7\\ \pm 6.8}} & {\tiny\Vectorstack{+7.9\\ \pm 5.5}} & {\tiny\Vectorstack{+8.2\\ \pm 6.2}} & {\tiny\Vectorstack{+12\\ \pm 13}} & {\tiny\Vectorstack{-3.3\\ \pm 9.6}} & {\tiny\Vectorstack{+10\\ \pm 7.1}} &  & {\tiny\Vectorstack{+6.6\\ \pm 7.2}} & {\tiny\Vectorstack{-5.5\\ \pm 6.6}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{+0.9\\ \pm 2.5}} & {\tiny\Vectorstack{-0.9\\ \pm 1.6}} & {\tiny\Vectorstack{+1.4\\ \pm 5.0}} & {\tiny\Vectorstack{+1.7\\ \pm 3.3}} & {\tiny\Vectorstack{+5.0\\ \pm 11}} & {\tiny\Vectorstack{-9.9\\ \pm 7.4}} & {\tiny\Vectorstack{+3.8\\ \pm 2.5}} & {\tiny\Vectorstack{-6.6\\ \pm 7.2}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-12\\ \pm 2.6}}} \\[2pt]
+\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+13\\ \pm 3.6}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 2.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+13\\ \pm 3.6}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+14\\ \pm 3.1}}} & {\tiny\Vectorstack{+17\\ \pm 11}} & {\tiny\Vectorstack{+2.2\\ \pm 7.2}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+16\\ \pm 3.3}}} & {\tiny\Vectorstack{+5.5\\ \pm 6.6}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+12\\ \pm 2.6}}} &
 \end{tabular}}

### 2020-04-11 implementation/src/ltag/datasets/manager.py

@@ -525,13 +525,16 @@ class GraphDatasetManager:
     node_counts = []
     edge_counts = []
     radii = []
-    diameters = []
     degrees = []
 
     for g in gs:
       node_counts.append(g.order())
       edge_counts.append(g.size())
       degrees += [d for n, d in g.degree()]
+      try:
+        radii.append(nx.algorithms.distance_measures.radius(g))
+      except Exception:
+        radii.append(np.inf)
 
     return dict(
       node_counts=utils.statistics(node_counts),
@@ -540,4 +543,5 @@ class GraphDatasetManager:
       dim_node_features=self.dim_node_features,
       dim_edge_features=self.dim_edge_features,
       num_node_labels=self.num_node_labels,
-      num_edge_labels=self.num_edge_labels)
+      num_edge_labels=self.num_edge_labels,
+      radii=utils.statistics(radii, mask_invalid=True))

### 2020-04-11 implementation/src/ltag/utils.py

@@ -21,13 +21,26 @@ class NumpyEncoder(json.JSONEncoder):
     return json.JSONEncoder.default(self, obj)
 
 def statistics(vals, mask_invalid=False):
-  return {
-    "mean": np.mean(np.ma.masked_invalid(vals) if mask_invalid else vals),
-    "std": np.std(np.ma.masked_invalid(vals) if mask_invalid else vals),
-    "min": np.min(vals),
-    "max": np.max(vals),
-    "count": len(vals) if isinstance(vals, Sized) else 1
-  }
+  if mask_invalid:
+    vals_masked = np.ma.masked_invalid(vals)
+    return {
+      "mean": np.mean(vals_masked),
+      "std": np.std(vals_masked),
+      "min": np.min(vals),
+      "max": np.max(vals),
+      "max_masked": np.max(vals_masked),
+      "min_masked": np.min(vals_masked),
+      "count": len(vals),
+      "count_masked": vals_masked.count()
+    }
+  else:
+    return {
+      "mean": np.mean(vals),
+      "std": np.std(vals),
+      "min": np.min(vals),
+      "max": np.max(vals),
+      "count": len(vals) if isinstance(vals, Sized) else 1
+    }
 
 @contextlib.contextmanager
 def local_seed(seed):

### 2020-04-12 thesis/content/chapter-appendix.tex
completed evaluation chapter

@@
 
-In the following \crefrange{tbl:appendix:diff-triangle}{tbl:appendix:diff-imdb} (\cpagerefrange{tbl:appendix:diff-triangle}{tbl:appendix:diff-imdb}), the differences are computed as $\textit{row accuracy} - \textit{column accuracy}$.
-For each row $i$ and column $j$ we highlight the corresponding cell $(i,j)$ in \textcolor{t_red}{red} or \textcolor{t_darkgreen}{green} iff.\ the learner~$i$ performs consistently \textcolor{t_red}{worse} (or \textcolor{t_darkgreen}{better} respectively) than $j$ with a significance level of $2\sigma$.
-To compute the deltas for 2-\acs{wl}-\acsp{gnn}, we use the same neighborhood radii as in \cref{tbl:eval:synthetic,tbl:eval:real}, i.e.\ $r = 2$ for the synthetic triangle detection dataset and $r = 8$, $5$, $2$, $1$ and $4$ for NCI1, PROTEINS, D\&D, REDDIT and IMDB respectively.
+In the following \crefrange{tbl:appendix:diff-triangle}{tbl:appendix:diff-imdb} (\cpagerefrange{tbl:appendix:diff-triangle}{tbl:appendix:diff-imdb}), we show accuracy differences as $\textit{row accuracy} - \textit{column accuracy}$.
+For each row $i$ and column $j$ the corresponding cell $(i,j)$ is highlighted in \textcolor{t_red}{red} or \textcolor{t_darkgreen}{green} iff.\ the learner~$i$ performs consistently \textcolor{t_red}{worse} (or \textcolor{t_darkgreen}{better} respectively) than $j$ with a significance level of $2\sigma$.
+To compute the deltas for 2-\acs{wl}-\acsp{gnn}, the same neighborhood radii as in \cref{tbl:eval:synthetic,tbl:eval:real} are used, i.e.\ $r = 2$ for the synthetic triangle detection dataset and $r = 8$, $5$, $2$, $1$ and $4$ for NCI1, PROTEINS, D\&D, REDDIT and IMDB respectively.
 

### 2020-04-12 thesis/content/chapter-evaluation.tex
completed evaluation chapter

@@
 The structure aware learners on the other hand all perform better than random guessing and are in fact mostly able to fit the training data perfectly.
-This shows that all generated graphs are 1-\acs{wl} distinguishable; the \ac{wl} subtree kernel, for example, can simply memorize the training graphs via their unique 1-\acs{wl} color distribution after $T = 5$ refinement steps.
+This shows that all generated graphs are 1-\acs{wl} distinguishable; the \ac{wl} subtree kernel \ac{svm}, for example, can simply ``memorize'' the training graphs via their unique 1-\acs{wl} color distribution after $T = 5$ refinement steps.
 
@@
 
-Finally, if we look at the two evaluated pooling layers, static $\mean$ pooling and \ac{sampool}, we see that the attention mechanism of the latter significantly improves the generalization performance of both 2-\acsp{gnn} and 2-\acs{wl}-\acsp{gnn}.
+Finally, if we look at the two evaluated pooling layers, static $\mean$ pooling and \ac{sampool}, we see that the attention mechanism of the latter tends to improve the generalization performance of both 2-\acsp{gnn} and 2-\acs{wl}-\acsp{gnn}.
 This indicates that \ac{sampool} is successful at filtering out the randomly generated noisy parts of a given graph and putting most attention to the relevant unicolored triangle.
@@
 	For REDDIT and IMDB however, we used the same splits.
+	Also note that some values are missing because of \acf{oom} errors;
+	the resource use for all evaluations was limited to \SI{64}{\giga\byte} main memory and \SI{11}{\giga\byte} GPU memory.
+	On average, each \ac{gnn} assessment (see \cref{algo:eval:assessment}) took about three days on an AMD 1800X 16-thread CPU and an Nvidia 1080~Ti GPU.% chktex 29
 }.
@@
 As we saw in the proof sketch of 2-\acs{wl}'s cycle counting ability (see \cfullref{prop:related:wl2-cycle-count}), those additional features/colors can carry important structural information about a graph.
+\newcommand{\wlRadiusPlot}[4][width=0.22\linewidth]{%
+\begin{tikzpicture}
+	\begin{axis}[
+		height=0.3\linewidth,
+		#1,
+		label style={font=\tiny},
+		tick label style={font=\tiny},
+		ymajorgrids,
+		ytick style={draw=none},
+		ylabel shift=-6pt,
+		xlabel shift=-5pt,
+		axis line style={gray},
+		mark size={1.5pt},
+		title style={yshift=-0.3em},
+		#3
+	]
+		\addplot [color=t_blue, only marks, mark=*]
+		plot[error bars/.cd, y dir=both, y explicit]
+		table [x=r, y=#2TrainMean, y error=#2TrainStd, col sep=semicolon] {data/wl2_radii_#4.csv};
+		\label{pgfplots:eval:wl-radius:#4-#2-train}
+		\addplot [color=t_red, only marks, mark=square*]
+		plot[error bars/.cd, y dir=both, y explicit]
+		table [x=r, y=#2TestMean, y error=#2TestStd, col sep=semicolon] {data/wl2_radii_#4.csv};
+		\label{pgfplots:eval:wl-radius:#4-#2-test}
+	\end{axis}
+\end{tikzpicture}} % chktex 31
+\begin{figure}[t]
+	\centering
+	\setlength{\tabcolsep}{1pt}
+	\begin{tabular}{m{1em}cccccc}
+		\raisebox{0.27\height}[0pt][0pt]{\rotatebox{90}{\small\textsc{$\mean$~pooling}}} &
+		\wlRadiusPlot{triangle}{title={\small TRIANGLE},ylabel={accuracy},xtick={1,2},xmin=0,xmax=3,ymin=80,ymax=100,try min ticks=4}{mean} &
+		\wlRadiusPlot{nci}{title={\small NCI1},xtick={1,8},xmin=-6,xmax=15,ymin=60,ymax=80,try min ticks=4}{mean} &
+		\wlRadiusPlot{proteins}{title={\small PROTEINS},xtick={1,5},xmin=-3,xmax=9,ymin=65,ymax=85,try min ticks=4}{mean} &
+		\wlRadiusPlot{dd}{title={\small D\&D},xtick={1,2},xmin=0,xmax=3,ymin=65,ymax=85,try min ticks=4}{mean} &
+		\wlRadiusPlot[width=0.33\linewidth]{imdb}{title={\small IMDB},xtick={1,2,4,6,8},ymin=60,ymax=80,try min ticks=4}{mean} &
+		\multirow{2}{*}[1.5em]{\hspace{8pt}\rotatebox[origin=c]{90}{\scriptsize%
+			\ref{pgfplots:eval:wl-radius:mean-triangle-test}~Test\quad % chktex 2
+			\ref{pgfplots:eval:wl-radius:mean-triangle-train}~Train % chktex 2
+		}} \\
+		\raisebox{0.33\height}[0pt][0pt]{\rotatebox{90}{\small\textsc{$\mathrm{SAM}$~pooling}}} &
+		\wlRadiusPlot{triangle}{ylabel={accuracy},xtick={1,2},xmin=0,xmax=3,ymin=80,ymax=100,try min ticks=4, xlabel={radius $r$}}{sam} &
+		\wlRadiusPlot{nci}{xtick={1,8},xmin=-6,xmax=15,ymin=60,ymax=80,try min ticks=4, xlabel={radius $r$}}{sam} &
+		\wlRadiusPlot{proteins}{xtick={1,5},xmin=-3,xmax=9,ymin=65,ymax=85,try min ticks=4, xlabel={radius $r$}}{sam} &
+		\wlRadiusPlot{dd}{xtick={1,2},xmin=0,xmax=3,ymin=65,ymax=85,try min ticks=4, xlabel={radius $r$}}{sam} &
+		\wlRadiusPlot[width=0.33\linewidth]{imdb}{xtick={1,2,4,6,8},ymin=60,ymax=80,try min ticks=4, xlabel={radius $r$}}{sam} &
+	\end{tabular}
+	\caption[Accuracy of the \ac{lta}-like 2-\acs{wl}-\acs{gnn} configurations for varying neighborhood radii.]{
+		Accuracy of the \ac{lta}-like 2-\acs{wl}-\acs{gnn} configurations for varying neighborhood radii $r$.
+		All datasets were evaluated on $r = 1$ and the highest radius for which the 2-\acs{wl} graph encodings would still fit into memory;
+		on the REDDIT dataset all radii larger than $1$ produced \ac{oom} errors, therefore it is not shown here.
+	}\label{fig:eval:wl2-gnn-radius}
+\end{figure}
+\Cref{fig:eval:wl2-gnn-radius} shows that adding this information via a neighborhood radius $r > 1$ does correlate with a higher training and test accuracy on the NCI1 and D\&D datasets; on the IMDB and PROTEINS datasets this is not the case.
+This difference is interesting because NCI1 (molecular structures) and D\&D (protein sequences) contain more cyclic graphs while IMDB (ego-network structures) and PROTEINS (protein sequences) consist of more tree- or list-like graphs (see \cref{sec:appendix:ds-stats}).
+Even though the PROTEINS and D\&D datasets both contain protein sequences, we find that the protein sequences in the PROTEINS dataset are very ``list-like'' with much fewer large cycles than in the proteins structures of the D\&D dataset (compare the vertex and edge count statistics of both datasets in \cref{tbl:appendix:ds-stats}).
+\Cref{fig:evaluation:proteins-dd-diff} illustrates this difference.
+This leads us the the hypothesis that 2-\acs{wl}-\acsp{gnn} with a neighborhood radius of $r > 1$ are able to improve their real-world performance over that achieved with $r = 1$ by detecting cyclic constituents in graphs.
+Due to the limited number of evaluation results, further investigations are required to verify this hypothesis.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.8\linewidth]{gfx/evaluation/proteins-dd-diff.pdf}
+	\caption{
+		Two samples illustrating the difference between the PROTEINS and D\&D datasets.
+	}\label{fig:evaluation:proteins-dd-diff}
+\end{figure}

### 2020-04-12 thesis/data/diffs/nci.tex
completed evaluation chapter

@@
 & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-GWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{Baseline} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{GIN} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)} \\
-\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)& & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 1.3}}} & {\tiny\Vectorstack{-2.8\\ \pm 1.5}} & {\tiny\Vectorstack{+2.3\\ \pm 2.4}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.2\\ \pm 2.9}}} & {\tiny\Vectorstack{-3.5\\ \pm 2.2}} & {\tiny\Vectorstack{-1.9\\ \pm 2.9}} & {\tiny\Vectorstack{-4.4\\ \pm 2.3}} & {\tiny\Vectorstack{+3.1\\ \pm 2.5}} & {\tiny\Vectorstack{+1.6\\ \pm 2.6}} & {\tiny\Vectorstack{+3.7\\ \pm 3.6}} & {\tiny\Vectorstack{+1.5\\ \pm 4.0}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=3$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 1.3}}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.1\\ \pm 1.6}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+13\\ \pm 1.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+17\\ \pm 2.9}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+7.3\\ \pm 2.7}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.9\\ \pm 2.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.5\\ \pm 2.0}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+14\\ \pm 2.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+12\\ \pm 2.5}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+15\\ \pm 4.2}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+12\\ \pm 4.1}}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)&{\tiny\Vectorstack{+2.8\\ \pm 1.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.1\\ \pm 1.6}}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+5.1\\ \pm 2.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+9.0\\ \pm 3.0}}} & {\tiny\Vectorstack{-0.7\\ \pm 2.7}} & {\tiny\Vectorstack{+0.9\\ \pm 2.0}} & {\tiny\Vectorstack{-1.6\\ \pm 2.3}} & {\tiny\Vectorstack{+5.9\\ \pm 3.2}} & {\tiny\Vectorstack{+4.4\\ \pm 2.5}} & {\tiny\Vectorstack{+6.5\\ \pm 4.2}} & {\tiny\Vectorstack{+4.3\\ \pm 4.3}} \\[2pt]
-\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-2.3\\ \pm 2.4}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-13\\ \pm 1.8}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-5.1\\ \pm 2.3}}} &  & {\tiny\Vectorstack{+3.9\\ \pm 4.1}} & {\tiny\Vectorstack{-5.8\\ \pm 3.6}} & {\tiny\Vectorstack{-4.3\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.7\\ \pm 2.5}}} & {\tiny\Vectorstack{+0.8\\ \pm 3.7}} & {\tiny\Vectorstack{-0.8\\ \pm 2.7}} & {\tiny\Vectorstack{+1.4\\ \pm 4.0}} & {\tiny\Vectorstack{-0.8\\ \pm 5.3}} \\[2pt]
-\textbf{Baseline} ($\mathrm{sum}$)&\cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.2\\ \pm 2.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-17\\ \pm 2.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-9.0\\ \pm 3.0}}} & {\tiny\Vectorstack{-3.9\\ \pm 4.1}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-9.8\\ \pm 2.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.2\\ \pm 2.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 2.6}}} & {\tiny\Vectorstack{-3.1\\ \pm 3.0}} & {\tiny\Vectorstack{-4.7\\ \pm 3.1}} & {\tiny\Vectorstack{-2.6\\ \pm 5.0}} & {\tiny\Vectorstack{-4.7\\ \pm 3.3}} \\[2pt]
-\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{+3.5\\ \pm 2.2}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-7.3\\ \pm 2.7}}} & {\tiny\Vectorstack{+0.7\\ \pm 2.7}} & {\tiny\Vectorstack{+5.8\\ \pm 3.6}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+9.8\\ \pm 2.4}}} &  & {\tiny\Vectorstack{+1.6\\ \pm 3.5}} & {\tiny\Vectorstack{-0.9\\ \pm 2.2}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.6\\ \pm 2.5}}} & {\tiny\Vectorstack{+5.1\\ \pm 3.2}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+7.2\\ \pm 3.3}}} & {\tiny\Vectorstack{+5.0\\ \pm 4.2}} \\[2pt]
-\textbf{2-GNN} ($\mean$)&{\tiny\Vectorstack{+1.9\\ \pm 2.9}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.9\\ \pm 2.3}}} & {\tiny\Vectorstack{-0.9\\ \pm 2.0}} & {\tiny\Vectorstack{+4.3\\ \pm 2.5}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.2\\ \pm 2.9}}} & {\tiny\Vectorstack{-1.6\\ \pm 3.5}} &  & {\tiny\Vectorstack{-2.4\\ \pm 2.6}} & {\tiny\Vectorstack{+5.1\\ \pm 3.5}} & {\tiny\Vectorstack{+3.5\\ \pm 2.2}} & {\tiny\Vectorstack{+5.6\\ \pm 4.8}} & {\tiny\Vectorstack{+3.4\\ \pm 4.2}} \\[2pt]
-\textbf{2-GNN} ($\mathrm{SAM}$)&{\tiny\Vectorstack{+4.4\\ \pm 2.3}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.5\\ \pm 2.0}}} & {\tiny\Vectorstack{+1.6\\ \pm 2.3}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.7\\ \pm 2.5}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 2.6}}} & {\tiny\Vectorstack{+0.9\\ \pm 2.2}} & {\tiny\Vectorstack{+2.4\\ \pm 2.6}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+7.5\\ \pm 3.1}}} & {\tiny\Vectorstack{+5.9\\ \pm 3.2}} & {\tiny\Vectorstack{+8.1\\ \pm 4.2}} & {\tiny\Vectorstack{+5.9\\ \pm 5.0}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)&{\tiny\Vectorstack{-3.1\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-14\\ \pm 2.8}}} & {\tiny\Vectorstack{-5.9\\ \pm 3.2}} & {\tiny\Vectorstack{-0.8\\ \pm 3.7}} & {\tiny\Vectorstack{+3.1\\ \pm 3.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.6\\ \pm 2.5}}} & {\tiny\Vectorstack{-5.1\\ \pm 3.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-7.5\\ \pm 3.1}}} &  & {\tiny\Vectorstack{-1.5\\ \pm 2.2}} & {\tiny\Vectorstack{+0.6\\ \pm 2.9}} & {\tiny\Vectorstack{-1.6\\ \pm 3.9}} \\[2pt]
-\textbf{2-WL-GNN\phantom{*}} ($\mean$)&{\tiny\Vectorstack{-1.6\\ \pm 2.6}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-12\\ \pm 2.5}}} & {\tiny\Vectorstack{-4.4\\ \pm 2.5}} & {\tiny\Vectorstack{+0.8\\ \pm 2.7}} & {\tiny\Vectorstack{+4.7\\ \pm 3.1}} & {\tiny\Vectorstack{-5.1\\ \pm 3.2}} & {\tiny\Vectorstack{-3.5\\ \pm 2.2}} & {\tiny\Vectorstack{-5.9\\ \pm 3.2}} & {\tiny\Vectorstack{+1.5\\ \pm 2.2}} &  & {\tiny\Vectorstack{+2.1\\ \pm 3.4}} & {\tiny\Vectorstack{-0.1\\ \pm 3.6}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{-3.7\\ \pm 3.6}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-15\\ \pm 4.2}}} & {\tiny\Vectorstack{-6.5\\ \pm 4.2}} & {\tiny\Vectorstack{-1.4\\ \pm 4.0}} & {\tiny\Vectorstack{+2.6\\ \pm 5.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-7.2\\ \pm 3.3}}} & {\tiny\Vectorstack{-5.6\\ \pm 4.8}} & {\tiny\Vectorstack{-8.1\\ \pm 4.2}} & {\tiny\Vectorstack{-0.6\\ \pm 2.9}} & {\tiny\Vectorstack{-2.1\\ \pm 3.4}} &  & {\tiny\Vectorstack{-2.2\\ \pm 5.6}} \\[2pt]
-\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{-1.5\\ \pm 4.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-12\\ \pm 4.1}}} & {\tiny\Vectorstack{-4.3\\ \pm 4.3}} & {\tiny\Vectorstack{+0.8\\ \pm 5.3}} & {\tiny\Vectorstack{+4.7\\ \pm 3.3}} & {\tiny\Vectorstack{-5.0\\ \pm 4.2}} & {\tiny\Vectorstack{-3.4\\ \pm 4.2}} & {\tiny\Vectorstack{-5.9\\ \pm 5.0}} & {\tiny\Vectorstack{+1.6\\ \pm 3.9}} & {\tiny\Vectorstack{+0.1\\ \pm 3.6}} & {\tiny\Vectorstack{+2.2\\ \pm 5.6}} &
+\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)& & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 1.3}}} & {\tiny\Vectorstack{-2.8\\ \pm 1.5}} & {\tiny\Vectorstack{+2.3\\ \pm 2.4}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.2\\ \pm 2.9}}} & {\tiny\Vectorstack{-3.5\\ \pm 2.2}} & {\tiny\Vectorstack{-1.9\\ \pm 2.9}} & {\tiny\Vectorstack{-4.4\\ \pm 2.3}} & {\tiny\Vectorstack{+3.1\\ \pm 2.5}} & {\tiny\Vectorstack{+1.6\\ \pm 2.6}} & {\tiny\Vectorstack{+3.7\\ \pm 3.6}} & {\tiny\Vectorstack{+0.2\\ \pm 3.5}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=3$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 1.3}}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.1\\ \pm 1.6}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+13\\ \pm 1.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+17\\ \pm 2.9}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+7.3\\ \pm 2.7}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.9\\ \pm 2.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.5\\ \pm 2.0}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+14\\ \pm 2.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+12\\ \pm 2.5}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+15\\ \pm 4.2}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 3.4}}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)&{\tiny\Vectorstack{+2.8\\ \pm 1.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.1\\ \pm 1.6}}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+5.1\\ \pm 2.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+9.0\\ \pm 3.0}}} & {\tiny\Vectorstack{-0.7\\ \pm 2.7}} & {\tiny\Vectorstack{+0.9\\ \pm 2.0}} & {\tiny\Vectorstack{-1.6\\ \pm 2.3}} & {\tiny\Vectorstack{+5.9\\ \pm 3.2}} & {\tiny\Vectorstack{+4.4\\ \pm 2.5}} & {\tiny\Vectorstack{+6.5\\ \pm 4.2}} & {\tiny\Vectorstack{+3.0\\ \pm 4.0}} \\[2pt]
+\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-2.3\\ \pm 2.4}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-13\\ \pm 1.8}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-5.1\\ \pm 2.3}}} &  & {\tiny\Vectorstack{+3.9\\ \pm 4.1}} & {\tiny\Vectorstack{-5.8\\ \pm 3.6}} & {\tiny\Vectorstack{-4.3\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.7\\ \pm 2.5}}} & {\tiny\Vectorstack{+0.8\\ \pm 3.7}} & {\tiny\Vectorstack{-0.8\\ \pm 2.7}} & {\tiny\Vectorstack{+1.4\\ \pm 4.0}} & {\tiny\Vectorstack{-2.1\\ \pm 4.7}} \\[2pt]
+\textbf{Baseline} ($\mathrm{sum}$)&\cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.2\\ \pm 2.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-17\\ \pm 2.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-9.0\\ \pm 3.0}}} & {\tiny\Vectorstack{-3.9\\ \pm 4.1}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-9.8\\ \pm 2.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.2\\ \pm 2.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 2.6}}} & {\tiny\Vectorstack{-3.1\\ \pm 3.0}} & {\tiny\Vectorstack{-4.7\\ \pm 3.1}} & {\tiny\Vectorstack{-2.6\\ \pm 5.0}} & {\tiny\Vectorstack{-6.1\\ \pm 4.4}} \\[2pt]
+\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{+3.5\\ \pm 2.2}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-7.3\\ \pm 2.7}}} & {\tiny\Vectorstack{+0.7\\ \pm 2.7}} & {\tiny\Vectorstack{+5.8\\ \pm 3.6}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+9.8\\ \pm 2.4}}} &  & {\tiny\Vectorstack{+1.6\\ \pm 3.5}} & {\tiny\Vectorstack{-0.9\\ \pm 2.2}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.6\\ \pm 2.5}}} & {\tiny\Vectorstack{+5.1\\ \pm 3.2}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+7.2\\ \pm 3.3}}} & {\tiny\Vectorstack{+3.7\\ \pm 4.8}} \\[2pt]
+\textbf{2-GNN} ($\mean$)&{\tiny\Vectorstack{+1.9\\ \pm 2.9}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.9\\ \pm 2.3}}} & {\tiny\Vectorstack{-0.9\\ \pm 2.0}} & {\tiny\Vectorstack{+4.3\\ \pm 2.5}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.2\\ \pm 2.9}}} & {\tiny\Vectorstack{-1.6\\ \pm 3.5}} &  & {\tiny\Vectorstack{-2.4\\ \pm 2.6}} & {\tiny\Vectorstack{+5.1\\ \pm 3.5}} & {\tiny\Vectorstack{+3.5\\ \pm 2.2}} & {\tiny\Vectorstack{+5.6\\ \pm 4.8}} & {\tiny\Vectorstack{+2.1\\ \pm 4.7}} \\[2pt]
+\textbf{2-GNN} ($\mathrm{SAM}$)&{\tiny\Vectorstack{+4.4\\ \pm 2.3}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.5\\ \pm 2.0}}} & {\tiny\Vectorstack{+1.6\\ \pm 2.3}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.7\\ \pm 2.5}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 2.6}}} & {\tiny\Vectorstack{+0.9\\ \pm 2.2}} & {\tiny\Vectorstack{+2.4\\ \pm 2.6}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+7.5\\ \pm 3.1}}} & {\tiny\Vectorstack{+5.9\\ \pm 3.2}} & {\tiny\Vectorstack{+8.1\\ \pm 4.2}} & {\tiny\Vectorstack{+4.6\\ \pm 4.9}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)&{\tiny\Vectorstack{-3.1\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-14\\ \pm 2.8}}} & {\tiny\Vectorstack{-5.9\\ \pm 3.2}} & {\tiny\Vectorstack{-0.8\\ \pm 3.7}} & {\tiny\Vectorstack{+3.1\\ \pm 3.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.6\\ \pm 2.5}}} & {\tiny\Vectorstack{-5.1\\ \pm 3.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-7.5\\ \pm 3.1}}} &  & {\tiny\Vectorstack{-1.5\\ \pm 2.2}} & {\tiny\Vectorstack{+0.6\\ \pm 2.9}} & {\tiny\Vectorstack{-2.9\\ \pm 4.5}} \\[2pt]
+\textbf{2-WL-GNN\phantom{*}} ($\mean$)&{\tiny\Vectorstack{-1.6\\ \pm 2.6}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-12\\ \pm 2.5}}} & {\tiny\Vectorstack{-4.4\\ \pm 2.5}} & {\tiny\Vectorstack{+0.8\\ \pm 2.7}} & {\tiny\Vectorstack{+4.7\\ \pm 3.1}} & {\tiny\Vectorstack{-5.1\\ \pm 3.2}} & {\tiny\Vectorstack{-3.5\\ \pm 2.2}} & {\tiny\Vectorstack{-5.9\\ \pm 3.2}} & {\tiny\Vectorstack{+1.5\\ \pm 2.2}} &  & {\tiny\Vectorstack{+2.1\\ \pm 3.4}} & {\tiny\Vectorstack{-1.4\\ \pm 4.3}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{-3.7\\ \pm 3.6}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-15\\ \pm 4.2}}} & {\tiny\Vectorstack{-6.5\\ \pm 4.2}} & {\tiny\Vectorstack{-1.4\\ \pm 4.0}} & {\tiny\Vectorstack{+2.6\\ \pm 5.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-7.2\\ \pm 3.3}}} & {\tiny\Vectorstack{-5.6\\ \pm 4.8}} & {\tiny\Vectorstack{-8.1\\ \pm 4.2}} & {\tiny\Vectorstack{-0.6\\ \pm 2.9}} & {\tiny\Vectorstack{-2.1\\ \pm 3.4}} &  & {\tiny\Vectorstack{-3.5\\ \pm 6.2}} \\[2pt]
+\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{-0.2\\ \pm 3.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 3.4}}} & {\tiny\Vectorstack{-3.0\\ \pm 4.0}} & {\tiny\Vectorstack{+2.1\\ \pm 4.7}} & {\tiny\Vectorstack{+6.1\\ \pm 4.4}} & {\tiny\Vectorstack{-3.7\\ \pm 4.8}} & {\tiny\Vectorstack{-2.1\\ \pm 4.7}} & {\tiny\Vectorstack{-4.6\\ \pm 4.9}} & {\tiny\Vectorstack{+2.9\\ \pm 4.5}} & {\tiny\Vectorstack{+1.4\\ \pm 4.3}} & {\tiny\Vectorstack{+3.5\\ \pm 6.2}} &
 \end{tabular}}

### 2020-04-12 thesis/data/diffs/reddit.tex
completed evaluation chapter

@@
 & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-GWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{Baseline} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{GIN} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)} \\
-\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)& & {\tiny\Vectorstack{-1.8\\ \pm 1.8}} & {\tiny\Vectorstack{+0.5\\ \pm 4.9}} & {\tiny\Vectorstack{+0.8\\ \pm 3.6}} & {\tiny\Vectorstack{+4.2\\ \pm 10}} & {\tiny\Vectorstack{-11\\ \pm 6.9}} & {\tiny\Vectorstack{+2.9\\ \pm 3.5}} & {\tiny\Vectorstack{-7.4\\ \pm 6.2}} & {\tiny\Vectorstack{-0.9\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-13\\ \pm 3.6}}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=3$)&{\tiny\Vectorstack{+1.8\\ \pm 1.8}} &  & {\tiny\Vectorstack{+2.3\\ \pm 4.7}} & {\tiny\Vectorstack{+2.6\\ \pm 3.5}} & {\tiny\Vectorstack{+5.9\\ \pm 11}} & {\tiny\Vectorstack{-9.0\\ \pm 7.1}} & {\tiny\Vectorstack{+4.7\\ \pm 3.3}} & {\tiny\Vectorstack{-5.7\\ \pm 6.8}} & {\tiny\Vectorstack{+0.9\\ \pm 1.6}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 2.8}}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)&{\tiny\Vectorstack{-0.5\\ \pm 4.9}} & {\tiny\Vectorstack{-2.3\\ \pm 4.7}} &  & {\tiny\Vectorstack{+0.3\\ \pm 4.2}} & {\tiny\Vectorstack{+3.7\\ \pm 11}} & {\tiny\Vectorstack{-11\\ \pm 7.3}} & {\tiny\Vectorstack{+2.4\\ \pm 5.5}} & {\tiny\Vectorstack{-7.9\\ \pm 5.5}} & {\tiny\Vectorstack{-1.4\\ \pm 5.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-13\\ \pm 3.6}}} \\[2pt]
-\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-0.8\\ \pm 3.6}} & {\tiny\Vectorstack{-2.6\\ \pm 3.5}} & {\tiny\Vectorstack{-0.3\\ \pm 4.2}} &  & {\tiny\Vectorstack{+3.3\\ \pm 11}} & {\tiny\Vectorstack{-12\\ \pm 7.6}} & {\tiny\Vectorstack{+2.1\\ \pm 2.4}} & {\tiny\Vectorstack{-8.2\\ \pm 6.2}} & {\tiny\Vectorstack{-1.7\\ \pm 3.3}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-14\\ \pm 3.1}}} \\[2pt]
-\textbf{Baseline} ($\mathrm{sum}$)&{\tiny\Vectorstack{-4.2\\ \pm 10}} & {\tiny\Vectorstack{-5.9\\ \pm 11}} & {\tiny\Vectorstack{-3.7\\ \pm 11}} & {\tiny\Vectorstack{-3.3\\ \pm 11}} &  & {\tiny\Vectorstack{-15\\ \pm 12}} & {\tiny\Vectorstack{-1.2\\ \pm 12}} & {\tiny\Vectorstack{-12\\ \pm 13}} & {\tiny\Vectorstack{-5.0\\ \pm 11}} & {\tiny\Vectorstack{-17\\ \pm 11}} \\[2pt]
-\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{+11\\ \pm 6.9}} & {\tiny\Vectorstack{+9.0\\ \pm 7.1}} & {\tiny\Vectorstack{+11\\ \pm 7.3}} & {\tiny\Vectorstack{+12\\ \pm 7.6}} & {\tiny\Vectorstack{+15\\ \pm 12}} &  & {\tiny\Vectorstack{+14\\ \pm 8.3}} & {\tiny\Vectorstack{+3.3\\ \pm 9.6}} & {\tiny\Vectorstack{+9.9\\ \pm 7.4}} & {\tiny\Vectorstack{-2.2\\ \pm 7.2}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)&{\tiny\Vectorstack{-2.9\\ \pm 3.5}} & {\tiny\Vectorstack{-4.7\\ \pm 3.3}} & {\tiny\Vectorstack{-2.4\\ \pm 5.5}} & {\tiny\Vectorstack{-2.1\\ \pm 2.4}} & {\tiny\Vectorstack{+1.2\\ \pm 12}} & {\tiny\Vectorstack{-14\\ \pm 8.3}} &  & {\tiny\Vectorstack{-10\\ \pm 7.1}} & {\tiny\Vectorstack{-3.8\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-16\\ \pm 3.3}}} \\[2pt]
-\textbf{2-WL-GNN\phantom{*}} ($\mean$)&{\tiny\Vectorstack{+7.4\\ \pm 6.2}} & {\tiny\Vectorstack{+5.7\\ \pm 6.8}} & {\tiny\Vectorstack{+7.9\\ \pm 5.5}} & {\tiny\Vectorstack{+8.2\\ \pm 6.2}} & {\tiny\Vectorstack{+12\\ \pm 13}} & {\tiny\Vectorstack{-3.3\\ \pm 9.6}} & {\tiny\Vectorstack{+10\\ \pm 7.1}} &  & {\tiny\Vectorstack{+6.6\\ \pm 7.2}} & {\tiny\Vectorstack{-5.5\\ \pm 6.6}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{+0.9\\ \pm 2.5}} & {\tiny\Vectorstack{-0.9\\ \pm 1.6}} & {\tiny\Vectorstack{+1.4\\ \pm 5.0}} & {\tiny\Vectorstack{+1.7\\ \pm 3.3}} & {\tiny\Vectorstack{+5.0\\ \pm 11}} & {\tiny\Vectorstack{-9.9\\ \pm 7.4}} & {\tiny\Vectorstack{+3.8\\ \pm 2.5}} & {\tiny\Vectorstack{-6.6\\ \pm 7.2}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-12\\ \pm 2.6}}} \\[2pt]
-\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+13\\ \pm 3.6}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 2.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+13\\ \pm 3.6}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+14\\ \pm 3.1}}} & {\tiny\Vectorstack{+17\\ \pm 11}} & {\tiny\Vectorstack{+2.2\\ \pm 7.2}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+16\\ \pm 3.3}}} & {\tiny\Vectorstack{+5.5\\ \pm 6.6}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+12\\ \pm 2.6}}} &
+\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)& & {\tiny\Vectorstack{-1.8\\ \pm 1.8}} & {\tiny\Vectorstack{+0.5\\ \pm 4.9}} & {\tiny\Vectorstack{+0.8\\ \pm 3.6}} & {\tiny\Vectorstack{+4.2\\ \pm 10}} & {\tiny\Vectorstack{-11\\ \pm 6.9}} & {\tiny\Vectorstack{+2.9\\ \pm 3.5}} & {\tiny\Vectorstack{-7.4\\ \pm 6.2}} & {\tiny\Vectorstack{-0.9\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-13\\ \pm 3.3}}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=3$)&{\tiny\Vectorstack{+1.8\\ \pm 1.8}} &  & {\tiny\Vectorstack{+2.3\\ \pm 4.7}} & {\tiny\Vectorstack{+2.6\\ \pm 3.5}} & {\tiny\Vectorstack{+5.9\\ \pm 11}} & {\tiny\Vectorstack{-9.0\\ \pm 7.1}} & {\tiny\Vectorstack{+4.7\\ \pm 3.3}} & {\tiny\Vectorstack{-5.7\\ \pm 6.8}} & {\tiny\Vectorstack{+0.9\\ \pm 1.6}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 2.6}}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)&{\tiny\Vectorstack{-0.5\\ \pm 4.9}} & {\tiny\Vectorstack{-2.3\\ \pm 4.7}} &  & {\tiny\Vectorstack{+0.3\\ \pm 4.2}} & {\tiny\Vectorstack{+3.7\\ \pm 11}} & {\tiny\Vectorstack{-11\\ \pm 7.3}} & {\tiny\Vectorstack{+2.4\\ \pm 5.5}} & {\tiny\Vectorstack{-7.9\\ \pm 5.5}} & {\tiny\Vectorstack{-1.4\\ \pm 5.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-13\\ \pm 3.3}}} \\[2pt]
+\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-0.8\\ \pm 3.6}} & {\tiny\Vectorstack{-2.6\\ \pm 3.5}} & {\tiny\Vectorstack{-0.3\\ \pm 4.2}} &  & {\tiny\Vectorstack{+3.3\\ \pm 11}} & {\tiny\Vectorstack{-12\\ \pm 7.6}} & {\tiny\Vectorstack{+2.1\\ \pm 2.4}} & {\tiny\Vectorstack{-8.2\\ \pm 6.2}} & {\tiny\Vectorstack{-1.7\\ \pm 3.3}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-14\\ \pm 2.5}}} \\[2pt]
+\textbf{Baseline} ($\mathrm{sum}$)&{\tiny\Vectorstack{-4.2\\ \pm 10}} & {\tiny\Vectorstack{-5.9\\ \pm 11}} & {\tiny\Vectorstack{-3.7\\ \pm 11}} & {\tiny\Vectorstack{-3.3\\ \pm 11}} &  & {\tiny\Vectorstack{-15\\ \pm 12}} & {\tiny\Vectorstack{-1.2\\ \pm 12}} & {\tiny\Vectorstack{-12\\ \pm 13}} & {\tiny\Vectorstack{-5.0\\ \pm 11}} & {\tiny\Vectorstack{-17\\ \pm 10}} \\[2pt]
+\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{+11\\ \pm 6.9}} & {\tiny\Vectorstack{+9.0\\ \pm 7.1}} & {\tiny\Vectorstack{+11\\ \pm 7.3}} & {\tiny\Vectorstack{+12\\ \pm 7.6}} & {\tiny\Vectorstack{+15\\ \pm 12}} &  & {\tiny\Vectorstack{+14\\ \pm 8.3}} & {\tiny\Vectorstack{+3.3\\ \pm 9.6}} & {\tiny\Vectorstack{+9.9\\ \pm 7.4}} & {\tiny\Vectorstack{-2.2\\ \pm 6.8}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)&{\tiny\Vectorstack{-2.9\\ \pm 3.5}} & {\tiny\Vectorstack{-4.7\\ \pm 3.3}} & {\tiny\Vectorstack{-2.4\\ \pm 5.5}} & {\tiny\Vectorstack{-2.1\\ \pm 2.4}} & {\tiny\Vectorstack{+1.2\\ \pm 12}} & {\tiny\Vectorstack{-14\\ \pm 8.3}} &  & {\tiny\Vectorstack{-10\\ \pm 7.1}} & {\tiny\Vectorstack{-3.8\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-16\\ \pm 2.8}}} \\[2pt]
+\textbf{2-WL-GNN\phantom{*}} ($\mean$)&{\tiny\Vectorstack{+7.4\\ \pm 6.2}} & {\tiny\Vectorstack{+5.7\\ \pm 6.8}} & {\tiny\Vectorstack{+7.9\\ \pm 5.5}} & {\tiny\Vectorstack{+8.2\\ \pm 6.2}} & {\tiny\Vectorstack{+12\\ \pm 13}} & {\tiny\Vectorstack{-3.3\\ \pm 9.6}} & {\tiny\Vectorstack{+10\\ \pm 7.1}} &  & {\tiny\Vectorstack{+6.6\\ \pm 7.2}} & {\tiny\Vectorstack{-5.5\\ \pm 6.1}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{+0.9\\ \pm 2.5}} & {\tiny\Vectorstack{-0.9\\ \pm 1.6}} & {\tiny\Vectorstack{+1.4\\ \pm 5.0}} & {\tiny\Vectorstack{+1.7\\ \pm 3.3}} & {\tiny\Vectorstack{+5.0\\ \pm 11}} & {\tiny\Vectorstack{-9.9\\ \pm 7.4}} & {\tiny\Vectorstack{+3.8\\ \pm 2.5}} & {\tiny\Vectorstack{-6.6\\ \pm 7.2}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-12\\ \pm 2.7}}} \\[2pt]
+\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+13\\ \pm 3.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 2.6}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+13\\ \pm 3.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+14\\ \pm 2.5}}} & {\tiny\Vectorstack{+17\\ \pm 10}} & {\tiny\Vectorstack{+2.2\\ \pm 6.8}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+16\\ \pm 2.8}}} & {\tiny\Vectorstack{+5.5\\ \pm 6.1}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+12\\ \pm 2.7}}} &
 \end{tabular}}

### 2020-04-12 implementation/src/ltag/playground.py

@@ -302,8 +302,13 @@ def ds_stats():
 # synthetic.balanced_triangle_classification_dataset(stored=True)().draw(153, label_colors=True)
 # synthetic.balanced_triangle_classification_dataset(stored=True)().export_dot(152)
 # synthetic.balanced_triangle_classification_dataset(stored=True)().export_dot(153)
-ds_stats()
-
+# ds_stats()
+# p = eval_ds.Proteins_5()
+# d = eval_ds.DD_2()
+# p.draw(402)
+# d.draw(2)
+# p.export_dot(402)
+# d.export_dot(2)
 #
 # list(dsm.get_all(output_type="grakel")[0])
 #

### 2020-04-13 thesis/content/abstract.tex
started with ltag definition + various minor fixes
improved intro and conclusion

@@
-% !TEX root = ../main.tex
+%!TEX root = ../main.tex
 %
@@
 \vspace*{-10mm}
+
+This thesis describes the research field of \acl{gcr} from the perspective of the \ac{lta} problem.
+It formally characterizes a selection of state-of-the-art \textit{graph kernels} and \acp{gnn} as instances of \ac{lta}.
+Those characterizations are shown to be limited by the way in which they decompose graphs.
+To overcome this limitation, an avenue for a more ``\acs{lta}-like'' \ac{gnn} is provided in form of so-called \textit{learned edge filters}.
+To realize edge filters, the novel \textit{2-\acs*{wl}-\acs*{gnn}} model is proposed;
+it is inspired by the two-dimensional \ac{wl} algorithm and proven to be strictly more expressive than existing \ac{gnn} approaches which are bounded by the more restrictive one-dimensional \ac{wl} algorithm.

### 2020-04-13 thesis/content/chapter-conclusion.tex
started with ltag definition + various minor fixes
improved intro and conclusion

@@
 
+To conclude the thesis, we now look back on the three research questions described in \cref{sec:intro:questions} and summarize the answers we gave to them in the previous chapters.
+Afterwards, a brief overview of future research directions based on our findings will be given.
+
 \section{Review}%
@@
 
+\paragraph{\circled{1}\; What constitutes an \ac{lta} method?}
+We began with a general definition of \ac{lta} in \cref{sec:ltag:definition}.
+There we proposed that its defining characteristic should be the \textit{localized explainability} of its predictions.
+This characteristic was formalized via the notion of \textit{\ac{lta} formulations} (see \cfullref{defn:ltag:formulation}) which requires that a model is expressible in terms of a decomposition function $\psi: \mathcal{G} \to \mathcal{P}(\mathcal{G})$, a local evaluation function $f: \mathcal{G} \to \mathcal{Y} \times \mathcal{R}$ and a weighted aggregation function $\mathcal{A}: {(\mathcal{Y} \times \mathcal{R}_{\geq 0})}^* \to \mathcal{Y}$.
+An ideal \ac{lta} method has such a formulation with a decomposition function $\psi$ that splits graphs into ``meaningful'' constituents in some domain-specific sense of the word.
+Since this ideal notion of \ac{lta} is generally quite fuzzy, we only distinguished between \acs{lta}-like and non-\acs{lta} methods in this thesis;
+a method was called non-\acs{lta} if it uses a trivial decomposition function that just splits a graph $G$ into the single ``constituent'' $G$.
+
+\paragraph{\circled{2}\; How do existing \ac{gcr} methods relate to \ac{lta}?}
+In \cref{sec:ltag:formulation:svm,sec:ltag:formulation:gcnn} we used our definition of \ac{lta} to check which of the existing \ac{gcr} approaches are compatible with it.
+For the case of an \ac{svm} using a graph kernel/embedding we found that it is an \acs{lta}-like method if the kernel is a so-called nontrivial \acf{sce} (see \cfullref{defn:ltag:substruct-embedding,thm:ltag:svm-ltag-formulation}).
+This \ac{sce} condition is satisfied by fingerprint embeddings, the \ac{wl} subtree kernel and the 2-LWL kernel which makes them \acs{lta}-like.
+However, \texttt{graph2vec} embeddings, the \ac{wl} shortest-path kernel and the 2-GWL kernel were found to be trivial or only partly nontrivial \acp{sce}, i.e.\ they are non-\acs{lta} methods.
+After considering those embedding approaches we looked at \ac{gcnn} and showed that they also have an \ac{lta} formulation under certain conditions (see \cfullref{thm:ltag:gcnn-ltag-formulation}).
+More specifically, we saw that the constituents used by a \ac{gcnn} are spanned by the \ac{bfs} subtrees of its input graph.
+
+\paragraph{\circled{3}\; What are limitations of existing graph \ac{lta} methods and how can they be overcome?}
+\Cref{sec:ltd:edge-filter} described that the subtree constituents are their primary limitation of \acp{gcnn} and that more flexible decompositions can be learned via an edge filtering strategy.
+To realize edge filtering, we proposed that informative edge feature vectors could be used as the input to a filtering classifier.
+To produce such feature vectors we first looked at 2-\acsp{gnn} and found that they have various theoretical limitations, i.e.\ the inability to distinguish regular graphs and to detect cycles (see \cfullref{prop:ltd:2gnn-regular-limit,prop:ltd:2gnn-cycle-limit}).
+We therefore proposed the 2-\acs{wl}-\acs{gnn} which does not have those limitations (see \cfullref{cor:ltd:wl2-gnn-regular}).
+
+\paragraph{Evaluation results}
+For the evaluation of our results we considered two aspects:
+Firstly, we looked at how 2-\acs{wl}-\acsp{gnn} compare to other \acp{gnn}.
+Secondly, we evaluated how \acs{lta}-like methods compare to non-\acs{lta} methods.
+Regarding the first aspect, we showed that the theoretical advantages of 2-\acs{wl}-\acsp{gnn} are clearly observable on the synthetic triangle detection dataset while on the evaluated real-world datasets we got results which are generally comparable with the best state-of-the-art approaches but not significantly better.
+Regarding the second aspect, we observed no general advantage or disadvantage of \acs{lta}-like methods.
+While the \acs{lta}-like configurations of 2-\acs{wl}-\acsp{gnn} generally performed worse than their non-\ac{lta} counterparts, the \acs{lta}-like \ac{wl} subtree kernel generally performed quite well.
+This shows that \ac{lta} is in principle suitable for graph classification tasks if the right decomposition, evaluation and aggregation functions are chosen.
+
 \section{Future Directions}%

### 2020-04-13 thesis/content/chapter-evaluation.tex
improved intro and conclusion

@@
 Motivated by the idea of dynamically learning decompositions via edge filters, we then proposed the novel 2-\acs{wl}-\acs{gnn} in \cref{sec:ltd}.
-The ides presented in both chapters will now be empirically evaluated.
+The ideas presented in both chapters will now be empirically evaluated.
 Do to so we differentiate between two mostly independent evaluation aspects:
@@
 	\item Then evaluation results on real-world datasets are described in \cref{sec:eval:real}.
-		There we will see how 2-\acs{wl}-\acsp{gnn} compare to other \acp{gnn} in practice as well as how \acs{lta}-like models compare to non-\acs{lta} model.
+		There we will see how 2-\acs{wl}-\acsp{gnn} compare to other \acp{gnn} in practice as well as how \acs{lta}-like models compare to non-\acs{lta} models.
 	\item Finally, \cref{sec:eval:lta} looks at the synthetic and real-world results from an \ac{lta} perspective.
@@
 	the resource use for all evaluations was limited to \SI{64}{\giga\byte} main memory and \SI{11}{\giga\byte} GPU memory.
-	On average, each \ac{gnn} assessment (see \cref{algo:eval:assessment}) took about three days on an AMD 1800X 16-thread CPU and an Nvidia 1080~Ti GPU.% chktex 29
+	On average, each \ac{gnn} assessment (as in \cref{algo:eval:assessment}) took about three days on an AMD 1800X 16-thread CPU and an Nvidia 1080~Ti GPU.% chktex 29
 }.

### 2020-04-13 thesis/content/chapter-introduction.tex
started with ltag definition + various minor fixes
continued with spatial gcnn lta relation
reorganized thesis: wl2gnn gets its own chapter
includes the intro for the wl2gcnn chapter
+ various improvements
improved intro and conclusion

@@
 
-\section{Motivation}%
-\label{sec:intro:motivation}
-
-The field of \ac{ml} on graph-structured data has applications in many domains due to the general expressive power of graphs.
-The three most common types of graph~\ac{ml} problems are
+The field of \ac{ml} on graph-structured data has recently become an active topic of research.
+One reason for this is the wide range of domains and problems that are expressible in terms of graphs.
+Three very common types of such graph~\ac{ml} problems~\cite{Wu2019} are:
 \begin{enumerate}[label=\textbf{\arabic*.}]
@@
 		Here a class or a score has to be predicted for each vertex of a graph.
-		In social graphs this corresponds to the prediction of properties of individuals, e.g.\ personal preferences or gender.
+		In social graphs this corresponds to the prediction of properties of individuals, e.g.\ personal preferences.
 		Another example is the prediction of the amount of traffic at the intersections of a street network.
 	\item \textbf{Graph classification \& regression:}
-		In this final problem type a single global class or continuous value has to be predicted for an input graph.
-		The canonical example for this is the prediction of properties of molecule graphs, e.g.\ the toxicity or solubility of a chemical.
+		In this final problem type, a single global class or continuous value has to be predicted for an input graph.
+		The canonical example for this is the prediction of properties of molecular graphs, e.g.\ the toxicity or solubility of a chemical.
 \end{enumerate}
-In this thesis we will focus on the last problem type, \acf{gcr}.
-An \ac{ml} method for this problem has to accept graphs of varying size and should be permutation invariant wrt.\ the order in which graph vertices are provided.
-Those requirements are not met by the commonly used learners that only accept fixed-size feature vectors as their input, e.g.\ \acp{lm}, \acp{svm} or \acp{mlp}.
+In this thesis we will focus specifically on the last problem type, \ac{gcr}.
+When comparing this problem with the more common problem of classification/regression on fixed-dimensional vector inputs $x \in \mathbb{R}^d$ we find two fundamental differences:
+\begin{enumerate*}[label={\circled{\small\arabic*}}]
+	\item The sizes of input graphs are generally not fixed; in \ac{gcr}, inputs can be arbitrarily small or large.
+	\item Moreover, while two distinct vectors $x \neq x'$ typically represent distinct inputs for which different predictions $y \neq y'$ can be produced, graphs generally cannot be uniquely encoded because the order in which vertices are provided does not matter.
+		A \ac{gcr} method must therefore be invariant w.r.t.\ vertex permutations, i.e.\ it must map two structurally identical graphs $G \simeq G'$ to the same prediction even if their encoding differs.
+\end{enumerate*}
 
-A \ac{gcr} method has to account for two central aspects of the problem:
-\begin{enumerate*}
-	\item Local structural analysis and
-	\item global aggregation
-\end{enumerate*}.
-The first aspect is about the extraction of relevant features of substructures of the input graph.
-The latter is about the way in which the local features are combined into a final class or regression value.
-The existing \ac{gcr} methods are mostly motivated by local structural graph analysis.
-The aspect of global aggregation on the other hand is less emphasized by those methods.
+Due to those two differences, the \ac{gcr} problem cannot be directly solved by any of the common learners for vectorial inputs, e.g.\ \acp{lm} or \acp{mlp}.
+Therefore, to tackle this problem, various graph-specific learners have been proposed over the recent years.
+This thesis provides a novel perspective on the \ac{gcr} problem by combining the existing approaches with another, currently unrelated, field of research called \ac{lta}~\cite{Melnikov2016}\cite{Melnikov2019}.
 
-There is however a separate branch of research that specifically looks at the problem of learning aggregation functions, called \ac{lta}.
-Current \ac{lta} approaches explicitly learn an aggregation functions for sets which can be interpreted as graphs without edges.
-The motivation for this thesis is to generalize \ac{lta} from sets to arbitrary graphs.
-The overall goal is to combine the aggregation learning perspective with existing \ac{gcr} methods.
+\section{Motivation}%
+\label{sec:intro:motivation}
 
-\section{Goals}%
-\label{sec:intro:goals}
+The existing \ac{gcr} approaches can be split into the family of so-called \textit{graph embeddings} and that of \acp{gnn}.
+The embedding approach maps graphs to vectors in order to then apply a standard vector classification or regression method.
+\Acp{gnn} on the other hand directly produce a prediction by using shared weights, similar to \acp*{cnn} on images~\cite{LeCun1998}.
 
-To extend \ac{lta} to graphs, three goals have to be achieved:
+Independent of those approaches, the family of \ac{lta} methods looks at the problem of aggregating variable-size \textit{compositions}, i.e.\ multisets, of so-called \textit{constituents}.
+Like \ac{gcr} approaches, an \ac{lta} method has to be able to process inputs of varying size in a permutation invariant fashion.
+Motivated by this similarity, this thesis extends the ideas of \ac{lta} to the domain of structured graph data.
+
+Such an extension is interesting because \ac{lta} models produce composition predictions which are the direct result of local constituent predictions.
+This means that \ac{lta} models provide insight into the influence of constituents on the final prediction, which makes \ac{lta} relevant from the perspective \ac{xai}~\cite{Gilpin2018}.
+\Cref{fig:intro:intuition} illustrates how a graph-variant of \ac{lta} could explain the prediction of some graph property.
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=0.8\linewidth]{gfx/introduction/intuition.pdf}
+	\caption{Intuition for how \ac{lta} could describe some graph property, e.g.\ the toxicity of a molecule, via a set of local constituent predictions.}\label{fig:intro:intuition}
+\end{figure}
+
+\section{Research Questions}%
+\label{sec:intro:questions}
+
+Based on the idea of extending \ac{lta} to graphs, we state the following three research questions which will be answered in this thesis:
 \begin{enumerate}[label=\textbf{\arabic*.}]
 	\item \textbf{Formalization of \ac{lta}:}
-		Before \ac{lta} can be extended, its essential characteristics have to be defined.
-		Those characteristics should provide the terminology to formally capture the differences and similarities between \ac{lta} and existing \ac{gcr} methods.
-	\item \textbf{Give an \ac{lta} interpretation of \ac{gcr} methods:}
-		Using the \ac{lta} formalization, representative \ac{gcr} approaches should be restated as \ac{lta} instances.
-		Currently there is no comprehensive formulation of the relation between both fields of research;
-		this is addressed by the the second goal.
-	\item \textbf{Define an \ac{lta} method for graphs:}
-		Using the \ac{lta} perspective on \ac{gcr}, hidden assumptions of the existing approaches should become clear and in which way they share the assumptions of \ac{lta}.
-		The last goal is to use those insights to formulate an \ac{lta}-\ac{gcr} method that combines ideas from the existing approaches with the \ac{lta} assumptions.
+		The previous work on \ac{lta} provides only a narrow definition of \ac{lta} for unstructured multiset inputs.
+		The question what its essential characteristics are, on a more general level, has not yet been considered.
+		Therefore we will provide the terminology and definitions to formally capture what constitutes an \ac{lta} method as opposed to a non-\ac{lta} method.
+	\item \textbf{An \ac{lta} interpretation of existing \ac{gcr} methods:}
+		Using the generalized \ac{lta} formalization, a selection of state-of-the-art \ac{gcr} approaches will be checked for their compatibility with \ac{lta}.
+		By doing so, we answer the question whether and how existing \ac{gcr} methods are able to produce predictions which are explainable by local constituent predictions.
+	\item \textbf{Definition of a novel \acs{lta}-inspired \ac{gcr} method:}
+		As a follow-up question to the previous one, we ask to which extent the existing \ac{gcr} methods are explainable and what their specific shortcomings are.
+		Based on those shortcomings we will propose a novel \acs{lta}-inspired approach to overcome them.
 \end{enumerate}
@@
 \paragraph{\Cref{sec:related}: \nameref{sec:related}}
+In order to answer our three research questions, we begin by with on overview of the previous work regarding \ac{lta} as well as \ac{gcr}.
+We begin with a brief description of \ac{lta}.
+Then the theoretical foundations of the existing \ac{gcr} approaches are introduced;
+one particularly important concept there will be the so-called \acl*{wl} coloring.
+Based on the theoretical foundations, we conclude the chapter by describing common graph embedding and \acl{gnn} methods.
 
 \paragraph{\Cref{sec:ltag}: \nameref{sec:ltag}}
+In this chapter we answer the first two of our research questions.
+In the first step a formal definition of \ac{lta} is provided, which answers the question of what constitutes an \ac{lta} method.
+This definition is then used to analyze all \ac{gcr} approaches described in \cref{sec:related} for their compatibility with \ac{lta}.
+
+\paragraph{\Cref{sec:ltd}: \nameref{sec:ltd}}
+Based on the \ac{lta} perspective on \ac{gcr} provided in \cref{sec:ltag}, we find that the existing approaches are all limited by their underlying solution to the so-called \ac*{ltd} problem.
+To overcome this limitation, and thereby answer our third research question, we propose the general idea of edge-filtered graph convolutions.
+To realize this general idea in a specific method, we then proceed by proposing a novel type of \ac*{gcnn}, the so-called \textit{2-\acs*{wl}-\acs{gnn}}, which can serve as the foundation for an edge-filtering-based solution to the \acs*{ltd} problem.
+Apart from its uses in the context of \ac{lta}/\acs*{ltd}, we show that 2-\acs*{wl}-\acsp{gnn} have additional theoretical advantages over existing approaches;
+this makes them interesting even from a non-\acs{lta} perspective.
 
 \paragraph{\Cref{sec:eval}: \nameref{sec:eval}}
+In this chapter we empirically evaluate how \ac{gcr} approaches that are compatible with our definition of \ac{lta} perform compared to non-\acs{lta} \ac{gcr} approaches.
+Additionally, we compare the proposed 2-\acs*{wl}-\acs{gnn} with other state-of-the-art \acp{gnn} to verify whether its theoretical advantages over previous approaches can be observed in practice.
 
 \paragraph{\Cref{sec:conclusion}: \nameref{sec:conclusion}}
+Finally, the results of this thesis are summarized and a brief outline of promising directions for future research is given.

### 2020-04-13 thesis/content/chapter-related-work.tex
improved intro and conclusion

@@
 
-Before combining \ac{lta} and \ac{gcr} as described in \cref{sec:intro:goals}, we first give an overview of the state-of-the-art in both fields of research.
+Before combining \ac{lta} and \ac{gcr}, as described in \cref{sec:intro:questions}, we first give an overview of the state-of-the-art in both fields of research.
 This is done in three steps:
@@
 		To solve the \ac{gcr} problem, relevant graphs characteristics have to be defined in order to determine the similarity and dissimilarity of graphs.
-		We will look at three approaches for graph characterization: The graph isomorphism test, the so-called \acl{wl} coloring and lastly the notion of graph spectra.W
+		We will look at three approaches for graph characterization: The strict graph isomorphism test, the so-called \acl{wl} coloring and lastly the notion of graph spectra.
 	\item Based on the described graph characterization approaches, an overview of current \ac{gcr} methods will then be given.
@@
 
-The last family of \ac{gcr} approaches we will look at is that of \acp{gnn}.
+The last family of \ac{gcr} approaches we will look at is that of \acfp{gnn}.
 The idea to feed a graph into a \ac{nn} was first described by \citet{Gori2005}.

### 2020-04-13 thesis/data/diffs/nci.tex
improved intro and conclusion

@@
 & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-GWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{Baseline} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{GIN} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)} \\
-\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)& & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 1.3}}} & {\tiny\Vectorstack{-2.8\\ \pm 1.5}} & {\tiny\Vectorstack{+2.3\\ \pm 2.4}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.2\\ \pm 2.9}}} & {\tiny\Vectorstack{-3.5\\ \pm 2.2}} & {\tiny\Vectorstack{-1.9\\ \pm 2.9}} & {\tiny\Vectorstack{-4.4\\ \pm 2.3}} & {\tiny\Vectorstack{+3.1\\ \pm 2.5}} & {\tiny\Vectorstack{+1.6\\ \pm 2.6}} & {\tiny\Vectorstack{+3.7\\ \pm 3.6}} & {\tiny\Vectorstack{+0.2\\ \pm 3.5}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=3$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 1.3}}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.1\\ \pm 1.6}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+13\\ \pm 1.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+17\\ \pm 2.9}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+7.3\\ \pm 2.7}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.9\\ \pm 2.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.5\\ \pm 2.0}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+14\\ \pm 2.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+12\\ \pm 2.5}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+15\\ \pm 4.2}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 3.4}}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)&{\tiny\Vectorstack{+2.8\\ \pm 1.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.1\\ \pm 1.6}}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+5.1\\ \pm 2.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+9.0\\ \pm 3.0}}} & {\tiny\Vectorstack{-0.7\\ \pm 2.7}} & {\tiny\Vectorstack{+0.9\\ \pm 2.0}} & {\tiny\Vectorstack{-1.6\\ \pm 2.3}} & {\tiny\Vectorstack{+5.9\\ \pm 3.2}} & {\tiny\Vectorstack{+4.4\\ \pm 2.5}} & {\tiny\Vectorstack{+6.5\\ \pm 4.2}} & {\tiny\Vectorstack{+3.0\\ \pm 4.0}} \\[2pt]
-\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-2.3\\ \pm 2.4}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-13\\ \pm 1.8}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-5.1\\ \pm 2.3}}} &  & {\tiny\Vectorstack{+3.9\\ \pm 4.1}} & {\tiny\Vectorstack{-5.8\\ \pm 3.6}} & {\tiny\Vectorstack{-4.3\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.7\\ \pm 2.5}}} & {\tiny\Vectorstack{+0.8\\ \pm 3.7}} & {\tiny\Vectorstack{-0.8\\ \pm 2.7}} & {\tiny\Vectorstack{+1.4\\ \pm 4.0}} & {\tiny\Vectorstack{-2.1\\ \pm 4.7}} \\[2pt]
-\textbf{Baseline} ($\mathrm{sum}$)&\cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.2\\ \pm 2.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-17\\ \pm 2.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-9.0\\ \pm 3.0}}} & {\tiny\Vectorstack{-3.9\\ \pm 4.1}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-9.8\\ \pm 2.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.2\\ \pm 2.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 2.6}}} & {\tiny\Vectorstack{-3.1\\ \pm 3.0}} & {\tiny\Vectorstack{-4.7\\ \pm 3.1}} & {\tiny\Vectorstack{-2.6\\ \pm 5.0}} & {\tiny\Vectorstack{-6.1\\ \pm 4.4}} \\[2pt]
-\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{+3.5\\ \pm 2.2}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-7.3\\ \pm 2.7}}} & {\tiny\Vectorstack{+0.7\\ \pm 2.7}} & {\tiny\Vectorstack{+5.8\\ \pm 3.6}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+9.8\\ \pm 2.4}}} &  & {\tiny\Vectorstack{+1.6\\ \pm 3.5}} & {\tiny\Vectorstack{-0.9\\ \pm 2.2}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.6\\ \pm 2.5}}} & {\tiny\Vectorstack{+5.1\\ \pm 3.2}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+7.2\\ \pm 3.3}}} & {\tiny\Vectorstack{+3.7\\ \pm 4.8}} \\[2pt]
-\textbf{2-GNN} ($\mean$)&{\tiny\Vectorstack{+1.9\\ \pm 2.9}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.9\\ \pm 2.3}}} & {\tiny\Vectorstack{-0.9\\ \pm 2.0}} & {\tiny\Vectorstack{+4.3\\ \pm 2.5}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.2\\ \pm 2.9}}} & {\tiny\Vectorstack{-1.6\\ \pm 3.5}} &  & {\tiny\Vectorstack{-2.4\\ \pm 2.6}} & {\tiny\Vectorstack{+5.1\\ \pm 3.5}} & {\tiny\Vectorstack{+3.5\\ \pm 2.2}} & {\tiny\Vectorstack{+5.6\\ \pm 4.8}} & {\tiny\Vectorstack{+2.1\\ \pm 4.7}} \\[2pt]
-\textbf{2-GNN} ($\mathrm{SAM}$)&{\tiny\Vectorstack{+4.4\\ \pm 2.3}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.5\\ \pm 2.0}}} & {\tiny\Vectorstack{+1.6\\ \pm 2.3}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.7\\ \pm 2.5}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 2.6}}} & {\tiny\Vectorstack{+0.9\\ \pm 2.2}} & {\tiny\Vectorstack{+2.4\\ \pm 2.6}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+7.5\\ \pm 3.1}}} & {\tiny\Vectorstack{+5.9\\ \pm 3.2}} & {\tiny\Vectorstack{+8.1\\ \pm 4.2}} & {\tiny\Vectorstack{+4.6\\ \pm 4.9}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)&{\tiny\Vectorstack{-3.1\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-14\\ \pm 2.8}}} & {\tiny\Vectorstack{-5.9\\ \pm 3.2}} & {\tiny\Vectorstack{-0.8\\ \pm 3.7}} & {\tiny\Vectorstack{+3.1\\ \pm 3.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.6\\ \pm 2.5}}} & {\tiny\Vectorstack{-5.1\\ \pm 3.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-7.5\\ \pm 3.1}}} &  & {\tiny\Vectorstack{-1.5\\ \pm 2.2}} & {\tiny\Vectorstack{+0.6\\ \pm 2.9}} & {\tiny\Vectorstack{-2.9\\ \pm 4.5}} \\[2pt]
-\textbf{2-WL-GNN\phantom{*}} ($\mean$)&{\tiny\Vectorstack{-1.6\\ \pm 2.6}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-12\\ \pm 2.5}}} & {\tiny\Vectorstack{-4.4\\ \pm 2.5}} & {\tiny\Vectorstack{+0.8\\ \pm 2.7}} & {\tiny\Vectorstack{+4.7\\ \pm 3.1}} & {\tiny\Vectorstack{-5.1\\ \pm 3.2}} & {\tiny\Vectorstack{-3.5\\ \pm 2.2}} & {\tiny\Vectorstack{-5.9\\ \pm 3.2}} & {\tiny\Vectorstack{+1.5\\ \pm 2.2}} &  & {\tiny\Vectorstack{+2.1\\ \pm 3.4}} & {\tiny\Vectorstack{-1.4\\ \pm 4.3}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{-3.7\\ \pm 3.6}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-15\\ \pm 4.2}}} & {\tiny\Vectorstack{-6.5\\ \pm 4.2}} & {\tiny\Vectorstack{-1.4\\ \pm 4.0}} & {\tiny\Vectorstack{+2.6\\ \pm 5.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-7.2\\ \pm 3.3}}} & {\tiny\Vectorstack{-5.6\\ \pm 4.8}} & {\tiny\Vectorstack{-8.1\\ \pm 4.2}} & {\tiny\Vectorstack{-0.6\\ \pm 2.9}} & {\tiny\Vectorstack{-2.1\\ \pm 3.4}} &  & {\tiny\Vectorstack{-3.5\\ \pm 6.2}} \\[2pt]
-\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{-0.2\\ \pm 3.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 3.4}}} & {\tiny\Vectorstack{-3.0\\ \pm 4.0}} & {\tiny\Vectorstack{+2.1\\ \pm 4.7}} & {\tiny\Vectorstack{+6.1\\ \pm 4.4}} & {\tiny\Vectorstack{-3.7\\ \pm 4.8}} & {\tiny\Vectorstack{-2.1\\ \pm 4.7}} & {\tiny\Vectorstack{-4.6\\ \pm 4.9}} & {\tiny\Vectorstack{+2.9\\ \pm 4.5}} & {\tiny\Vectorstack{+1.4\\ \pm 4.3}} & {\tiny\Vectorstack{+3.5\\ \pm 6.2}} &
+\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)& & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 1.3}}} & {\tiny\Vectorstack{-2.8\\ \pm 1.5}} & {\tiny\Vectorstack{+2.3\\ \pm 2.4}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.2\\ \pm 2.9}}} & {\tiny\Vectorstack{-3.5\\ \pm 2.2}} & {\tiny\Vectorstack{-1.9\\ \pm 2.9}} & {\tiny\Vectorstack{-4.4\\ \pm 2.3}} & {\tiny\Vectorstack{+3.1\\ \pm 2.5}} & {\tiny\Vectorstack{+1.6\\ \pm 2.6}} & {\tiny\Vectorstack{+3.7\\ \pm 3.6}} & {\tiny\Vectorstack{+0.4\\ \pm 2.1}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=3$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 1.3}}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.1\\ \pm 1.6}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+13\\ \pm 1.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+17\\ \pm 2.9}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+7.3\\ \pm 2.7}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.9\\ \pm 2.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.5\\ \pm 2.0}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+14\\ \pm 2.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+12\\ \pm 2.5}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+15\\ \pm 4.2}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 2.0}}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)&{\tiny\Vectorstack{+2.8\\ \pm 1.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.1\\ \pm 1.6}}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+5.1\\ \pm 2.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+9.0\\ \pm 3.0}}} & {\tiny\Vectorstack{-0.7\\ \pm 2.7}} & {\tiny\Vectorstack{+0.9\\ \pm 2.0}} & {\tiny\Vectorstack{-1.6\\ \pm 2.3}} & {\tiny\Vectorstack{+5.9\\ \pm 3.2}} & {\tiny\Vectorstack{+4.4\\ \pm 2.5}} & {\tiny\Vectorstack{+6.5\\ \pm 4.2}} & {\tiny\Vectorstack{+3.2\\ \pm 2.5}} \\[2pt]
+\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-2.3\\ \pm 2.4}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-13\\ \pm 1.8}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-5.1\\ \pm 2.3}}} &  & {\tiny\Vectorstack{+3.9\\ \pm 4.1}} & {\tiny\Vectorstack{-5.8\\ \pm 3.6}} & {\tiny\Vectorstack{-4.3\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.7\\ \pm 2.5}}} & {\tiny\Vectorstack{+0.8\\ \pm 3.7}} & {\tiny\Vectorstack{-0.8\\ \pm 2.7}} & {\tiny\Vectorstack{+1.4\\ \pm 4.0}} & {\tiny\Vectorstack{-1.9\\ \pm 3.1}} \\[2pt]
+\textbf{Baseline} ($\mathrm{sum}$)&\cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.2\\ \pm 2.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-17\\ \pm 2.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-9.0\\ \pm 3.0}}} & {\tiny\Vectorstack{-3.9\\ \pm 4.1}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-9.8\\ \pm 2.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.2\\ \pm 2.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 2.6}}} & {\tiny\Vectorstack{-3.1\\ \pm 3.0}} & {\tiny\Vectorstack{-4.7\\ \pm 3.1}} & {\tiny\Vectorstack{-2.6\\ \pm 5.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-5.8\\ \pm 2.8}}} \\[2pt]
+\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{+3.5\\ \pm 2.2}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-7.3\\ \pm 2.7}}} & {\tiny\Vectorstack{+0.7\\ \pm 2.7}} & {\tiny\Vectorstack{+5.8\\ \pm 3.6}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+9.8\\ \pm 2.4}}} &  & {\tiny\Vectorstack{+1.6\\ \pm 3.5}} & {\tiny\Vectorstack{-0.9\\ \pm 2.2}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.6\\ \pm 2.5}}} & {\tiny\Vectorstack{+5.1\\ \pm 3.2}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+7.2\\ \pm 3.3}}} & {\tiny\Vectorstack{+3.9\\ \pm 3.3}} \\[2pt]
+\textbf{2-GNN} ($\mean$)&{\tiny\Vectorstack{+1.9\\ \pm 2.9}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.9\\ \pm 2.3}}} & {\tiny\Vectorstack{-0.9\\ \pm 2.0}} & {\tiny\Vectorstack{+4.3\\ \pm 2.5}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.2\\ \pm 2.9}}} & {\tiny\Vectorstack{-1.6\\ \pm 3.5}} &  & {\tiny\Vectorstack{-2.4\\ \pm 2.6}} & {\tiny\Vectorstack{+5.1\\ \pm 3.5}} & {\tiny\Vectorstack{+3.5\\ \pm 2.2}} & {\tiny\Vectorstack{+5.6\\ \pm 4.8}} & {\tiny\Vectorstack{+2.4\\ \pm 2.6}} \\[2pt]
+\textbf{2-GNN} ($\mathrm{SAM}$)&{\tiny\Vectorstack{+4.4\\ \pm 2.3}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.5\\ \pm 2.0}}} & {\tiny\Vectorstack{+1.6\\ \pm 2.3}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.7\\ \pm 2.5}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 2.6}}} & {\tiny\Vectorstack{+0.9\\ \pm 2.2}} & {\tiny\Vectorstack{+2.4\\ \pm 2.6}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+7.5\\ \pm 3.1}}} & {\tiny\Vectorstack{+5.9\\ \pm 3.2}} & {\tiny\Vectorstack{+8.1\\ \pm 4.2}} & {\tiny\Vectorstack{+4.8\\ \pm 3.2}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)&{\tiny\Vectorstack{-3.1\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-14\\ \pm 2.8}}} & {\tiny\Vectorstack{-5.9\\ \pm 3.2}} & {\tiny\Vectorstack{-0.8\\ \pm 3.7}} & {\tiny\Vectorstack{+3.1\\ \pm 3.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.6\\ \pm 2.5}}} & {\tiny\Vectorstack{-5.1\\ \pm 3.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-7.5\\ \pm 3.1}}} &  & {\tiny\Vectorstack{-1.5\\ \pm 2.2}} & {\tiny\Vectorstack{+0.6\\ \pm 2.9}} & {\tiny\Vectorstack{-2.7\\ \pm 2.3}} \\[2pt]
+\textbf{2-WL-GNN\phantom{*}} ($\mean$)&{\tiny\Vectorstack{-1.6\\ \pm 2.6}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-12\\ \pm 2.5}}} & {\tiny\Vectorstack{-4.4\\ \pm 2.5}} & {\tiny\Vectorstack{+0.8\\ \pm 2.7}} & {\tiny\Vectorstack{+4.7\\ \pm 3.1}} & {\tiny\Vectorstack{-5.1\\ \pm 3.2}} & {\tiny\Vectorstack{-3.5\\ \pm 2.2}} & {\tiny\Vectorstack{-5.9\\ \pm 3.2}} & {\tiny\Vectorstack{+1.5\\ \pm 2.2}} &  & {\tiny\Vectorstack{+2.1\\ \pm 3.4}} & {\tiny\Vectorstack{-1.1\\ \pm 2.0}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{-3.7\\ \pm 3.6}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-15\\ \pm 4.2}}} & {\tiny\Vectorstack{-6.5\\ \pm 4.2}} & {\tiny\Vectorstack{-1.4\\ \pm 4.0}} & {\tiny\Vectorstack{+2.6\\ \pm 5.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-7.2\\ \pm 3.3}}} & {\tiny\Vectorstack{-5.6\\ \pm 4.8}} & {\tiny\Vectorstack{-8.1\\ \pm 4.2}} & {\tiny\Vectorstack{-0.6\\ \pm 2.9}} & {\tiny\Vectorstack{-2.1\\ \pm 3.4}} &  & {\tiny\Vectorstack{-3.3\\ \pm 4.3}} \\[2pt]
+\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{-0.4\\ \pm 2.1}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 2.0}}} & {\tiny\Vectorstack{-3.2\\ \pm 2.5}} & {\tiny\Vectorstack{+1.9\\ \pm 3.1}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+5.8\\ \pm 2.8}}} & {\tiny\Vectorstack{-3.9\\ \pm 3.3}} & {\tiny\Vectorstack{-2.4\\ \pm 2.6}} & {\tiny\Vectorstack{-4.8\\ \pm 3.2}} & {\tiny\Vectorstack{+2.7\\ \pm 2.3}} & {\tiny\Vectorstack{+1.1\\ \pm 2.0}} & {\tiny\Vectorstack{+3.3\\ \pm 4.3}} &
 \end{tabular}}

### 2020-04-13 implementation/src/ltag/playground.py

@@ -303,12 +303,25 @@ def ds_stats():
 # synthetic.balanced_triangle_classification_dataset(stored=True)().export_dot(152)
 # synthetic.balanced_triangle_classification_dataset(stored=True)().export_dot(153)
 # ds_stats()
-# p = eval_ds.Proteins_5()
-# d = eval_ds.DD_2()
-# p.draw(402)
-# d.draw(2)
-# p.export_dot(402)
-# d.export_dot(2)
+p = eval_ds.Proteins_5()
+d = eval_ds.DD_2()
+nci = eval_ds.NCI1_8()
+imdb = eval_ds.IMDB_8()
+reddit = eval_ds.RedditBinary_1()
+
+nci.draw(30)
+nci.draw(33)
+nci.draw(204)
+nci.draw(1300)
+nci.draw(4101)
+
+imdb.draw(107)
+imdb.draw(999)
+imdb.export_dot(999)
+
+reddit.draw(2)
+reddit.draw(100)
+reddit.export_dot(100)
 #
 # list(dsm.get_all(output_type="grakel")[0])
 #

### 2020-04-14 thesis/content/chapter-conclusion.tex
added future directions

--- 2019-10-29 thesis/content/chapter-conclusion.tex
@@
 
+To conclude the thesis, we now look back on the three research questions described in \cref{sec:intro:questions} and summarize the answers that were given to them in the previous chapters.
+Afterwards, a brief overview of future research directions based on our findings will be given.
+
 \section{Review}%
@@
 
+\paragraph{\circled{1}\; What constitutes an \ac{lta} method?}
+We began with a general definition of \ac{lta} in \cref{sec:ltag:definition}.
+There we proposed that its defining characteristic is the \textit{localized explainability} of its predictions.
+This characteristic was formalized via the notion of \textit{\ac{lta} formulations} (see \cfullref{defn:ltag:formulation}) which requires that a model is expressible in terms of a decomposition function $\psi: \mathcal{G} \to \mathcal{P}(\mathcal{G})$, a local evaluation function $f: \mathcal{G} \to \mathcal{Y} \times \mathbb{R}$ and a weighted aggregation function $\mathcal{A}: {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^* \to \mathcal{Y}$.
+An ideal \ac{lta} method has such a formulation with a decomposition function $\psi$ that splits graphs into ``meaningful'' constituents in some domain-specific sense of the word.
+Since this ideal notion of \ac{lta} is generally quite fuzzy, we only distinguished between \acs{lta}-like and non-\acs{lta} methods in this thesis;
+a method was called non-\acs{lta} if it uses a trivial decomposition function that just splits a graph $G$ into the single global constituent $G$.
+
+\paragraph{\circled{2}\; How do existing \ac{gcr} methods relate to \ac{lta}?}
+In \cref{sec:ltag:formulation:svm,sec:ltag:formulation:gcnn} we used our definition of \ac{lta} to check which of the existing \ac{gcr} approaches are compatible with it.
+For the case of an \ac{svm} using a graph kernel/embedding, we found that it is an \acs{lta}-like method if the kernel is a so-called nontrivial \acf{sce} (see \cfullref{defn:ltag:substruct-embedding,thm:ltag:svm-ltag-formulation}).
+This \ac{sce} condition is satisfied by fingerprint embeddings, the \ac{wl} subtree kernel and the 2-LWL kernel, which makes them \acs{lta}-like.
+However, \texttt{graph2vec} embeddings, the \ac{wl} shortest path kernel and the 2-GWL kernel were found to be trivial or only partly nontrivial \acp{sce}, i.e.\ they are non-\acs{lta} methods.
+After considering those embedding approaches we looked at \ac{gcnn} and showed that they also have an \ac{lta} formulation under certain conditions (see \cfullref{thm:ltag:gcnn-ltag-formulation}).
+
+\paragraph{\circled{3}\; What are limitations of existing graph \ac{lta} methods and how can they be overcome?}
+While graph embeddings techniques can, in principle, decompose a given graph into arbitrary constituents (see fingerprint embeddings), the \ac{lta} formulation of \acp{gcnn} uses constituents that are spanned by the \ac{bfs} subtrees of its input graph.
+In \cref{sec:ltd:edge-filter} we addressed this limitation of \acp{gcnn} by proposing that more flexible decompositions can be learned via an edge filtering strategy.
+The idea behind this is to selectively prune subtree constituents by learning to ignore certain edges based on informative edge feature vectors.
+To produce such edge feature vectors we first looked at 2-\acsp{gnn} and found that they have various theoretical limitations, i.e.\ the inability to distinguish regular graphs and to detect cycles (see \cfullref{prop:ltd:2gnn-regular-limit,prop:ltd:2gnn-cycle-limit}).
+We therefore proposed the novel 2-\acs{wl}-\acs{gnn} which does not have those limitations (see \cfullref{cor:ltd:wl2-gnn-regular}).
+It additionally also has a strictly greater expressive power than all the 1-\acs{wl} bounded vertex neighborhood aggregation \acp{gnn}, e.g.\ \ac{gcn} or \ac{gin} (see \cfullref{cor:ltd:wl2-gnn-more-wl1-power}).
+
+\paragraph{Evaluation results}
+For the evaluation of our results we considered two aspects:
+Firstly, we looked at how 2-\acs{wl}-\acsp{gnn} compare to other \acp{gnn}.
+Secondly, we evaluated how \acs{lta}-like methods compare to non-\acs{lta} methods.
+Regarding the first aspect, we showed that the theoretical advantages of 2-\acs{wl}-\acsp{gnn} are clearly observable on the synthetic triangle detection dataset while on the evaluated real-world datasets we got results which are generally comparable with the best state-of-the-art approaches but not significantly better.
+Regarding the second aspect, we observed no general advantage or disadvantage of \acs{lta}-like methods.
+While the \acs{lta}-like configurations of 2-\acs{wl}-\acsp{gnn} generally performed worse than their non-\ac{lta} counterparts, the \acs{lta}-like \ac{wl} subtree kernel generally performed quite well.
+This shows that \ac{lta} is in principle suitable for graph classification tasks if the right decomposition, evaluation and aggregation functions are chosen.
+
 \section{Future Directions}%
 \label{sec:conclusion:todo}
+
+Based on our results, let us now consider which follow-up questions could be tackled in future research.
+
+\paragraph{Further theoretical analysis of 2-\acs{wl}-\acsp{gnn}}
+Even though we proved that our proposed 2-\acs{wl} inspired convolution approach has certain theoretical advantages over the existing methods, we did not fully characterize its expressive and computational power w.r.t.\ the 2-\acs{wl} algorithm itself.
+As we empirically showed in \cref{sec:eval:synthetic}, 2-\acs{wl}-\acsp{gnn} are able to learn to detect certain triangles in graphs.
+We did however not consider the aspect of $m$-cycle detection or counting for $m > 3$.
+Further work is required to adapt the theoretical results for the 2-\acs{wl} algorithm to the graph convolution setting.
+One particularly interesting question there would be to analyze how exactly the neighborhood radius $r$ relates to the expressive power of a 2-\acs{wl}-\acs{gnn}.
+
+\paragraph{Further empirical real-world evaluation of 2-\acs{wl}-\acsp{gnn}}
+In our evaluations on real-world datasets we did not observe a clear advantage of 2-\acs{wl}-\acsp{gnn} over other methods.
+We did however find indicators that for the hypothesis that 2-\acs{wl}-\acsp{gnn} are able to perform better on datasets containing graphs with cycles when increasing the neighborhood radius.
+The cycle detection abilities of 2-\acs{wl} could therefore potentially be relevant for real-world problems.
+Future research could evaluate 2-\acs{wl}-\acsp{gnn} on a more diverse set of problem domains, e.g.\ software verification on \acfp{cfg}, and with other pooling layers to determine to which extent the theoretical advantages of 2-\acs{wl}-\acsp{gnn} can lead to an advantage in practice.
+
+\paragraph{Evaluation of the explainability of \acs{lta}-like models}
+Our definition of \ac{lta} was motivated by the idea that the local evaluations of constituents provide an explanation for the global aggregated prediction.
+However, in our evaluation of \ac{lta} we focused mostly on the question how \acs{lta}-like approaches perform compared to non-\acs{lta} approaches with the result that \acs{lta}-like models does not appear to have a general advantage or disadvantage regarding classification accuracy.
+In the next step it should be evaluated whether the constituent scores of an \acs{lta}-like model are meaningful and therefore provide an advantage from the perspective of \acl{xai}.
+
+\paragraph{Solving the \ac{ltd} problem via edge filtering}
+The motivation behind the proposed 2-\acs{wl}-\acs{gnn} was to use it to obtain informative edge feature vectors with which an edge filter could be trained to dynamically decompose graphs.
+The results in this thesis provide the first step towards the realization of this idea.
+Further research is required to determine whether this approach is suitable to learn meaningful constituents in real-world problem domains.

### 2020-04-14 thesis/content/chapter-introduction.tex
added future directions

@@
 Such an extension is interesting because \ac{lta} models produce composition predictions which are the direct result of local constituent predictions.
-This means that \ac{lta} models provide insight into the influence of constituents on the final prediction, which makes \ac{lta} relevant from the perspective \ac{xai}~\cite{Gilpin2018}.
+This means that \ac{lta} models provide insight into the influence of constituents on the final prediction, which makes \ac{lta} relevant from the perspective \acl{xai}~\cite{Gilpin2018}.
 \Cref{fig:intro:intuition} illustrates how a graph-variant of \ac{lta} could explain the prediction of some graph property.

### 2020-04-14 thesis/content/chapter-related-work.tex
added future directions

@@
 	\item \textbf{Aggregation:}
-		A variadic aggregation function $\mathcal{A}: \mathcal{Y}^{*} \to \mathcal{Y}$ that estimates composite scores has to be learned, i.e.\@ $y \approx \hat{y} = \mathcal{A}(y_{1}, \dots, y_{n})$.
+		A variadic aggregation function $\mathcal{A}: \mathcal{Y}^{*} \to \mathcal{Y}$ that estimates composition scores has to be learned, i.e.\@ $y \approx \hat{y} = \mathcal{A}(y_{1}, \dots, y_{n})$.
 		Typically the aggregation function $\mathcal{A}$ should be associative and commutative to fit with the multiset-structure of compositions.
@@
 		In case the constituent scores $y_i$ are not given, they have to be derived from a constituent representation, e.g.\ a vector $x_i \in \mathcal{X}$.
-		To learn this derivation function $f: \mathcal{X} \to \mathcal{Y}$, only the constituent vectors ${\ldblbrace x_i \rdblbrace}_{i = 1}^{n}$ and the composite score $y$ is given.
+		To learn this derivation function $f: \mathcal{X} \to \mathcal{Y}$, only the constituent vectors ${\ldblbrace x_i \rdblbrace}_{i = 1}^{n}$ and the composition score $y$ is given.
 		Thus the constituent scores $y_i$ need to be \textit{disaggregated} from $y$ in order to learn $f$.
@@
 The first approach uses \textit{uninorms}~\cite{Melnikov2016} to do so.
-There the basic idea is to express composite scores as fuzzy truth assignments $y \in [0, 1]$.
-Such a composite assignment $y$ is modeled as the result of a parameterized logical expression of constituent assignments $y_i \in [0, 1]$.
+There the basic idea is to express composition scores as fuzzy truth assignments $y \in [0, 1]$.
+Such a score assignment $y$ is modeled as the result of a parameterized logical expression of constituent assignments $y_i \in [0, 1]$.
 As the logical expression that thus effectively aggregates the constituents, a uninorm $U_{\lambda}$ is used.

### 2020-04-14 thesis/content/chapter-conclusion.tex
corrections

@@
 However, \texttt{graph2vec} embeddings, the \ac{wl} shortest path kernel and the 2-GWL kernel were found to be trivial or only partly nontrivial \acp{sce}, i.e.\ they are non-\acs{lta} methods.
-After considering those embedding approaches we looked at \ac{gcnn} and showed that they also have an \ac{lta} formulation under certain conditions (see \cfullref{thm:ltag:gcnn-ltag-formulation}).
+After considering those embedding approaches we looked at \acp{gcnn} and showed that they also have an \ac{lta} formulation under certain conditions (see \cfullref{thm:ltag:gcnn-ltag-formulation}).
 
@@
 The idea behind this is to selectively prune subtree constituents by learning to ignore certain edges based on informative edge feature vectors.
-To produce such edge feature vectors we first looked at 2-\acsp{gnn} and found that they have various theoretical limitations, i.e.\ the inability to distinguish regular graphs and to detect cycles (see \cfullref{prop:ltd:2gnn-regular-limit,prop:ltd:2gnn-cycle-limit}).
+To produce such edge feature vectors, we first looked at 2-\acsp{gnn} and found that they have various theoretical limitations, i.e.\ the inability to distinguish regular graphs and to detect cycles (see \cfullref{prop:ltd:2gnn-regular-limit,prop:ltd:2gnn-cycle-limit}).
 We therefore proposed the novel 2-\acs{wl}-\acs{gnn} which does not have those limitations (see \cfullref{cor:ltd:wl2-gnn-regular}).
@@
 In our evaluations on real-world datasets we did not observe a clear advantage of 2-\acs{wl}-\acsp{gnn} over other methods.
-We did however find indicators that for the hypothesis that 2-\acs{wl}-\acsp{gnn} are able to perform better on datasets containing graphs with cycles when increasing the neighborhood radius.
+We did however find indicators for the hypothesis that 2-\acs{wl}-\acsp{gnn} are able to perform better on datasets containing graphs with cycles when increasing the neighborhood radius.
 The cycle detection abilities of 2-\acs{wl} could therefore potentially be relevant for real-world problems.
-Future research could evaluate 2-\acs{wl}-\acsp{gnn} on a more diverse set of problem domains, e.g.\ software verification on \acfp{cfg}, and with other pooling layers to determine to which extent the theoretical advantages of 2-\acs{wl}-\acsp{gnn} can lead to an advantage in practice.
+Future research could evaluate 2-\acs{wl}-\acsp{gnn} on a more diverse set of problem domains, e.g.\ software analysis on \acfp{cfg}, and with other pooling layers to determine to which extent the theoretical advantages of 2-\acs{wl}-\acsp{gnn} can lead to an advantage in practice.
 
@@
 Our definition of \ac{lta} was motivated by the idea that the local evaluations of constituents provide an explanation for the global aggregated prediction.
-However, in our evaluation of \ac{lta} we focused mostly on the question how \acs{lta}-like approaches perform compared to non-\acs{lta} approaches with the result that \acs{lta}-like models does not appear to have a general advantage or disadvantage regarding classification accuracy.
+However, in our evaluation of \ac{lta} we focused mostly on the question how \acs{lta}-like approaches perform compared to non-\acs{lta} approaches, with the result that \acs{lta}-like models does not appear to have a general advantage or disadvantage regarding classification accuracy.
 In the next step it should be evaluated whether the constituent scores of an \acs{lta}-like model are meaningful and therefore provide an advantage from the perspective of \acl{xai}.

### 2020-04-14 thesis/content/chapter-evaluation.tex
corrections

@@
 The ideas presented in both chapters will now be empirically evaluated.
-Do to so we differentiate between two mostly independent evaluation aspects:
+To do so we differentiate between two mostly independent evaluation aspects:
 \begin{enumerate}[label={\textbf{\arabic*.}}]
@@
 		Even though it was motivated by \ac{lta}, a 2-\acs{wl}-\acs{gnn} is not generally more ``\acs{lta}-like'' than other approaches.
-		Nonetheless, due to the theoretical advantages described in \cref{sec:ltd:wl2gnn:properties} it is an interesting approach independently from its potential applications in \ac{lta} (see \cref{sec:ltd:edge-filter}).
+		Nonetheless, due to the theoretical advantages described in \cref{sec:ltd:wl2gnn:properties}, it is an interesting approach independently from its potential applications in \ac{lta} (see \cref{sec:ltd:edge-filter}).
 		Thus the first aspect of our evaluation is to compare 2-\acs{wl}-\acsp{gnn} with the other previously described \ac{gcr} methods in a general non-\acs{lta} fashion, i.e.\ with an added \ac{mlp} after the pooling layer since this is how \acp{gnn} are typically evaluated in other works.
@@
 	\item We then present results on synthetically generated data in \cref{sec:eval:synthetic}.
-	 	There we will illustrate the higher expressive power of 2-\acs{wl}-\acsp{gnn} when compared to other \ac{gcr} approaches which confirms the theoretical results from \cref{sec:ltd:wl2gnn:properties}.
+	 	There we will illustrate the higher expressive power of 2-\acs{wl}-\acsp{gnn} when compared to other \ac{gcr} approaches, which confirms the theoretical results from \cref{sec:ltd:wl2gnn:properties}.
 	\item Then evaluation results on real-world datasets are described in \cref{sec:eval:real}.
@@
 We base our evaluations on this assessment strategy with $k = 10$ folds and $r = 3$ repeats per fold to smooth out differences caused by random weight initializations.
-For each dataset the same folds are used across the evaluated models; class proportions are preserved within each fold by using stratified splits.
+For each dataset, the same folds are used across the evaluated models; class proportions are preserved within each fold by using stratified splits.
 To keep the total runtime of the experiments feasible, a single 90\%/10\% holdout split into training and validation data is used instead of cross-validation.
@@
 
-Using this assessment strategy we evaluate \acp{svm} with the following graph kernels:
+Using this assessment strategy, we evaluate \acp{svm} with the following graph kernels:
 \begin{enumerate}[label={\textbf{\arabic*.}},itemsep=2pt,parsep=2pt]
-	\item \textbf{\ac{wl} subtree kernel (\acs{wl}\textsubscript{ST})} with the iteration counts $T \in \{ 1,2,3,4,5 \}$ to evaluate the influence of the depth of \ac{bfs} subtrees which span \ac{lta} constituents.
+	\item \textbf{\ac{wl} subtree kernel (\acs{wl}\textsubscript{ST})} with the iteration counts $T \in \{ 1,2,3,4,5 \}$, to evaluate the influence of the depth of \ac{bfs} subtrees which span \ac{lta} constituents.
 	\item \textbf{\ac{wl} shortest path kernel (\acs{wl}\textsubscript{SP})} with the iteration count $T = 3$.
@@
 	\item \textbf{Structure unaware baseline:}
-		\citet{Errica2020} describe a simple model which simply applies a standard \ac{mlp} to each individual vertex feature vector, then sums up the resulting feature vectors and applies another \ac{mlp} to the vector sum.
+		\citet{Errica2020} describe a simple model which applies a standard \ac{mlp} to each individual vertex feature vector, then sums up the resulting feature vectors and applies another \ac{mlp} to the vector sum.
 		This approach does not use any structural information and therefore serves as a baseline to detect whether a \ac{gnn} is able to exploit graph structure.
@@
 	\item \textbf{2-\acs{gnn}} is evaluated with both a static $\mean$ pooling layer and with \ac{sampool} (see \cfullref{defn:ltag:sam-pool}).
-		After the pooling layer a \ac{mlp} is used to produce the final prediction.
+		After the pooling layer, an \ac{mlp} is used to produce the final prediction.
 	\item \textbf{2-\acs{wl}-\ac{gnn}} \textit{(our method)} is also evaluated with $\mean$ pooling and \ac{sampool}, just like 2-\ac{gnn}.
-		To test the \ac{lta} assumption we evaluate 2-\acs{wl}-\acs{gnn} in an \acs{lta}-like configuration and in the standard non-\acs{lta} configuration.
+		To test the \ac{lta} assumption, we evaluate 2-\acs{wl}-\acs{gnn} in an \acs{lta}-like configuration and in the standard non-\acs{lta} configuration.
 		The \acs{lta}-like configuration uses a stack of 2-\acs{wl} convolutions that produce a local prediction $y_{ij} \in [0, 1]$ for each edge $e_{ij}$ instead of a local feature vector $z_{ij} \in \mathbb{R}^{d^{(T)}}$ and it does not have an \ac{mlp} after the pooling/aggregation layer.
@@
 The baseline and \ac{gin} results are obtained using the PyTorch-based implementation provided by \citet{Errica2020}.
-For both 2-\acs{gnn} and 2-\acs{wl}-\ac{gnn} a custom TensorFlow-based implementation is used.
+For both 2-\acs{gnn} and 2-\acs{wl}-\ac{gnn}, a custom TensorFlow-based implementation is used.
 The code for all conducted experiments as well as the used dataset splits are available GitHub\footnote{\url{https://github.com/Cortys/master-thesis}}.
@@
 
-Based on this problem we generated a synthetic triangle detection dataset.
+Based on this problem, we generated a synthetic triangle detection dataset.
 It contains randomly generated graphs with varying vertex counts and vertex color proportions (see \cref{sec:appendix:ds-stats} for a more detailed description).
@@
 Let us now take a look at the 2-\acs{wl} inspired kernels: 2-LWL and 2-GWL.\@
-Interestingly both kernels do not appear to generalize better the 1-\acs{wl} bounded methods;
+Interestingly, both kernels do not appear to generalize better the 1-\acs{wl} bounded methods;
 we explain this by the fairly small size of the triangle detection dataset (228 graphs).
-Even though both kernels embed graphs into a space with dimensions that indicate the presence of a unicolored triangle (see \cfullref{prop:related:wl2-cycle-count}), there are many of those triangle-indicating embedding dimensions s.t.\ the indicating dimensions found in a given training split might not overlap with those in the test split.
+Even though both kernels embed graphs into a space with dimensions that indicate the presence of a unicolored triangle (see \cfullref{prop:related:wl2-cycle-count}), there are so many of those triangle-indicating embedding dimensions that the indicating dimensions found in a given training split might not overlap with those in the test split.
 
-Looking at the 2-\acs{wl} inspired \acp{gnn} (2-\acs{gnn}, 2-\acs{wl}-\acs{gnn}), we find that the proposed 2-\acs{wl}-\acs{gnn} significantly outperforms all other evaluated methods which confirms our theoretical results from \cref{sec:ltd:wl2gnn:properties}.
+Looking at the 2-\acs{wl} inspired \acp{gnn} (2-\acs{gnn}, 2-\acs{wl}-\acs{gnn}), we find that the proposed 2-\acs{wl}-\acs{gnn} significantly outperforms all other evaluated methods, which confirms our theoretical results from \cref{sec:ltd:wl2gnn:properties}.
 We evaluated 2-\acs{wl}-\acsp{gnn} in an \acs{lta}-like configuration without a final \ac{mlp} and in a non-\acs{lta} configuration with an \ac{mlp} after the pooling layer.
@@
 		The REDDIT and IMDB datasets~\cite{Yanardag2015} contain graphs representing the relations between users in online discussions and movie actors respectively.
-		For REDDIT the type of a community has to be predicted based on a discussion thread; for IMDB the goal is to determine movie genres.
+		For REDDIT, the type of a community has to be predicted based on a discussion thread; for IMDB, the goal is to determine movie genres.
 \end{enumerate}
@@
 Regarding the \acs{lta} assumption (see \cfullref{defn:ltag:lta-assumption}), this is evidence that the decomposition function $\psi$ of 2-\acs{wl}-\acsp{gnn} (see \cfullref{thm:ltag:gcnn-ltag-formulation}) does not produce constituents $c_i \in \psi(G)$ whose local evaluations $f(c_i) \in \mathcal{Y}$ are as indicative of the graph $G$'s class $y \in \mathcal{Y}$ as an arbitrary feature vector $z_i \in \mathbb{R}^{d^{(T)}}$.
-This does however not imply, that the evaluated real-world problem domains are generally incompatible with the \ac{lta} assumption.
+This does however not imply that the evaluated real-world problem domains are generally incompatible with the \ac{lta} assumption.
 When considering the results of the \ac{lta}-like \acs{wl}\textsubscript{ST} model, we find that it is mostly comparable with the alternative non-\acs{lta} approaches and, in the case of NCI1, even significantly better.
@@
 
-We now analyze for which degree of constituent locality \acs{wl}\textsubscript{ST} and 2-\acs{wl}-\acsp{gnn} perform best by evaluating them using varying tree depths $T$.
+We now analyze for which degree of constituent locality \acs{wl}\textsubscript{ST} and 2-\acs{wl}-\acsp{gnn} perform best, by evaluating them using varying tree depths $T$.
 For \acs{wl}\textsubscript{ST} we can set $T$ directly via the number of refinement steps.
@@
 This implies that the best-performing subtree constituents do not span entire connected components but smaller substructures on average.
-The REDDIT dataset shows this very clearly; there a subtree must have a depth of at least $5$ in order to span an average graph while the best-performing subtree constituents only have a depth of $T = 2$.
+The REDDIT dataset shows this very clearly; there a subtree must have a depth of at least $5$ in order to span an average graph, while the best-performing subtree constituents only have a depth of $T = 2$.
 This indicates that the REDDIT community detection problem is better described by small localized constituents than by large non-localized ones.
@@
 
-In addition to its influence on constituent sizes, a higher neighborhood radius also introduces additional feature vectors into the convolved feature matrices $Z^{(t)} \in \mathbb{R}^{\left|\mathcal{E}_{G^r}\right| \times d^{(t)}}$.
+Apart from its influence on constituent sizes, a higher neighborhood radius also introduces additional feature vectors into the convolved feature matrices $Z^{(t)} \in \mathbb{R}^{\left|\mathcal{E}_{G^r}\right| \times d^{(t)}}$.
 As we saw in the proof sketch of 2-\acs{wl}'s cycle counting ability (see \cfullref{prop:related:wl2-cycle-count}), those additional features/colors can carry important structural information about a graph.
@@
 \Cref{fig:eval:wl2-gnn-radius} shows that adding this information via a neighborhood radius $r > 1$ does correlate with a higher training and test accuracy on the NCI1 and D\&D datasets; on the IMDB and PROTEINS datasets this is not the case.
-This difference is interesting because NCI1 (molecular structures) and D\&D (protein sequences) contain more cyclic graphs while IMDB (ego-network structures) and PROTEINS (protein sequences) consist of more tree- or list-like graphs (see \cref{sec:appendix:ds-stats}).
+This difference is interesting because NCI1 (molecular structures) and D\&D (protein sequences) contain more cyclic graphs, while IMDB (ego-network structures) and PROTEINS (protein sequences) consist of more tree- or list-like graphs (see \cref{sec:appendix:ds-stats}).
 Even though the PROTEINS and D\&D datasets both contain protein sequences, we find that the protein sequences in the PROTEINS dataset are very ``list-like'' with much fewer large cycles than in the proteins structures of the D\&D dataset (compare the vertex and edge count statistics of both datasets in \cref{tbl:appendix:ds-stats}).

### 2020-04-14 thesis/content/chapter-graph-lta.tex
improved intro and conclusion
corrections

@@
 In the previous chapter an introduction to two separate fields of research was given:
-\begin{enumerate*}
-	\item \Acf{lta},
-	\item \Acf{gcr}
-\end{enumerate*}.
-In this chapter we will combine them and define an extension of \ac{lta} to the \ac{gcr} problem.
+\begin{enumerate*}[label={\circled{\small\arabic*}}]
+	\item \Acf{lta}.
+	\item \Acf{gcr}.
+\end{enumerate*}
+In this chapter we combine them and define an extension of \ac{lta} to the \ac{gcr} problem.
 This will be done in three steps:
@@
 	\item \textbf{Local evaluation:}
-		The constituents $c_{G, i} \in \psi(G)$ must be evaluated via some function $f: \mathcal{G} \to \mathcal{Y} \times \mathbb{R}$.
-		This \textit{evaluation function} assigns a prediction $\hat{y}_{G, i} \in \mathcal{Y}$ and a weight $w_{G, i} \in \mathbb{R}$ to each constituent.
+		The constituents $c_{G, i} \in \psi(G)$ must be evaluated via some function $f: \mathcal{G} \to \mathcal{Y} \times \mathbb{R}_{\geq 0}$.
+		This \textit{evaluation function} assigns a prediction $\hat{y}_{G, i} \in \mathcal{Y}$ and a weight $w_{G, i} \in \mathbb{R}_{\geq 0}$ to each constituent.
 		A constituent's weight $w_{G, i}$ can intuitively be interpreted as a measure of the confidence that the local prediction $\hat{y}_{G, i}$ is indicative of the composition's global target score $y_G$.
@@
 		For structured data however, where the decomposition $\psi(G)$ is not given as part of the input, this assumption does not necessarily hold.
-		By weighting the constituents, an \ac{lta} model can reduce the relevance or even ignore constituents that turn out to be irrelevant for the compositions target score $y_G$.
+		By weighting the constituents, an \ac{lta} model can reduce the relevance or even ignore constituents that turn out to be irrelevant for the composition's target score $y_G$.
 	\item \textbf{Aggregation:}
-		Lastly a \textit{weighted aggregation function} $\mathcal{A}: {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*} \to \mathcal{Y}$ must be applied.
+		Lastly, a \textit{weighted aggregation function} $\mathcal{A}: {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*} \to \mathcal{Y}$ must be applied.
 		It combines the multiset of local constituent predictions and non-negative weights into a single global composition prediction.
@@
 			\item Another basic weighted aggregator is the \textit{weighted majority vote} function $\wmaj({\ldblbrace (y_i, w_i) \rdblbrace}_{i=1}^n) \coloneqq {\arg\max}_{y \in \mathcal{Y}} \sum_{y_i = y} w_i$ which returns the input with the highest total weight.
-				Unlike $\wmean$ it can also be applied to score domains $\mathcal{Y}$ without a multiplication operator, e.g.\ sets of discrete classes.
-			\item Alternatively an unweighted aggregation function like $\min$, $\max$, $\mean$ or \ac{owa}\footnote{
+				Unlike $\wmean$, it can also be applied to score domains $\mathcal{Y}$ without a multiplication operator, e.g.\ sets of discrete classes.
+			\item Alternatively, an unweighted aggregation function like $\min$, $\max$, $\mean$ or \ac{owa}\footnote{
 				Even though \ac{owa} uses weights, it does not get those weights as part of the input and is therefore considered to be unweighted in this context.
@@
 Based on the notion of decomposition, local evaluation and aggregation we can now define the concept of \textit{\ac{lta} formulations}.
-\begin{defn}
+\begin{defn}\label{defn:ltag:formulation}
 	A model $h: \mathcal{G} \to \mathcal{Y}$ is in an \textit{\ac{lta} formulation} iff.\ it is expressed as
@@
 \end{defn}
-Note that every model $h: \mathcal{G} \to \mathcal{Y}$ has a trivial recursive \ac{lta} formulation by choosing $\psi(G) = \{ G \}$, $f(G) = (h(G), 1)$ and an arbitrary weighted aggregation function $\mathcal{A}$.
+Note that every model $h: \mathcal{G} \to \mathcal{Y}$ has a trivial recursive \ac{lta} formulation by choosing $\psi(G) = \{ G \}$, $f(G) = (h(G), \eta)$ and an arbitrary weighted aggregation function $\mathcal{A}$.
 Those trivial \ac{lta} formulations do not split compositions into locally evaluated constituents and therefore intuitively do not fulfill the postulated localized explainability characteristic of \ac{lta}.
-However since there is no commonly accepted formal criterion to decide whether a model's decisions are explainable~\cite{Lipton2018}, we do not attempt to strictly distinguish between \ac{lta} and non-\acs{lta} methods.
-Instead the notion of \ac{lta} formulations should be seen as way to identify how ``\acs{lta}-like'' a model is:
+However, since there is no commonly accepted formal criterion to decide whether a model's decisions are explainable~\cite{Lipton2018}, we do not attempt to strictly distinguish between \ac{lta} and non-\acs{lta} methods.
+Instead, the notion of \ac{lta} formulations should be seen as way to identify how ``\acs{lta}-like'' a model is.
 \begin{itemize}
@@
 \begin{defn}\label{defn:ltag:lta-assumption}
-	A problem domain $\mathcal{D}$ satisfies the \textit{\ac{lta} assumption} iff.\ there is an \ac{lta} method which produces models with an equal or lower out-of-sample error than the models produced by non-\acs{lta} methods for most training samples $\mathcal{D}_{\mathrm{train}} \subseteq \mathcal{D}$.
+	A problem domain $\mathcal{D}$ satisfies the \textit{\ac{lta} assumption} iff.\ there is an \ac{lta} learner which produces models with an equal or lower out-of-sample error than the models produced by non-\acs{lta} learners for most training samples $\mathcal{D}_{\mathrm{train}} \subseteq \mathcal{D}$.
 \end{defn}
 Due to the fuzziness of the class of \ac{lta} methods, the \ac{lta} assumption is naturally also a fuzzy concept.
-Nonetheless evidence for its truthiness in a given domain $\mathcal{D}$ can be empirically obtained by comparing candidate \ac{lta} methods with the best known non-\ac{lta} method for $\mathcal{D}$, assuming that some cut-off condition for the required ``\ac{lta}-ness'' of an \ac{lta} method is agreed upon.
+Nonetheless, evidence for its truthiness in a given domain $\mathcal{D}$ can be empirically obtained by comparing candidate \ac{lta} methods with the best known non-\ac{lta} method for $\mathcal{D}$, assuming that some cut-off condition for the required ``\ac{lta}-ness'', i.e.\ constituent interpretability, of an \ac{lta} method is agreed upon.
 
@@
 Based on the general definition of \ac{lta} from the last section, we will now see to what extent the \ac{gcr} methods described in \cref{sec:related:gcr} can be interpreted as \ac{lta} instances.
-This section explores the relation between \acp{svm} that use graph embeddings and \ac{lta}.
-In \cref{sec:related:gcr:embed,sec:related:gcr:kernel} three different ways to map a given graph $G$ to a vector $\varphi(G) \in \mathbb{R}^d$ were described:
-\begin{enumerate*}
+This section explores the relation between \ac{lta} and \acp{svm} that use graph embeddings.
+In \cref{sec:related:gcr:embed,sec:related:gcr:kernel}, three different ways to map a given graph $G$ to a vector $\varphi(G) \in \mathbb{R}^d$ were described:
+\begin{enumerate*}[label={\circled{\small\arabic*}}]
 	\item Fingerprint embeddings,
@@
 \begin{defn}
-	Given a a multiset $A = \ldblbrace \underbrace{a_1, \dots, a_1}_{\gamma_A(a_1)\text{ times}},\allowbreak \dots, \underbrace{a_n, \dots, a_n}_{\gamma_A(a_n)\text{ times}} \rdblbrace \subseteq D$, the so-called \textit{multiplicity function} $\gamma_A: D \to \mathbb{N}_0$ maps each element of the domain $D$ to its multiplicity in $A$ with $\gamma_A(x) = 0 \Leftrightarrow x \notin A$.
+	Given a multiset $A = \ldblbrace \underbrace{a_1, \dots, a_1}_{\gamma_A(a_1)\text{ times}},\allowbreak \dots, \underbrace{a_n, \dots, a_n}_{\gamma_A(a_n)\text{ times}} \rdblbrace \subseteq D$, the so-called \textit{multiplicity function} $\gamma_A: D \to \mathbb{N}_0$ maps each element of the domain $D$ to its multiplicity in $A$, with $\gamma_A(x) = 0 \Leftrightarrow x \notin A$.
 \end{defn}
@@
 	there are decomposition functions $\psi_{\varphi, i}: \mathcal{G} \to \mathcal{P}(\mathcal{G})$ and multiplicity functions $\gamma_{\varphi, i}: \mathcal{G} \to \mathbb{N}_0$ for all embedding components $i \in [d]$ s.t.\ %
-	$\forall G \in \mathcal{G}, i \in [d]: \varphi(G)[i] = \sum_{c \in \psi_{\varphi, i}(G)} \gamma_{\varphi, i}(c)$ where $\gamma_{\varphi, i}$ is the multiplicity function of a multiset of the constituents $\psi_{\varphi, i}(G)$.
+	$\forall G \in \mathcal{G}, i \in [d]: \varphi(G)[i] = \sum_{c \in \psi_{\varphi, i}(G)} \gamma_{\varphi, i}(c)$, where $\gamma_{\varphi, i}$ is the multiplicity function of a multiset of the constituents $\psi_{\varphi, i}(G)$.
 	We call $\psi_{\varphi, i}$ an \textit{underlying decomposition} of the $i$-th component of the embedding $\varphi$.
-	Similarly the joint decomposition $\psi_{\varphi}(G) \coloneqq \bigcup_{i=1}^d \psi_{\varphi,i}(G)$ is an \textit{underlying decomposition} of $\varphi$.
+	Similarly, the joint decomposition $\psi_{\varphi}(G) \coloneqq \bigcup_{i=1}^d \psi_{\varphi,i}(G)$ is an \textit{underlying decomposition} of $\varphi$.
 \end{defn}
-Intuitively \cref{defn:ltag:substruct-embedding} states that the value of each \ac{sce} component must be derived from the number of constituents produced by some decomposition function where each constituent might be counted multiple times.
+Intuitively, \cref{defn:ltag:substruct-embedding} states that the value of each \ac{sce} component must be derived from the number of constituents produced by some decomposition function, where each constituent might be counted multiple times.
 Based on this requirement we now proof the main theorem which shows the relation between \acp{svm} and \ac{lta}.
@@
 \begin{thm}\label{thm:ltag:svm-ltag-formulation}
-	A binary \ac{svm} graph classifier $h$ that applies a \ac{sce} $\varphi$ to its inputs has an \ac{lta} formulation that uses an underlying decomposition $\psi_{\varphi}$ of $\varphi$.
+	A binary \ac{svm} graph classifier $h$ that applies an \ac{sce} $\varphi$ to its inputs has an \ac{lta} formulation that uses an underlying decomposition $\psi_{\varphi}$ of $\varphi$.
 \end{thm}
 \begin{proof}
-	Let $h: \mathcal{G} \to {\{ -1, +1 \}}$ be a binary graph classifier expressed as $h = h_{\mathrm{\acs*{svm}}} \circ \varphi$ where $\varphi: \mathcal{G} \to \mathbb{N}_0^{d}$ is an \ac{sce} and $h_{\mathrm{\acs*{svm}}}: \mathbb{R}^{d} \to {\{ -1, +1 \}}$ a standard \ac{svm} classifier.
-	Additionally, let $\psi_{\varphi}$ be some underlying decomposition of $\varphi$ with ${\{ \gamma_{\varphi, i} \}}_{i=1}^d$ being the corresponding multiplicity functions.
+	Let $h: \mathcal{G} \to {\{ -1, +1 \}}$ be a binary graph classifier expressed as $h = h_{\mathrm{\acs*{svm}}} \circ \varphi$, where $\varphi: \mathcal{G} \to \mathbb{N}_0^{d}$ is an \ac{sce} and $h_{\mathrm{\acs*{svm}}}: \mathbb{R}^{d} \to {\{ -1, +1 \}}$ a standard \ac{svm} classifier.
+	Additionally, let $\psi_{\varphi}$ be some underlying decomposition of $\varphi$, with ${\{ \gamma_{\varphi, i} \}}_{i=1}^d$ being the corresponding multiplicity functions.
 
@@
 	To complete the proof it now remains to show that $f_h$ and $\mathcal{A}_h$ are in fact a local evaluation function and a weighted aggregation function respectively.
-	To see that $f_h$ perform local evaluation, note that $f_h(c_t) \coloneqq (\sgn z_t, {|z_t|})$ with $z_t \coloneqq \sum_{i=1}^{d} \gamma_{\varphi,i}(c_t) \beta_i$ only depends on the shared multiplicity functions $\gamma_{\varphi,i}$, the constants $\beta_i$ and the constituent $c_t$;
+	To see that $f_h$ performs local evaluation, note that $f_h(c_t) \coloneqq (\sgn z_t, {|z_t|})$, with $z_t \coloneqq \sum_{i=1}^{d} \gamma_{\varphi,i}(c_t) \beta_i$ only depending on the shared multiplicity functions $\gamma_{\varphi,i}$, the constants $\beta_i$ and the constituent $c_t$;
 	apart from $c_t$ no other information from the input graph $G$ is required by $f_h$ which makes it a local evaluation function.
@@
 	\item \textbf{Fingerprint embeddings:}\label{itm:ltag:fingerprint-lta-formulation}
-		Here the $i$-th embedding component represent the number of occurrences of a substructure $S_i$, i.e.\ ${\varphi}_{\mathrm{FP}}(G)[i] \coloneqq \mathit{count}_{S_i}(G)$ (see \cfullref{defn:related:wl-count-detect}).
-		Such a fingerprint embedding naturally is an \ac{sce} with the decomposition functions $\psi_{\mathrm{FP}, i}(G) \coloneqq \{ G[\set(s)]\, |\, s \in \mathcal{V}_G^{*} \land G[s] \equiv S_i \}$ and the multiplicity functions $\gamma_{\mathrm{FP}, i}(c) \coloneqq \mathbbm{1}[c \simeq S_i]$ since those functions produce the subgraphs that are counted by ${\varphi}_{\mathrm{FP}}(G)[i]$ with multiplicity $1$.
+		Here the $i$-th embedding component represents the number of occurrences of a substructure $S_i$, i.e.\ ${\varphi}_{\mathrm{FP}}(G)[i] \coloneqq \mathit{count}_{S_i}(G)$ (see \cfullref{defn:related:wl-count-detect}).
+		Such a fingerprint embedding naturally is an \ac{sce} with the decomposition functions $\psi_{\mathrm{FP}, i}(G) \coloneqq \{ G[\set(s)]\, |\, s \in \mathcal{V}_G^{*} \land G[s] \equiv S_i \}$ and the multiplicity functions $\gamma_{\mathrm{FP}, i}(c) \coloneqq \mathbbm{1}[c \simeq S_i]$ since those decompositions produce the subgraphs that are counted by ${\varphi}_{\mathrm{FP}}(G)[i]$ with multiplicity $1$.
 % chktex 21
@@
 		The individual components of the \ac{wl} subtree kernel's embedding $\varphi_{\mathrm{ST}}$ represent the number of occurrences of a color $\kappa_i \in \mathcal{C}$, i.e.\ $\varphi_{\mathrm{ST}}(G)[i, t] \coloneqq \mathit{dist}_{\chi_{G,1}^{(t)}}(\kappa_i)$ as defined in \cfullref[ on ]{eq:related:wl-subtree-kernel}\footnote{
-			To avoid confusion $\kappa_i \in \mathcal{C}$ is used for colors and $c_j \in \mathcal{G}$ for constituents in this context.
+			To avoid confusion, $\kappa_i \in \mathcal{C}$ is used for colors and $c_j \in \mathcal{G}$ for constituents in this context.
 		}.
-		The occurrence of a color $\kappa_i$ at some vertex $v_j \in \mathcal{V}_G$ in the $t$-th refinement step (i.e.\ $\chi_{G,1}^{(t)}(v_j) = \kappa_i$) in turn implies that the \ac{bfs} tree $S_{j,t} \coloneqq \text{\acs*{bfs}}(G, v_j, t)$ of depth $t$ rooted at $v_j$ is isomorphic to some \ac{wl} color subtree $S_{\kappa_i}$ (i.e.\ $S_{j,t} \simeq S_{\kappa_i}$).
+		The occurrence of a color $\kappa_i$ at some vertex $v_j \in \mathcal{V}_G$ in the $t$-th refinement step (i.e.\ $\chi_{G,1}^{(t)}(v_j) = \kappa_i$) in turn implies that the \ac{bfs} tree $S_{j,t} \coloneqq \mathrm{\acs*{bfs}}(G, v_j, t)$ of depth $t$ rooted at $v_j$ is isomorphic to some \ac{wl} color subtree $S_{\kappa_i}$ (i.e.\ $S_{j,t} \simeq S_{\kappa_i}$).
 		This relation between \ac{wl} colors $\kappa_i$ and subtrees $S_{\kappa_i}$ was illustrated in \cfullref[ on ]{fig:related:wl-subtree}.
@@
 				The subtrees $S_{j,t}$ that span each constituent $c_{j,t}$ are visualized by the colored arrows that originate from the colored root vertices.
-				Because $S_{1,1}$ and $S_{2,1}$ span the same constituent $c = c_{1,1} = c_{2,1} = G[\{ v_1, v_2, v_3 \}]$, its multiplicity is $\gamma_{\varphi, \textcolor{t_red}{\texttt{B}}, 1}(c) = 2$; % chktex 25
+				Because $S_{1,1}$ and $S_{2,1}$ span the same constituent $c = c_{1,1} = c_{2,1} = G[\{ v_1, v_2, v_3 \}]$, its multiplicity is $\gamma_{\varphi, \colorlabel{t_red}{\scriptsize B}, 1}(c) = 2$; % chktex 25
 				the multiplicity of all other constituents is $1$.
@@
 
-		To fix the previous intuition we ``convert'' the subtrees $S_{j, t}$ into proper subgraph constituents via $c_{j,t} \coloneqq G[\mathcal{V}_{S_{j,t}}]$, i.e.\ the subgraphs that are spanned by the subtrees.
+		To fix the previous intuition, we ``convert'' the subtrees $S_{j, t}$ into proper subgraph constituents via $c_{j,t} \coloneqq G[\mathcal{V}_{S_{j,t}}]$, i.e.\ the subgraphs that are spanned by the subtrees.
 		The resulting decomposition functions are $\psi_{\varphi, i, t}(G) \coloneqq \{ c_{j,t}\, |\, {v_j \in \mathcal{V}_G}\,\land {S_{j,t} \simeq S_{\kappa_i}} \}$. % chktex 21
@@
 		\begin{align}
-			&\gamma_{\varphi, i, t}(c) \coloneqq {\left|\{ S = \text{\acs*{bfs}}(c, v_{\mathrm{root}}, t)\, |\, v_{\mathrm{root}} \in \mathcal{V}_c \land S \simeq S_{\kappa_i} \land \mathit{complete}(S) \}\right|}\label{eq:ltag:wlst-constituent-subtree-count} \\
+			&\gamma_{\varphi, i, t}(c) \coloneqq {\left|\{ S = \mathrm{\acs*{bfs}}(c, v_{\mathrm{root}}, t)\, |\, v_{\mathrm{root}} \in \mathcal{V}_c \land S \simeq S_{\kappa_i} \land \mathit{complete}(S) \}\right|}\label{eq:ltag:wlst-constituent-subtree-count} \\
 			&\text{with }\mathit{complete}(S) \Leftrightarrow \forall\ \text{leaf nodes $v_{\mathrm{leaf}}$ of the tree $S$}: {|\Gamma_c(v_{\mathrm{leaf}})|} = {|\Gamma_G(v_{\mathrm{leaf}})|}
@@
 		The purpose of the $\mathit{complete}(S)$ condition is to only count the \ac{bfs} subtrees of the constituent $c$ that are also \ac{bfs} subtrees of the complete graph $G$.
-		To see why this is required, note that in \cref{fig:ltag:wl1-constituents} the $c_{1,1}$/$c_{2,1}$ constituent contains the color subtree $S_{\textcolor{t_red}{\texttt{B}}}$ three times (rooted at $v_1$, $v_2$ and $v_3$) even though $G$ only contains it twice (rooted at $v_1$ and $v_2$).
+		To see why this is required, note that in \cref{fig:ltag:wl1-constituents} the $c_{1,1}$/$c_{2,1}$ constituent contains the color subtree $S_{\colorlabel{t_red}{\scriptsize B}}$ three times (rooted at $v_1$, $v_2$ and $v_3$) even though $G$ only contains it twice (rooted at $v_1$ and $v_2$).
 % chktex 25
@@
 		To determine the number of such vertex pairs via an \ac{sce} multiplicity function $\gamma_{\varphi, t_i, a_i, b_i, d_i}$, each connected pair of vertices $v_a, v_b$ and the shortest path between them must occur together in at least one constituent, otherwise a multiplicity function cannot compute whether $v_a$ and $v_b$ are in fact $d_i$ hops apart.
-		One simple decomposition which guarantees that all shortest paths are part of at least one constituent simply splits a given graph into its connected components.
+		One simple decomposition, which guarantees that all shortest paths are part of at least one constituent, simply splits a given graph into its connected components.
 		Even though such a decomposition is non-trivial since it uses at least some structural information to determine the set of constituents, the fact that any pair of connected vertices must co-occur within a single constituent means that constituents must span arbitrarily large distances within a given graph.
@@
 		Analogous to \ac{wl} subtree embeddings, those $t$-neighborhoods form the localized constituents of the $k$-LWL kernel via a \ac{bfs} over neighboring $k$-multisets.
-		Consequently \acp{svm} using the $k$-LWL kernel can be interpreted as ``\acs{lta}-like'' methods with nontrivial localized decompositions.
+		Consequently, \acp{svm} using the $k$-LWL kernel can be interpreted as ``\acs{lta}-like'' methods with nontrivial localized decompositions.
 		\badgebox[t_green]{\acs*{lta}-like}
@@
 		The only difference between the $k$-LWL and the $k$-GWL kernel is their definition of the $k$-multiset neighborhood.
-		Namely the refined color of a multiset $s \subseteq \mathcal{V}_G$ in $k$-GWL depends on the the colors of all vertices $\mathcal{V}_G$ since all multisets $s' = s \setminus \{ v \} \cup \{ u \}$ with $v \in s$ and $u \in \mathcal{V}_G$ are neighbors of $s$.
+		Namely, the refined color of a multiset $s \subseteq \mathcal{V}_G$ in $k$-GWL depends on the the colors of all vertices $\mathcal{V}_G$ since all multisets $s' = s \setminus \{ v \} \cup \{ u \}$ with $v \in s$ and $u \in \mathcal{V}_G$ are neighbors of $s$.
 		Because of those global multiset neighborhoods, all the color subtrees of a graph span the entire graph.
@@
 	\item \textbf{Convolution:}
-		First a stack of $T$ convolution layers is applied to the input feature matrix $X \in \mathbb{R}^{n \times d^{(0)}}$ where each row is a feature vector $x_i \in \mathbb{R}^{d^{(0)}}$ for the vertex $v_i \in \mathcal{V}_G$ or the vertex $k$-multiset $s_i \subseteq \mathcal{V}_G$ in case of $k$-\acsp{gnn}.
-		The convolution layers produce a convolved feature matrix $Z \in \mathbb{R}^{n \times d^{(T)}}$ where each row is a convolved feature vector $z_i \in \mathbb{R}^{d^{(\mathrm{pool})}}$.
+		First a stack of $T$ convolution layers is applied to the input feature matrix $X \in \mathbb{R}^{n \times d^{(0)}}$ where each row is a feature vector $x_i \in \mathbb{R}^{d^{(0)}}$ for the vertex $v_i \in \mathcal{V}_G$ or, in case of $k$-\acsp{gnn}, the vertex $k$-multiset $s_i \subseteq \mathcal{V}_G$.
+		The convolution layers produce a convolved feature matrix $Z \in \mathbb{R}^{n \times d^{(T)}}$, where each row is a convolved feature vector $z_i \in \mathbb{R}^{d^{(\mathrm{pool})}}$.
 	\item \textbf{Pooling:}
@@
 	\item \textbf{\ac{mlp}:}
-		Lastly the final output $y_G \in \mathcal{Y}$ is computed by applying a standard \ac{mlp} to the graph feature vector $z_G$.
+		Lastly, the final output $y_G \in \mathcal{Y}$ is computed by applying a standard \ac{mlp} to the graph feature vector $z_G$.
 \end{enumerate}
@@
 \Cref{fig:ltag:gcnn-structure} illustrates this three step architecture for vertex neighborhood convolutions like those used by \acp{gcn}~\cite{Kipf2017} or \acp{gin}~\cite{Xu2018}.
-Intuitively the convolution layers implicitly define an \ac{lta} decomposition function $\psi$ and a local evaluation function $f$ while the pooling layer roughly corresponds to an \ac{lta} aggregation function $\mathcal{A}$.
+Intuitively, the convolution layers implicitly define an \ac{lta} decomposition function $\psi$ and a local evaluation function $f$ while the pooling layer roughly corresponds to an \ac{lta} aggregation function $\mathcal{A}$.
 This intuition is now formalized.
@@
 \begin{lem}\label{lem:ltag:general-spectral-conv-nonlocality}
-	For all graphs $G$ consisting of the connected components $C_1, \dots, C_m$ there is an $\eta \in \mathbb{R}$ s.t.\ if $\hat{g}(0) = \eta$, the convolved feature vector $z_i$ of each $v_i \in \mathcal{V}_{C_l}$ with $l \in [m]$ depends on all input features $x_j$ of the vertices $v_j \in \mathcal{V}_{C_l}$ from the same component.
+	For all graphs $G$, consisting of the connected components $C_1, \dots, C_m$, there is an $\eta \in \mathbb{R}$ s.t.\ if $\hat{g}(0) = \eta$, the convolved feature vector $z_i$ of each $v_i \in \mathcal{V}_{C_l}$ with $l \in [m]$ depends on all input features $x_j$ of the vertices $v_j \in \mathcal{V}_{C_l}$ from the same component.
 \end{lem}
@@
 	}.
-	Consequently in \cref{eq:ltag:general-spectral-conv} the first $m$ summands of each $\alpha_j$ sum up to $\hat{g}(0) u_{l,j} u_{l,i} = \eta$ for all $v_i \in C_l$ with $l \in [m]$.
+	Consequently, in \cref{eq:ltag:general-spectral-conv} the first $m$ summands of each $\alpha_j$ sum up to $\hat{g}(0) u_{l,j} u_{l,i} = \eta$ for all $v_i \in C_l$ with $l \in [m]$.
 	The lemma then follows by choosing $\eta$ s.t.\ the last $n - m$ summands of every $\alpha_j$ do not sum up to $\eta$.
@@
 Analogously to single vertex convolutions, the $k$-multiset convolutions of $k$-\acp{gnn} produce local evaluation vectors $z_i$ for the $T$-neighborhood of each $k$-multiset $s_i \subseteq \mathcal{V}_G$.
-By \cfullref[ on ]{eq:related:kgnn-layer} the constituent $c_i$ spanned by the vertices that have an influence on $z_i$ can thus be written as $c_i = G\left[ \bigcup_{v \in s_i} \Gamma_G^T(v) \right]$.
+By \cfullref[ on ]{eq:related:kgnn-layer}, the constituent $c_i$ spanned by the vertices that have an influence on $z_i$ can thus be written as $c_i = G\left[ \bigcup_{v \in s_i} \Gamma_G^T(v) \right]$.
 This constituent definition is identical to that used in the \ac{lta} formulation of $k$-LWL \acp{svm} (see \cfullref{itm:ltag:klwl-lta-formulation}).
@@
 Both, the single vertex neighborhood convolution as well as the vertex $k$-multiset neighborhood convolution produce vectors $z_i \in \mathbb{R}^{d^{(T)}}$, which are each based on the constituent $c_i$ spanned by the $T$-neighborhood of some vertex $v_i$ or vertex $k$-multiset $s_i$.
-Consequently we define the \ac{lta} decomposition function of such neighborhood convolution \acp{gcnn} as $\psi(G) \coloneqq {\{ c_i \}}_{i=1}^n$.
+Consequently, we define the \ac{lta} decomposition function of such neighborhood convolution \acp{gcnn} as $\psi(G) \coloneqq {\{ c_i \}}_{i=1}^n$.
 
@@
 If there is a translation function $\tau_{1}: \mathbb{R}^{d^{(T)}} \to \mathcal{Y} \times \mathbb{R}_{\geq 0}$, the convolved vectors $z_i \in \mathbb{R}^{d^{(T)}}$ can be interpreted as constituent evaluations.
-One trivial example for such a translation function would be the identity $\tau_{1}(z) = z$ for a graph regression problem with the score domain $\mathcal{Y} = [0, 1]$ and where the last convolution layer uses a logistic activation function with two output dimensions s.t.\ $z \in {[0,1]}^2$.
-Assuming that there is a translation function $\tau_{1}$ we can define a multi-score evaluation function $f^{*}(c) \coloneqq {\ldblbrace \tau_{1}(z_i)\, |\, {i \in [n]} \land {c = c_i} \rdblbrace} = {\ldblbrace \tau_{1}(\mathit{Conv}(c)[v_i])\, |\, v_i \in \mathit{root}(c) \rdblbrace}$.
+One trivial example for such a translation function would be the identity $\tau_{1}(z) = z$ for a graph regression problem with the score domain $\mathcal{Y} = [0, 1]$ and where the last convolution layer uses a logistic activation function with two output dimensions s.t.\ $z \in {[0,1]} \times \mathbb{R}$.
+Assuming that there is a translation function $\tau_{1}$, we can define a multi-score evaluation function $f^{*}(c) \coloneqq {\ldblbrace \tau_{1}(z_i)\, |\, {i \in [n]} \land {c = c_i} \rdblbrace} = {\ldblbrace \tau_{1}(\mathit{Conv}(c)[v_i])\, |\, v_i \in \mathit{root}(c) \rdblbrace}$.
 The function $\mathit{root}(c) = \{ v_i\, |\, {i \in [n]} \land {c = c_i} \}$ returns the set of root vertices $v_i$ (or root vertex $k$-multisets in case of $k$-\acs{gnn}) whose \ac{bfs} subtree of depth $T$ in $G$ spans exactly the vertices in $c$.
-To compute $\mathit{root}(c)$ given only a single constituent $c$ and not the entire graph $G$, it has to be determined which \ac{bfs} subtrees of $c$ are also \ac{bfs} subtrees of $G$.
+To compute $\mathit{root}(c)$, given only a single constituent $c$ and not the entire graph $G$, it has to be determined which \ac{bfs} subtrees of $c$ are also \ac{bfs} subtrees of $G$.
 We already saw how to do this in the \ac{lta} formulation of \ac{wl} subtree kernel \acp{svm} (see $\mathit{complete}$ in \cfullref{eq:ltag:wlst-constituent-subtree-count}).
@@
 
-Based on the decomposition function $\psi: \mathcal{G} \to \mathcal{P}(\mathcal{G})$ and the multi-score evaluation function $f^{*}: \mathcal{G} \to {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*}$ which are implicitly defined by the stack of convolution layers $\mathit{Conv}$ of a given \ac{gcnn}, we now complete our \ac{gcnn} \ac{lta} formulation by defining an evaluation function $f: \mathcal{G} \to \mathcal{Y} \times \mathbb{R}_{\geq 0}$ as well as an aggregation function $\mathcal{A}: {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*} \to \mathcal{Y}$ based on the pooling layer $\mathit{Pool}: \mathbb{R}^{n \times d^{(T)}} \to \mathbb{R}^{d^{(\mathrm{pool})}}$.
+Based on the decomposition function $\psi: \mathcal{G} \to \mathcal{P}(\mathcal{G})$ and the multi-score evaluation function $f^{*}: \mathcal{G} \to {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*}$, which are implicitly defined by the stack of convolution layers $\mathit{Conv}$ of a given \ac{gcnn}, we now complete our \ac{gcnn} \ac{lta} formulation by defining an evaluation function $f: \mathcal{G} \to \mathcal{Y} \times \mathbb{R}_{\geq 0}$ as well as an aggregation function $\mathcal{A}: {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*} \to \mathcal{Y}$ based on the pooling layer $\mathit{Pool}: \mathbb{R}^{n \times d^{(T)}} \to \mathbb{R}^{d^{(\mathrm{pool})}}$.
 \begin{defn}\label{defn:ltag:assoc-algebra}
@@
 \end{proof}
-Note that \cref{thm:ltag:gcnn-ltag-formulation} requires \acp{gcnn} without a final \ac{mlp} which is unlike the architecture proposed in the \ac{gcn}~\cite{Kipf2017} and \ac{gin}~\cite{Xu2018} papers.
+Note that \cref{thm:ltag:gcnn-ltag-formulation} requires \acp{gcnn} without a final \ac{mlp}, which is unlike the architecture proposed in the \ac{gcn}~\cite{Kipf2017} and \ac{gin}~\cite{Xu2018} papers.
 The reason for this is that the composition $\mathrm{\acs*{mlp}} \circ \mathit{Pool}$ cannot be guaranteed to be homomorphic to a weighted aggregation function $\mathcal{A}$ due to the well-known universal approximation theorem~\cite{Hornik1991}.
@@
 		\end{alignat*}
-		From an \ac{lta} perspective the non-associativity of SortPooling and \ac{sagpool} implies that, if they were used in a \ac{gcnn}, it could happen that some subtree evaluations $(y_i, w_i) \in f^{*}(c)$ of a single constituent $c$ are considered by the aggregation function $\mathcal{A}$ while some are ignored depending on the subtree evaluations of other constituents, i.e.\ the constituent evaluation $f(c)$ could then depend on other constituents.
+		From an \ac{lta} perspective, the non-associativity of SortPooling and \ac{sagpool} implies that, if they were used in a \ac{gcnn}, it could happen that some subtree evaluations $(y_i, w_i) \in f^{*}(c)$ of a single constituent $c$ are considered by the aggregation function $\mathcal{A}$ while some are ignored depending on the subtree evaluations of other constituents, i.e.\ the constituent evaluation $f(c)$ could then depend on other constituents.
 		This is forbidden in our definition of \ac{lta}.

### 2020-04-14 thesis/content/chapter-introduction.tex
corrections

@@
 		The previous work on \ac{lta} provides only a narrow definition of \ac{lta} for unstructured multiset inputs.
-		The question what its essential characteristics are, on a more general level, has not yet been considered.
+		The question, what its essential characteristics are on a more general level, has not yet been considered.
 		Therefore we will provide the terminology and definitions to formally capture what constitutes an \ac{lta} method as opposed to a non-\ac{lta} method.
@@
 	\item \textbf{Definition of a novel \acs{lta}-inspired \ac{gcr} method:}
-		As a follow-up question to the previous one, we ask to which extent the existing \ac{gcr} methods are explainable and what their specific shortcomings are.
+		As a follow-up question to the previous one, we ask to which extent the existing \ac{gcr} methods are explainable via the \ac{lta} perspective and what their specific shortcomings are.
 		Based on those shortcomings we will propose a novel \acs{lta}-inspired approach to overcome them.
@@
 \paragraph{\Cref{sec:related}: \nameref{sec:related}}
-In order to answer our three research questions, we begin by with on overview of the previous work regarding \ac{lta} as well as \ac{gcr}.
+In order to answer our three research questions, we begin with an overview of the previous work regarding \ac{lta} as well as \ac{gcr}.
 We begin with a brief description of \ac{lta}.
@@
 In the first step a formal definition of \ac{lta} is provided, which answers the question of what constitutes an \ac{lta} method.
-This definition is then used to analyze all \ac{gcr} approaches described in \cref{sec:related} for their compatibility with \ac{lta}.
+This definition is then used to analyze all \ac{gcr} approaches described in \cref{sec:related} for their compatibility with \ac{lta}, which answers the second research question.
 
 \paragraph{\Cref{sec:ltd}: \nameref{sec:ltd}}
-Based on the \ac{lta} perspective on \ac{gcr} provided in \cref{sec:ltag}, we find that the existing approaches are all limited by their underlying solution to the so-called \ac*{ltd} problem.
+Based on the \ac{lta} perspective on \ac{gcr} provided in \cref{sec:ltag}, we find that the existing \ac{gnn} approaches are all limited by their underlying solution to the so-called \ac*{ltd} problem.
 To overcome this limitation, and thereby answer our third research question, we propose the general idea of edge-filtered graph convolutions.
-To realize this general idea in a specific method, we then proceed by proposing a novel type of \ac*{gcnn}, the so-called \textit{2-\acs*{wl}-\acs{gnn}}, which can serve as the foundation for an edge-filtering-based solution to the \acs*{ltd} problem.
+To realize this general idea in a specific method, we then proceed by proposing a novel type of \ac*{gcnn}, the so-called \textit{2-\acs*{wl}-\acs{gnn}}, which can serve as the foundation for an edge-filtering based solution to the \acs*{ltd} problem.
 Apart from its uses in the context of \ac{lta}/\acs*{ltd}, we show that 2-\acs*{wl}-\acsp{gnn} have additional theoretical advantages over existing approaches;

### 2020-04-14 thesis/content/chapter-ltd.tex
improved intro and conclusion
corrections

@@
 
-As per our definition of \ac{lta} from \cref{sec:ltag:definition}, the quality of an \ac{lta} formulation is determined by its chosen solution for the \ac{ltd} problem.
+As per our definition of \ac{lta} from \cref{sec:ltag:definition}, the quality of an \ac{lta} formulation is determined by its chosen solution to the \ac{ltd} problem.
 Since a comprehensive analysis of this problem would be beyond the scope of this thesis, we focus on the relation between \ac{ltd} and \acp{gcnn}.
@@
 		A graph decomposition obtained via dynamic edge filtering which cannot be modeled by edge prefiltering.
-		Depending on the \ac{bfs} root node $v_{\mathrm{root}}$ different edges are removed.
+		Depending on the \ac{bfs} root node $v_{\mathrm{root}}$, different edges are removed.
 		For $v_{\mathrm{root}} \in \textcolor{t_blue}{\{ v_3, v_4 \}}$ the edges $\{ (v_3, v_1), (v_3, v_2), (v_4, v_5), (v_4, v_6) \}$ are filtered out while for $v_{\mathrm{root}} \in \textcolor{t_red}{\{ v_1, v_2, v_5, v_6 \}}$ the edge $(v_3, v_4)$ is removed.
@@
 % chktex 25
-In 2-\acs{wl} on the other hand, the neighbors of $\textcolor{t_blue}{e_{ij}}$ are the edge pairs ${\left\{(\textcolor{t_red}{e_{il}}, \textcolor{t_darkgreen}{e_{lj}})\right\}}_{v_l \in \mathcal{V}_G}$, i.e.\ all possible paths of length two that start at $v_i$ and end at $v_j$.
+In 2-\acs{wl} on the other hand, the neighbors of $\textcolor{t_blue}{e_{ij}}$ are the edge pairs ${\left\{(\textcolor{t_red}{e_{il}}, \textcolor{t_darkgreen}{e_{lj}})\right\}}_{v_l \in \mathcal{V}_G}$, i.e.\ all possible walks of length two that start at $v_i$ and end at $v_j$.
 % chktex 21 chktex 25
@@
 \begin{proof}
-	By \cref{defn:ltd:edge-graph} it holds that $\forall e_{ij} \in \mathcal{E}_G: {\Gamma_G(v_i) \cup \Gamma_G(v_j)} = \Gamma_{G^{\mathcal{E}}}(e_{ij})$.
+	By \cref{defn:ltd:edge-graph}, it holds that $\forall e_{ij} \in \mathcal{E}_G: {\Gamma_G(v_i) \cup \Gamma_G(v_j)} = \Gamma_{G^{\mathcal{E}}}(e_{ij})$.
 	Therefore the 2-\acs{gnn} convolution operation defined in \cfullref[ on ]{eq:related:kgnn-layer} can be rewritten as a vertex neighborhood convolution operator
@@
 
-As described in \cref{sec:related:character:wl} the discriminative power of a model is, by itself, not necessarily relevant for real-world graph datasets.
+As described in \cref{sec:related:character:wl}, the discriminative power of a model is, by itself, not necessarily relevant for real-world \ac{gcr} problems.
 In \cfullref[ on ]{prop:related:wl2-cycle-count} we saw however that 2-\acs{wl} not only has a higher discriminative power than 1-\acs{wl} but also that it is able to count the number of $m$-cycles in a given graph for all $m \leq 7$.
 This is relevant because cycle counts are a commonly used metric in real-world domains such as social network and molecular structure analysis~\cite{Milo2002}\cite{Newman2003}\cite{Welser2007}\cite{Adamson1973}\cite{Kekule1866}.
-To conclude this section we now show that 2-\acsp{gnn} not only have a lower discriminative power than 2-\acs{wl} but are also unable to detect cycles.
-\begin{prop}
+To conclude this section, we now show that 2-\acsp{gnn} not only have a lower discriminative power than 2-\acs{wl} but are also unable to detect cycles.
+\begin{prop}\label{prop:ltd:2gnn-cycle-limit}
 	2-\acsp{gnn} cannot detect $m$-cycles for all $m \geq 3$.
@@
 	We define $c_3 \coloneqq \frac{n}{3}$ and $c_m \coloneqq \frac{n}{m}$.
-	Based on that we define the following two graphs:
+	Based on that, we define the following two graphs:
 	Let $G_3$ be a graph consisting of $c_3$ disconnected cycles of length $3$, analogously let $G_m$ be a graph consisting of $c_m$ disconnected cycles of length $m$.
@@
 	Let us assume that $h_2$ is able to detect cycles of length $3$, i.e.\ triangles.
-	By \cfullref[ on ]{defn:related:wl-count-detect} this would imply that there is a function $g: \mathcal{Y} \to \{ 0, 1\}$ s.t.\ $g(h_2(G_3)) = g(y) = 1$ and $g(h_2(G_m)) = g(y) = 0\ \lightning$.
+	By \cfullref[ on ]{defn:related:wl-count-detect}, this would imply that there is a function $g: \mathcal{Y} \to \{ 0, 1\}$ s.t.\ $g(h_2(G_3)) = g(y) = 1$ and $g(h_2(G_m)) = g(y) = 0\ \lightning$.
 	Conversely, assuming that $h_2$ is able to detect cycles of length $m > 3$, we obtain the contradiction $g(h_2(G_3)) = g(y) = 0$  and $g(h_2(G_m)) = g(y) = 1\ \lightning$.
@@
 In this section we compared 2-\acsp{gnn} with the 2-\acs{wl} algorithm and found that they have a significantly lower discriminative and computational power than 2-\acs{wl}.
-Motivated by those limitations we describe a novel graph convolution operator which is closer to 2-\acs{wl} in the following section.
+Motivated by those limitations we describe a novel graph convolution operator, which is closer to 2-\acs{wl}, in the following section.
 
@@
 This will be done in three steps:
-\begin{enumerate*}
+\begin{enumerate*}[label={\circled{\small\arabic*}}]
 	\item We begin by formally defining the 2-\acs{wl} convolution operator.
-	\item Then the discriminative and computational power of this operator are analyzed.
-	\item Lastly we describe how 2-\acs{wl} convolutions can be implemented efficiently on a modern \ac{gpgpu}.
+	\item Then the discriminative and computational power of this operator is analyzed.
+	\item Lastly, we describe how 2-\acs{wl} convolutions can be implemented efficiently on a modern \ac{gpgpu}.
 \end{enumerate*}
@@
 		Since we assume that graphs are undirected, $e_{ij}$ and $e_{ji}$ have identical feature vectors $x[e_{ij}] = x[e_{ji}] \in \mathcal{X}_{\mathcal{E}}$ and the same 2-\acs{wl} neighborhood.
-		Instead of refining/convolving the feature vectors of 2-tuples $(v_i, v_j)$ we can therefore refine/convolve the feature vectors of 2-multisets $\ldblbrace v_i, v_j \rdblbrace$.
+		Instead of refining/convolving the feature vectors of 2-tuples $(v_i, v_j)$, we can therefore refine/convolve the feature vectors of 2-multisets $\ldblbrace v_i, v_j \rdblbrace$.
 		This halves the number of feature vectors without affecting the discriminative or computational power of the 2-\acs{wl} convolution operator.
@@
 		Using the 2-multiset simplification, the original 2-\acs{wl} algorithm refines the color of all multisets $e_{ij} \in \mathcal{V}_G^2$ by hashing its current color and the colors of all neighbors ${\{ \ldblbrace e_{il}, e_{lj} \rdblbrace \}}_{v_l \in \mathcal{V}_G}$.
-		This means that the time complexity of a single refinement step is $\mathcal{O}(n^3)$ for $n \coloneqq \left| \mathcal{V}_G \right|$ which quickly becomes infeasible for large graphs.
+		This means that the time complexity of a single refinement step is $\mathcal{O}(n^3)$ for $n \coloneqq \left| \mathcal{V}_G \right|$, which quickly becomes infeasible for large graphs.
 
@@
 		\end{defn}
-		For the neighborhood radius $r = 1$ only the self-loop edges ${\{ e_{ii} \}}_{v_i \in \mathcal{V}_G}$ and the edges $\mathcal{E}_{G}$ are considered;
-		for $r > 1$ edges between indirectly connected vertices are considered as well.
+		For the neighborhood radius $r = 1$, only the self-loop edges ${\{ e_{ii} \}}_{v_i \in \mathcal{V}_G}$ and the edges $\mathcal{E}_{G}$ are considered;
+		for $r > 1$, edges between indirectly connected vertices are considered as well.
 		\Cref{fig:ltd:neighborhood-radius} illustrates this for varying neighborhood radii.
@@
 				The self-loop edge at each of the vertices is not explicitly shown.
-				For $r = 3$ all possible edges between the six vertices will be considered just as in the original 2-\acs{wl} algorithm.
+				For $r = 3$ all possible edges between the six vertices will be considered, just as in the original 2-\acs{wl} algorithm.
 			}\label{fig:ltd:neighborhood-radius}
@@
 		Let us now consider what the reduced number of considered edges and the reduced number of edge neighbors implies for the runtime of a refinement step.
-		If $G$ is a sparse graph with the maximum vertex degree $d \coloneqq \max_{v \in \mathcal{V}_G} \left|\Gamma_G(v)\right|$, the number of considered edges is bounded by $\mathcal{O}(n d^r)$ where each edge has at most $\mathcal{O}(d^r)$ neighbors.
+		If $G$ is a sparse graph with the maximum vertex degree $d \coloneqq \max_{v \in \mathcal{V}_G} \left|\Gamma_G(v)\right|$, the number of considered edges is bounded by $\mathcal{O}(n d^r)$, where each edge has at most $\mathcal{O}(d^r)$ neighbors.
 		Consequently the time complexity of a refinement step becomes $\mathcal{O}(n d^{2r})$ which is a significant improvement over the $\mathcal{O}(n^3)$ bound of a full 2-\acs{wl} refinement step (assuming $d \ll n$).
 \end{enumerate}
-Based on the 2-multiset and the neighborhood localization simplifications we can now define the 2-\ac{wl} convolution operator and the corresponding so-called \textit{2-\acs{wl}-\acs{gnn}}.
+Based on the 2-multiset and the neighborhood localization simplifications, we now define the 2-\ac{wl} convolution operator and the corresponding so-called \textit{2-\acs{wl}-\acs{gnn}}.
 \begin{defn}\label{defn:ltd:wl2-conv-init}
 	The \textit{initial feature matrix $Z_G^{(0)}$} of the 2-\acs{wl} convolution operator with the neighborhood radius $r \in \mathbb{N}$ contains both the vertex features $x[v_i] \in \mathcal{X}_{\mathcal{V}} = \mathbb{R}^{d_{\mathcal{V}}}$ as well as the edge features $x[e_{ij}] \in \mathcal{X}_{\mathcal{E}} = \mathbb{R}^{d_{\mathcal{E}}}$ of a given graph $G$.
-	More specifically $Z_G^{(0)} \in \mathbb{R}^{{|\mathcal{E}_{G^r}|} \times (d_{\mathcal{V}} + d_{\mathcal{E}})}$ assigns a row vector $Z_G^{(0)}[e_{ij}]$ to all edges $e_{ij} \in \mathcal{E}_{G^r}$.
+	More specifically, $Z_G^{(0)} \in \mathbb{R}^{{|\mathcal{E}_{G^r}|} \times (d_{\mathcal{V}} + d_{\mathcal{E}})}$ assigns a row vector $Z_G^{(0)}[e_{ij}]$ to all edges $e_{ij} \in \mathcal{E}_{G^r}$.
 	Those initial edge feature vectors are defined by the following vector concatenation:
@@
 	\item \textbf{Simulation of \acp{mlp}:}
-		By choosing $W_{\mathrm{F}}^{(t)} = \mathbf{0}$ the convolution behaves like a regular fully connected layer.
+		By choosing $W_{\mathrm{F}}^{(t)} = \mathbf{0}$, the convolution behaves like a regular fully connected layer.
 		A stack of 2-\acs{wl} convolution layers can therefore simulate arbitrary \acp{mlp}.
@@
 		A 2-\acs{gnn} use a similar formulation but leaves out the inner nonlinearity $\sigma_{\Gamma}$;
-		as we saw in \cref{fig:ltd:2gnn-2wl-diff} this causes 2-\acsp{gnn} to lose the edge pair information due to the commutativity and associativity of $+$.
+		as we saw in \cref{fig:ltd:2gnn-2wl-diff}, this causes 2-\acsp{gnn} to lose the edge pair information due to the commutativity and associativity of $+$.
 		2-\acs{wl}-\acsp{gnn} do not generally have this problem because $\kappa^{(t)}(\colorlabel{t_blue}{A}, \ldblbrace \colorlabel{t_red}{B}, \colorlabel{t_red}{B} \rdblbrace) + \kappa^{(t)}(\colorlabel{t_blue}{A}, \ldblbrace \colorlabel{t_darkgreen}{C}, \colorlabel{t_darkgreen}{C} \rdblbrace) \neq \kappa^{(t)}(\colorlabel{t_blue}{A}, \ldblbrace \colorlabel{t_red}{B}, \colorlabel{t_darkgreen}{C} \rdblbrace) + \kappa^{(t)}(\colorlabel{t_blue}{A}, \ldblbrace \colorlabel{t_red}{B}, \colorlabel{t_darkgreen}{C} \rdblbrace)$ if $\sigma_{\Gamma}$ is chosen to be a nonlinear activation function.
@@
 		We can interpret $z_{ij} W_{\mathrm{F}}^{(t)} \in \mathbb{R}^{d^{(t)}}$ as a row vector of feature dimension weights.
-		The feature dimensions of neighbors are rescaled via those weights which allows the model to aggregate different types of neighbors depending on $z_{ij}$.
+		The feature dimensions of neighbors are rescaled via those weights, which allows the model to aggregate different types of neighbors depending on $z_{ij}$.
 \end{enumerate}
@@
 
-In this section we will analyze the power of \acp{gnn} which use the 2-\acs{wl} convolution operator that was just defined.
+In this section we will now analyze the power of \acp{gnn} which use the 2-\acs{wl} convolution operator that was just defined.
 Our goal is to show that such 2-\acs{wl}-\acsp{gnn} have a strictly larger discriminative power than 1-\acs{wl}.
@@
 \end{defn}
-\Cref{defn:ltd:vert-conv} includes \acp{gcn}~\cite{Kipf2017} where the \ac{mlp} only consists of a single layer and the weights are $w_{ij} = {\left(\left| \Gamma_G(v_i) \right| + 1\right)}^{-\frac{1}{2}}{\left(\left| \Gamma_G(v_j) \right| + 1\right)}^{-\frac{1}{2}}$ (see \cfullref{eq:related:gcn-wl1}).
+\Cref{defn:ltd:vert-conv} includes \acp{gcn}~\cite{Kipf2017}, where the \ac{mlp} only consists of a single layer and the weights are $w_{ij} = {\left(\left| \Gamma_G(v_i) \right| + 1\right)}^{-\frac{1}{2}}{\left(\left| \Gamma_G(v_j) \right| + 1\right)}^{-\frac{1}{2}}$ (see \cfullref{eq:related:gcn-wl1}).
 \Acp{gin}~\cite{Xu2018} also trivially satisfy the definition (see \cfullref{eq:related:gin-layer}).
@@
 	Let $G \in \mathcal{G}$ be an arbitrary input graph with $n \coloneqq \left|\mathcal{V}_G\right|$ and $m \coloneqq n + \left|\mathcal{E}_G\right|$.
-	By definition $h_1$ is a \ac{gnn} of the form $\mathit{Pool}_1(\mathit{Conv}_1(G))$ where $\mathit{Conv}_1$ is a stack of $T$ weighted vertex neighborhood sum convolutions ${\left\{ S^{(t)}: \mathbb{R}^{n \times d^{(t-1)}} \to \mathbb{R}^{n \times d^{(t)}} \right\}}_{t=1}^T$ with each corresponding $\mathrm{\acs*{mlp}}^{(t)}$ having $K$ layers.
+	By definition, $h_1$ is a \ac{gnn} of the form $\mathit{Pool}_1(\mathit{Conv}_1(G))$, where $\mathit{Conv}_1$ is a stack of $T$ weighted vertex neighborhood sum convolutions ${\left\{ S^{(t)}: \mathbb{R}^{n \times d^{(t-1)}} \to \mathbb{R}^{n \times d^{(t)}} \right\}}_{t=1}^T$ with each corresponding $\mathrm{\acs*{mlp}}^{(t)}$ having $K$ layers.
 % chktex 21
@@
 
-	Let $h_2$ be a \ac{gnn} of the form $\mathit{Pool}_2(\mathit{Conv}_2(G))$ where $\mathit{Conv}_2$ is a stack of $(2 + K) T$ 2-\acs{wl} convolution layers ${\left\{ S^{(t, k)}: \mathbb{R}^{m \times d^{(t,k-1)}} \to \mathbb{R}^{m \times d^{(t,k)}} \right\}}_{(t, k) \in [T] \times [2+K]}$ with the neighborhood radius $r \coloneqq 1$.
+	Let $h_2$ be a \ac{gnn} of the form $\mathit{Pool}_2(\mathit{Conv}_2(G))$, where $\mathit{Conv}_2$ is a stack of $(2 + K) T$ 2-\acs{wl} convolution layers ${\left\{ S^{(t, k)}: \mathbb{R}^{m \times d^{(t,k-1)}} \to \mathbb{R}^{m \times d^{(t,k)}} \right\}}_{(t, k) \in [T] \times [2+K]}$ with the neighborhood radius $r \coloneqq 1$.
 % chktex 21
@@
 % chktex 21
-	Intuitively $\mathit{Conv}_2$ simulates each layer of $\mathit{Conv}_1$ via a stack of $2 + K$ 2-\acs{wl} convolution layers.
+	Intuitively, $\mathit{Conv}_2$ simulates each layer of $\mathit{Conv}_1$ via a stack of $2 + K$ 2-\acs{wl} convolution layers.
 	This is illustrated in \cref{fig:ltd:wl2-simulation}.
@@
 	\Cref{thm:ltd:wl2-simulation} then follows if there is a function $\varphi$ s.t.\ $\forall v_i \in \mathcal{V}_G: \mathit{Conv}_1(G)[v_i] = \varphi(\mathit{Conv}_2(G)[e_{ii}])$ and $\forall e_{ij} \in \mathcal{E}_G: \varphi(\mathit{Conv}_2(G)[e_{ij}]) = \mathtt{nil}$.
-	To guarantee that there is such a function $\varphi$ we now inductively prove the following three invariants which have to hold for all $t \in \{ 0, \dots, T \}$:
+	To guarantee that there is such a function $\varphi$, we now inductively prove the following three invariants which have to hold for all $t \in \{ 0, \dots, T \}$:
 	\begin{enumerate}[label={(P\arabic*)}]
@@
 	Assuming the invariants hold for $t - 1$ we now show that they also hold for $t$.
-	As illustrated in \cref{fig:ltd:wl2-simulation} the layers $S^{(t, 1)}$ and $S^{(t, 2)}$ should compute the weighted vertex neighborhood sums
+	As illustrated in \cref{fig:ltd:wl2-simulation}, the layers $S^{(t, 1)}$ and $S^{(t, 2)}$ should compute the weighted vertex neighborhood sums
 	\begin{align*}
@@
 	\end{gather*}
-	The vertex neighborhood summation is completed via $S^{(t, 2)}$ which is defined by
+	The vertex neighborhood summation is completed via $S^{(t, 2)}$, which is defined by
 	\begin{gather*}\allowdisplaybreaks
@@
 	Using the two layers $S^{(t, 1)}$ and $S^{(t, 2)}$ that we just defined, the weighted vertex neighborhood sum for all $v_i \in \mathcal{V}_G$ is contained in $Z^{(t,2)}[e_{ii}]$.
-	Additionally for all $e_{ij} \in \mathcal{E}_{G^1}$, the indicators ${Z^{(t,2)}[e_{ij}]}_1 = \mathbbm{1}[i = j]$ and the weights ${Z^{(t,2)}[e_{ij}]}_{d^{(t)} + 2} = w_{ij}$ are preserved.
+	Additionally, for all $e_{ij} \in \mathcal{E}_{G^1}$, the indicators ${Z^{(t,2)}[e_{ij}]}_1 = \mathbbm{1}[i = j]$ and the weights ${Z^{(t,2)}[e_{ij}]}_{d^{(t)} + 2} = w_{ij}$ are preserved.
 	This means that \cref{inv:ltd:wl2-simulation:indicator,inv:ltd:wl2-simulation:weight} are satisfied after $S^{(t, 2)}$.
@@
 
-To complete our analysis of the expressive power of 2-\acs{wl}-\acsp{gnn} we now show that they are not just as powerful as 1-\acs{wl} but in fact more powerful than 1-\acs{wl}.
+To complete our analysis of the expressive power of 2-\acs{wl}-\acsp{gnn}, we now show that they are not just as powerful as 1-\acs{wl} but in fact more powerful than 1-\acs{wl}.
 \begin{prop}\label{cor:ltd:wl2-gnn-regular}
@@
 	Let $h_2 = \mathit{Pool} \circ S$ be a 2-\acs{wl}-\acs{gnn} with the neighborhood radius $r = 1$ which consists of a single 2-\acs{wl} convolution layer $S: \mathbb{R}^{* \times 2} \to \mathbb{R}^{* \times 1}$ and the pooling layer $\mathit{Pool} = \min$.
-	In accordance with \cref{defn:ltd:wl2-conv-init} we set the initial feature vectors of the vertices $v_i$ of $G$ and $H$ to $Z^{(0)}[e_{ii}] \coloneqq (1, 0)$ and the initial feature vectors of their edges $e_{ij}$ to $Z^{(0)}[e_{ij}] \coloneqq (0, 1)$.
+	In accordance with \cref{defn:ltd:wl2-conv-init}, we set the initial feature vectors of the vertices $v_i$ of $G$ and $H$ to $Z^{(0)}[e_{ii}] \coloneqq (1, 0)$ and the initial feature vectors of their edges $e_{ij}$ to $Z^{(0)}[e_{ij}] \coloneqq (0, 1)$.
 	Let the weight matrices of $S$ be $W_{\mathrm{L}} \coloneqq \mathbf{0}$ and $W_{\mathrm{F}} = W_{\Gamma} \coloneqq \begin{pmatrix} 1\\1 \end{pmatrix}$.
@@
 % chktex 25
-	However for the non-self-loops of $G^1$ and $H^1$, i.e.\ the edges of $G$ and $H$, we get differing convolved feature vectors.
+	However, for the non-self-loops of $G^1$ and $H^1$, i.e.\ the edges of $G$ and $H$, we get differing convolved feature vectors.
 	The 2-\acs{wl} neighbors of $e_{ij} \in \mathcal{E}_G$ are $\{ \ldblbrace e_{ii},  e_{ij} \rdblbrace,\allowbreak \ldblbrace e_{ij}, e_{jj} \rdblbrace \}$.
 	The 2-\acs{wl} neighbors of $e_{ij}' \in \mathcal{E}_H$ are $\{ \ldblbrace e_{ii}',  e_{ij}' \rdblbrace,\allowbreak \ldblbrace e_{ij}',  e_{jj}' \rdblbrace,\allowbreak \ldblbrace e_{il}',  e_{lj}' \rdblbrace \}$ where $v_l' \in \mathcal{V}_H$ is the common neighbor of $v_i'$ and $v_j'$.
-	The different neighborhood sizes of the edges of $G$ and $H$ imply that $\forall e_{ij} \in \mathcal{E}_G: Z^{(1)}[e_{ij}] = 4$ while $\forall e_{ij}' \in \mathcal{E}_H: Z^{(1)}[e_{ij}] = 6$.
+	The different neighborhood sizes of the edges of $G$ and $H$ imply that $\forall e_{ij} \in \mathcal{E}_G: Z^{(1)}[e_{ij}] = 4$, while $\forall e_{ij}' \in \mathcal{E}_H: Z^{(1)}[e_{ij}] = 6$.
 	Thus $h_2(G) = \min\{ 4, 6 \} \neq \min\{ 6, 6 \} = h_2(H)$, which concludes the proof. % chktex 21
@@
 It is easy to see that 2-\acs{wl}-\acsp{gnn} \textit{generally} cannot have the same power as 2-\acs{wl} due to the neighborhood localization simplification;
-e.g.\ for a small neighborhood radius of $r = 1$, nonexistent edges $e_{ij} \notin \mathcal{E}_G$ do not have a feature vector which implies that the proof of 2-\acs{wl}'s $m$-cycle counting ability no longer holds for $m \geq 4$ (see \cfullref{prop:related:wl2-cycle-count}).
+e.g.\ for a small neighborhood radius of $r = 1$, nonexistent edges $e_{ij} \notin \mathcal{E}_G$ do not have a feature vector, which implies that the proof of 2-\acs{wl}'s $m$-cycle counting ability no longer holds for $m \geq 4$ (see \cfullref{prop:related:wl2-cycle-count}).
 We leave a more thorough discussion of the relation between 2-\acs{wl}-\acsp{gnn} and 2-\acs{wl} for future work.
@@
 For this purpose we propose a sparse 2-\acs{wl} graph representation which is inspired by the coordinate list adjacency format described by \citet{Fey2019}.
-Given a neighborhood radius $r$ we encode a graph $G$ using the following two matrices:
+Given a neighborhood radius $r$, we encode a graph $G$ using the following two matrices:
 \begin{enumerate}
@@
 		Each neighbor row is a vector $(r_{\mathrm{L}}, r_{\Gamma,1}, r_{\Gamma,2}) \in {[m]}^3$ of three index pointers to rows in $Z_G^{(0)}$.
-		$r_{\mathrm{L}}$ points to the row index of the feature vector of $e_{ij}$ while $r_{\Gamma,1}$ and $r_{\Gamma,2}$ point to the indices of $e_{il}$ and $e_{lj}$ respectively.
+		$r_{\mathrm{L}}$ points to the row index of the feature vector of $e_{ij}$, while $r_{\Gamma,1}$ and $r_{\Gamma,2}$ point to the indices of $e_{il}$ and $e_{lj}$ respectively.
 		We will refer to the three column vectors of $R_G$ as $R_{G, \mathrm{L}}$, $R_{G, \Gamma, 1}$ and $R_{G, \Gamma, 2}$.

### 2020-04-14 thesis/content/chapter-related-work.tex
corrections

@@
 
-Before combining \ac{lta} and \ac{gcr}, as described in \cref{sec:intro:questions}, we first give an overview of the state-of-the-art in both fields of research.
+Before combining \ac{lta} and \ac{gcr} as described in \cref{sec:intro:questions}, we first give an overview of the state-of-the-art in both fields of research.
 This is done in three steps:
@@
 There an input instance is understood as a composition $C = \ldblbrace c_1, \dots, c_n \rdblbrace$ of so-called constituents, i.e.\ as a variable-size multiset (denoted as $\ldblbrace \cdot \rdblbrace$).
-The assumption in \ac{lta} problems is that for all constituents $c_i \in C$ a local score $y_i \in \mathcal{Y}$ is either given or computable.
+The assumption in \ac{lta} problems is that, for all constituents $c_i \in C$, a local score $y_i \in \mathcal{Y}$ is either given or computable.
 The set of those local scores should be indicative of the overall score $y \in \mathcal{Y}$ of the composition $C$.
@@
 \end{figure}
-Overall \ac{lta} can be understood as the joint problem of learning the aggregation function $\mathcal{A}$ and the local score derivation function $f$.
+Overall, \ac{lta} can be understood as the joint problem of learning the aggregation function $\mathcal{A}$ and the local score derivation function $f$.
 Two main approaches to represent the aggregation function in \ac{lta} problems have been explored.
@@
 
-Since t-norms and t-conorms are commutative and associative they can also be applied to non-empty sets of arbitrary size, i.e.
+Since t-norms and t-conorms are commutative and associative, they can also be applied to non-empty sets of arbitrary size, i.e.
 $T(\ldblbrace y_1, \dots, y_n \rdblbrace) = T(y_1, T(\ldblbrace y_2, \dots, y_n \rdblbrace))$ with fixpoint $T(\ldblbrace y \rdblbrace) = y$.
-Using this extension, a uninorm $U_\lambda$ can be applied to sets which turns it into a parameterized aggregation function $\mathcal{A}_\lambda: {[0, 1]}^* \to [0, 1]$.
+Using this extension, a uninorm $U_\lambda$ can be applied to sets, which turns it into a parameterized aggregation function $\mathcal{A}_\lambda: {[0, 1]}^* \to [0, 1]$.
 In this simple model the \ac{lta} aggregation problem boils down to the optimization of $\lambda$.
-The \ac{lta} disaggregation problem is solved by jointly optimizing a \acl{lm}, i.e.\ the constituent scores ${\ldblbrace y_i \in [0, 1] \rdblbrace}_{c_i \in C}$ are described by $y_i = {\left( 1 + \exp(-\theta^\top x_i) \right)}^{-1}$.
-Overall an \ac{lta} model is therefore described by the uninorm parameter $\lambda$ and the regression coefficients $\theta$.
+The \ac{lta} disaggregation problem is solved by jointly optimizing a \acl{lm}; i.e.\ the constituent scores ${\ldblbrace y_i \in [0, 1] \rdblbrace}_{c_i \in C}$ are described by $y_i = {\left( 1 + \exp(-\theta^\top x_i) \right)}^{-1}$.
+Overall, such an \ac{lta} model is therefore described by the uninorm parameter $\lambda$ and the regression coefficients $\theta$.
 
@@
 To deal with varying composition sizes $n$, the weights $\lambda_1, \dots, \lambda_n$ can however not be statically assigned.
-Instead they are interpolated using a so-called \ac{bum} function $q: [0, 1] \to [0, 1]$.
-It takes constituent positions that are normalized to the unit interval, i.e.\ $\frac{i}{n} \in [0, 1]$.
-The \ac{bum} function $q$ is then used to interpolate a weight for any normalized sort position via $\lambda_i \coloneqq q\left(\frac{i}{n}\right) - q\left(\frac{i - 1}{n}\right)$.
+Instead, they are interpolated using a so-called \ac{bum} function $q: [0, 1] \to [0, 1]$ which takes constituent positions that are normalized to the unit interval, i.e.\ $\frac{i}{n} \in [0, 1]$.
+The \ac{bum} function $q$ is used to interpolate a weight for any normalized sort position via $\lambda_i \coloneqq q\left(\frac{i}{n}\right) - q\left(\frac{i - 1}{n}\right)$.
 Because $q$ is monotone with $q(0) = 0$ and $q(1) = 1$, it always holds that ${\|\lambda\|}_1 = q(1) - q(0) = 1$. % chktex 21
@@
 \end{figure}
-In the \ac{owa} approach the \ac{bum} function $q$ is modeled as a piecewise linear spline.
+In the \ac{owa} \ac{lta} approach the \ac{bum} function $q$ is modeled as a piecewise linear spline.
 This spline is described by $m+1$ points ${\left\{ \left( \frac{j}{m}, a_j \right) \right\}}_{j = 0}^{m}$, the so-called knots of the spline. % chktex 21
@@
 The \ac{lta} aggregation problem is therefore solved by optimizing $a \in \mathbb{R}^{m + 1}$ under this constraint.
-The disaggregation problem is tackled by adding the scores $y_1, \dots, y_M \in \mathbb{R}$ to the learnable parameters of the model where $M$ is assumed to be the finite number of constituents.
+The disaggregation problem is tackled by adding the scores $y_1, \dots, y_M \in \mathbb{R}$ to the learnable parameters of the model, where $M$ is assumed to be the finite number of constituents.
 Currently the \ac{owa} approach requires all possible constituents to be part of the training dataset since it does not consider constituent features $x_i \in \mathcal{X}$ to predict the scores of previously unseen constituents.
@@
 	A \textit{graph} $G \coloneqq (\mathcal{V}_G, \mathcal{E}_G)$ consists of a finite set of vertices $v_i \in \mathcal{V}_G$ and a set of edges $e_{ij} = (v_i, v_j) \in \mathcal{E}_G \subseteq \mathcal{V}_G^2$.
-	Optionally discrete vertex labels $l_G[v_i] \in L_{\mathcal{V}}$ or edge labels $l_G[e_{ij}] \in L_{\mathcal{E}}$ may be associated with all vertices $v_i \in \mathcal{V}_G$ and edges $e_{ij}\in \mathcal{E}_G$ respectively.
+	Optionally, discrete vertex labels $l_G[v_i] \in L_{\mathcal{V}}$ or edge labels $l_G[e_{ij}] \in L_{\mathcal{E}}$ may be associated with all vertices $v_i \in \mathcal{V}_G$ and edges $e_{ij}\in \mathcal{E}_G$ respectively.
 	Also continuous feature vectors $x_G[v_{i}] \in \mathcal{X}_{\mathcal{V}}$, $x_G[e_{ij}] \in \mathcal{X}_{\mathcal{E}}$ may be given.
@@
 
-	In this thesis all graphs $G$ are assumed to be undirected if not explicitly stated otherwise, i.e.\ ${e_{ij} \in \mathcal{E}_G} \leftrightarrow\allowbreak {e_{ji} \in \mathcal{E}_G} \allowbreak\land l_G[e_{ij}] = l_G[e_{ji}] \allowbreak\land x_G[e_{ij}] = x_G[e_{ji}]$.
+	In this thesis, all graphs $G$ are assumed to be undirected if not explicitly stated otherwise, i.e.\ ${e_{ij} \in \mathcal{E}_G} \leftrightarrow\allowbreak {e_{ji} \in \mathcal{E}_G} \allowbreak\land l_G[e_{ij}] = l_G[e_{ji}] \allowbreak\land x_G[e_{ij}] = x_G[e_{ji}]$.
 	We denote the set of all undirected graphs as $\mathcal{G}$.
-	Additionally we denote $G[S] \coloneqq (S, \mathcal{E}_G \cap S^2)$ as the subgraph of $G$ induced by $S \subseteq \mathcal{V}_G$.
+	Additionally, we denote $G[S] \coloneqq (S, \mathcal{E}_G \cap S^2)$ as the subgraph of $G$ induced by $S \subseteq \mathcal{V}_G$.
 \end{defn}
@@
 Consequently there are $n!$ equivalent encodings of $G$. % chktex 40
-To represent those encodings we introduce the notion of ordered induced subgraphs.
+To represent those encodings, we introduce the notion of ordered induced subgraphs.
 \begin{defn}\label{defn:related:ordered-subgraph}
-	For all vertex $k$-tuples $v = (v_1, \dots, v_k) \in \mathcal{V}_G^k$ let $\hat{v} = \set(v) \coloneqq \{ v_i\, |\, i \in [k] \}$.
+	For all vertex $k$-tuples $v = (v_1, \dots, v_k) \in \mathcal{V}_G^k$, let $\hat{v} = \set(v) \coloneqq \{ v_i\, |\, i \in [k] \}$.
 	Then $G[v] \coloneqq (\hat{v}, \mathcal{E}_G \cap \hat{v}^2, v)$ is called the \textit{ordered subgraph} of $G$ induced by $v$.
@@
 In addition to the binary aspect of \ac{wl} distinguishability and its relation to the \ac{gi} problem, \ac{wl} colorings are also useful for more fuzzy graph similarity comparisons as we will see in \cref{sec:related:gcr:kernel} when we look at graph kernels.
-Before that however, the details of \ac{wl} color refinement strategy have to be described.
+Before that however, the details of the \ac{wl} color refinement strategy have to be described.
 We begin with the color refinement algorithm for the most simple case of $k = 1$.
@@
 % chktex 25
-In each iteration of the 1-\acs{wl} color refinement algorithm, the following neighborhood aggregation scheme is used to compute a new color for all vertices:
+In each iteration of the 1-\acs{wl} color refinement algorithm, the following neighborhood aggregation scheme is used to compute a new color for each vertex:
 \begin{defn}\label{defn:related:wl1-refine}
@@
 By extending \ac{wl} to higher dimensions, such 1-\acs{wl} indistinguishable cases can however be handled.
-Analogous to the 1-dimensional \cref{defn:related:wl1-refine}, the $k$-dimensional color refinement step is defined by
+Analogous to the 1-dimensional variant in \cref{defn:related:wl1-refine}, the $k$-dimensional color refinement step is defined by
 \begin{defn}\label{defn:related:wlk-refine}
@@
 \end{defn}
-In 1-\acs{wl} a vertex color is refined by combining the colors of neighboring vertices.
-In $k$-\acs{wl} the color of a $k$-tuple $s \in \mathcal{V}_G^k$ is refined by combining the colors of its neighborhood which is defined as the set of all $k$-tuples in which at most one vertex differs from $s$.
+In 1-\acs{wl}, a vertex color is refined by combining the colors of neighboring vertices.
+In $k$-\acs{wl}, the color of a $k$-tuple $s \in \mathcal{V}_G^k$ is refined by combining the colors of its neighborhood which is defined as the set of all $k$-tuples in which at most one vertex differs from $s$.
 Note that each vertex $k$-tuple has one neighbor for each $u \in \mathcal{V}_G$, each of which is a $k$-tuple of vertex $k$-tuples.
@@
 For $k = 1$ the equivalence class of a single vertex $v$ directly corresponds to its label $l_G[v]$.
-More generally for $k > 1$ this means that
+More generally, for $k > 1$ this means that
 \begin{align}
@@
 \Cref{fig:related:wl2-problem-solution} illustrates \cref{cor:related:wl-monotonous} by showing how 2-WL is able to distinguish the two 1-WL indistinguishable graphs from \cref{fig:related:wl1-problem}.
-Since the time complexity of $k$-\acs{wl} grows exponentially with $k$~\cite[cor.~1.9.7]{Immerman1990}, it does not provide an efficient universal solution to \ac{gi}.
+Since the time complexity of $k$-\acs{wl} grows exponentially with $k$~\cite[cor.~1.9.7]{Immerman1990}, it does not provide an efficient universal solution to the \ac{gi} problem.
 However it turns out that almost all graphs are \ac{wl} distinguishable even for a small constant $k$.
@@
 Though, by \cref{prop:related:wl-upper}, there are of course still 2-\acs{wl} indistinguishable graphs;
-among others those are the strongly regular graphs $\mathit{srg}_{n, d, \lambda, \mu}(G)$.
+among others, those are the strongly regular graphs $\mathit{srg}_{n, d, \lambda, \mu}(G)$.
 \begin{defn}
@@
 \end{defn}
-Generally this restriction of 2-\acs{wl} is not an issue since strongly regular graphs do not appear in typical \ac{gcr} datasets.
-By going to $k = 3$, even some strongly regular graphs as well as all planar graphs can be distinguished though~\cite{Kiefer2017}.
-Apart from that there are currently few results regarding the classes of distinguishable graphs and computable functions for even higher \ac{wl}-dimensions.
+Generally speaking, this restriction of 2-\acs{wl} is not an issue since strongly regular graphs do not appear in typical \ac{gcr} datasets.
+By going to $k = 3$, even some strongly regular graphs, as well as all planar graphs, can be distinguished though~\cite{Kiefer2017}.
+Apart from that, there are currently few results regarding the classes of distinguishable graphs and computable functions for even higher \ac{wl}-dimensions.
 
@@
 Let us now look at an alternative perspective on graph characterization which is provided by \textit{spectral graph theory}.
-While \ac{wl} characterizes a graph via its color distribution $\mathit{dist}_{\chi_{G,k}^{*}}$, the spectral approach uses its so-called \textit{spectrum} $\lambda_G = (\lambda_{G, 1}, \dots, \lambda_{G, n}) \in \mathbb{R}^n$ where $n = |\mathcal{V}_G|$.
+While \ac{wl} characterizes a graph via its color distribution $\mathit{dist}_{\chi_{G,k}^{*}}$, the spectral approach uses its so-called \textit{spectrum} $\lambda_G = (\lambda_{G, 1}, \dots, \lambda_{G, n}) \in \mathbb{R}^n$, where $n = |\mathcal{V}_G|$.
 The key idea behind this is to interpret the adjacency matrix $A_G \in \mathbb{R}^{n \times n}$ of $G$ not simply as an encoding of the edges ($A_{G,i,j} = \mathbbm{1}[(v_i, v_j) \in \mathcal{E}_G]$) but as a linear operator $A_G: (\mathcal{V}_G \to \mathbb{R}) \to (\mathcal{V}_G \to \mathbb{R})$ acting on the vector space of real-valued functions with a vertex domain.
-This means that $A_G$ transforms so-called \textit{graph signals} $x: \mathcal{V}_G \to \mathbb{R}$ which are functions assigning \textit{signal strengths} to vertices.
+This means that $A_G$ transforms so-called \textit{graph signals} $x: \mathcal{V}_G \to \mathbb{R}$, which are functions assigning \textit{signal strengths} to vertices.
 By applying $A_G x$ the signal strength $x[v_i]$ of each vertex is added to the signal strengths of its neighbors $\Gamma_G(v_i)$.
@@
 \paragraph{The graph \acl{ft}}
-To extend the notion of the \ac{ft} from real functions $f: \mathbb{R} \to \mathbb{R}$ to graph signals $x: \mathcal{V}_G \to \mathbb{R}$ we need to define a graph variant of the Laplace operator $\Delta$.
+To extend the notion of the \ac{ft} from real functions $f: \mathbb{R} \to \mathbb{R}$ to graph signals $x: \mathcal{V}_G \to \mathbb{R}$, we need to define a graph variant of the Laplace operator $\Delta$.
 It turns out that there are multiple possible ways to do so, the most simple being the so-called \textit{combinatorial graph Laplacian}.
@@
 The second derivative of a function $f$ essentially averages the value differences in the neighborhood of a point $t$.
-For real-valued functions this neighborhood only consists of the two infinitesimally close points to the left and to the right of $t$, i.e.\ $t - h$ and $t + h$.
+For real-valued functions, this neighborhood only consists of the two infinitesimally close points to the left and to the right of $t$, i.e.\ $t - h$ and $t + h$.
 The combinatorial graph Laplacian represents the same operation, where each point/vertex $v$ might however have more than two neighbors $u \in \Gamma_G(v)$ that need to be averaged.
 
-The \ac{ft} of a graph signal $x$ then is $\hat{x}[i] = \langle u_{G,i}, x \rangle$ with $u_G = {\{ u_{G,i} \}}_{i=1}^{n}$ being the eigenvectors of $L_G$, s.t.\ $x = \sum_{i=1}^n \hat{x}[i] u_{G,i}$.
+The \ac{ft} of a graph signal $x$ then is $\hat{x}[i] = \langle u_{G,i}, x \rangle$, with $u_G = {\{ u_{G,i} \}}_{i=1}^{n}$ being the eigenvectors of $L_G$, s.t.\ $x = \sum_{i=1}^n \hat{x}[i] u_{G,i}$.
 Since the set of Laplacian eigenvectors is finite, we can express the graph \ac{ft} $\mathcal{F}_G$ as a change of basis matrix $U_G = \begin{pmatrix} u_{G,1,1} & \cdots & u_{G,1,n} \\ \vdots & \ddots & \vdots \\ u_{G,n,1} & \cdots & u_{G,n,n} \end{pmatrix} \in \mathbb{R}^{n \times n}$.
-Similarly the inverse \ac{ft} can be expressed as $\mathcal{F}^{-1} = U_G^{-1} = U_G^{\top}$ because $U_G$ is a real orthogonal matrix.
+Similarly the inverse \ac{ft} can be expressed as $\mathcal{F}_G^{-1} = U_G^{-1} = U_G^{\top}$ because $U_G$ is a real orthogonal matrix.
 To see why this is the case, note that $L_G$ is guaranteed to only have real eigenvectors $u_G$ and eigenvalues $\lambda_G$ due to the fact that we only consider undirected graphs with symmetric adjacency matrices $A_G$.
@@
 		Comparison between the basis functions/vectors of the real domain \ac{ft} and the graph \ac{ft}.
-		For the eigenfunctions on the upper half only the real cosine components of the complex exponentials are shown.
+		For the eigenfunctions on the upper half, only the real cosine components of the complex exponentials are shown.
 		\source[based on]{Shuman2013}
@@
 }.
-Consequently a small graph $G$ will typically have a large spectral distance to a large graph $H$ (assuming ${|\mathcal{E}_G|} \ll {|\mathcal{E}_H|}$) irrespective of their similarity when ignoring the scale difference.
+Consequently a small graph $G$ will typically have a large spectral distance to a large graph $H$ (assuming ${|\mathcal{E}_G|} \ll {|\mathcal{E}_H|}$), irrespective of their similarity when ignoring the scale difference.
 In domains where the absolute size of a graph should not influence its spectral characterization, it can therefore be useful to normalize the spectrum.
@@
 The existing approaches to tackle the \ac{gcr} problem can be categorized into three main families:
-\begin{enumerate*}
+\begin{enumerate*}[label={\circled{\small\arabic*}}]
 	\item Explicit graph embeddings,
@@
 The embedding of a graph $G$ is a binary vector ${\varphi}_{\mathrm{FP}}(G) \in {\{0, 1\}}^d$ with ${\varphi}_{\mathrm{FP}}(G)[i] \coloneqq \mathit{contain}_{S_i}(G)$ (see~\cref{defn:related:wl-count-detect}).
-Alternatively a fingerprint embedding can additionally encode multiplicities via ${\varphi}_{\mathrm{FP}}(G)[i] \coloneqq \mathit{count}_{S_i}(G) \in \mathbb{N}_0$.
+Alternatively, a fingerprint embedding can additionally encode multiplicities via ${\varphi}_{\mathrm{FP}}(G)[i] \coloneqq \mathit{count}_{S_i}(G) \in \mathbb{N}_0$.
 
@@
 It gets a sequence of words $(w_0, \dots, w_n)$ as input and outputs embedding vectors $\varphi(w_0), \dots, \varphi(w_n) \in \mathbb{R}^d$.
-To do this the context $\Gamma_k(w_i) = \{ w_{i-k}, \dots, w_{i + k} \}$ is computed for all words where $w_i$ is the so-called \textit{context root}.
+To do this, the context $\Gamma_k(w_i) = \{ w_{i-k}, \dots, w_{i + k} \}$ is computed for all words, where $w_i$ is the so-called \textit{context root}.
 The word contexts are then used to optimize the following log-likelihood objective:
 \begin{gather}
-	\max_{\varphi, \varphi_{\Gamma}} \sum_{i = 1}^n \log P(\Gamma_k(w_i) | w_i)\ =\ \max_{\varphi, \varphi_{\Gamma}} \sum_{i = 1}^n \smashoperator[r]{\sum_{w_j \in \Gamma_k(w_i)}}\, \left[ {\varphi(w_i)}^\top \varphi_{\Gamma}(w_j) - \log Z_{w_i} \right] \label{eq:related:word2vec}\\
+	\max_{\varphi, \varphi_{\Gamma}} \sum_{i = 1}^n \log P(\Gamma_k(w_i) | w_i)\ =\ \max_{\varphi, \varphi_{\Gamma}} \sum_{i = 1}^n \smashoperator[r]{\sum_{w_j \in \Gamma_k(w_i)}}\, \left( {\varphi(w_i)}^\top \varphi_{\Gamma}(w_j) - \log Z_{w_i} \right) \label{eq:related:word2vec}\\
 	\text{with } P(\Gamma_k(w_i) | w_i) \coloneqq \prod_{\mathclap{w_j \in \Gamma_k(w_i)}}\, \overbrace{\frac{\exp\left({\varphi(w_i)}^\top \varphi_{\Gamma}(w_j)\right)}{Z_{w_i}}}^{P(w_j | w_i)}
@@
 \texttt{word2vec} essentially uses an expectation maximization scheme to maximize the probabilities $P(w_j | w_i)$ of observing the context words $w_j \in \Gamma_k(w_i)$ of all words $w_i$.
-Those probabilities are described by the overlap of the embeddings $\varphi(w_i)$ of words $w_i$ and the embeddings $\varphi_{\Gamma}(w_j)$ of their context words $w_j$.
-Intuitively this means that words with similar contexts will be mapped close to each other in the embedding space.
+Those probabilities are described by the overlaps of the embedding $\varphi(w_i)$ of a word $w_i$ and the embeddings $\varphi_{\Gamma}(w_j)$ of its context words $w_j$.
+Intuitively, this means that words with similar contexts will be mapped close to each other in the embedding space.
 Note that \texttt{word2vec} actually finds two embeddings $\varphi(w)$ and $\varphi_{\Gamma}(w)$ for each word of which only the first is returned.
@@
 The problem with this na{\"\i}ve extension is that the sizes of $k$-neighborhoods in arbitrary graphs can be much larger and often tend to grow exponentially with $k$.
-To deal with this computational problem the so-called DeepWalk~\cite{Perozzi2014} and \texttt{node2vec}~\cite{Grover2016} methods perform random walks of fixed length to effectively take samples from the neighborhood of vertices.
+To deal with this computational problem, the so-called DeepWalk~\cite{Perozzi2014} and \texttt{node2vec}~\cite{Grover2016} methods perform random walks of fixed length to effectively take samples from the neighborhood of vertices.
 Both methods only differ in the transition matrix that is used for the random walk.
-Another difference of DeepWalk and \texttt{node2vec} compared to \texttt{word2vec} is the so-called \textit{feature space symmetry} which states that the context root interpretation ($\varphi$) of a vertex should be symmetric to its context element interpretation ($\varphi_{\Gamma}$), i.e.\ $\varphi = \varphi_{\Gamma}$.
+Another difference of DeepWalk and \texttt{node2vec}, compared to \texttt{word2vec}, is the so-called \textit{feature space symmetry} which states that the context root interpretation ($\varphi$) of a vertex should be symmetric to its context element interpretation ($\varphi_{\Gamma}$), i.e.\ $\varphi = \varphi_{\Gamma}$.
 The combination of random walk context sampling and and the feature space symmetry assumption can be used to compute vertex embeddings even for very large graphs.
@@
 One way to do this is via the \texttt{graph2vec}~\cite{Narayanan2017} method.
-It is inspired by \texttt{doc2vec}~\cite{Le2014} which in turn is based on \texttt{word2vec}.
-\texttt{graph2vec} gets a set of graphs $\mathcal{G}' = \{ G_1, \dots G_N \}$ as input and outputs graph embeddings $\varphi(G_1), \dots, \varphi(G_N)$.
-While in \texttt{word2vec} every word can be a context root as well as a context element, \texttt{graph2vec} uses the graphs $\mathcal{G}'$ as context roots.
+It is inspired by \texttt{doc2vec}~\cite{Le2014}, which in turn is based on \texttt{word2vec}.
+\texttt{graph2vec} gets a set of graphs $\mathcal{G}_{\mathcal{D}} = \{ G_1, \dots G_N \}$ as input and outputs graph embeddings $\varphi(G_1), \dots, \varphi(G_N)$.
+While in \texttt{word2vec} every word can be a context root as well as a context element, \texttt{graph2vec} uses the graphs $\mathcal{G}_{\mathcal{D}}$ as context roots.
 The context $\Gamma_T(G_i)$ of a graph $G_i$ is defined as
@@
 \end{equation}
-Intuitively this context can be understood as the set of 1-\acs{wl}-distinguishable subgraphs of $G_i$ with diameter $\leq 2T$.
-Since \ac{wl} is used to identify distinct subgraphs, \texttt{graph2vec} can only be applied to graphs with discrete vertex labels.
-Using the previous definitions, the context root embedding function has the signature $\varphi: \mathcal{G} \to \mathbb{R}^d$ while the context element embedding function is of type $\varphi_{\Gamma}: \bigcup_{G_i \in \mathcal{G}} \Gamma_T(G_i) \to \mathbb{R}^d$.
-To find those embeddings the \texttt{word2vec} objective from \cref{eq:related:word2vec} is reused.
-Analogous to \texttt{word2vec}, \texttt{graph2vec} therefore embeds graphs that share subgraphs close to each other, whereas graphs that do not share substructures tend to be embedded further away from each other.
+This context describes a graph as the set of its 1-\acs{wl} colors that occur within the first $T$ refinement steps.
+Since \ac{wl} colors do not incorporate continuous feature vectors, \texttt{graph2vec} can only be applied to graphs with discrete vertex labels.
+Using the previous definitions, the context root embedding function has the signature $\varphi: \mathcal{G} \to \mathbb{R}^d$, while the context element embedding function is of type $\mathcal{C} \to \mathbb{R}^d$.
+To find those embeddings, the \texttt{word2vec} objective from \cref{eq:related:word2vec} is reused.
+Analogous to \texttt{word2vec}, \texttt{graph2vec} therefore embeds graphs that share the same \ac{wl} colors close to each other, whereas graphs that do not share colors tend to be embedded further away from each other.
 
@@
 Note that the underlying mapping $\varphi_{\mathrm{ST}}$ cannot be computed independently for each graph $G \in \mathcal{G}_{\mathcal{D}} \subseteq \mathcal{G}$ in a given training dataset $\mathcal{D}$ since the set of colors $c \in \mathcal{C}$ introduced by the hashing function $h$ is conjointly determined by all graphs $\mathcal{G}_{\mathcal{D}}$ (see \cref{defn:related:wl1-refine}).
-Consequently the dimensionality $d$ of $\varphi_{\mathrm{ST}}(G) \in \mathbb{N}^d$ varies, depending on the entire dataset $\mathcal{D}$.
+Consequently, the dimensionality $d$ of $\varphi_{\mathrm{ST}}(G) \in \mathbb{N}^d$ varies, depending on the entire dataset $\mathcal{D}$.
 
@@
 even for $k = 2$ the cost can become infeasibly high when working with large graphs.
-To tackle this problem \citet{Morris2017} proposed a combination of two optimizations:
+To tackle this problem, \citet{Morris2017} proposed a combination of two optimizations:
 \begin{enumerate}[label=\textbf{\arabic*.}]
@@
 		\end{align}
-		Even though this simplification generally reduces the discriminative power of the kernel, for $k = 2$ specifically no information and therefore no power is lost by replacing the tuples $(v, u)$ and $(u, v)$ with $\ldblbrace v, u \rdblbrace$ since the undirectedness of graphs implies that $\chi_{G,2}^{(t)}(v, u) = \chi_{G,2}^{(t)}(u, v)$ even when doing tuple-based color refinement.
+		Even though this simplification generally reduces the discriminative power of the kernel, for $k = 2$ specifically, no information and therefore no power is lost by replacing the tuples $(v, u)$ and $(u, v)$ with $\ldblbrace v, u \rdblbrace$ since the undirectedness of graphs implies that $\chi_{G,2}^{(t)}(v, u) = \chi_{G,2}^{(t)}(u, v)$ even when doing tuple-based color refinement.
 	\item \textbf{Neighborhood localization:}
@@
 We call the kernel that only uses the first optimization \textit{$k$-GWL}~($k$-dim.\ global \ac{wl}) and the kernel that uses both optimizations \textit{$k$-LWL}~($k$-dim.\ local \ac{wl}).
-\citet{Morris2017} have empirically shown that focusing on local graph structures via LWL often actually performs better than the computationally more expensive GWL kernel.
+\citet{Morris2017} have empirically shown that focusing on local graph structures via LWL often actually results in a better predictive performance than that achieved by the computationally more expensive GWL kernel.
 
@@
 Since then many variants and extensions of that idea have been proposed~\cite{Wu2019}.
-We will focus specifically on the so-called \acp{gcnn} which can be divided into two variants:
+We will focus specifically on the so-called \acp{gcnn}, which can be divided into two variants:
 Spectral and spatial \acp{gcnn}.
@@
 To see the connection to the convolution theorem, remember that the Laplacian eigenvector matrices $U_G$ and $U_G^{\top}$ correspond to the \ac{ft} $\mathcal{F}$ and inverse \ac{ft} $\mathcal{F}^{-1}$ respectively.
-In this formulation of convolution the Fourier transformed kernel $\hat{g}$ can be interpreted as a spectral filter, i.e.\ it dampens or amplifies certain eigenvector components of a graph.
+In this formulation of convolution, the Fourier transformed kernel $\hat{g}$ can be interpreted as a spectral filter, i.e.\ it dampens or amplifies certain eigenvector components of a graph.
 \citet{Bruna2013}\cite{Henaff2015} first described a \ac{gnn} architecture that learns such a spectral filter $\hat{g}$.
@@
 It describes a spectral filter as a linear combination $\hat{g}_{\theta}(\lambda) \coloneqq \sum_{k=0}^{K-1} \theta_k T_k(\lambda)$ with $\theta \in \mathbb{R}^K$.
-This restricts $\hat{g}_{\theta}$ to be a polynomial which allows us to rewrite the graph convolution from \cref{eq:related:graph-conv-general} as
+This restricts $\hat{g}_{\theta}$ to be a polynomial, which allows us to rewrite the graph convolution from \cref{eq:related:graph-conv-general} as
 \begin{align}
@@
 because $L_G = U_G^{\top} \Lambda_G U_G$ is an eigendecomposition of the Laplacian.
-The $\tfrac{2}{\max \lambda_{G}} \Lambda_G - I$ term normalizes the eigenvalues to $[-1, 1]$ which prevents vanishing and exploding gradients.
+The $\tfrac{2}{\max \lambda_{G}} \Lambda_G - I$ term normalizes the eigenvalues to $[-1, 1]$, which prevents vanishing and exploding gradients.
 The advantage of this formulation is that it can be evaluated without actually having to compute the expensive eigendecomposition of $L_G$.
@@
 \end{align}
-In this formulation the spectral filter is effectively just a linear function $\hat{g}_{\theta}(\lambda) = \theta \lambda$.
+In this formulation, the spectral filter is effectively just a linear function $\hat{g}_{\theta}(\lambda) = \theta \lambda$.
 Via this simplified notion of convolution for a single feature signal $x_{G,j}$, a convolutional \acl{nn} layer over all features ${\{ x_{G,j} \}}_{j=1}^{d}$ can analogously be defined as
@@
 Instead of interpreting convolutions as applications of linear spectral filters, we can interpret them as vertex neighborhood aggregations, similar to 1-\acs{wl}.
-From this perspective a \ac{gcn} can be seen as a so-called spatial \ac{gcnn} because it operates directly in the vertex domain.
+From this perspective, a \ac{gcn} can be seen as a so-called spatial \ac{gcnn} because it operates directly in the vertex domain.
 To see why this is the case, we rewrite \cref{eq:related:gcn-layer} and compare it with the 1-\acs{wl} color refinement strategy (see \cfullref{defn:related:wl1-refine}):
@@
 The comparison shows that \ac{gcn} convolutions can be understood as iterative refinements of continuous ``color vectors'' $Z_G^{(t)}[v] \in \mathbb{R}^{d^{(t)}}$.
-Then the continuous analogue to the injective 1-\acs{wl} hash function $h: \mathcal{C}^{*} \to \mathcal{C}$ can be defined as $h_{\mathrm{\acs*{gcn}}}(z_0, z_1, \dots, z_m) = \sigma(\sum_{j=0}^m \mu(v_0) \mu(v_j) z_j \Theta^{(t)})$ which computes a single color vector from multiple color vectors $z_j$ where $v_j$ are the corresponding vertices.
+Then the continuous analogue to the injective 1-\acs{wl} hash function $h: \mathcal{C}^{*} \to \mathcal{C}$ can be defined as $h_{\mathrm{\acs*{gcn}}}(z_0, z_1, \dots, z_m) = \sigma(\sum_{j=0}^m \mu(v_0) \mu(v_j) z_j \Theta^{(t)})$ which computes a single color vector from multiple color vectors $z_j$ with the corresponding vertices $v_j$.
 
@@
 \end{prop}
-By \cref{prop:related:gcnn-wl1-limit} the \ac{gcn} architecture is strictly less powerful than 1-\acs{wl} due to the fact that $h_{\mathrm{\acs*{gcn}}}$ is not an injective hash function.
-To overcome this limitation \citet{Xu2018} proposed an alternative vertex neighborhood convolution architecture called \ac{gin}:
+By \cref{prop:related:gcnn-wl1-limit}, the \ac{gcn} architecture is strictly less powerful than 1-\acs{wl} due to the fact that $h_{\mathrm{\acs*{gcn}}}$ is not an injective hash function.
+To overcome this limitation, \citet{Xu2018} proposed an alternative vertex neighborhood convolution architecture called \ac{gin}:
 \begin{align}
@@
 \end{align}
-By leaving out the normalization factors $\mu(v)$ and by using a \ac{mlp} instead of the single fully-connected layer $\sigma \circ \Theta^{(t)}$ the resulting hashing function $h_{\mathrm{\acs*{gin}}}(z_1, \dots, z_m) = \mathrm{\acs*{mlp}}^{(t)}(\sum_{j=1}^{m} z_j)$ becomes injective\footnote{
+By leaving out the normalization factors $\mu(v)$ and by using a \ac{mlp} instead of the single fully-connected layer $\sigma \circ \Theta^{(t)}$, the resulting hashing function $h_{\mathrm{\acs*{gin}}}(z_1, \dots, z_m) = \mathrm{\acs*{mlp}}^{(t)}(\sum_{j=1}^{m} z_j)$ becomes injective\footnote{
 	Under the assumption that the set of possible input feature matrices $X \in \mathbb{R}^{n \times d^{(0)}}$ is countable.
@@
 }, i.e.\ it assigns a unique color vector to each multiset of color vectors.
-By \cref{prop:related:gcnn-wl1-limit} this implies that \ac{gin} has the same discriminative power as 1-\ac{wl}.
-As expected, \ac{gin} therefore fits training data better than non-injective \acp{gcnn} like \ac{gcn}.
-Interestingly \ac{gin} additionally seems to generalize better on test data than non-injective methods~\cite{Xu2018}\cite{Errica2020}.
+By \cref{prop:related:gcnn-wl1-limit}, this implies that \ac{gin} has the same discriminative power as 1-\ac{wl}.
+As expected, \ac{gin} therefore fits training data better than less powerful, non-injective \acp{gcnn} like \ac{gcn}.
+Interestingly, \acp{gin} additionally seem to generalize better on test data than non-injective methods~\cite{Xu2018}\cite{Errica2020}.
 It is not yet fully understood why this is the case.
@@
 \paragraph{Higher dimensional \ac{wl} \acp{gnn}}
-The majority of \ac{gnn} methods uses a vertex neighborhood convolution scheme which limits their discriminative and, more importantly, computational power to that of 1-\acs{wl} (see \cfullref{defn:related:wl-count-detect}).
+The majority of \ac{gnn} methods uses a vertex neighborhood convolution scheme which limits their discriminative and, more importantly, computational power to that of 1-\acs{wl} (see \cfullref{defn:related:wl-compute}).
 To go beyond the limits of 1-\acs{wl}, \citet{Morris2019} proposed the first \ac{gcnn} architecture inspired by the higher dimensional $k$-\acs{wl} algorithms.
@@
 \begin{align}
-	Z_G^{(t)}[s] \coloneqq \sigma\left( {Z_G^{(t-1)}[s]} W^{(t)} + \smashoperator[lr]{\sum_{{v \in s}, {u \in \Gamma_G(v)}}} {Z_G^{(t-1)}[s \setminus \{ v \} \cup \{ u \}]} W_{\Gamma}^{(t)} \right) \text{.} \label{eq:related:kgnn-layer}
+	Z_G^{(t)}[s] \coloneqq \sigma\left( {Z_G^{(t-1)}[s]} W^{(t)} + \smashoperator[lr]{\sum_{{v \in s},\,{u \in \Gamma_G(v)}}} {Z_G^{(t-1)}[s \setminus \{ v \} \cup \{ u \}]} W_{\Gamma}^{(t)} \right) \text{.} \label{eq:related:kgnn-layer}
 \end{align}
@@
 This separation of the parameter matrix effectively fulfills the same purpose as the \ac{mlp} in \ac{gin}; it makes the implicitly defined $k$-\acs{gnn} hash function injective.
-This can be seen by realizing that a stack of $k$-\acs{gnn} convolution layers with $W_{\Gamma}^{(t)} = \mathbf{0}$ for all but the first layer simulates a \ac{mlp}.
+This can be seen by realizing that a stack of $k$-\acs{gnn} convolution layers, with $W_{\Gamma}^{(t)} = \mathbf{0}$ for all layers, simulates a \ac{mlp}.
 
@@
 To do this, so-called \textit{graph pooling layers} are used.
-Generally speaking there are two types of pooling approaches:
+Generally speaking, there are two types of pooling approaches:
 \begin{enumerate}[label=\textbf{\arabic*.}]
 	\item \textbf{Hierarchical Pooling:}
-		This type of pooling is similar to that found in \acp{cnn} on pictures where the resolution is iteratively reduced.
-		Similarly hierarchical graph pooling iteratively \textit{coarsens} a graph every couple of convolutional layers.
+		This type of pooling is similar to that found in \acp{cnn} which iteratively reduce the resolution of images.
+		Similarl,y hierarchical graph pooling iteratively \textit{coarsens} a graph every couple of convolutional layers.
 		Graph coarsening works by merging vertices in a spectrum-preserving manner~\cite{Loukas2018}.
@@
 	\item \textbf{Global Pooling}:
-		Alternatively pooling can also be performed in a single step after the convolutional layers.
+		Alternatively, pooling can also be performed in a single step after the convolutional layers.
 		This type of pooling takes all $m$ feature vectors and directly maps them to a single graph feature vector.
@@
 The so-called \ac{sagpool}~\cite{Lee2019} learns those attention scores via a separate stack of convolutional layers.
-The set of convolved feature vectors is then filtered down from the size $m$ to $p = \lceil r m \rceil$ with $r \in [0, 1]$ by removing the vectors of the vertices with the lowest attention scores.
+The set of convolved feature vectors is then filtered down from the size $m$ to $p \coloneqq \lceil r m \rceil$, with $r \in [0, 1]$, by removing the vectors of the vertices with the lowest attention scores.
 Let $z_1, \dots, z_p$ be the remaining feature vectors.

### 2020-04-15 thesis/content/chapter-ltd.tex
added another illustration to theorem 4.11 this hopefully makes the defined convolution matrices more intuitive

@@
 			&= \begin{cases}
-				\textcolor{t_blue}{(1, 0, w_{ii}, 0)} + \textcolor{t_darkgreen}{\left(0, 0, 0, 2\textcolor{t_red}{w_{ii}} Z^{(t-1)}[v_i]\right)} + \smashoperator[lr]{\sum_{v_l \in \Gamma_{G^1}(v_i)}} \textcolor{t_darkgreen}{\left(0, w_{il} \textcolor{t_red}{Z^{(t-1)}[v_i]}, 0, 0\right)} & \text{if $i = j$} \\ % chktex 25
+				\textcolor{t_blue}{(1, 0, w_{ii}, 0)} + \textcolor{t_darkgreen}{\left(0, 0, 0, 2\textcolor{t_red}{w_{ii}} Z^{(t-1)}[v_i]\right)} + \smashoperator[lr]{\sum_{v_l \in \Gamma_{G^1}(v_i)}} \textcolor{t_darkgreen}{\left(0, \tfrac{2}{\textcolor{t_red}{2}} \textcolor{t_red}{Z^{(t-1)}[v_i]} w_{il}, 0, 0\right)} & \text{if $i = j$} \\ % chktex 25
 				\textcolor{t_blue}{(0, 0, w_{ij}, 0)} + \textcolor{t_darkgreen}{\left(0, 0, 0, \textcolor{t_red}{w_{ij}} Z^{(t-1)}[v_i]\right)} + \textcolor{t_darkgreen}{\left(0, 0, 0, \textcolor{t_red}{w_{ij}} Z^{(t-1)}[v_j]\right)} & \text{else} % chktex 25
@@
 	\end{gather*}
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=\linewidth]{gfx/ltd/wl2-vert-agg-intuition.pdf}
+		\caption[Intuition for how $S^{(t, 1)}$ and $S^{(t, 2)}$ compute vertex neighborhood sums.]{
+			Intuition for how $S^{(t, 1)}$ and $S^{(t, 2)}$ compute vertex neighborhood sums in two steps.
+			For simplicity, weights are ignored, i.e.\ all $w_{ij} = 1$.
+			For each self-loop/edge $e_{ij} \in \mathcal{E}_{G^1}$, the set of localized 2-\acs{wl} neighbors $\Gamma_{G^1}(v_i) \cap \Gamma_{G^1}(v_j)$ is shown in the middle, after the first step.
+			Note that the colors in this illustration are unrelated to the colored parts in the previous equations.
+		}\label{fig:ltd:wl2-vert-agg-intuition}
+	\end{figure}
 	Using the two layers $S^{(t, 1)}$ and $S^{(t, 2)}$ that we just defined, the weighted vertex neighborhood sum for all $v_i \in \mathcal{V}_G$ is contained in $Z^{(t,2)}[e_{ii}]$.
@@
 	This means that \cref{inv:ltd:wl2-simulation:indicator,inv:ltd:wl2-simulation:weight} are satisfied after $S^{(t, 2)}$.
+
 	To complete the induction step, it now remains to show that all three invariants hold after applying the layers $S^{(t, 2+1)}, \dots, S^{(t, 2+K)}$.
@@
 	\end{align*}
-	By our previous definition of $\mathit{Pool}_2$, this in turn implies that $\mathit{Pool}_2(Z_G^{(T,2+K)}) = \mathit{Pool}_1(Z_G^{(T)}) \iff h_2(G) = h_1(G)$ which concludes the proof.
+	By our previous definition of $\mathit{Pool}_2$, this in turn implies that $\mathit{Pool}_2(Z_G^{(T,2+K)}) = \mathit{Pool}_1(Z_G^{(T)}) \iff h_2(G) = h_1(G)$, which concludes the proof.
 \end{proof}

### 2020-04-16 thesis/content/chapter-evaluation.tex
thesis final

@@
 		}} \\
-		\raisebox{0.33\height}[0pt][0pt]{\rotatebox{90}{\small\textsc{$\mathrm{SAM}$~pooling}}} &
-		\wlRadiusPlot{triangle}{ylabel={accuracy},xtick={1,2},xmin=0,xmax=3,ymin=80,ymax=100,try min ticks=4, xlabel={radius $r$}}{sam} &
+		\raisebox{0.35\height}[0pt][0pt]{\rotatebox{90}{\small\textsc{$\mathrm{SAM}$~pooling}}} &
+		\wlRadiusPlot{triangle}{title={\hphantom{\small TRIANGLE}},ylabel={accuracy},xtick={1,2},xmin=0,xmax=3,ymin=80,ymax=100,try min ticks=4, xlabel={radius $r$}}{sam} &
 		\wlRadiusPlot{nci}{xtick={1,8},xmin=-6,xmax=15,ymin=60,ymax=80,try min ticks=4, xlabel={radius $r$}}{sam} &
-		\wlRadiusPlot{proteins}{xtick={1,5},xmin=-3,xmax=9,ymin=65,ymax=85,try min ticks=4, xlabel={radius $r$}}{sam} &
+		\wlRadiusPlot{proteins}{title={\hphantom{\small PROTEINS}},xtick={1,5},xmin=-3,xmax=9,ymin=65,ymax=85,try min ticks=4, xlabel={radius $r$}}{sam} &
 		\wlRadiusPlot{dd}{xtick={1,2},xmin=0,xmax=3,ymin=65,ymax=85,try min ticks=4, xlabel={radius $r$}}{sam} &

### 2020-04-16 thesis/content/chapter-graph-lta.tex
thesis final

@@
 
-	To complete the proof it now remains to show that $f_h$ and $\mathcal{A}_h$ are in fact a local evaluation function and a weighted aggregation function respectively.
+	To complete the proof, it now remains to show that $f_h$ and $\mathcal{A}_h$ are in fact a local evaluation function and a weighted aggregation function respectively.
 	To see that $f_h$ performs local evaluation, note that $f_h(c_t) \coloneqq (\sgn z_t, {|z_t|})$, with $z_t \coloneqq \sum_{i=1}^{d} \gamma_{\varphi,i}(c_t) \beta_i$ only depending on the shared multiplicity functions $\gamma_{\varphi,i}$, the constants $\beta_i$ and the constituent $c_t$;
-	apart from $c_t$ no other information from the input graph $G$ is required by $f_h$ which makes it a local evaluation function.
+	apart from $c_t$ no other information from the input graph $G$ is required by $f_h$, which makes it a local evaluation function.
 	To see why $\mathcal{A}_h$ satisfies \cref{defn:ltag:weighted-agg}, note that it inherits the idempotency property from $\wmaj$ because $\mathcal{A}_h$ ignores the bias ``pseudo-constituent'' $(y_b, w_b)$ for all threshold weights $\max w_t \geq \eta > w_b$, similarly the zero invariance of $\wmaj$ is also directly inherited.
@@
 				The subtrees $S_{j,t}$ that span each constituent $c_{j,t}$ are visualized by the colored arrows that originate from the colored root vertices.
-				Because $S_{1,1}$ and $S_{2,1}$ span the same constituent $c = c_{1,1} = c_{2,1} = G[\{ v_1, v_2, v_3 \}]$, its multiplicity is $\gamma_{\varphi, \colorlabel{t_red}{\scriptsize B}, 1}(c) = 2$; % chktex 25
+				Because $S_{1,1}$ and $S_{2,1}$ span the same constituent $c = c_{1,1} = c_{2,1} = G[\{ v_1, v_2, v_3 \}]$, its multiplicity is $\gamma_{\varphi, \colorlabel{t_red}{\tiny B}, 1}(c) = 2$; % chktex 25
 				the multiplicity of all other constituents is $1$.
@@
 		The purpose of the $\mathit{complete}(S)$ condition is to only count the \ac{bfs} subtrees of the constituent $c$ that are also \ac{bfs} subtrees of the complete graph $G$.
-		To see why this is required, note that in \cref{fig:ltag:wl1-constituents} the $c_{1,1}$/$c_{2,1}$ constituent contains the color subtree $S_{\colorlabel{t_red}{\scriptsize B}}$ three times (rooted at $v_1$, $v_2$ and $v_3$) even though $G$ only contains it twice (rooted at $v_1$ and $v_2$).
+		To see why this is required, note that in \cref{fig:ltag:wl1-constituents} the $c_{1,1}$/$c_{2,1}$ constituent contains the color subtree $S_{\colorlabel{t_red}{\tiny B}}$ three times (rooted at $v_1$, $v_2$ and $v_3$) even though $G$ only contains it twice (rooted at $v_1$ and $v_2$).
 % chktex 25
@@
 
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=0.73\linewidth]{gfx/graph-lta/gcnn-lta.pdf}
+	\caption[Illustration of the \ac{lta} formulation for \acp{gcnn}.]{
+		Illustration of the \ac{lta} formulation for \acp{gcnn}, as shown in \cref{thm:ltag:gcnn-ltag-formulation}.
+		Constituent weights are left out for simplicity.
+	}\label{fig:ltag:gcnn-lta}
+\end{figure}
 To conclude our analysis of the relation between \ac{lta} and \acp{gcnn}, we check which of the graph pooling layers described in \cref{sec:related:gcr:nn} satisfy \cref{defn:ltag:assoc-aggregation-pooling} are therefore allowed in \acs{lta}-like \acp{gcnn} by \cref{thm:ltag:gcnn-ltag-formulation}.

### 2020-09-02 thesis/data/diffs/imdb.tex
added epoch timing evaluation infrastructure

@@
 {\setlength\tabcolsep{2.5pt}\setlength{\extrarowheight}{2pt}%
-\begin{tabular}{lccccccccccccc}
-& \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=5$)} & \rotatebox[origin=l]{90}{\textbf{WL\textsubscript{SP}} ($T=5$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-GWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{Baseline} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{GIN} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)} \\
-\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)& & {\tiny\Vectorstack{-1.7\\ \pm 3.2}} & {\tiny\Vectorstack{-2.6\\ \pm 3.5}} & {\tiny\Vectorstack{-1.2\\ \pm 2.6}} & {\tiny\Vectorstack{+0.6\\ \pm 3.5}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+20\\ \pm 4.6}}} & {\tiny\Vectorstack{+4.2\\ \pm 6.1}} & {\tiny\Vectorstack{-0.4\\ \pm 3.1}} & {\tiny\Vectorstack{+0.1\\ \pm 2.4}} & {\tiny\Vectorstack{+0.1\\ \pm 3.5}} & {\tiny\Vectorstack{-0.2\\ \pm 4.4}} & {\tiny\Vectorstack{-1.2\\ \pm 3.1}} & {\tiny\Vectorstack{-0.1\\ \pm 3.8}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=5$)&{\tiny\Vectorstack{+1.7\\ \pm 3.2}} &  & {\tiny\Vectorstack{-0.9\\ \pm 3.7}} & {\tiny\Vectorstack{+0.5\\ \pm 3.3}} & {\tiny\Vectorstack{+2.3\\ \pm 3.2}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+22\\ \pm 4.6}}} & {\tiny\Vectorstack{+5.9\\ \pm 6.1}} & {\tiny\Vectorstack{+1.3\\ \pm 3.8}} & {\tiny\Vectorstack{+1.8\\ \pm 4.3}} & {\tiny\Vectorstack{+1.8\\ \pm 4.1}} & {\tiny\Vectorstack{+1.5\\ \pm 4.6}} & {\tiny\Vectorstack{+0.5\\ \pm 4.1}} & {\tiny\Vectorstack{+1.6\\ \pm 5.1}} \\[2pt]
-\textbf{WL\textsubscript{SP}} ($T=5$)&{\tiny\Vectorstack{+2.6\\ \pm 3.5}} & {\tiny\Vectorstack{+0.9\\ \pm 3.7}} &  & {\tiny\Vectorstack{+1.4\\ \pm 4.9}} & {\tiny\Vectorstack{+3.2\\ \pm 5.0}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+23\\ \pm 5.0}}} & {\tiny\Vectorstack{+6.8\\ \pm 6.5}} & {\tiny\Vectorstack{+2.2\\ \pm 4.8}} & {\tiny\Vectorstack{+2.7\\ \pm 4.4}} & {\tiny\Vectorstack{+2.7\\ \pm 5.0}} & {\tiny\Vectorstack{+2.4\\ \pm 5.7}} & {\tiny\Vectorstack{+1.4\\ \pm 5.1}} & {\tiny\Vectorstack{+2.5\\ \pm 5.6}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)&{\tiny\Vectorstack{+1.2\\ \pm 2.6}} & {\tiny\Vectorstack{-0.5\\ \pm 3.3}} & {\tiny\Vectorstack{-1.4\\ \pm 4.9}} &  & {\tiny\Vectorstack{+1.8\\ \pm 2.6}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+22\\ \pm 5.7}}} & {\tiny\Vectorstack{+5.4\\ \pm 7.2}} & {\tiny\Vectorstack{+0.8\\ \pm 2.0}} & {\tiny\Vectorstack{+1.3\\ \pm 3.2}} & {\tiny\Vectorstack{+1.3\\ \pm 3.2}} & {\tiny\Vectorstack{+1.0\\ \pm 3.5}} & {\tiny\Vectorstack{+0.0\\ \pm 3.0}} & {\tiny\Vectorstack{+1.1\\ \pm 3.7}} \\[2pt]
-\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-0.6\\ \pm 3.5}} & {\tiny\Vectorstack{-2.3\\ \pm 3.2}} & {\tiny\Vectorstack{-3.2\\ \pm 5.0}} & {\tiny\Vectorstack{-1.8\\ \pm 2.6}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+20\\ \pm 5.6}}} & {\tiny\Vectorstack{+3.6\\ \pm 7.1}} & {\tiny\Vectorstack{-1.0\\ \pm 2.6}} & {\tiny\Vectorstack{-0.5\\ \pm 4.1}} & {\tiny\Vectorstack{-0.5\\ \pm 3.3}} & {\tiny\Vectorstack{-0.8\\ \pm 3.1}} & {\tiny\Vectorstack{-1.8\\ \pm 3.0}} & {\tiny\Vectorstack{-0.7\\ \pm 4.3}} \\[2pt]
-\textbf{Baseline} ($\mathrm{sum}$)&\cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-20\\ \pm 4.6}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-22\\ \pm 4.6}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-23\\ \pm 5.0}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 5.7}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-20\\ \pm 5.6}}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-16\\ \pm 6.3}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 6.0}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-20\\ \pm 5.6}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-20\\ \pm 5.8}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 6.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 5.5}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-20\\ \pm 6.9}}} \\[2pt]
-\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{-4.2\\ \pm 6.1}} & {\tiny\Vectorstack{-5.9\\ \pm 6.1}} & {\tiny\Vectorstack{-6.8\\ \pm 6.5}} & {\tiny\Vectorstack{-5.4\\ \pm 7.2}} & {\tiny\Vectorstack{-3.6\\ \pm 7.1}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+16\\ \pm 6.3}}} &  & {\tiny\Vectorstack{-4.6\\ \pm 7.5}} & {\tiny\Vectorstack{-4.1\\ \pm 7.1}} & {\tiny\Vectorstack{-4.1\\ \pm 7.3}} & {\tiny\Vectorstack{-4.4\\ \pm 7.9}} & {\tiny\Vectorstack{-5.4\\ \pm 7.0}} & {\tiny\Vectorstack{-4.3\\ \pm 8.4}} \\[2pt]
-\textbf{2-GNN} ($\mean$)&{\tiny\Vectorstack{+0.4\\ \pm 3.1}} & {\tiny\Vectorstack{-1.3\\ \pm 3.8}} & {\tiny\Vectorstack{-2.2\\ \pm 4.8}} & {\tiny\Vectorstack{-0.8\\ \pm 2.0}} & {\tiny\Vectorstack{+1.0\\ \pm 2.6}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+21\\ \pm 6.0}}} & {\tiny\Vectorstack{+4.6\\ \pm 7.5}} &  & {\tiny\Vectorstack{+0.5\\ \pm 2.8}} & {\tiny\Vectorstack{+0.5\\ \pm 2.5}} & {\tiny\Vectorstack{+0.2\\ \pm 2.8}} & {\tiny\Vectorstack{-0.8\\ \pm 2.6}} & {\tiny\Vectorstack{+0.2\\ \pm 2.4}} \\[2pt]
-\textbf{2-GNN} ($\mathrm{SAM}$)&{\tiny\Vectorstack{-0.1\\ \pm 2.4}} & {\tiny\Vectorstack{-1.8\\ \pm 4.3}} & {\tiny\Vectorstack{-2.7\\ \pm 4.4}} & {\tiny\Vectorstack{-1.3\\ \pm 3.2}} & {\tiny\Vectorstack{+0.5\\ \pm 4.1}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+20\\ \pm 5.6}}} & {\tiny\Vectorstack{+4.1\\ \pm 7.1}} & {\tiny\Vectorstack{-0.5\\ \pm 2.8}} &  & {\tiny\Vectorstack{+0.0\\ \pm 2.8}} & {\tiny\Vectorstack{-0.3\\ \pm 3.5}} & {\tiny\Vectorstack{-1.3\\ \pm 2.5}} & {\tiny\Vectorstack{-0.2\\ \pm 2.7}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)&{\tiny\Vectorstack{-0.1\\ \pm 3.5}} & {\tiny\Vectorstack{-1.8\\ \pm 4.1}} & {\tiny\Vectorstack{-2.7\\ \pm 5.0}} & {\tiny\Vectorstack{-1.3\\ \pm 3.2}} & {\tiny\Vectorstack{+0.5\\ \pm 3.3}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+20\\ \pm 5.8}}} & {\tiny\Vectorstack{+4.1\\ \pm 7.3}} & {\tiny\Vectorstack{-0.5\\ \pm 2.5}} & {\tiny\Vectorstack{-0.0\\ \pm 2.8}} &  & {\tiny\Vectorstack{-0.3\\ \pm 1.7}} & {\tiny\Vectorstack{-1.3\\ \pm 1.1}} & {\tiny\Vectorstack{-0.3\\ \pm 2.2}} \\[2pt]
-\textbf{2-WL-GNN\phantom{*}} ($\mean$)&{\tiny\Vectorstack{+0.2\\ \pm 4.4}} & {\tiny\Vectorstack{-1.5\\ \pm 4.6}} & {\tiny\Vectorstack{-2.4\\ \pm 5.7}} & {\tiny\Vectorstack{-1.0\\ \pm 3.5}} & {\tiny\Vectorstack{+0.8\\ \pm 3.1}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+21\\ \pm 6.4}}} & {\tiny\Vectorstack{+4.4\\ \pm 7.9}} & {\tiny\Vectorstack{-0.2\\ \pm 2.8}} & {\tiny\Vectorstack{+0.3\\ \pm 3.5}} & {\tiny\Vectorstack{+0.3\\ \pm 1.7}} &  & {\tiny\Vectorstack{-1.0\\ \pm 2.0}} & {\tiny\Vectorstack{+0.1\\ \pm 3.4}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{+1.2\\ \pm 3.1}} & {\tiny\Vectorstack{-0.5\\ \pm 4.1}} & {\tiny\Vectorstack{-1.4\\ \pm 5.1}} & {\tiny\Vectorstack{-0.0\\ \pm 3.0}} & {\tiny\Vectorstack{+1.8\\ \pm 3.0}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+21\\ \pm 5.5}}} & {\tiny\Vectorstack{+5.4\\ \pm 7.0}} & {\tiny\Vectorstack{+0.8\\ \pm 2.6}} & {\tiny\Vectorstack{+1.3\\ \pm 2.5}} & {\tiny\Vectorstack{+1.3\\ \pm 1.1}} & {\tiny\Vectorstack{+1.0\\ \pm 2.0}} &  & {\tiny\Vectorstack{+1.0\\ \pm 2.6}} \\[2pt]
-\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{+0.1\\ \pm 3.8}} & {\tiny\Vectorstack{-1.6\\ \pm 5.1}} & {\tiny\Vectorstack{-2.5\\ \pm 5.6}} & {\tiny\Vectorstack{-1.1\\ \pm 3.7}} & {\tiny\Vectorstack{+0.7\\ \pm 4.3}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+20\\ \pm 6.9}}} & {\tiny\Vectorstack{+4.3\\ \pm 8.4}} & {\tiny\Vectorstack{-0.2\\ \pm 2.4}} & {\tiny\Vectorstack{+0.2\\ \pm 2.7}} & {\tiny\Vectorstack{+0.3\\ \pm 2.2}} & {\tiny\Vectorstack{-0.1\\ \pm 3.4}} & {\tiny\Vectorstack{-1.0\\ \pm 2.6}} &
+\begin{tabular}{lcccccccccc}
+& \rotatebox[origin=l]{90}{\textbf{WL\textsubscript{ST}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{WL\textsubscript{SP}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-LWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-GWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{Baseline} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{GIN} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\wmean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN} ($\wmean$)} \\
+\textbf{WL\textsubscript{ST}} ($T=3$)& & {\tiny\Vectorstack{-1.5\\ \pm 4.3}} & {\tiny\Vectorstack{+0.7\\ \pm 2.0}} & {\tiny\Vectorstack{+2.5\\ \pm 3.4}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+22\\ \pm 4.9}}} & {\tiny\Vectorstack{+6.1\\ \pm 6.4}} & {\tiny\Vectorstack{+1.5\\ \pm 2.6}} & {\tiny\Vectorstack{+2.0\\ \pm 2.7}} & {\tiny\Vectorstack{+1.7\\ \pm 3.7}} & {\tiny\Vectorstack{+1.8\\ \pm 3.8}} \\[2pt]
+\textbf{WL\textsubscript{SP}} ($T=3$)&{\tiny\Vectorstack{+1.5\\ \pm 4.3}} &  & {\tiny\Vectorstack{+2.2\\ \pm 5.0}} & {\tiny\Vectorstack{+4.0\\ \pm 5.2}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+24\\ \pm 5.9}}} & {\tiny\Vectorstack{+7.6\\ \pm 7.4}} & {\tiny\Vectorstack{+3.0\\ \pm 5.2}} & {\tiny\Vectorstack{+3.5\\ \pm 5.7}} & {\tiny\Vectorstack{+3.2\\ \pm 6.4}} & {\tiny\Vectorstack{+3.3\\ \pm 6.7}} \\[2pt]
+\textbf{2-LWL} ($T=3$)&{\tiny\Vectorstack{-0.7\\ \pm 2.0}} & {\tiny\Vectorstack{-2.2\\ \pm 5.0}} &  & {\tiny\Vectorstack{+1.8\\ \pm 2.6}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+22\\ \pm 5.7}}} & {\tiny\Vectorstack{+5.4\\ \pm 7.2}} & {\tiny\Vectorstack{+0.8\\ \pm 2.0}} & {\tiny\Vectorstack{+1.3\\ \pm 3.2}} & {\tiny\Vectorstack{+1.0\\ \pm 3.5}} & {\tiny\Vectorstack{+1.1\\ \pm 3.7}} \\[2pt]
+\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-2.5\\ \pm 3.4}} & {\tiny\Vectorstack{-4.0\\ \pm 5.2}} & {\tiny\Vectorstack{-1.8\\ \pm 2.6}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+20\\ \pm 5.6}}} & {\tiny\Vectorstack{+3.6\\ \pm 7.1}} & {\tiny\Vectorstack{-1.0\\ \pm 2.6}} & {\tiny\Vectorstack{-0.5\\ \pm 4.1}} & {\tiny\Vectorstack{-0.8\\ \pm 3.1}} & {\tiny\Vectorstack{-0.7\\ \pm 4.3}} \\[2pt]
+\textbf{Baseline} ($\mathrm{sum}$)&\cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-22\\ \pm 4.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-24\\ \pm 5.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 5.7}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-20\\ \pm 5.6}}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-16\\ \pm 6.3}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 6.0}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-20\\ \pm 5.6}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 6.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-20\\ \pm 6.9}}} \\[2pt]
+\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{-6.1\\ \pm 6.4}} & {\tiny\Vectorstack{-7.6\\ \pm 7.4}} & {\tiny\Vectorstack{-5.4\\ \pm 7.2}} & {\tiny\Vectorstack{-3.6\\ \pm 7.1}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+16\\ \pm 6.3}}} &  & {\tiny\Vectorstack{-4.6\\ \pm 7.5}} & {\tiny\Vectorstack{-4.1\\ \pm 7.1}} & {\tiny\Vectorstack{-4.4\\ \pm 7.9}} & {\tiny\Vectorstack{-4.3\\ \pm 8.4}} \\[2pt]
+\textbf{2-GNN} ($\mean$)&{\tiny\Vectorstack{-1.5\\ \pm 2.6}} & {\tiny\Vectorstack{-3.0\\ \pm 5.2}} & {\tiny\Vectorstack{-0.8\\ \pm 2.0}} & {\tiny\Vectorstack{+1.0\\ \pm 2.6}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+21\\ \pm 6.0}}} & {\tiny\Vectorstack{+4.6\\ \pm 7.5}} &  & {\tiny\Vectorstack{+0.5\\ \pm 2.8}} & {\tiny\Vectorstack{+0.2\\ \pm 2.8}} & {\tiny\Vectorstack{+0.2\\ \pm 2.4}} \\[2pt]
+\textbf{2-GNN} ($\wmean$)&{\tiny\Vectorstack{-2.0\\ \pm 2.7}} & {\tiny\Vectorstack{-3.5\\ \pm 5.7}} & {\tiny\Vectorstack{-1.3\\ \pm 3.2}} & {\tiny\Vectorstack{+0.5\\ \pm 4.1}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+20\\ \pm 5.6}}} & {\tiny\Vectorstack{+4.1\\ \pm 7.1}} & {\tiny\Vectorstack{-0.5\\ \pm 2.8}} &  & {\tiny\Vectorstack{-0.3\\ \pm 3.5}} & {\tiny\Vectorstack{-0.2\\ \pm 2.7}} \\[2pt]
+\textbf{2-WL-GNN} ($\mean$)&{\tiny\Vectorstack{-1.7\\ \pm 3.7}} & {\tiny\Vectorstack{-3.2\\ \pm 6.4}} & {\tiny\Vectorstack{-1.0\\ \pm 3.5}} & {\tiny\Vectorstack{+0.8\\ \pm 3.1}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+21\\ \pm 6.4}}} & {\tiny\Vectorstack{+4.4\\ \pm 7.9}} & {\tiny\Vectorstack{-0.2\\ \pm 2.8}} & {\tiny\Vectorstack{+0.3\\ \pm 3.5}} &  & {\tiny\Vectorstack{+0.1\\ \pm 3.4}} \\[2pt]
+\textbf{2-WL-GNN} ($\wmean$)&{\tiny\Vectorstack{-1.8\\ \pm 3.8}} & {\tiny\Vectorstack{-3.3\\ \pm 6.7}} & {\tiny\Vectorstack{-1.1\\ \pm 3.7}} & {\tiny\Vectorstack{+0.7\\ \pm 4.3}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+20\\ \pm 6.9}}} & {\tiny\Vectorstack{+4.3\\ \pm 8.4}} & {\tiny\Vectorstack{-0.2\\ \pm 2.4}} & {\tiny\Vectorstack{+0.2\\ \pm 2.7}} & {\tiny\Vectorstack{-0.1\\ \pm 3.4}} &
 \end{tabular}}

### 2020-09-02 thesis/data/diffs/nci.tex
added epoch timing evaluation infrastructure

--- 2020-04-12 thesis/data/diffs/nci.tex
@@
 {\setlength\tabcolsep{2.5pt}\setlength{\extrarowheight}{2pt}%
-\begin{tabular}{lcccccccccccc}
-& \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-GWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{Baseline} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{GIN} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)} \\
-\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)& & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 1.3}}} & {\tiny\Vectorstack{-2.8\\ \pm 1.5}} & {\tiny\Vectorstack{+2.3\\ \pm 2.4}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.2\\ \pm 2.9}}} & {\tiny\Vectorstack{-3.5\\ \pm 2.2}} & {\tiny\Vectorstack{-1.9\\ \pm 2.9}} & {\tiny\Vectorstack{-4.4\\ \pm 2.3}} & {\tiny\Vectorstack{+3.1\\ \pm 2.5}} & {\tiny\Vectorstack{+1.6\\ \pm 2.6}} & {\tiny\Vectorstack{+3.7\\ \pm 3.6}} & {\tiny\Vectorstack{+0.2\\ \pm 3.5}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=3$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 1.3}}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.1\\ \pm 1.6}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+13\\ \pm 1.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+17\\ \pm 2.9}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+7.3\\ \pm 2.7}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.9\\ \pm 2.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.5\\ \pm 2.0}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+14\\ \pm 2.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+12\\ \pm 2.5}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+15\\ \pm 4.2}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 3.4}}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)&{\tiny\Vectorstack{+2.8\\ \pm 1.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.1\\ \pm 1.6}}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+5.1\\ \pm 2.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+9.0\\ \pm 3.0}}} & {\tiny\Vectorstack{-0.7\\ \pm 2.7}} & {\tiny\Vectorstack{+0.9\\ \pm 2.0}} & {\tiny\Vectorstack{-1.6\\ \pm 2.3}} & {\tiny\Vectorstack{+5.9\\ \pm 3.2}} & {\tiny\Vectorstack{+4.4\\ \pm 2.5}} & {\tiny\Vectorstack{+6.5\\ \pm 4.2}} & {\tiny\Vectorstack{+3.0\\ \pm 4.0}} \\[2pt]
-\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-2.3\\ \pm 2.4}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-13\\ \pm 1.8}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-5.1\\ \pm 2.3}}} &  & {\tiny\Vectorstack{+3.9\\ \pm 4.1}} & {\tiny\Vectorstack{-5.8\\ \pm 3.6}} & {\tiny\Vectorstack{-4.3\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.7\\ \pm 2.5}}} & {\tiny\Vectorstack{+0.8\\ \pm 3.7}} & {\tiny\Vectorstack{-0.8\\ \pm 2.7}} & {\tiny\Vectorstack{+1.4\\ \pm 4.0}} & {\tiny\Vectorstack{-2.1\\ \pm 4.7}} \\[2pt]
-\textbf{Baseline} ($\mathrm{sum}$)&\cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.2\\ \pm 2.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-17\\ \pm 2.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-9.0\\ \pm 3.0}}} & {\tiny\Vectorstack{-3.9\\ \pm 4.1}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-9.8\\ \pm 2.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.2\\ \pm 2.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 2.6}}} & {\tiny\Vectorstack{-3.1\\ \pm 3.0}} & {\tiny\Vectorstack{-4.7\\ \pm 3.1}} & {\tiny\Vectorstack{-2.6\\ \pm 5.0}} & {\tiny\Vectorstack{-6.1\\ \pm 4.4}} \\[2pt]
-\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{+3.5\\ \pm 2.2}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-7.3\\ \pm 2.7}}} & {\tiny\Vectorstack{+0.7\\ \pm 2.7}} & {\tiny\Vectorstack{+5.8\\ \pm 3.6}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+9.8\\ \pm 2.4}}} &  & {\tiny\Vectorstack{+1.6\\ \pm 3.5}} & {\tiny\Vectorstack{-0.9\\ \pm 2.2}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.6\\ \pm 2.5}}} & {\tiny\Vectorstack{+5.1\\ \pm 3.2}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+7.2\\ \pm 3.3}}} & {\tiny\Vectorstack{+3.7\\ \pm 4.8}} \\[2pt]
-\textbf{2-GNN} ($\mean$)&{\tiny\Vectorstack{+1.9\\ \pm 2.9}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.9\\ \pm 2.3}}} & {\tiny\Vectorstack{-0.9\\ \pm 2.0}} & {\tiny\Vectorstack{+4.3\\ \pm 2.5}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.2\\ \pm 2.9}}} & {\tiny\Vectorstack{-1.6\\ \pm 3.5}} &  & {\tiny\Vectorstack{-2.4\\ \pm 2.6}} & {\tiny\Vectorstack{+5.1\\ \pm 3.5}} & {\tiny\Vectorstack{+3.5\\ \pm 2.2}} & {\tiny\Vectorstack{+5.6\\ \pm 4.8}} & {\tiny\Vectorstack{+2.1\\ \pm 4.7}} \\[2pt]
-\textbf{2-GNN} ($\mathrm{SAM}$)&{\tiny\Vectorstack{+4.4\\ \pm 2.3}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.5\\ \pm 2.0}}} & {\tiny\Vectorstack{+1.6\\ \pm 2.3}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.7\\ \pm 2.5}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 2.6}}} & {\tiny\Vectorstack{+0.9\\ \pm 2.2}} & {\tiny\Vectorstack{+2.4\\ \pm 2.6}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+7.5\\ \pm 3.1}}} & {\tiny\Vectorstack{+5.9\\ \pm 3.2}} & {\tiny\Vectorstack{+8.1\\ \pm 4.2}} & {\tiny\Vectorstack{+4.6\\ \pm 4.9}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)&{\tiny\Vectorstack{-3.1\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-14\\ \pm 2.8}}} & {\tiny\Vectorstack{-5.9\\ \pm 3.2}} & {\tiny\Vectorstack{-0.8\\ \pm 3.7}} & {\tiny\Vectorstack{+3.1\\ \pm 3.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.6\\ \pm 2.5}}} & {\tiny\Vectorstack{-5.1\\ \pm 3.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-7.5\\ \pm 3.1}}} &  & {\tiny\Vectorstack{-1.5\\ \pm 2.2}} & {\tiny\Vectorstack{+0.6\\ \pm 2.9}} & {\tiny\Vectorstack{-2.9\\ \pm 4.5}} \\[2pt]
-\textbf{2-WL-GNN\phantom{*}} ($\mean$)&{\tiny\Vectorstack{-1.6\\ \pm 2.6}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-12\\ \pm 2.5}}} & {\tiny\Vectorstack{-4.4\\ \pm 2.5}} & {\tiny\Vectorstack{+0.8\\ \pm 2.7}} & {\tiny\Vectorstack{+4.7\\ \pm 3.1}} & {\tiny\Vectorstack{-5.1\\ \pm 3.2}} & {\tiny\Vectorstack{-3.5\\ \pm 2.2}} & {\tiny\Vectorstack{-5.9\\ \pm 3.2}} & {\tiny\Vectorstack{+1.5\\ \pm 2.2}} &  & {\tiny\Vectorstack{+2.1\\ \pm 3.4}} & {\tiny\Vectorstack{-1.4\\ \pm 4.3}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{-3.7\\ \pm 3.6}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-15\\ \pm 4.2}}} & {\tiny\Vectorstack{-6.5\\ \pm 4.2}} & {\tiny\Vectorstack{-1.4\\ \pm 4.0}} & {\tiny\Vectorstack{+2.6\\ \pm 5.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-7.2\\ \pm 3.3}}} & {\tiny\Vectorstack{-5.6\\ \pm 4.8}} & {\tiny\Vectorstack{-8.1\\ \pm 4.2}} & {\tiny\Vectorstack{-0.6\\ \pm 2.9}} & {\tiny\Vectorstack{-2.1\\ \pm 3.4}} &  & {\tiny\Vectorstack{-3.5\\ \pm 6.2}} \\[2pt]
-\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{-0.2\\ \pm 3.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 3.4}}} & {\tiny\Vectorstack{-3.0\\ \pm 4.0}} & {\tiny\Vectorstack{+2.1\\ \pm 4.7}} & {\tiny\Vectorstack{+6.1\\ \pm 4.4}} & {\tiny\Vectorstack{-3.7\\ \pm 4.8}} & {\tiny\Vectorstack{-2.1\\ \pm 4.7}} & {\tiny\Vectorstack{-4.6\\ \pm 4.9}} & {\tiny\Vectorstack{+2.9\\ \pm 4.5}} & {\tiny\Vectorstack{+1.4\\ \pm 4.3}} & {\tiny\Vectorstack{+3.5\\ \pm 6.2}} &
+\begin{tabular}{lcccccccccc}
+& \rotatebox[origin=l]{90}{\textbf{WL\textsubscript{ST}} ($T=1$)} & \rotatebox[origin=l]{90}{\textbf{WL\textsubscript{ST}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-LWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-GWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{Baseline} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{GIN} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\wmean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN} ($\wmean$)} \\
+\textbf{WL\textsubscript{ST}} ($T=1$)& & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 1.3}}} & {\tiny\Vectorstack{-2.8\\ \pm 1.5}} & {\tiny\Vectorstack{+2.3\\ \pm 2.4}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.2\\ \pm 2.9}}} & {\tiny\Vectorstack{-3.5\\ \pm 2.2}} & {\tiny\Vectorstack{-1.9\\ \pm 2.9}} & {\tiny\Vectorstack{-4.4\\ \pm 2.3}} & {\tiny\Vectorstack{+1.6\\ \pm 2.6}} & {\tiny\Vectorstack{+0.4\\ \pm 2.1}} \\[2pt]
+\textbf{WL\textsubscript{ST}} ($T=3$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 1.3}}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.1\\ \pm 1.6}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+13\\ \pm 1.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+17\\ \pm 2.9}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+7.3\\ \pm 2.7}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.9\\ \pm 2.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.5\\ \pm 2.0}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+12\\ \pm 2.5}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 2.0}}} \\[2pt]
+\textbf{2-LWL} ($T=3$)&{\tiny\Vectorstack{+2.8\\ \pm 1.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.1\\ \pm 1.6}}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+5.1\\ \pm 2.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+9.0\\ \pm 3.0}}} & {\tiny\Vectorstack{-0.7\\ \pm 2.7}} & {\tiny\Vectorstack{+0.9\\ \pm 2.0}} & {\tiny\Vectorstack{-1.6\\ \pm 2.3}} & {\tiny\Vectorstack{+4.4\\ \pm 2.5}} & {\tiny\Vectorstack{+3.2\\ \pm 2.5}} \\[2pt]
+\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-2.3\\ \pm 2.4}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-13\\ \pm 1.8}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-5.1\\ \pm 2.3}}} &  & {\tiny\Vectorstack{+3.9\\ \pm 4.1}} & {\tiny\Vectorstack{-5.8\\ \pm 3.6}} & {\tiny\Vectorstack{-4.3\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.7\\ \pm 2.5}}} & {\tiny\Vectorstack{-0.8\\ \pm 2.7}} & {\tiny\Vectorstack{-1.9\\ \pm 3.1}} \\[2pt]
+\textbf{Baseline} ($\mathrm{sum}$)&\cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.2\\ \pm 2.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-17\\ \pm 2.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-9.0\\ \pm 3.0}}} & {\tiny\Vectorstack{-3.9\\ \pm 4.1}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-9.8\\ \pm 2.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.2\\ \pm 2.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 2.6}}} & {\tiny\Vectorstack{-4.7\\ \pm 3.1}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-5.8\\ \pm 2.8}}} \\[2pt]
+\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{+3.5\\ \pm 2.2}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-7.3\\ \pm 2.7}}} & {\tiny\Vectorstack{+0.7\\ \pm 2.7}} & {\tiny\Vectorstack{+5.8\\ \pm 3.6}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+9.8\\ \pm 2.4}}} &  & {\tiny\Vectorstack{+1.6\\ \pm 3.5}} & {\tiny\Vectorstack{-0.9\\ \pm 2.2}} & {\tiny\Vectorstack{+5.1\\ \pm 3.2}} & {\tiny\Vectorstack{+3.9\\ \pm 3.3}} \\[2pt]
+\textbf{2-GNN} ($\mean$)&{\tiny\Vectorstack{+1.9\\ \pm 2.9}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.9\\ \pm 2.3}}} & {\tiny\Vectorstack{-0.9\\ \pm 2.0}} & {\tiny\Vectorstack{+4.3\\ \pm 2.5}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.2\\ \pm 2.9}}} & {\tiny\Vectorstack{-1.6\\ \pm 3.5}} &  & {\tiny\Vectorstack{-2.4\\ \pm 2.6}} & {\tiny\Vectorstack{+3.5\\ \pm 2.2}} & {\tiny\Vectorstack{+2.4\\ \pm 2.6}} \\[2pt]
+\textbf{2-GNN} ($\wmean$)&{\tiny\Vectorstack{+4.4\\ \pm 2.3}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.5\\ \pm 2.0}}} & {\tiny\Vectorstack{+1.6\\ \pm 2.3}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.7\\ \pm 2.5}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 2.6}}} & {\tiny\Vectorstack{+0.9\\ \pm 2.2}} & {\tiny\Vectorstack{+2.4\\ \pm 2.6}} &  & {\tiny\Vectorstack{+5.9\\ \pm 3.2}} & {\tiny\Vectorstack{+4.8\\ \pm 3.2}} \\[2pt]
+\textbf{2-WL-GNN} ($\mean$)&{\tiny\Vectorstack{-1.6\\ \pm 2.6}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-12\\ \pm 2.5}}} & {\tiny\Vectorstack{-4.4\\ \pm 2.5}} & {\tiny\Vectorstack{+0.8\\ \pm 2.7}} & {\tiny\Vectorstack{+4.7\\ \pm 3.1}} & {\tiny\Vectorstack{-5.1\\ \pm 3.2}} & {\tiny\Vectorstack{-3.5\\ \pm 2.2}} & {\tiny\Vectorstack{-5.9\\ \pm 3.2}} &  & {\tiny\Vectorstack{-1.1\\ \pm 2.0}} \\[2pt]
+\textbf{2-WL-GNN} ($\wmean$)&{\tiny\Vectorstack{-0.4\\ \pm 2.1}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 2.0}}} & {\tiny\Vectorstack{-3.2\\ \pm 2.5}} & {\tiny\Vectorstack{+1.9\\ \pm 3.1}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+5.8\\ \pm 2.8}}} & {\tiny\Vectorstack{-3.9\\ \pm 3.3}} & {\tiny\Vectorstack{-2.4\\ \pm 2.6}} & {\tiny\Vectorstack{-4.8\\ \pm 3.2}} & {\tiny\Vectorstack{+1.1\\ \pm 2.0}} &
 \end{tabular}}

### 2020-09-02 thesis/data/diffs/reddit.tex
added epoch timing evaluation infrastructure

@@
 {\setlength\tabcolsep{2.5pt}\setlength{\extrarowheight}{2pt}%
-\begin{tabular}{lcccccccccc}
-& \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-GWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{Baseline} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{GIN} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)} \\
-\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)& & {\tiny\Vectorstack{-1.8\\ \pm 1.8}} & {\tiny\Vectorstack{+0.5\\ \pm 4.9}} & {\tiny\Vectorstack{+0.8\\ \pm 3.6}} & {\tiny\Vectorstack{+4.2\\ \pm 10}} & {\tiny\Vectorstack{-11\\ \pm 6.9}} & {\tiny\Vectorstack{+2.9\\ \pm 3.5}} & {\tiny\Vectorstack{-7.4\\ \pm 6.2}} & {\tiny\Vectorstack{-0.9\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-13\\ \pm 3.3}}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=3$)&{\tiny\Vectorstack{+1.8\\ \pm 1.8}} &  & {\tiny\Vectorstack{+2.3\\ \pm 4.7}} & {\tiny\Vectorstack{+2.6\\ \pm 3.5}} & {\tiny\Vectorstack{+5.9\\ \pm 11}} & {\tiny\Vectorstack{-9.0\\ \pm 7.1}} & {\tiny\Vectorstack{+4.7\\ \pm 3.3}} & {\tiny\Vectorstack{-5.7\\ \pm 6.8}} & {\tiny\Vectorstack{+0.9\\ \pm 1.6}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 2.6}}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)&{\tiny\Vectorstack{-0.5\\ \pm 4.9}} & {\tiny\Vectorstack{-2.3\\ \pm 4.7}} &  & {\tiny\Vectorstack{+0.3\\ \pm 4.2}} & {\tiny\Vectorstack{+3.7\\ \pm 11}} & {\tiny\Vectorstack{-11\\ \pm 7.3}} & {\tiny\Vectorstack{+2.4\\ \pm 5.5}} & {\tiny\Vectorstack{-7.9\\ \pm 5.5}} & {\tiny\Vectorstack{-1.4\\ \pm 5.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-13\\ \pm 3.3}}} \\[2pt]
-\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-0.8\\ \pm 3.6}} & {\tiny\Vectorstack{-2.6\\ \pm 3.5}} & {\tiny\Vectorstack{-0.3\\ \pm 4.2}} &  & {\tiny\Vectorstack{+3.3\\ \pm 11}} & {\tiny\Vectorstack{-12\\ \pm 7.6}} & {\tiny\Vectorstack{+2.1\\ \pm 2.4}} & {\tiny\Vectorstack{-8.2\\ \pm 6.2}} & {\tiny\Vectorstack{-1.7\\ \pm 3.3}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-14\\ \pm 2.5}}} \\[2pt]
-\textbf{Baseline} ($\mathrm{sum}$)&{\tiny\Vectorstack{-4.2\\ \pm 10}} & {\tiny\Vectorstack{-5.9\\ \pm 11}} & {\tiny\Vectorstack{-3.7\\ \pm 11}} & {\tiny\Vectorstack{-3.3\\ \pm 11}} &  & {\tiny\Vectorstack{-15\\ \pm 12}} & {\tiny\Vectorstack{-1.2\\ \pm 12}} & {\tiny\Vectorstack{-12\\ \pm 13}} & {\tiny\Vectorstack{-5.0\\ \pm 11}} & {\tiny\Vectorstack{-17\\ \pm 10}} \\[2pt]
-\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{+11\\ \pm 6.9}} & {\tiny\Vectorstack{+9.0\\ \pm 7.1}} & {\tiny\Vectorstack{+11\\ \pm 7.3}} & {\tiny\Vectorstack{+12\\ \pm 7.6}} & {\tiny\Vectorstack{+15\\ \pm 12}} &  & {\tiny\Vectorstack{+14\\ \pm 8.3}} & {\tiny\Vectorstack{+3.3\\ \pm 9.6}} & {\tiny\Vectorstack{+9.9\\ \pm 7.4}} & {\tiny\Vectorstack{-2.2\\ \pm 6.8}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)&{\tiny\Vectorstack{-2.9\\ \pm 3.5}} & {\tiny\Vectorstack{-4.7\\ \pm 3.3}} & {\tiny\Vectorstack{-2.4\\ \pm 5.5}} & {\tiny\Vectorstack{-2.1\\ \pm 2.4}} & {\tiny\Vectorstack{+1.2\\ \pm 12}} & {\tiny\Vectorstack{-14\\ \pm 8.3}} &  & {\tiny\Vectorstack{-10\\ \pm 7.1}} & {\tiny\Vectorstack{-3.8\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-16\\ \pm 2.8}}} \\[2pt]
-\textbf{2-WL-GNN\phantom{*}} ($\mean$)&{\tiny\Vectorstack{+7.4\\ \pm 6.2}} & {\tiny\Vectorstack{+5.7\\ \pm 6.8}} & {\tiny\Vectorstack{+7.9\\ \pm 5.5}} & {\tiny\Vectorstack{+8.2\\ \pm 6.2}} & {\tiny\Vectorstack{+12\\ \pm 13}} & {\tiny\Vectorstack{-3.3\\ \pm 9.6}} & {\tiny\Vectorstack{+10\\ \pm 7.1}} &  & {\tiny\Vectorstack{+6.6\\ \pm 7.2}} & {\tiny\Vectorstack{-5.5\\ \pm 6.1}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{+0.9\\ \pm 2.5}} & {\tiny\Vectorstack{-0.9\\ \pm 1.6}} & {\tiny\Vectorstack{+1.4\\ \pm 5.0}} & {\tiny\Vectorstack{+1.7\\ \pm 3.3}} & {\tiny\Vectorstack{+5.0\\ \pm 11}} & {\tiny\Vectorstack{-9.9\\ \pm 7.4}} & {\tiny\Vectorstack{+3.8\\ \pm 2.5}} & {\tiny\Vectorstack{-6.6\\ \pm 7.2}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-12\\ \pm 2.7}}} \\[2pt]
-\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+13\\ \pm 3.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 2.6}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+13\\ \pm 3.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+14\\ \pm 2.5}}} & {\tiny\Vectorstack{+17\\ \pm 10}} & {\tiny\Vectorstack{+2.2\\ \pm 6.8}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+16\\ \pm 2.8}}} & {\tiny\Vectorstack{+5.5\\ \pm 6.1}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+12\\ \pm 2.7}}} &
+\begin{tabular}{lcccccccc}
+& \rotatebox[origin=l]{90}{\textbf{WL\textsubscript{ST}} ($T=1$)} & \rotatebox[origin=l]{90}{\textbf{WL\textsubscript{ST}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-LWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-GWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{Baseline} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{GIN} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN} ($\wmean$)} \\
+\textbf{WL\textsubscript{ST}} ($T=1$)& & {\tiny\Vectorstack{-1.8\\ \pm 1.8}} & {\tiny\Vectorstack{+0.5\\ \pm 4.9}} & {\tiny\Vectorstack{+0.8\\ \pm 3.6}} & {\tiny\Vectorstack{+4.2\\ \pm 10}} & {\tiny\Vectorstack{-11\\ \pm 6.9}} & {\tiny\Vectorstack{-7.4\\ \pm 6.2}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-13\\ \pm 3.5}}} \\[2pt]
+\textbf{WL\textsubscript{ST}} ($T=3$)&{\tiny\Vectorstack{+1.8\\ \pm 1.8}} &  & {\tiny\Vectorstack{+2.3\\ \pm 4.7}} & {\tiny\Vectorstack{+2.6\\ \pm 3.5}} & {\tiny\Vectorstack{+5.9\\ \pm 11}} & {\tiny\Vectorstack{-9.0\\ \pm 7.1}} & {\tiny\Vectorstack{-5.7\\ \pm 6.8}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 2.9}}} \\[2pt]
+\textbf{2-LWL} ($T=3$)&{\tiny\Vectorstack{-0.5\\ \pm 4.9}} & {\tiny\Vectorstack{-2.3\\ \pm 4.7}} &  & {\tiny\Vectorstack{+0.3\\ \pm 4.2}} & {\tiny\Vectorstack{+3.7\\ \pm 11}} & {\tiny\Vectorstack{-11\\ \pm 7.3}} & {\tiny\Vectorstack{-7.9\\ \pm 5.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-14\\ \pm 3.4}}} \\[2pt]
+\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-0.8\\ \pm 3.6}} & {\tiny\Vectorstack{-2.6\\ \pm 3.5}} & {\tiny\Vectorstack{-0.3\\ \pm 4.2}} &  & {\tiny\Vectorstack{+3.3\\ \pm 11}} & {\tiny\Vectorstack{-12\\ \pm 7.6}} & {\tiny\Vectorstack{-8.2\\ \pm 6.2}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-14\\ \pm 2.7}}} \\[2pt]
+\textbf{Baseline} ($\mathrm{sum}$)&{\tiny\Vectorstack{-4.2\\ \pm 10}} & {\tiny\Vectorstack{-5.9\\ \pm 11}} & {\tiny\Vectorstack{-3.7\\ \pm 11}} & {\tiny\Vectorstack{-3.3\\ \pm 11}} &  & {\tiny\Vectorstack{-15\\ \pm 12}} & {\tiny\Vectorstack{-12\\ \pm 13}} & {\tiny\Vectorstack{-17\\ \pm 10}} \\[2pt]
+\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{+11\\ \pm 6.9}} & {\tiny\Vectorstack{+9.0\\ \pm 7.1}} & {\tiny\Vectorstack{+11\\ \pm 7.3}} & {\tiny\Vectorstack{+12\\ \pm 7.6}} & {\tiny\Vectorstack{+15\\ \pm 12}} &  & {\tiny\Vectorstack{+3.3\\ \pm 9.6}} & {\tiny\Vectorstack{-2.4\\ \pm 7.0}} \\[2pt]
+\textbf{2-WL-GNN} ($\mean$)&{\tiny\Vectorstack{+7.4\\ \pm 6.2}} & {\tiny\Vectorstack{+5.7\\ \pm 6.8}} & {\tiny\Vectorstack{+7.9\\ \pm 5.5}} & {\tiny\Vectorstack{+8.2\\ \pm 6.2}} & {\tiny\Vectorstack{+12\\ \pm 13}} & {\tiny\Vectorstack{-3.3\\ \pm 9.6}} &  & {\tiny\Vectorstack{-5.7\\ \pm 6.5}} \\[2pt]
+\textbf{2-WL-GNN} ($\wmean$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+13\\ \pm 3.5}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 2.9}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+14\\ \pm 3.4}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+14\\ \pm 2.7}}} & {\tiny\Vectorstack{+17\\ \pm 10}} & {\tiny\Vectorstack{+2.4\\ \pm 7.0}} & {\tiny\Vectorstack{+5.7\\ \pm 6.5}} &
 \end{tabular}}

### 2020-09-02 thesis/data/diffs/triangle.tex
added epoch timing evaluation infrastructure

@@
 {\setlength\tabcolsep{2.5pt}\setlength{\extrarowheight}{2pt}%
-\begin{tabular}{lccccccccccccc}
-& \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=5$)} & \rotatebox[origin=l]{90}{\textbf{WL\textsubscript{SP}} ($T=5$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-GWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{Baseline} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{GIN} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)} \\
-\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)& & {\tiny\Vectorstack{+2.2\\ \pm 9.2}} & {\tiny\Vectorstack{-4.9\\ \pm 12}} & {\tiny\Vectorstack{+8.2\\ \pm 15}} & {\tiny\Vectorstack{+2.9\\ \pm 12}} & {\tiny\Vectorstack{+20\\ \pm 19}} & {\tiny\Vectorstack{-5.2\\ \pm 16}} & {\tiny\Vectorstack{-12\\ \pm 18}} & {\tiny\Vectorstack{-17\\ \pm 19}} & {\tiny\Vectorstack{-26\\ \pm 15}} & {\tiny\Vectorstack{-28\\ \pm 17}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-26\\ \pm 13}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-35\\ \pm 13}}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=5$)&{\tiny\Vectorstack{-2.2\\ \pm 9.2}} &  & {\tiny\Vectorstack{-7.1\\ \pm 14}} & {\tiny\Vectorstack{+6.1\\ \pm 9.8}} & {\tiny\Vectorstack{+0.8\\ \pm 13}} & {\tiny\Vectorstack{+18\\ \pm 14}} & {\tiny\Vectorstack{-7.4\\ \pm 11}} & {\tiny\Vectorstack{-14\\ \pm 16}} & {\tiny\Vectorstack{-19\\ \pm 16}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-28\\ \pm 14}}} & {\tiny\Vectorstack{-30\\ \pm 17}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-28\\ \pm 9.6}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-37\\ \pm 11}}} \\[2pt]
-\textbf{WL\textsubscript{SP}} ($T=5$)&{\tiny\Vectorstack{+4.9\\ \pm 12}} & {\tiny\Vectorstack{+7.1\\ \pm 14}} &  & {\tiny\Vectorstack{+13\\ \pm 13}} & {\tiny\Vectorstack{+7.9\\ \pm 10}} & {\tiny\Vectorstack{+25\\ \pm 16}} & {\tiny\Vectorstack{-0.3\\ \pm 13}} & {\tiny\Vectorstack{-7.1\\ \pm 10}} & {\tiny\Vectorstack{-12\\ \pm 12}} & {\tiny\Vectorstack{-21\\ \pm 11}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-23\\ \pm 11}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 10}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-30\\ \pm 8.9}}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)&{\tiny\Vectorstack{-8.2\\ \pm 15}} & {\tiny\Vectorstack{-6.1\\ \pm 9.8}} & {\tiny\Vectorstack{-13\\ \pm 13}} &  & {\tiny\Vectorstack{-5.3\\ \pm 11}} & {\tiny\Vectorstack{+12\\ \pm 7.5}} & {\tiny\Vectorstack{-13\\ \pm 6.7}} & {\tiny\Vectorstack{-20\\ \pm 13}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-25\\ \pm 9.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-34\\ \pm 8.6}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-36\\ \pm 10}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-34\\ \pm 7.0}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-43\\ \pm 6.8}}} \\[2pt]
-\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-2.9\\ \pm 12}} & {\tiny\Vectorstack{-0.8\\ \pm 13}} & {\tiny\Vectorstack{-7.9\\ \pm 10}} & {\tiny\Vectorstack{+5.3\\ \pm 11}} &  & {\tiny\Vectorstack{+17\\ \pm 15}} & {\tiny\Vectorstack{-8.2\\ \pm 13}} & {\tiny\Vectorstack{-15\\ \pm 14}} & {\tiny\Vectorstack{-20\\ \pm 13}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-29\\ \pm 8.8}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-31\\ \pm 9.0}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-29\\ \pm 10}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-38\\ \pm 9.0}}} \\[2pt]
-\textbf{Baseline} ($\mathrm{sum}$)&{\tiny\Vectorstack{-20\\ \pm 19}} & {\tiny\Vectorstack{-18\\ \pm 14}} & {\tiny\Vectorstack{-25\\ \pm 16}} & {\tiny\Vectorstack{-12\\ \pm 7.5}} & {\tiny\Vectorstack{-17\\ \pm 15}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-25\\ \pm 8.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-32\\ \pm 14}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-37\\ \pm 9.2}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-46\\ \pm 9.8}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-48\\ \pm 14}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-46\\ \pm 9.2}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-55\\ \pm 8.5}}} \\[2pt]
-\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{+5.2\\ \pm 16}} & {\tiny\Vectorstack{+7.4\\ \pm 11}} & {\tiny\Vectorstack{+0.3\\ \pm 13}} & {\tiny\Vectorstack{+13\\ \pm 6.7}} & {\tiny\Vectorstack{+8.2\\ \pm 13}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+25\\ \pm 8.4}}} &  & {\tiny\Vectorstack{-6.8\\ \pm 13}} & {\tiny\Vectorstack{-12\\ \pm 9.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 9.9}}} & {\tiny\Vectorstack{-23\\ \pm 14}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 8.3}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-29\\ \pm 7.3}}} \\[2pt]
-\textbf{2-GNN} ($\mean$)&{\tiny\Vectorstack{+12\\ \pm 18}} & {\tiny\Vectorstack{+14\\ \pm 16}} & {\tiny\Vectorstack{+7.1\\ \pm 10}} & {\tiny\Vectorstack{+20\\ \pm 13}} & {\tiny\Vectorstack{+15\\ \pm 14}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+32\\ \pm 14}}} & {\tiny\Vectorstack{+6.8\\ \pm 13}} &  & {\tiny\Vectorstack{-5.0\\ \pm 6.6}} & {\tiny\Vectorstack{-14\\ \pm 9.6}} & {\tiny\Vectorstack{-16\\ \pm 11}} & {\tiny\Vectorstack{-14\\ \pm 8.8}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-23\\ \pm 9.8}}} \\[2pt]
-\textbf{2-GNN} ($\mathrm{SAM}$)&{\tiny\Vectorstack{+17\\ \pm 19}} & {\tiny\Vectorstack{+19\\ \pm 16}} & {\tiny\Vectorstack{+12\\ \pm 12}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+25\\ \pm 9.4}}} & {\tiny\Vectorstack{+20\\ \pm 13}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+37\\ \pm 9.2}}} & {\tiny\Vectorstack{+12\\ \pm 9.5}} & {\tiny\Vectorstack{+5.0\\ \pm 6.6}} &  & {\tiny\Vectorstack{-8.9\\ \pm 6.3}} & {\tiny\Vectorstack{-11\\ \pm 8.8}} & {\tiny\Vectorstack{-9.2\\ \pm 7.1}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-18\\ \pm 7.2}}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)&{\tiny\Vectorstack{+26\\ \pm 15}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+28\\ \pm 14}}} & {\tiny\Vectorstack{+21\\ \pm 11}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+34\\ \pm 8.6}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+29\\ \pm 8.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+46\\ \pm 9.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+21\\ \pm 9.9}}} & {\tiny\Vectorstack{+14\\ \pm 9.6}} & {\tiny\Vectorstack{+8.9\\ \pm 6.3}} &  & {\tiny\Vectorstack{-2.2\\ \pm 7.2}} & {\tiny\Vectorstack{-0.3\\ \pm 6.0}} & {\tiny\Vectorstack{-8.7\\ \pm 4.6}} \\[2pt]
-\textbf{2-WL-GNN\phantom{*}} ($\mean$)&{\tiny\Vectorstack{+28\\ \pm 17}} & {\tiny\Vectorstack{+30\\ \pm 17}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+23\\ \pm 11}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+36\\ \pm 10}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+31\\ \pm 9.0}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+48\\ \pm 14}}} & {\tiny\Vectorstack{+23\\ \pm 14}} & {\tiny\Vectorstack{+16\\ \pm 11}} & {\tiny\Vectorstack{+11\\ \pm 8.8}} & {\tiny\Vectorstack{+2.2\\ \pm 7.2}} &  & {\tiny\Vectorstack{+1.9\\ \pm 9.7}} & {\tiny\Vectorstack{-6.5\\ \pm 8.7}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+26\\ \pm 13}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+28\\ \pm 9.6}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+21\\ \pm 10}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+34\\ \pm 7.0}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+29\\ \pm 10}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+46\\ \pm 9.2}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+21\\ \pm 8.3}}} & {\tiny\Vectorstack{+14\\ \pm 8.8}} & {\tiny\Vectorstack{+9.2\\ \pm 7.1}} & {\tiny\Vectorstack{+0.3\\ \pm 6.0}} & {\tiny\Vectorstack{-1.9\\ \pm 9.7}} &  & {\tiny\Vectorstack{-8.5\\ \pm 4.7}} \\[2pt]
-\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+35\\ \pm 13}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+37\\ \pm 11}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+30\\ \pm 8.9}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+43\\ \pm 6.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+38\\ \pm 9.0}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+55\\ \pm 8.5}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+29\\ \pm 7.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+23\\ \pm 9.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+18\\ \pm 7.2}}} & {\tiny\Vectorstack{+8.7\\ \pm 4.6}} & {\tiny\Vectorstack{+6.5\\ \pm 8.7}} & {\tiny\Vectorstack{+8.5\\ \pm 4.7}} &
+\begin{tabular}{lcccccccccc}
+& \rotatebox[origin=l]{90}{\textbf{WL\textsubscript{ST}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{WL\textsubscript{SP}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-LWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-GWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{Baseline} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{GIN} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\wmean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN} ($\wmean$)} \\
+\textbf{WL\textsubscript{ST}} ($T=3$)& & {\tiny\Vectorstack{-11\\ \pm 13}} & {\tiny\Vectorstack{+0.4\\ \pm 9.6}} & {\tiny\Vectorstack{-4.9\\ \pm 16}} & {\tiny\Vectorstack{+12\\ \pm 12}} & {\tiny\Vectorstack{-13\\ \pm 10}} & {\tiny\Vectorstack{-20\\ \pm 14}} & {\tiny\Vectorstack{-25\\ \pm 14}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-36\\ \pm 17}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-42\\ \pm 11}}} \\[2pt]
+\textbf{WL\textsubscript{SP}} ($T=3$)&{\tiny\Vectorstack{+11\\ \pm 13}} &  & {\tiny\Vectorstack{+11\\ \pm 13}} & {\tiny\Vectorstack{+6.1\\ \pm 9.2}} & {\tiny\Vectorstack{+23\\ \pm 16}} & {\tiny\Vectorstack{-2.0\\ \pm 14}} & {\tiny\Vectorstack{-8.8\\ \pm 13}} & {\tiny\Vectorstack{-14\\ \pm 14}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-25\\ \pm 12}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-31\\ \pm 11}}} \\[2pt]
+\textbf{2-LWL} ($T=3$)&{\tiny\Vectorstack{-0.4\\ \pm 9.6}} & {\tiny\Vectorstack{-11\\ \pm 13}} &  & {\tiny\Vectorstack{-5.3\\ \pm 11}} & {\tiny\Vectorstack{+12\\ \pm 7.5}} & {\tiny\Vectorstack{-13\\ \pm 6.7}} & {\tiny\Vectorstack{-20\\ \pm 13}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-25\\ \pm 9.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-36\\ \pm 10}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-43\\ \pm 6.8}}} \\[2pt]
+\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{+4.9\\ \pm 16}} & {\tiny\Vectorstack{-6.1\\ \pm 9.2}} & {\tiny\Vectorstack{+5.3\\ \pm 11}} &  & {\tiny\Vectorstack{+17\\ \pm 15}} & {\tiny\Vectorstack{-8.2\\ \pm 13}} & {\tiny\Vectorstack{-15\\ \pm 14}} & {\tiny\Vectorstack{-20\\ \pm 13}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-31\\ \pm 9.0}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-38\\ \pm 9.0}}} \\[2pt]
+\textbf{Baseline} ($\mathrm{sum}$)&{\tiny\Vectorstack{-12\\ \pm 12}} & {\tiny\Vectorstack{-23\\ \pm 16}} & {\tiny\Vectorstack{-12\\ \pm 7.5}} & {\tiny\Vectorstack{-17\\ \pm 15}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-25\\ \pm 8.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-32\\ \pm 14}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-37\\ \pm 9.2}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-48\\ \pm 14}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-55\\ \pm 8.5}}} \\[2pt]
+\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{+13\\ \pm 10}} & {\tiny\Vectorstack{+2.0\\ \pm 14}} & {\tiny\Vectorstack{+13\\ \pm 6.7}} & {\tiny\Vectorstack{+8.2\\ \pm 13}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+25\\ \pm 8.4}}} &  & {\tiny\Vectorstack{-6.8\\ \pm 13}} & {\tiny\Vectorstack{-12\\ \pm 9.5}} & {\tiny\Vectorstack{-23\\ \pm 14}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-29\\ \pm 7.3}}} \\[2pt]
+\textbf{2-GNN} ($\mean$)&{\tiny\Vectorstack{+20\\ \pm 14}} & {\tiny\Vectorstack{+8.8\\ \pm 13}} & {\tiny\Vectorstack{+20\\ \pm 13}} & {\tiny\Vectorstack{+15\\ \pm 14}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+32\\ \pm 14}}} & {\tiny\Vectorstack{+6.8\\ \pm 13}} &  & {\tiny\Vectorstack{-5.0\\ \pm 6.6}} & {\tiny\Vectorstack{-16\\ \pm 11}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-23\\ \pm 9.8}}} \\[2pt]
+\textbf{2-GNN} ($\wmean$)&{\tiny\Vectorstack{+25\\ \pm 14}} & {\tiny\Vectorstack{+14\\ \pm 14}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+25\\ \pm 9.4}}} & {\tiny\Vectorstack{+20\\ \pm 13}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+37\\ \pm 9.2}}} & {\tiny\Vectorstack{+12\\ \pm 9.5}} & {\tiny\Vectorstack{+5.0\\ \pm 6.6}} &  & {\tiny\Vectorstack{-11\\ \pm 8.8}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-18\\ \pm 7.2}}} \\[2pt]
+\textbf{2-WL-GNN} ($\mean$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+36\\ \pm 17}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+25\\ \pm 12}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+36\\ \pm 10}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+31\\ \pm 9.0}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+48\\ \pm 14}}} & {\tiny\Vectorstack{+23\\ \pm 14}} & {\tiny\Vectorstack{+16\\ \pm 11}} & {\tiny\Vectorstack{+11\\ \pm 8.8}} &  & {\tiny\Vectorstack{-6.5\\ \pm 8.7}} \\[2pt]
+\textbf{2-WL-GNN} ($\wmean$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+42\\ \pm 11}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+31\\ \pm 11}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+43\\ \pm 6.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+38\\ \pm 9.0}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+55\\ \pm 8.5}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+29\\ \pm 7.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+23\\ \pm 9.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+18\\ \pm 7.2}}} & {\tiny\Vectorstack{+6.5\\ \pm 8.7}} &
 \end{tabular}}

### 2020-09-02 implementation/src/ltag/chaining/pipeline.py

@@ -9,7 +9,7 @@ def tolerant(f):
   if hasattr(f, "__tolerant__"):
     return f
 
-  spec = inspect.getfullargspec(f)
+  spec = inspect.getfullargspec(f.__init__ if inspect.isclass(f) else f)
   f_varkws = spec.varkw is not None
 
   if f_varkws:

### 2020-09-02 implementation/src/ltag/evaluate_datasets.py

@@ -39,6 +39,8 @@ def summarize(mf, dsm, **kwargs):
   return summary.summarize_evaluation(
     evaluate.find_eval_dir(mf, dsm), **kwargs)
 
+def epoch_times(mf, dsm, **kwargs):
+  return evaluate.evaluate_epoch_time(mf, dsm)
 
 if __name__ == "__main__":
   parser = argparse.ArgumentParser(
@@ -52,6 +54,9 @@ if __name__ == "__main__":
   parser.add_argument(
     "-q", "--quick", action="store_true",
     help="Only do a quick run to check whether there might be OOM issues.")
+  parser.add_argument(
+    "-t", "--time", action="store_true",
+    help="Only measure the epoch times of the first hp config.")
   parser.add_argument(
     "-s", "--summarize", action="store_true",
     help="Only run the summarizer on existing evaluations.")
@@ -69,6 +74,9 @@ if __name__ == "__main__":
   elif args.summarize:
     type = "summarization"
     f = summarize
+  elif args.time:
+    type = "epoch time measurement"
+    f = epoch_times
   elif args.dry:
     type = "dry evaluation"
     f = fy.partial(resume, dry=True)

### 2020-09-02 implementation/src/ltag/evaluation/evaluate.py

@@ -12,7 +12,7 @@ from tensorflow import keras
 import funcy as fy
 import tensorflow as tf
 
-from ltag.utils import NumpyEncoder
+from ltag.utils import statistics, NumpyEncoder
 import ltag.chaining.pipeline as cp
 import ltag.evaluation.summary as summary
 
@@ -28,12 +28,25 @@ log_dir_base = Path("../logs")
 def time_str():
   return datetime.now().strftime("%Y-%m-%d_%H-%M-%S")
 
+class EpochTimeCallback(keras.callbacks.Callback):
+  def __init__(self):
+    self.times = []
+    self.currTime = timer()
+
+  def on_epoch_end(self, epoch, logs={}):
+    currTime = timer()
+    self.times.append(currTime - self.currTime)
+    self.currTime = currTime
+
+  def on_train_end(self, logs={}):
+    pass
+
 def train(
   model, train_ds, val_ds=None, label=None,
   log_dir_base=log_dir_base,
   epochs=1000, patience=50, restore_best=False,
   stopping_min_delta=0.0001,
-  verbose=2):
+  verbose=2, measure_epoch_times=False):
   label = "_" + label if label is not None else ""
 
   t = time_str()
@@ -47,10 +60,21 @@ def train(
     min_delta=stopping_min_delta,
     restore_best_weights=restore_best)
 
-  return model.fit(
+  if measure_epoch_times:
+    tc = EpochTimeCallback()
+    callbacks = [tb, es, tc]
+  else:
+    callbacks = [tb, es]
+
+  hist = model.fit(
     train_ds, validation_data=val_ds,
-    epochs=epochs, callbacks=[tb, es],
+    epochs=epochs, callbacks=callbacks,
     verbose=verbose)
+
+  if measure_epoch_times:
+    return hist, tc.times
+  else:
+    return hist
 
 @cp.tolerant
 def evaluation_step(
@@ -345,3 +369,35 @@ def quick_evaluate(model_factory, ds_manager, **kwargs):
     model_factory, ds_manager,
     epochs=1, repeat=1, winner_repeat=1, label="quick",
     **fy.omit(kwargs, ["epochs", "repeat", "winner_repeat", "label"]))
+
+def evaluate_epoch_time(model_factory, ds_manager, hp_args=None):
+  eval_dir = find_eval_dir(model_factory, ds_manager, "time_eval")
+  if not eval_dir.exists():
+    os.makedirs(eval_dir)
+
[…]

### 2020-09-02 implementation/src/ltag/playground.py

@@ -179,7 +179,7 @@ def imdb_experient():
   print(model.evaluate(test))
 
 def wl2_power_experiment():
-  model_class = gnn_models.SagK2GNN
+  model_class = gnn_models.AvgCWL2GCN
   # model_class = gnn_models.with_fc(model_class)
   dsm = synthetic.threesix_dataset(stored=True)(
     wl2_neighborhood=1)  # ok(3, 2, 1)
@@ -294,7 +294,7 @@ def ds_stats():
     json.dump(s, f, cls=NumpyEncoder, indent="\t")
   print("Done.")
 
-# wl2_power_experiment()
+wl2_power_experiment()
 # synthetic_experiment2()
 # imdb_experient()
 
@@ -303,25 +303,25 @@ def ds_stats():
 # synthetic.balanced_triangle_classification_dataset(stored=True)().export_dot(152)
 # synthetic.balanced_triangle_classification_dataset(stored=True)().export_dot(153)
 # ds_stats()
-p = eval_ds.Proteins_5()
-d = eval_ds.DD_2()
-nci = eval_ds.NCI1_8()
-imdb = eval_ds.IMDB_8()
-reddit = eval_ds.RedditBinary_1()
+# p = eval_ds.Proteins_5()
+# d = eval_ds.DD_2()
+# nci = eval_ds.NCI1_8()
+# imdb = eval_ds.IMDB_8()
+# reddit = eval_ds.RedditBinary_1()
 
-nci.draw(30)
-nci.draw(33)
-nci.draw(204)
-nci.draw(1300)
-nci.draw(4101)
-
-imdb.draw(107)
-imdb.draw(999)
-imdb.export_dot(999)
-
-reddit.draw(2)
-reddit.draw(100)
-reddit.export_dot(100)
+# nci.draw(30)
+# nci.draw(33)
+# nci.draw(204)
+# nci.draw(1300)
+# nci.draw(4101)
+#
+# imdb.draw(107)
+# imdb.draw(999)
+# imdb.export_dot(999)
+#
+# reddit.draw(2)
+# reddit.draw(100)
+# reddit.export_dot(100)
 #
 # list(dsm.get_all(output_type="grakel")[0])
 #

### 2020-09-03 implementation/src/ltag/datasets/utils.py
implemented GIN in TF for a meaningful runtime comparison with 2-WL-GNNs

@@ -61,6 +61,53 @@ def eid_lookup(e_ids, i, j):
   return e_ids[(i, j)]
 
 @cp.tolerant
+def wl1_encode(g, dim_node_features=None, num_node_labels=None):
+  if dim_node_features is None:
+    dim_node_features = 0
+
+  if num_node_labels is None:
+    num_node_labels = 0
+
+  dim_node_wl1 = dim_node_features + num_node_labels
+
+  n_zero_f = np.zeros(dim_node_features)
+  I_n = np.eye(num_node_labels)
+
+  x = []
+  ref_a = []
+  ref_b = []
+
+  n_ids = {}
+  i = 0
+
+  for node, data in g.nodes(data=True):
+    if dim_node_wl1 == 0:
+      f = [1]
+      lab = []
+    else:
+      f = data.get("features", n_zero_f)
+      if num_node_labels > 0:
+        lab = I_n[data["label"] - 1]
+      else:
+        lab = []
+
+    n_ids[node] = i
+    i += 1
+    x.append(np.concatenate((lab, f)))
+
+  for (a, b) in g.edges():
+    ref_a.append(n_ids[a])
+    ref_b.append(n_ids[b])
+
+  n = g.order()
+  return (
+    np.array(x),
+    np.array(ref_a),
+    None,
+    np.array(ref_b),
+    n)
+
+@cp.tolerant
 def edge_neighborhood_encode(
   g,
   dim_node_features=None, dim_edge_features=None,
@@ -78,7 +125,7 @@ def edge_neighborhood_encode(
     dim_edge_features = 0
 
   if num_node_labels is None:
-    dim_edge_features = 0
+    num_node_labels = 0
 
   if num_edge_labels is None:
     num_edge_labels = 0
@@ -256,12 +303,12 @@ def wl2_encode(
 
   return res
 
-def make_wl2_batch(
-  encoded_graphs, dim_wl2, with_indices=False,
-  compact=False, en_encode=False):
+def make_wl_batch(
+  encoded_graphs, dim_wl, with_indices=False,
+  compact=False, en_encode=False, wl1_encode=False):
   "Takes a sequence of graphs that were encoded via `wl2_encode` or\
    `edge_neighborhood_encode` (if en_encode=True)."
-  if not compact and not en_encode:
+  if not compact and not en_encode and not wl1_encode:
     max_ref_dim = np.max([g[3] for g in encoded_graphs])
 
   x_len = 0
@@ -272,12 +319,12 @@ def make_wl2_batch(
     x_len += len(x)
[…]

### 2020-09-03 implementation/src/ltag/evaluation/evaluate.py

@@ -8,6 +8,7 @@ from timeit import default_timer as timer
 import warnings
 from pathlib import Path
 from datetime import datetime
+import numpy as np
 from tensorflow import keras
 import funcy as fy
 import tensorflow as tf
@@ -386,7 +387,10 @@ def evaluate_epoch_time(model_factory, ds_manager, hp_args=None):
   model = model_ctr(**list(hps)[0])
   train_ds = ds_manager.get_all(output_type=model_factory.input_type)
 
-  model.summary()
+  wc = int(np.sum([
+    keras.backend.count_params(w)
+    for w in model.trainable_weights]))
+  print(f"\nModel {model_factory.name} trainable weights = {wc}\n")
 
   _, times = train(
     model, train_ds,

### 2020-09-03 implementation/src/ltag/models/gnn.py

@@ -8,6 +8,7 @@ import funcy as fy
 import ltag.chaining.model as cm
 import ltag.layers.dense as ld
 import ltag.layers.wl2 as lwl2
+import ltag.layers.wl1 as lwl1
 from ltag.layers.DenseLayer import DenseLayer
 
 @cm.model_inputs
@@ -50,12 +51,23 @@ def gnn_en_inputs():
 
   return X, ref_a, backref, e_map, v_count
 
+@cm.model_inputs
+def gnn_wl1_inputs():
+  X = keras.Input(shape=(None,), dtype=tf.float32, name="X")
+  ref_a = keras.Input(shape=(), dtype=tf.int32, name="ref_a")
+  ref_b = keras.Input(shape=(), dtype=tf.int32, name="ref_b")
+  v_map = keras.Input(shape=(), dtype=tf.int32, name="v_map")
+  v_count = keras.Input(shape=(), dtype=tf.int32, name="v_count")
+
+  return X, ref_a, ref_b, v_map, v_count
+
 
 input_types = {
   "dense": gnn_dense_inputs,
   "wl2": gnn_wl2_inputs,
   "wl2c": gnn_wl2c_inputs,
-  "en": gnn_en_inputs
+  "en": gnn_en_inputs,
+  "wl1": gnn_wl1_inputs
 }
 
 @cm.model_step
@@ -128,6 +140,11 @@ K2GNN = gnn_model("K2GNN", [
   select_features],
   input_type="en")
 
+GIN = gnn_model("GIN", [
+  with_layers(lwl1.GINLayer, prefix="conv"),
+  select_features],
+  input_type="wl1")
+
 # Averaging GNNs:
 
 AvgDenseGCN = DenseGCN.extend("AvgDenseGCN", [
@@ -140,6 +157,8 @@ AvgWL2GCN = WL2GCN.extend("AvgWL2GCN", [
 AvgCWL2GCN = CWL2GCN.extend("AvgCWL2GCN", [
   cm.with_layer(lwl2.AvgPooling, with_inputs=True)])
 AvgK2GNN = K2GNN.extend("AvgK2GNN", [
+  cm.with_layer(lwl2.AvgPooling, with_inputs=True)])
+AvgGIN = GIN.extend("AvgGIN", [
   cm.with_layer(lwl2.AvgPooling, with_inputs=True)])
 
 # Max GNNs:

### 2020-09-03 implementation/src/ltag/ops.py

@@ -94,18 +94,36 @@ def wl1_convolution_compact(X, ref_a, backref):
 
   return X_agg
 
+@tf.function
+def gin_convolution_compact(X, ref_a, ref_b):
+  X_a = tf.gather(X, ref_a, axis=0)
+  X_b = tf.gather(X, ref_b, axis=0)
 
-# import ltag.datasets.synthetic.datasets as synthetic
-#
-# ds = synthetic.threesix_dataset()(wl2_batch_size={
-#   "batch_graph_count": 100
-# })
-#
-# (X, ref_a, ref_b, backref, e_map, v_count), y = list(
-#   ds.get_all(output_type="wl2c"))[0]
-#
-# c = wl2_convolution_compact(X, ref_a, ref_b, backref, tf.nn.relu)
-#
-# e_map
-#
-# print(c)
+  X_shape = tf.shape(X)
+
+  idx_a = tf.expand_dims(ref_a, axis=-1)
+  idx_b = tf.expand_dims(ref_b, axis=-1)
+  X_agg = (
+    X
+    + tf.scatter_nd(idx_a, X_b, shape=X_shape)
+    + tf.scatter_nd(idx_b, X_a, shape=X_shape))
+
+  return X_agg
+
+
+import ltag.datasets.synthetic.datasets as synthetic
+
+ds = synthetic.threesix_dataset()(wl2_batch_size={
+  "batch_graph_count": 100
+})
+
+(X, ref_a, ref_b, v_map, v_count), y = list(
+  ds.get_all(output_type="wl1"))[0]
+
+synthetic.threesix_dataset()().get_all(output_type="wl1")
+
+import numpy as np
+np.array(list(zip(ref_a, ref_b)))
+
+gin_convolution_compact(
+tf.constant(np.ones((12, 1))), ref_a, ref_b)

### 2020-09-05 thesis/data/diffs/imdb.tex
implemented epoch time evaluation result postprocessing

@@
 {\setlength\tabcolsep{2.5pt}\setlength{\extrarowheight}{2pt}%
-\begin{tabular}{lcccccccccc}
-& \rotatebox[origin=l]{90}{\textbf{WL\textsubscript{ST}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{WL\textsubscript{SP}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-LWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-GWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{Baseline} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{GIN} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\wmean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN} ($\wmean$)} \\
-\textbf{WL\textsubscript{ST}} ($T=3$)& & {\tiny\Vectorstack{-1.5\\ \pm 4.3}} & {\tiny\Vectorstack{+0.7\\ \pm 2.0}} & {\tiny\Vectorstack{+2.5\\ \pm 3.4}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+22\\ \pm 4.9}}} & {\tiny\Vectorstack{+6.1\\ \pm 6.4}} & {\tiny\Vectorstack{+1.5\\ \pm 2.6}} & {\tiny\Vectorstack{+2.0\\ \pm 2.7}} & {\tiny\Vectorstack{+1.7\\ \pm 3.7}} & {\tiny\Vectorstack{+1.8\\ \pm 3.8}} \\[2pt]
-\textbf{WL\textsubscript{SP}} ($T=3$)&{\tiny\Vectorstack{+1.5\\ \pm 4.3}} &  & {\tiny\Vectorstack{+2.2\\ \pm 5.0}} & {\tiny\Vectorstack{+4.0\\ \pm 5.2}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+24\\ \pm 5.9}}} & {\tiny\Vectorstack{+7.6\\ \pm 7.4}} & {\tiny\Vectorstack{+3.0\\ \pm 5.2}} & {\tiny\Vectorstack{+3.5\\ \pm 5.7}} & {\tiny\Vectorstack{+3.2\\ \pm 6.4}} & {\tiny\Vectorstack{+3.3\\ \pm 6.7}} \\[2pt]
-\textbf{2-LWL} ($T=3$)&{\tiny\Vectorstack{-0.7\\ \pm 2.0}} & {\tiny\Vectorstack{-2.2\\ \pm 5.0}} &  & {\tiny\Vectorstack{+1.8\\ \pm 2.6}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+22\\ \pm 5.7}}} & {\tiny\Vectorstack{+5.4\\ \pm 7.2}} & {\tiny\Vectorstack{+0.8\\ \pm 2.0}} & {\tiny\Vectorstack{+1.3\\ \pm 3.2}} & {\tiny\Vectorstack{+1.0\\ \pm 3.5}} & {\tiny\Vectorstack{+1.1\\ \pm 3.7}} \\[2pt]
-\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-2.5\\ \pm 3.4}} & {\tiny\Vectorstack{-4.0\\ \pm 5.2}} & {\tiny\Vectorstack{-1.8\\ \pm 2.6}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+20\\ \pm 5.6}}} & {\tiny\Vectorstack{+3.6\\ \pm 7.1}} & {\tiny\Vectorstack{-1.0\\ \pm 2.6}} & {\tiny\Vectorstack{-0.5\\ \pm 4.1}} & {\tiny\Vectorstack{-0.8\\ \pm 3.1}} & {\tiny\Vectorstack{-0.7\\ \pm 4.3}} \\[2pt]
-\textbf{Baseline} ($\mathrm{sum}$)&\cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-22\\ \pm 4.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-24\\ \pm 5.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 5.7}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-20\\ \pm 5.6}}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-16\\ \pm 6.3}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 6.0}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-20\\ \pm 5.6}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 6.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-20\\ \pm 6.9}}} \\[2pt]
-\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{-6.1\\ \pm 6.4}} & {\tiny\Vectorstack{-7.6\\ \pm 7.4}} & {\tiny\Vectorstack{-5.4\\ \pm 7.2}} & {\tiny\Vectorstack{-3.6\\ \pm 7.1}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+16\\ \pm 6.3}}} &  & {\tiny\Vectorstack{-4.6\\ \pm 7.5}} & {\tiny\Vectorstack{-4.1\\ \pm 7.1}} & {\tiny\Vectorstack{-4.4\\ \pm 7.9}} & {\tiny\Vectorstack{-4.3\\ \pm 8.4}} \\[2pt]
-\textbf{2-GNN} ($\mean$)&{\tiny\Vectorstack{-1.5\\ \pm 2.6}} & {\tiny\Vectorstack{-3.0\\ \pm 5.2}} & {\tiny\Vectorstack{-0.8\\ \pm 2.0}} & {\tiny\Vectorstack{+1.0\\ \pm 2.6}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+21\\ \pm 6.0}}} & {\tiny\Vectorstack{+4.6\\ \pm 7.5}} &  & {\tiny\Vectorstack{+0.5\\ \pm 2.8}} & {\tiny\Vectorstack{+0.2\\ \pm 2.8}} & {\tiny\Vectorstack{+0.2\\ \pm 2.4}} \\[2pt]
-\textbf{2-GNN} ($\wmean$)&{\tiny\Vectorstack{-2.0\\ \pm 2.7}} & {\tiny\Vectorstack{-3.5\\ \pm 5.7}} & {\tiny\Vectorstack{-1.3\\ \pm 3.2}} & {\tiny\Vectorstack{+0.5\\ \pm 4.1}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+20\\ \pm 5.6}}} & {\tiny\Vectorstack{+4.1\\ \pm 7.1}} & {\tiny\Vectorstack{-0.5\\ \pm 2.8}} &  & {\tiny\Vectorstack{-0.3\\ \pm 3.5}} & {\tiny\Vectorstack{-0.2\\ \pm 2.7}} \\[2pt]
-\textbf{2-WL-GNN} ($\mean$)&{\tiny\Vectorstack{-1.7\\ \pm 3.7}} & {\tiny\Vectorstack{-3.2\\ \pm 6.4}} & {\tiny\Vectorstack{-1.0\\ \pm 3.5}} & {\tiny\Vectorstack{+0.8\\ \pm 3.1}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+21\\ \pm 6.4}}} & {\tiny\Vectorstack{+4.4\\ \pm 7.9}} & {\tiny\Vectorstack{-0.2\\ \pm 2.8}} & {\tiny\Vectorstack{+0.3\\ \pm 3.5}} &  & {\tiny\Vectorstack{+0.1\\ \pm 3.4}} \\[2pt]
-\textbf{2-WL-GNN} ($\wmean$)&{\tiny\Vectorstack{-1.8\\ \pm 3.8}} & {\tiny\Vectorstack{-3.3\\ \pm 6.7}} & {\tiny\Vectorstack{-1.1\\ \pm 3.7}} & {\tiny\Vectorstack{+0.7\\ \pm 4.3}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+20\\ \pm 6.9}}} & {\tiny\Vectorstack{+4.3\\ \pm 8.4}} & {\tiny\Vectorstack{-0.2\\ \pm 2.4}} & {\tiny\Vectorstack{+0.2\\ \pm 2.7}} & {\tiny\Vectorstack{-0.1\\ \pm 3.4}} &
+\begin{tabular}{lcccccccccccc}
+& \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{WL\textsubscript{SP}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-GWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{Baseline} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{GIN} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)} \\
+\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=3$)& & {\tiny\Vectorstack{-1.5\\ \pm 4.3}} & {\tiny\Vectorstack{+0.7\\ \pm 2.0}} & {\tiny\Vectorstack{+2.5\\ \pm 3.4}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+22\\ \pm 4.9}}} & {\tiny\Vectorstack{+6.1\\ \pm 6.4}} & {\tiny\Vectorstack{+1.5\\ \pm 2.6}} & {\tiny\Vectorstack{+2.0\\ \pm 2.7}} & {\tiny\Vectorstack{+2.0\\ \pm 3.2}} & {\tiny\Vectorstack{+1.7\\ \pm 3.7}} & {\tiny\Vectorstack{+0.7\\ \pm 3.2}} & {\tiny\Vectorstack{+1.8\\ \pm 3.8}} \\[2pt]
+\textbf{WL\textsubscript{SP}} ($T=3$)&{\tiny\Vectorstack{+1.5\\ \pm 4.3}} &  & {\tiny\Vectorstack{+2.2\\ \pm 5.0}} & {\tiny\Vectorstack{+4.0\\ \pm 5.2}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+24\\ \pm 5.9}}} & {\tiny\Vectorstack{+7.6\\ \pm 7.4}} & {\tiny\Vectorstack{+3.0\\ \pm 5.2}} & {\tiny\Vectorstack{+3.5\\ \pm 5.7}} & {\tiny\Vectorstack{+3.5\\ \pm 6.0}} & {\tiny\Vectorstack{+3.2\\ \pm 6.4}} & {\tiny\Vectorstack{+2.2\\ \pm 6.1}} & {\tiny\Vectorstack{+3.3\\ \pm 6.7}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)&{\tiny\Vectorstack{-0.7\\ \pm 2.0}} & {\tiny\Vectorstack{-2.2\\ \pm 5.0}} &  & {\tiny\Vectorstack{+1.8\\ \pm 2.6}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+22\\ \pm 5.7}}} & {\tiny\Vectorstack{+5.4\\ \pm 7.2}} & {\tiny\Vectorstack{+0.8\\ \pm 2.0}} & {\tiny\Vectorstack{+1.3\\ \pm 3.2}} & {\tiny\Vectorstack{+1.3\\ \pm 3.2}} & {\tiny\Vectorstack{+1.0\\ \pm 3.5}} & {\tiny\Vectorstack{+0.0\\ \pm 3.0}} & {\tiny\Vectorstack{+1.1\\ \pm 3.7}} \\[2pt]
+\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-2.5\\ \pm 3.4}} & {\tiny\Vectorstack{-4.0\\ \pm 5.2}} & {\tiny\Vectorstack{-1.8\\ \pm 2.6}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+20\\ \pm 5.6}}} & {\tiny\Vectorstack{+3.6\\ \pm 7.1}} & {\tiny\Vectorstack{-1.0\\ \pm 2.6}} & {\tiny\Vectorstack{-0.5\\ \pm 4.1}} & {\tiny\Vectorstack{-0.5\\ \pm 3.3}} & {\tiny\Vectorstack{-0.8\\ \pm 3.1}} & {\tiny\Vectorstack{-1.8\\ \pm 3.0}} & {\tiny\Vectorstack{-0.7\\ \pm 4.3}} \\[2pt]
+\textbf{Baseline} ($\mathrm{sum}$)&\cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-22\\ \pm 4.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-24\\ \pm 5.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 5.7}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-20\\ \pm 5.6}}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-16\\ \pm 6.3}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 6.0}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-20\\ \pm 5.6}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-20\\ \pm 5.8}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 6.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 5.5}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-20\\ \pm 6.9}}} \\[2pt]
+\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{-6.1\\ \pm 6.4}} & {\tiny\Vectorstack{-7.6\\ \pm 7.4}} & {\tiny\Vectorstack{-5.4\\ \pm 7.2}} & {\tiny\Vectorstack{-3.6\\ \pm 7.1}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+16\\ \pm 6.3}}} &  & {\tiny\Vectorstack{-4.6\\ \pm 7.5}} & {\tiny\Vectorstack{-4.1\\ \pm 7.1}} & {\tiny\Vectorstack{-4.1\\ \pm 7.3}} & {\tiny\Vectorstack{-4.4\\ \pm 7.9}} & {\tiny\Vectorstack{-5.4\\ \pm 7.0}} & {\tiny\Vectorstack{-4.3\\ \pm 8.4}} \\[2pt]
+\textbf{2-GNN} ($\mean$)&{\tiny\Vectorstack{-1.5\\ \pm 2.6}} & {\tiny\Vectorstack{-3.0\\ \pm 5.2}} & {\tiny\Vectorstack{-0.8\\ \pm 2.0}} & {\tiny\Vectorstack{+1.0\\ \pm 2.6}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+21\\ \pm 6.0}}} & {\tiny\Vectorstack{+4.6\\ \pm 7.5}} &  & {\tiny\Vectorstack{+0.5\\ \pm 2.8}} & {\tiny\Vectorstack{+0.5\\ \pm 2.5}} & {\tiny\Vectorstack{+0.2\\ \pm 2.8}} & {\tiny\Vectorstack{-0.8\\ \pm 2.6}} & {\tiny\Vectorstack{+0.2\\ \pm 2.4}} \\[2pt]
+\textbf{2-GNN} ($\mathrm{SAM}$)&{\tiny\Vectorstack{-2.0\\ \pm 2.7}} & {\tiny\Vectorstack{-3.5\\ \pm 5.7}} & {\tiny\Vectorstack{-1.3\\ \pm 3.2}} & {\tiny\Vectorstack{+0.5\\ \pm 4.1}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+20\\ \pm 5.6}}} & {\tiny\Vectorstack{+4.1\\ \pm 7.1}} & {\tiny\Vectorstack{-0.5\\ \pm 2.8}} &  & {\tiny\Vectorstack{+0.0\\ \pm 2.8}} & {\tiny\Vectorstack{-0.3\\ \pm 3.5}} & {\tiny\Vectorstack{-1.3\\ \pm 2.5}} & {\tiny\Vectorstack{-0.2\\ \pm 2.7}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)&{\tiny\Vectorstack{-2.0\\ \pm 3.2}} & {\tiny\Vectorstack{-3.5\\ \pm 6.0}} & {\tiny\Vectorstack{-1.3\\ \pm 3.2}} & {\tiny\Vectorstack{+0.5\\ \pm 3.3}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+20\\ \pm 5.8}}} & {\tiny\Vectorstack{+4.1\\ \pm 7.3}} & {\tiny\Vectorstack{-0.5\\ \pm 2.5}} & {\tiny\Vectorstack{-0.0\\ \pm 2.8}} &  & {\tiny\Vectorstack{-0.3\\ \pm 1.7}} & {\tiny\Vectorstack{-1.3\\ \pm 1.1}} & {\tiny\Vectorstack{-0.3\\ \pm 2.2}} \\[2pt]
+\textbf{2-WL-GNN\phantom{*}} ($\mean$)&{\tiny\Vectorstack{-1.7\\ \pm 3.7}} & {\tiny\Vectorstack{-3.2\\ \pm 6.4}} & {\tiny\Vectorstack{-1.0\\ \pm 3.5}} & {\tiny\Vectorstack{+0.8\\ \pm 3.1}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+21\\ \pm 6.4}}} & {\tiny\Vectorstack{+4.4\\ \pm 7.9}} & {\tiny\Vectorstack{-0.2\\ \pm 2.8}} & {\tiny\Vectorstack{+0.3\\ \pm 3.5}} & {\tiny\Vectorstack{+0.3\\ \pm 1.7}} &  & {\tiny\Vectorstack{-1.0\\ \pm 2.0}} & {\tiny\Vectorstack{+0.1\\ \pm 3.4}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{-0.7\\ \pm 3.2}} & {\tiny\Vectorstack{-2.2\\ \pm 6.1}} & {\tiny\Vectorstack{-0.0\\ \pm 3.0}} & {\tiny\Vectorstack{+1.8\\ \pm 3.0}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+21\\ \pm 5.5}}} & {\tiny\Vectorstack{+5.4\\ \pm 7.0}} & {\tiny\Vectorstack{+0.8\\ \pm 2.6}} & {\tiny\Vectorstack{+1.3\\ \pm 2.5}} & {\tiny\Vectorstack{+1.3\\ \pm 1.1}} & {\tiny\Vectorstack{+1.0\\ \pm 2.0}} &  & {\tiny\Vectorstack{+1.0\\ \pm 2.6}} \\[2pt]
+\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{-1.8\\ \pm 3.8}} & {\tiny\Vectorstack{-3.3\\ \pm 6.7}} & {\tiny\Vectorstack{-1.1\\ \pm 3.7}} & {\tiny\Vectorstack{+0.7\\ \pm 4.3}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+20\\ \pm 6.9}}} & {\tiny\Vectorstack{+4.3\\ \pm 8.4}} & {\tiny\Vectorstack{-0.2\\ \pm 2.4}} & {\tiny\Vectorstack{+0.2\\ \pm 2.7}} & {\tiny\Vectorstack{+0.3\\ \pm 2.2}} & {\tiny\Vectorstack{-0.1\\ \pm 3.4}} & {\tiny\Vectorstack{-1.0\\ \pm 2.6}} &
 \end{tabular}}

### 2020-09-05 thesis/data/diffs/reddit.tex
implemented epoch time evaluation result postprocessing

--- 2020-04-12 thesis/data/diffs/reddit.tex
@@
 & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-GWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{Baseline} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{GIN} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)} \\
-\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)& & {\tiny\Vectorstack{-1.8\\ \pm 1.8}} & {\tiny\Vectorstack{+0.5\\ \pm 4.9}} & {\tiny\Vectorstack{+0.8\\ \pm 3.6}} & {\tiny\Vectorstack{+4.2\\ \pm 10}} & {\tiny\Vectorstack{-11\\ \pm 6.9}} & {\tiny\Vectorstack{+2.9\\ \pm 3.5}} & {\tiny\Vectorstack{-7.4\\ \pm 6.2}} & {\tiny\Vectorstack{-0.9\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-13\\ \pm 3.3}}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=3$)&{\tiny\Vectorstack{+1.8\\ \pm 1.8}} &  & {\tiny\Vectorstack{+2.3\\ \pm 4.7}} & {\tiny\Vectorstack{+2.6\\ \pm 3.5}} & {\tiny\Vectorstack{+5.9\\ \pm 11}} & {\tiny\Vectorstack{-9.0\\ \pm 7.1}} & {\tiny\Vectorstack{+4.7\\ \pm 3.3}} & {\tiny\Vectorstack{-5.7\\ \pm 6.8}} & {\tiny\Vectorstack{+0.9\\ \pm 1.6}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 2.6}}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)&{\tiny\Vectorstack{-0.5\\ \pm 4.9}} & {\tiny\Vectorstack{-2.3\\ \pm 4.7}} &  & {\tiny\Vectorstack{+0.3\\ \pm 4.2}} & {\tiny\Vectorstack{+3.7\\ \pm 11}} & {\tiny\Vectorstack{-11\\ \pm 7.3}} & {\tiny\Vectorstack{+2.4\\ \pm 5.5}} & {\tiny\Vectorstack{-7.9\\ \pm 5.5}} & {\tiny\Vectorstack{-1.4\\ \pm 5.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-13\\ \pm 3.3}}} \\[2pt]
-\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-0.8\\ \pm 3.6}} & {\tiny\Vectorstack{-2.6\\ \pm 3.5}} & {\tiny\Vectorstack{-0.3\\ \pm 4.2}} &  & {\tiny\Vectorstack{+3.3\\ \pm 11}} & {\tiny\Vectorstack{-12\\ \pm 7.6}} & {\tiny\Vectorstack{+2.1\\ \pm 2.4}} & {\tiny\Vectorstack{-8.2\\ \pm 6.2}} & {\tiny\Vectorstack{-1.7\\ \pm 3.3}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-14\\ \pm 2.5}}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)& & {\tiny\Vectorstack{-1.8\\ \pm 1.8}} & {\tiny\Vectorstack{+0.5\\ \pm 4.9}} & {\tiny\Vectorstack{+0.8\\ \pm 3.6}} & {\tiny\Vectorstack{+4.2\\ \pm 10}} & {\tiny\Vectorstack{-11\\ \pm 6.9}} & {\tiny\Vectorstack{+2.9\\ \pm 3.5}} & {\tiny\Vectorstack{-7.4\\ \pm 6.2}} & {\tiny\Vectorstack{-0.9\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-13\\ \pm 3.5}}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=3$)&{\tiny\Vectorstack{+1.8\\ \pm 1.8}} &  & {\tiny\Vectorstack{+2.3\\ \pm 4.7}} & {\tiny\Vectorstack{+2.6\\ \pm 3.5}} & {\tiny\Vectorstack{+5.9\\ \pm 11}} & {\tiny\Vectorstack{-9.0\\ \pm 7.1}} & {\tiny\Vectorstack{+4.7\\ \pm 3.3}} & {\tiny\Vectorstack{-5.7\\ \pm 6.8}} & {\tiny\Vectorstack{+0.9\\ \pm 1.6}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 2.9}}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)&{\tiny\Vectorstack{-0.5\\ \pm 4.9}} & {\tiny\Vectorstack{-2.3\\ \pm 4.7}} &  & {\tiny\Vectorstack{+0.3\\ \pm 4.2}} & {\tiny\Vectorstack{+3.7\\ \pm 11}} & {\tiny\Vectorstack{-11\\ \pm 7.3}} & {\tiny\Vectorstack{+2.4\\ \pm 5.5}} & {\tiny\Vectorstack{-7.9\\ \pm 5.5}} & {\tiny\Vectorstack{-1.4\\ \pm 5.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-14\\ \pm 3.4}}} \\[2pt]
+\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-0.8\\ \pm 3.6}} & {\tiny\Vectorstack{-2.6\\ \pm 3.5}} & {\tiny\Vectorstack{-0.3\\ \pm 4.2}} &  & {\tiny\Vectorstack{+3.3\\ \pm 11}} & {\tiny\Vectorstack{-12\\ \pm 7.6}} & {\tiny\Vectorstack{+2.1\\ \pm 2.4}} & {\tiny\Vectorstack{-8.2\\ \pm 6.2}} & {\tiny\Vectorstack{-1.7\\ \pm 3.3}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-14\\ \pm 2.7}}} \\[2pt]
 \textbf{Baseline} ($\mathrm{sum}$)&{\tiny\Vectorstack{-4.2\\ \pm 10}} & {\tiny\Vectorstack{-5.9\\ \pm 11}} & {\tiny\Vectorstack{-3.7\\ \pm 11}} & {\tiny\Vectorstack{-3.3\\ \pm 11}} &  & {\tiny\Vectorstack{-15\\ \pm 12}} & {\tiny\Vectorstack{-1.2\\ \pm 12}} & {\tiny\Vectorstack{-12\\ \pm 13}} & {\tiny\Vectorstack{-5.0\\ \pm 11}} & {\tiny\Vectorstack{-17\\ \pm 10}} \\[2pt]
-\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{+11\\ \pm 6.9}} & {\tiny\Vectorstack{+9.0\\ \pm 7.1}} & {\tiny\Vectorstack{+11\\ \pm 7.3}} & {\tiny\Vectorstack{+12\\ \pm 7.6}} & {\tiny\Vectorstack{+15\\ \pm 12}} &  & {\tiny\Vectorstack{+14\\ \pm 8.3}} & {\tiny\Vectorstack{+3.3\\ \pm 9.6}} & {\tiny\Vectorstack{+9.9\\ \pm 7.4}} & {\tiny\Vectorstack{-2.2\\ \pm 6.8}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)&{\tiny\Vectorstack{-2.9\\ \pm 3.5}} & {\tiny\Vectorstack{-4.7\\ \pm 3.3}} & {\tiny\Vectorstack{-2.4\\ \pm 5.5}} & {\tiny\Vectorstack{-2.1\\ \pm 2.4}} & {\tiny\Vectorstack{+1.2\\ \pm 12}} & {\tiny\Vectorstack{-14\\ \pm 8.3}} &  & {\tiny\Vectorstack{-10\\ \pm 7.1}} & {\tiny\Vectorstack{-3.8\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-16\\ \pm 2.8}}} \\[2pt]
-\textbf{2-WL-GNN\phantom{*}} ($\mean$)&{\tiny\Vectorstack{+7.4\\ \pm 6.2}} & {\tiny\Vectorstack{+5.7\\ \pm 6.8}} & {\tiny\Vectorstack{+7.9\\ \pm 5.5}} & {\tiny\Vectorstack{+8.2\\ \pm 6.2}} & {\tiny\Vectorstack{+12\\ \pm 13}} & {\tiny\Vectorstack{-3.3\\ \pm 9.6}} & {\tiny\Vectorstack{+10\\ \pm 7.1}} &  & {\tiny\Vectorstack{+6.6\\ \pm 7.2}} & {\tiny\Vectorstack{-5.5\\ \pm 6.1}} \\[2pt]
-\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{+0.9\\ \pm 2.5}} & {\tiny\Vectorstack{-0.9\\ \pm 1.6}} & {\tiny\Vectorstack{+1.4\\ \pm 5.0}} & {\tiny\Vectorstack{+1.7\\ \pm 3.3}} & {\tiny\Vectorstack{+5.0\\ \pm 11}} & {\tiny\Vectorstack{-9.9\\ \pm 7.4}} & {\tiny\Vectorstack{+3.8\\ \pm 2.5}} & {\tiny\Vectorstack{-6.6\\ \pm 7.2}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-12\\ \pm 2.7}}} \\[2pt]
-\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+13\\ \pm 3.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 2.6}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+13\\ \pm 3.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+14\\ \pm 2.5}}} & {\tiny\Vectorstack{+17\\ \pm 10}} & {\tiny\Vectorstack{+2.2\\ \pm 6.8}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+16\\ \pm 2.8}}} & {\tiny\Vectorstack{+5.5\\ \pm 6.1}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+12\\ \pm 2.7}}} &
+\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{+11\\ \pm 6.9}} & {\tiny\Vectorstack{+9.0\\ \pm 7.1}} & {\tiny\Vectorstack{+11\\ \pm 7.3}} & {\tiny\Vectorstack{+12\\ \pm 7.6}} & {\tiny\Vectorstack{+15\\ \pm 12}} &  & {\tiny\Vectorstack{+14\\ \pm 8.3}} & {\tiny\Vectorstack{+3.3\\ \pm 9.6}} & {\tiny\Vectorstack{+9.9\\ \pm 7.4}} & {\tiny\Vectorstack{-2.4\\ \pm 7.0}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)&{\tiny\Vectorstack{-2.9\\ \pm 3.5}} & {\tiny\Vectorstack{-4.7\\ \pm 3.3}} & {\tiny\Vectorstack{-2.4\\ \pm 5.5}} & {\tiny\Vectorstack{-2.1\\ \pm 2.4}} & {\tiny\Vectorstack{+1.2\\ \pm 12}} & {\tiny\Vectorstack{-14\\ \pm 8.3}} &  & {\tiny\Vectorstack{-10\\ \pm 7.1}} & {\tiny\Vectorstack{-3.8\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-16\\ \pm 3.0}}} \\[2pt]
+\textbf{2-WL-GNN\phantom{*}} ($\mean$)&{\tiny\Vectorstack{+7.4\\ \pm 6.2}} & {\tiny\Vectorstack{+5.7\\ \pm 6.8}} & {\tiny\Vectorstack{+7.9\\ \pm 5.5}} & {\tiny\Vectorstack{+8.2\\ \pm 6.2}} & {\tiny\Vectorstack{+12\\ \pm 13}} & {\tiny\Vectorstack{-3.3\\ \pm 9.6}} & {\tiny\Vectorstack{+10\\ \pm 7.1}} &  & {\tiny\Vectorstack{+6.6\\ \pm 7.2}} & {\tiny\Vectorstack{-5.7\\ \pm 6.5}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{+0.9\\ \pm 2.5}} & {\tiny\Vectorstack{-0.9\\ \pm 1.6}} & {\tiny\Vectorstack{+1.4\\ \pm 5.0}} & {\tiny\Vectorstack{+1.7\\ \pm 3.3}} & {\tiny\Vectorstack{+5.0\\ \pm 11}} & {\tiny\Vectorstack{-9.9\\ \pm 7.4}} & {\tiny\Vectorstack{+3.8\\ \pm 2.5}} & {\tiny\Vectorstack{-6.6\\ \pm 7.2}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-12\\ \pm 2.9}}} \\[2pt]
+\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+13\\ \pm 3.5}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 2.9}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+14\\ \pm 3.4}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+14\\ \pm 2.7}}} & {\tiny\Vectorstack{+17\\ \pm 10}} & {\tiny\Vectorstack{+2.4\\ \pm 7.0}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+16\\ \pm 3.0}}} & {\tiny\Vectorstack{+5.7\\ \pm 6.5}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+12\\ \pm 2.9}}} &
 \end{tabular}}

### 2020-09-05 thesis/data/diffs/triangle.tex
implemented epoch time evaluation result postprocessing

@@
 {\setlength\tabcolsep{2.5pt}\setlength{\extrarowheight}{2pt}%
-\begin{tabular}{lcccccccccc}
-& \rotatebox[origin=l]{90}{\textbf{WL\textsubscript{ST}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{WL\textsubscript{SP}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-LWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-GWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{Baseline} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{GIN} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\wmean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN} ($\wmean$)} \\
-\textbf{WL\textsubscript{ST}} ($T=3$)& & {\tiny\Vectorstack{-11\\ \pm 13}} & {\tiny\Vectorstack{+0.4\\ \pm 9.6}} & {\tiny\Vectorstack{-4.9\\ \pm 16}} & {\tiny\Vectorstack{+12\\ \pm 12}} & {\tiny\Vectorstack{-13\\ \pm 10}} & {\tiny\Vectorstack{-20\\ \pm 14}} & {\tiny\Vectorstack{-25\\ \pm 14}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-36\\ \pm 17}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-42\\ \pm 11}}} \\[2pt]
-\textbf{WL\textsubscript{SP}} ($T=3$)&{\tiny\Vectorstack{+11\\ \pm 13}} &  & {\tiny\Vectorstack{+11\\ \pm 13}} & {\tiny\Vectorstack{+6.1\\ \pm 9.2}} & {\tiny\Vectorstack{+23\\ \pm 16}} & {\tiny\Vectorstack{-2.0\\ \pm 14}} & {\tiny\Vectorstack{-8.8\\ \pm 13}} & {\tiny\Vectorstack{-14\\ \pm 14}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-25\\ \pm 12}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-31\\ \pm 11}}} \\[2pt]
-\textbf{2-LWL} ($T=3$)&{\tiny\Vectorstack{-0.4\\ \pm 9.6}} & {\tiny\Vectorstack{-11\\ \pm 13}} &  & {\tiny\Vectorstack{-5.3\\ \pm 11}} & {\tiny\Vectorstack{+12\\ \pm 7.5}} & {\tiny\Vectorstack{-13\\ \pm 6.7}} & {\tiny\Vectorstack{-20\\ \pm 13}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-25\\ \pm 9.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-36\\ \pm 10}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-43\\ \pm 6.8}}} \\[2pt]
-\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{+4.9\\ \pm 16}} & {\tiny\Vectorstack{-6.1\\ \pm 9.2}} & {\tiny\Vectorstack{+5.3\\ \pm 11}} &  & {\tiny\Vectorstack{+17\\ \pm 15}} & {\tiny\Vectorstack{-8.2\\ \pm 13}} & {\tiny\Vectorstack{-15\\ \pm 14}} & {\tiny\Vectorstack{-20\\ \pm 13}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-31\\ \pm 9.0}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-38\\ \pm 9.0}}} \\[2pt]
-\textbf{Baseline} ($\mathrm{sum}$)&{\tiny\Vectorstack{-12\\ \pm 12}} & {\tiny\Vectorstack{-23\\ \pm 16}} & {\tiny\Vectorstack{-12\\ \pm 7.5}} & {\tiny\Vectorstack{-17\\ \pm 15}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-25\\ \pm 8.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-32\\ \pm 14}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-37\\ \pm 9.2}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-48\\ \pm 14}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-55\\ \pm 8.5}}} \\[2pt]
-\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{+13\\ \pm 10}} & {\tiny\Vectorstack{+2.0\\ \pm 14}} & {\tiny\Vectorstack{+13\\ \pm 6.7}} & {\tiny\Vectorstack{+8.2\\ \pm 13}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+25\\ \pm 8.4}}} &  & {\tiny\Vectorstack{-6.8\\ \pm 13}} & {\tiny\Vectorstack{-12\\ \pm 9.5}} & {\tiny\Vectorstack{-23\\ \pm 14}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-29\\ \pm 7.3}}} \\[2pt]
-\textbf{2-GNN} ($\mean$)&{\tiny\Vectorstack{+20\\ \pm 14}} & {\tiny\Vectorstack{+8.8\\ \pm 13}} & {\tiny\Vectorstack{+20\\ \pm 13}} & {\tiny\Vectorstack{+15\\ \pm 14}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+32\\ \pm 14}}} & {\tiny\Vectorstack{+6.8\\ \pm 13}} &  & {\tiny\Vectorstack{-5.0\\ \pm 6.6}} & {\tiny\Vectorstack{-16\\ \pm 11}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-23\\ \pm 9.8}}} \\[2pt]
-\textbf{2-GNN} ($\wmean$)&{\tiny\Vectorstack{+25\\ \pm 14}} & {\tiny\Vectorstack{+14\\ \pm 14}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+25\\ \pm 9.4}}} & {\tiny\Vectorstack{+20\\ \pm 13}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+37\\ \pm 9.2}}} & {\tiny\Vectorstack{+12\\ \pm 9.5}} & {\tiny\Vectorstack{+5.0\\ \pm 6.6}} &  & {\tiny\Vectorstack{-11\\ \pm 8.8}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-18\\ \pm 7.2}}} \\[2pt]
-\textbf{2-WL-GNN} ($\mean$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+36\\ \pm 17}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+25\\ \pm 12}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+36\\ \pm 10}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+31\\ \pm 9.0}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+48\\ \pm 14}}} & {\tiny\Vectorstack{+23\\ \pm 14}} & {\tiny\Vectorstack{+16\\ \pm 11}} & {\tiny\Vectorstack{+11\\ \pm 8.8}} &  & {\tiny\Vectorstack{-6.5\\ \pm 8.7}} \\[2pt]
-\textbf{2-WL-GNN} ($\wmean$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+42\\ \pm 11}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+31\\ \pm 11}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+43\\ \pm 6.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+38\\ \pm 9.0}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+55\\ \pm 8.5}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+29\\ \pm 7.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+23\\ \pm 9.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+18\\ \pm 7.2}}} & {\tiny\Vectorstack{+6.5\\ \pm 8.7}} &
+\begin{tabular}{lcccccccccccc}
+& \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{WL\textsubscript{SP}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-GWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{Baseline} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{GIN} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)} \\
+\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=3$)& & {\tiny\Vectorstack{-11\\ \pm 13}} & {\tiny\Vectorstack{+0.4\\ \pm 9.6}} & {\tiny\Vectorstack{-4.9\\ \pm 16}} & {\tiny\Vectorstack{+12\\ \pm 12}} & {\tiny\Vectorstack{-13\\ \pm 10}} & {\tiny\Vectorstack{-20\\ \pm 14}} & {\tiny\Vectorstack{-25\\ \pm 14}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-34\\ \pm 14}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-36\\ \pm 17}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-34\\ \pm 8.6}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-42\\ \pm 11}}} \\[2pt]
+\textbf{WL\textsubscript{SP}} ($T=3$)&{\tiny\Vectorstack{+11\\ \pm 13}} &  & {\tiny\Vectorstack{+11\\ \pm 13}} & {\tiny\Vectorstack{+6.1\\ \pm 9.2}} & {\tiny\Vectorstack{+23\\ \pm 16}} & {\tiny\Vectorstack{-2.0\\ \pm 14}} & {\tiny\Vectorstack{-8.8\\ \pm 13}} & {\tiny\Vectorstack{-14\\ \pm 14}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-23\\ \pm 11}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-25\\ \pm 12}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-23\\ \pm 8.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-31\\ \pm 11}}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)&{\tiny\Vectorstack{-0.4\\ \pm 9.6}} & {\tiny\Vectorstack{-11\\ \pm 13}} &  & {\tiny\Vectorstack{-5.3\\ \pm 11}} & {\tiny\Vectorstack{+12\\ \pm 7.5}} & {\tiny\Vectorstack{-13\\ \pm 6.7}} & {\tiny\Vectorstack{-20\\ \pm 13}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-25\\ \pm 9.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-34\\ \pm 8.6}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-36\\ \pm 10}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-34\\ \pm 7.0}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-43\\ \pm 6.8}}} \\[2pt]
+\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{+4.9\\ \pm 16}} & {\tiny\Vectorstack{-6.1\\ \pm 9.2}} & {\tiny\Vectorstack{+5.3\\ \pm 11}} &  & {\tiny\Vectorstack{+17\\ \pm 15}} & {\tiny\Vectorstack{-8.2\\ \pm 13}} & {\tiny\Vectorstack{-15\\ \pm 14}} & {\tiny\Vectorstack{-20\\ \pm 13}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-29\\ \pm 8.8}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-31\\ \pm 9.0}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-29\\ \pm 10}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-38\\ \pm 9.0}}} \\[2pt]
+\textbf{Baseline} ($\mathrm{sum}$)&{\tiny\Vectorstack{-12\\ \pm 12}} & {\tiny\Vectorstack{-23\\ \pm 16}} & {\tiny\Vectorstack{-12\\ \pm 7.5}} & {\tiny\Vectorstack{-17\\ \pm 15}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-25\\ \pm 8.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-32\\ \pm 14}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-37\\ \pm 9.2}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-46\\ \pm 9.8}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-48\\ \pm 14}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-46\\ \pm 9.2}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-55\\ \pm 8.5}}} \\[2pt]
+\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{+13\\ \pm 10}} & {\tiny\Vectorstack{+2.0\\ \pm 14}} & {\tiny\Vectorstack{+13\\ \pm 6.7}} & {\tiny\Vectorstack{+8.2\\ \pm 13}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+25\\ \pm 8.4}}} &  & {\tiny\Vectorstack{-6.8\\ \pm 13}} & {\tiny\Vectorstack{-12\\ \pm 9.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 9.9}}} & {\tiny\Vectorstack{-23\\ \pm 14}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 8.3}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-29\\ \pm 7.3}}} \\[2pt]
+\textbf{2-GNN} ($\mean$)&{\tiny\Vectorstack{+20\\ \pm 14}} & {\tiny\Vectorstack{+8.8\\ \pm 13}} & {\tiny\Vectorstack{+20\\ \pm 13}} & {\tiny\Vectorstack{+15\\ \pm 14}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+32\\ \pm 14}}} & {\tiny\Vectorstack{+6.8\\ \pm 13}} &  & {\tiny\Vectorstack{-5.0\\ \pm 6.6}} & {\tiny\Vectorstack{-14\\ \pm 9.6}} & {\tiny\Vectorstack{-16\\ \pm 11}} & {\tiny\Vectorstack{-14\\ \pm 8.8}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-23\\ \pm 9.8}}} \\[2pt]
+\textbf{2-GNN} ($\mathrm{SAM}$)&{\tiny\Vectorstack{+25\\ \pm 14}} & {\tiny\Vectorstack{+14\\ \pm 14}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+25\\ \pm 9.4}}} & {\tiny\Vectorstack{+20\\ \pm 13}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+37\\ \pm 9.2}}} & {\tiny\Vectorstack{+12\\ \pm 9.5}} & {\tiny\Vectorstack{+5.0\\ \pm 6.6}} &  & {\tiny\Vectorstack{-8.9\\ \pm 6.3}} & {\tiny\Vectorstack{-11\\ \pm 8.8}} & {\tiny\Vectorstack{-9.2\\ \pm 7.1}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-18\\ \pm 7.2}}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+34\\ \pm 14}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+23\\ \pm 11}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+34\\ \pm 8.6}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+29\\ \pm 8.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+46\\ \pm 9.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+21\\ \pm 9.9}}} & {\tiny\Vectorstack{+14\\ \pm 9.6}} & {\tiny\Vectorstack{+8.9\\ \pm 6.3}} &  & {\tiny\Vectorstack{-2.2\\ \pm 7.2}} & {\tiny\Vectorstack{-0.3\\ \pm 6.0}} & {\tiny\Vectorstack{-8.7\\ \pm 4.6}} \\[2pt]
+\textbf{2-WL-GNN\phantom{*}} ($\mean$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+36\\ \pm 17}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+25\\ \pm 12}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+36\\ \pm 10}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+31\\ \pm 9.0}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+48\\ \pm 14}}} & {\tiny\Vectorstack{+23\\ \pm 14}} & {\tiny\Vectorstack{+16\\ \pm 11}} & {\tiny\Vectorstack{+11\\ \pm 8.8}} & {\tiny\Vectorstack{+2.2\\ \pm 7.2}} &  & {\tiny\Vectorstack{+1.9\\ \pm 9.7}} & {\tiny\Vectorstack{-6.5\\ \pm 8.7}} \\[2pt]
+\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+34\\ \pm 8.6}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+23\\ \pm 8.4}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+34\\ \pm 7.0}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+29\\ \pm 10}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+46\\ \pm 9.2}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+21\\ \pm 8.3}}} & {\tiny\Vectorstack{+14\\ \pm 8.8}} & {\tiny\Vectorstack{+9.2\\ \pm 7.1}} & {\tiny\Vectorstack{+0.3\\ \pm 6.0}} & {\tiny\Vectorstack{-1.9\\ \pm 9.7}} &  & {\tiny\Vectorstack{-8.5\\ \pm 4.7}} \\[2pt]
+\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+42\\ \pm 11}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+31\\ \pm 11}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+43\\ \pm 6.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+38\\ \pm 9.0}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+55\\ \pm 8.5}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+29\\ \pm 7.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+23\\ \pm 9.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+18\\ \pm 7.2}}} & {\tiny\Vectorstack{+8.7\\ \pm 4.6}} & {\tiny\Vectorstack{+6.5\\ \pm 8.7}} & {\tiny\Vectorstack{+8.5\\ \pm 4.7}} &
 \end{tabular}}
